---
title: "analyses-C1A"
author: "Justin Isip"
date: "2024-01-10"
output: html_document
---

# Read in the data and load packages
```{r}
diversity <- readRDS("./data/diversity-2023-09-14-02-33-21.rds")
library(tidyverse)
library(lme4)
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(sjPlot) # for visualising results
library(effects) # for extracting model effects
library(merTools) # useful for a few things, but we're using it for extracting estimates for plotting
library(emmeans) # testing multiple comparisons (but also useful for plotting)
library(coefplot) # visualising model coefficients
library(cli)
library(ggeffects)
library(ggpubr)
library(lmerTest)
library(influence.ME)
library(robustlmm)
library(performance)
```

# Load functions
```{r}
source("/Applications/PhD/Andy/data_prep_functions.R") 
```

### Load in model plot function
```{r}

## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(2,2))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}
```

# Correct for sampling effort
```{r}
if(check_sampling_effort_nas(diversity) ==
   FALSE) stop("Some studies have NA for sampling effort in only some sites!")

diversity <- correct_for_sampling_effort(diversity) # correct for sampling effort

print(summarise_diversity(diversity)) # quick summary
```

# Merge sites
```{r}
merged <- merge_sites(diversity, verbose = TRUE) 
```


# First, I will find the studies which contain data for multiple insect groups but subset at the source level, rather than the study level.

# The idea being that some sources may contain multiple separate studies with different sampling methods for different insect groups. 

# An example could be, within a source, Study 1: Malaise traps for Diptera, Study 2: Pitfall traps for Coleoptera, Study 3: Sweep nets for Lepidoptera, they will be seperate studies with different sampling methods within a source, but would have been conducted at the EXACT SAME set of sites and habitats. Therefore they are still precisely matched and suitable for my analysis.

# I may be missing data which has multiple groups at the source level but not at the study level, and this data would not be extracted if I subset ONLY at the study level. 

```{r}
source_multiple <- merged %>% 
  filter(Higher_taxon %in% c("Diptera", "Coleoptera", "Lepidoptera", "Hymenoptera", "Hemiptera")) %>% 
  group_by(Source_ID) %>%  # Subset multiple orders at the source level
  mutate(
    Multiple_groups = 
      case_when(
        (length(unique(Higher_taxon))) > 1 ~ "YES",
        (length(unique(Higher_taxon))) == 1 ~ "NO")) %>%  # when the number of unique higher taxon > 1, its a multi group study.
  filter(Multiple_groups == "YES") %>% 
  droplevels()

ss_multiple <- merged %>% 
  filter(Higher_taxon %in% c("Diptera", "Coleoptera", "Lepidoptera", "Hymenoptera", "Hemiptera")) %>% 
  group_by(SS) %>%  # Subset multiple orders at the study level
  mutate(
    Multiple_groups = 
      case_when(
        (length(unique(Higher_taxon))) > 1 ~ "YES",
        (length(unique(Higher_taxon))) == 1 ~ "NO")) %>%  
  filter(Multiple_groups == "YES") %>% 
  droplevels()


# Which data is subset at the source level but not at the study level?
source_only <- anti_join(source_multiple, ss_multiple, 
                by = c("Source_ID", "SS", "Order", "Sampling_method", "Site_name", "Habitat_as_described", "Latitude", "Longitude")) 
# matching based on EXACT SAME site_name, habitat_as_described and lat/long coordinates!

# Simplify source level df for merging
source_only <- source_only %>%
  mutate(site_habitat_combo = paste(Site_name, Habitat_as_described),# Concatenate site and habitat_as_described
         lat_long_combo = paste(Latitude, Longitude)) %>% # Concatenate lat and long
  dplyr::select(Source_ID, SS, Order, Sampling_method, site_habitat_combo, lat_long_combo) %>% # Select the columns of interest
  unique()
  
# Simplify the study level df for merging
studies_only <- ss_multiple %>%
  mutate(site_habitat_combo = paste(Site_name, Habitat_as_described),
         lat_long_combo = paste(Latitude, Longitude)) %>% 
  dplyr::select(Source_ID, SS, Order, Sampling_method, site_habitat_combo, lat_long_combo) %>%
  unique()

# Merge the studies_only df and source_only df only by rows that have an EXACT MATCH in their values for site_habitat_combo (i.e have the EXACT same sites and habitats)
same_site_habitat <- inner_join(studies_only, source_only, by = c("site_habitat_combo", "lat_long_combo")) 

# SS.x is subset at study level, SS.y is at the source level, here we can see which site x habitat combos match 
same_site_habitat %>% dplyr::select(SS.x, Order.x, Sampling_method.x, site_habitat_combo, SS.y, Order.y, Sampling_method.y) %>% unique() %>% View()

# the 10 additional studies subset at the source level 
same_site_habitat %>% dplyr::select(Source_ID.y, SS.y, Order.y, Sampling_method.y) %>% unique() %>% View()

# Check there's no missing lat/long or site/habitat data
sum(is.na(studies_only$lat_long_combo))
sum(is.na(source_only$site_habitat_combo))

# The unique studies in this column are the studies which have the EXACT same sites and habitats subset at the source level
data2add <- unique(same_site_habitat$SS.y) %>% droplevels()
print(data2add)

rm(same_site_habitat, source_only, studies_only)
```

# Merge source level with study level multi-group data

```{r}
# Filter for those additional studies subset at source level
source_multiple <- source_multiple %>% filter(SS %in% data2add) %>% droplevels() 

# Check its pulled the right studies
source_multiple %>% select(Source_ID, SS) %>% unique() %>% View()
print(data2add)

# Merge the additional 10 studies subset at the source level to the dataset subset at the study level
overall <- rbind(source_multiple, ss_multiple)

# Check
length(unique(source_multiple$SS))
length(unique(ss_multiple$SS))
length(unique(overall$SS)) # 113 studies overall

rm(source_multiple, ss_multiple, data2add, diversity)
```

# Now that I have my overall dataset (with multiple groups subset at both the source and study level)
# I will update the sampling methods for a number of studies based on checking the source publication. The idea being that some studies in my dataset have an obscure/unspecified sampling method (e.g. specimen collection, systematic searching, various, multiple, mixed trap etc).

```{r}
# First ungroup the df
overall <- overall %>% ungroup()

overall %>% select(SS, Sampling_method) %>% unique() %>%  View()  # have a look at the sampling methods for each study

# Update the sampling methods
overall <- overall %>%
  mutate(Sampling_method = case_when(
    SS == "AD1_2002__Vazquez 1" ~ "flower visitation observation",
    SS == "AD1_2009__Vergara 1" ~ "flower visitation observation",
    SS == "AD1_2011__Bates 1" ~ "Pan traps/sweep nets/hand searching", # multiple
    SS == "AD1_2011__Weiner 1" ~ "flower visitation observation",
    SS == "AD1_2013__Grass 1" ~ "flower visitation observation",
    SS == "AR1_2008__Basset 1" ~ "Malaise traps/yellow pan traps/pitfall traps", # multiple
    SS == "CC1_2005__Barratt 1" ~ "soil sample",
    SS == "CC1_2005__Barratt 2" ~ "soil sample",
    SS == "CC1_2008__Schon 1" ~ "soil cores along transects",
    SS == "CC1_2008__Schon 2" ~ "soil cores along transects",
    SS == "CC1_2010__Schon 2" ~ "soil cores along transects",
    SS == "CC1_2012__Waite 2" ~ "bole traps/flight intercept trap/foliage sampling", # multiple
    SS == "CC1_2013__Litchwark 1" ~ "flower visitation observation",
    SS == "CC1_2014__Rader 1" ~ "flight intercept trap/pan trap", # multiple
    SS == "DG1_2012__Ge 1" ~ "soil cores along transects",
    SS == "GP1_2009__Vasconcelos 2" ~ "litter collection",
    SS == "HP1_2006__Lachat 1" ~ "Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap", # multiple
    SS == "MH1_2009__Turner 1" ~ "Leaf litter sampling/ferns cut and bagged/fogging for canopy species", # multiple
    SS == "SC1_2005__Richardson 1" ~ "litter collection",
    SS == "SJ2_2019__Eckert 1" ~ "pit-fall traps", # multiple "Pitfall trapping/direct sampling (i.e. active searching) within quadrats" but didn't include direct sampling as not a PREDICTS sampling method
    SS == "VB1_2005__Eggleton 2" ~ "Hand sorting in quadrat and winkler bags", # multiple
    SS == "VB1_2012a_Carpenter 1" ~ "Soil cores/pitfall traps", # multiple
    SS == "VB1_2012__LeightonGoodall 1" ~ "Leaf litter sampling/soil sampling", # multiple
    SS == "GP1_2009__Vasconcelos 1" ~ "litter collection", # this study was subset at source level
    SS == "VB1_2012__Carpenter 5" ~ "soil sampling/litter collection/pitfall traps", # multiple - this study was subset at source level
    SS == "AD1_2006__Blanche 1" ~ "flower visitation observation",
    SS == "AD1_2006__Blanche 2" ~ "flower visitation observation",
    SS == "AD1_2007__Meyer 2" ~ "flower visitation observation",
    SS == "SH1_2011__Todd 1" ~ "flight intercept/pan trap/pitfall trap", # multiple
    SS == "VB1_2020__Burton 2" ~ "litter collection", # multiple
    TRUE ~ Sampling_method
  ))

# Recheck
overall %>% select(SS, Order, Sampling_method) %>% unique() %>%  View()

# Merge similar methods, e.g. sweep netting and sweep net.
overall <- overall %>%
mutate(Sampling_method = recode_factor(Sampling_method,
                                       "sweep netting" = "sweep net", # Combine sweep net and sweep netting
                                       "sweep_netting" = "sweep net",
                                       "soil core" = "soil sample",
                                      "soil cores along transects" = "soil sample", # Combine soil cores, soil cores along transect and soil sample
                                      "mixed traps" = "multiple", # Combine mixed traps with multiple traps
                                      "window trap" = "flight trap" # Combine window trap with flight trap as they are the same thing
                                      ))
# Check unique sampling methods
unique(overall$Sampling_method)
```

# Some studies sample larvae, speaking with Gavin Broad (Entomologist at the NHM), you would only expect to sample larvae in the soil, therefore all other instances of larvae caught with any sampling method other than soil sampling are bycatch. I will filter PREDICTS only for larvae sampled in the soil below.

```{r}
# Lets have a look at all instances of larvae
overall %>% 
  filter(grepl('larv|immatu|pupa|nymphe|nymph_', Taxon_name_entered, ignore.case = T)) %>% # different words for larvae 
  dplyr::select(SS, Higher_taxon, Taxon_name_entered, Order, Genus, Species, Sampling_method) %>% 
  filter(Higher_taxon %in% c("Diptera", "Coleoptera", "Lepidoptera", "Hymenoptera", "Hemiptera")) %>%  # orders relevant to my analysis
  unique %>% 
  View() 

# Larvae_keep contains data only on larvae where they have been sampled appropriately in the soil.
larvae_keep <- overall %>%  
  filter(grepl('larv|immatu|pupa|nymphe|nymph_', Taxon_name_entered, ignore.case = T) & Sampling_method %in% 
           c("soil sample",
             "soil sampling/litter collection/pitfall traps",
             "Soil cores/pitfall traps",
             "Leaf litter sampling/soil sampling")) %>%  
              droplevels()

# Those studies that have multiple sampling methods including soil sampling, do not have many instances of larvae so it should not be a problem to include them.
```

# Now, based on the expert spreadsheet sent over by Andrew (containing the sampling methods used by experts for their order(s) of interest). 
# I will, for each sampling method, filter for the orders which are not bycatch.

```{r}

# active visual surveys good for all five orders
active_visual_surveys <- overall %>% filter(Order %in% c("Lepidoptera", "Coleoptera", "Diptera", "Hymenoptera", "Hemiptera") 
                                            & Sampling_method %in% c("sweep net", 
                                                                     "aerial transect", 
                                                                     "flower visitation observation", 
                                                                     "visual encounter survey", 
                                                                     "systematic searching")) 
# these five PREDICTS methods constitute active visual surveys in the expert spreadsheet.


# light trap - good for all except Hemiptera
light_trap <- overall %>% filter(Order %in% c("Lepidoptera", "Coleoptera", "Diptera", "Hymenoptera") & Sampling_method == "light trap")

 # pan trap - good for all (Lepidoptera added after reverification 17/01)
pan_trap <- overall %>% filter(Order %in% c("Coleoptera", "Diptera", "Hymenoptera", "Hemiptera", "Lepidoptera") & Sampling_method %in% c("pan traps", 
                                                                                                "Pan traps/sweep nets/hand searching",
                                                                                                "Malaise traps/yellow pan traps/pitfall traps",
                                                                                                "flight intercept trap/pan trap",
                                                                                                "flight intercept/pan trap/pitfall trap"))

# You can see for pan traps there are instances of multiple sampling methods for a given study, when I merge all of the cleaned methods I will pass a function that will remove duplicate rows, so this won't be an issue.

# flight intercept / harp trap - good for Hymenoptera, Coleoptera and Diptera (added after reverification 17/01)
flight_trap <- overall %>% filter(Order %in% c("Coleoptera", "Hymenoptera", "Diptera") & Sampling_method %in% 
                                    c("flight trap",
                                      "bole traps/flight intercept trap/foliage sampling",
                                      "flight intercept trap/pan trap",
                                      "Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap",
                                      "flight intercept/pan trap/pitfall trap",
                                      "harp trap"))

# Malaise trap - good for all except Lepidoptera
malaise_trap <- overall %>% filter(Order %in% c("Coleoptera", "Diptera", "Hymenoptera", "Hemiptera") & Sampling_method %in% c("malaise traps",
                                                                                                          "Malaise traps/yellow pan traps/pitfall traps",
                                                                  "Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap"))


# Winkler bags or litter collection good for beetles and ants, other orders bycatch 
winkler_beetles <- overall %>% filter(Order == "Coleoptera" & Sampling_method %in% c("litter collection", 
                                                                                     "soil sampling/litter collection/pitfall traps",
                                                                                     "Hand sorting in quadrat and winkler bags", 
                                                                                     "Leaf litter sampling/ferns cut and bagged/fogging for canopy species",
                                                                                     "Leaf litter sampling/soil sampling"))

# Have to create a seperate object for ants in order to filter for Formicidae 
winkler_ants <- overall %>% filter(Order == "Hymenoptera" & Family == "Formicidae" & Sampling_method %in% c("litter collection", 
                                                                                     "soil sampling/litter collection/pitfall traps",
                                                                                     "Hand sorting in quadrat and winkler bags", 
                                                                                     "Leaf litter sampling/ferns cut and bagged/fogging for canopy species",
                                                                                     "Leaf litter sampling/soil sampling"))

# All expert methods have now been cleaned for non-bycatch data.
```

# Some of the PREDICTS sampling methods could not be assigned to an expert sampling method.
# These include - soil sampling, trap nests, bole trap, suction samplers, live traps and baited traps.
# Where a study had multiple sampling methods, if any of those multiple sampling methods was suitable for any of the groups then it was included.
# I will clean these methods based on comms with Victoria Burton, Gavin Broad and Ben Woodcock. 
```{r}
# Soil sample only good for ants, all other groups bycatch.
soil_sampling <- overall %>% 
  filter(Order == "Hymenoptera" & Family == "Formicidae" & Sampling_method %in% c("soil sample",
                                                                                  "soil sampling/litter collection/pitfall traps",
                                                                                  "Soil cores/pitfall traps",
                                                                                  "Leaf litter sampling/soil sampling"))


# Glue trap good for Hymenoptera, Diptera and Hemiptera, other orders bycatch.
glue_trap <- overall %>% filter(Order %in% c("Hymenoptera", "Diptera", "Hemiptera") & Sampling_method == "glue traps")

# Bole trap only good for Coleoptera and Hemiptera
bole_trap <- overall %>%  filter(Order %in% c("Coleoptera", "Hemiptera") & Sampling_method %in% c("bole trap", 
                                                                           "bole traps/flight intercept trap/foliage sampling"))


# BW says suction samplers good for Coleoptera and Hemiptera but not Hymenoptera and Lepidoptera.
suction <- overall %>% filter(Order %in% c("Coleoptera", "Hemiptera") & Sampling_method == "suction samplers")

# Trap nests only good for Hymenoptera says VB.
trap_nests <- overall %>% filter(Order == "Hymenoptera" & Sampling_method == "trap nests")

# Pitfall traps good for only Coleoptera and ants says VB (surprisingly not good for Hemiptera) - includes baited-pitfall too
pitfall_beetles <- overall %>% filter(Order == "Coleoptera" & 
                                        Sampling_method %in% c("Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap",
                                                               "Malaise traps/yellow pan traps/pitfall traps", 
                                                               "Soil cores/pitfall traps", 
                                                               "soil sampling/litter collection/pitfall traps", 
                                                               "pit-fall traps", 
                                                               "baited pit-fall traps", 
                                                               "flight intercept/pan trap/pitfall trap"))

pitfall_ants <- overall %>% filter(Order == "Hymenoptera" & Family == "Formicidae" 
                                   & Sampling_method %in% c("Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap",
                                                               "Malaise traps/yellow pan traps/pitfall traps", 
                                                               "Soil cores/pitfall traps", 
                                                               "soil sampling/litter collection/pitfall traps", 
                                                               "pit-fall traps", 
                                                               "baited pit-fall traps", 
                                                               "flight intercept/pan trap/pitfall trap"))


# Below is code to filter the overall dataset for the levels of a given sampling method, in this example pitfall traps.
overall %>%  
  filter(grepl('sort', Sampling_method, ignore.case = T)) %>% 
  droplevels() %>% 
  dplyr::select(Sampling_method) %>% 
  unique() %>% 
  View()

# 2023 Smith is the only baited-traps paper, Diptera and Hymenoptera are OK (Hymenoptera > 20% of the individuals) but Coleoptera, Hemiptera, and Lepidoptera bycatch. 
baited_trap <- overall %>% filter(Order %in% c("Diptera", "Hymenoptera") & Sampling_method == "baited traps") 

# Live traps (TN_1_2007_Bouyer 2 and TN_1_2007_Bouyer 3) all taxa are targeted - checked source paper.
live_trap <- overall %>% filter(Sampling_method == "live traps")

# All non-expert methods have now been cleaned for non-bycatch data.
```

# Now that I've removed bycatch from all of the sampling methods in PREDICTS, I will combine them all with the larvae_keep data.
```{r}
clean <- bind_rows(active_visual_surveys,
                   baited_trap,
                   bole_trap,
                   flight_trap,
                   glue_trap,
                   larvae_keep,
                   light_trap,
                   live_trap,
                   malaise_trap,
                   pan_trap,
                   pitfall_ants,
                   pitfall_beetles,
                   soil_sampling,
                   suction,
                   trap_nests,
                   winkler_ants,
                   winkler_beetles) 
# ^ each of these R objects is a sampling method in PREDICTS which has been filtered only for the orders which are NOT bycatch.

# Remove duplicate rows as some data will have been duplicated due to the studies with multiple sampling methods.
clean <- unique(clean)

# Relevel the 'Order' variable alphabetically
clean$Order <- factor(clean$Order, levels = sort(levels(clean$Order)))

# Relevel the 'Higher_taxon' variable alphabetically
clean$Higher_taxon <- factor(clean$Higher_taxon, levels = sort(levels(clean$Higher_taxon)))

# check its worked
levels(clean$Order)
levels(clean$Higher_taxon)

# Tidy the environment
rm(active_visual_surveys,baited_trap,bole_trap,flight_trap,glue_trap,larvae_keep,light_trap,live_trap,malaise_trap,pan_trap,pitfall_ants,pitfall_beetles,soil_sampling,suction,trap_nests,winkler_ants,winkler_beetles)
```

# Lets have a look at the appropriate sampling methods for each of the Orders.
```{r}
clean %>% group_by(Order) %>% reframe(Sampling_method) %>% unique() %>% View() 

table(clean$Order, clean$Predominant_habitat)

levels(clean$Sampling_method) # 31 methods in total
```

# I will see if there are any sampling methods which are VALID for an Order if it was used in ANY study where that 
# Order is the Order contributing most individuals to the data (not just the studies that have multiple orders)
```{r}
# This code lets me see which studies caught a lot of individuals for a given order 
merged %>% 
  filter(Class == "Insecta") %>% 
  dplyr::select(Source_ID, SS, Order, Sampling_method, Taxon_name_entered) %>% 
  unique() %>% 
  group_by(SS, Order, Sampling_method) %>% 
  summarise(Count = n()) %>% 
  filter(Order == "Hemiptera") %>% 
  View()
```


##%######################################################%##
#                                                          #
####                 Prepare model data                 ####
#                                                          #
##%######################################################%##

### Load add_order_fold function to split the cleaned  data into exclusive insect orders 
```{r}

add_order_fold <- function(data, folds = "default", verbose = TRUE){
  
  # By default (and nothing else is yet coded up), this puts each row into
  # an insect order that each have a decent, number of rows of data. 
  # This allows site-level values to be obtained within
  # each of the order folds in turn.
  
  # Set up order_fold, a factor that holds the fold identity.
  data$order_fold <- rep(NA, nrow(data))
  
  # Rows are assigned to folds based on the higher taxonomy information.
  if (folds == "default"){
    data$order_fold[data$Higher_taxon == "Diptera"] <- "Diptera"
    data$order_fold[data$Higher_taxon == "Coleoptera"] <- "Coleoptera"
    data$order_fold[data$Higher_taxon == "Lepidoptera"] <- "Lepidoptera"
    data$order_fold[data$Higher_taxon == "Hymenoptera"] <- "Hymenoptera"
    data$order_fold[data$Higher_taxon == "Hemiptera"] <- "Hemiptera"
    data$order_fold[data$Higher_taxon == "Arachnida"] <- "Arachnida"
    data$order_fold[data$Higher_taxon == "Odonata" |
                    data$Higher_taxon == "Trichoptera" |
                    data$Higher_taxon == "Ephemeroptera" |
                    data$Higher_taxon == "Plecoptera" ] <- "Aquatics"
    data$order_fold[data$Phylum == "Arthropoda" & 
                      data$Higher_taxon != "Diptera" & 
                      data$Higher_taxon != "Coleoptera" &
                      data$Higher_taxon != "Lepidoptera" &
                      data$Higher_taxon != "Hymenoptera" &
                      data$Higher_taxon != "Hemiptera" &
                      data$Higher_taxon != "Arachnida" &
                      data$Higher_taxon != "Odonata" &
                      data$Higher_taxon != "Trichoptera" &
                      data$Higher_taxon != "Ephemeroptera" &
                      data$Higher_taxon != "Plecoptera"] <- "OtherArthropods"
  }else{
    stop("Only the default folds have been coded up so far!")
  }
  
  data$order_fold <- as.factor(data$order_fold)
  
  if (verbose == TRUE){
    # By default, report on the breakdown of records per fold.
    cat("Numbers of rows within each order fold are as follows:\n\n")
    print(table(data$order_fold))
  }
  
  return(data)
}

```

### Load get_site_abundances_order function to calculate total abundance for each site x order_fold combination
```{r}
# Calculate total abundance for each site --------------------------------------
# Modified to clarify reporting is on site x fold combinations
get_site_abundances_order <- function(data, 
                                order_folds = FALSE,
                                verbose = TRUE){
  # This calculates site metrics for each site or, if desired, each order fold
  # within each site.
  
  # Make it a data frame to avoid having to deal with a tibble.
  data <- as.data.frame(data)
  
  # Set up a variable for uniquely identifying the site x order_fold combination
  # or, if order_folds == FALSE, the site. This is what grouping is then done
  # by.
  print(table(data$order_fold))
  
  if (order_folds == TRUE){
    data$the_order_folds <- data$order_fold
  }else{
    data$the_order_folds <- ""
  }
  
  # Just for debugging
  print(table(data$the_order_folds))
  
  # Start to calculate site-level metrics.
  sites <- data %>%
    
    # pull out only the merged diversity data
    distinct(merge_ID, .keep_all = TRUE) %>%
    
    # Re-make SSB and SSBS values since we've now dropped a bunch of values; and
    # also add SSBST which is that order x site combo. The trimws is because the
    # SSBST would otherwise end in a space if order_folds == FALSE.
    mutate(SS = paste(Source_ID, Study_number),
           SSB = paste(SS, Block),
           SSBS = paste(SSB, Site_number),
           SSBST = trimws(paste(SSBS, the_order_folds))) %>%
    
    # group by SSBST (each unique value corresponds to a unique site x fold)
    group_by(SSBST) %>%
    
    # Now add up all the abundance measurements within each site and also 
    # estimate coverage.
    mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                   sum(merged_diversity, na.rm = TRUE),
                                   # If the diversity metric type isn't 
                                   # Abundance, then leave the TotalAbundance 
                                   # measurement as NA.
                                   NA),
           
           SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                    merged_diversity,
                                    # for abundance and occurrence measurements,
                                    # count the number of unique species names 
                                    # that are present at the site 
                                    n_distinct(Best_guess_binomial[merged_diversity > 0])),
           
           coverage = get.SC2(merged_diversity)) %>%
    
    # ungroup
    ungroup() %>%
    
    # pull out unique site/fold combinations.
    distinct(SSBST, .keep_all = TRUE) %>%
    
    # now group by Study ID and fold.
    group_by(SS, the_order_folds) %>%
    
        # pull out the maximum abundance for each study/fold combination.
    mutate(MaxAbundance = max(TotalAbundance),
           MinNonZero = min(TotalAbundance[TotalAbundance > 0])) %>%
    
    # ungroup.
    ungroup() %>%
    
    # now rescale total abundance, so that within each study, the maximum is 1.
    mutate(RescaledAbundance = TotalAbundance/MaxAbundance) %>%
    
    mutate(sqrtRescaledAbundance = sqrt(RescaledAbundance)) 
  
    # If a study reports abundance as a count, add 1 to all values; otherwise
  # add 1/2 the minimum non-zero value; then log-transform to get a useful
  # logAbundance value that may be more symmetric than sqrtRescaledAbundance.
  sites$to_add <- ifelse(sites$Diversity_metric_unit %in%
                           c("individuals", "times observed"), 1,
                         sites$MinNonZero/2)
  sites$logAbundance <- log(sites$TotalAbundance + sites$to_add)
  
  
  # Remove NA values and drop levels
  sites <- as.data.frame(sites)
  sites <- droplevels(subset(sites, !is.na(sqrtRescaledAbundance)))
  
  if (verbose == TRUE){
    # Summarise structure of site-level data frame
    cat("\nThere are abundance data for:\n")
    cat(paste("   ", sum(!is.na(sites$sqrtRescaledAbundance)), "sites, from\n"))
    cat(paste("   ", 
              length(unique(sites$SS[!is.na(sites$sqrtRescaledAbundance)])), 
              "studies and\n"))
    cat(paste("   ", 
              length(unique(sites$Source_ID[!is.na(sites$sqrtRescaledAbundance)])),
              "sources.\n"))
  }
  
  # Set provenance attributes
  attr(sites, which = "diversity_extract") <- 
    attr(data, which = "diversity_extract")
  attr(sites, which = "is_merged") <- attr(data, which = "is_merged")
  attr(sites, which = "order_folds") <- order_folds 
  attr(sites, which = "when_Created") <- Sys.time()
  
  return(sites)
}

```

### Add_order_fold to allow taxonomic cross-validation 
```{r}
# Just adds order_fold column
clean <- add_order_fold(clean, verbose = TRUE)
levels(clean$order_fold)

# Col = 159k, Dip = 26k, Hemi = 25k, Hym = 71k, Lep = 40k records.
```

### Calculate site level diversity metrics using get_site_abundances_order and order_fold for each site x order combination
```{r}

clean <- get_site_abundances_order(clean, order_folds = TRUE, verbose = TRUE)

table(clean$order_fold) # number of rows for each site x order combination

# 5453 sites, from 102 studies and 68 sources.
```

### Fix up your explanatory variables 
```{r}

# Rename predominant habitat since its not really habitat we're looking at
clean <- rename(clean,
                Predominant_land_use = Predominant_habitat)


# Relevel the land use and use intensity classes
clean <- clean %>%
  
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    
    # indeterminate secondary veg and cannot decide get NA
   # Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"), - removed as we are collapsing secondary
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(clean$Predominant_land_use, clean$Use_intensity)
```

### Collinearity
```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(clean[ , c("Predominant_land_use", "Use_intensity")])
```

### Complete cases
```{r}

# This will drop all the rows with NAs for abundance, land use or use intensity so we have complete data for modelling
model_data_ab <- drop_na(clean, 
                         logAbundance, Predominant_land_use,
                         Use_intensity)

model_data_sr <- drop_na(clean, 
                         SpeciesRichness, Predominant_land_use,
                         Use_intensity)

sum(is.na(clean$logAbundance))
sum(is.na(clean$SpeciesRichness))
sum(is.na(clean$Predominant_land_use)) 
# There are no NAs for either of our explanatory variables so model_data_ab and model_data_sr are the same dataframe
```

### Prepare the abundance model dataframe
```{r}

# No longer needed as lme4 has improved, can use logAbundance provided in get_site_abundances()
# model_data_ab$logRescaledAbundance <- log(model_data_ab$RescaledAbundance + 1)
# sum(model_data_ab$RescaledAbundance == 1)

# Check the distributions
hist(model_data_ab$RescaledAbundance) ; hist(model_data_ab$logAbundance) ; hist(model_data_ab$sqrtRescaledAbundance)

plot(density(model_data_ab$logAbundance)) 

# Have a look at sample sizes
table(model_data_ab$Predominant_land_use, model_data_ab$Use_intensity)

# Collapse intense and light land use intensities to have a more even spread of the data
model_data_ab <- 
  model_data_ab %>%
  mutate(
    Use_intensity = recode_factor(Use_intensity,
                                  "Intense use" = "High_use",
                                  "Light use" = "High_use"),
# reset reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )


table(model_data_ab$Predominant_land_use, model_data_ab$Use_intensity) # Only MSV HIGH USE (n = 73) and Pasture MINIMAL USE (n = 85) less than 100.

# Order / land use split
table(model_data_ab$Order,  model_data_ab$Predominant_land_use)

# Relevel the land use and use intensity classes
model_data_ab <- model_data_ab %>%
  
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Young secondary vegetation" = "Secondary vegetation", 
                                         "Intermediate secondary vegetation" = "Secondary vegetation",
                                         "Mature secondary vegetation" = "Secondary vegetation",
                                         "Secondary vegetation (indeterminate age)" = "Secondary vegetation"),
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(model_data_ab$Predominant_land_use, model_data_ab$Use_intensity) # Only pasture minimal use (n = 85) is less than 100 sites.

# Lets have a look at the order x land use split again
table(model_data_ab$Order, model_data_ab$Predominant_land_use) 

# Best way to proceed: Mick Crawley reckons an average of 30 data points to fit a parameter is comfy, 10 might work, and 3 if you're as lucky as a human can be. On that basis, I think you'll be fine - but obviously some combinations of order 1, order 2 and land use will be based on very few data indeed (so will have suitably wide CIs).

# Lets look at Orders sampled under the same LUs and UIs
model_data_ab %>% dplyr::select(Source_ID, SS, Order, Sampling_method, Predominant_land_use, Use_intensity) %>% unique() %>% View()

# Are there any studies which only compare different stages of secondary vegetation? - as these would stop being informative now that I've collapsed secondary
# This code groups the data by SS and then filters for studies where "all" specified SV LUs are present. 
clean %>%
  group_by(SS) %>% 
  filter(all(c("Young secondary vegetation", "Intermediate secondary vegetation", "Mature secondary vegetation") %in%   Predominant_land_use)) %>%  
  dplyr::select(SS, Predominant_land_use) %>% 
  unique() %>% 
  View()
```

# Lets summarise the cleaned model dataset (without bycatch) now that we've prepped it for modelling
```{r}
length(unique(merged$SS)) # 975 studies in all of PREDICTS
length(unique(overall$SS)) # 113 for multi groups
length(unique(overall$SSBS)) # 4189 SITES for multi groups
# 106 SS for non bycatch multi groups before model prep
length(unique(clean$SS)) # 102 ss after model prep
length(unique(clean$SSBS)) # 2580 sites after model prep
```


##%######################################################%##
#                                                          #
####               Order comparisons for                ####
####         abundance, modelling and plotting          ####
#                                                          #
##%######################################################%##

# Load functions fot plotting and predicting effects
```{r}
source("/Applications/PhD/Projects/multi-groups/scripts/combine_and_plot_function.R")
source("/Applications/PhD/Projects/multi-groups/scripts/predict_this_model.R")
```

# diptera_hymenoptera abundance prepare df
```{r}
# Filter for studies that only compare hymenoptera and diptera
dip_hym_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Filter our abundance model dataset for those studies:
diptera_hymenoptera <- model_data_ab %>% filter(SS %in% dip_hym_ss) %>% 
  # then filter the two orders we want to compare, as some studies will have include more than those two orders
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  droplevels()


table(diptera_hymenoptera$Predominant_land_use)

```

# diptera_hymenoptera linear mixed effects modelling
```{r}
# maximal model
dhy_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hymenoptera)

# remove block
dhy_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# remove study
dhy_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_hymenoptera)

# Compare models 
AIC(dhy_m1, dhy_m2, dhy_m3) # dhy_m2 best model

rm(dhy_m1, dhy_m3)

Anova(dhy_m2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dhy_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just land use as predictor
dhy_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just Higher_taxon as the predictor
dhy_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Make a null, intercept only model
dhy_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Compare the models
AIC(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7)
model.sel(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7)

# dhy_m2 still the best

# model estimates
summary(dhy_m2)
anova(dhy_m2)

```

### Simulate diptera_hymenoptera effects (total abundance % difference)
```{r}

# Predict Diptera effects
dip_eff <- predict_this_model(dhy_m2, order = "Diptera", percentage_diff = "YES", baseline = "percentage")

# relevel model with hymenoptera as reference level
diptera_hymenoptera <- within(diptera_hymenoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model
hym_1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# Predict hymenoptera effects
hym_eff <- predict_this_model(hym_1, order = "Hymenoptera", , percentage_diff = "YES", baseline = "percentage")

# Combine effects and plot them
diptera_hymenoptera_plot_AB <- combine_and_plot(dip_eff, hym_eff, diversity = "Abundance")

# Extract f ratio, df and p value
anova(dhy_m2)

# Check residuals
hist(residuals(dhy_m2))

# Check r2
r.squaredGLMM(dhy_m2)

# Extract correlation coefficients
corr_dip_hym <- as.matrix(summary(dhy_m2)$coefficients)

# Split the coefficients into the two orders
diptera <- corr_dip_hym[2:6, 1]
hymenoptera <- corr_dip_hym[8:nrow(corr_dip_hym), 1] 

# Plot the coefficients
plot(diptera, hymenoptera)
cor(diptera, hymenoptera)

rm(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7, hym_1, hym_eff, diptera_hymenoptera, dip_eff, corr_dip_hym, diptera, hymenoptera)
```


# Log scale plotting and modelling 



# diptera_hymenoptera abundance prepare df
```{r}
# Filter for studies that only compare hymenoptera and diptera
dip_hym_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Filter our abundance model dataset for those studies
diptera_hymenoptera <- model_data_ab %>% filter(SS %in% dip_hym_ss) %>% 
  # then filter the two orders we want to compare, as some studies will have include more than those two orders
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  droplevels()

# Number of sites
table(diptera_hymenoptera$Predominant_land_use)
```

# diptera_hymenoptera linear mixed effects modelling
```{r}
# maximal model
dhy_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hymenoptera)

# remove block
dhy_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# remove study
dhy_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_hymenoptera)
  
# Compare models 
AIC(dhy_m1, dhy_m2, dhy_m3) # dhy_m2 best model

rm(dhy_m1, dhy_m3)

Anova(dhy_m2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dhy_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just land use as predictor
dhy_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just Higher_taxon as the predictor
dhy_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Make a null, intercept only model
dhy_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Compare the models
AIC(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7)
model.sel(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7)

# dhy_m2 still the best

# model estimates
summary(dhy_m2)
anova(dhy_m2)

```

### Simulate diptera_hymenoptera effects
```{r}
# Predict Diptera effects
dip_eff <- predict_this_model(dhy_m2, order = "Diptera", baseline = "median")

# relevel model with hymenoptera as reference level
diptera_hymenoptera <- within(diptera_hymenoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# Check its worked
levels(diptera_hymenoptera$Higher_taxon)

# maximal model
hym_1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# Predict hymenoptera effects
hym_eff <- predict_this_model(hym_1, order = "Hymenoptera", baseline = "median")

# Plot effects on the log abundance scale 

# Combine effects and plot them
diptera_hymenoptera_plot_AB_log <- combine_and_plot(dip_eff, hym_eff, diversity = "Log_Abundance")

# Extract f ratio, df and p value
anova(dhy_m2)

# Check residuals
hist(residuals(dhy_m2))

# Check r2
r.squaredGLMM(dhy_m2)

# Extract correlation coefficients
corr_dip_hym <- as.matrix(summary(dhy_m2)$coefficients) 

dip <- corr_dip_hym[2:6,1]
hym <- corr_dip_hym[8:nrow(corr_dip_hym), 1]

# Plot the coefficients
plot(dip, hym)
cor(dip, hym)

# regression through the origin
m <- lm(dip ~ hym - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # coef = - 0.41

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.93
r <- sqrt(r2) # 0.96

# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # - 0.96 corr through the origin


rm(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7, hym_1, hym_eff, diptera_hymenoptera, dip_eff, corr_dip_hym, dip, hym, m, r2, r, a)
```



# diptera_coleoptera abundance prepare df
```{r}
# Studies
dip_col_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Coleoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
diptera_coleoptera <- 
  model_data_ab %>% 
  filter(SS %in% dip_col_ss) %>% 
  filter(Order %in% c("Diptera", "Coleoptera")) %>% 
  droplevels()

# Number of sites
table(diptera_coleoptera$Predominant_land_use)
```

# diptera_coleoptera linear mixed effects modelling
```{r}
## Backwards stepwise selection of random effects 

# maximal model
dc_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_coleoptera)

# remove block
dc_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_coleoptera)

# remove study
dc_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_coleoptera)

# Compare models 
AIC(dc_m1, dc_m2, dc_m3) # dc_m2 best model

rm(dc_m1, dc_m3)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dc_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_coleoptera)

## Just land use as predictor
dc_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = diptera_coleoptera)

## Just Higher_taxon as the predictor
dc_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_coleoptera)

## Make a null, intercept only model
dc_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS), data = diptera_coleoptera)

## Compare the models
AIC(dc_m2, dc_m4, dc_m5, dc_m6, dc_m7) # dcm6 best model without interaction and only higher_taxon as predictor which suggests interaction isn't important nor is land use as a predictor.
model.sel(dc_m2, dc_m4, dc_m5, dc_m6, dc_m7)

# Proceed with model 2 to plot interaction
summary(dc_m2) # yep no significant interactions
```

### Simulate diptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(dc_m2, order = "Coleoptera")

# relevel model with Diptera as reference level
diptera_coleoptera <- within(diptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Diptera"))

# maximal model
dip_m2 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_coleoptera)

# Predict Diptera effects - remember that this overwrites previous effects of Diptera model
dip_eff <- predict_this_model(dip_m2, order = "Diptera")

# Combine effects and relevel Diptera as reference
diptera_coleoptera_plot_AB_log <- combine_and_plot(dip_eff, col_eff, relevel = "Diptera", diversity = "Log_Abundance")

# Extract f ratio, df and p value
anova(dc_m2)
anova(dip_m2)

# Check residuals
hist(residuals(dc_m2))

# Check r2
r.squaredGLMM(dc_m2)
r.squaredGLMM(dip_m2)

# Extract correlation coefficients with each order as the reference
corr_col_ref <- as.matrix(summary(dc_m2)$coefficients)
corr_dip_ref <- as.matrix(summary(dip_m2)$coefficients)

# Split the coefficients into two groups
# coleoptera <- corr[2:6, 1] 
# diptera <- corr[8:nrow(corr), 1] 

# Split the coefficients into two groups based on each order as the reference
coleoptera <- corr_col_ref[2:6, 1] 
diptera <- corr_dip_ref[2:6, 1]  

plot(coleoptera, diptera)
cor(coleoptera, diptera) # 0.18

# regression through the origin
m <- lm(coleoptera ~ diptera - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # coef = 0.05

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.005
r <- sqrt(r2) # 0.08
D
# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # r = 0.08

rm(col_eff, corr, dc_m2, dc_m4, dc_m5, dc_m6, dc_m7, dip_eff, dip_m2, coleoptera, diptera, diptera_coleoptera, m, r2, r, a, corr_col_ref, corr_dip_ref)
```



# diptera_hemiptera abundance prepare df
```{r}
# Studies
dip_hemi_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hemiptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
diptera_hemiptera <- model_data_ab %>% filter(SS %in% dip_hemi_ss) %>% filter(Order %in% c("Diptera", "Hemiptera"))
```

# diptera_hemiptera linear mixed effects modelling
```{r}
# maximal model
dh_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hemiptera)

# remove block
dh_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hemiptera)

# remove study
dh_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_hemiptera)

# Compare models 
AIC(dh_m1, dh_m2, dh_m3) # dh_m1 best model

rm(dh_m2, dh_m3)

# model estimates
summary(dh_m1)
Anova(dh_m1)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dh_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = diptera_hemiptera)

## Just land use as predictor
dh_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS) + (1 | SSB), data = diptera_hemiptera)

## Just Higher_taxon as the predictor
dh_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = diptera_hemiptera)

## Make a null, intercept only model
dh_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS) + (1 | SSB), data = diptera_hemiptera)

## Compare the models
AIC(dh_m1, dh_m4, dh_m5, dh_m6, dh_m7)
model.sel(dh_m1, dh_m4, dh_m5, dh_m6, dh_m7) # dh_m1 still best with interaction
```

# Simulate diptera effects and plotting, but because there is no plantation data, I will do this manually.
```{r}
summary(dh_m1) # no coefficient for plantation as there is no data

# Predict diptera effects

# gather the effects and confidence intervals using simulation. Simulate 1000 times
  dip_eff <- FEsim(dh_m1, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  dip_eff$term <- as.factor(dip_eff$term)
  
  # pull out effects
  dip_eff <- dip_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  dip_eff$term <- recode_factor(dip_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  dip_eff <- dip_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(dip_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  dip_eff <- dip_eff %>% mutate(Order = "Diptera")
  
  # make the order column a factor so it can be recoded
  dip_eff$Order <- as.factor(dip_eff$Order)

    
# relevel model with Hemiptera as reference level
diptera_hemiptera <- within(diptera_hemiptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model with Hemiptera as ref
hemi_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hemiptera)

summary(hemi_m1)
Anova(hemi_m1)

# Predict Hemiptera effects 

# gather the effects and confidence intervals using simulation. Simulate 1000 times using the re leveled model
  hemi_eff <- FEsim(hemi_m1, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  hemi_eff$term <- as.factor(hemi_eff$term)
  
  # pull out effects
  hemi_eff <- hemi_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  hemi_eff$term <- recode_factor(hemi_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hemi_eff <- hemi_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(hemi_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  hemi_eff <- hemi_eff %>% mutate(Order = "Hemiptera")
  
  # make the order column a factor so it can be recoded
  hemi_eff$Order <- as.factor(hemi_eff$Order)

# Combine effects
combined <- rbind(dip_eff, hemi_eff)


comb_plot_limits <- c("Primary", "Secondary vegetation", "Pasture", "Cropland", "Urban") # LU categories for plotting

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845")
  
  
# Plot the effects for both orders
diptera_hemiptera_plot_AB_log <- 
    combined %>%
  ggplot()+
        aes(x = Predominant_land_use, y = median, colour = Order, group = Order)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Upper_ci, ymax = Lower_ci), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        scale_x_discrete(limits=comb_plot_limits)+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        #geom_vline(xintercept = 5.5, linetype = 2)+
        #geom_vline(xintercept = 6.5, linetype = 2)+
        #geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("Log abundance") +
        scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 8)))
  

# Extract f ratio, df and p value
anova(dh_m1)

# Check residuals
hist(residuals(dh_m1))

# Check r2
r.squaredGLMM(dh_m1)

# Extract correlation coefficients
corr_dip_ref <- as.matrix(summary(dh_m1)$coefficients)
corr_hemi_ref <- as.matrix(summary(hemi_m1)$coefficients)

# Split the coefficients into two groups
# diptera <- corr[2:5, 1] 
# hemiptera <- corr[7:nrow(corr), 1] 

diptera <- corr_dip_ref[2:5, 1]
hemiptera <- corr_hemi_ref[2:5, 1]

plot(diptera, hemiptera)
cor(diptera, hemiptera) # - 0.97

# regression through the origin
m <- lm(diptera ~ hemiptera - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # coef = - 0.91

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.946
r <- sqrt(r2) # 0.97

# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # r = - 0.97

rm(combined, corr, dh_m1, dh_m4, dh_m5, dh_m6, dh_m7, dip_eff, diptera_hemiptera, hemi_eff, hemi_m1, diptera, hemiptera, corr_dip_ref, corr_hemi_ref, m, r2, r, a)
```


# diptera_lepidoptera abundance prepare df
```{r}
# Studies
dip_lep_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
diptera_lepidoptera <- model_data_ab %>% filter(SS %in% dip_lep_ss) %>% filter(Order %in% c("Diptera", "Lepidoptera")) 

# Number of sites per LU
table(diptera_lepidoptera$Predominant_land_use)
```

# diptera_lepidoptera linear mixed effects modelling
```{r}
# maximal model
dl_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_lepidoptera)

# remove block
dl_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_lepidoptera)

# remove study
dl_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_lepidoptera)

# Compare models 
AIC(dl_m1, dl_m2, dl_m3) # dl_m3 best model

rm(dl_m1, dl_m2)

model_plot(dl_m2)
check_model(dl_m2)

# model estimates
summary(dl_m3)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dl_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_lepidoptera)

## Just land use as predictor
dl_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = diptera_lepidoptera)

## Just Higher_taxon as the predictor
dl_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_lepidoptera)

## Make a null, intercept only model
dl_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS), data = diptera_lepidoptera)

## Compare the models
AIC(dl_m3, dl_m4, dl_m5, dl_m6, dl_m7)
model.sel(dl_m3, dl_m4, dl_m5, dl_m6, dl_m7)

# model estimates
summary(dl_m3)
```

### Simulate diptera_lepidoptera effects
```{r}
# Predict Diptera effects 
dip_eff <- predict_this_model(dl_m3, order = "Diptera")

# relevel model with Lepidoptera as reference level
diptera_lepidoptera <- within(diptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_lepidoptera)

check_model(lep_m1)
summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
diptera_lepidoptera_plot_AB_log <- combine_and_plot(dip_eff, lep_eff, diversity = "Log_Abundance")

# Extract f ratio, df and p value
anova(dl_m3)

# Check residuals
hist(residuals(dl_m3))

# Check r2
r.squaredGLMM(dl_m3)

# Extract correlation coefficients
corr_dip_ref <- as.matrix(summary(dl_m3)$coefficients)
corr_lep_ref <- as.matrix(summary(lep_m1)$coefficients)

# Split the coefficients into two groups
# diptera <- corr[2:6, 1] 
# lepidoptera <- corr[8:nrow(corr), 1] 

diptera <- corr_dip_ref[2:6, 1] 
lepidoptera <- corr_lep_ref[2:6, 1] 

plot(diptera, lepidoptera)
cor(diptera, lepidoptera) # - 0.97

# regression through the origin
m <- lm(diptera ~ lepidoptera - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # coef = - 1.13

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.79
r <- sqrt(r2) # 0.89

# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # r = - 0.89

rm(col_eff, corr, dl_m3, dl_m4, dl_m5, dl_m6, dl_m7, lep_eff, dip_eff, lep_m1, lepidoptera, diptera, diptera_lepidoptera, corr_dip_ref, corr_lep_ref, m, r2, r, a)
```


## Hymenoptera

# hymenoptera_coleoptera abundance prepare df
```{r}
hym_col_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
hymenoptera_coleoptera <- model_data_ab %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Hymenoptera", "Coleoptera")) 

# LU split
table(hymenoptera_coleoptera$Predominant_land_use)
```

# hymenoptera_coleoptera linear mixed effects modelling
```{r}
# maximal model
hc_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_coleoptera)

# remove block
hc_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

# remove study
hc_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_coleoptera)

# Compare models 
AIC(hc_m1, hc_m2, hc_m3) # hc_m1 marginally better

rm(hc_m2, hc_m3)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hc_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_coleoptera)

## Just land use as predictor
hc_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = hymenoptera_coleoptera)

## Just Higher_taxon as the predictor
hc_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_coleoptera)

## Make a null, intercept only model
hc_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS), data = hymenoptera_coleoptera)

## Compare the models
AIC(hc_m1, hc_m4, hc_m5, hc_m6, hc_m7)
model.sel(hc_m1, hc_m4, hc_m5, hc_m6, hc_m7)

# hc_m1 still optimal
model_plot(hc_m1)

# model estimates
summary(hc_m1)
```

### Simulate hymenoptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(hc_m1, order = "Coleoptera")

# relevel model with Hymenoptera as reference level
hymenoptera_coleoptera <- within(hymenoptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model with Hymenoptera as ref
hym_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hym_m1, order = "Hymenoptera")

# Combine effects and plot
hymenoptera_coleoptera_plot_AB_log <- combine_and_plot(col_eff, hym_eff, diversity = "Log_Abundance")

# Extract f ratio, df and p value
anova(hc_m1)

# Check residuals
hist(residuals(hc_m1))

# Check r2
r.squaredGLMM(hc_m1)

# Extract correlation coefficients
corr_col_ref <- as.matrix(summary(hc_m1)$coefficients)
corr_hym_ref <- as.matrix(summary(hym_m1)$coefficients)

# Split the coefficients into two groups
# coleoptera <- corr[2:6, 1] 
# hymenoptera <- corr[8:nrow(corr), 1] 

coleoptera <- corr_col_ref[2:6, 1] 
hymenoptera <- corr_hym_ref[2:6, 1] 


plot(coleoptera, hymenoptera)
cor(coleoptera, hymenoptera)

# regression through the origin
m <- lm(coleoptera ~ hymenoptera - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # coef = 0.1927

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.1890454
r <- sqrt(r2) # 0.4347936

# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # r = 0.4347936

rm(col_eff, corr, hc_m1, hc_m4, hc_m5, hc_m6, hc_m7, hym_eff, hym_m1, coleoptera, hymenoptera, hymenoptera_coleoptera, corr_col_ref, corr_hym_ref, m, r2, r, a)
```



# hymenoptera_hemiptera abundance prepare df
```{r}
# Studies
hym_hemi_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hymenoptera", "Hemiptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
hymenoptera_hemiptera <- model_data_ab %>% filter(SS %in% hym_hemi_ss) %>% filter(Order %in% c("Hymenoptera", "Hemiptera")) 

# LU split
table(hymenoptera_hemiptera$Predominant_land_use)
```
# hymenoptera_hemiptera linear mixed effects modelling
```{r}
# maximal model
hh_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_hemiptera)

# remove block
hh_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_hemiptera)

# remove study
hh_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_hemiptera)

# Compare models 
AIC(hh_m1, hh_m2, hh_m3) # hh_m1 best model

model_plot(hh_m1)

rm(hh_m2, hh_m3)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hh_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = hymenoptera_hemiptera)

## Just land use as predictor
hh_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS) + (1 | SSB), data = hymenoptera_hemiptera)

## Just Higher_taxon as the predictor
hh_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = hymenoptera_hemiptera)

## Make a null, intercept only model
hh_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS) + (1 | SSB), data = hymenoptera_hemiptera)

## Compare the models
AIC(hh_m1, hh_m4, hh_m5, hh_m6, hh_m7)
model.sel(hh_m1, hh_m4, hh_m5, hh_m6, hh_m7)

# hh_m1 still optimal
# model estimates
summary(hh_m1)
```
# Simulate hymenoptera_hemiptera effects and plotting, but because there is no plantation data, I will do this manually.
```{r}
# Predict Hemiptera effects

# gather the effects and confidence intervals using simulation. Simulate 1000 times
  hemi_eff <- FEsim(hh_m1, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  hemi_eff$term <- as.factor(hemi_eff$term)
  
  # pull out effects
  hemi_eff <- hemi_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  hemi_eff$term <- recode_factor(hemi_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hemi_eff <- hemi_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(hemi_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  hemi_eff <- hemi_eff %>% mutate(Order = "Hemiptera")
  
  # make the order column a factor so it can be recoded
  hemi_eff$Order <- as.factor(hemi_eff$Order)

    
# relevel model with Hymenoptera as reference level
hymenoptera_hemiptera <- within(hymenoptera_hemiptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model with Hymenoptera as ref
hym_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_hemiptera)

summary(hym_m1)
Anova(hym_m1)

# Predict Hymenoptera effects 

# gather the effects and confidence intervals using simulation. Simulate 1000 times using the re leveled model
  hym_eff <- FEsim(hym_m1, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  hym_eff$term <- as.factor(hym_eff$term)
  
  # pull out effects
  hym_eff <- hym_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  hym_eff$term <- recode_factor(hym_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hym_eff <- hym_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(hym_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  hym_eff <- hym_eff %>% mutate(Order = "Hymenoptera")
  
  # make the order column a factor so it can be recoded
  hym_eff$Order <- as.factor(hym_eff$Order)

# Combine effects
combined <- rbind(hemi_eff, hym_eff)


comb_plot_limits <- c("Primary", "Secondary vegetation", "Pasture", "Cropland", "Urban") # LU categories for plotting

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845")
  
  
# Plot the effects for both orders
hymenoptera_hemiptera_plot_AB_log <- 
    combined %>%
  ggplot()+
        aes(x = Predominant_land_use, y = median, colour = Order, group = Order)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Upper_ci, ymax = Lower_ci), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        scale_x_discrete(limits=comb_plot_limits)+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        #geom_vline(xintercept = 5.5, linetype = 2)+
        #geom_vline(xintercept = 6.5, linetype = 2)+
        #geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("Log abundance") +
        scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 8)))
  

# Extract f ratio, df and p value
anova(hh_m1)

# Check residuals
hist(residuals(hh_m1))

# Check r2
r.squaredGLMM(hh_m1)

# Extract correlation coefficients
corr_hemi_ref <- as.matrix(summary(hh_m1)$coefficients)
corr_hym_ref <- as.matrix(summary(hym_m1)$coefficients)

# Split the coefficients into two groups
# hemiptera <- corr[2:5, 1] 
# hymenoptera <- corr[7:nrow(corr), 1] 

hemiptera <- corr_hemi_ref[2:5, 1] 
hymenoptera <- corr_hym_ref[2:5, 1] 

plot(hymenoptera, hemiptera)
cor(hymenoptera, hemiptera) # - 0.9634868

# regression through the origin
m <- lm(hemiptera ~ hymenoptera - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # coef = - 0.8932

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.9125051
r <- sqrt(r2) # 0.9552513

# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # r = - 0.9552513

rm(combined, corr, hemi_eff, hh_m1, hh_m4, hh_m5, hh_m6, hh_m7, hym_eff, hym_m1, hymenoptera_hemiptera, hemiptera, hymenoptera, corr_hemi_ref, corr_hym_ref, m, r2, r, a)
```



# hymenoptera_lepidoptera abundance prepare df
```{r}
# Studies
hym_lep_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hymenoptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
hymenoptera_lepidoptera <- model_data_ab %>% filter(SS %in% hym_lep_ss) %>% filter(Order %in% c("Hymenoptera", "Lepidoptera")) 

# LU split
table(hymenoptera_lepidoptera$Predominant_land_use)
```
# hymenoptera_lepidoptera linear mixed effects modelling
```{r}
# maximal model
hl_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_lepidoptera)

# remove block 
hl_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_lepidoptera)

# remove study
hl_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_lepidoptera)

# Compare models 
AIC(hl_m1, hl_m2, hl_m3) # hl_m2 best model

model_plot(hl_m2)

rm(hl_m1, hl_m3)


## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hl_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_lepidoptera)

## Just land use as predictor
hl_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = hymenoptera_lepidoptera)

## Just Higher_taxon as the predictor
hl_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_lepidoptera)

## Make a null, intercept only model
hl_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS), data = hymenoptera_lepidoptera)

## Compare the models
AIC(hl_m2, hl_m4, hl_m5, hl_m6, hl_m7)
model.sel(hl_m2, hl_m4, hl_m5, hl_m6, hl_m7)

# hl_m2 still optimal
# model estimates
summary(hl_m2)
```
### Simulate hymenoptera_lepidoptera effects
```{r}
# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hl_m2, order = "Hymenoptera")

# relevel model with Lepidoptera as reference level
hymenoptera_lepidoptera <- within(hymenoptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_lepidoptera)

model_plot(lep_m1)
summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
hymenoptera_lepidoptera_plot_AB_log <- combine_and_plot(hym_eff, lep_eff, diversity = "Log_Abundance")

# Extract f ratio, df and p value
anova(hl_m2)

# Check residuals
hist(residuals(hl_m2))

# Check r2
r.squaredGLMM(hl_m2)

# Extract correlation coefficients
corr_hym_ref <- as.matrix(summary(hl_m2)$coefficients)
corr_lep_ref <- as.matrix(summary(lep_m1)$coefficients)

# Split the coefficients into two groups
# hymenoptera <- corr[2:6, 1] 
# lepidoptera <- corr[8:nrow(corr), 1] 

hymenoptera <- corr_hym_ref[2:6, 1] 
lepidoptera <- corr_lep_ref[2:6, 1] 

plot(hymenoptera, lepidoptera)
cor(hymenoptera, lepidoptera)

# regression through the origin
m <- lm(hymenoptera ~ lepidoptera - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # coef = - 0.3293

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.4689337
r <- sqrt(r2) # 0.6847874

# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # r = - 0.6847874

rm(hym_eff, corr, hl_m2, hl_m4, hl_m5, hl_m6, hl_m7, lep_eff, lep_m1, hymenoptera, lepidoptera, hymenoptera_lepidoptera, corr_hym_ref, corr_lep_ref, m, r2, r, a)
```



## Hemiptera


# hemiptera_lepidoptera abundance prepare df
```{r}
# Studies
hemi_lep_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
hemiptera_lepidoptera <- model_data_ab %>% filter(SS %in% hemi_lep_ss) %>% filter(Order %in% c("Hemiptera", "Lepidoptera"))

# LU split
table(hemiptera_lepidoptera$Predominant_land_use)
```
# hemiptera_lepidoptera linear mixed effects modelling
```{r}
# maximal model
hl_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hemiptera_lepidoptera)

# remove block
hl_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_lepidoptera)

# remove study
hl_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_lepidoptera)

# Compare models 
AIC(hl_m1, hl_m2, hl_m3) # hl_m3 best model

model_plot(hl_m3)
rm(hl_m1, hl_m2)

# model estimates
summary(hl_m3)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hl_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = hemiptera_lepidoptera)

## Just land use as predictor
hl_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS) + (1 | SSB), data = hemiptera_lepidoptera)

## Just Higher_taxon as the predictor
hl_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = hemiptera_lepidoptera)

## Make a null, intercept only model
hl_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS) + (1 | SSB), data = hemiptera_lepidoptera)

## Compare the models
AIC(hl_m3, hl_m4, hl_m5, hl_m6, hl_m7)
model.sel(hl_m3, hl_m4, hl_m5, hl_m6, hl_m7)

# hl_m3 still optimal
```
### Simulate hemiptera_lepidoptera effects, but because there is no plantation data, I will do this manually.
```{r}
# Predict Hemiptera effects

# gather the effects and confidence intervals using simulation. Simulate 1000 times
  hemi_eff <- FEsim(hl_m3, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  hemi_eff$term <- as.factor(hemi_eff$term)
  
  # pull out effects
  hemi_eff <- hemi_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  hemi_eff$term <- recode_factor(hemi_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hemi_eff <- hemi_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(hemi_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  hemi_eff <- hemi_eff %>% mutate(Order = "Hemiptera")
  
  # make the order column a factor so it can be recoded
  hemi_eff$Order <- as.factor(hemi_eff$Order)

    
# relevel model with Lepidoptera as reference level
hemiptera_lepidoptera <- within(hemiptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_lepidoptera)

summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 

# gather the effects and confidence intervals using simulation. Simulate 1000 times using the re leveled model
  lep_eff <- FEsim(lep_m1, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  lep_eff$term <- as.factor(lep_eff$term)
  
  # pull out effects
  lep_eff <- lep_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  lep_eff$term <- recode_factor(lep_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  lep_eff <- lep_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(lep_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  lep_eff <- lep_eff %>% mutate(Order = "Lepidoptera")
  
  # make the order column a factor so it can be recoded
  lep_eff$Order <- as.factor(lep_eff$Order)

# Combine effects
combined <- rbind(hemi_eff, lep_eff)


comb_plot_limits <- c("Primary", "Secondary vegetation", "Pasture", "Cropland", "Urban") # LU categories for plotting

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845")
  
  
# Plot the effects for both orders
hemiptera_lepidoptera_plot_AB_log <- 
    combined %>%
  ggplot()+
        aes(x = Predominant_land_use, y = median, colour = Order, group = Order)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Upper_ci, ymax = Lower_ci), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        scale_x_discrete(limits=comb_plot_limits)+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        #geom_vline(xintercept = 5.5, linetype = 2)+
        #geom_vline(xintercept = 6.5, linetype = 2)+
        #geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("Log abundance") +
        scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 8)))
  

# Extract f ratio, df and p value
anova(hl_m3)

# Check residuals
hist(residuals(hl_m3))

# Check r2
r.squaredGLMM(hl_m3)

# Extract correlation coefficients
corr_hemi_ref <- as.matrix(summary(hl_m3)$coefficients)
corr_lep_ref <- as.matrix(summary(lep_m1)$coefficients)

# Split the coefficients into two groups
# hemiptera <- corr[2:5, 1] 
# lepidoptera <- corr[7:nrow(corr), 1] 

hemiptera <- corr_hemi_ref[2:5, 1] 
lepidoptera <- corr_lep_ref[2:5, 1] 

plot(lepidoptera, hemiptera)
cor(lepidoptera, hemiptera) # - 0.77

# regression through the origin
m <- lm(hemiptera ~ lepidoptera - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # coef = - 0.8408

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.4137117
r <- sqrt(r2) # 0.6432043

# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # r = - 0.6432043

rm(combined, corr, hemi_eff, hemiptera_lepidoptera, hl_m3, lep_eff, lep_m1, hemiptera, lepidoptera, corr_hemi_ref, corr_lep_ref, m, r2, r, a)
```



# hemiptera_coleoptera abundance prepare df
```{r}
# Studies
hemi_col_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Coleoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
hemiptera_coleoptera <- model_data_ab %>% filter(SS %in% hemi_col_ss) %>% filter(Order %in% c("Hemiptera", "Coleoptera")) 

# LU split
table(hemiptera_coleoptera$Predominant_land_use)

```
# hemiptera_coleoptera linear mixed effects modelling
```{r}
# maximal model
hc_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hemiptera_coleoptera)

# remove block
hc_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_coleoptera)

# remove study
hc_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_coleoptera)

# Compare models 
AIC(hc_m1, hc_m2, hc_m3) # hc_m1 best model

model_plot(hc_m1)

rm(hc_m2, hc_m3)

# model estimates
summary(hc_m1)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hc_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = hemiptera_coleoptera)

## Just land use as predictor
hc_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS) + (1 | SSB), data = hemiptera_coleoptera)

## Just Higher_taxon as the predictor
hc_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = hemiptera_coleoptera)

## Make a null, intercept only model
hc_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS) + (1 | SSB), data = hemiptera_coleoptera)

## Compare the models
AIC(hc_m1, hc_m4, hc_m5, hc_m6, hc_m7)
model.sel(hc_m1, hc_m4, hc_m5, hc_m6, hc_m7)

# hc_m1 still optimal
```
### Simulate hemiptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(hc_m1, order = "Coleoptera")

# relevel model with Hemiptera as reference level
hemiptera_coleoptera <- within(hemiptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model with Hemiptera as ref
hem_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
    data = hemiptera_coleoptera)

summary(hem_m1)
Anova(hem_m1)

# Predict Hemiptera effects 
hem_eff <- predict_this_model(hem_m1, order = "Hemiptera")

# Combine effects and plot
hemiptera_coleoptera_plot_AB_log <- combine_and_plot(col_eff, hem_eff, diversity = "Log_Abundance")

# Extract f ratio, df and p value
anova(hc_m1)

# Check residuals
hist(residuals(hc_m1))

# Check r2
r.squaredGLMM(hc_m1)

# Extract correlation coefficients
corr_col_ref <- as.matrix(summary(hc_m1)$coefficients)
corr_hemi_ref <- as.matrix(summary(hem_m1)$coefficients)

# Split the coefficients into two groups
# coleoptera <- corr[2:6, 1] 
# hemiptera <- corr[8:nrow(corr), 1] 

coleoptera <- corr_col_ref[2:6, 1] 
hemiptera <- corr_hemi_ref[2:6, 1] 

plot(coleoptera, hemiptera)
cor(coleoptera, hemiptera) # - 0.87

# regression through the origin
m <- lm(coleoptera ~ hemiptera - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # coef = - 1.1625

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.7717527
r <- sqrt(r2) # 0.8784946

# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # r = - 0.8784946

rm(col_eff, corr, hc_m1, hc_m4, hc_m5, hc_m6, hc_m7, hem_eff, hem_m1, coleoptera, hemiptera, hemiptera_coleoptera, corr_col_ref, corr_hemi_ref, m, r2, r, a)
```



### coleoptera_lepidoptera abundance prepare df
```{r}
# Studies
col_lep_ss <-  model_data_ab %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
coleoptera_lepidoptera <- model_data_ab %>% filter(SS %in% col_lep_ss) %>% filter(Order %in% c("Coleoptera", "Lepidoptera"))

# LU split
table(coleoptera_lepidoptera$Predominant_land_use)
```
# coleoptera_lepidoptera linear mixed effects modelling
```{r}
# maximal model
cl_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = coleoptera_lepidoptera)

# remove block
cl_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = coleoptera_lepidoptera)

# remove study
cl_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = coleoptera_lepidoptera)

# Compare models 
AIC(cl_m1, cl_m2, cl_m3) # cl_m2 best model

model_plot(cl_m2)

rm(cl_m1, cl_m3)

# model estimates
summary(cl_m2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
cl_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = coleoptera_lepidoptera)

## Just land use as predictor
cl_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS) + (1 | SSB), data = coleoptera_lepidoptera)

## Just Higher_taxon as the predictor
cl_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS) + (1 | SSB), data = coleoptera_lepidoptera)

## Make a null, intercept only model
cl_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS) + (1 | SSB), data = coleoptera_lepidoptera)

## Compare the models
AIC(cl_m2, cl_m4, cl_m5, cl_m6, cl_m7)
model.sel(cl_m2, cl_m4, cl_m5, cl_m6, cl_m7)

# cl_m2 still optimal
```
### Simulate coleoptera_lepidoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(cl_m2, order = "Coleoptera")

# relevel model with Lepidoptera as reference level
coleoptera_lepidoptera <- within(coleoptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
    data = coleoptera_lepidoptera)

model_plot(lep_m1)
summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
coleoptera_lepidoptera_plot_AB_log <- combine_and_plot(col_eff, lep_eff, diversity = "Log_Abundance")

# Extract f ratio, df and p value
anova(cl_m2)

# Check residuals
hist(residuals(cl_m2))

# Check r2
r.squaredGLMM(cl_m2)

# Extract correlation coefficients
corr_col_ref <- as.matrix(summary(cl_m2)$coefficients)
corr_lep_ref <- as.matrix(summary(lep_m1)$coefficients)

# Split the coefficients into two groups
#coleoptera <- corr[2:6, 1] 
#lepidoptera <- corr[8:nrow(corr), 1] 

coleoptera <- corr_col_ref[2:6, 1] 
lepidoptera <- corr_lep_ref[2:6, 1] 

plot(coleoptera, lepidoptera)
cor(coleoptera, lepidoptera) # - 0.81

# regression through the origin
m <- lm(coleoptera ~ lepidoptera - 1)

# This model has only one coefficient, with no intercept, it only contains the slope
summary(m) # - 0.33

# get r2 from the summary; square-root it and give the result the correct sign to get the correlation coefficient through the origin
r2 <- summary(m)$r.squared # 0.1664792
r <- sqrt(r2) # 0.4080186

# and change the sign if you need to 
a <- ifelse(coef(m) > 0, r, -r) # - 0.4080186

rm(col_eff, corr, cl_m2, cl_m4, cl_m5, cl_m6, cl_m7, lep_eff, lep_m1, coleoptera, lepidoptera, coleoptera_lepidoptera, corr_col_ref, corr_lep_ref, m, r2, r, a)
```



### Order comparisons for Species Richness

### Prepare the SR model data
```{r}
# Log transform species richness
model_data_sr$logSR <- log(model_data_sr$SpeciesRichness + 1)

# Check the distributions
hist(model_data_sr$SpeciesRichness) ; hist(model_data_sr$logSR) ; hist(model_data_ab$logAbundance) ; hist(log(model_data_ab$TotalAbundance + 1))

# Have a look at sample sizes
table(model_data_sr$Predominant_land_use, model_data_sr$Use_intensity)

# Collapse intense and light land use intensities to have a more even spread of the data
model_data_sr <- 
  model_data_sr %>%
  mutate(
    Use_intensity = recode_factor(Use_intensity,
                                  "Intense use" = "High_use",
                                  "Light use" = "High_use"),
# reset reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )
# Have a look at sample sizes
table(model_data_sr$Predominant_land_use, model_data_sr$Use_intensity)

# Order / land use split
table(model_data_sr$Order,  model_data_sr$Predominant_land_use)

# Relevel the land use and use intensity classes
model_data_sr <- model_data_sr %>%
  
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Young secondary vegetation" = "Secondary vegetation", 
                                         "Intermediate secondary vegetation" = "Secondary vegetation",
                                         "Mature secondary vegetation" = "Secondary vegetation",
                                         "Secondary vegetation (indeterminate age)" = "Secondary vegetation"),
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(model_data_sr$Predominant_land_use, model_data_sr$Use_intensity) # Only pasture minimal use (n = 85) is less than 100 sites.

# Lets have a look at the order x land use split again
table(model_data_sr$Order, model_data_sr$Predominant_land_use) 

# Best way to proceed: Mick Crawley reckons an average of 30 data points to fit a parameter is comfy, 10 might work, and 3 if you're as lucky as a human can be. On that basis, I think you'll be fine - but obviously some combinations of order 1, order 2 and land use will be based on very few data indeed (so will have suitably wide CIs).
```


## Diptera


# diptera_hymenoptera species richness prepare df
```{r}
# Studies
dip_hym_ss_sr <-  model_data_sr %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Its the same sets of studies for abundance and richness so I don't need to recalculate the studies for species richness data
cbind(dip_hym_ss_sr, dip_hym_ss) %>% View()

# Model data
diptera_hymenoptera <- model_data_sr %>% filter(SS %in% dip_hym_ss_sr) %>% filter(Order %in% c("Diptera", "Hymenoptera"))

# LU split
table(diptera_hymenoptera$Predominant_land_use) # LU split is same as for abundance data.

```
# diptera_hymenoptera linear mixed effects modelling
```{r}
# maximal model
dhy_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hymenoptera)

# remove block
dhy_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# remove study
dhy_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_hymenoptera)

# Compare models 
AIC(dhy_m1, dhy_m2, dhy_m3) # dhy_m2 best model

model_plot(dhy_m2)

rm(dhy_m1, dhy_m3)

summary(dhy_m2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dhy_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just land use as predictor
dhy_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just Higher_taxon as the predictor
dhy_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Make a null, intercept only model
dhy_m7 <- lmer(logSR ~ (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Compare the models
model.sel(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7)

# dhy_m2 still the best
```
### Simulate diptera_hymenoptera effects
```{r}
# Predict Diptera effects
dip_eff <- predict_this_model(dhy_m2, order = "Diptera", baseline = "median")

# relevel model with hymenoptera as reference level
diptera_hymenoptera <- within(diptera_hymenoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model
hym_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# Predict hymenoptera effects
hym_eff <- predict_this_model(hym_m1, order = "Hymenoptera", baseline = "median")

# Combine effects and plot them
diptera_hymenoptera_plot_SR_log <- combine_and_plot(dip_eff, hym_eff, diversity = "Log_SR")

# Extract f ratio, df and p value
anova(dhy_m2)

# Check residuals
hist(residuals(dhy_m2))

# Check r2
r.squaredGLMM(dhy_m2)

# Extract correlation coefficients
corr <- as.matrix(summary(dhy_m2)$coefficients)

# Split the coefficients into two groups
diptera <- corr[2:6, 1] 
hymenoptera <- corr[8:nrow(corr), 1] 

plot(diptera, hymenoptera)
cor(diptera, hymenoptera)

rm(dip_eff, corr, dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7, hym_eff, hym_m1, hymenoptera, diptera, diptera_hymenoptera)

diptera_hymenoptera_plot_AB_log ; diptera_hymenoptera_plot_SR_log
```



# diptera_coleoptera species richness
```{r}
# Model data
diptera_coleoptera <- model_data_sr %>% filter(SS %in% dip_col_ss) %>% filter(Order %in% c("Diptera", "Coleoptera")) 

# LU split
table(diptera_coleoptera$Predominant_land_use) # just checking the LU split is the same as in the abundance data - which it is.

# maximal model
dc_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_coleoptera)

# remove block
dc_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_coleoptera)

# remove study
dc_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_coleoptera)

# Compare models 
AIC(dc_m1, dc_m2, dc_m3) # dc_m2 best model

model_plot(dc_m2)
rm(dc_m1, dc_m3)

# model estimates
summary(dc_m2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dc_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_coleoptera)

## Just land use as predictor
dc_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = diptera_coleoptera)

## Just Higher_taxon as the predictor
dc_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_coleoptera)

## Make a null, intercept only model
dc_m7 <- lmer(logSR ~ (1 | Source_ID) + (1|SS), data = diptera_coleoptera)

## Compare the models
AIC(dc_m2, dc_m4, dc_m5, dc_m6, dc_m7)
model.sel(dc_m2, dc_m4, dc_m5, dc_m6, dc_m7)

# dc_m2 still optimal
```
### Simulate diptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(dc_m2, order = "Coleoptera")

# relevel model with Diptera as reference level
diptera_coleoptera <- within(diptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Diptera"))

# maximal model
dip_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_coleoptera)

summary(dip_m1)
Anova(dip_m1)

# Predict Diptera effects - remember that this overwrites previous effects of Diptera model
dip_eff <- predict_this_model(dip_m1, order = "Diptera")

# Combine effects and relevel Diptera as reference
diptera_coleoptera_plot_SR_log <- combine_and_plot(dip_eff, col_eff, relevel = "Diptera", diversity = "Log_SR")

# Extract f ratio, df and p value
anova(dc_m2)

# Check residuals
hist(residuals(dc_m2))

# Check r2
r.squaredGLMM(dc_m2)

# Extract correlation coefficients
corr <- as.matrix(summary(dc_m2)$coefficients)

# Split the coefficients into two groups
col <- corr[2:6, 1] 
dip <- corr[8:nrow(corr), 1] 

plot(col, dip)
cor(col, dip)

# Compare biodiversity metrics
diptera_coleoptera_plot_AB_log ; diptera_coleoptera_plot_SR_log

rm(col_eff, corr, dc_m2, dc_m4, dc_m5, dc_m6, dc_m7, dip_eff, dip_m1, col, dip, diptera_coleoptera)
```


# diptera_hemiptera species richness
```{r}
# Model data
diptera_hemiptera <- model_data_sr %>% filter(SS %in% dip_hemi_ss) %>% filter(Order %in% c("Diptera", "Hemiptera"))

# maximal model
dh_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hemiptera)

# remove block
dh_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hemiptera)

# remove study
dh_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_hemiptera)

# Compare models 
AIC(dh_m1, dh_m2, dh_m3) # dh_m3 best model

model_plot(dh_m3)

rm(dh_m1, dh_m2)

# model estimates
summary(dh_m3)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dh_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID), data = diptera_hemiptera)

## Just land use as predictor
dh_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID), data = diptera_hemiptera)

## Just Higher_taxon as the predictor
dh_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID), data = diptera_hemiptera)

## Make a null, intercept only model
dh_m7 <- lmer(logSR ~ (1 | Source_ID), data = diptera_hemiptera)

## Compare the models
AIC(dh_m3, dh_m4, dh_m5, dh_m6, dh_m7)
model.sel(dh_m3, dh_m4, dh_m5, dh_m6, dh_m7)

# dh_m3 still optimal
```
# Simulate diptera effects and plotting, but because there is no plantation data, I will do this manually.
```{r}
summary(dh_m3) # no coefficient for plantation as there is no data

# Predict diptera effects

# gather the effects and confidence intervals using simulation. Simulate 1000 times
  dip_eff <- FEsim(dh_m3, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  dip_eff$term <- as.factor(dip_eff$term)
  
  # pull out effects
  dip_eff <- dip_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  dip_eff$term <- recode_factor(dip_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  dip_eff <- dip_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(dip_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  dip_eff <- dip_eff %>% mutate(Order = "Diptera")
  
  # make the order column a factor so it can be recoded
  dip_eff$Order <- as.factor(dip_eff$Order)

    
# relevel model with Hemiptera as reference level
diptera_hemiptera <- within(diptera_hemiptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model with Hemiptera as ref
hemi_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
              data = diptera_hemiptera)

summary(hemi_m1)
Anova(hemi_m1)

# Predict Hemiptera effects 

# gather the effects and confidence intervals using simulation. Simulate 1000 times using the re leveled model
  hemi_eff <- FEsim(hemi_m1, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  hemi_eff$term <- as.factor(hemi_eff$term)
  
  # pull out effects
  hemi_eff <- hemi_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  hemi_eff$term <- recode_factor(hemi_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hemi_eff <- hemi_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(hemi_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  hemi_eff <- hemi_eff %>% mutate(Order = "Hemiptera")
  
  # make the order column a factor so it can be recoded
  hemi_eff$Order <- as.factor(hemi_eff$Order)

# Combine effects
combined <- rbind(dip_eff, hemi_eff)


comb_plot_limits <- c("Primary", "Secondary vegetation", "Pasture", "Cropland", "Urban") # LU categories for plotting

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845")
  
  
# Plot the effects for both orders
diptera_hemiptera_plot_SR_log <- 
    combined %>%
  ggplot()+
        aes(x = Predominant_land_use, y = median, colour = Order, group = Order)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Upper_ci, ymax = Lower_ci), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        scale_x_discrete(limits=comb_plot_limits)+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        #geom_vline(xintercept = 5.5, linetype = 2)+
        #geom_vline(xintercept = 6.5, linetype = 2)+
        #geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("Log Species Richness") +
        scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 8)))
  

# Extract f ratio, df and p value
anova(dh_m3)

# Check residuals
hist(residuals(dh_m3))

# Check r2
r.squaredGLMM(dh_m3)

# Extract correlation coefficients
corr <- as.matrix(summary(dh_m3)$coefficients)

# Split the coefficients into two groups
diptera <- corr[2:5, 1] 
hemiptera <- corr[7:nrow(corr), 1] 

plot(diptera, hemiptera)
cor(diptera, hemiptera)

diptera_hemiptera_plot_AB_log ; diptera_hemiptera_plot_SR_log

rm(combined, corr, dh_m3, dh_m4, dh_m5, dh_m6, dh_m7, dip_eff, diptera_hemiptera, hemi_eff, hemi_m1, diptera, hemiptera)
```



# diptera_lepidoptera species richness
```{r}
# Model data
diptera_lepidoptera <- model_data_sr %>% filter(SS %in% dip_lep_ss) %>% filter(Order %in% c("Diptera", "Lepidoptera")) 

# maximal model
dl_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_lepidoptera)

# remove block
dl_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_lepidoptera)

# remove study
dl_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_lepidoptera)

# Compare models 
AIC(dl_m1, dl_m2, dl_m3) # dl_m3 best model

model_plot(dl_m3)
rm(dl_m2, dl_m1)

# model estimates
summary(dl_m3)
Anova(dl_m3)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dl_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID), data = diptera_lepidoptera)

## Just land use as predictor
dl_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID), data = diptera_lepidoptera)

## Just Higher_taxon as the predictor
dl_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID), data = diptera_lepidoptera)

## Make a null, intercept only model
dl_m7 <- lmer(logSR ~ (1 | Source_ID), data = diptera_lepidoptera)

## Compare the models
AIC(dl_m3, dl_m4, dl_m5, dl_m6, dl_m7)
model.sel(dl_m3, dl_m4, dl_m5, dl_m6, dl_m7)

# dl_m6 has lower AIC than dl_m3 
```
### Simulate diptera_lepidoptera effects
```{r}
# Predict Diptera effects 
dip_eff <- predict_this_model(dl_m3, order = "Diptera")

# relevel model with Lepidoptera as reference level
diptera_lepidoptera <- within(diptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_lepidoptera)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
diptera_lepidoptera_plot_SR_log <- combine_and_plot(dip_eff, lep_eff, diversity = "Log_SR")

# Extract f ratio, df and p value
anova(dl_m3)

# Check residuals
hist(residuals(dl_m3))

# Check r2
r.squaredGLMM(dl_m3)

# Extract correlation coefficients
corr <- as.matrix(summary(dl_m3)$coefficients)

# Split the coefficients into two groups
dip <- corr[2:6, 1] 
lep <- corr[8:nrow(corr), 1] 

plot(dip, lep)
cor(dip, lep)

diptera_lepidoptera_plot_AB_log ; diptera_lepidoptera_plot_SR_log

rm(dip_eff, corr, dl_m3, dl_m4, dl_m5, dl_m6, dl_m7, lep_eff, lep_m1, dip, lep, diptera_lepidoptera)
```



## Hymenoptera

# hymenoptera_coleoptera species richness
```{r}
# Model data
hymenoptera_coleoptera <- model_data_sr %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Hymenoptera", "Coleoptera")) 

# maximal model
hc_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_coleoptera)

# remove block
hc_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

# remove study
hc_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_coleoptera)

# Compare models 
AIC(hc_m1, hc_m2, hc_m3) # hc_m2 best model

model_plot(hc_m2)
rm(hc_m1, hc_m3)

# model estimates
summary(hc_m2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hc_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_coleoptera)

## Just land use as predictor
hc_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = hymenoptera_coleoptera)

## Just Higher_taxon as the predictor
hc_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_coleoptera)

## Make a null, intercept only model
hc_m7 <- lmer(logSR ~ (1 | Source_ID) + (1|SS), data = hymenoptera_coleoptera)

## Compare the models
AIC(hc_m2, hc_m4, hc_m5, hc_m6, hc_m7)
model.sel(hc_m2, hc_m4, hc_m5, hc_m6, hc_m7)

# hc_m2 still optimal
```
### Simulate hymenoptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(hc_m2, order = "Coleoptera")

# relevel model with Hymenoptera as reference level
hymenoptera_coleoptera <- within(hymenoptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model with Hymenoptera as ref
hym_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hym_m1, order = "Hymenoptera")

# Combine effects and plot
hymenoptera_coleoptera_plot_SR_log <- combine_and_plot(col_eff, hym_eff, diversity = "Log_SR")

# Extract f ratio, df and p value
anova(hc_m2)

# Check residuals
hist(residuals(hc_m2))

# Check r2
r.squaredGLMM(hc_m2)

# Extract correlation coefficients
corr <- as.matrix(summary(hc_m2)$coefficients)

# Split the coefficients into two groups
col <- corr[2:6, 1] 
hym <- corr[8:nrow(corr), 1] 

plot(col, hym)
cor(col, hym)

hymenoptera_coleoptera_plot_AB_log ; hymenoptera_coleoptera_plot_SR_log

rm(col_eff, corr, hc_m2, hc_m4, hc_m5, hc_m6, hc_m7, hym_eff, hym_m1, col, hym, hymenoptera_coleoptera)
```



# hymenoptera_hemiptera species richness
```{r}
# Model data
hymenoptera_hemiptera <- model_data_sr %>% filter(SS %in% hym_hemi_ss) %>% filter(Order %in% c("Hymenoptera", "Hemiptera")) 

# maximal model
hh_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_hemiptera)

# remove block
hh_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_hemiptera)

# remove study
hh_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_hemiptera)

# Compare models 
AIC(hh_m1, hh_m2, hh_m3) # hh_m3 best model

model_plot(hh_m3)
rm(hh_m1, hh_m2)

# model estimates
summary(hh_m3)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hh_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_hemiptera)

## Just land use as predictor
hh_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = hymenoptera_hemiptera)

## Just Higher_taxon as the predictor
hh_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_hemiptera)

## Make a null, intercept only model
hh_m7 <- lmer(logSR ~ (1 | Source_ID) + (1|SS), data = hymenoptera_hemiptera)

## Compare the models
AIC(hh_m3, hh_m4, hh_m5, hh_m6, hh_m7)
model.sel(hh_m3, hh_m4, hh_m5, hh_m6, hh_m7)

# hh_m3 still optimal

```
# Simulate hymenoptera_hemiptera effects and plotting, but because there is no plantation data, I will do this manually.
```{r}
# Predict Hemiptera effects

# gather the effects and confidence intervals using simulation. Simulate 1000 times
  hemi_eff <- FEsim(hh_m3, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  hemi_eff$term <- as.factor(hemi_eff$term)
  
  # pull out effects
  hemi_eff <- hemi_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  hemi_eff$term <- recode_factor(hemi_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hemi_eff <- hemi_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(hemi_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  hemi_eff <- hemi_eff %>% mutate(Order = "Hemiptera")
  
  # make the order column a factor so it can be recoded
  hemi_eff$Order <- as.factor(hemi_eff$Order)

    
# relevel model with Hymenoptera as reference level
hymenoptera_hemiptera <- within(hymenoptera_hemiptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model with Hymenoptera as ref
hym_m1  <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_hemiptera)

summary(hym_m1)
Anova(hym_m1)

# Predict Hymenoptera effects 

# gather the effects and confidence intervals using simulation. Simulate 1000 times using the re leveled model
  hym_eff <- FEsim(hym_m1, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  hym_eff$term <- as.factor(hym_eff$term)
  
  # pull out effects
  hym_eff <- hym_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  hym_eff$term <- recode_factor(hym_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hym_eff <- hym_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(hym_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  hym_eff <- hym_eff %>% mutate(Order = "Hymenoptera")
  
  # make the order column a factor so it can be recoded
  hym_eff$Order <- as.factor(hym_eff$Order)

# Combine effects
combined <- rbind(hemi_eff, hym_eff)


comb_plot_limits <- c("Primary", "Secondary vegetation", "Pasture", "Cropland", "Urban") # LU categories for plotting

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845")
  
  
# Plot the effects for both orders
hymenoptera_hemiptera_plot_SR_log <- 
    combined %>%
  ggplot()+
        aes(x = Predominant_land_use, y = median, colour = Order, group = Order)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Upper_ci, ymax = Lower_ci), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        scale_x_discrete(limits=comb_plot_limits)+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        #geom_vline(xintercept = 5.5, linetype = 2)+
        #geom_vline(xintercept = 6.5, linetype = 2)+
        #geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("Log Species Richness") +
        scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 8)))
  

# Extract f ratio, df and p value
anova(hh_m3)

# Check residuals
hist(residuals(hh_m3))

# Check r2
r.squaredGLMM(hh_m3)

# Extract correlation coefficients
corr <- as.matrix(summary(hh_m3)$coefficients)

# Split the coefficients into two groups
hemiptera <- corr[2:5, 1] 
hymenoptera <- corr[7:nrow(corr), 1] 

plot(hymenoptera, hemiptera)
cor(hymenoptera, hemiptera)

hymenoptera_hemiptera_plot_AB_log ; hymenoptera_hemiptera_plot_SR_log

rm(combined, corr, hemi_eff, hh_m3, hh_m4, hh_m5, hh_m6, hh_m7, hym_eff, hym_m1, hymenoptera_hemiptera, hemiptera, hymenoptera)
```



# hymenoptera_lepidoptera species_richness
```{r}
# Model data
hymenoptera_lepidoptera <- model_data_sr %>% filter(SS %in% hym_lep_ss) %>% filter(Order %in% c("Hymenoptera", "Lepidoptera")) 

# maximal model
hl_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_lepidoptera)

# remove block
hl_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_lepidoptera)

# remove study
hl_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_lepidoptera)

# Compare models 
AIC(hl_m1, hl_m2, hl_m3) # hl_m2 best model

model_plot(hl_m2)
rm(hl_m1, hl_m3)

# model estimates
summary(hl_m2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hl_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_lepidoptera)

## Just land use as predictor
hl_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = hymenoptera_lepidoptera)

## Just Higher_taxon as the predictor
hl_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = hymenoptera_lepidoptera)

## Make a null, intercept only model
hl_m7 <- lmer(logSR ~ (1 | Source_ID) + (1|SS), data = hymenoptera_lepidoptera)

## Compare the models
AIC(hl_m2, hl_m4, hl_m5, hl_m6, hl_m7)
model.sel(hl_m2, hl_m4, hl_m5, hl_m6, hl_m7)

# hl_m2 still optimal
```
### Simulate hymenoptera_lepidoptera effects
```{r}
# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hl_m2, order = "Hymenoptera")

# relevel model with Lepidoptera as reference level
hymenoptera_lepidoptera <- within(hymenoptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_lepidoptera)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
hymenoptera_lepidoptera_plot_SR_log <- combine_and_plot(hym_eff, lep_eff, diversity = "Log_SR")

# Extract f ratio, df and p value
anova(hl_m2)

# Check residuals
hist(residuals(hl_m2))

# Check r2
r.squaredGLMM(hl_m2)

# Extract correlation coefficients
corr <- as.matrix(summary(hl_m2)$coefficients)

# Split the coefficients into two groups
hym <- corr[2:6, 1] 
lep <- corr[8:nrow(corr), 1] 

plot(hym, lep)
cor(hym, lep)

hymenoptera_lepidoptera_plot_AB_log ; hymenoptera_lepidoptera_plot_SR_log

rm(hym_eff, corr, hl_m2, hl_m4, hl_m5, hl_m6, hl_m7, lep_eff, lep_m1, hym, lep, hymenoptera_lepidoptera)
```




## Hemiptera

# hemiptera_lepidoptera species richness
```{r}
# Model data
hemiptera_lepidoptera <- model_data_sr %>% filter(SS %in% hemi_lep_ss) %>% filter(Order %in% c("Hemiptera", "Lepidoptera")) 

# maximal model
hl_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hemiptera_lepidoptera)

# remove block
hl_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_lepidoptera)

# remove study
hl_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_lepidoptera)

# Compare models 
AIC(hl_m1, hl_m2, hl_m3) # hl_m3 best model

model_plot(hl_m3)
rm(hl_m2, hl_m1)

# model estimates
summary(hl_m3)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hl_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID), data = hemiptera_lepidoptera)

## Just land use as predictor
hl_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID), data = hemiptera_lepidoptera)

## Just Higher_taxon as the predictor
hl_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID), data = hemiptera_lepidoptera)

## Make a null, intercept only model
hl_m7 <- lmer(logSR ~ (1 | Source_ID), data = hemiptera_lepidoptera)

## Compare the models
AIC(hl_m3, hl_m4, hl_m5, hl_m6, hl_m7)
model.sel(hl_m3, hl_m4, hl_m5, hl_m6, hl_m7)

# hl_m3 still optimal
```
### Simulate hemiptera_lepidoptera effects, but because there is no plantation data, I will do this manually.
```{r}
# Predict Hemiptera effects

# gather the effects and confidence intervals using simulation. Simulate 1000 times
  hemi_eff <- FEsim(hl_m3, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  hemi_eff$term <- as.factor(hemi_eff$term)
  
  # pull out effects
  hemi_eff <- hemi_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  hemi_eff$term <- recode_factor(hemi_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hemi_eff <- hemi_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(hemi_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  hemi_eff <- hemi_eff %>% mutate(Order = "Hemiptera")
  
  # make the order column a factor so it can be recoded
  hemi_eff$Order <- as.factor(hemi_eff$Order)

    
# relevel model with Lepidoptera as reference level
hemiptera_lepidoptera <- within(hemiptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_lepidoptera)

summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 

# gather the effects and confidence intervals using simulation. Simulate 1000 times using the re leveled model
  lep_eff <- FEsim(lep_m1, n.sims = 1000)
  
  # make the term column a factor so it can be recoded
  lep_eff$term <- as.factor(lep_eff$term)
  
  # pull out effects
  lep_eff <- lep_eff %>%
    slice(1:5) 
  
  # rename factors for ease of visualisation
  lep_eff$term <- recode_factor(lep_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                #"Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  lep_eff <- lep_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  # Rename the sim eff term column name for joining
  colnames(lep_eff)[1] <- "Predominant_land_use"
  
  # Add in a column of Order to make sure there is a grouping variable
  lep_eff <- lep_eff %>% mutate(Order = "Lepidoptera")
  
  # make the order column a factor so it can be recoded
  lep_eff$Order <- as.factor(lep_eff$Order)

# Combine effects
combined <- rbind(hemi_eff, lep_eff)


comb_plot_limits <- c("Primary", "Secondary vegetation", "Pasture", "Cropland", "Urban") # LU categories for plotting

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845")
  
  
# Plot the effects for both orders
hemiptera_lepidoptera_plot_SR_log <- 
    combined %>%
  ggplot()+
        aes(x = Predominant_land_use, y = median, colour = Order, group = Order)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Upper_ci, ymax = Lower_ci), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        scale_x_discrete(limits=comb_plot_limits)+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        #geom_vline(xintercept = 5.5, linetype = 2)+
        #geom_vline(xintercept = 6.5, linetype = 2)+
        #geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("Log Species Richness") +
        scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 8)))
  

# Extract f ratio, df and p value
anova(hl_m3)

# Check residuals
hist(residuals(hl_m3))

# Check r2
r.squaredGLMM(hl_m3)

# Extract correlation coefficients
corr <- as.matrix(summary(hl_m3)$coefficients)

# Split the coefficients into two groups
hemiptera <- corr[2:5, 1] 
lepidoptera <- corr[7:nrow(corr), 1] 

plot(lepidoptera, hemiptera)
cor(lepidoptera, hemiptera)

hemiptera_lepidoptera_plot_AB_log ; hemiptera_lepidoptera_plot_SR_log

rm(combined, corr, hemi_eff, hemiptera_lepidoptera, hl_m3, hl_m4, hl_m5, hl_m6, hl_m7, lep_eff, lep_m1, hemiptera, lepidoptera)
```



# hemiptera_coleoptera species_richness
```{r}
# Model data
hemiptera_coleoptera <- model_data_sr %>% filter(SS %in% hemi_col_ss) %>% filter(Order %in% c("Hemiptera", "Coleoptera")) 

# maximal model
hc_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hemiptera_coleoptera)

# remove block
hc_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_coleoptera)

# remove study
hc_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_coleoptera)

# Compare models 
AIC(hc_m1, hc_m2, hc_m3) # hc_m1 best model

model_plot(hc_m1)
rm(hc_m2, hc_m3)

# model estimates
summary(hc_m1)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
hc_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS) + (1|SSB), data = hemiptera_coleoptera)

## Just land use as predictor
hc_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID) + (1|SS) + (1|SSB), data = hemiptera_coleoptera)

## Just Higher_taxon as the predictor
hc_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID) + (1|SS) + (1|SSB), data = hemiptera_coleoptera)

## Make a null, intercept only model
hc_m7 <- lmer(logSR ~ (1 | Source_ID) + (1|SS) + (1|SSB), data = hemiptera_coleoptera)

## Compare the models
AIC(hc_m1, hc_m4, hc_m5, hc_m6, hc_m7)
model.sel(hc_m1, hc_m4, hc_m5, hc_m6, hc_m7)

# hc_m1 still optimal
```
### Simulate hemiptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(hc_m1, order = "Coleoptera")

# relevel model with Hemiptera as reference level
hemiptera_coleoptera <- within(hemiptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model with Hemiptera as ref
hem_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
    data = hemiptera_coleoptera)

model_plot(hem_m1)
summary(hem_m1)
Anova(hem_m1)

# Predict Hemiptera effects 
hem_eff <- predict_this_model(hem_m1, order = "Hemiptera")

# Combine effects and plot
hemiptera_coleoptera_plot_SR_log <- combine_and_plot(col_eff, hem_eff, diversity = "Log_SR")

# Extract f ratio, df and p value
anova(hc_m1)

# Check residuals
hist(residuals(hc_m1))

# Check r2
r.squaredGLMM(hc_m1)

# Extract correlation coefficients
corr <- as.matrix(summary(hc_m1)$coefficients)

# Split the coefficients into two groups
col <- corr[2:6, 1] 
hemi<- corr[8:nrow(corr), 1] 

plot(col, hemi)
cor(col, hemi)

hemiptera_coleoptera_plot_AB_log ; hemiptera_coleoptera_plot_SR_log

rm(col_eff, corr, hc_m1, hc_m4, hc_m5, hc_m6, hc_m7, hem_eff, hem_m1, col, hemi, hemiptera_coleoptera)
```



# coleoptera_lepidoptera species richness
```{r}
# Model data
coleoptera_lepidoptera <- model_data_sr %>% filter(SS %in% col_lep_ss) %>% filter(Order %in% c("Coleoptera", "Lepidoptera")) 

# maximal model
cl_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = coleoptera_lepidoptera)

# remove block
cl_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = coleoptera_lepidoptera)

# remove study
cl_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = coleoptera_lepidoptera)

# Compare models 
AIC(cl_m1, cl_m2, cl_m3) # cl_m1 best model

model_plot(cl_m1)
rm(cl_m2, cl_m3)

# model estimates
summary(cl_m1)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
cl_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1 | SS), data = coleoptera_lepidoptera)

## Just land use as predictor
cl_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID) + (1 | SS), data = coleoptera_lepidoptera)

## Just Higher_taxon as the predictor
cl_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID) + (1 | SS), data = coleoptera_lepidoptera)

## Make a null, intercept only model
cl_m7 <- lmer(logSR ~ (1 | Source_ID) + (1 | SS), data = coleoptera_lepidoptera)

## Compare the models
AIC(cl_m1, cl_m4, cl_m5, cl_m6, cl_m7)
model.sel(cl_m1, cl_m4, cl_m5, cl_m6, cl_m7)

# cl_m1 still optimal
```
### Simulate coleoptera_lepidoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(cl_m1, order = "Coleoptera")

# relevel model with Lepidoptera as reference level
coleoptera_lepidoptera <- within(coleoptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = coleoptera_lepidoptera)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
coleoptera_lepidoptera_plot_SR_log <- combine_and_plot(col_eff, lep_eff, diversity = "Log_SR")

# Extract f ratio, df and p value
anova(cl_m1)

# Check residuals
hist(residuals(cl_m1))

# Check r2
r.squaredGLMM(cl_m1)

# Extract correlation coefficients
corr <- as.matrix(summary(cl_m1)$coefficients)

# Split the coefficients into two groups
col <- corr[2:6, 1] 
lep <- corr[8:nrow(corr), 1] 

plot(col, lep)
cor(col, lep)

coleoptera_lepidoptera_plot_AB_log ; coleoptera_lepidoptera_plot_SR_log

rm(col_eff, corr, cl_m1, cl_m4, cl_m5, cl_m6, cl_m7, lep_eff, lep_m1, col, lep, coleoptera_lepidoptera)
```



## GGSAVE
```{r}
dip_col <- ggarrange(diptera_coleoptera_plot_AB_log, diptera_coleoptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/dip_col.png", dip_col, height = 7, width = 7)

dip_hemi <- ggarrange(diptera_hemiptera_plot_AB_log, diptera_hemiptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/dip_hemi.png", dip_hemi, height = 7, width = 7)

dip_hym <- ggarrange(diptera_hymenoptera_plot_AB_log, diptera_hymenoptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/dip_hym.png", dip_hym, height = 7, width = 7)
     
dip_lep <- ggarrange(diptera_lepidoptera_plot_AB_log, diptera_lepidoptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/dip_lep.png", dip_lep, height = 7, width = 7)         

hym_lep <- ggarrange(hymenoptera_lepidoptera_plot_AB_log, hymenoptera_lepidoptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/hym_lep.png", hym_lep, height = 7, width = 7)    

hym_hemi <- ggarrange(hymenoptera_hemiptera_plot_AB_log, hymenoptera_hemiptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/hym_hemi.png", hym_hemi, height = 7, width = 7)    

hym_col <- ggarrange(hymenoptera_coleoptera_plot_AB_log, hymenoptera_coleoptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/hym_col.png", hym_col, height = 7, width = 7)    

hemi_lep <- ggarrange(hemiptera_lepidoptera_plot_AB_log, hemiptera_lepidoptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/hemi_lep.png", hemi_lep, height = 7, width = 7)  

hemi_col <- ggarrange(hemiptera_coleoptera_plot_AB_log, hemiptera_coleoptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/hemi_col.png", hemi_col, height = 7, width = 7)    

col_lep <- ggarrange(coleoptera_lepidoptera_plot_AB_log, coleoptera_lepidoptera_plot_SR_log, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("./figures/log_figures_bycatch_removed/col_lep.png", col_lep, height = 7, width = 7)  

```

### Calculate number of studies and sites for 5/4 order model
```{r}

# 7 studies, 109 sites that compare all five orders
five_ss <- model_data_ab %>% dplyr::select(SS, Order) %>% unique() %>% group_by(SS) %>% tally() %>% filter(n == 5) %>% pull(SS) %>% unique()
five_ssbs <- model_data_ab %>% dplyr::select(SSBS, Order) %>% unique() %>%  group_by(SSBS) %>% 
  tally() %>% filter(n == 5) %>% pull(SSBS) %>% unique()
# Subset our model data for studies that include all five orders
all_five <- model_data_ab %>% filter(SS %in% five_ss) 

# 9 studies, 130 sites that compare 4 orders
four_ss <- model_data_ab %>% dplyr::select(SS, Order) %>% unique() %>% group_by(SS) %>% tally() %>% filter(n == 4) %>% pull(SS) %>% unique()
all_four <- model_data_ab %>% filter(SS %in% four_ss) 
four_ssbs <- model_data_ab %>% dplyr::select(SSBS, Order) %>% unique() %>%  group_by(SSBS) %>% 
  tally() %>% filter(n == 4) %>% pull(SSBS) %>% unique()

# 16 studies and 239 sites that compare >= 4 orders
fourplus_ss <- model_data_ab %>% dplyr::select(SS, Order) %>% unique() %>% group_by(SS) %>% tally() %>% filter(n >= 4) %>% pull(SS) %>% unique()
fourplus_studies <- model_data_ab %>% filter(SS %in% fourplus_ss)
fourplus_ssbs <- model_data_ab %>% dplyr::select(SSBS, Order) %>% unique() %>%  group_by(SSBS) %>% 
  tally() %>% filter(n >= 4) %>% pull(SSBS) %>% unique()

```