---
title: "SSI_analysis"
author: "Justin Isip"
date: "2023-04-07"
output: html_document
---


## Requires your dataframe to have a column called LU_nvh which has the levels "intact" and "converted". Make sure that intact is the reference level.
## Alternatively, if you're looking at intensification or another continuous scale, this should be a continuous variable also with the column name LU_nvh.

# Load packages
```{r}
library(tidyverse)
library(arm)
```

# Subset the df 
```{r}
df <-
  model_data_ab %>% 
  # Select the variables of interest
  dplyr::select(SS, SSBS, Order, Predominant_land_use, TotalAbundance, Sampling_method, Region) %>% 
  # Create a new column for the two coarse land-use classes
  mutate(
    LU_nvh =
      case_when(
        Predominant_land_use == "Primary" | Predominant_land_use == "Young secondary vegetation" | Predominant_land_use == "Intermediate secondary vegetation" |
          Predominant_land_use == "Mature secondary vegetation" ~ "intact",
        TRUE ~ "converted"
      )) %>%
      distinct()
```

# Relevel and factor
```{r}
df$LU_nvh <- as.factor(df$LU_nvh)
df$LU_nvh <- relevel(df$LU_nvh, ref = "intact")
```

# Models
```{r}
# bayesglm(Abundance of order_i in study_j ~ Intact vs converted, family = poisson, data = data[order == i & study == j, ])


# First -inf is AD1_2002_Vazquez 1 for hemiptera

hemi_vaz <- df %>% filter(SS == "AD1_2002__Vazquez 1") %>% filter(Order == "Hemiptera")
m1 <- bayesglm(TotalAbundance ~ LU_nvh, family = poisson(link = "log"), data = hemi_vaz)
summary(m1)


# -inf in AD1_2002_Vazquez 1 for lepidoptera

lep_vaz <- df %>% filter(SS == "AD1_2002__Vazquez 1") %>% filter(Order == "Lepidoptera")
m2 <- bayesglm(TotalAbundance ~ LU_nvh, family = poisson(link = "log"), data = lep_vaz)
summary(m2)


# inf in DI1_2012__Muchane 1 for lepidoptera

lep_muc <- df %>% filter(SS == "DI1_2012__Muchane 1") %>% filter(Order == "Lepidoptera")
m3 <- bayesglm(TotalAbundance ~ LU_nvh, family = poisson(link = "log"), data = lep_muc)
summary(m3)

# -inf in VB1_2008a_Smith 1 for hemiptera

hemi_smith <- df %>% filter(SS == "VB1_2008a_Smith 1") %>% filter(Order == "Hemiptera")
m4 <- bayesglm(TotalAbundance ~ LU_nvh, family = poisson(link = "log"), data = hemi_smith)
summary(m4)





df %>% filter(SS == "AD1_2002__Vazquez 1") %>% View()

# inf in X for X

X_X <- df %>% filter(SS == "") %>% filter(Order == "")
m <- bayesglm(TotalAbundance ~ LU_nvh, family = poisson(link = "log"), data = x)
summary(m)

```


# How many models I need to run to calc LRR for all the orders in a given study for my whole df
```{r}
# 146 models to run for the whole table
sum(!is.na(a$LRR_Col)) # 36
sum(!is.na(a$LRR_Dip)) # 26
sum(!is.na(a$LRR_Hemi)) # 26
sum(!is.na(a$LRR_Hym)) # 33
sum(!is.na(a$LRR_Lep)) # 25
```

# Dispersion test function
```{r}
dispersiontest <- function(object, trafo = NULL, alternative = c("greater", "two.sided", "less"))
{
  if(!inherits(object, "glm") || family(object)$family != "poisson")
    stop("only Poisson GLMs can be tested")
  alternative <- match.arg(alternative)
  otrafo <- trafo
  if(is.numeric(otrafo)) trafo <- function(x) x^otrafo
  
  y <- if(is.null(object$y)) model.response(model.frame(object)) else object$y
  yhat <- fitted(object)
  aux <- ((y - yhat)^2 - y)/yhat
  
  if(is.null(trafo)) {
    STAT <- sqrt(length(aux)) * mean(aux)/sd(aux)
    NVAL <- c(dispersion = 1)
    EST <- c(dispersion = mean(aux) + 1)    
  } else {
    auxreg <- lm(aux ~ 0 + I(trafo(yhat)/yhat))
    STAT <- as.vector(summary(auxreg)$coef[1,3])
    NVAL <- c(alpha = 0)
    EST <- c(alpha = as.vector(coef(auxreg)[1]))
  }
  
  rval <- list(statistic = c(z = STAT),
               p.value = switch(alternative,
                                "greater" = pnorm(STAT, lower.tail = FALSE),
                                "two.sided" = pnorm(abs(STAT), lower.tail = FALSE)*2,
                                "less" = pnorm(STAT)),
               estimate = EST,
               null.value = NVAL,
               alternative = alternative,
               method = switch(alternative,
                               "greater" = "Overdispersion test",
                               "two.sided" = "Dispersion test",
                               "less" = "Underdispersion test"),
               data.name = deparse(substitute(object)))
  class(rval) <- "htest"
  return(rval)
}
```


# SSI calc function
```{r}
SSI_calc<-function(data, fac = TRUE, type = "abundance", remove_uncertain = TRUE){
  
  ## function to extract (and back-transform) estimates from the model output
  get_estimates<-function(m){
      fam<-m$family$family
      if(fam=="poisson"|fam=="quasipoisson"){
        transform_fun<-exp
      }else{
        transform_fun<-inv.logit
      }
      
      intercept<-coef(m)[1]
      estimate<-coef(m)[2]
      exp_estimate<-transform_fun(coef(m)[2])
      exp_intercept<-transform_fun(coef(m)[1])
      se <- summary(m)$coefficients[2,2]
      ci <- transform_fun(c(coef(m)[2]-1.96*se, coef(m)[2]+1.96*se))
    return(list("intercept"=intercept,"estimate"=estimate,"exp_estimate" = exp_estimate,"exp_intercept" = exp_intercept,"se"= se, "ci" = ci))
  } 
  
  require(arm)
  require(boot)
  require(AER)
  
  data$present<-data$Measurement>0 ## add in a binary value for species presence at a site
  
  if(type=="abundance-present"){
    data<-data[data$present=="TRUE",] # only use data where the species is present
  }
  
  uncert<-sum(data$Resolution_entered=="Uncertain")
  if(remove_uncertain ==TRUE){
    data<-data[data$Resolution_entered!="Uncertain",] # remove species with uncertain taxa name
    print(paste("removing",uncert, "rows with uncertain taxa name"))
  }else{
    print(paste("Warning:", uncert, "rows with uncertain taxa name")) ## otherwise warn that they're still in there
  }
  
  if(type%in%c("abundance","abundance-present")){
    occ<-sum(data$Diversity_metric_type!="Abundance") ## how many rows are not abundance measurements?
    print(paste("removing", occ, "rows with non-abundance data")) ## warn that they're being removed
    data<-data[data$Diversity_metric_type=="Abundance",] # only use abundance measurements
  }
  
  data<-droplevels(data)
  ##Check that you have enough data for each species
  
  speciestable<-as.data.frame.table(tapply(data$LU_nvh, data$Best_guess_binomial, function(x)length(unique(x))))
  sprem<-sum(speciestable$Freq<2)
  
  species<-speciestable$Var1[speciestable$Freq>1]
  print(paste("removing",sprem, "species without data from intact and converted sites"))
  
  
  
  if(fac == TRUE){ # set up the dataframes for output
    species.coef<-as.data.frame(matrix(NA,nrow = length(species),ncol = 27, dimnames = list(c(),c("Species", "Genus","Family","COLTAXON","err_family","ranefs","intercept","est","exp_est","exp_intercept","se","lower","upper","significance","Sources","Studies","Sites","meanIntact","meanConverted","subAbund","newsubAbund","Sites_intact","Sites_converted","Sites_intact_zero","Sites_converted_zero","Perc_intact_sites_zero","Perc_converted_sites_zero"))))
  }else{
    species.coef<-as.data.frame(matrix(NA,nrow = length(species),ncol = 30, dimnames = list(c(),c("Species", "Genus","Family","COLTAXON","err_family","ranefs","intercept","est","exp_est","exp_intercept","se","lower","upper","significance","Sources","Studies","Sites","meanMinimal","meanLight","meanIntense","subAbund","Sites_minimal","Sites_light","Sites_intense","Sites_minimal_zero","Sites_light_zero","Sites_intense_zero","Perc_minimal_sites_zero","Perc_light_sites_zero","Perc_intense_sites_zero"))))
  }
    
  for(i in 1:length(species)){
    print(i)
    dat.i<-data[data$Best_guess_binomial==species[i],] # subset to the correct species
    dat.i<-droplevels(dat.i)

    species.coef$subAbund[i] <- sum(dat.i$Measurement)
    
    # get the easy data out
    species.coef$Species[i]<-paste(species[i])
    species.coef$Family[i]<-paste(dat.i$Family[1])
    species.coef$COLTAXON[i]<-paste(dat.i$Taxon[1])
    species.coef$Genus[i]<-paste(strsplit(as.character(species[i]),"_")[[1]][1])
    species.coef$Sources[i] <- length(unique(dat.i$Source_ID))
    species.coef$Studies[i] <- length(unique(dat.i$Study_name))
    species.coef$Sites[i] <- nrow(dat.i)
    
    
    if(fac == TRUE){
      species.coef$meanIntact[i] <- mean(dat.i$Measurement[dat.i$LU_nvh=="intact"])
      species.coef$meanConverted[i] <- mean(dat.i$Measurement[dat.i$LU_nvh=="converted"])
      species.coef$Sites_intact[i] <- sum(dat.i$LU_nvh=="intact")
      species.coef$Sites_converted[i] <- sum(dat.i$LU_nvh=="converted")
      species.coef$Sites_intact_zero[i]<-nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh=="intact",])
      species.coef$Sites_converted_zero[i]<-nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh=="converted",])
      species.coef$Perc_intact_sites_zero[i]<-(nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh=="intact",])/sum(dat.i$LU_nvh=="intact"))*100
      species.coef$Perc_converted_sites_zero[i]<-(nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh=="converted",])/sum(dat.i$LU_nvh=="converted"))*100
    }else{
      species.coef$meanMinimal[i] <- sum(dat.i$Measurement[dat.i$LU_nvh==1])/sum(dat.i$LU_nvh==1)
      species.coef$meanLight[i] <- sum(dat.i$Measurement[dat.i$LU_nvh%in%2:3])/sum(dat.i$LU_nvh==2)
      species.coef$meanIntense[i] <- sum(dat.i$Measurement[dat.i$LU_nvh%in%2:3])/sum(dat.i$LU_nvh==3)
      species.coef$Sites_minimal[i] <- sum(dat.i$Measurement[dat.i$LU_nvh==1])
      species.coef$Sites_light[i] <- sum(dat.i$Measurement[dat.i$LU_nvh==2])
      species.coef$Sites_intense[i] <- sum(dat.i$Measurement[dat.i$LU_nvh==3])
      species.coef$Sites_minimal_zero[i]<- nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh==1,])
      species.coef$Sites_light_zero[i]<- nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh==2,])
      species.coef$Sites_intense_zero[i]<- nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh==3,])
      species.coef$Perc_minimal_sites_zero[i]<- (nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh==1,])/sum(dat.i$LU_nvh==1))*100
      species.coef$Perc_light_sites_zero[i]<- (nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh==2,])/sum(dat.i$LU_nvh==2))*100
      species.coef$Perc_intense_sites_zero[i]<- (nrow(dat.i[dat.i$Measurement==0&dat.i$LU_nvh==3,])/sum(dat.i$LU_nvh==3))*100
      
    }
    
    # do the models and get the estimates
    if(type%in%c("abundance","abundance-present")){      
      species.coef$newsubAbund[i] <- sum(dat.i$Measurement)
      
      if(length(unique(dat.i$SS))>1){ # if there's more than one study
        m<-bayesglm(Measurement~LU_nvh+SS, family = poisson(link="log"),data = dat.i)
        if(dispersiontest(m)$p.value<0.05){ # if the model is overdispersed, fit a quasipoisson model
          m<-bayesglm(Measurement~LU_nvh+SS, family = quasipoisson(link="log"),data = dat.i)
        }
      }else{
        m<-bayesglm(Measurement~LU_nvh, family =  poisson(link="log"),data = dat.i)
        if(dispersiontest(m)$p.value<0.05){
          m<-bayesglm(Measurement~LU_nvh, family = quasipoisson(link="log"),data = dat.i)
        }
      }
    }
    
    if(type=="occurrence"){
      species.coef$newsubAbund[i] <- NA
      if(length(unique(dat.i$SS))>1){ # if there's more than one study
        m<-bayesglm(present~LU_nvh+SS, family = binomial(link = "logit"),data = dat.i)
        if(dispersiontest(m)$p.value<0.05){ # if the model is overdispersed, fit a quasibinomial model
          m<-bayesglm(present~LU_nvh+SS, family = quasibinomial(link = "logit"),data = dat.i)
        }
      }else{
        m<-bayesglm(present~LU_nvh, family = binomial(link = "logit"),data = dat.i)
        if(dispersiontest(m)$p.value<0.05){
          m<-bayesglm(present~LU_nvh, family = quasibinomial(link = "logit"),data = dat.i)
        }
      }
    } # use this otherwise predicted probabilities of 0 and 1 become a major problem
    
    ranef.form<-"none" ## relic from when I was trying it out with random effects as well
    species.coef$ranefs[i]<-ranef.form
    species.coef$significance[i]<-summary(m)$coefficients[2,4] # slope significance
    ests<-get_estimates(m)
    species.coef$intercept[i]<-ests$intercept
    species.coef$est[i]<-ests$estimate
    species.coef$exp_est[i] <- ests$exp_estimate
    species.coef$exp_intercept[i] <- ests$exp_intercept
    species.coef$se[i] <- ests$se
    species.coef$lower[i] <- ests$ci[1]
    species.coef$upper[i]<-ests$ci[2]
    species.coef$err_family[i]<-m$family$family
    
  }
  return(species.coef)
}
```

