---
title: "Untitled"
output: html_document
date: "2026-03-11"
---

# Clear the work space

```{r}
rm(list=ls())
```

# Read in the packages 
```{r}
library(readxl)
library(tidyverse)
library(here)
library(car)
library(readr)
library(qs)
library(rgbif)
library(lme4)
library(stringr)
library(lmerTest)
library(merTools) 
library(performance)
library(DHARMa)
library(here)
```

# Read in Logghe et al 2025 which has body size data for ~ 4800 arthropod species and makes the bulk of the GLiTRS trait data set

```{r}
# Upload logghe et al
logghe <- read_excel(here("data", "Body_Size_data.xlsx"))

# keep only first 8 cols
logghe <- logghe %>% dplyr::select(1:8)

# Make first row the header
colnames(logghe) <- as.character(logghe[1, ])
logghe <- logghe[-1, ] # drop first row

# Create a genus and species column
logghe <- logghe %>%
  separate(scientificName, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop")

# Combine genus and species to create Best_guess_binomial which I will use to merge 
logghe <- logghe %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Reorder columns
logghe <- logghe %>% dplyr::select(order, family, Genus, Species, Best_guess_binomial, Trait_Value) %>% 
  rename(Family = family) %>% 
  rename(Order = order)

logghe$Trait_Value <-  as.numeric(logghe$Trait_Value)
```

# Read and tidy Geoffrey Su's ant trait data
```{r}
# Load the data
geoffreys_ants <- read_excel(here("data", "SuFormacidaeBodyLength-2014-05-23.xlsx")) %>%  strings2factors()

summary(geoffreys_ants)

# Rename body length average column
geoffreys_ants <- geoffreys_ants %>% rename(avg_body_length_mm = "Body.Length..mm...worker..AVG")

# Lets remove those rows with NAs for body length
geoffreys_ants <- geoffreys_ants %>%
  filter(!is.na(avg_body_length_mm) & avg_body_length_mm != 0) # 559 rows with no body length data, 2 rows where BL = 0

# Combine genus and species  
geoffreys_ants <- geoffreys_ants %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Simplify ants
geoffreys_ants <- geoffreys_ants %>% dplyr::select(Genus, Best_guess_binomial, avg_body_length_mm)

# Tidy
geoffreys_ants <- geoffreys_ants %>%
  mutate(
    Order = "Hymenoptera",
    Family = "Formicidae"
  ) %>% 
  rename(Trait_Value = avg_body_length_mm) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop") %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)
# ants ready
```

# Read and tidy Melanie Edgars beetle trait data
```{r}
# Load Melanies beetle trait data
melanie_beetles <- read_excel(here("data", "EdgarBeetleLength-2014-01-09.xlsx"))

# Combine genus and species for merging
melanie_beetles <- melanie_beetles %>% mutate(Best_guess_binomial = str_c(GENUS, SPECIES, sep = " "))

# Calculate average b.length
melanie_beetles <- melanie_beetles %>%
  separate(B.LENGTH, into = c("b_min", "b_max"), sep = "-", fill = "right", remove = FALSE) %>% # fill = "right" allows single values to stay in b_min while b_max becomes NA.
  mutate(
    b_min = as.numeric(trimws(b_min)), # trimws() removes any leading/trailing spaces 
    b_max = as.numeric(trimws(b_max)),
    b.length_avg = if_else(
      is.na(b_max),
      b_min,  # If no range, use the min value
      (b_min + b_max) / 2  # Else if range, compute the average
    ))

# Prepare and tidy beetles
melanie_beetles <- melanie_beetles %>% 
  mutate(Order = "Coleoptera") %>% 
  dplyr::select(Order, FAMILY, GENUS, SPECIES, Best_guess_binomial, b.length_avg) %>% 
  rename(Family = FAMILY) %>% 
  rename(Genus = GENUS) %>% 
  rename(Species = SPECIES) %>% 
  rename(Trait_Value = b.length_avg)
```

# Read and tidy Rachel Tans bee trait data
```{r}
# Load Rachel bee trait data
rachel_bee <- read_excel(here("data", "BeeTraits_P2_RachelTan.xlsx"))

# Create a genus and species column
rachel_bee <- rachel_bee %>%
  separate(CorrectedName, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop")

# Filter for columns of interest
rachel_bee <- rachel_bee %>% dplyr::select(HigherTaxon, Family, Genus, Species, CorrectedName, MidRangeBodyLength) %>% drop_na() # drop those sp. with NA for body length

# Remove rows identified as "sp." or inside brackets e.g., (Dialictus) or (Psithyrus) which are subgenera and not a true species name
rachel_bee <- rachel_bee %>%
  filter(
    !str_detect(Species, "sp\\.") &         # remove rows with "sp."
    !str_detect(Species, "\\(.*\\)")        # remove rows with anything in brackets
  )

# Rename for merging
rachel_bee <- rachel_bee %>% 
  rename(Best_guess_binomial = CorrectedName) %>% 
  rename(Order = HigherTaxon) %>% 
  rename(Trait_Value = MidRangeBodyLength)

```

# Read and tidy global ants database

```{r}
# Read in Global Ants trait data
globalants <- read.csv(here("data", "globalants.csv"))

# Filter for columns of interest
globalants <- globalants %>% dplyr::select(Genus, Species, Whole.body.length..mm.) %>% droplevels()

# Filter out those rows without a body length value
globalants <- globalants %>% filter(`Whole.body.length..mm.` != "Not in paper")

# Create best guess binomial
globalants <- globalants %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Convert body length to numeric for summarizing
globalants <- globalants %>%
  mutate(Whole.body.length..mm. = as.numeric(Whole.body.length..mm.))

# Calculate an average for duplicate species entries
globalants <- globalants %>%
  group_by(Best_guess_binomial) %>%
  summarise(
    avg_body_length_mm = mean(`Whole.body.length..mm.`, na.rm = TRUE),
    .groups = "drop"
  )

# Update rows with dodgy names, row 31
globalants$Best_guess_binomial[31] <- "Formica oreas"
# row 45
globalants$Best_guess_binomial[45] <- "Leptothorax muscorum"

# Recreate genus and species
globalants <- globalants %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE) %>% 
  mutate(Order = "Hymenoptera", Family = "Formicidae") %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, avg_body_length_mm) %>% 
  rename(Trait_Value = avg_body_length_mm)

```

# Data from AnthropInsect

```{r}
# Read in the trait data
anthroptraits <- read_xlsx(here("data", "AnthropInsect_V1.xlsx")) 

# first remove eusocial insects as their body size values consider the queen's size.
anthroptraits <- anthroptraits %>% filter(Sociability != "Eusocial")

# Select cols of interest, filter out missing body size values, rename Species
anthroptraits <- anthroptraits %>% dplyr::select(Order, Family, Genus, Species, Body_size_adult_mean) %>% filter(!is.na(Body_size_adult_mean)) %>% rename(Best_guess_binomial = Species)

# Recreate species column
anthroptraits <- anthroptraits %>%
  mutate(Species = sapply(strsplit(Best_guess_binomial, " "), `[`, 2))

# Reorder cols
anthroptraits <- anthroptraits %>% dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Body_size_adult_mean) %>% rename(Trait_Value = Body_size_adult_mean)
```

# Bee Traits from Emma Flynn

```{r}
# Read in the trait data
flynn_bees <- read.csv(here("data", "FlynnHymenoptera.csv"))

# select cols of interest and filter out missing data
flynn_bees <- flynn_bees %>% dplyr::select(Family, Genus, Species, Body.Length.Mean..mm.) %>% filter(!is.na(Body.Length.Mean..mm.))

# There are some rows which only have a size value at the genus level

# Lets create Order and Best_guess_binomial, those rows without a species name will get an NA, then organise the cols
flynn_bees <- flynn_bees %>% mutate(Order = "Hymenoptera", Best_guess_binomial = str_c(Genus, Species, sep = " ")) %>% dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Body.Length.Mean..mm.) %>% 
  rename(Trait_Value = Body.Length.Mean..mm.)

```

# Odonata Phenotypic Database

```{r}
# Read in the trait data
odonata <- read.csv(here("data", "odonata_phenotypic_database.csv")) %>% strings2factors()

summary(odonata) # barely any data for forewing length so we'll use body lengths

odonata <- odonata %>% dplyr::select(Family, Genus, Species, GenusSpecies, body_lengths) %>% filter(!is.na(body_lengths))

odonata <- odonata %>% mutate(Order = "Odonata") %>% dplyr::select(Order, Family, Genus, Species, GenusSpecies, body_lengths) %>% rename(Best_guess_binomial = GenusSpecies, Trait_Value = body_lengths)
```

# Morphological trait database of saproxylic beetles

```{r}
# Read in the data
saproxylic_beetles <- read.csv(here("data", "Hagge_etal_saproxylic_beetle_morphological_trait_database_20210426.csv"))

summary(saproxylic_beetles)

# Select cols of interest
saproxylic_beetles <- saproxylic_beetles %>% dplyr::select(family, species, body_length) %>% filter(!is.na(body_length))

saproxylic_beetles <- saproxylic_beetles %>% 
  mutate(Order = "Coleoptera") %>% 
  separate(species, into = c("Genus", "Species"), sep = "_") %>% # seperate species into genus and species
  mutate(Best_guess_binomial = paste(Genus, Species, sep = " ")) # create best_guess_binomial

saproxylic_beetles <- saproxylic_beetles %>% 
  dplyr::select(Order, family, Genus, Species, Best_guess_binomial, body_length) %>% 
  rename(Family = family, Trait_Value = body_length)

```

# Data from: A summary of eight traits of Coleoptera, Hemiptera, Orthoptera and Araneae, occurring in grasslands in Germany

```{r}
# Read in the trait data
col_hemi_german_grasslands <- read_delim(here("data", "ArthropodSpeciesTraits.txt")) %>% strings2factors()

col_hemi_german_grasslands <- col_hemi_german_grasslands %>%  separate(SpeciesID, into = c("Genus", "Species"), sep = " ", remove = FALSE)

# Dicymbium nigrum brevisetosum gets discarded as it has three names but it is a spider so no big issue

col_hemi_german_grasslands <- col_hemi_german_grasslands %>% dplyr::select(Order, Family, Genus, Species, SpeciesID,  Body_Size) %>% rename(Best_guess_binomial = SpeciesID, Trait_Value = Body_Size)
```

# PREDICTS Trait dataset: TraitDataFournier2017

```{r}
fournier_predicts <- read.csv(here("data", "TraitData_Fournier2017.csv")) 

# 1. Set the column names to the first row
colnames(fournier_predicts) <- as.character(fournier_predicts[1, ])

# 2. Remove the first row
fournier_predicts <- fournier_predicts[-1, ]

# Remove redundant rows
fournier_predicts <- fournier_predicts[1:23, ]

# Two columns had the same name - fix
colnames(fournier_predicts) <- make.unique(colnames(fournier_predicts))

# Create order col
fournier_predicts <- fournier_predicts %>% mutate(Order = "Orthoptera")

# Select cols of interest
fournier_predicts <- fournier_predicts %>% dplyr::select(Order, Family, Genus, Species.1, Species, BS)

# Convert BS (body size) values from cm to mm
fournier_predicts <- fournier_predicts %>%
  mutate(BS = as.numeric(BS) * 10) # replaces BS with mm values

fournier_predicts <- fournier_predicts %>% rename(Species = Species.1, Best_guess_binomial = Species, Trait_Value = BS)

```

# PREDICTS Trait dataset: Perrin2020

```{r}
perrin2020_traits <- read_xlsx(here("data", "TraitData_Perrin2020.xlsx"))

# select cols of interest
perrin2020_traits <- perrin2020_traits %>% dplyr::select(Species, `Mean body-size (mm)`)

# Update row 7 with subspecies name
perrin2020_traits$Species[7] <- "Bodiloides ictericus"

# row 14
perrin2020_traits$Species[14] <- "Copris hispanus"

# row 19
perrin2020_traits$Species[19] <- "Gymnopleurus flagellatus"

# Create cols
perrin2020_traits <- perrin2020_traits %>% rename(Best_guess_binomial = Species, Trait_Value = `Mean body-size (mm)`) %>% separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE) %>% mutate(Order = "Coleoptera", Family = "Scarabaeidae")

# Organise cols
perrin2020_traits <- perrin2020_traits %>% dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)
```

# PREDICTS Trait dataset: Pianoetal2020

```{r}
piano2020_traits <- read_xlsx(here("data", "Pianoetal_traits.xlsx"))

piano2020_traits <- piano2020_traits %>% separate(`Species binomial (Genus species)`, into = c("Genus", "Species"), sep = " ", remove = FALSE) 

# Select cols of interest
piano2020_traits <- piano2020_traits %>% dplyr::select(Taxa, Family, Genus, Species, `Species binomial (Genus species)`, `Body Size Trait`)

# Rename
piano2020_traits <- piano2020_traits %>% rename(Order = Taxa, Best_guess_binomial = `Species binomial (Genus species)`, Trait_Value = `Body Size Trait`)
```

# PREDICTS Trait dataset: Gagne&Fahrig2011

```{r}
gagnefahrig2011_traits <- read_xlsx(here("data", "Gagne&Fahrig2011.xlsx"))

# Update row 29 with subspecies name
gagnefahrig2011_traits$`Species binomial (Genus species)`[29] <- "Bembidion quadrimaculatum"

# row 75
gagnefahrig2011_traits$`Species binomial (Genus species)`[75] <- "Poecilus lucublandus"

# row 88
gagnefahrig2011_traits$`Species binomial (Genus species)`[88] <- "Sphaeroderus canadensis"

# row 89
gagnefahrig2011_traits$`Species binomial (Genus species)`[89] <- "Sphaeroderus stenostomus"


gagnefahrig2011_traits <- gagnefahrig2011_traits %>% 
  separate(`Species binomial (Genus species)`, into = c("Genus", "Species"), sep = " ", remove = FALSE) %>% 
  mutate(Order = "Coleoptera") %>% 
  rename(Best_guess_binomial = `Species binomial (Genus species)`) %>% 
  rename(Trait_Value = `Body Length`)

gagnefahrig2011_traits <- gagnefahrig2011_traits %>% dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)
```

# PREDICTS Trait dataset: Diptera traits from Jess Wood

```{r}
# Read in the data
diptera_wood_traits <- read.csv(here("data", "WoodDiptera.csv"))

summary(diptera_wood_traits)

# filter out NAs for body length
diptera_wood_traits <- diptera_wood_traits %>% filter(!is.na(Adult.body.length.average))

# Convert rows with no species data into NAs prior to creating BGB
diptera_wood_traits <- diptera_wood_traits %>%
  mutate(Species = na_if(Species, ""))

# create cols
diptera_wood_traits <- diptera_wood_traits %>%
  mutate(Order ="Diptera") %>% 
  mutate(Best_guess_binomial = ifelse(is.na(Species), NA, paste(Genus, Species)))

# Rename col
diptera_wood_traits <- diptera_wood_traits %>% rename(Trait_Value = Adult.body.length.average)

# select cols
diptera_wood_traits <- diptera_wood_traits %>% dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)

```

# Bee functional Trait Database

```{r}
# Load the data
bee_functional_traits_database <- read.csv(here("data", "Sup Table 5 Trait Data Template.csv")) %>% strings2factors()

# Those species in the database with less than two words in their acceptednameusage
bee_functional_traits_database %>%
  filter(str_count(acceptedNameUsage, "\\S+") < 2)

# There are 196 rows which have a name like "C.OCTODENTATA" in the acceptedNameUsage column, they all use ITD as their measurement type for body size so not to worry for now..

# filter only for body length or body size data for now
bee_functional_traits_database <- bee_functional_traits_database %>% filter(measurementType %in% c("body length", "body size"))

# Select the cols of interest
bee_functional_traits_database <- bee_functional_traits_database %>% dplyr::select(acceptedNameUsage, measurementValue)

# Create the cols we need
bee_functional_traits_database <- bee_functional_traits_database %>%
  mutate(
    Genus = word(acceptedNameUsage, 1),
    Species = word(acceptedNameUsage, 2),
    Best_guess_binomial = paste(Genus, Species)) %>% 
    mutate(Order = "Hymenoptera") %>% 
    rename(Trait_Value = measurementValue) %>% 
    dplyr::select(Order, Genus, Species, Best_guess_binomial, Trait_Value)

# Get a list of all of the unique species names to match to the GBIF backbone
species_list <- unique(paste(bee_functional_traits_database$Genus, bee_functional_traits_database$Species))

# Extract order and family names for all of the species in my df from GBIF
taxonomy_gbif <- bind_rows(lapply(species_list, function(sp) {
  result <- name_backbone(name = sp, rank = "species")
  data.frame(
    Best_guess_binomial = sp,
    Family = ifelse(!is.null(result$family), result$family, NA)
  )
}))

# 4 species not matched to the gbif taxonomy 

# Merge back into original df
bee_functional_traits_database <- bee_functional_traits_database %>%
  left_join(taxonomy_gbif, by = "Best_guess_binomial")

# Reorder columns
bee_functional_traits_database <- bee_functional_traits_database %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)

rm(taxonomy_gbif, species_list)
```

# **** (DON'T RUN THIS CHUNK!! TAKES AGES SO HAVE SAVED THE OUTPUT READY TO LOAD IN THE NEXT CHUNK) ****
# Lets have a look at the GRIT database 

```{r, eval = FALSE}
grit <- read.csv(here("data", "grit.csv")) %>% strings2factors()

summary(grit)

# select cols of interest
grit <- grit %>% dplyr::select(verbatimIdentification, measurementType, measurementMethod, measurementValue., measurementUnit, measurementStatistic, associatedReferences)

# select body size values of interest
grit <- grit %>% filter(measurementType %in% c ("Body Length", "Body Mass", "Body size", "Body Size", "body total length", "Estimated Dry Body Mass", 
                                                "Forewing Length", "Wingspan")) %>% droplevels()

# Remove rows that have measurementstatistic == variance, as these rows which have a body size values that represent the var around the mean presumably rather than the actual measurement
# having a look at them, they are much lower than the mean, min or max values
grit <- grit %>% 
  filter(measurementStatistic != "variance") # 276 rows removed

# Split VerbatimIdentification into Genus and Species
grit <- grit %>%
  separate(verbatimIdentification, 
           into = c("Genus", "Species"), 
           sep = "[ _]",   # matches either a space or underscore
           remove = FALSE) # keep the original column; remove if not needed 


# 1262 rows have underscores in their name
sum(grepl("_", grit$verbatimIdentification))

# convert all underscores to spaces
grit <- grit %>%
  mutate(verbatimIdentification = str_replace(verbatimIdentification, "_", " "))

# Sense check
sum(grepl("_", grit$verbatimIdentification))  # should return 0 if all fixed

# Get a list of all of the unique species names to match to the GBIF backbone
species_list <- unique(paste(grit$Genus, grit$Species))

# Extract order and family names for all of the species in my df from GBIF
taxonomy_gbif <- bind_rows(lapply(species_list, function(sp) {
  result <- name_backbone(name = sp, rank = "species")
  data.frame(
    verbatimIdentification = sp,
    Family = ifelse(!is.null(result$family), result$family, NA),
    Order  = ifelse(!is.null(result$order),  result$order,  NA)
  )
}))

# Merge back into original df
grit <- grit %>%
  left_join(taxonomy_gbif, by = "verbatimIdentification")

# lets rename measurement value first
grit <- grit %>% 
  rename(measurementValue =  "measurementValue.")

# convert cm to mm 
grit <- grit %>%
  mutate(
    measurementValue = case_when(
      measurementUnit == "cm" ~ measurementValue * 10,
      TRUE ~ measurementValue  # leave all other rows unchanged
    ),
    measurementUnit = case_when(
      measurementUnit == "cm" ~ "mm",
      TRUE ~ measurementUnit  # leave all other rows unchanged
    )
  )

# sense check
# Should return 0 if all cm rows have been converted
sum(grit$measurementUnit == "cm", na.rm = TRUE)

# Check the distribution of units remaining
table(grit$measurementUnit)

# save cleaned grit database so we don't have to rerun the time consuming taxonomy matching to gbif backbone
#write_csv(grit, "/Users/jusi/Desktop/multi-groups/data/grit_cleaned.csv")

# clean environment
rm(species_list, taxonomy_gbif)
```


# So far I've updated the taxonomic information in the GRIT database using the GBIF backbone and converted cm to mm..
# time to clean it up further and get one body size value per species.. 

```{r}
# Read in data frame which is from the previous chunk but has been cleaned
grit <- read.csv(here("data", "grit_cleaned.csv")) %>% strings2factors()

# Filter only for body length variables (rather than mass, wingspan or forewing length)
grit_BS_only <- grit %>% filter(measurementType %in% c("Body Length", "Body size", "Body Size", "body total length")) %>% droplevels()

# check they are all in mm
table(grit_BS_only$measurementUnit)

# check there are only four types of body size metrics so I can calculate one body size value for each taxa
table(grit_BS_only$measurementStatistic)

# There are four potential measurement types for body size for each sp
# Individual, mean, max and min
# Each sp can have multiple different measurement types for body size, so for example species A has a mean, a max and a min.
# Species B has just an individual measurement and a mean
# I want to calculate a single body size value for each species
# The temporary plan is as follows:
# If the species has a mean value, it gets the mean of the means.
# If the species has no mean value but has a max or min value, it gets the mean of the max and/or min
# If hte species has no mean value and no max or min value, it gets the mean of the individual value. 

one_value_per_sp <- grit_BS_only %>%
  group_by(verbatimIdentification) %>% # group by species
  summarise(
    measurementValue = case_when(
      # Priority 1: mean of all "mean" values if any exist
      any(measurementStatistic == "mean", na.rm = TRUE) ~
        mean(measurementValue[measurementStatistic == "mean"], na.rm = TRUE),
      
      # Priority 2: mean of "max" and/or "min" values if no "mean" exists
      any(measurementStatistic %in% c("max", "min"), na.rm = TRUE) ~
        mean(measurementValue[measurementStatistic %in% c("max", "min")], na.rm = TRUE),
      
      # Priority 3: fallback to mean of "individual" value, if there are no means, min or max values
      any(measurementStatistic == "individual", na.rm = TRUE) ~
        mean(measurementValue[measurementStatistic == "individual"], na.rm = TRUE),
      # If nothing matches, return NA
      TRUE ~ NA_real_
    ),
    measurementStatistic = case_when(
      any(measurementStatistic == "mean",             na.rm = TRUE) ~ "mean",
      any(measurementStatistic %in% c("max", "min"),  na.rm = TRUE) ~ "mean_of_max_min", # update measurement statistic to represent the mean of the min and max body size value
      any(measurementStatistic == "individual",        na.rm = TRUE) ~ "individual",
      TRUE ~ NA_character_
    ),
    .groups = "drop"
  )

# Get one row per species (e.g. family, order) from the overall df for body lengths for grit
distinct_taxonomy_grit <- grit_BS_only %>%
  distinct(Order, Family, Genus, Species, verbatimIdentification) 

# So that we can merge the summarised mean body lengths back into the df giving us one mean body length for each taxa
grit_BS_only_cleaned <- distinct_taxonomy_grit %>%
  left_join(one_value_per_sp, by = "verbatimIdentification")

# use anti_join() to find rows in grit_BS_only_cleaned that have no match in logghe:
grit_BS_only_cleaned <- anti_join(
  grit_BS_only_cleaned,
  logghe,
  by = c("verbatimIdentification" = "Best_guess_binomial")
)

# Tidy up the df
grit_BS_only_cleaned <- grit_BS_only_cleaned %>% 
  rename(Best_guess_binomial = verbatimIdentification) %>% 
  rename(Trait_Value = measurementValue) %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)

# Clean up environment
rm(distinct_taxonomy_grit, grit_BS_only, one_value_per_sp)
```

# Lets now merge all the trait databases

```{r}
# create a list of all my trait datasets
df_list <- list(anthroptraits, 
                      bee_functional_traits_database, 
                      col_hemi_german_grasslands, 
                      diptera_wood_traits,
                      flynn_bees,
                      fournier_predicts,
                      gagnefahrig2011_traits,
                      geoffreys_ants,
                      globalants,
                      grit_BS_only_cleaned,
                      logghe,
                      melanie_beetles,
                      odonata,
                      perrin2020_traits,
                      piano2020_traits,
                      rachel_bee,
                      saproxylic_beetles)  

# Convert the column Trait_value to a numeric across all dataframes and bind all the trait datasets together
combined <- bind_rows(lapply(df_list, function(df) {
  df %>% mutate(Trait_Value = as.numeric(as.character(Trait_Value)))
}))

# Sensitivity checks
class(combined$Trait_Value)  # should return "numeric"

# Check row count matches 
sum(nrow(anthroptraits),
    nrow(bee_functional_traits_database),
    nrow(col_hemi_german_grasslands),
    nrow(diptera_wood_traits),
    nrow(flynn_bees),
    nrow(fournier_predicts),
    nrow(gagnefahrig2011_traits),
    nrow(geoffreys_ants),
    nrow(globalants),
    nrow(grit_BS_only_cleaned),
    nrow(logghe),
    nrow(melanie_beetles),
    nrow(odonata),
    nrow(perrin2020_traits),
    nrow(piano2020_traits),
    nrow(rachel_bee),
    nrow(saproxylic_beetles)
    )

# Check if there are any NAs in the separate datasets
sum(is.na(anthroptraits$Trait_Value))
sum(is.na(bee_functional_traits_database$Trait_Value))
sum(is.na(col_hemi_german_grasslands$Trait_Value))
sum(is.na(diptera_wood_traits$Trait_Value))
sum(is.na(flynn_bees$Trait_Value))
sum(is.na(fournier_predicts$Trait_Value))
sum(is.na(gagnefahrig2011_traits$Trait_Value))
sum(is.na(geoffreys_ants$Trait_Value))
sum(is.na(globalants$Trait_Value))
sum(is.na(grit_BS_only_cleaned$Trait_Value))
sum(is.na(logghe$Trait_Value))
sum(is.na(melanie_beetles$Trait_Value))
sum(is.na(odonata$Trait_Value))
sum(is.na(perrin2020_traits$Trait_Value))
sum(is.na(piano2020_traits$Trait_Value))
sum(is.na(rachel_bee$Trait_Value))
sum(is.na(saproxylic_beetles$Trait_Value))

# the row counts add up and no unexpected NAs were introduced, the merge has most likely worked correctly.

# Any NAs in the combined dataset?
sum(is.na(combined$Trait_Value))

rm(anthroptraits, 
    bee_functional_traits_database, 
                      col_hemi_german_grasslands, 
                      diptera_wood_traits,
                      df_list,
                      flynn_bees,
                      fournier_predicts,
                      gagnefahrig2011_traits,
                      geoffreys_ants,
                      globalants,
                      grit_BS_only_cleaned,
                      logghe,
                      melanie_beetles,
                      odonata,
                      perrin2020_traits,
                      piano2020_traits,
                      rachel_bee,
                      saproxylic_beetles)  
```


# Dealing with duplicates in the combined overall trait dataset

```{r}
# Looking at the combined trait dataset, there are duplicate rows for the same taxa, sometimes with differing body size values

# Lets calculate the mean body size value across duplicate rows
combined <- combined %>%
  group_by(Order, Family, Genus, Species, Best_guess_binomial) %>%  # add any other metadata columns here
  summarise(Trait_Value = mean(Trait_Value, na.rm = TRUE),
            .groups = "drop")

# Sensitivity check
n_distinct(combined$Best_guess_binomial) # 12016 distinct taxa
nrow(combined) # but 12407 total rows, i.e. more than one row per taxa

12407 - 12016  # 391 row discrepancy

# Number of total rows > number of distinct species in the cleaned data set - what"s going on?

# Sum rows have body size data just at the genus level , with no species info - lets filter those rows for now

# Rows where species or Best_guess_binomial are empty/NA (genus level only)
combined_genus_only <- combined %>%
  filter(is.na(Best_guess_binomial) | Best_guess_binomial == "" |
         is.na(Species) | Species == "")

# Rows where species and Best_guess_binomial are present (species level)
combined_sp_only <- combined %>%
  filter(!is.na(Best_guess_binomial) & Best_guess_binomial != "" &
         !is.na(Species) & Species != "")

nrow(combined_sp_only) # the overall combined trait data set which has only species level data (i.e. genus level only values removed) has 12210 rows
n_distinct(combined_sp_only$Best_guess_binomial) # but it has 12015 distinct taxa names

# There is a discrepancy between the total number of rows and the number of distinct best_guess_binomials (12210 - 12015 = 195). Let's find out why:

# Find which taxa are duplicated:
combined_duplicate_species <- combined_sp_only %>%
  group_by(Best_guess_binomial) %>%
  filter(n() > 1) %>%
  arrange(Best_guess_binomial) 

# The duplicates are for taxa who have been assigned to different families and therefore the group_by(all taxonomic columns in the df) fails as the duplicate rows
# are assigned to different families, it is tricky to work out which is the correct taxonomy, so I think best to rematch to the GBIF taxonomy backbone and go with that...

# Get a list of all of the unique duplicate species names to match to the GBIF backbone
species_list <- unique(paste(combined_duplicate_species$Genus, combined_duplicate_species$Species))

# Extract order and family names for all of the duplicate species in my df from GBIF
taxonomy_gbif <- bind_rows(lapply(species_list, function(sp) {
  result <- name_backbone(name = sp, rank = "species")
  data.frame(
    Best_guess_binomial = sp,
    Family = ifelse(!is.null(result$family), result$family, NA)
  )
}))

# Merge the family names from GBIF back into original df
combined_duplicate_species <- combined_duplicate_species %>%
  left_join(taxonomy_gbif, by = "Best_guess_binomial")

# Keep only the family.y column (which is the gbif taxonomy backbone) and order the columns in the format of all the other trait datasets  
combined_duplicate_species <- combined_duplicate_species %>% 
  dplyr::select(Order, Family.y, Genus, Species, Best_guess_binomial, Trait_Value) %>%  
  rename(Family = Family.y)

# Now that the duplicate taxa have the correct taxonomic info for family, lets recalculate their mean body size so that we have one value per species
combined_duplicate_species <- combined_duplicate_species %>% 
  group_by(Order, Family, Genus, Species, Best_guess_binomial) %>% 
  summarise(Trait_Value = mean(Trait_Value, na.rm = TRUE),
            .groups = "drop") # remove all grouping after summarising

# the number of rows for combined_duplicate_species is 194 but there are 191 distinct taxa, hmm?
n_distinct(combined_duplicate_species$Best_guess_binomial) # 191

# Still some duplicates with synonym names, we can use GBIF to get the current accepted name
combined_duplicate_species %>%
  group_by(Best_guess_binomial) %>%
  filter(n() > 1) %>%
  arrange(Best_guess_binomial) %>% 
  View()

# Lets get the duplicate species names
duplicate_names <- combined_duplicate_species %>%
  group_by(Best_guess_binomial) %>%
  filter(n() > 1) %>%
  distinct(Best_guess_binomial) %>%
  pull(Best_guess_binomial)
 
# Look up each duplicate in the GBIF backbone
accepted_names <- lapply(duplicate_names, function(species) {
  result <- name_backbone(name = species, rank = "species")
  data.frame(
    Best_guess_binomial  = species,
    accepted_name        = ifelse(is.null(result$species),      NA_character_, result$species),
    status               = ifelse(is.null(result$status),       NA_character_, result$status),
    accepted_usage_key   = ifelse(is.null(result$acceptedUsageKey), NA_integer_, result$acceptedUsageKey),
    confidence           = ifelse(is.null(result$confidence),   NA_integer_,   result$confidence),
    stringsAsFactors     = FALSE
  )
})

# Turn into a df
accepted_names_df <- bind_rows(accepted_names)

# Duplicated species df
duplicated_species_df <- combined_duplicate_species %>%
  group_by(Best_guess_binomial) %>%
  filter(n() > 1) %>%
  ungroup()

# Update the names
duplicated_species_df <- duplicated_species_df %>%
  left_join(accepted_names_df %>% 
              select(Best_guess_binomial, accepted_name, status),
            by = "Best_guess_binomial") %>%
  mutate(
    Best_guess_binomial = case_when(
      status == "SYNONYM" & !is.na(accepted_name) ~ accepted_name,
      TRUE ~ Best_guess_binomial
    )
  ) %>%
  select(-accepted_name, -status)  

# Remove Erigone atra that has an NA for Order and Family
duplicated_species_df <- duplicated_species_df[-6,]

# Add the missing family which I searched on GBIF online 
duplicated_species_df[1,2] <- "Linyphiidae"

# Remove the row for Roeseliana roeselii which has a body length of 88mm for one row and 15mm for another row, checking Google, 88mm is too large.
duplicated_species_df <- duplicated_species_df[-4,]
duplicated_species_df[2,3] <- "Brachypera"
duplicated_species_df[2,4] <- "zoilus"
duplicated_species_df[3,3] <- "Brachypera"
duplicated_species_df[3,4] <- "zoilus"

# lets recalculate mean body size for Brachypera zoilus, giving us three rows, one for each duplicate taxa
duplicated_species_df <- duplicated_species_df %>% 
  group_by(Order, Family, Genus, Species, Best_guess_binomial) %>% 
  summarise(Trait_Value = mean(Trait_Value, na.rm = TRUE),
            .groups = "drop")

# Lets remove the duplicate species names from the overall combined duplicate species df
combined_duplicate_species <- combined_duplicate_species %>% filter(!Best_guess_binomial %in% duplicate_names)

# And remerge the correctly updated rows
combined_duplicate_species <- bind_rows(combined_duplicate_species, duplicated_species_df)


# combined_duplicate_species is the final cleaned df for all duplicate species
n_distinct(combined_duplicate_species$Best_guess_binomial) # all rows are distinct taxa

rm(accepted_names, accepted_names_df, duplicated_species_df, duplicate_names, species_list, taxonomy_gbif)
```


# Now that I've cleaned the species only dataframe of all duplicates,  I now want to remerge those duplicate rows back into the overall df for species only

```{r}
# Step 1: Remove the duplicate taxa from the combined_sp_only - I have to anti_join by BGB in order to just remove the rows by their name rather than all their taxonomic info
combined_sp_only <- combined_sp_only %>% anti_join(combined_duplicate_species, by = "Best_guess_binomial")

# Step 2: Bind the corrected rows back in
combined_sp_only_cleaned <- bind_rows(combined_sp_only, combined_duplicate_species)

# Step 3: remove the rows for Metrioptera roeselii and Hypera punctata which are still in the dataset as they have the old synonym names
combined_sp_only_cleaned <- combined_sp_only_cleaned %>%
  group_by(Best_guess_binomial) %>%
  filter(!n() > 1) %>%
  ungroup()

n_distinct(combined_sp_only_cleaned$Best_guess_binomial) # all rows are now distinct taxa, n = 12014

# Spot check a previously duplicated taxon e.g. Acrossus luridus
combined_sp_only_cleaned %>%
  filter(Best_guess_binomial == "Acrossus luridus") %>% 
  View()

# No remaining duplicates
combined_sp_only_cleaned %>%
  group_by(Best_guess_binomial) %>%
  filter(n() > 1) %>%
  arrange(Best_guess_binomial) %>% 
  View()

# sensitivity checks
# Lets look at which rows differ between the final cleaned dataset and the original dataset for species only
which_differs <- combined_sp_only_cleaned %>% anti_join(combined_sp_only) 
# looks like all the duplicate rows
n_distinct(which_differs$Best_guess_binomial) # 191 distinct taxa
n_distinct(combined_duplicate_species$Best_guess_binomial) # 191 distinct duplicate taxa

rm(combined, combined_duplicate_species, combined_sp_only, which_differs)
```


# Now we've cleaned and combined all of the trait datasets, lets merge it into PREDICTS

# Read in PREDICTS data 
```{r}
# All insect data effort-corrected and merged
insects <-  qread("/Users/jusi/Desktop/multi-groups/data/predicts-prepared-whole-database.qs")
```


# Clean and tidy PREDICTS data then merge trait dataset based on taxon resolution
```{r}
# Get relevant taxon cols for all insects
taxon_insects <- insects %>% dplyr::select(SS, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, Merged_diversity, Diversity_metric_unit, Country)

# Get rid of rows which aren't resolved to order level
taxon_insects <- taxon_insects %>% filter(Order != "") # 1056 rows not resolved to order

# Recreate genus and species column as sometimes BGB is present in PREDICTS but no info for genus or sp.
taxon_insects <- taxon_insects %>%
  mutate(Best_guess_binomial = as.character(Best_guess_binomial)) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "merge", fill = "right")

# duplicated rows
combined_sp_only_cleaned[duplicated(combined_sp_only_cleaned["Best_guess_binomial"]) | 
                                  duplicated(combined_sp_only_cleaned["Best_guess_binomial"], fromLast = TRUE), ] %>%  View()

# Get species trait values for merging
sp_value <- combined_sp_only_cleaned %>% dplyr::select(Best_guess_binomial, Trait_Value) 

# First merge all species level data
taxon_insects <- left_join(taxon_insects, sp_value, by = "Best_guess_binomial")

rm(sp_value)

# Merge combined_genus_only with overall_combined df for all species
combined_genus_and_species <- bind_rows(combined_sp_only_cleaned, combined_genus_only)

# Calculate average body length for genera
genus_value <- combined_genus_and_species %>%
  group_by(Genus) %>%
  summarise(genus_avg_body_length = mean(Trait_Value, na.rm = TRUE)) # 2381 genus data before, now 4213 genus values..

# Merge based on genus where best_guess_binomial is empty
taxon_insects <- taxon_insects %>%
  left_join(genus_value, by = "Genus", suffix = c("", "_genus")) 

rm(genus_value)

# Average body length per family based on species
family_value <- combined_genus_and_species %>%
  group_by(Family) %>%
  summarise(family_avg_body_length = mean(Trait_Value, na.rm = TRUE))

# Merge based on family where genus and species is empty
taxon_insects <- taxon_insects %>%
  left_join(family_value, by = "Family", suffix = c("", "_family")) 

rm(family_value)

# priority-based fallback assignment where final body length values are selected based on the first non-missing (or non-blank) column in a defined order.
taxon_insects <- taxon_insects %>%
  mutate(final_body_length = coalesce(
    Trait_Value,
    genus_avg_body_length,
    family_avg_body_length
  ))

rm(combined_sp_only_cleaned, combined_genus_only)
```

# Rainford et al 2014 provides min and max body length data for 760 insect families, I will calculate the average body length between the min and the max and then merge those values onto those rows in my data frame that do not yet have a body length value. 
 
```{r}
# Upload 
rainford <- read_excel(here("data", "rainford_body_size.xlsx"))

# Make first row the header
colnames(rainford) <- as.character(rainford[1, ])
rainford <- rainford[-1, ] # drop first row

# Extract cols of interest
rainford <- rainford %>% dplyr::select("Order", "Family", "Min (mm)", "Max (mm)") %>% rename(min_bl = "Min (mm)") %>% rename(max_bl = "Max (mm)")

# Drop rows with all NAs
rainford <- rainford %>%
  filter(!if_all(everything(), is.na))

# Convert cols to numeric
rainford$min_bl <- as.numeric(rainford$min_bl) # Odonata (Pseudolestidae) has an NA for BL hence the warning message
rainford$max_bl <- as.numeric(rainford$max_bl)

# Drop rows only to Order level
rainford <- rainford %>%
  filter(!is.na(Family))

# Calculate average body length
rainford <- rainford %>%
  mutate(Rainford_family_avg = (min_bl + max_bl) / 2)
         
# Simplify for merging
rainford <- rainford %>% 
  dplyr::select(c(-min_bl, -max_bl, -Order))
```

# Merge Rainford 

```{r}
# Step 1: Filter rows in taxon_insects where there is no body size value at the species, genus or family level and therefore we can merge a family body size value from Rainford
needs_family_avg <- taxon_insects %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length))

# Step 2: Join family_avg body size values from Rainford based on Family column
needs_family_avg <- needs_family_avg %>%
  left_join(rainford, dplyr::select(Family, Rainford_family_avg), by = "Family")

# Step 3: Combine updated rows (need_family_avg) back into original data (taxon_insects)
# First, remove the rows you updated from taxon_insects
remaining_rows <- taxon_insects %>%
  filter(!(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length)))

# sense check remaining rows should be taxon_insects - needs_family_avg
nrow(taxon_insects) - nrow(needs_family_avg)

# Bind them together
taxon_insects_updated <- bind_rows(remaining_rows, needs_family_avg)

# How many rows still without a family average
sum(!is.na(taxon_insects_updated$Rainford_family_avg)) # 134120 rows befpre, 32163 after!

# How many rows have data based on each level of taxonomic resolution
taxon_insects_updated %>%
  summarise(across(
    c(Trait_Value,
    genus_avg_body_length,
    family_avg_body_length,
    final_body_length,
    Rainford_family_avg),
    list(
      non_NAs = ~sum(!is.na(.))
    )
  ))

rm(needs_family_avg, remaining_rows, taxon_insects)
```


# Now that we have merged all the trait data, coalesce the family averages from Rainford into the final_body_length column (for those rows that don't yet have a body length value at the family level) - if they already do then we stick with the earlier value from Logghe and the PREDICTS datasets.

```{r}
# Replace those rows with missing values in final_body_length with the corresponding family_avg
taxon_insects_updated <- taxon_insects_updated %>%
  mutate(final_body_length = coalesce(final_body_length, Rainford_family_avg)) # coalesce(a,b) returns a if its not NA otherwise it returns b

# How many rows do we still not have a body size value for and are still NA - 71437 before/ 59656 after!
no_body_size <- taxon_insects_updated %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length) & is.na(final_body_length) & is.na(Rainford_family_avg)) %>% 
  droplevels()

# But after removing rows with no info for family, genus or species it drops to 8855 rows which can be improved
no_body_size <- no_body_size %>%
  filter(
    (!is.na(Family)  & Family  != "") |
    (!is.na(Genus)   & Genus   != "") |
    (!is.na(Species) & Species != "")
  )

no_body_size %>%
  count(Order, sort = TRUE) # mostly beetles which can still be improved..

# Having a look at it, there's not much which is useful and can be improved.. lots of taxa with numbers in their name, larvae and rows with only family info but no genus or species info. 

# Rename but don't filter out rows which still don't have a body size value as they could be matched with the new data
all_insects_size <- taxon_insects_updated %>%
  #filter(!is.na(final_body_length)) %>% # 1195212
  droplevels()
```

# First split sites based on their latitude to tropical or temperate

```{r}
# Filter overall predicts for sites and their latitude
latitude <- insects %>% dplyr::select(SSBS, Latitude) 

# There are duplicated sites
any(duplicated(latitude$SSBS))

# Lets get a list of distinct sites
latitude <- latitude %>% distinct(SSBS, .keep_all = TRUE)

# Merge latitude values into size data frame for all of the sites
all_insects_size <- all_insects_size %>%
  left_join(latitude, by = "SSBS")

# Add a new column classifying each site into tropical or temperate
all_insects_size$climate_zone <- ifelse(all_insects_size$Latitude >= -23.4368 & all_insects_size$Latitude <= 23.4368, 
                          "tropical", 
                          "temperate")
```

# Read in allometry coefficients
```{r}
coefficients <- read_excel(here("data", "allometry_coefficients.xlsx"))
```

# Merge allometry coefficients to size data
```{r}
# Step 1: Join by order and climate_zone
all_insects_size <- all_insects_size %>%
  left_join(coefficients, by = c("Order", "climate_zone"))

# Step 2: Assign All_insects coefficients for all other taxa without order specific equations
# First extract coefficients for All_insects
fallback <- coefficients %>%
  filter(Order == "All_insects") %>%
  dplyr::select(climate_zone, a_fallback = a, b_fallback = b)

# Step 3: Join coefficients for All_insects based on whether its a tropical or temperate site
all_insects_size <- all_insects_size %>%
  left_join(fallback, by = "climate_zone") %>%
  mutate(
    a = if_else(is.na(a), a_fallback, a), # if a is NA, assign it a for all_insects, else keep it as its order specific a
    b = if_else(is.na(b), b_fallback, b)
  ) %>%
  dplyr::select(-a_fallback, -b_fallback)

# Have a look
all_insects_size %>% dplyr::select(SSBS, Order, final_body_length, Latitude, climate_zone, a, b) %>% unique() %>%  View()
```

# Calculate mass in mg using allometric specific coefficients for each order
```{r}
all_insects_size <- all_insects_size %>%
  mutate(
    log10_mass = a + b * log10(final_body_length),
    mass_mg = 10^log10_mass
  )

summary(all_insects_size$final_body_length) # no zero values for final_body_length
any(all_insects_size$final_body_length == 0) # FALSE

rm(coefficients, fallback, latitude, rainford, insects, no_body_size, taxon_insects_updated)
```

# Predict DBw from wingspan or forewing length for Lepidoptera (DON'T NEED TO RUN LINES 1148-1590 CAN READ IN THE COMBINED LEP TRAIT DATASET LINE 1594)

# First lets work with LepTraits

```{r}
# Check matching for Leptraits
# Read in the data
leptraits <- read.csv(here("data", "consensus.csv")) 

# WS_L = The lowest measurement reported for unspecified sex wingspan in centimeters.
# WS_U = The highest (or sole) measurement reported for unspecified sex wingspan in centimeters.

# Create average wingspan based on the lowest and highest wingspan measurement reported for each taxa
leptraits <- leptraits %>%
  mutate(WS_AVG = ifelse(
    is.na(WS_L) & is.na(WS_U), NA, # if both are an NA = NA
    ifelse(is.na(WS_L), WS_U, # if WS_L is an NA then give me WS_U
           ifelse(is.na(WS_U), WS_L, # if WS_U is an NA then give me WS_L
                  (WS_L + WS_U)/2))) # else give me the average beteween WS_L and WS_U
  )

# select cols of interest
leptraits <- leptraits %>% dplyr::select(Family, Genus, Species, WS_L, WS_U, WS_AVG) 

# filter out rows which don't have wingspan data
leptraits <- leptraits %>% 
  filter(!is.na(WS_AVG)) %>% 
  mutate(Order = "Lepidoptera") %>% 
  rename(Best_guess_binomial = Species) %>% 
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE) %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, WS_L, WS_U, WS_AVG)

# Allometry for calculating dry body weight from wingspan from Garcia Barrios
# log10(dry body weight) = - -3.211 + 2.962 * log10(wingspan)

# Garcia Barrios allometries are predicted using linear measurements in mm so need to convert wingspan to mm first as LepTraits data are in cm
leptraits <- leptraits %>%
  mutate(WS_mm = WS_AVG * 10)

# Calculate log10 dBW using intercept and coefficient from Garcia Barrios and then raise to the power of 10 to get dry body weight in mg
leptraits <- leptraits %>%
  mutate(
    log10_dBW = -3.211 + 2.962 * log10(WS_mm),
    dBW_mg = 10^log10_dBW
  )

# Accordingly, fresh body weight was estimated as a function of BDw only from the male data, as: Log10(fresh body weight) = 0.4829 + 0.9841·Log10(dry body weight), R2 = 0.987, P < 0.001 (n = 183 species). 
leptraits <- leptraits %>%
  mutate(
    log10_fBW = 0.4829 + 0.9841 * log10_dBW,
    fBW_mg = 10^log10_fBW
  )
```

# Check matching for MiddletonWelling

```{r}
# Read in the data
middletonwellingtraits <- read_xlsx(here("data", "MiddletonWelling-2014-07-08.xlsx")) %>% strings2factors()

glimpse(middletonwellingtraits)

# There are 1500 NAs for forewing length rather than 850 NAs for wingspan, lets work with wingspan.. 

# select columns of interest
middletonwellingtraits <- middletonwellingtraits %>% 
  dplyr::select(Family, Genus, Species, `Wingspan male`, `Wingspan female`, `Measurement type`) %>% 
  rename(Wingspan_male = `Wingspan male`, Wingspan_female = `Wingspan female`) # rename columns for easier functionality

# First lets remove NAs
middletonwellingtraits <- middletonwellingtraits %>%
  filter(!is.na(Wingspan_male) | !is.na(Wingspan_female))

# The | (OR) operator means a row is kept if either Wingspan_male or Wingspan_female is not NA — only rows where both are NA are removed.

# Check no rows remain where both columns are NA
middletonwellingtraits %>%
  filter(is.na(Wingspan_male) & is.na(Wingspan_female)) %>%
  nrow()  # should return 0

# First lets create BGB
middletonwellingtraits <- middletonwellingtraits %>%
  mutate(Best_guess_binomial = paste(Genus, Species, sep = " "))


# Lets remove those rows where wingspan is a single number (based on it being a date in the .xsls file) and not a range (which are the correct values)

# First I'll filter out those rows where wingspan values are a single number (a date) for males 
# You can use grepl() to detect hyphens, since ranges will always contain a "-"
# !grepl("-", ...) keeps only values that do not contain a hyphen (i.e. single numbers)
# !is.na(...) excludes NA values
wingspan_male_single <- middletonwellingtraits %>%
  filter(!grepl("-", Wingspan_male) & !is.na(Wingspan_male)) %>%
  select(Best_guess_binomial, Wingspan_male)

# Now for females
wingspan_female_single <- middletonwellingtraits %>%
  filter(!grepl("-", Wingspan_female) & !is.na(Wingspan_female)) %>%
  select(Best_guess_binomial, Wingspan_female)

# # Check for any rows where Genus or Species was empty/NA before combining
middletonwellingtraits %>%
  filter(is.na(Genus) | Genus == "" | is.na(Species) | Species == "") %>%
  select(Genus, Species, Best_guess_binomial)

# Remove data which is based on dates
middletonwellingtraits <- middletonwellingtraits %>%
  anti_join(wingspan_male_single,   by = c("Best_guess_binomial", "Wingspan_male")) %>%
  anti_join(wingspan_female_single, by = c("Best_guess_binomial", "Wingspan_female"))

# First I need to convert wingspan to character first in order to calculate the average wingspan of males and females
# Why do we convert to character first before calculating the averages?
# strsplit() only accepts character input
# Wingspan_male was originally a factor variable, converting a factor directly to numeric gives you the integer codes (i.e. the factor levels: 1,2,3 etc..) rather than the actual values. Converting to character first ensures you get the correct numeric value. 
middletonwellingtraits <- middletonwellingtraits %>%
  mutate(
    Wingspan_male   = as.character(Wingspan_male),
    Wingspan_female = as.character(Wingspan_female)
  )

# Calculate the average wingspan for males and females separately
middletonwellingtraits <- middletonwellingtraits %>%
  mutate(
    Wingspan_male_avg = sapply(strsplit(Wingspan_male, "-"), # splits the range into two numbers
                               function(x) mean(as.numeric(x), na.rm = TRUE)), # make sure both numbers are numeric and calculate the mean
    Wingspan_female_avg = sapply(strsplit(Wingspan_female, "-"), 
                                 function(x) mean(as.numeric(x), na.rm = TRUE))
  )

# Calculate the combined average wingspan between males and females
middletonwellingtraits <- middletonwellingtraits %>%
  mutate(
    Wingspan_CombinedSex_avg = rowMeans(cbind(Wingspan_male_avg, Wingspan_female_avg), 
                                        na.rm = TRUE)
  )
 
# Often the average male wingspan and the average female wingspan is the same

# Lets add the taxonomic order as a column using the gbif backbone
# Look up each species in the GBIF backbone
gbif_lookup <- lapply(middletonwellingtraits$Best_guess_binomial, function(species) {
  result <- name_backbone(name = species, rank = "species")
  data.frame(
    Best_guess_binomial = species,
    order               = ifelse(is.null(result$order), NA_character_, result$order),
    stringsAsFactors    = FALSE
  )
})

# Combine results into a single data frame
gbif_lookup_df <- bind_rows(gbif_lookup)

# Join the order back onto middletonwellingtraits
middletonwellingtraits <- middletonwellingtraits %>%
  left_join(gbif_lookup_df, by = "Best_guess_binomial")

# Ok about 20 rows have an NA but that's fine, I know that the families are lepidoptera and it won't matter much for matching.

# Allometry for calculating dry body weight from wingspan from Garcia Barrios
# log10(dry body weight) = - -3.211 + 2.962 * log10(wingspan)

# Calculate log10 dBW using intercept and coefficient from Garcia Barrios and then raise to the power of 10 to get dry body weight in mg
middletonwellingtraits <- middletonwellingtraits %>%
  mutate(
    log10_dBW = -3.211 + 2.962 * log10(Wingspan_CombinedSex_avg),
    dBW_mg = 10^log10_dBW
  )

# Accordingly, fresh body weight was estimated as a function of BDw only from the male data, as: Log10(fresh body weight) = 0.4829 + 0.9841·Log10(dry body weight), R2 = 0.987, P < 0.001 (n = 183 species). 
middletonwellingtraits <- middletonwellingtraits %>%
  mutate(
    log10_fBW = 0.4829 + 0.9841 * log10_dBW,
    fBW_mg = 10^log10_fBW
  )

rm(gbif_lookup_df, wingspan_female_single, wingspan_male_single, gbif_lookup)
```

# Lets extract wingspan values from GRIT

```{r}
summary(grit)
# wingspan n = 2183
# forewing length n = 1784

# Having a look at the DB it seems there are some sp which have a wingspan value, a forewing length value or both
# I will separate grit into a wingspan data set and a forewing length data set and then calculate the allometries separately for both.

grit_wingspan <- grit %>% filter(measurementType == "Wingspan")
grit_forewing <- grit %>% filter(measurementType == "Forewing Length")

# Lets deal with wingspan first
grit_wingspan <- grit_wingspan %>% dplyr::select(Order, Family, Genus, Species, verbatimIdentification, measurementValue, measurementUnit, measurementStatistic)

# For wingspan, there are three potential measurement types
# max, mean and min wingspan for each species
table(grit_wingspan$measurementStatistic)

# Have a look
grit_wingspan %>% dplyr::select(verbatimIdentification, measurementValue, measurementStatistic) %>% View()

# I want to calculate a single wingspan value for each species
# Each species can have multiple rows of either a mean, min or max wingspan value.
# The temporary plan is as follows:
# If the species has a mean value, it gets the mean of the means.
# If the species has no mean value but has a max or min value, it gets the mean of the max and/or min

one_wingspan_per_species <- grit_wingspan %>%
  group_by(verbatimIdentification) %>%
  summarise(
    measurementValue = case_when(
      # Priority 1: mean of all "mean" wingspan values if any exist
      any(measurementStatistic == "mean", na.rm = TRUE) ~
        mean(measurementValue[measurementStatistic == "mean"], na.rm = TRUE),

      # Priority 2: mean of "max" and/or "min" wingspan values if no "mean" exists
      any(measurementStatistic %in% c("max", "min"), na.rm = TRUE) ~
        mean(measurementValue[measurementStatistic %in% c("max", "min")], na.rm = TRUE),

      # If nothing matches, return NA
      TRUE ~ NA_real_
    ),
    measurementStatistic = case_when(
      any(measurementStatistic == "mean",            na.rm = TRUE) ~ "mean",
      any(measurementStatistic %in% c("max", "min"), na.rm = TRUE) ~ "mean_of_max_min",
      TRUE ~ NA_character_
    ),
    .groups = "drop"
  )

n_distinct(grit_wingspan$verbatimIdentification) # 1278 looks like it worked!


# Get one row per species (e.g. family, order) from the overall df for merging
one_row_per_species <- grit_wingspan %>%
  distinct(Order, Family, Genus, Species, verbatimIdentification) 

# Merge
grit_wingspan_cleaned <- one_row_per_species %>%
  left_join(one_wingspan_per_species, by = "verbatimIdentification")

# Clean up environment
rm(grit_wingspan, one_row_per_species, one_wingspan_per_species)

# Allometry for calculating dry body weight from wingspan from Garcia Barrios
# log10(dry body weight) = - -3.211 + 2.962 * log10(wingspan)

# Calculate log10 dBW using intercept and coefficient from Garcia Barrios and then raise to the power of 10 to get dry body weight in mg
grit_wingspan_cleaned <- grit_wingspan_cleaned %>%
  mutate(
    log10_dBW = -3.211 + 2.962 * log10(measurementValue),
    dBW_mg = 10^log10_dBW
  )

# Accordingly, fresh body weight was estimated as a function of BDw only from the male data, as: Log10(fresh body weight) = 0.4829 + 0.9841·Log10(dry body weight), R2 = 0.987, P < 0.001 (n = 183 species). 
grit_wingspan_cleaned <- grit_wingspan_cleaned %>%
  mutate(
    log10_fBW = 0.4829 + 0.9841 * log10_dBW,
    fBW_mg = 10^log10_fBW
  )
```

# Ok now lets deal with forewing length data in GRIT

```{r}
# For forewing length, there are two potential measurement types: max and min
table(grit_forewing$measurementStatistic)

# Have a look
grit_forewing %>% dplyr::select(verbatimIdentification, measurementValue, measurementStatistic) %>% View()

# I want to calculate a single forewing length value for each species
# It looks like each species has a min and a max value, there are 892 for each * 2 = 1784
# The temporary plan is as follows:
# each species gets the mean of the min and max value

one_forewing_per_species <- grit_forewing %>%
  group_by(verbatimIdentification) %>%
  summarise(
    measurementValue = case_when(
      # Priority 1: mean of "max" and/or "min" forewing length
      any(measurementStatistic %in% c("max", "min"), na.rm = TRUE) ~
        mean(measurementValue[measurementStatistic %in% c("max", "min")], na.rm = TRUE),

      # If nothing matches, return NA
      TRUE ~ NA_real_
    ),
    measurementStatistic = case_when(
      any(measurementStatistic %in% c("max", "min"), na.rm = TRUE) ~ "mean_of_max_min",
      TRUE ~ NA_character_
    ),
    .groups = "drop"
  )

n_distinct(grit_forewing$verbatimIdentification) # 892 looks like it worked!

# Get one row per species (e.g. family, order) from the overall df for merging
one_row_per_species <- grit_forewing %>%
  distinct(Order, Family, Genus, Species, verbatimIdentification) 

# Merge
grit_forewing_cleaned <- one_row_per_species %>%
  left_join(one_forewing_per_species, by = "verbatimIdentification")

# Clean up environment
rm(grit_forewing, one_row_per_species, one_forewing_per_species)

# Allometry for calculating dry body weight from forewing length from Garcia Barrios
# log10(dry body weight) = - 2.053 + 2.727 * log10(forewing length)

# Calculate log10 dBW using intercept and coefficient from Garcia Barrios and then raise to the power of 10 to get dry body weight in mg
grit_forewing_cleaned <- grit_forewing_cleaned %>%
  mutate(
    log10_dBW = - 2.053 + 2.727 * log10(measurementValue),
    dBW_mg = 10^log10_dBW
  )

# Accordingly, fresh body weight was estimated as a function of BDw only from the male data, as: Log10(fresh body weight) = 0.4829 + 0.9841·Log10(dry body weight), R2 = 0.987, P < 0.001 (n = 183 species). 
grit_forewing_cleaned <- grit_forewing_cleaned %>%
  mutate(
    log10_fBW = 0.4829 + 0.9841 * log10_dBW,
    fBW_mg = 10^log10_fBW
  )
```

# Lets deal with the duplicates between Lepidoptera trait datasets

```{r}
# Collapse duplicates within each dataset individually
leptraits_collapsed <- leptraits %>%
  group_by(Best_guess_binomial, Genus) %>%
  summarise(fBW_mg = mean(fBW_mg, na.rm = TRUE), .groups = "drop") %>%
  mutate(source = "leptraits")

middletonwellingtraits_collapsed <- middletonwellingtraits %>%
  group_by(Best_guess_binomial, Genus) %>%
  summarise(fBW_mg = mean(fBW_mg, na.rm = TRUE), .groups = "drop") %>%
  mutate(source = "middletonwelling")

grit_forewing_collapsed <- grit_forewing_cleaned %>%
  rename(Best_guess_binomial = verbatimIdentification) %>% 
  group_by(Best_guess_binomial, Genus) %>%
  summarise(fBW_mg = mean(fBW_mg, na.rm = TRUE), .groups = "drop") %>%
  mutate(source = "grit_forewing")

grit_wingspan_collapsed <- grit_wingspan_cleaned %>%
  rename(Best_guess_binomial = verbatimIdentification) %>% 
  group_by(Best_guess_binomial, Genus) %>%
  summarise(fBW_mg = mean(fBW_mg, na.rm = TRUE), .groups = "drop") %>%
  mutate(source = "grit_wingspan")

# Combine all four collapsed datasets
lep_combined <- bind_rows(
  leptraits_collapsed,
  middletonwellingtraits_collapsed,
  grit_forewing_collapsed,
  grit_wingspan_collapsed
)

# Collapse any cross-dataset duplicates by averaging fBW_mg across sources
lep_combined_clean <- lep_combined %>%
  group_by(Best_guess_binomial) %>%
  summarise(
    fBW_mg          = mean(fBW_mg, na.rm = TRUE),
    Genus = first(Genus), # retains genus from first available source
    n_sources        = n(),                          # how many datasets contributed
    sources          = paste(source, collapse = ", ") # which datasets contributed
  ) %>%
  ungroup()

# Check row counts at each stage
cat("Rows in leptraits after collapse:            ", nrow(leptraits_collapsed), "\n") # 6643
cat("Rows in middletonwelling after collapse:     ", nrow(middletonwellingtraits_collapsed), "\n") # 965
cat("Rows in grit_forewing after collapse:        ", nrow(grit_forewing_collapsed), "\n") # 892
cat("Rows in grit_wingspan after collapse:        ", nrow(grit_wingspan_collapsed), "\n") # 1278
cat("Total rows combined:                         ", nrow(lep_combined), "\n") # 9778
cat("Rows after cross-dataset duplicate collapse:           ", nrow(lep_combined_clean), "\n") # 8874

# Check how many species appeared in more than one dataset = 781
lep_combined_clean %>%
  filter(n_sources > 1) %>%
  arrange(desc(n_sources)) %>% 
  View()

# Check no duplicates remain
cat("Duplicates remaining:", 
    sum(duplicated(lep_combined_clean$Best_guess_binomial)), "\n")  # should be 0

# Check fBW_mg values look sensible
summary(lep_combined_clean$fBW_mg)

leptraits %>% filter(Best_guess_binomial == "Baliochila aslanga") %>% View()


# save combined trait dataset for leps
# write_csv(lep_combined_clean, "/Users/jusi/Desktop/multi-groups/data/lep_combined_cleaned.csv")


rm(leptraits, middletonwellingtraits, grit_forewing_cleaned, grit_wingspan_cleaned)
```

# First lets filter all_insects_size for those Lepidoptera rows that can be improved

```{r}
# filter all_insects_size for leps with no body size value
no_lep_body_size <- all_insects_size %>% filter(Order == "Lepidoptera") %>% 
# filter out rows with empty values for family, genus and species
# The | (OR) operator means a row is kept if at least one of Family, Genus or Species is not NA or an empty string — only rows where all three are NA/empty string are removed.
  dplyr::filter(
    !is.na(Family)              & Family              != "",
    !is.na(Genus)               & Genus               != "",
    !is.na(Species)             & Species             != "",
    !is.na(Best_guess_binomial) & Best_guess_binomial != ""
  ) %>% 
# now filter rows with no trait data yet
  dplyr::filter(
    is.na(Trait_Value),
    is.na(genus_avg_body_length),
    is.na(family_avg_body_length),
    is.na(Rainford_family_avg),
    is.na(final_body_length)
  )

# 208 rows with currently no body size value

# 2 species, nothing to worry about
n_distinct(no_lep_body_size$Best_guess_binomial)

# Lets filter for those rows which have ONLY a family-level body length value, but are taxonomically
# resolve to genus or species and therefore can be improved!
can_be_improved <- all_insects_size %>%
  filter(Order == "Lepidoptera") %>% 
  filter(
    (!is.na(family_avg_body_length) | !is.na(Rainford_family_avg)),  # contains a family value from either column
    is.na(Trait_Value),                   # no species level value
    is.na(genus_avg_body_length),         # no genus level value
    (!is.na(Genus) & Genus != "") | (!is.na(Best_guess_binomial) & Best_guess_binomial != "") # genus or species present
  )

rm(no_lep_body_size)
```

# Start merging the combined lep trait dataset to those rows that can_be_improved

```{r}
# Re-read in trait dataset 
lep_combined_clean <- read.csv(here("data", "lep_combined_cleaned.csv")) %>% strings2factors()

# Join species-level fBW_mg onto can_be_improved by Best_guess_binomial
can_be_improved <- can_be_improved %>%
  left_join(lep_combined_clean %>% 
              select(Best_guess_binomial, fBW_mg),
            by = "Best_guess_binomial")

# For rows that didn't match at species level, calculate genus-level averages from lep_combined_clean and match on Genus:
# # Calculate genus-level average FBW_mg from lep_combined_clean
lep_combined_clean_genus_avg <- lep_combined_clean %>%
  group_by(Genus) %>%
  summarise(FBW_mg_genus_avg = mean(fBW_mg, na.rm = TRUE), .groups = "drop")

# Join genus-level averages for rows that didn't get a species-level match
can_be_improved <- can_be_improved %>%
  left_join(lep_combined_clean_genus_avg, by = "Genus")

# Update mass_mg with the best available match, prioritising species over genus level:
can_be_improved <- can_be_improved %>%
  mutate(
    # Record what level the mass_mg value was matched at
    mass_mg_match_level = case_when(
      !is.na(fBW_mg)           ~ "species",
      !is.na(FBW_mg_genus_avg) ~ "genus",
      TRUE                     ~ "family"   # no match found, retains original family-level value
    ),
    # Update mass_mg with the best available match
    mass_mg = case_when(
      !is.na(fBW_mg)           ~ fBW_mg,
      !is.na(FBW_mg_genus_avg) ~ FBW_mg_genus_avg,
      TRUE                     ~ mass_mg   # retain original family-level mass_mg
    )
  ) # %>%
  # select(-fBW_mg, -FBW_mg_genus_avg)  # remove the lookup columns


# Check the breakdown of match levels
can_be_improved %>%
  count(mass_mg_match_level, sort = TRUE) %>%
  mutate(percentage = round(n / sum(n) * 100, 1)) %>% 
  View()

# Check no unexpected NAs were introduced in mass_mg
cat("NAs in mass_mg:", sum(is.na(can_be_improved$mass_mg)), "\n")
```


# Lastly, lets remerge the updated rows back into all_insects_size

```{r}
# Since can_be_improved is a subset of all_insects_size, you can use a shared unique identifier to remove the original rows and bind the updated ones back in:

# Step 1 - Remove the original family-level rows from all_insects_size
all_insects_size_updated <- all_insects_size %>%
  anti_join(can_be_improved, by = "Best_guess_binomial")

# Step 2 - Bind the updated rows back in
all_insects_size_updated <- bind_rows(all_insects_size_updated, can_be_improved)

# sense check the remerge:
# Check row counts add up correctly
cat("Original all_insects_size rows:", nrow(all_insects_size), "\n")
cat("Rows removed (can_be_improved):", nrow(can_be_improved), "\n")
cat("Rows after removal:            ", nrow(all_insects_size_updated) - nrow(can_be_improved), "\n")
cat("Final rows after remerge:      ", nrow(all_insects_size_updated), "\n")
# Final rows should equal original all_insects_size rows

# Lets fill in the mass_mg_match_level column for the rows which were not in can_be_improved
# You can use case_when() with !is.na(mass_mg_match_level) as the first condition to preserve the existing values from can_be_improved before assigning new ones:

all_insects_size_updated <- all_insects_size_updated %>%
  mutate(
    mass_mg_match_level = case_when(
      # Preserve existing values from can_be_improved
      !is.na(mass_mg_match_level)                                        ~ mass_mg_match_level,
      # Species-level: has a value in Trait_Value
      !is.na(Trait_Value)                                                ~ "species",
      # Genus-level: no Trait_Value but has genus_avg_body_length
      is.na(Trait_Value) & !is.na(genus_avg_body_length)                ~ "genus",
      # Family-level: no Trait_Value or genus, but has family_avg_body_length or Rainford_family_avg
      is.na(Trait_Value) & is.na(genus_avg_body_length) & 
        (!is.na(family_avg_body_length) | !is.na(Rainford_family_avg))  ~ "family",
      # No match found at any level
      TRUE                                                               ~ NA_character_
    )
  )

# Check the breakdown of match levels across the full dataset
all_insects_size_updated %>%
  count(mass_mg_match_level, sort = TRUE) %>%
  mutate(percentage = round(n / sum(n) * 100, 1)) %>% 
  View()

# Check how many rows still have no match level assigned
cat("Rows with no match level:", sum(is.na(all_insects_size_updated$mass_mg_match_level)), "\n")

# If there are rows with no match level, inspect them to understand why
all_insects_size_updated %>%
  filter(is.na(mass_mg_match_level)) %>%
  View()

rm(all_insects_size)
```

# Summary stats for methods

```{r}
# Read in old data frame
all_insects_size_old <- read.csv(here("data", "all_insects_size_old.csv")) %>% strings2factors()

# Create mass_mg_match_level for the old data frame
all_insects_size_old <- all_insects_size_old %>%
  mutate(
    mass_mg_match_level = case_when(
      !is.na(Trait_Value)                                                ~ "species",
      is.na(Trait_Value) & !is.na(genus_avg_body_length)                ~ "genus",
      is.na(Trait_Value) & is.na(genus_avg_body_length) &
        (!is.na(family_avg_body_length) | !is.na(Rainford_family_avg))  ~ "family",
      TRUE                                                               ~ NA_character_
    )
  )

# Summarise match levels for old dataset
summary_old <- all_insects_size_old %>%
  filter(!is.na(mass_mg_match_level)) %>%
  group_by(Order) %>%
  summarise(
    n_species_old  = n_distinct(Best_guess_binomial[mass_mg_match_level == "species"]),
    n_genera_old   = n_distinct(Genus[mass_mg_match_level == "genus"]),
    n_families_old = n_distinct(Family[mass_mg_match_level == "family"]),
    .groups        = "drop"
  )

# Summarise match levels for new dataset
summary_new <- all_insects_size_updated %>%
  filter(!is.na(mass_mg_match_level)) %>%
  group_by(Order) %>%
  summarise(
    n_species_new  = n_distinct(Best_guess_binomial[mass_mg_match_level == "species"]),
    n_genera_new   = n_distinct(Genus[mass_mg_match_level == "genus"]),
    n_families_new = n_distinct(Family[mass_mg_match_level == "family"]),
    .groups        = "drop"
  )

# Join old and new summaries and calculate the improvement
comparison <- summary_old %>%
  left_join(summary_new, by = "Order") %>%
  mutate(
    species_gain = n_species_new  - n_species_old,
    genera_gain  = n_genera_new   - n_genera_old,
    family_gain  = n_families_new - n_families_old
  ) %>%
  arrange(desc(species_gain))

print(comparison)

# Table X highlighting how many species are matched to species-level data, how many genera are matched to genus-level data and how many families are matched to family level data.
 order_summary_old <- all_insects_size_old %>%
  # Keep only rows with any body length data available
  filter(
    !is.na(Trait_Value)            |
    !is.na(genus_avg_body_length)  |
    !is.na(family_avg_body_length) |
    !is.na(Rainford_family_avg)
  ) %>%
  group_by(Order) %>%
  summarise(
    n_species  = n_distinct(Best_guess_binomial[!is.na(Trait_Value)]),
    n_genera   = n_distinct(Genus[!is.na(genus_avg_body_length)]),
    n_families = n_distinct(Family[!is.na(family_avg_body_length) | !is.na(Rainford_family_avg)]),
    .groups = "drop"
  ) %>%
  arrange(desc(n_species))

# Do the same for the old data set
order_summary_new <- all_insects_size_updated %>% # feed old df
  filter(!is.na(mass_mg_match_level)) %>%
  group_by(Order) %>%
  summarise(
    n_species  = n_distinct(Best_guess_binomial[mass_mg_match_level == "species"]),
    n_genera   = n_distinct(Genus[mass_mg_match_level == "genus"]),
    n_families = n_distinct(Family[mass_mg_match_level == "family"]),
    .groups    = "drop"
  ) %>%
  arrange(desc(n_species))


# I noticed a MASSIVE discrepancy between the number of genera matched for Coleoptera between the old (n = 320) and new (n = 836) data frame, between using the actual columns such as !is.na(Trait_Value) to calculate the matching in the old 
# method versus using filter(!is.na(mass_mg_match_level)) in the new method. 
# 836 - 320 = 516 genera that have a genus_avg_body_length value but were ACTUALLY matched at the species level.
all_insects_size_updated %>%
  filter(
    Order == "Coleoptera",
    !is.na(genus_avg_body_length),
    mass_mg_match_level == "species"
  ) %>%
  summarise(n_genera = n_distinct(Genus)) %>% 
  View()

# The discrepancy is because the simple subtraction 836 - 320 = 516 assumed the two groups were mutually exclusive, but they are not. A genus like Carabus can contain some species matched at species-level AND other species within the same genus matched at genus-level. This means 262 genera appear in both the 778 and 320 counts: 778 + 320 - 836 = 262 genera appearing in both groups.

# Filter for Coleoptera rows which have a genus value
coleoptera <- all_insects_size_updated %>%
  filter(Order == "Coleoptera", !is.na(genus_avg_body_length))

# Genera with ONLY species-level matches
only_species <- coleoptera %>%
  group_by(Genus) %>%
  filter(all(mass_mg_match_level == "species")) %>%
  ungroup() %>%
  summarise(n_genera = n_distinct(Genus))

# Genera with ONLY genus-level matches
only_genus <- coleoptera %>%
  group_by(Genus) %>%
  filter(all(mass_mg_match_level == "genus")) %>%
  ungroup() %>%
  summarise(n_genera = n_distinct(Genus))

# Genera with BOTH species and genus level matches
# i.e. different species within same genus matched at different levels
both <- coleoptera %>%
  group_by(Genus) %>%
  filter(
    any(mass_mg_match_level == "species") & 
    any(mass_mg_match_level == "genus")
  ) %>%
  ungroup() %>%
  summarise(n_genera = n_distinct(Genus))

cat("Genera with ONLY species-level matches: ", only_species$n_genera, "\n")
cat("Genera with ONLY genus-level matches:   ", only_genus$n_genera,   "\n")
cat("Genera with BOTH levels:                ", both$n_genera,         "\n")
cat("Total (should equal 836):               ", 
    only_species$n_genera + only_genus$n_genera + both$n_genera, "\n")

# (!is.na(genus_avg_body_length)) counts every genus that has a genus average available, regardless of whether it was ultimately used as the final match. 
#  When genus_avg_body_length was calculated and joined earlier, it was joined to all rows in that genus, including rows that also have a species-level Trait_Value. These rows do not have an NA for genus_avg_body_length but were actually matched at the species level. Method 2 (mass_mg_match_level == "genus") only counts genera where the genus average was the best available match — i.e. no species-level data existed for those rows.
# 	                                                              n_genera
# Genera with genus_avg_body_length available (Method 1):	        836
# Genera actually matched at genus level (Method 2):              320
# Difference: genera where species-level data took precedence:    516

# Method 2 is the correct one for your table, as you are asking how many genera were matched to genus-level data, not how many genera have a genus-level value available.


# Compare species-level counts before and after fix for Lepidoptera
cat("Using Trait_Value (old - incorrect for Lepidoptera):\n")
all_insects_size_old %>%
  filter(Order == "Lepidoptera") %>%
  summarise(n_species = n_distinct(Best_guess_binomial[!is.na(Trait_Value)])) %>%
  print()

cat("Using mass_mg_match_level (new - correct):\n")
all_insects_size_updated %>%
  filter(Order == "Lepidoptera") %>%
  summarise(n_species = n_distinct(Best_guess_binomial[mass_mg_match_level == "species"])) %>%
  print()
# The new count should be higher as it includes species matched via FBW_mg

# Calculate how many species are matched to species-level data
match_level_count <- all_insects_size_updated %>%
  filter(mass_mg_match_level == "species") %>%
  summarise(n = n_distinct(Best_guess_binomial)) %>%
  pull(n)

trait_value_count <- all_insects_size_updated %>%
  filter(!is.na(Trait_Value)) %>%
  summarise(n = n_distinct(Best_guess_binomial)) %>%
  pull(n)

cat("Via mass_mg_match_level:", match_level_count, "\n")
cat("Via Trait_Value:        ", trait_value_count, "\n")
cat("Match:", match_level_count == trait_value_count, "\n")

# The mismatch will be because rows from can_be_improved have mass_mg_match_level == "species" (matched via FBW_mg from the lep trait datasets) but 
# Trait_Value is still NA as the species-level match didn't come from the original Trait_Value column

# Group 1: mass_mg_match_level == "species" but Trait_Value is NA
# These are the rows from can_be_improved rows matched via FBW_mg
matched_via_fBW_mg <- all_insects_size_updated %>%
  filter(mass_mg_match_level == "species" & is.na(Trait_Value))

cat("Rows matched as species via FBW_mg (not Trait_Value):", nrow(matched_via_fBW_mg), "\n")
cat("Unique species:                                       ", n_distinct(matched_via_fBW_mg$Best_guess_binomial), "\n")

# Confirm these are all Lepidoptera from can_be_improved
matched_via_fBW_mg %>%
  count(Order, sort = TRUE)

# Group 2: Trait_Value is not NA but mass_mg_match_level != "species"
# These would be unexpected and worth investigating
unexpected <- all_insects_size_updated %>%
  filter(!is.na(Trait_Value) & mass_mg_match_level != "species")

cat("\nRows with Trait_Value but not a species-level match:", nrow(unexpected), "\n")

# Check what match level they have been assigned
unexpected %>%
  count(mass_mg_match_level, Order, sort = TRUE)

# Correct species-level count combining both sources
all_insects_size_updated %>%
  filter(
    mass_mg_match_level == "species"  # captures both Trait_Value and FBW_mg matches
  ) %>%
  summarise(
    n_species_total       = n_distinct(Best_guess_binomial),
    n_via_trait_value     = n_distinct(Best_guess_binomial[!is.na(Trait_Value)]),
    n_via_fBW_mg          = n_distinct(Best_guess_binomial[is.na(Trait_Value)])
  ) %>% 
  View()

# I now want to calculate how many species are matched to genus-level data and family-level data.
all_insects_size_updated %>%
  filter(mass_mg_match_level %in% c("genus", "family")) %>%
  group_by(mass_mg_match_level) %>%
  summarise(n_species = n_distinct(Best_guess_binomial), .groups = "drop") %>% 
  View()

all_insects_size_old %>%
  filter(mass_mg_match_level %in% c("genus", "family")) %>%
  group_by(mass_mg_match_level) %>%
  summarise(n_species = n_distinct(Best_guess_binomial), .groups = "drop") %>% 
  View()

# If you also want to include species-level for a full picture of the dataset:
all_insects_size_updated %>%
  group_by(mass_mg_match_level) %>%
  summarise(n_species = n_distinct(Best_guess_binomial), .groups = "drop") %>%
  mutate(percentage = round(n_species / sum(n_species) * 100, 1)) %>%
  arrange(desc(n_species)) %>%
  View()

# Total species and rows by match level with cumulative coverage
all_insects_size_updated %>%
  group_by(mass_mg_match_level) %>%
  summarise(
    n_rows    = n(),
    n_species = n_distinct(Best_guess_binomial),
    .groups   = "drop"
  ) %>%
  mutate(
    pct_rows    = round(n_rows    / sum(n_rows)    * 100, 1),
    pct_species = round(n_species / sum(n_species) * 100, 1)
  ) %>%
  arrange(desc(n_species)) %>% 
  View()

# For each order: total species, species at each match level and % coverage
all_insects_size_updated %>%
  group_by(Order) %>%
  mutate(total_species_in_order = n_distinct(Best_guess_binomial)) %>%
  group_by(Order, mass_mg_match_level, total_species_in_order) %>%
  summarise(n_species = n_distinct(Best_guess_binomial), .groups = "drop") %>%
  mutate(pct = round(n_species / total_species_in_order * 100, 1)) %>%
  pivot_wider(
    names_from   = mass_mg_match_level,
    values_from  = c(n_species, pct),
    values_fill  = 0
  ) %>%
  arrange(desc(total_species_in_order)) %>% 
  View()

# How much does mass_mg vary within each match level?
# High variation at family level suggests family-level matching is imprecise
all_insects_size_updated %>%
  group_by(mass_mg_match_level) %>%
  summarise(
    mean_mass_mg   = round(mean(mass_mg,   na.rm = TRUE), 3),
    median_mass_mg = round(median(mass_mg, na.rm = TRUE), 3),
    sd_mass_mg     = round(sd(mass_mg,     na.rm = TRUE), 3),
    min_mass_mg    = round(min(mass_mg,    na.rm = TRUE), 3),
    max_mass_mg    = round(max(mass_mg,    na.rm = TRUE), 3),
    cv             = round(sd(mass_mg, na.rm = TRUE) / 
                           mean(mass_mg, na.rm = TRUE) * 100, 1),  # coefficient of variation
    .groups        = "drop"
  ) %>% 
  View()
# Higher CV at family level = more uncertainty in mass_mg values at that level


rm(unexpected, summary_new, summary_old, order_summary_new, order_summary_old, only_species, only_genus, 
   leptraits_collapsed, middletonwellingtraits_collapsed, grit_forewing_collapsed, grit_wingspan_collapsed,
   lep_combined, grit, comparison, matched_via_fBW_mg, lep_combined_clean_genus_avg, both, can_be_improved, all_insects_size_old, coleoptera, match_level_count, trait_value_count)
```





# Analysis looking at ALL of the insects 


# CREATE BIOMASS MODEL DATA

# Load in model plot function
```{r}
## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(2,2))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}
```

# first lets calculate one weighted mean length value per site

```{r}
# Create a separate data frame for the overall site level analysis
overall_site_size <- all_insects_size_updated

# Calculate the biomass of each taxa
overall_site_size <- overall_site_size %>% 
  mutate(
    Biomass = mass_mg * Merged_diversity
  )

# there are 34 studies where all sites have zero total biomass, so min(TotalSiteBiomass[TotalSiteBiomass > 0]) has no values to take the minimum of and returns Inf
# Inspecting these studies reveals that none of the rows/taxa/sites have a body length value.
# I will remove rows with no body length value
overall_site_size <- overall_site_size %>%
  filter(!is.na(mass_mg_match_level))

# Calculate TotalSiteBiomass, SmallestNonZeroTotalSiteBiomass for a site within a study and Log_plus_what
overall_site_size <- overall_site_size %>%
  group_by(SSBS) %>%
  mutate(TotalSiteBiomass = sum(Biomass, na.rm = TRUE)) %>%  # Calculate total site biomass
  ungroup() %>%
  group_by(SS) %>% # Now group by study
  mutate(
    SmallestNonZeroTotalSiteBiomass = min(TotalSiteBiomass[TotalSiteBiomass > 0]), # Which site has the smallest total biomass greater than 0 per study? 
    Log_plus_what = ifelse(
    Diversity_metric_unit %in% c("individuals", "times observed"), min(mass_mg, na.rm = TRUE),  
    # If abundance data are counts, assign the smallest mass_mg for an individual of any species in the study
    SmallestNonZeroTotalSiteBiomass / 2)) %>%  # else 1/2 the SmallestNonZeroTotalSiteBiomass for any site within the study 
  ungroup() %>% 
  group_by(SSBS) %>% 
  mutate(LogBiomass = log(TotalSiteBiomass + Log_plus_what)) %>% 
  ungroup()

# Confirm no studies with all zero biomass remain
overall_site_size %>%
  group_by(SS) %>%
  summarise(all_zero = all(TotalSiteBiomass == 0), .groups = "drop") %>%
  filter(all_zero) %>%
  nrow()
# Should return 0

# Calculate one LogBiomass value per site
LogBiomass_summary <- overall_site_size %>%
  group_by(SSBS) %>%
  summarise(
    n_TSB = n_distinct(TotalSiteBiomass), 
    n_LPW = n_distinct(Log_plus_what),
    is_consistent = (n_TSB == 1 & n_LPW == 1), # sense check that all TotalSiteBiomass and Log_plus_what values are the same for all rows in the site so that we can pull the first value
    TotalSiteBiomass = first(TotalSiteBiomass),
    Log_plus_what = first(Log_plus_what),
    LogBiomass = log(first(TotalSiteBiomass) + first(Log_plus_what)),
    .groups = "drop"
  )

# All sites are consistent
LogBiomass_summary %>% filter(is_consistent != TRUE)

# Any inconsistent sites?
inconsistent_groups <- overall_site_size %>%
  group_by(SSBS) %>%
  summarise(
    n_TSB = n_distinct(TotalSiteBiomass),
    n_LPW = n_distinct(Log_plus_what)
  ) %>%
  filter(n_TSB > 1 | n_LPW > 1) 

# No rows where LogBiomass is 0, NA, or blank
LogBiomass_summary %>%
  filter(is.na(LogBiomass) | LogBiomass == 0 | LogBiomass == "") %>% View()

rm(inconsistent_groups)
```

# Merge with PREDICTS insects
```{r}
# All insect data effort-corrected and merged
insects <-  qread("/Users/jusi/Desktop/multi-groups/data/predicts-prepared-whole-database.qs")

# Use distinct() to keep only the first appearance of each site within the overall PREDICTS div extract for insects
insects_single <- insects %>%
  distinct(SSBS, .keep_all = TRUE)

# merge LogBiomass_summary values
final_size <- insects_single %>%
  left_join(LogBiomass_summary, by = "SSBS") %>% 
  droplevels()

# Have a look at sites with no LogBiomass
no_biomass <- final_size %>%
  filter(
    is.na(LogBiomass) |         # remove NA
    is.nan(LogBiomass)|        # remove NaN
    LogBiomass == ""            # remove empty strings
  )  # I think this is because these rows do not have a body size value, therefore no LogBiomass value. 

# filter out sites with no LogBiomass 
final_size <- final_size %>%
  filter(
    !is.na(LogBiomass),         # remove NA
    !is.nan(LogBiomass),        # remove NaN
    LogBiomass != ""            # remove empty strings
  )

# 14009 sites in total

rm(insects_single, no_biomass, LogBiomass_summary)
```

# Start the modelling of all of the insects

```{r}
# This PREDICTS data file is diversity extract of all insect data which has been merged and effort corrected.
# It has also had a set of sampling methods updated, see PreparePREDICTSData.R
# It has not had accidentally sampled larvae removed as this analysis does not focus on multiple groups.

# Rename predominant habitat since its not really habitat we're looking at
final_size <- rename(final_size,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
final_size <- final_size %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(final_size$Predominant_land_use, final_size$Use_intensity)
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
final_size <- final_size %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# Have a look at sample sizes
table(final_size$LUI)

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 73 sites, merge with light use
final_size <- final_size %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Have a look at sample sizes
table(final_size$LUI)

# Set the reference levels for LUI
final_size <- final_size %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )

# Have a look at sample sizes
table(final_size$LUI)
```

# Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(final_size[ , c("Predominant_land_use", "Use_intensity")])
```

# Complete cases and model data

```{r}
# Lets have a look at rows with an NA for LU and LUI
final_size %>%
  filter(if_any(c(Predominant_land_use, Use_intensity, LUI), is.na)) %>% 
  dplyr::select(SS, SSBS,Predominant_land_use, Use_intensity, LUI) %>% 
  View()

final_size %>%
  filter(!if_any(c(Predominant_land_use, Use_intensity, LUI), is.na)) %>% 
  dplyr::select(SS, SSBS,Predominant_land_use, Use_intensity, LUI) %>% 
  View()

# drop all rows with NAs for land use and use intensity and drop empty levels
model_data_biomass <- drop_na(final_size, Predominant_land_use, Use_intensity) %>% droplevels() 

# check distributions
hist(model_data_biomass$LogBiomass)

# Have a look at sample sizes
table(model_data_biomass$Predominant_land_use, model_data_biomass$Use_intensity)
table(model_data_biomass$LUI)

length(unique(model_data_biomass$SS)) # 444 studies 
length(unique(model_data_biomass$SSBS)) # 11516 sites

rm(final_size, overall_site_size)
```

# Have a look at temperate versus tropical split for log biomass model data

```{r}
# Filter overall predicts for sites and their latitude
latitude <- model_data_biomass %>% dplyr::select(SSBS, Latitude) 

# There are duplicated sites
any(duplicated(latitude$SSBS))

# Lets get a list of distinct sites
latitude <- latitude %>% distinct(SSBS, .keep_all = TRUE)

# Add a new column classifying each site into tropical or temperate
latitude$climate_zone <- ifelse(latitude$Latitude >= -23.4368 & latitude$Latitude <= 23.4368, 
                          "tropical", 
                          "temperate")

table(latitude$climate_zone)

rm(latitude)
```



# CREATE AVERAGE LIVE MASS MODEL DATA

```{r}
# Create a separate data frame for average weighted mean size of an insect at a site
avg_mass <- all_insects_size_updated

# Calculate the biomass of each taxa
avg_mass <- avg_mass %>% 
  mutate(
    Biomass = mass_mg * Merged_diversity
  )

# there are 34 studies where all sites have zero total biomass, so min(TotalSiteBiomass[TotalSiteBiomass > 0]) has no values to take the minimum of and returns Inf
# Inspecting these studies reveals that none of the rows/taxa/sites have a body length value.
# I will remove rows with no body length value
avg_mass <- avg_mass %>%
  filter(!is.na(mass_mg_match_level))

# Calculate AvgSiteBiomass
site_avg_mass <- avg_mass %>%
  group_by(SSBS) %>%
  summarise(TotalSiteBiomass = sum(Biomass, na.rm = TRUE),
         TotalSiteAbundance = sum(Merged_diversity, na.rm = TRUE),
         LogAvgSiteBiomass = log(TotalSiteBiomass / TotalSiteAbundance)
         ) %>%  
  ungroup() 

# View rows where LogAvgSiteBiomass is 0, NA, or blank - i.e. no total abundance!
site_avg_mass %>%
  filter(is.na(LogAvgSiteBiomass) | LogAvgSiteBiomass == 0 | LogAvgSiteBiomass == "") %>% View()

avg_mass %>% filter(SSBS == "AD1_2006__Diekoetter 1  1") %>% View()
```

# Check if there are any studies where the average mass does not vary among sites and exclude them from the analysis
# If mass is the same in all sites, then either its a 1-species study or its a study where they're all in the same family/genus and we don't have any size data more finely resolved
# Either way, such studies cannot contribute to the model for average mass, so their inclusion would bias the results (it would flatten all the coefficients towards zero) so lets remove them

```{r}
# Get a list of the unique studies and their sites
unique_SS <- avg_mass %>% dplyr::select(SS, SSBS) %>% unique()

# First lets merge SS into out site_avg_mass data frame
site_avg_mass <- site_avg_mass %>% 
  left_join(unique_SS %>% dplyr::select(SSBS, SS), by = "SSBS") 

# Filter out those sites with NaN for LogAvgSiteBiomass due to log(0)
site_avg_mass <- site_avg_mass %>% filter(!is.nan(LogAvgSiteBiomass))

# Identify studies that contain sites with no variation in LogAvgSiteBiomass
no_variation <- site_avg_mass %>%
  group_by(SS) %>%
  summarise(n_unique = n_distinct(round(LogAvgSiteBiomass, 5))) %>% # round all values to 5 digits
  filter(n_unique == 1)

# Lets remove those rows which come from studies with no variation in LogAvgSiteBiomass (i.e. the same average mass value across sites)
filtered_site_avg_mass <- site_avg_mass %>% filter(!SS %in% no_variation$SS)

# Sense check
any(filtered_site_avg_mass$SS %in% no_variation$SS) # are any of these studies still present in my filtered data frame? 
```

# Calculating what % of taxa at a site have a unique mass_mg value. This captures both dimensions of the quality of the size data in that 1. If mass_mg values across taxa at a site are resolved to species level then the % of taxa with a unique mass_mg value will be high AND the taxa will all have different mass_mg values because they are resolved to species level.

```{r}
# Calculate the number of unique mass_mg values per site as a % of the overall number of rows per site
UniqueSiteMassValues <- avg_mass %>%
  group_by(SSBS) %>%
  summarise(
    unique_mass_mg = n_distinct(mass_mg),
    total_rows = n(),
    percent_unique = (unique_mass_mg / total_rows) * 100
  )

# Merge SS
UniqueSiteMassValues <- UniqueSiteMassValues %>% 
  left_join(unique_SS %>% dplyr::select(SSBS, SS), by = "SSBS") %>% 
  dplyr::select(SS, SSBS, unique_mass_mg, total_rows, percent_unique)

# Filter out those studies already identified with no variation in average mass among sites
FilteredUniqueSiteMassValues <- UniqueSiteMassValues %>% filter(!SS %in% no_variation$SS)
```

# Merge with PREDICTS insects
```{r}
# Use distinct() to keep only the first appearance of each site within the overall PREDICTS div extract for insects
insects_single <- insects %>% 
  distinct(SSBS, .keep_all = TRUE)

 # Drop the extra SS column 
filtered_site_avg_mass <- filtered_site_avg_mass %>% dplyr::select(-SS)

# merge site level avg mass values
final_avg_mass <- insects_single %>%
  left_join(filtered_site_avg_mass, by = "SSBS") %>% 
  droplevels()

# Have a look at sites with no value for average biomass, either 0 total abundance or no body size value assigned
no_avg_biomass <- final_avg_mass %>%
  filter(
    is.na(LogAvgSiteBiomass) |         # remove NA
    is.nan(LogAvgSiteBiomass)|        # remove NaN
    LogAvgSiteBiomass == 0 |            # remove 0
    LogAvgSiteBiomass == ""            # remove empty strings
  ) 

# filter out sites with no avg mass value 
final_avg_mass <- final_avg_mass %>%
  filter(
    !is.na(LogAvgSiteBiomass),         # remove NA
    !is.nan(LogAvgSiteBiomass),        # remove NaN
    LogAvgSiteBiomass != 0,            # remove 0
    LogAvgSiteBiomass != ""            # remove empty strings
  )

# 10646 sites in total before / 11086 sites with updated traits = extra 440 sites

# These are the studies with the same average mass value across sites
site_avg_mass %>% filter(SS %in% no_variation$SS) %>% View()

rm(insects_single, avg_mass, no_avg_biomass, site_avg_mass, unique_SS, no_variation, filtered_site_avg_mass, UniqueSiteMassValues)
```

# Tidy the LU classes

```{r}
# Rename predominant habitat since its not really habitat we're looking at
final_avg_mass <- rename(final_avg_mass,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
final_avg_mass <- final_avg_mass %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(final_avg_mass$Predominant_land_use, final_avg_mass$Use_intensity)
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
final_avg_mass <- final_avg_mass %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# Have a look at sample sizes
table(final_avg_mass$LUI)

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 61 sites, merge with light use
final_avg_mass <- final_avg_mass %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Have a look at sample sizes
table(final_avg_mass$LUI)

# Set the reference levels for LUI
final_avg_mass <- final_avg_mass %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )
```

# Create model data for average mass

```{r}
# drop all rows with NAs for land use and use intensity and drop empty levels
model_data_avg_mass <- drop_na(final_avg_mass, Predominant_land_use, Use_intensity) %>% droplevels() 

# check distributions
hist(model_data_avg_mass$LogAvgSiteBiomass)

# Have a look at sample sizes
table(model_data_avg_mass$Predominant_land_use, model_data_avg_mass$Use_intensity)
table(model_data_avg_mass$LUI)

length(unique(model_data_avg_mass$SS)) # 372 studies -> 390 studies
length(unique(model_data_avg_mass$SSBS)) # 8621 sites -> 8864 sites

rm(final_avg_mass)
```



# CREATE ABUNDANCE MODEL DATA 

```{r}
# All-insect set of site-level log(abundance) and species-richness data
# Has filtered out accidentally sampled larvae
abundance <-  qread("/Users/jusi/Desktop/multi-groups/data/abundance-data-prepared-all-insects.qs")

# Use dplyr to reshape the abundance data
# For each SSBS: sum the total abundance values then take the mean log_plus_what (they'll all be the same - comes from the minimum non-zero value across the entire study) 
# logAllInsectAbundance = log(sumTotalAbundance + mean_log_plus_what) - an overall abundance value for all of the insects at each site!
abundance_summary <- abundance %>%
  group_by(SSBS) %>%
  summarise(
    sumTotalAbundance = sum(Total_abundance, na.rm = TRUE),
    mean_log_plus_what = mean(Log_plus_what, na.rm = TRUE)
  ) %>%
  mutate(logAllInsectAbundance = log(sumTotalAbundance + mean_log_plus_what))

# Filter PREDICTS for same set of sites as in my biomass model data
filtered_abundance <-  abundance_summary %>%
  filter(SSBS %in% model_data_biomass$SSBS) %>%
  distinct(SSBS, .keep_all = TRUE)

# Append use intensity based on SSBS
filtered_abundance <- filtered_abundance %>%
  left_join(
    model_data_biomass %>% dplyr::select(SS, SSBS, Predominant_land_use, Use_intensity, Source_ID, SSB),
    by = "SSBS"
  )
```

# Start the modelling of all of the insects (Abundance responses)

```{r}
# Relevel the land use and use intensity classes
filtered_abundance <- filtered_abundance %>%
  mutate(
    # convert back to factors and set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(filtered_abundance$Predominant_land_use, filtered_abundance$Use_intensity)

levels(filtered_abundance$Predominant_land_use)
levels(filtered_abundance$Use_intensity)
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
filtered_abundance <- filtered_abundance %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 61 sites, merge with light use
filtered_abundance <- filtered_abundance %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Set the reference levels for LUI
filtered_abundance <- filtered_abundance %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )
```

# Create model data for abundance

```{r}
# drop all rows with NAs for land use, use intensity and log_abundance_px and drop empty levels
model_data_abundance <- drop_na(filtered_abundance, Predominant_land_use, Use_intensity, logAllInsectAbundance) %>% droplevels() 

# check distributions
hist(model_data_abundance$logAllInsectAbundance)
hist(abundance$Log_abundance_pX)

# Have a look at sample sizes
table(model_data_abundance$Predominant_land_use, model_data_abundance$Use_intensity)
table(model_data_abundance$LUI)

length(unique(model_data_abundance$SS)) # 444 studies 
length(unique(model_data_abundance$SSBS)) # 11516 sites

  length(unique(model_data_biomass$SS)) # 444 studies
 length(unique(model_data_biomass$SSBS)) # 11516 sites

  length(unique(model_data_avg_mass$SS)) # 372 studies
  length(unique(model_data_avg_mass$SSBS)) # 8621 sites
  
       

rm(abundance, abundance_summary, filtered_abundance)
```


# Now that we've created the model data for each of the variables (biomass, average live mass and abundance) I'll create the overall plot for all insects ~ LU 

# Modelling LogBiomass ~ LU and creating the LU plot for all insects

```{r}
# maximal model
m1 <- lmer(LogBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_biomass)

# remove block
m2 <-lmer(LogBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS),
    data = model_data_biomass)

# remove study
m3 <-lmer(LogBiomass ~ Predominant_land_use + (1 | Source_ID),
    data = model_data_biomass)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
coef(summary(m1))

# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(biomass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))

       
  #ggsave(filename = "./figures/C3_updated_figures/biomass_LU.png",  plot = p, width = 6, height = 4, dpi = 300)

biomass_eff <- eff %>% mutate(Variable = "Total biomass")

rm(eff, m1, m2, m3, p)     
```

# Modelling AVG_MASS ~ LU and creating the LU plot for all insects

```{r}
# maximal model
m1 <- lmer(LogAvgSiteBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_avg_mass)

# remove block
m2 <-lmer(LogAvgSiteBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS),
    data = model_data_avg_mass)

# remove study
m3 <-lmer(LogAvgSiteBiomass ~ Predominant_land_use + (1 | Source_ID),
    data = model_data_avg_mass)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))

# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(average mass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
  
     
  #ggsave(filename = "./figures/C3_updated_figures/avg_mass_LU.png",  plot = p, width = 6, height = 4, dpi = 300)
  avgmass_eff <- eff %>% mutate(Variable = "Average live mass")
  
rm(eff, m1, m2, m3, p)     
```

# Modelling ABUNDANCE ~ LU and creating the LU plot for all insects

```{r}
# maximal model
m1 <- lmer(logAllInsectAbundance ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_abundance)

# remove block
m2 <-lmer(logAllInsectAbundance ~ Predominant_land_use + (1 | Source_ID) + (1 | SS),
    data = model_data_abundance)

# remove study
m3 <-lmer(logAllInsectAbundance ~ Predominant_land_use + (1 | Source_ID),
    data = model_data_abundance)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))

# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(total abundance) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
  
     
     abundance_eff <- eff %>% mutate(Variable = "Species abundance")
  
rm(eff, m1, m2, m3, p)     
```

# Combine Biomass, Mass and Abundance into one plot with three different colours

```{r}
biomass_eff <- biomass_eff %>% mutate("Response variable" = "Total biomass")
avgmass_eff <- avgmass_eff %>% mutate("Response variable" = "Average mass")
abundance_eff <- abundance_eff %>% mutate("Response variable" = "Species abundance")


combined <- rbind(biomass_eff, avgmass_eff, abundance_eff)


p <- combined %>%
  ggplot(aes(x = Predominant_land_use,
             y = Percent_diff,
             colour = Variable,
             group = Variable)) +
  
  # zero line
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  
  # confidence intervals for each Variable
  geom_linerange(aes(ymin = Percent_lower,
                     ymax = Percent_upper),
                 linewidth = 1,
                 position = position_dodge(width = 0.75)) +
  
  # points for each Variable
  geom_point(size = 3,
             position = position_dodge(width = 0.75)) +
  
  # manual colours
  scale_colour_manual(values = c(
    "Total biomass" = "purple",
    "Average live mass" = "blue",
    "Species abundance" = "red"
  )) +
  
  # aesthetics & theme
  theme_classic() +
  geom_vline(xintercept = 1.5, linetype = 2) +
  geom_vline(xintercept = 2.5, linetype = 2) +
  geom_vline(xintercept = 3.5, linetype = 2) +
  geom_vline(xintercept = 4.5, linetype = 2) +
  geom_vline(xintercept = 5.5, linetype = 2) +
  geom_vline(xintercept = 6.5, linetype = 2) +
  geom_vline(xintercept = 7.5, linetype = 2) +
  
  theme(
    axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "bottom"
  ) +
  
  xlab("Land use intensity class") +
  ylab("Difference in log response variable (%)") +
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 6, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = 8)) +
  theme(legend.title = element_text(face = "bold"),
        legend.text = element_text(size = 10)) 


 #ggsave(filename = "./figures/C3_updated_figures/combined-LU-plot.png",  plot = p, width = 8, height = 5, dpi = 300)

rm(abundance_eff, avgmass_eff, biomass_eff, combined)
```


# Next lets do the plot for all insects ~ LUI

# Modelling LogBiomass ~ LUI (LU + UI) and creating the LUI plot for all insects

```{r}
# maximal model
m1 <- lmer(LogBiomass ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_biomass)


m2 <-lmer(LogBiomass ~ LUI + (1 | Source_ID) + (1 | SS),
    data = model_data_biomass)

# remove study
m3 <-lmer(LogBiomass ~ LUI + (1 | Source_ID),
    data = model_data_biomass)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary (minimal)",
                            "LUIPrimary_Light_use" = "Primary (light)",
                                "LUIPrimary_Intense_use" = "Primary (intense)",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV (minimal)",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV (light)",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV (intense)",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV (minimal)",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV (high)",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV (minimal)",
                            "LUIMature_secondary_vegetation_High_use" = "MSV (high)",
                            "LUIPlantation_forest_Minimal_use" = "Plantation (minimal)",
                            "LUIPlantation_forest_Light_use" = "Plantation (light)",
                            "LUIPlantation_forest_Intense_use" = "Plantation (intense)",
                            "LUIPasture_Minimal_use" = "Pasture (minimal)",
                            "LUIPasture_Light_use" = "Pasture (light)",
                            "LUIPasture_Intense_use" = "Pasture (intense)",
                            "LUICropland_Minimal_use" = "Cropland (minimal)",
                            "LUICropland_Light_use" = "Cropland (light)",
                            "LUICropland_Intense_use" = "Cropland (intense)",
                            "LUIUrban_Minimal_use" = "Urban (minimal)",
                            "LUIUrban_Light_use" = "Urban (light)",
                            "LUIUrban_Intense_use" = "Urban (intense)")
                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    BiomassLUIPlot <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, colour = "purple")+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1, colour = "purple")+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(biomass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
#       ggsave(filename = "./figures/C3_updated_figures/biomass_LUI.png",  plot = p, width = 8, height = 7, dpi = 300)

biomass_LUI_eff <- eff %>% mutate(Variable = "Total biomass")    
  
rm(eff, m1, m2, m3)     
```

# Modelling AVG_MASS ~ LUI (LU + UI) and creating the LUI plot for all insects

```{r}
# maximal model
m1 <- lmer(LogAvgSiteBiomass ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_avg_mass)

# remove block
m2 <-lmer(LogAvgSiteBiomass ~ LUI + (1 | Source_ID) + (1 | SS),
    data = model_data_avg_mass)

# remove study
m3 <-lmer(LogAvgSiteBiomass ~ LUI + (1 | Source_ID),
    data = model_data_avg_mass)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary (minimal)",
                            "LUIPrimary_Light_use" = "Primary (light)",
                                "LUIPrimary_Intense_use" = "Primary (intense)",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV (minimal)",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV (light)",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV (intense)",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV (minimal)",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV (high)",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV (minimal)",
                            "LUIMature_secondary_vegetation_High_use" = "MSV (high)",
                            "LUIPlantation_forest_Minimal_use" = "Plantation (minimal)",
                            "LUIPlantation_forest_Light_use" = "Plantation (light)",
                            "LUIPlantation_forest_Intense_use" = "Plantation (intense)",
                            "LUIPasture_Minimal_use" = "Pasture (minimal)",
                            "LUIPasture_Light_use" = "Pasture (light)",
                            "LUIPasture_Intense_use" = "Pasture (intense)",
                            "LUICropland_Minimal_use" = "Cropland (minimal)",
                            "LUICropland_Light_use" = "Cropland (light)",
                            "LUICropland_Intense_use" = "Cropland (intense)",
                            "LUIUrban_Minimal_use" = "Urban (minimal)",
                            "LUIUrban_Light_use" = "Urban (light)",
                            "LUIUrban_Intense_use" = "Urban (intense)")
                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    MassLUIPlot <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, colour = "blue")+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1, colour = "blue")+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(average mass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10))) +
      scale_y_continuous(
  breaks = seq(-25, 50, by = 25)   # add more ticks
)

     
       #ggsave(filename = "../figures/avg_mass_LUI.png",  plot = p, width = 8, height = 7, dpi = 300)
       
       mass_LUI_eff <- eff %>% mutate(Variable = "Average live mass")
       #mass_LUI_eff <- eff %>% mutate(Data = "All of the data")
     
rm(eff, m1, m2, m3)     
```

# Modelling ABUNDANCE ~ LUI (LU + UI) and creating the LUI plot for all insects

```{r}
# maximal model
m1 <- lmer(logAllInsectAbundance ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_abundance)

# remove block
m2 <-lmer(logAllInsectAbundance ~ LUI + (1 | Source_ID) + (1 | SS),
    data = model_data_abundance)

# remove study
m3 <-lmer(logAllInsectAbundance ~ LUI + (1 | Source_ID),
    data = model_data_abundance)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary (minimal)",
                            "LUIPrimary_Light_use" = "Primary (light)",
                                "LUIPrimary_Intense_use" = "Primary (intense)",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV (minimal)",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV (light)",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV (intense)",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV (minimal)",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV (high)",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV (minimal)",
                            "LUIMature_secondary_vegetation_High_use" = "MSV (high)",
                            "LUIPlantation_forest_Minimal_use" = "Plantation (minimal)",
                            "LUIPlantation_forest_Light_use" = "Plantation (light)",
                            "LUIPlantation_forest_Intense_use" = "Plantation (intense)",
                            "LUIPasture_Minimal_use" = "Pasture (minimal)",
                            "LUIPasture_Light_use" = "Pasture (light)",
                            "LUIPasture_Intense_use" = "Pasture (intense)",
                            "LUICropland_Minimal_use" = "Cropland (minimal)",
                            "LUICropland_Light_use" = "Cropland (light)",
                            "LUICropland_Intense_use" = "Cropland (intense)",
                            "LUIUrban_Minimal_use" = "Urban (minimal)",
                            "LUIUrban_Light_use" = "Urban (light)",
                            "LUIUrban_Intense_use" = "Urban (intense)")

                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    AbundanceLUIPlot <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, colour = "red")+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1, colour = "red")+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(total abundance) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10))) 
      
     
       #ggsave(filename = "../figures/abundance_LUI.png",  plot = p, width = 8, height = 7, dpi = 300)
    
rm(eff, m1, m2, m3)     
```


# Combine all three LUI plots for biomass, mass and abundance. Stack them above one another with the x labels only appearing once

```{r}
library(patchwork)

p1 <- AbundanceLUIPlot + theme(axis.title.y = element_text(size = 12), axis.text.x  = element_text(size = 12))
p2 <- MassLUIPlot + theme(axis.title.y  = element_text(size = 12), axis.text.x  = element_text(size = 12))
p3 <- BiomassLUIPlot + theme(axis.title.y = element_text(size = 12), axis.text.x  = element_text(size = 12))

final_plot <- (p1 / p2 / p3) &
  theme(legend.position = "bottom")

# Now blank FIRST TWO x axes again (after global theme):
final_plot[[1]] <- final_plot[[1]] + theme(axis.text.x = element_blank(),
                                           axis.title.x = element_blank())
final_plot[[2]] <- final_plot[[2]] + theme(axis.text.x = element_blank(),
                                           axis.title.x = element_blank())

  #ggsave(filename = "./figures/C3_updated_figures/combined-LUI-plot.png",  plot = final_plot, width = 18, height = 10, dpi = 300)

rm(p1, p2, p3, MassLUIPlot, AbundanceLUIPlot, BiomassLUIPlot, final_plot)
```



# SENSITIVITY ANALYSIS

# Figure which removes studies with < 10% unique size values across taxa within a site for LU

```{r}
# Pull studies with less than 10 percent unique size values among taxa sampled at a site
lessthan10 <- FilteredUniqueSiteMassValues %>%
  filter(percent_unique < 10) %>%
  pull(SS) %>% 
  unique()

# Create the model data by filtering out those studies
model10 <- model_data_avg_mass %>%
  filter(!SS %in% lessthan10)

# LU model with reduced studies
reduced_model <- lmer(LogAvgSiteBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model10)

# model estimates
summary(reduced_model)
anova(reduced_model)
model_plot(reduced_model)


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(reduced_model, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(average mass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
  #ggsave(filename = "../figures/avg_mass_LU_10percent.png",  plot = p, width = 6, height = 4, dpi = 300)

mass_sensitivity_eff <- eff %>% mutate(Analysis = "Sensitivity analysis")
  
rm(eff, reduced_model, model10, lessthan10, p)     

```

# # Figure which removes studies with < 10% unique size values across taxa within a site for LUI 

```{r}
lessthan10 <- FilteredUniqueSiteMassValues %>%
  filter(percent_unique < 10) %>%
  pull(SS) %>% 
  unique()

model10 <- model_data_avg_mass %>%
  filter(!SS %in% lessthan10)

reduced_model <- lmer(LogAvgSiteBiomass ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model10)

# model estimates
summary(reduced_model)
anova(reduced_model)
      
      
      # gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(reduced_model, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
 # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary (minimal)",
                            "LUIPrimary_Light_use" = "Primary (light)",
                                "LUIPrimary_Intense_use" = "Primary (intense)",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV (minimal)",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV (light)",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV (intense)",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV (minimal)",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV (high)",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV (minimal)",
                            "LUIMature_secondary_vegetation_High_use" = "MSV (high)",
                            "LUIPlantation_forest_Minimal_use" = "Plantation (minimal)",
                            "LUIPlantation_forest_Light_use" = "Plantation (light)",
                            "LUIPlantation_forest_Intense_use" = "Plantation (intense)",
                            "LUIPasture_Minimal_use" = "Pasture (minimal)",
                            "LUIPasture_Light_use" = "Pasture (light)",
                            "LUIPasture_Intense_use" = "Pasture (intense)",
                            "LUICropland_Minimal_use" = "Cropland (minimal)",
                            "LUICropland_Light_use" = "Cropland (light)",
                            "LUICropland_Intense_use" = "Cropland (intense)",
                            "LUIUrban_Minimal_use" = "Urban (minimal)",
                            "LUIUrban_Light_use" = "Urban (light)",
                            "LUIUrban_Intense_use" = "Urban (intense)")

                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    p <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, colour = "red")+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1, colour = "red")+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(average mass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
       #ggsave(filename = "../figures/avg_mass_LUI_10percent.png",  plot = p, width = 8, height = 7, dpi = 300)
    
    mass_LUI_sensitivity_eff <- eff %>% mutate(Data = "Studies with homogenous size data excluded")
     
rm(eff, m1, m2, m3, p)    

```

# Combine coefficients from non-sensitivity and sensitivity analysis of mass (LU)

```{r}
avgmass_eff <- avgmass_eff %>% mutate(Data = "All of the data")
mass_sensitivity_eff <- mass_sensitivity_eff %>% mutate(Data = "Studies with homogenous size data excluded")

combined <- rbind(mass_sensitivity_eff, avgmass_eff)

p <- combined %>%
  ggplot(aes(x = Predominant_land_use,
             y = Percent_diff,
             colour = Data,
             group = Data)) +
  
  # zero line
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  
  # confidence intervals for each Variable
  geom_linerange(aes(ymin = Percent_lower,
                     ymax = Percent_upper),
                 linewidth = 1,
                 position = position_dodge(width = 0.75)) +
  
  # points for each Variable
  geom_point(size = 3,
             position = position_dodge(width = 0.75)) +
  
  # manual colours
  scale_colour_manual(values = c(
    "All of the data" = "black",
    "Studies with homogenous size data excluded" = "red"
  )) +
  
  # aesthetics & theme
  theme_classic() +
  geom_vline(xintercept = 1.5, linetype = 2) +
  geom_vline(xintercept = 2.5, linetype = 2) +
  geom_vline(xintercept = 3.5, linetype = 2) +
  geom_vline(xintercept = 4.5, linetype = 2) +
  geom_vline(xintercept = 5.5, linetype = 2) +
  geom_vline(xintercept = 6.5, linetype = 2) +
  geom_vline(xintercept = 7.5, linetype = 2) +
  
  theme(
    axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "bottom"
  ) +
  
  xlab("Land use intensity class") +
  ylab("log(average mass) difference (%)") +
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 6, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = 8)) +
  theme(legend.title = element_text(face = "bold"),
        legend.text = element_text(size = 10)) 

  #ggsave(filename = "../figures/C3-sensitivity-mass.png",  plot = p, width = 8, height = 5, dpi = 300)
  #saveRDS(combined, "../figures/C3 figures/C3-mass-LU-sensitivity-df.rds")


```

# Combine coefficients from non-sensitivity and sensitivity analysis of mass (LUI)

```{r}
avgmass_eff <- avgmass_eff %>% mutate(Data = "All of the data")
mass_sensitivity_eff <- mass_sensitivity_eff %>% mutate(Data = "Studies with homogenous size data excluded")

combined <- rbind(mass_LUI_eff, mass_LUI_sensitivity_eff)

p <- combined %>%
  ggplot(aes(x = Predominant_land_use,
             y = Percent_diff,
             colour = Data,
             group = Data)) +
  
  # zero line
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  
  # confidence intervals for each Variable
  geom_linerange(aes(ymin = Percent_lower,
                     ymax = Percent_upper),
                 linewidth = 1,
                 position = position_dodge(width = 0.75)) +
  
  # points for each Variable
  geom_point(size = 3,
             position = position_dodge(width = 0.75)) +
  
  # manual colours
  scale_colour_manual(values = c(
    "All of the data" = "black",
    "Studies with homogenous size data excluded" = "red"
  )) +
  
  # aesthetics & theme
  theme_classic() +
   geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
  
  theme(
    axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "bottom"
  ) +
  
  xlab("Land use intensity class") +
  ylab("log(average mass) difference (%)") +
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 6, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = 8)) +
  theme(legend.title = element_text(face = "bold"),
        legend.text = element_text(size = 10)) 

  #ggsave(filename = "../figures/C3 figures/C3-sensitivity-mass-LUI.png",  plot = p, width = 10, height = 6, dpi = 300)
  #saveRDS(combined, "../figures/C3 figures/C3-mass-LU-sensitivity-df.rds")
  
combined <- read_rds(file.choose())  

  combined <- combined %>% mutate(percentdiff = round(Percent_upper - Percent_lower, 2)


```



# Start with biomass ~ LU * Order analysis

```{r}
# Create a separate data frame
orders_size <- all_insects_size_updated 

# Lets look at the number of sites per order
orders_size %>%
  group_by(Order) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Orders to keep and assign all others to "other_orders"
target_orders <- c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera", "Hemiptera")

orders_size <- orders_size %>%
  mutate(Order = case_when(
    Order %in% target_orders ~ Order,
    TRUE ~ "Other_orders"
  ))

unique(orders_size$Order)
```

# (NEW CODE) the big five orders - up to five size values per site

```{r}
# Calculate the biomass of each taxa
orders_size <- orders_size %>% 
  mutate(
    Biomass = mass_mg * Merged_diversity
  )

# there are 34 studies where all sites have zero total biomass, so min(TotalSiteBiomass[TotalSiteBiomass > 0]) has no values to take the minimum of and returns Inf
# Inspecting these studies reveals that none of the rows/taxa/sites have a body length value.
# I will remove rows with no body length value
orders_size <- orders_size %>%
  filter(!is.na(mass_mg_match_level))

orders_size <- orders_size %>%
  group_by(SSBS, Order) %>% # Calculate for each site x order combination
  mutate(TotalSiteOrderBiomass = sum(Biomass, na.rm = TRUE)) %>%  # Calculate total site biomass for each order 
  ungroup() %>%
  group_by(SS) %>% # Now group by study
  mutate(
    SmallestNonZeroTotalSiteOrderBiomass = min(TotalSiteOrderBiomass[TotalSiteOrderBiomass > 0]), # Which site x order combo has the smallest total biomass greater than 0 per study? 
    Log_plus_what = ifelse(
    Diversity_metric_unit %in% c("individuals", "times observed"), min(mass_mg, na.rm = TRUE),  
    # If abundance data are counts, assign the smallest mass_mg for an individual of any species in the study
    SmallestNonZeroTotalSiteOrderBiomass / 2)) %>%  # else 1/2 the SmallestNonZeroTotalSiteOrderBiomass for any site within the study 
  ungroup() %>% 
  group_by(SSBS, Order) %>%  # Calculate for each site x order combination
  mutate(LogBiomass = log(TotalSiteOrderBiomass + Log_plus_what)) %>% 
  ungroup()

# Calculate one LogBiomass value per site x order combination
LogBiomass_summary <- orders_size %>%
  group_by(SSBS, Order) %>%
  summarise(
    n_TSOB = n_distinct(TotalSiteOrderBiomass), 
    n_LPW = n_distinct(Log_plus_what),
    is_consistent = (n_TSOB == 1 & n_LPW == 1), # sense check that all TotalSiteOrderBiomass and Log_plus_what values are the same for all rows in the site x order combination
    TotalSiteBiomass = first(TotalSiteOrderBiomass),
    Log_plus_what = first(Log_plus_what),
    LogBiomass = log(first(TotalSiteBiomass) + first(Log_plus_what)),
    .groups = "drop"
  )

# All sites are consistent
LogBiomass_summary %>% filter(is_consistent != TRUE)

inconsistent_groups <- orders_size %>%
  group_by(SSBS, Order) %>%
  summarise(
    n_TSB = n_distinct(TotalSiteOrderBiomass),
    n_LPW = n_distinct(Log_plus_what)
  ) %>%
  filter(n_TSB > 1 | n_LPW > 1) 

# No rows where LogBiomass is 0, NA, or blank
LogBiomass_summary %>%
  filter(is.na(LogBiomass) | LogBiomass == 0 | LogBiomass == "") %>% View()

rm(inconsistent_groups)
```

# Merge with PREDICTS but this time keep a row for each order within each site!
```{r}
# Create the other orders column so we filter for the correct number of sites
insects <- insects %>%
  mutate(Order = case_when(
    Order %in% target_orders ~ Order,
    TRUE ~ "Other_orders"
  ))

# Use distinct() to keep only the first appearance of each Order within each site
ssbs_by_order <- insects %>%
  distinct(SSBS, Order, .keep_all = TRUE) # 23,381 sites!

# merge LogBiomass_summary values
BiomassByOrder <- ssbs_by_order %>%
  left_join(LogBiomass_summary, by = c("SSBS", "Order")) %>% 
  droplevels()

# Have a look at sites with no LogBiomass
no_biomass <- BiomassByOrder %>%
  filter(
    is.na(LogBiomass) |         # remove NA
    is.nan(LogBiomass)|        # remove NaN
    LogBiomass == ""            # remove empty strings
  )  # I think this is because these rows do not have a body size value, therefore no LogBiomass value. 

all_insects_size_updated %>% filter(SS == "AC1_2022__Kinneen 1") %>% View()

# filter out sites with no LogBiomass
BiomassByOrder <- BiomassByOrder %>%
  filter(
    !is.na(LogBiomass),         # remove NA
    !is.nan(LogBiomass),        # remove NaN
    LogBiomass != ""            # remove empty strings
  )

sum(!is.na(LogBiomass_summary$LogBiomass)) # number of site x order combos with data, sensitivity check as should match number of sites in BiomassByOrder 

rm(ssbs_by_order, no_biomass, LogBiomass_summary, orders_size)
```

# Prep model data
```{r}
# Create model data
big_five <- BiomassByOrder 

# Lets look at the number of sites per order
big_five %>%
  group_by(Order) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>% 
  View()

# Rename predominant habitat since its not really habitat we're looking at
big_five <- rename(big_five,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
big_five <- big_five %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   #Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
   
       # collapse secondary age categories into one and include SV (indeterminate age)
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Young secondary vegetation" = "Secondary vegetation", 
                                         "Intermediate secondary vegetation" = "Secondary vegetation",
                                         "Mature secondary vegetation" = "Secondary vegetation",
                                         "Secondary vegetation (indeterminate age)" = "Secondary vegetation"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(big_five$Order, big_five$Predominant_land_use)

```

# Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(big_five[ , c("Predominant_land_use", "Use_intensity")])
```

# Complete cases and summary stats

```{r}
# Lets have a look at rows with an NA for land use or use intensity
big_five %>%
  filter(if_any(c(Predominant_land_use, Use_intensity), is.na)) %>% 
  dplyr::select(SS, SSBS, Predominant_land_use, LogBiomass, Use_intensity) %>% 
  View()

# drop all rows with NAs for land use and use intensity and drop empty levels
big_five <- drop_na(big_five, Predominant_land_use, Use_intensity) %>% droplevels() 

# check distribution of LogBiomass
hist(big_five$LogBiomass)

# Have a look at sample sizes
table(big_five$Predominant_land_use, big_five$Use_intensity)
table(big_five$Order, big_five$Predominant_land_use)

length(unique(big_five$SS)) # 455 studies
length(unique(big_five$SSBS)) # 12232 sites

# Save the data frame for ease in the future
#saveRDS(big_five, "BigFiveBiomass.rds")
```

# Modelling biomass ~ LU * Order

```{r}
# Set order as a factor variable for modelling
big_five <- big_five %>%
  mutate(Order = as.factor(Order))

# maximal model
m1 <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = big_five)

# remove block
m2 <-lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS),
    data = big_five)

# remove study
m3 <-lmer(LogBiomass ~ Predominant_land_use + (1 | Source_ID),
    data = big_five)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)

rm(m2, m3)
```

# Coleoptera effects
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
col_eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  col_eff$term <- as.factor(col_eff$term)
  
   # pull out effects
  col_eff <- col_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  col_eff$term <- recode_factor(col_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  col_eff <- col_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    col_eff <- col_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      col_eff[1,7] <- 0 #shift median
      col_eff[1,8] <- 0 #shift upper
      col_eff[1,9] <- 0 #shift lower
      
      colnames(col_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   col_eff <- col_eff %>% mutate(Order = "Coleoptera")

  # make the order column a factor so it can be recoded
   col_eff$Order <- as.factor(col_eff$Order)      
      
rm(m1)
```

# Diptera effects
```{r}
# Relevel model with Diptera as reference level
dip_ref <- within(big_five, Order <- relevel(Order, ref = "Diptera"))

# dip model
dip_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = dip_ref)

summary(dip_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
dip_eff <- FEsim(dip_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  dip_eff$term <- as.factor(dip_eff$term)
  
   # pull out effects
  dip_eff <- dip_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  dip_eff$term <- recode_factor(dip_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  dip_eff <- dip_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    dip_eff <- dip_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      dip_eff[1,7] <- 0 #shift median
      dip_eff[1,8] <- 0 #shift upper
      dip_eff[1,9] <- 0 #shift lower
      
      colnames(dip_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   dip_eff <- dip_eff %>% mutate(Order = "Diptera")

  # make the order column a factor so it can be recoded
   dip_eff$Order <- as.factor(dip_eff$Order)      
      
rm(dip_ref, dip_m)
```

# Hemiptera effects
```{r}
# Relevel model with Hemiptera as reference level
hemi_ref <- within(big_five, Order <- relevel(Order, ref = "Hemiptera"))

# dip model
hemi_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = hemi_ref)

summary(hemi_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
hemi_eff <- FEsim(hemi_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  hemi_eff$term <- as.factor(hemi_eff$term)
  
   # pull out effects
  hemi_eff <- hemi_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  hemi_eff$term <- recode_factor(hemi_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hemi_eff <- hemi_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    hemi_eff <- hemi_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      hemi_eff[1,7] <- 0 #shift median
      hemi_eff[1,8] <- 0 #shift upper
      hemi_eff[1,9] <- 0 #shift lower
      
      colnames(hemi_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   hemi_eff <- hemi_eff %>% mutate(Order = "Hemiptera")

  # make the order column a factor so it can be recoded
   hemi_eff$Order <- as.factor(hemi_eff$Order)      
      
rm(hemi_ref, hemi_m)
```

# Hymenoptera effects
```{r}
# Relevel model with Hymenoptera as reference level
hym_ref <- within(big_five, Order <- relevel(Order, ref = "Hymenoptera"))

# dip model
hym_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = hym_ref)

summary(hym_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
hym_eff <- FEsim(hym_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  hym_eff$term <- as.factor(hym_eff$term)
  
   # pull out effects
  hym_eff <- hym_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  hym_eff$term <- recode_factor(hym_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hym_eff <- hym_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    hym_eff <- hym_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      hym_eff[1,7] <- 0 #shift median
      hym_eff[1,8] <- 0 #shift upper
      hym_eff[1,9] <- 0 #shift lower
      
      colnames(hym_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   hym_eff <- hym_eff %>% mutate(Order = "Hymenoptera")

  # make the order column a factor so it can be recoded
   hym_eff$Order <- as.factor(hym_eff$Order)     
      
rm(hym_ref, hym_m)
```

# Lepidoptera effects
```{r}
# Relevel model with Lepidoptera as reference level
lep_ref <- within(big_five, Order <- relevel(Order, ref = "Lepidoptera"))

# dip model
lep_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = lep_ref)

summary(lep_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
lep_eff <- FEsim(lep_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  lep_eff$term <- as.factor(lep_eff$term)
  
   # pull out effects
  lep_eff <- lep_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  lep_eff$term <- recode_factor(lep_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  lep_eff <- lep_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    lep_eff <- lep_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      lep_eff[1,7] <- 0 #shift median
      lep_eff[1,8] <- 0 #shift upper
      lep_eff[1,9] <- 0 #shift lower
      
      colnames(lep_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   lep_eff <- lep_eff %>% mutate(Order = "Lepidoptera")

  # make the order column a factor so it can be recoded
   lep_eff$Order <- as.factor(lep_eff$Order)     
      
rm(lep_ref, lep_m)
```

# Other_orders effects
```{r}
# Relevel model with Other_orders as reference level
other_ref <- within(big_five, Order <- relevel(Order, ref = "Other_orders"))

# other_orders model
other_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = other_ref)

summary(other_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
other_eff <- FEsim(other_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  other_eff$term <- as.factor(other_eff$term)
  
   # pull out effects
  other_eff <- other_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  other_eff$term <- recode_factor(other_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  other_eff <- other_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    other_eff <- other_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      other_eff[1,7] <- 0 #shift median
      other_eff[1,8] <- 0 #shift upper
      other_eff[1,9] <- 0 #shift lower
      
      colnames(other_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   other_eff <- other_eff %>% mutate(Order = "Other_orders")

  # make the order column a factor so it can be recoded
   other_eff$Order <- as.factor(other_eff$Order)     
      
rm(other_ref, other_m)
```

# Combine the effects and plot them
```{r}
combined <- rbind(col_eff, dip_eff, hemi_eff, hym_eff, lep_eff, other_eff)

# Add land-use factors for plotting
comb_plot_limits <- c("Primary", "Secondary vegetation", "Plantation", "Pasture", "Cropland", "Urban") 

# Add specific colours for each order
group_colours <- c(
  "#1b9e77", "#d95f02", "#7570b3",
  "#e7298a", "#66a61e", "#e6ab02"
)

# Remove rows for plantation for Leps and other_orders which have crazy CIs
combined <- combined[-c(27,33),]

  p <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = Percent_diff, colour = Order, group = Order) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper),
                 position = position_dodge(width = 0.5), linewidth = 1) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  theme(
    axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "bottom"        # 👈 Move legend
  ) +
  xlab("Land use intensity class") +
  ylab("log(biomass) difference (%)") +
  scale_colour_manual(values = group_colours) +
  guides(color = guide_legend(override.aes = list(shape = 19, size = 3))) +  # 👈 round icons
  scale_colour_manual(
  values = group_colours,
  labels = c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera",
             "Lepidoptera", "Other orders") # Change Other_orders to Other orders
) +
theme(
  legend.title = element_text(face = "bold")
) +
  scale_y_continuous(breaks = seq(-100, 150, by = 25))
 
        
#ggsave(filename = "./figures/C3_updated_figures/big_five_LU_second.png",  plot = p, width = 8, height = 6, dpi = 300)

rm(col_eff, dip_eff, hemi_eff, hym_eff, lep_eff, p, other_eff)
```




# SAVE DFS TO SPEED UP IN THE FUTURE

```{r}
# Save combined coefficients for the big five orders
# saveRDS(combined, "/Users/jusi/Desktop/multi-groups/data/C3_updated_data/combined_coefficients_bigfive.rds")

# Save model data for the big five orders
# saveRDS(big_five, "/Users/jusi/Desktop/multi-groups/data/C3_updated_data/model_data_bigfive.rds")

# Save abundance model data
# saveRDS(model_data_abundance, "/Users/jusi/Desktop/multi-groups/data/C3_updated_data/model_data_abundance.rds")

# Save average live mass model data
# saveRDS(model_data_avg_mass, "/Users/jusi/Desktop/multi-groups/data/C3_updated_data/model_data_avg_mass.rds")

# Save biomass model data
# saveRDS(model_data_biomass, "/Users/jusi/Desktop/multi-groups/data/C3_updated_data/model_data_biomass.rds")

```

