---
title: "Untitled"
output: html_document
date: "2026-02-09"
---

# Clear the work space

```{r}
rm(list=ls())
```

# Load required packages

```{r, include =FALSE}
library(tidyverse) # data processing
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(sjPlot) # for visualising results
library(effects) # for extracting model effects
library(merTools) # useful for a few things, but we're using it for extracting estimates for plotting
library(emmeans) # testing multiple comparisons (but also useful for plotting)
library(coefplot) # visualising model coefficients
library(lmerTest) # for p values and degrees of freedom from anova table
library(ggpubr) # arranging plots
library(glmmTMB)
library(glmm.hp)
```

# Load the datasets

```{r}
model_data_ab <- readRDS("./C2 data/ModelDataAbundanceNoBug.rds")
model_data_sr <- readRDS("./C2 data/ModelDataSpeciesRichnessNoBug.rds")
```

# Convert to factors 
```{r include = FALSE}
str(model_data_ab)
model_data_ab <- model_data_ab %>%
  mutate(across(c(SS, SSB, SSBS), as.factor))

str(model_data_sr)
model_data_sr <- model_data_sr %>%
  mutate(across(c(SS, SSB, SSBS), as.factor))

length(unique(model_data_ab$SS))
length(unique(model_data_sr$SS))
length(unique(model_data_ab$SSBS))
length(unique(model_data_sr$SSBS))
```

# First lets deal with the non-integer count data in the model data 

```{r}
# Are there non count data in model_data_ab?
unique(levels(model_data_ab$Diversity_metric_unit)) # yes

# First lets filter for integer only data
model_data_ab_integer_only <- model_data_ab %>% filter(Diversity_metric_unit == "individuals") %>% droplevels()

# Check again
unique(levels(model_data_ab_integer_only$Diversity_metric_unit))

# Are there non count data in model_data_sr?
unique(levels(model_data_sr$Diversity_metric_unit)) # yes

# First lets filter for integer only data
model_data_sr_integer_only <- model_data_ab %>% filter(Diversity_metric_unit == "individuals") %>% droplevels()

# Check again
unique(levels(model_data_sr_integer_only$Diversity_metric_unit))
```

# Great now both the abundance and species richness model data frames are for count data only
# But are there still non-integers despite the fact we've filtered only for individuals?
```{r}
# If you subtract each total abundance value in my data frame from the rounded version of the same value, is that value less than 0.000000001? (across all values)
all(abs(model_data_ab_integer_only$TotalAbundance -
        round(model_data_ab_integer_only$TotalAbundance)) < 1e-8,
    na.rm = TRUE) # FALSE

all(abs(model_data_sr_integer_only$SpeciesRichness -
        round(model_data_sr_integer_only$SpeciesRichness)) < 1e-8,
    na.rm = TRUE) # TRUE

# There are still non-integers in the abundance data, lets filter for them

# Check which values are non-integers
still_non_integer_abundance <- model_data_ab_integer_only %>%
  dplyr::filter(abs(TotalAbundance - round(TotalAbundance)) > 1e-8) %>% dplyr::select(SS, SSBS, SSBST, Best_guess_binomial, Diversity_metric, Diversity_metric_unit, Diversity_metric_type,Effort_corrected_measurement, Corrected_sampling_effort,  merged_diversity, TotalAbundance, SpeciesRichness, to_add, logAbundance) %>%  droplevels()

# Which studies?
unique(still_non_integer_abundance$SS)

# Ok lets round all of the Total Abundance values to the nearest integer, any value greater than 0.5 should become 1!
still_non_integer_abundance <- still_non_integer_abundance %>%
  mutate(
    RoundedTotalAbundance = if_else(
      TotalAbundance >= 0.5,
      ceiling(TotalAbundance),
      0
    )
  )

# Ok lets merge these rounded values back into the overall df for those SSBSTs
rounded_abundances <- still_non_integer_abundance %>%  dplyr::select(SSBST, RoundedTotalAbundance) # 407 SSBSTs have had their abundances rounded

# Merge into original data frame
model_data_ab_integer_only <- model_data_ab_integer_only %>% 
    left_join(rounded_abundances, by = ("SSBST")) %>%  # join the values in by their ssbst
  mutate(
    TotalAbundance = coalesce(RoundedTotalAbundance, TotalAbundance) # if a rounded values exists, assign it to TotalAbundance otherwise keep the original value
  ) %>% 
  dplyr::select(-RoundedTotalAbundance)

# Ok now all the abundance values should be integers now that the non-integers have been rounded
all(abs(model_data_ab_integer_only$TotalAbundance -
        round(model_data_ab_integer_only$TotalAbundance)) < 1e-8,
    na.rm = TRUE) # TRUE - all the values in TotalAbundance are integers

# Remove old data frames
rm(model_data_ab, model_data_sr, still_non_integer_abundance, rounded_abundances)
```

# Function to robustly extract delta-method R2m, regardless of model structure returned from r.squaredGLMM()
```{r}
get_R2m_delta <- function(x) {
  if (is.matrix(x)) {
    rn <- rownames(x)
    cn <- colnames(x)
    if (!is.null(rn) && "delta" %in% rn && !is.null(cn) && "R2m" %in% cn) {
      return(as.numeric(x["delta", "R2m"]))
    }
  }
  if (is.numeric(x) && !is.null(names(x)) && "R2m" %in% names(x)) {
    # lme4-style returns named numeric vector c(R2m, R2c)
    return(as.numeric(x["R2m"]))
  }
  stop("Unexpected R2 structure returned by r.squaredGLMM.")
}
```

# Function to calculate interaction importance
```{r}

compute_interaction_importance <- function(m_interaction, m_additive, m_plu_only) {
  
  R2_int <- r.squaredGLMM(m_interaction)
  R2_add <- r.squaredGLMM(m_additive)
  R2_plu <- r.squaredGLMM(m_plu_only)

  R2m_int <- get_R2m_delta(R2_int)
  R2m_add <- get_R2m_delta(R2_add)
  R2m_plu <- get_R2m_delta(R2_plu)

  r2m_for_interaction <- max(R2m_int - R2m_add, 0) # if the r2m for the additive model > interactive model, give me 0.
  interaction_importance <- r2m_for_interaction / (R2m_plu + r2m_for_interaction)

  list(
    R2m_int_delta = R2m_int,
    R2m_add_delta = R2m_add,
    R2m_plu_delta = R2m_plu,
    r2m_for_interaction = r2m_for_interaction,
    interaction_importance = interaction_importance
  )
}
```



# Ok now all the abundance and species richness values for both model datasets are integers we can start calculating the interaction importance

#### Hymenoptera vs Coleoptera

#### Abundance 
```{r}
hym_col_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique()

# Model data
hymenoptera_coleoptera <- model_data_ab_integer_only %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.47

rm(m_additive, m_interaction, m_plu_only, results)
```

#### Species richness 
```{r}
# Model data
hymenoptera_coleoptera <- model_data_sr_integer_only %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Hymenoptera", "Coleoptera"))

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.42

rm(m_additive, m_interaction, m_plu_only, results, hymenoptera_coleoptera)
```


#### Hymenoptera vs Diptera

#### Abundance
```{r}
# Filter for studies that only compare hymenoptera and diptera
dip_hym_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Filter our abundance model dataset for those studies:
diptera_hymenoptera <- model_data_ab_integer_only %>% filter(SS %in% dip_hym_ss) %>% 
  # then filter the two orders we want to compare, as some studies will have include more than those two orders
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.76

rm(m_additive, m_interaction, m_plu_only, results)
```

#### Species richness 
```{r}
# Model data
diptera_hymenoptera <- model_data_sr_integer_only %>% filter(SS %in% dip_hym_ss) %>% filter(Order %in% c("Diptera", "Hymenoptera"))

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.61

rm(m_additive, m_interaction, m_plu_only, results, diptera_hymenoptera)
```

#### Hymenoptera vs Hemiptera

#### Abundance
```{r}
#Extract the studies and create the model data
hemi_hym_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hemiptera_hymenoptera <- model_data_ab_integer_only %>% filter(SS %in% hemi_hym_ss) %>% filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% droplevels()

# There are only 4 observations for each order in pasture
table(hemiptera_hymenoptera$Higher_taxon, hemiptera_hymenoptera$Predominant_land_use) 

# Lets remove pasture as there are too few rows to permit robust modelling
hemiptera_hymenoptera <- hemiptera_hymenoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.73

rm(m_additive, m_interaction, m_plu_only, results)
```

#### Species richness 
```{r}
#Extract the studies and create the model data
hemi_hym_ss <-  model_data_sr_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hemiptera_hymenoptera <- model_data_sr_integer_only %>% filter(SS %in% hemi_hym_ss) %>% filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% droplevels()

# Lets remove pasture as there are too few rows to permit robust modelling
hemiptera_hymenoptera <- hemiptera_hymenoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.69

rm(m_additive, m_interaction, m_plu_only, results, hemiptera_hymenoptera)
```


#### Hymenoptera vs Lepidoptera 

#### Abundance

```{r}
#Extract the studies and create the model data
lep_hym_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Lepidoptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
lepidoptera_hymenoptera <- model_data_ab_integer_only %>% filter(SS %in% lep_hym_ss) %>% filter(Order %in% c("Lepidoptera", "Hymenoptera")) %>% droplevels()

# There are only 1 observations for each order in plantation, 7 in urban might be ok
table(lepidoptera_hymenoptera$Higher_taxon, lepidoptera_hymenoptera$Predominant_land_use) 

# Lets remove plantation as there are too few rows to permit robust modelling
lepidoptera_hymenoptera <- lepidoptera_hymenoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0

rm(m_additive, m_interaction, m_plu_only, results)
```

#### Species richness 
```{r}
# Model data
lepidoptera_hymenoptera <- model_data_sr_integer_only %>% filter(SS %in% lep_hym_ss) %>% filter(Order %in% c("Lepidoptera", "Hymenoptera")) %>% droplevels()

# Lets remove pasture as there are too few rows to permit robust modelling
lepidoptera_hymenoptera <- lepidoptera_hymenoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.53

rm(m_additive, m_interaction, m_plu_only, results, lepidoptera_hymenoptera)
```

#### Coleoptera vs Diptera 

#### Abundance

```{r}
#Extract the studies and create the model data
col_dip_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Diptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
coleoptera_diptera <- model_data_ab_integer_only %>% filter(SS %in% col_dip_ss) %>% filter(Order %in% c("Coleoptera", "Diptera")) %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.76

rm(m_additive, m_interaction, m_plu_only, results)
```

#### Species richness 
```{r}
# Model data
coleoptera_diptera <- model_data_sr_integer_only %>% filter(SS %in% col_dip_ss) %>% filter(Order %in% c("Coleoptera", "Diptera")) %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.97

rm(m_additive, m_interaction, m_plu_only, results, coleoptera_diptera)
```


#### Coleoptera vs Hemiptera

#### Abundance

```{r}
#Extract the studies and create the model data
col_hemi_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Hemiptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
coleoptera_hemiptera <- model_data_ab_integer_only %>% filter(SS %in% col_hemi_ss) %>% filter(Order %in% c("Coleoptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data (n = 4)
table(coleoptera_hemiptera$Higher_taxon, coleoptera_hemiptera$Predominant_land_use) 
coleoptera_hemiptera <- coleoptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.92

rm(m_additive, m_interaction, m_plu_only, results)
```

#### Species richness 
```{r}
# Model data
coleoptera_hemiptera <- model_data_sr_integer_only %>% filter(SS %in% col_hemi_ss) %>% filter(Order %in% c("Coleoptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data (n = 4)
coleoptera_hemiptera <- coleoptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.36

rm(m_additive, m_interaction, m_plu_only, results, coleoptera_hemiptera)
```


#### Coleoptera vs Lepidoptera

#### Abundance
```{r}
#Extract the studies and create the model data
col_lep_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
coleoptera_lepidoptera <- model_data_ab_integer_only %>% filter(SS %in% col_lep_ss) %>% filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% droplevels()

# Remove pasture as has too few data (n = 1)
table(coleoptera_lepidoptera$Higher_taxon, coleoptera_lepidoptera$Predominant_land_use) 
coleoptera_lepidoptera <- coleoptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.93

rm(m_additive, m_interaction, m_plu_only, results)
```

#### Species richness 

```{r}
# Model data
coleoptera_lepidoptera <- model_data_sr_integer_only %>% filter(SS %in% col_lep_ss) %>% filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data
coleoptera_lepidoptera <- coleoptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.97

rm(m_additive, m_interaction, m_plu_only, results, coleoptera_lepidoptera)
```

#### Diptera vs Hemiptera 

#### Abundance

```{r}
#Extract the studies and create the model data
dip_hemi_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hemiptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_hemiptera <- model_data_ab_integer_only %>% filter(SS %in% dip_hemi_ss) %>% filter(Order %in% c("Diptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data (n = 1)
table(diptera_hemiptera$Higher_taxon, diptera_hemiptera$Predominant_land_use) 
diptera_hemiptera <- diptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.93

rm(m_additive, m_interaction, m_plu_only, results)
```


#### Species richness 

```{r}
# Model data
diptera_hemiptera <- model_data_sr_integer_only %>% filter(SS %in% dip_hemi_ss) %>% filter(Order %in% c("Diptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data
diptera_hemiptera <- diptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.99

rm(m_additive, m_interaction, m_plu_only, results, diptera_hemiptera)
```

#### Diptera vs Lepidoptera

#### Abundance

```{r}
#Extract the studies and create the model data
dip_lep_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_lepidoptera <- model_data_ab_integer_only %>% filter(SS %in% dip_lep_ss) %>% filter(Order %in% c("Diptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data (n = 1)
table(diptera_lepidoptera$Higher_taxon, diptera_lepidoptera$Predominant_land_use) 
diptera_lepidoptera <- diptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | SS),
  family = nbinom2(),
  data = diptera_lepidoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
   (1 | SS),
  family = nbinom2(),
  data = diptera_lepidoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
   (1 | SS),
  family = nbinom2(),
  data = diptera_lepidoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.93

rm(m_additive, m_interaction, m_plu_only, results)
```


#### Species richness 

```{r}
# Model data
diptera_lepidoptera <- model_data_sr_integer_only %>% filter(SS %in% dip_lep_ss) %>% filter(Order %in% c("Diptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data
diptera_lepidoptera <- diptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
  (1 | SS),
  family = nbinom2(),
  data = diptera_lepidoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
   (1 | SS),
  family = nbinom2(),
  data = diptera_lepidoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
  (1 | SS),
  family = nbinom2(),
  data = diptera_lepidoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.58

rm(m_additive, m_interaction, m_plu_only, results, diptera_lepidoptera)
```

#### Lepidoptera vs Hemiptera 

#### Abundance

```{r}
#Extract the studies and create the model data
hemi_lep_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hemiptera_lepidoptera <- model_data_ab_integer_only %>% filter(SS %in% hemi_lep_ss) %>% filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data (n = 1)
table(hemiptera_lepidoptera$Higher_taxon, hemiptera_lepidoptera$Predominant_land_use) 
hemiptera_lepidoptera <- hemiptera_lepidoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon +
   (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  TotalAbundance ~ Predominant_land_use +
   (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.93

rm(m_additive, m_interaction, m_plu_only, results)
```

#### Species richness

```{r}
# Model data
hemiptera_lepidoptera <- model_data_sr_integer_only %>% filter(SS %in% hemi_lep_ss) %>% filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% droplevels()

# Remove pasture as has too few data
hemiptera_lepidoptera <- hemiptera_lepidoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Interaction model (includes main effects + interaction)
m_interaction <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
  (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

# Additive model (main effects only, no interaction)
m_additive <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use + Higher_taxon +
   (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

# Predominant_land_use-only model (only land use as fixed effect)
m_plu_only <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use +
  (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

results <- compute_interaction_importance(m_interaction, m_additive, m_plu_only)
results

# Interaction importance = 0.58

rm(m_additive, m_interaction, m_plu_only, results, diptera_lepidoptera)

```

