---
title: "Predicts multi-group data"
author: "Justin Isip"
date: "2023-02-17"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Exploring multi-group insect data in PREDICTS

The complete dataset contains 4838399 records. Of this 1427446 are observations of insects. There are 1851 insects observations not resolved to order level. 

In this chapter I will be targeting studies within the PREDICTS database that sample more than one insect group (i.e. two or more groups sampled simultaneously within the same study, hereafter referred to as multi-group). The number of observations for the multi-group subset is 406574, which is ~ 28 % of the insect PREDICTS database. 

## Read in the data, all that jazz
```{r, echo = FALSE, message = FALSE}
diversity <- readRDS("./data/diversity-2022-04-13-02-33-10.rds")
library(tidyverse)
library(gt)
library(gtExtras)
library(webshot2)

# Obs of all insects
insects <- diversity %>% filter(Class == "Insecta")

# Subset predicts for multi-group insects
multi <- 
diversity %>%
  filter(Class == "Insecta") %>% 
  mutate_all(na_if,"") %>%
  group_by(SS) %>% 
  mutate(
    Multiple_groups = 
       case_when(
        (length(unique(Higher_taxon))) > 1 ~ "YES",
        (length(unique(Higher_taxon))) == 1 ~ "NO")) %>% 
  # Filter the data for studies with multiple groups
  filter(Multiple_groups == "YES") %>%
  droplevels() %>%
  ungroup() 

# Obs per order (all insects)
higher_tax_all <- 
    insects %>%
    mutate_all(na_if,"") %>% 
    group_by(Higher_taxon) %>%
    summarise(n()) %>%
    rename(number="n()") %>%
    arrange(desc(number)) %>%
    filter(Higher_taxon != "NA") # drops the Na's, 1851 obs


# Obs per order (all insects)
higher_tax_multi <- 
    multi %>%
    mutate_all(na_if,"") %>% 
    group_by(Higher_taxon) %>%
    summarise(n()) %>%
    rename(number="n()") %>%
    arrange(desc(number)) %>%
    filter(Higher_taxon != "NA") # drops the Na's, 1823 obs

# Calculate % of multi in whole insects subset
nrow(multi) / nrow(insects)


```

# Plot of number of records for each order (all insects and multi)
```{r}
higher_tax_all %>%
    ggplot(aes(x = reorder(Higher_taxon, number),
    y = number)) + 
    geom_bar(stat = "identity") +
    coord_flip() + 
    xlab("Order") +
    ylab("Number")

higher_tax_multi %>%
    ggplot(aes(x = reorder(Higher_taxon, number),
    y = number)) + 
    geom_bar(stat = "identity") +
    coord_flip() + 
    xlab("Order") +
    ylab("Number of records") 
```

# Create tables of number of obs for each order (multi)
```{r}
tb_1 <- higher_tax_multi %>% slice(1:14) %>% gt() %>% 
        cols_label(Higher_taxon = "Order", number = "Number of observations") %>% 
        tab_options(table.font.names = 'Palatino linotype', column_labels.font.weight = "bold")
  

tb_2 <- higher_tax_multi %>% slice(15:27) %>% gt() %>% 
        cols_label(Higher_taxon = "Order", number = "Number of observations") %>% 
        tab_options(table.font.names = 'Palatino linotype', column_labels.font.weight = "bold")

tb_1;tb_2
```

# Calculate families, orders and number of obs
```{r}
length(unique(multi$Family)) # 417 families
length(unique(multi$Order)) # 28 order

# Number of observations for each family within each order of interest
n_family_order <- multi %>% 
                  group_by(Order, Family) %>% 
                  summarise(n()) %>% 
                  rename(number="n()") %>%
                  filter(Family != "NA") %>% 
                  filter(Order %in% c("Arachnida", "Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))
```



## Prep the ssbst dataframe to calculate site level summaries

## Load functions

## Load order fold
```{r}
add_order_fold <- function(data, folds = "default", verbose = TRUE){
  #
  # By default (and nothing else is yet coded up), this puts each row into
  # an insect order that each have a decent, number of rows of data. 
  # This allows site-level values to be obtained within
  # each of the order folds in turn.
  
  # Set up order_fold, a factor that holds the fold identity.
  data$order_fold <- rep(NA, nrow(data))
  
  # Rows are assigned to folds based on the higher taxonomy information.
  if (folds == "default"){
    data$order_fold[data$Higher_taxon == "Diptera"] <- "Diptera"
    data$order_fold[data$Higher_taxon == "Coleoptera"] <- "Coleoptera"
    data$order_fold[data$Higher_taxon == "Lepidoptera"] <- "Lepidoptera"
    data$order_fold[data$Higher_taxon == "Hymenoptera"] <- "Hymenoptera"
    data$order_fold[data$Higher_taxon == "Hemiptera"] <- "Hemiptera"
    data$order_fold[data$Higher_taxon == "Arachnida"] <- "Arachnida"
    data$order_fold[data$Higher_taxon == "Odonata" |
                    data$Higher_taxon == "Trichoptera" |
                    data$Higher_taxon == "Ephemeroptera" |
                    data$Higher_taxon == "Plecoptera" ] <- "Aquatics"
    data$order_fold[data$Phylum == "Arthropoda" & 
                      data$Higher_taxon != "Diptera" & 
                      data$Higher_taxon != "Coleoptera" &
                      data$Higher_taxon != "Lepidoptera" &
                      data$Higher_taxon != "Hymenoptera" &
                      data$Higher_taxon != "Hemiptera" &
                      data$Higher_taxon != "Arachnida" &
                      data$Higher_taxon != "Odonata" &
                      data$Higher_taxon != "Trichoptera" &
                      data$Higher_taxon != "Ephemeroptera" &
                      data$Higher_taxon != "Plecoptera"] <- "OtherArthropods"
  }else{
    stop("Only the default folds have been coded up so far!")
  }
  
  data$order_fold <- as.factor(data$order_fold)
  
  if (verbose == TRUE){
    # By default, report on the breakdown of records per fold.
    cat("Numbers of rows within each order fold are as follows:\n\n")
    print(table(data$order_fold))
  }
  
  return(data)
}
```
### Load function to calculate total abundance for each site x order_fold combination
```{r}
# Calculate total abundance for each site --------------------------------------
# Modified to clarify reporting is on site x fold combinations
get_site_abundances_order <- function(data, 
                                order_folds = FALSE,
                                verbose = TRUE){
  # This calculates site metrics for each site or, if desired, each order fold
  # within each site.
  
  # Make it a data frame to avoid having to deal with a tibble.
  data <- as.data.frame(data)
  
  # Set up a variable for uniquely identifying the site x order_fold combination
  # or, if order_folds == FALSE, the site. This is what grouping is then done
  # by.
  print(table(data$order_fold))
  
  if (order_folds == TRUE){
    data$the_order_folds <- data$order_fold
  }else{
    data$the_order_folds <- ""
  }
  
  # Just for debugging
  print(table(data$the_order_folds))
  
  # Start to calculate site-level metrics.
  sites <- data %>%
    
    # pull out only the merged diversity data
    distinct(merge_ID, .keep_all = TRUE) %>%
    
    # Re-make SSB and SSBS values since we've now dropped a bunch of values; and
    # also add SSBST which is that order x site combo. The trimws is because the
    # SSBST would otherwise end in a space if order_folds == FALSE.
    mutate(SS = paste(Source_ID, Study_number),
           SSB = paste(SS, Block),
           SSBS = paste(SSB, Site_number),
           SSBST = trimws(paste(SSBS, the_order_folds))) %>%
    
    # group by SSBST (each unique value corresponds to a unique site x fold)
    group_by(SSBST) %>%
    
    # Now add up all the abundance measurements within each site and also 
    # estimate coverage.
    mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                   sum(merged_diversity, na.rm = TRUE),
                                   # If the diversity metric type isn't 
                                   # Abundance, then leave the TotalAbundance 
                                   # measurement as NA.
                                   NA),
           
           SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                    merged_diversity,
                                    # for abundance and occurrence measurements,
                                    # count the number of unique species names 
                                    # that are present at the site 
                                    n_distinct(Best_guess_binomial[merged_diversity > 0])),
           
           coverage = get.SC2(merged_diversity)) %>%
    
    # ungroup
    ungroup() %>%
    
    # pull out unique site/fold combinations.
    distinct(SSBST, .keep_all = TRUE) %>%
    
    # now group by Study ID and fold.
    group_by(SS, the_order_folds) %>%
    
        # pull out the maximum abundance for each study/fold combination.
    mutate(MaxAbundance = max(TotalAbundance),
           MinNonZero = min(TotalAbundance[TotalAbundance > 0])) %>%
    
    # ungroup.
    ungroup() %>%
    
    # now rescale total abundance, so that within each study, the maximum is 1.
    mutate(RescaledAbundance = TotalAbundance/MaxAbundance) %>%
    
    mutate(sqrtRescaledAbundance = sqrt(RescaledAbundance)) 
  
    # If a study reports abundance as a count, add 1 to all values; otherwise
  # add 1/2 the minimum non-zero value; then log-transform to get a useful
  # logAbundance value that may be more symmetric than sqrtRescaledAbundance.
  sites$to_add <- ifelse(sites$Diversity_metric_unit %in%
                           c("individuals", "times observed"), 1,
                         sites$MinNonZero/2)
  sites$logAbundance <- log(sites$TotalAbundance + sites$to_add)
  
  
  # Remove NA values and drop levels
  sites <- as.data.frame(sites)
  sites <- droplevels(subset(sites, !is.na(sqrtRescaledAbundance)))
  
  if (verbose == TRUE){
    # Summarise structure of site-level data frame
    cat("\nThere are abundance data for:\n")
    cat(paste("   ", sum(!is.na(sites$sqrtRescaledAbundance)), "sites, from\n"))
    cat(paste("   ", 
              length(unique(sites$SS[!is.na(sites$sqrtRescaledAbundance)])), 
              "studies and\n"))
    cat(paste("   ", 
              length(unique(sites$Source_ID[!is.na(sites$sqrtRescaledAbundance)])),
              "sources.\n"))
  }
  
  # Set provenance attributes
  attr(sites, which = "diversity_extract") <- 
    attr(data, which = "diversity_extract")
  attr(sites, which = "is_merged") <- attr(data, which = "is_merged")
  attr(sites, which = "order_folds") <- order_folds 
  attr(sites, which = "when_Created") <- Sys.time()
  
  return(sites)
}

```
### Load Andys functions
```{r}
source("/Applications/PhD/Andy/data_prep_functions.R")
source("/Applications/PhD/Andy/site_comparison_functions.R")
```



### Correct for sampling effort
```{r}
if(check_sampling_effort_nas(diversity) ==
   FALSE) stop("Some studies have NA for sampling effort in only some sites!")

diversity <- correct_for_sampling_effort(diversity)
print(summarise_diversity(diversity))
```

### Merge sites
```{r}
merged <- merge_sites(diversity, verbose = TRUE) 
```

### Filter merged sites for groups of interest
```{r}
insects <-
  merged %>%
  filter(Class == "Insecta") %>% # 
  group_by(SS) %>% # group by SS
  mutate(
    Multiple_groups = 
      case_when(
        (length(unique(Higher_taxon))) > 1 ~ "YES",
        (length(unique(Higher_taxon))) == 1 ~ "NO")) %>% # for each study create a new column called multiple groups = YES/NO - 
  # YES if there's more than one unique higher taxon within that study, NO if not
  # Filter the data for studies with multiple groups
  filter(Multiple_groups == "YES") %>% # filter for only the rows for studies (YES) with multiple groups
  ungroup() %>% 
  droplevels() 
```

### Add_order_fold to allow taxonomic cross-validation and split insects into exclusive groups
```{r}
insects <- add_order_fold(insects, verbose = TRUE) # I've sense checked we don't lose rows 
```

### Calculate site level diversity metrics using get_site_abundances_order and order_fold for each site x order combination
```{r}
insects_ssbst <- get_site_abundances_order(insects, 
                                  order_folds = TRUE, verbose = TRUE)


table(insects_ssbst$order_fold) # number of rows for each site x order combination
```

### Calculate site level summary data
```{r}
# Prep the dataframe
insects_ssbst <- 
  insects_ssbst %>%
  rename(Predominant_land_use = Predominant_habitat) %>%
  filter(order_fold %in% c("Arachnida", "Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera")) %>%
  filter(Predominant_land_use != "Secondary vegetation (indeterminate age)") %>%
  filter(Predominant_land_use != "Cannot decide") %>%
    mutate(
    Predominant_land_use = recode_factor(Predominant_land_use, 
    "Primary forest" = "Primary", 
    "Primary non-forest" = "Primary"))


# Number of sites per order
n_sites_order <- insects_ssbst %>% 
                  group_by(order_fold) %>% 
                  summarise(n())

# Number of sites per land use per order
n_land_use_order <- insects_ssbst %>% 
                    group_by(order_fold) %>% 
                    count(Predominant_habitat)

```

### Build a map of sites in the multi group analysis
```{r}

# Load packages
library(rworldmap)
library(rworldxtra)

# Pull unique site coordinates
data_coordinates <- model_data_ab %>%
  dplyr::select(Longitude, Latitude, order_fold) %>%
  unique()

# build base map for fertiliser/climate plot
get_basemap <- function(){
  
  # download full basemap
  base_map <- getMap(resolution = "high")
  
  # convert to correction projection
  proj4string(base_map) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
  
  # return basemap
  return(base_map)
}

# bring in basemap for underlying surface
base_map <- get_basemap() %>%
  fortify()

# build map for distribution of sites
site_distribution <- data_coordinates %>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = base_map, fill = "lightgrey") +
  geom_point(aes(x = Longitude, y = Latitude, colour = order_fold, fill = order_fold), alpha = 0.3, size = 2) +
  coord_map(projection = "mollweide") +
  theme(axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        axis.line = element_blank(),
        text = element_text(size = 13),
        panel.grid = element_blank(), panel.background = element_rect(fill = "grey98"),
        strip.text.x = element_text(size = 14)) +
        theme(legend.position = "none") +
        scale_colour_manual(values=group_colours) 


### Make the legend
### Make a new plot, extract just the legend from it,so I can combine with my other nicer plots
legend <- ggplot(model_data_ab, aes(x = Predominant_land_use, fill = order_fold)) + 
  scale_fill_manual(values=group_colours)+
  guides(fill=guide_legend(title = "Order", title.position = "top", nrow=1, byrow=TRUE))+
  theme(legend.position = "bottom",
        legend.text = element_text( face = "bold", size = 8),
        legend.title = element_text(face = "bold", size = 9))+
  geom_bar()

 legend 
  
```

```{r}
                  
```






