---
title: "Untitled"
output: html_document
date: "2025-04-07"
---
#### Clear the work space

```{r}
rm(list=ls())
```

# Read in the packages 
```{r}
library(lme4)
library(readxl)
library(tidyverse)
library(stringr)
library(car)
library(qs)
library(lmerTest)
library(merTools) 
library(performance)
```

# Load in model plot function
```{r}
## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(2,2))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}
```

# Read in Logghe and tidy
```{r}
# Upload logghe et al
logghe <- read_excel("data/Body_Size_data.xlsx")

# keep only first 8 cols
logghe <- logghe %>% dplyr::select(1:8)

# Make first row the header
colnames(logghe) <- as.character(logghe[1, ])
logghe <- logghe[-1, ] # drop first row

# Create a genus and species column
logghe <- logghe %>%
  separate(scientificName, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop")

# Combine genus and species for merging
logghe <- logghe %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Reorder columns
logghe <- logghe %>% dplyr::select(order, family, Genus, Species, Best_guess_binomial, Trait_Value) %>% 
  rename(Family = family) %>% 
  rename(Order = order)

logghe$Trait_Value <-  as.numeric(logghe$Trait_Value)
```

# Read and tidy Geoffrey Su's ant trait data
```{r}
# Load the data
trait_ants <- read_excel("data/SuFormacidaeBodyLength-2014-05-23.xlsx")

# Rename body length average column
trait_ants <- trait_ants %>% rename(avg_body_length_mm = "Body.Length..mm...worker..AVG")

# Lets remove those rows with NAs for body length
trait_ants <- trait_ants %>%
  filter(!is.na(avg_body_length_mm) & avg_body_length_mm != 0) # 559 rows with no body length data, 2 rows with BL = 0

# Combine genus and species  
trait_ants <- trait_ants %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Simplify ants
trait_ants <- trait_ants %>% dplyr::select(Genus, Best_guess_binomial, avg_body_length_mm)

# Tidy
trait_ants <- trait_ants %>%
  mutate(
    Order = "Hymenoptera",
    Family = "Formicidae"
  ) %>% 
  rename(Trait_Value = avg_body_length_mm) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop") %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)
# ants ready
```

# Read and tidy Melanie Edgars beetle trait data
```{r}
# Load Melanies beetle trait data
melanie_beetles <- read_excel("data/EdgarBeetleLength-2014-01-09.xlsx")

# Combine genus and species for merging
melanie_beetles <- melanie_beetles %>% mutate(Best_guess_binomial = str_c(GENUS, SPECIES, sep = " "))

# Calculate average b.length
melanie_beetles <- melanie_beetles %>%
  separate(B.LENGTH, into = c("b_min", "b_max"), sep = "-", fill = "right", remove = FALSE) %>% # fill = "right" allows single values to stay in b_min while b_max becomes NA.
  mutate(
    b_min = as.numeric(trimws(b_min)), # trimws() removes any leading/trailing spaces 
    b_max = as.numeric(trimws(b_max)),
    b.length_avg = if_else(
      is.na(b_max),
      b_min,  # If no range, use the min value
      (b_min + b_max) / 2  # Else if range, compute the average
    ))

# Simplify beetles
melanie_beetles <- melanie_beetles %>% 
  mutate(Order = "Coleoptera") %>% 
  dplyr::select(Order, FAMILY, GENUS, SPECIES, Best_guess_binomial, b.length_avg) %>% 
  rename(Family = FAMILY) %>% 
  rename(Genus = GENUS) %>% 
  rename(Species = SPECIES) %>% 
  rename(Trait_Value = b.length_avg)
```

# Read and tidy Rachel Tans bee trait data
```{r}
# Load Rachel bee trait data
rachel_bee <- read_excel("data/BeeTraits_P2_RachelTan.xlsx")

# Create a genus and species column
rachel_bee <- rachel_bee %>%
  separate(CorrectedName, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop")

# Filter for columns of interest
rachel_bee <- rachel_bee %>% dplyr::select(HigherTaxon, Family, Genus, Species, CorrectedName, MidRangeBodyLength) %>% drop_na() # couple sp. with NA for body length

# Remove rows identified as "sp." or inside brackets e.g., (Dialictus) or (Psithyrus) which are subgenera
rachel_bee <- rachel_bee %>%
  filter(
    !str_detect(Species, "sp\\.") &         # remove rows with "sp."
    !str_detect(Species, "\\(.*\\)")        # remove rows with anything in brackets
  )

# Rename for merging
rachel_bee <- rachel_bee %>% 
  rename(Best_guess_binomial = CorrectedName) %>% 
  rename(Order = HigherTaxon) %>% 
  rename(Trait_Value = MidRangeBodyLength)

```

# Merge Logghe, Geoffrey, Melanies and Rachels trait datasets to have an overall body size trait dataset
```{r}
# merge all trait datasets
all_traits <- bind_rows(logghe, melanie_beetles, rachel_bee, trait_ants)

# There are duplicate species entries
all_traits[duplicated(all_traits["Best_guess_binomial"]) | 
                                  duplicated(all_traits["Best_guess_binomial"], fromLast = TRUE), ] %>%  View()

# Sense check - there are 634 duplicates
all_traits[duplicated(all_traits["Best_guess_binomial"]),] 

# For those species with duplicate entries, calculate an average Trait_Value which is the mean of the duplicate entries
all_traits_clean <- all_traits %>%
  group_by(Best_guess_binomial) %>%
  summarise(
    Trait_Value = mean(Trait_Value, na.rm = TRUE),  # average Trait_Value
    across(.cols = -Trait_Value, .fns = ~ first(.x)),  # keep first value for other columns
    .groups = "drop"
  ) %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)

nrow(all_traits) - nrow(all_traits_clean) # 634

rm(all_traits, logghe, melanie_beetles, rachel_bee, trait_ants)
```

# Read in PREDICTS data 
```{r}
# All insect data effort-corrected and merged
insects <-  qread("/Users/jusi/Desktop/multi-groups/data/predicts-prepared-whole-database.qs")
```

# Clean and tidy PREDICTS data then merge trait dataset based on taxon resolution
```{r}
# Get relevant taxon cols for all insects
taxon_insects <- insects %>% dplyr::select(SS, SSBS, Order, Family, Genus, Species, Best_guess_binomial, Merged_diversity)

# Get rid of rows which aren't at family level, i.e. at superfamily or higher
#taxon_insects <- taxon_insects %>% filter(Family != "") # 58403 rows - Keep order rows for now!

# Get rid of rows which aren't resolved to order level
taxon_insects <- taxon_insects %>% filter(Order != "") # 1056 rows not resolved to order

# Recreate genus and species column as sometimes BGB is present in PREDICTS but no info for genus or sp.
taxon_insects <- taxon_insects %>%
  mutate(Best_guess_binomial = as.character(Best_guess_binomial)) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "merge", fill = "right")

# Get species trait values for merging
sp_value <- all_traits_clean %>% dplyr::select(Best_guess_binomial, Trait_Value) 

# First merge Logghe species level data
taxon_insects <- left_join(taxon_insects, sp_value, by = "Best_guess_binomial")

rm(sp_value)

# Calculate average body length for genera
genus_value <- all_traits_clean %>%
  group_by(Genus) %>%
  summarise(genus_avg_body_length = mean(Trait_Value, na.rm = TRUE)) # 2381 genus data

# Merge based on genus where best_guess_binomial is empty
taxon_insects <- taxon_insects %>%
  left_join(genus_value, by = "Genus", suffix = c("", "_genus")) 

rm(genus_value)

# Average body length per family based on species
family_value <- all_traits_clean %>%
  group_by(Family) %>%
  summarise(family_avg_body_length = mean(Trait_Value, na.rm = TRUE))

# Merge based on family where genus and best_guess_binomial is empty
taxon_insects <- taxon_insects %>%
  left_join(family_value, by = "Family", suffix = c("", "_family")) 

rm(family_value)

# priority-based fallback assignment where final body length values are selected based on the first non-missing (or non-blank) column in a defined order.
taxon_insects <- taxon_insects %>%
  mutate(final_body_length = coalesce(
    Trait_Value,
    genus_avg_body_length,
    family_avg_body_length
  ))

# How many rows have data based on each level of taxonomic resolution
taxon_insects %>%
  summarise(across(
    c(Trait_Value,
    genus_avg_body_length,
    family_avg_body_length,
    final_body_length),
    list(
      non_NAs = ~sum(!is.na(.))
    )
  ))
```

# Tidy Rainford
```{r}
# Upload 
rainford <- read_excel("data/rainford_body_size.xlsx")

# Make first row the header
colnames(rainford) <- as.character(rainford[1, ])
rainford <- rainford[-1, ] # drop first row

# Extract cols of interest
rainford <- rainford %>% dplyr::select("Order", "Family", "Min (mm)", "Max (mm)") %>% rename(min_bl = "Min (mm)") %>% rename(max_bl = "Max (mm)")

# Drop rows with all NAs
rainford <- rainford %>%
  filter(!if_all(everything(), is.na))

# Convert cols to numeric
rainford$min_bl <- as.numeric(rainford$min_bl)
rainford$max_bl <- as.numeric(rainford$max_bl)

# Drop rows only to Order level
rainford <- rainford %>%
  filter(!is.na(Family))

# Calculate average body length
rainford <- rainford %>%
  mutate(family_avg = (min_bl + max_bl) / 2)
         
# Simplify for merging
rainford <- rainford %>% 
  dplyr::select(c(-min_bl, -max_bl))
```

# Merge Rainford 

```{r}
# Step 1: Filter rows in taxon_insects where there is no body size value at the species, genus or family level and therefore we can merge a family body size value from Rainford
needs_family_avg <- taxon_insects %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length))

# Step 2: Join family_avg from Rainford based on Family
needs_family_avg <- needs_family_avg %>%
  left_join(rainford, dplyr::select(Family, family_avg), by = "Family")

# Step 3: Combine updated rows (need_family_avg) back into original data (taxon_insects)
# First, remove the rows you updated from taxon_insects
remaining_rows <- taxon_insects %>%
  filter(!(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length)))

# sense check remaining rows should be taxon_insects - needs_family_avg
nrow(taxon_insects) - nrow(needs_family_avg)

# Bind them together
taxon_insects_updated <- bind_rows(remaining_rows, needs_family_avg)

# How many rows still without a family average
sum(!is.na(taxon_insects_updated$family_avg)) # 134120

# How many rows have data based on each level of taxonomic resolution
taxon_insects_updated %>%
  summarise(across(
    c(Trait_Value,
    genus_avg_body_length,
    family_avg_body_length,
    final_body_length,
    family_avg),
    list(
      non_NAs = ~sum(!is.na(.))
    )
  ))

rm(needs_family_avg, remaining_rows, taxon_insects)
```

# Now that we have merged all the trait data, coalesce the family averages from Rainford into the final_body_length column

```{r}
# Replace missing values in final_body_length with the corresponding family_avg
taxon_insects_updated <- taxon_insects_updated %>%
  mutate(final_body_length = coalesce(final_body_length, family_avg)) # coalesce(a,b) returns a if its not NA otherwise it returns b

# How many rows are still NA for body size - 71437
taxon_insects_updated %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length) & is.na(final_body_length) & is.na(family_avg)) %>%
  nrow()

# Have a look at the rows
taxon_insects_updated %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length) & is.na(final_body_length) & is.na(family_avg)) %>% 
  View()

# Filter out rows which still don't have a body size value
all_insects_size <- taxon_insects_updated %>%
  filter(!is.na(final_body_length)) %>% # 1195212
  droplevels()

```

# Some rows have NA for Order

```{r}
# Have a look at an example
all_insects_size %>% filter(SS == "YL1_2023__Simba 1") %>% View()

# Coalesce order column based on other cols
all_insects_size <- all_insects_size %>%
  mutate(
    Order = coalesce(
      na_if(as.character(Order), ""),
      na_if(as.character(Order.x), ""),
      na_if(as.character(Order.y), "")
    )
  )

# Remove redundant cols
all_insects_size <- all_insects_size %>%
  dplyr::select(-Order.x, -Order.y)

```

# First split sites based on their latitude to tropical or temperate

```{r}
# Filter overall predicts for latitude
latitude <- insects %>% dplyr::select(SSBS, Latitude) 

# There are duplicated sites
any(duplicated(latitude$SSBS))

# Lets get a list of distinct sites
latitude <- latitude %>% distinct(SSBS, .keep_all = TRUE)

# Merge latitude values into size data frame for all of the sites
all_insects_size <- all_insects_size %>%
  left_join(latitude, by = "SSBS")

# Add a new column classifying each site into tropical or temperate
all_insects_size$climate_zone <- ifelse(all_insects_size$Latitude >= -23.4368 & all_insects_size$Latitude <= 23.4368, 
                          "tropical", 
                          "temperate")
```

# Read in allometry coefficients
```{r}
coefficients <- read_excel("data/allometry_coefficients.xlsx")
```


# Merge allometry coefficients to size data
```{r}
# Step 1: Join by order and climate_zone
all_insects_size <- all_insects_size %>%
  left_join(coefficients, by = c("Order", "climate_zone"))

# Step 2: Join fallback values for rows with missing a/b
# First extract fallback values for All_insects
fallback <- coefficients %>%
  filter(Order == "All_insects") %>%
  dplyr::select(climate_zone, a_fallback = a, b_fallback = b)

# Step 3: Join fallback by climate_zone
all_insects_size <- all_insects_size %>%
  left_join(fallback, by = "climate_zone") %>%
  mutate(
    a = if_else(is.na(a), a_fallback, a),
    b = if_else(is.na(b), b_fallback, b)
  ) %>%
  dplyr::select(-a_fallback, -b_fallback)

# Have a look
all_insects_size %>% dplyr::select(SSBS, Order, final_body_length, Latitude, climate_zone, a, b) %>% unique() %>%  View()
```

# Calculate mass in mg using allometric specific coefficients for each order
```{r}
all_insects_size <- all_insects_size %>%
  mutate(
    log10_mass = a + b * log10(final_body_length),
    mass_mg = 10^log10_mass
  )

summary(all_insects_size$final_body_length) # no zero values for final_body_length

# Have a look
all_insects_size %>% dplyr::select(SSBS, Order, final_body_length, Latitude, climate_zone, a, b, log10_mass, mass_mg) %>% View()
```


# Analysis looking at ALL of the insects 
# first lets calculate one weighted mean length value per site
```{r}
# Create a separate data frame for the overall site level analysis as we'll use all_insects_size to calculate order by order
overall_site_size <- all_insects_size

# Calculate abundance weighted log mass
overall_site_size <- overall_site_size %>%
  group_by(SSBS) %>%
  mutate(
    relativeAb_logMass = (log10_mass * Merged_diversity) / sum(Merged_diversity, na.rm = TRUE)
  ) %>%
  ungroup()

# Have a look at the worked exampled
overall_site_size %>% filter(SSBS == "AD1_2009__Vergara 1  6") %>% View()

# Calculate total log mass per site
logMass_summary <- overall_site_size %>%
  group_by(SSBS) %>%
  summarise(
    total_div = sum(Merged_diversity, na.rm = TRUE),
    relAb_sum = sum(relativeAb_logMass, na.rm = TRUE),
    logMass_total = ifelse(total_div > 0, relAb_sum + log(total_div), NA_real_) # this addresses -Inf values due to sites with 0 merged diversity, i.e. taking a log of 0.
  ) %>%
  dplyr::select(SSBS, logMass_total)

# Have a look at the worked exampled
logMass_summary %>% filter(SSBS == "AD1_2009__Vergara 1  6") %>% View()

# View rows where logMass_total is 0, NA, or blank
logMass_summary %>%
  filter(is.na(logMass_total) | logMass_total == 0 | logMass_total == "") %>% View()

rm(coefficients, fallback, latitude)
```

# Merge with PREDICTS insects
```{r}
# Use distinct() to keep only the first appearance of each site within the overall PREDICTS div extract for insects
insects_single <- insects %>%
  distinct(SSBS, .keep_all = TRUE)

# merge logMass_total values
final_size <- insects_single %>%
  left_join(logMass_summary, by = "SSBS") %>% 
  droplevels()

# filter out sites with no logMass_total 
final_size <- final_size %>%
  filter(
    !is.na(logMass_total),         # remove NA
    !is.nan(logMass_total),        # remove NaN
    logMass_total != 0,            # remove 0
    logMass_total != ""            # remove empty strings
  )

# 12918 sites in total

rm(insects_single)


table(all_insects_size$Order)
```

# Load additional explanatory variables
```{r}
# Read in the data
add_var <-  qread("/Users/jusi/Desktop/multi-groups/data/abundance-model-data-prepared 2.qs")

# Subset for cols of interest
add_var <- add_var %>% dplyr::select(SSBS, t30, Landscape_Natural_Habitat) 

# Creating log_t30_scaled
add_var$log_t30 <- log1p(add_var$t30)
add_var$log_t30_scaled <- scale(add_var$log_t30)

# Creating nat_hab_scaled
add_var$nathab_scaled <- scale(add_var$Landscape_Natural_Habitat)

# merge logMass_total values
final_size <- final_size %>%
  left_join(add_var, by = "SSBS") %>% 
  droplevels()

# have a look
final_size %>% dplyr::select(SSBS, t30, log_t30, log_t30_scaled, Landscape_Natural_Habitat, nathab_scaled) %>% View()


final_size %>%
  summarise(
    t30_non_NA = sum(!is.na(t30)),
    log_t30_non_NA = sum(!is.na(log_t30)),
    log_t30_scaled_non_NA = sum(!is.na(log_t30_scaled)),
    Landscape_Natural_Habitat_non_NA = sum(!is.na(Landscape_Natural_Habitat)),
    nathab_scaled_non_NA = sum(!is.na(nathab_scaled))
  )
# 12036/12918 for log_t30 and 12033/12918 for nathab scaled.

```

# Start the modelling of all of the insects

```{r}
# This PREDICTS data file is diversity extract of all insect data which has been merged and effort corrected.
# It has also had a set of sampling methods updated, see PreparePREDICTSData.R
# It has not had accidentally sampled larvae removed as this analysis does not focus on multiple groups.

# Rename predominant habitat since its not really habitat we're looking at
final_size <- rename(final_size,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
final_size <- final_size %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(final_size$Predominant_land_use, final_size$Use_intensity)
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
final_size <- final_size %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# Have a look at sample sizes
table(final_size$LUI)

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 61 sites, merge with light use
final_size <- final_size %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Have a look at sample sizes
table(final_size$LUI)

# Set the reference levels for LUI
final_size <- final_size %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )

# Have a look at sample sizes
table(final_size$LUI)

final_size %>% dplyr::select(SS, SSBS, Order, Predominant_land_use, Use_intensity, LUI, logMass_total) %>% View()
```

# Create NatOrHumDom

```{r}
final_size <- final_size %>%
  mutate(
    NatOrHumDom = case_when(
      is.na(Predominant_land_use) ~ NA_character_,
      Predominant_land_use %in% c("Primary", "Young_secondary_vegetation", "Intermediate secondary vegetation", "Mature secondary vegetation") ~ "natural",
      TRUE ~ "human_dominated"
    )
  )
# Have a look
final_size %>% dplyr::select(Predominant_land_use, NatOrHumDom) %>% View()

```

# Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(final_size[ , c("Predominant_land_use", "Use_intensity")])
```

# Complete cases and model data

```{r}
# Lets have a look at rows with an NA for our variables of interest
final_size %>%
  filter(if_any(c(Predominant_land_use, Use_intensity, LUI, log_t30_scaled, nathab_scaled, NatOrHumDom), is.na)) %>% 
  dplyr::select(SS, SSBS,Predominant_land_use, Use_intensity, LUI, log_t30_scaled, nathab_scaled, NatOrHumDom) %>% 
  View()

final_size %>%
  filter(!if_any(c(Predominant_land_use, Use_intensity, LUI, log_t30_scaled, nathab_scaled, NatOrHumDom), is.na)) %>% 
  dplyr::select(SS, SSBS,Predominant_land_use, Use_intensity, LUI, log_t30_scaled, nathab_scaled, NatOrHumDom) %>% 
  View()

# drop all rows with NAs for land use and use intensity and drop empty levels
model_data_size <- drop_na(final_size, Predominant_land_use, Use_intensity) %>% droplevels() 

# drop all rows for additional variables for modelling
model_data_t30 <- drop_na(final_size, Predominant_land_use, Use_intensity, log_t30_scaled, nathab_scaled) %>% droplevels()

# check distributions
hist(model_data_size$logMass_total)
hist(model_data_t30$log_t30_scaled)
hist(model_data_t30$nathab_scaled)

# Have a look at sample sizes
table(model_data_size$Predominant_land_use, model_data_size$Use_intensity)
table(model_data_size$LUI)

length(unique(model_data_size$SS)) # 444 studies 
length(unique(model_data_size$SSBS)) # 10520 sites

length(unique(model_data_t30$SS)) # 402 studies 
length(unique(model_data_t30$SSBS)) # 9773 sites
```

# Modelling size ~ LU and creating the LU plot for all insects

```{r}
# maximal model
m1 <- lmer(logMass_total ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_size)

# remove block
m2 <-lmer(logMass_total ~ Predominant_land_use + (1 | Source_ID) + (1 | SS),
    data = model_data_size)

# remove study
m3 <-lmer(logMass_total ~ Predominant_land_use + (1 | Source_ID),
    data = model_data_size)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log10(biomass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
  #ggsave(filename = "figures/biomass_LU.png",  plot = p, width = 6, height = 4, dpi = 300)

  
rm(eff, m1, m2, m3, p)     
```

# Modelling size ~ LUI (LU + UI) and creating the LUI plot for all insects

```{r}
# maximal model
m1 <- lmer(logMass_total ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_size)

# remove block
m2 <-lmer(logMass_total ~ LUI + (1 | Source_ID) + (1 | SS),
    data = model_data_size)

# remove study
m3 <-lmer(logMass_total ~ LUI + (1 | Source_ID),
    data = model_data_size)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary_minimal",
                            "LUIPrimary_Light_use" = "Primary_light",
                                "LUIPrimary_Intense_use" = "Primary_intense",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV_minimal",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV_light",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV_intense",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV_minimal",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV_high",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV_minimal",
                            "LUIMature_secondary_vegetation_High_use" = "MSV_high",
                            "LUIPlantation_forest_Minimal_use" = "Plantation_minimal",
                            "LUIPlantation_forest_Light_use" = "Plantation_light",
                            "LUIPlantation_forest_Intense_use" = "Plantation_intense",
                            "LUIPasture_Minimal_use" = "Pasture_minimal",
                            "LUIPasture_Light_use" = "Pasture_light",
                            "LUIPasture_Intense_use" = "Pasture_intense",
                            "LUICropland_Minimal_use" = "Cropland_minimal",
                            "LUICropland_Light_use" = "Cropland_light",
                            "LUICropland_Intense_use" = "Cropland_intense",
                            "LUIUrban_Minimal_use" = "Urban_minimal",
                            "LUIUrban_Light_use" = "Urban_light",
                            "LUIUrban_Intense_use" = "Urban_intense")
                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    p <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3)+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log10(biomass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
       ggsave(filename = "figures/biomass_LUI.png",  plot = p, width = 8, height = 7, dpi = 300)
     
rm(eff, m1, m2, m3, p)     
```

# Modelling size ~ LUI + logt30 + nathab + NatOHumDom

```{r}
# nathab_scaled = amount of natural habitat in the surrounding area
# log_t30_scaled = log(number of years since the broad landscape around the site first got 30% converted to human use)
# NatOrHumDom = natural or human dominated, natural = primary or secondary, all others human dominated.


# maximal model
m1 <- lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled + log_t30_scaled:NatOrHumDom + (1 | Source_ID) + (1 | SS) + (1 | SSB), data = model_data_t30)

# remove block
m2 <- lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled + log_t30_scaled:NatOrHumDom + (1 | Source_ID) + (1 | SS), 
           data = model_data_t30)

# remove study
m3 <-lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled + log_t30_scaled:NatOrHumDom + (1 | Source_ID),
    data = model_data_t30)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary_minimal_use",
                                "LUIPrimary_High_use" = "Primary_high_use",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "Young_secondary_vegetation_Minimal_use",
                            "LUIYoung_secondary_vegetation_High_use" = "Young_secondary_vegetation_High_use",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "Intermediate_secondary_vegetation_Minimal_use",
                            "LUIIntermediate_secondary_vegetation_High_use" = "Intermediate_secondary_vegetation_High_use",
                            "LUIMature_secondary_vegetation_Minimal_use" = "Mature_secondary_vegetation_Minimal_use",
                            "LUIMature_secondary_vegetation_High_use" = "Mature_secondary_vegetation_High_use",
                            "LUIPlantation_forest_Minimal_use" = "Plantation_forest_Minimal_use",
                            "LUIPlantation_forest_High_use" = "Plantation_forest_High_use",
                            "LUIPasture_Minimal_use" = "Pasture_Minimal_use",
                            "LUIPasture_High_use" = "Pasture_High_use",
                            "LUICropland_Minimal_use" = "Cropland_Minimal_use",
                            "LUICropland_High_use" = "Cropland_High_use")
                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3)+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 12.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("Total log mass difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
rm(eff, m1, m2, m3, final_size, logMass_summary, model_data_size, overall_site_size)     
```

# Model simplification
```{r}
# maximal model
m1 <- lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled + log_t30_scaled:NatOrHumDom + (1 | Source_ID) + (1 | SS) + (1 | SSB), data = model_data_t30)

# Model without interaction
m2 <- lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled +
             (1 | Source_ID) + (1 | SS) + (1 | SSB),
           data = model_data_t30)

summary(m1)

anova(m1, m2)  # Compare full model (with interaction) to simpler model

# p-value is > 0.05, the interaction is not contributing significantly → proceed with m2.

# Step 2: Remove log_t30_scaled 
m3 <- lmer(logMass_total ~ LUI + nathab_scaled +
             (1 | Source_ID) + (1 | SS) + (1 | SSB),
           data = model_data_t30)

anova(m2, m3)

# Step 3: Remove nathab_scaled 
m4 <- lmer(logMass_total ~ LUI +
             (1 | Source_ID) + (1 | SS) + (1 | SSB),
           data = model_data_t30)

anova(m3, m4) # only just not significant

AIC(m1, m2, m3, m4)

summary(m4)

# Stepwise backward selection for fixed effects
step(m1)

```

# Have a look at model residuals
```{r}

residuals_m4 <- resid(m4)

# 2. Plot residuals vs fitted values
plot(fitted(m4), residuals_m4,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red")

hist(residuals_m4, breaks = 50, main = "Histogram of Residuals",
     xlab = "Residuals", col = "lightblue")

qqnorm(residuals_m4)
qqline(residuals_m4, col = "red")

library(performance)
check_model(m4)

rm(add_var, m1, m2, m3, m4)
```

# Start with biomass ~ LU * Order analysis

```{r}
# Create a separate data frame
orders_size <- all_insects_size 

# Lets look at the number of sites per order
orders_size %>%
  group_by(Order) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Orders to keep and assign all others to "other_orders"
target_orders <- c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera", "Hemiptera")

orders_size <- orders_size %>%
  mutate(Order = case_when(
    Order %in% target_orders ~ Order,
    TRUE ~ "Other_orders"
  ))

unique(orders_size$Order)
```

# the big five orders - up to five size values per site
```{r}
# Calculate abundance weighted log mass per Order
orders_size <- orders_size %>%
  group_by(SSBS, Order) %>%
  mutate(
    relativeAb_logMass = (log10_mass * Merged_diversity) / sum(Merged_diversity, na.rm = TRUE)
  ) %>%
  ungroup()

# Have a look at the worked exampled
orders_size %>% filter(SSBS == "AD1_2009__Vergara 1  6") %>% View()

# Calculate total log mass per ORDER within each site
logMass_summary <- orders_size %>%
  group_by(SSBS, Order) %>%
  summarise(
    total_div = sum(Merged_diversity, na.rm = TRUE),
    relAb_sum = sum(relativeAb_logMass, na.rm = TRUE),
    logMass_total = ifelse(total_div > 0, relAb_sum + log10(total_div), NA_real_) # this addresses -Inf values due to sites with 0 merged diversity, i.e. taking a log of 0.
  ) %>%
  dplyr::select(SSBS, Order, logMass_total)

# Have a look at the worked exampled
logMass_summary %>% filter(SSBS == "AD1_2009__Vergara 1  6") %>% View()

# View rows where logMass_total is 0, NA, or blank
logMass_summary %>%
  filter(is.na(logMass_total) | logMass_total == 0 | logMass_total == "") %>% View() # NAs are for Orders with 0 abundance across all sampled taxa for that order, so results in an NA rather than a 0

# Have a look 
orders_size %>% filter(SSBS == "AD1_2002__Vazquez 1 1 2") %>% View()
```

# Merge with PREDICTS but this time keep a row for each order within each site!
```{r}
# Create the other orders column so we filter for the correct number of sites
insects <- insects %>%
  mutate(Order = case_when(
    Order %in% target_orders ~ Order,
    TRUE ~ "Other_orders"
  ))

# Use distinct() to keep only the first appearance of each Order within each site
ssbs_by_order <- insects %>%
  distinct(SSBS, Order, .keep_all = TRUE) # 23,381 sites!

# merge logMass_total values
mass_by_order <- ssbs_by_order %>%
  left_join(logMass_summary, by = c("SSBS", "Order")) %>% 
  droplevels()


# filter out sites with no logMass_total
mass_by_order <- mass_by_order %>%
  filter(
    !is.na(logMass_total),         # remove NA
    !is.nan(logMass_total),        # remove NaN
    logMass_total != 0,            # remove 0
    logMass_total != ""            # remove empty strings
  )

sum(!is.na(logMass_summary$logMass_total)) # number of site x order combos with data, sensitivity check as should match number of sites in mass_by_order 

rm(ssbs_by_order)
```

# Prep model data
```{r}
# Create model data
big_five <- mass_by_order 

# Lets look at the number of sites per order
big_five %>%
  group_by(Order) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Rename predominant habitat since its not really habitat we're looking at
big_five <- rename(big_five,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
big_five <- big_five %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(big_five$Predominant_land_use, big_five$Use_intensity)
```

# Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(big_five[ , c("Predominant_land_use", "Use_intensity")])
```

# Complete cases and summary stats

```{r}
# Lets have a look at rows with an NA for land use or use intensity
big_five %>%
  filter(if_any(c(Predominant_land_use, Use_intensity), is.na)) %>% 
  dplyr::select(SS, SSBS, Predominant_land_use, logMass_total, Use_intensity) %>% 
  View()

# drop all rows with NAs for land use and use intensity and drop empty levels
big_five <- drop_na(big_five, Predominant_land_use, Use_intensity) %>% droplevels() 

# check distribution of logMass
hist(big_five$logMass_total)

# Have a look at sample sizes
table(big_five$Predominant_land_use, big_five$Use_intensity)

length(unique(big_five$SS)) # 444 studies
length(unique(big_five$SSBS)) # 10520 sites
```

# Modelling biomass ~ LU * Order

```{r}
# Set order as a factor variable for modelling
big_five <- big_five %>%
  mutate(Order = as.factor(Order))

# maximal model
m1 <- lmer(logMass_total ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = big_five)

# remove block
m2 <-lmer(logMass_total ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS),
    data = big_five)

# remove study
m3 <-lmer(logMass_total ~ Predominant_land_use + (1 | Source_ID),
    data = big_five)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)

rm(m2, m3)
```

# Coleoptera effects
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
col_eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  col_eff$term <- as.factor(col_eff$term)
  
   # pull out effects
  col_eff <- col_eff %>%
    slice(1:8) 
  
  # rename factors for ease of visualisation
  col_eff$term <- recode_factor(col_eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  col_eff <- col_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    col_eff <- col_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      col_eff[1,7] <- 0 #shift median
      col_eff[1,8] <- 0 #shift upper
      col_eff[1,9] <- 0 #shift lower
      
      colnames(col_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   col_eff <- col_eff %>% mutate(Order = "Coleoptera")

  # make the order column a factor so it can be recoded
   col_eff$Order <- as.factor(col_eff$Order)      
      
rm(m1)
```

# Diptera effects
```{r}
# Relevel model with Diptera as reference level
dip_ref <- within(big_five, Order <- relevel(Order, ref = "Diptera"))

# dip model
dip_m <- lmer(logMass_total ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = dip_ref)

summary(dip_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
dip_eff <- FEsim(dip_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  dip_eff$term <- as.factor(dip_eff$term)
  
   # pull out effects
  dip_eff <- dip_eff %>%
    slice(1:8) 
  
  # rename factors for ease of visualisation
  dip_eff$term <- recode_factor(dip_eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  dip_eff <- dip_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    dip_eff <- dip_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      dip_eff[1,7] <- 0 #shift median
      dip_eff[1,8] <- 0 #shift upper
      dip_eff[1,9] <- 0 #shift lower
      
      colnames(dip_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   dip_eff <- dip_eff %>% mutate(Order = "Diptera")

  # make the order column a factor so it can be recoded
   dip_eff$Order <- as.factor(dip_eff$Order)      
      
rm(dip_ref, dip_m)
```

# Hemiptera effects
```{r}
# Relevel model with Hemiptera as reference level
hemi_ref <- within(big_five, Order <- relevel(Order, ref = "Hemiptera"))

# dip model
hemi_m <- lmer(logMass_total ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = hemi_ref)

summary(hemi_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
hemi_eff <- FEsim(hemi_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  hemi_eff$term <- as.factor(hemi_eff$term)
  
   # pull out effects
  hemi_eff <- hemi_eff %>%
    slice(1:8) 
  
  # rename factors for ease of visualisation
  hemi_eff$term <- recode_factor(hemi_eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hemi_eff <- hemi_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    hemi_eff <- hemi_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      hemi_eff[1,7] <- 0 #shift median
      hemi_eff[1,8] <- 0 #shift upper
      hemi_eff[1,9] <- 0 #shift lower
      
      colnames(hemi_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   hemi_eff <- hemi_eff %>% mutate(Order = "Hemiptera")

  # make the order column a factor so it can be recoded
   hemi_eff$Order <- as.factor(hemi_eff$Order)      
      
rm(hemi_ref, hemi_m)
```

# Hymenoptera effects
```{r}
# Relevel model with Hymenoptera as reference level
hym_ref <- within(big_five, Order <- relevel(Order, ref = "Hymenoptera"))

# dip model
hym_m <- lmer(logMass_total ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = hym_ref)

summary(hym_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
hym_eff <- FEsim(hym_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  hym_eff$term <- as.factor(hym_eff$term)
  
   # pull out effects
  hym_eff <- hym_eff %>%
    slice(1:8) 
  
  # rename factors for ease of visualisation
  hym_eff$term <- recode_factor(hym_eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hym_eff <- hym_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    hym_eff <- hym_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      hym_eff[1,7] <- 0 #shift median
      hym_eff[1,8] <- 0 #shift upper
      hym_eff[1,9] <- 0 #shift lower
      
      colnames(hym_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   hym_eff <- hym_eff %>% mutate(Order = "Hymenoptera")

  # make the order column a factor so it can be recoded
   hym_eff$Order <- as.factor(hym_eff$Order)     
      
rm(hym_ref, hym_m)
```

# Lepidoptera effects
```{r}
# Relevel model with Lepidoptera as reference level
lep_ref <- within(big_five, Order <- relevel(Order, ref = "Lepidoptera"))

# dip model
lep_m <- lmer(logMass_total ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = lep_ref)

summary(lep_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
lep_eff <- FEsim(lep_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  lep_eff$term <- as.factor(lep_eff$term)
  
   # pull out effects
  lep_eff <- lep_eff %>%
    slice(1:8) 
  
  # rename factors for ease of visualisation
  lep_eff$term <- recode_factor(lep_eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  lep_eff <- lep_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    lep_eff <- lep_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      lep_eff[1,7] <- 0 #shift median
      lep_eff[1,8] <- 0 #shift upper
      lep_eff[1,9] <- 0 #shift lower
      
      colnames(lep_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   lep_eff <- lep_eff %>% mutate(Order = "Lepidoptera")

  # make the order column a factor so it can be recoded
   lep_eff$Order <- as.factor(lep_eff$Order)     
      
rm(lep_ref, lep_m)
```

# Other_orders effects
```{r}
# Relevel model with Other_orders as reference level
other_ref <- within(big_five, Order <- relevel(Order, ref = "Other_orders"))

# other_orders model
other_m <- lmer(logMass_total ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = other_ref)

summary(other_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
other_eff <- FEsim(other_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  other_eff$term <- as.factor(other_eff$term)
  
   # pull out effects
  other_eff <- other_eff %>%
    slice(1:8) 
  
  # rename factors for ease of visualisation
  other_eff$term <- recode_factor(other_eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  other_eff <- other_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    other_eff <- other_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      other_eff[1,7] <- 0 #shift median
      other_eff[1,8] <- 0 #shift upper
      other_eff[1,9] <- 0 #shift lower
      
      colnames(other_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   other_eff <- other_eff %>% mutate(Order = "Other_orders")

  # make the order column a factor so it can be recoded
   other_eff$Order <- as.factor(other_eff$Order)     
      
rm(other_ref, lep_m)
```

# Combine the effects and plot them
```{r}
combined <- rbind(col_eff, dip_eff, hemi_eff, hym_eff, lep_eff, other_eff)

# Add land-use factors for plotting
comb_plot_limits <- c("Primary", "YSV", "ISV", "MSV", "Plantation", "Pasture", "Cropland", "Urban") 

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845", Other_orders = "coral4")

       p <- combined %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff, colour = Order, group = Order)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        scale_x_discrete(limits=comb_plot_limits)+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("Total log mass difference (%)") +
        scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
        
#ggsave(filename = "figures/big_five_LU.png",  plot = p, width = 14, height = 10, dpi = 300)

rm(col_eff, combined, dip_eff, hemi_eff, hym_eff, latitude, lep_eff, logMass_summary, coefficients)
```


# Average mass of an insect at a site (FOR ALL INSECTS)

```{r}
# Create a separate data frame for average weighted mean size of an insect at a site
avg_mass <- all_insects_size

# Calculate abundance weighted log10 average size/mass of an insect at a site
# So going from log10_mass -> ab_weighted_log10_mass
avg_mass <- avg_mass %>%
  group_by(SSBS) %>%
  mutate(
    relativeAb_logMass = (log10_mass * Merged_diversity) / sum(Merged_diversity, na.rm = TRUE)
  ) %>%
  ungroup()

# Now, to calculate the ab weighted average mass of an insect at a site we need to sum all of the mass values at a site and then divide by the sites total abundance, this will give us the ab weighted average mass value of an insect at a site.
site_avg_mass <- avg_mass %>%
  group_by(SSBS) %>%
  summarise(
    sum_relativeAb_logMass = sum(relativeAb_logMass, na.rm = TRUE),
    total_Merged_diversity = sum(Merged_diversity, na.rm = TRUE),
    average_relativeAb_logMass = sum_relativeAb_logMass / total_Merged_diversity
  ) 

# View rows where logMass_total is 0, NA, or blank
site_avg_mass %>%
  filter(is.na(average_relativeAb_logMass) | average_relativeAb_logMass == 0 | average_relativeAb_logMass == "") %>% View()
```

# Merge with PREDICTS insects
```{r}
# Use distinct() to keep only the first appearance of each site within the overall PREDICTS div extract for insects
insects_single <- insects %>%
  distinct(SSBS, .keep_all = TRUE)

# merge site level avg mass values
final_avg_mass <- insects_single %>%
  left_join(site_avg_mass, by = "SSBS") %>% 
  droplevels()

# filter out sites with no avg mass value 
final_avg_mass <- final_avg_mass %>%
  filter(
    !is.na(average_relativeAb_logMass),         # remove NA
    !is.nan(average_relativeAb_logMass),        # remove NaN
    average_relativeAb_logMass != 0,            # remove 0
    average_relativeAb_logMass != ""            # remove empty strings
  )

# 12918 sites in total
```

# Start the modelling of AVG MASS for ALL INSECTS

```{r}
# Rename predominant habitat since its not really habitat we're looking at
final_avg_mass <- rename(final_avg_mass,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
final_avg_mass <- final_avg_mass %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
final_avg_mass <- final_avg_mass %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# Have a look at sample sizes
table(final_avg_mass$LUI)

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 61 sites, merge with light use
final_avg_mass <- final_avg_mass %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Have a look at sample sizes
table(final_avg_mass$LUI)

# Set the reference levels for LUI
final_avg_mass <- final_avg_mass %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )

# Have a look at sample sizes
table(final_avg_mass$LUI)
```


# Create model data for average mass

```{r}
# drop all rows with NAs for land use and use intensity and drop empty levels
model_data_mass <- drop_na(final_avg_mass, Predominant_land_use, Use_intensity) %>% droplevels() 

# check distributions
hist(model_data_mass$average_relativeAb_logMass)
hist(avg_mass$relativeAb_logMass)


# Have a look at sample sizes
table(model_data_mass$Predominant_land_use, model_data_mass$Use_intensity)
table(model_data_mass$LUI)

length(unique(model_data_mass$SS)) # 444 studies 
length(unique(model_data_mass$SSBS)) # 10520 sites
```

```{r}
# Complete cases and model data

# drop all rows with NAs for land use and use intensity and drop empty levels
model_data_mass <- drop_na(final_avg_mass, Predominant_land_use, Use_intensity) %>% droplevels() 
```


# Modelling AVG_MASS ~ LU and creating the LU plot for all insects

```{r}
# maximal model
m1 <- lmer(average_relativeAb_logMass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_mass)

# remove block
m2 <-lmer(average_relativeAb_logMass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS),
    data = model_data_mass)

# remove study
m3 <-lmer(average_relativeAb_logMass ~ Predominant_land_use + (1 | Source_ID),
    data = model_data_mass)

AIC(m1, m2, m3) # m2 best model

# model estimates
summary(m2)
anova(m2)
model_plot(m2)
coef(summary(m2))

# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m2, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log10(average mass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
  #ggsave(filename = "figures/avg_mass_LU.png",  plot = p, width = 6, height = 4, dpi = 300)

  
rm(eff, m1, m2, m3, p, final_avg_mass, model_data_mass)     
```

# Total abundance responses to LU

```{r}
# All-insect set of site-level log(abundance) and species-richness data
# Has filtered out accidentally sampled larvae
abundance <-  qread("/Users/jusi/Desktop/multi-groups/data/abundance-data-prepared-all-insects.qs")

# Filter PREDICTS for same set of sites as in my biomass model data
filtered_abundance <- abundance %>%
  filter(SSBS %in% model_data_size$SSBS) %>%
  distinct(SSBS, .keep_all = TRUE)

# Append use intensity based on SSBS
filtered_abundance <- filtered_abundance %>%
  left_join(
    model_data_size %>% dplyr::select(SSBS, Use_intensity, Source_ID, SSB),
    by = "SSBS"
  )
```

# Start the modelling of all of the insects (Abundance responses)

```{r}
# Rename predominant habitat since its not really habitat we're looking at
filtered_abundance <- rename(filtered_abundance,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
filtered_abundance <- filtered_abundance %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(as.character(Use_intensity), "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(filtered_abundance$Predominant_land_use, filtered_abundance$Use_intensity)
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
filtered_abundance <- filtered_abundance %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# Have a look at sample sizes
table(filtered_abundance$LUI)

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 61 sites, merge with light use
filtered_abundance <- filtered_abundance %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Have a look at sample sizes
table(filtered_abundance$LUI)

# Set the reference levels for LUI
filtered_abundance <- filtered_abundance %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )

# Have a look at sample sizes
table(filtered_abundance$LUI)

filtered_abundance %>% dplyr::select(SS, SSBS, Taxon_fold, Predominant_land_use, Use_intensity, LUI, Log_abundance_pX) %>% View()
```

# Create model data for abundance

```{r}
# drop all rows with NAs for land use, use intensity and log_abundance_px and drop empty levels
model_data_abundance <- drop_na(filtered_abundance, Predominant_land_use, Use_intensity, Log_abundance_pX) %>% droplevels() 

# check distributions
hist(model_data_abundance$Log_abundance_pX)

# Have a look at sample sizes
table(model_data_abundance$Predominant_land_use, model_data_abundance$Use_intensity)
table(model_data_abundance$LUI)

length(unique(model_data_abundance$SS)) # 405 studies 
length(unique(model_data_abundance$SSBS)) # 9794 sites
```

# Modelling ABUNDANCE ~ LU and creating the LU plot for all insects

```{r}
# maximal model
m1 <- lmer(Log_abundance_pX ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_abundance)

# remove block
m2 <-lmer(Log_abundance_pX ~ Predominant_land_use + (1 | Source_ID) + (1 | SS),
    data = model_data_abundance)

# remove study
m3 <-lmer(Log_abundance_pX ~ Predominant_land_use + (1 | Source_ID),
    data = model_data_abundance)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log10(total abundance) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
  #ggsave(filename = "figures/abundance_LU.png",  plot = p, width = 6, height = 4, dpi = 300)

  
rm(eff, m1, m2, m3, p)     
```

# Modelling ABUNDANCE ~ LUI (LU + UI) and creating the LUI plot for all insects

```{r}
# maximal model
m1 <- lmer(Log_abundance_pX ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_abundance)

# remove block
m2 <-lmer(Log_abundance_pX ~ LUI + (1 | Source_ID) + (1 | SS),
    data = model_data_abundance)

# remove study
m3 <-lmer(Log_abundance_pX ~ LUI + (1 | Source_ID),
    data = model_data_abundance)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary_minimal",
                            "LUIPrimary_Light_use" = "Primary_light",
                                "LUIPrimary_Intense_use" = "Primary_intense",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV_minimal",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV_light",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV_intense",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV_minimal",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV_high",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV_minimal",
                            "LUIMature_secondary_vegetation_High_use" = "MSV_high",
                            "LUIPlantation_forest_Minimal_use" = "Plantation_minimal",
                            "LUIPlantation_forest_Light_use" = "Plantation_light",
                            "LUIPlantation_forest_Intense_use" = "Plantation_intense",
                            "LUIPasture_Minimal_use" = "Pasture_minimal",
                            "LUIPasture_Light_use" = "Pasture_light",
                            "LUIPasture_Intense_use" = "Pasture_intense",
                            "LUICropland_Minimal_use" = "Cropland_minimal",
                            "LUICropland_Light_use" = "Cropland_light",
                            "LUICropland_Intense_use" = "Cropland_intense",
                            "LUIUrban_Minimal_use" = "Urban_minimal",
                            "LUIUrban_Light_use" = "Urban_light",
                            "LUIUrban_Intense_use" = "Urban_intense")
                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    p <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3)+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log10(total abundance) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
       #ggsave(filename = "figures/abundance_LUI.png",  plot = p, width = 8, height = 7, dpi = 300)
     
rm(eff, m1, m2, m3, p)     
```




# Calculate taxonomic precision at a site
```{r}
# Get a taxonomic precision score for each site based on how well resolved each taxa at the site is!
# use the same machinery as calculating community weighted mean size

# Simplify df
test <- all_insects_size %>% dplyr::select(SS, SSBS, Order, Family, Best_guess_binomial, Genus, Species, Merged_diversity)

# Assign score based on taxonomy
test <- test %>%
  mutate(taxonomic_precision = case_when(
    !(is.na(Species) | Species == "") ~ 3,
    !(is.na(Genus) | Genus == "")     ~ 2,
    !(is.na(Family) | Family == "")   ~ 1,
    !(is.na(Order) | Order == "")     ~ 0,
    TRUE                              ~ NA_real_
  ))

# Get precision score per site
weighted_taxonomic_precision <- test %>%
  group_by(SSBS) %>%
  summarise(
    weighted_precision = sum(taxonomic_precision * Merged_diversity, na.rm = TRUE) / 
                         sum(Merged_diversity, na.rm = TRUE),
    .groups = "drop"
  )

# Classify each score
weighted_taxonomic_precision <- weighted_taxonomic_precision %>%
  mutate(precision_rank = case_when(
    weighted_precision < 1 ~ "Order",
    weighted_precision < 2 ~ "Family",
    weighted_precision < 3 ~ "Genus",
    weighted_precision == 3 ~ "Species",
    TRUE ~ "Unknown"
  ))

# Join this back to main df to assign score to each row
test <- test %>%
  left_join(weighted_taxonomic_precision, by = "SSBS")

# Filter for low-precision sites and count families, i.e. the number of rows resolved only to the family level of a given family
 test %>%
  filter(precision_rank %in% c("Order", "Family")) %>%
  count(Family, sort = TRUE) 
 
test %>% filter(SSBS == "AD1_2002__Kruess 1  7") %>% View()
```

```{r}
# Do the same but see which groups we do/don't have size data for
# The species we most want to get size values for are the really abundant ones in PREDICTS
# Particularly those where the abundances changes the most in response to land use

 # A list of species with > 1000 merged diversity and no size data
unique_species_high_div <- sp_no_size %>%
  filter(Merged_diversity > 10) %>%
  distinct(Best_guess_binomial, .keep_all = T) %>% 
  filter(!str_detect(Best_guess_binomial, "0"))  # Species entries with incorrect names!

sp_no_size %>%
  #filter(Merged_diversity > 100) %>%
  distinct(Best_guess_binomial, .keep_all = T) %>% 
  View()

# Family x genera combinations and number of rows without species level size data
 sp_no_size %>%
  group_by(Family, Genus) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) 
 
sp_no_size %>% filter(Best_guess_binomial == "Drosophila willistoni") %>% distinct() 

sp_no_size %>%
  filter(Merged_diversity > 100) %>%
  distinct(Best_guess_binomial, .keep_all = T)

rm(sp_no_size, weighted_taxonomic_precision, test)
```

# How much of the dataset is resolved to sp, genus and family level.

```{r}
# Count total number of rows
total_rows <- nrow(all_insects_size)

# Count how many are resolved to each level
species_count <- sum(!is.na(all_insects_size$Trait_Value))
genus_count <- sum(is.na(all_insects_size$Trait_Value) & !is.na(all_insects_size$genus_avg_body_length))
family_count <- sum(is.na(all_insects_size$Trait_Value) & is.na(all_insects_size$genus_avg_body_length) &
                    (!is.na(all_insects_size$family_avg_body_length) | !is.na(all_insects_size$family_avg)))

# Calculate proportions
species_prop <- species_count / total_rows
genus_prop <- genus_count / total_rows
family_prop <- family_count / total_rows

# Display results
cat("Proportion resolved to species level:", species_prop, "\n")
cat("Proportion resolved to genus level:", genus_prop, "\n")
cat("Proportion resolved to family level:", family_prop, "\n")

```

# How many studies sample only bees or ants

```{r}
bee_families <- c("Apidae", "Halictidae", "Megachilidae", "Andrenidae", "Colletidae", "Melittidae", "Stenotritidae")
ant_family <- "Formicidae"

study_family_summary <- all_insects_size %>%
  group_by(SS) %>%
  summarize(unique_families = list(unique(Family)), .groups = "drop")

# Now evaluate each row's family list separately
only_bee <- study_family_summary %>%
  rowwise() %>%
  filter(all(unique_families %in% bee_families)) %>%
  ungroup() %>%
  nrow()

only_ants <- study_family_summary %>%
  rowwise() %>%
  filter(all(unique_families %in% ant_family)) %>%
  ungroup() %>%
  nrow()

cat("Number of studies sampling only bees:", only_bees, "\n")
cat("Number of studies sampling only ants:", only_ants, "\n")
```

# Extract studies for only bees or only ants

```{r}
# Get study IDs that sample only bees
only_bee_studies <- study_family_summary %>%
  rowwise() %>%
  filter(all(unique_families %in% bee_families)) %>%
  pull(SS)

# Get study IDs that sample only ants
only_ant_studies <- study_family_summary %>%
  rowwise() %>%
  filter(all(unique_families %in% ant_family)) %>%
  pull(SS)

# Filter the original data for just those studies

# Filter rows for studies that sampled only bees
bee_only_data <- all_insects_size %>%
  filter(SS %in% only_bee_studies)

# Filter rows for studies that sampled only ants
ant_only_data <- all_insects_size %>%
  filter(SS %in% only_ant_studies)

```

# Function to calculate taxonomic resolution

```{r}
calculate_resolution_proportions <- function(data) {
  total_rows <- nrow(data)
  
  species_count <- sum(!is.na(data$Trait_Value))
  genus_count <- sum(is.na(data$Trait_Value) & !is.na(data$genus_avg_body_length))
  family_count <- sum(is.na(data$Trait_Value) & is.na(data$genus_avg_body_length) &
                        (!is.na(data$family_avg_body_length) | !is.na(data$family_avg)))
  
  species_prop <- species_count / total_rows
  genus_prop <- genus_count / total_rows
  family_prop <- family_count / total_rows
  
  return(list(
    species_count = species_count,
    genus_count = genus_count,
    family_count = family_count,
    species_prop = species_prop,
    genus_prop = genus_prop,
    family_prop = family_prop
  ))
}

# For bee-only studies
bee_resolution <- calculate_resolution_proportions(bee_only_data)

# For ant-only studies
ant_resolution <- calculate_resolution_proportions(ant_only_data)

# Print results
cat("Bee-only studies:\n")
cat("Species-level:", bee_resolution$species_count, 
    " (", round(bee_resolution$species_prop * 100, 1), "%)\n", sep = "")
cat("Genus-level:", bee_resolution$genus_count, 
    " (", round(bee_resolution$genus_prop * 100, 1), "%)\n", sep = "")
cat("Family-level:", bee_resolution$family_count, 
    " (", round(bee_resolution$family_prop * 100, 1), "%)\n\n", sep = "")

cat("Ant-only studies:\n")
cat("Species-level:", ant_resolution$species_count, 
    " (", round(ant_resolution$species_prop * 100, 1), "%)\n", sep = "")
cat("Genus-level:", ant_resolution$genus_count, 
    " (", round(ant_resolution$genus_prop * 100, 1), "%)\n", sep = "")
cat("Family-level:", ant_resolution$family_count, 
    " (", round(ant_resolution$family_prop * 100, 1), "%)\n", sep = "")

```

# From PREDICTS lets work out which groups we do not have size data for yet

```{r}
# From predicts, get me a list of those species without size data

# Rows resolved to sp level but no species data and have an abundance greater than 1
sp_no_size <- all_insects_size %>%
  filter(
    !is.na(Species), # its a species
    is.na(Trait_Value),                       # Trait_Value is an NA
    Merged_diversity > 1                       # Merged_diversity is greater than 1
  )

# Get a list of unique species
unique_species <- unique(sp_no_size$Best_guess_binomial)


summary_df <- all_insects_size %>%
  filter(is.na(Trait_Value)) %>%  # Keep only rows without a Trait_Value
  group_by(Genus, Species) %>%
  summarise(
    count_Trait = n(),            # Count of NA Trait_Value entries
    .groups = "drop"
  ) %>%
  arrange(desc(count_Trait)) 

library(dplyr)

# Step 1: For each Genus, count total rows and how many have either Trait_Value or genus_avg_body_length
genus_data_gap <- all_insects_size %>%
  group_by(Genus) %>%
  summarise(
    total_records = n(),
    #n_without_sp_value = sum(is.na(Trait_Value)),
    n_without_genus_value = sum(is.na(genus_avg_body_length)),
    .groups = "drop"
  ) %>%
  arrange(desc(total_records))  # Most common genera without any trait info


all_insects_size %>% count(Order) 

# View top genera missing trait data
head(genus_data_gap)


```

# Animal traits database

```{r}
# Load the datbase
animaltraits <- read_excel("data/animaltraitsdatabase.xlsx")

# Filter for insects
just_insects <- animaltraits %>% filter(class == "Insecta")

# Columns of interest
just_insects <- just_insects %>% dplyr::select(genus, species, `body mass`, `body mass - units`, `original body mass`, `original body mass - units`)

# Remove those rows without data
just_insects <- just_insects %>%
  filter(!if_all(c("body mass", "body mass - units", "original body mass", "original body mass - units"), is.na))

# Convert kg into mg
just_insects <- just_insects %>%
  rename(body_mass = `body mass`) %>% # deal with space between body and mass
  mutate(body_mass_mg = body_mass * 1e6)

# Calculate an average for duplicate species entries
just_insects <- just_insects %>%
  group_by(species) %>%
  summarise(
    avg_body_mass_mg = mean(body_mass_mg, na.rm = TRUE),
    .groups = "drop"
  )

# Recreate genus column
just_insects <- just_insects %>%
  mutate(genus = sub(" .*", "", species)) # subtract everything before the first space

# Which species are in animaltraits that are also species in my df without a size value?
sp_overlap <- just_insects[just_insects$species %in% unique_species, ]

length(unique(sp_overlap$species)) # 21 species added from AnimalTraits
```

# Do the same but work out if there are any genera which overlap between PREDICTS and Animal Traits

```{r}
# First get a list of all the genera in PREDICTS without a size value
missing_genus_avg <- all_insects_size %>%
  filter(is.na(genus_avg_body_length)) %>%  # Keep only rows with missing genus_avg_body_length
  group_by(Genus) %>%
  summarise(n_missing = n(), .groups = "drop") %>%
  arrange(desc(n_missing))  # Optional: sort by most missing

# Which genera are in Animal Traits that are also genera in my df without a size value?
genera_overlap <- just_insects[just_insects$genus %in% missing_genus_avg$Genus, ]

length(unique(genera_overlap$genus)) # 20 genera added from AnimalTraits

rm(animaltraits, genera_overlap, genus_data_gap, just_insects, missing_genus_avg, sp_overlap, summary_df)
```

# Lets see the overlap between Global Ants and PREDICTS ant taxa with missing size data

```{r}
# Ant species without a size value in PREDICTS
ants_no_size <- all_insects_size %>%
  filter(Family == "Formicidae") %>% 
  filter(
    !is.na(Species), # its a species
    is.na(Trait_Value),                       # Trait_Value is an NA
    Merged_diversity > 1                       # Merged_diversity is greater than 1
  )

# Read in Globala Ants trait data
globalants <- read.csv("data/globalants.csv")

# Filter for columns of interest
globalants <- globalants %>% dplyr::select(Genus, Species, Whole.body.length..mm.) %>% droplevels()

# Create best guess binomial
globalants <- globalants %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Filter out those rows without a body length value
globalants <- globalants %>% filter(`Whole.body.length..mm.` != "Not in paper")

# Convert variable to numeric for summarizing
globalants <- globalants %>%
  mutate(Whole.body.length..mm. = as.numeric(Whole.body.length..mm.))

# Calculate an average for duplicate species entries
avg_BL_ants <- globalants %>%
  group_by(Best_guess_binomial) %>%
  summarise(
    avg_body_length_mm = mean(`Whole.body.length..mm.`, na.rm = TRUE),
    .groups = "drop"
  )

# Recreate genus column
avg_BL_ants <- avg_BL_ants %>%
  mutate(genus = sub(" .*", "", Best_guess_binomial))  # subtract everything before the first space

# Which ant species are in globalants that are also species in my df without a size value?
ant_overlap <- avg_BL_ants[avg_BL_ants$Best_guess_binomial %in% ants_no_size$Best_guess_binomial, ] # 13 species
```

# Same again but for genera

```{r}
# First get a list of all the genera in PREDICTS without a size value
ant_no_genus_length <- all_insects_size %>%
  filter(Family == "Formicidae") %>% 
  filter(is.na(genus_avg_body_length)) %>%  # Keep only rows with missing genus_avg_body_length
  group_by(Genus) %>%
  summarise(n_missing = n(), .groups = "drop") %>%
  arrange(desc(n_missing))  # Optional: sort by most missing

# Genera without a genera value
length(unique(ant_no_genus_length$Genus)) # 58 genera without a size value

# Get a list of the unique ant genera with size data from Global Ants
unique_ant_genera <- avg_BL_ants %>% dplyr::select(genus) %>% distinct()

# Which genera are in Global Ants that are also genera in my df without a size value?
ant_genera_overlap <- ant_no_genus_length[ant_no_genus_length$Genus %in% unique_ant_genera$genus, ] # 6 genera we can add ant data for

rm(ant_genera_overlap, ant_no_genus_length, ant_overlap, ants_no_size, avg_BL_ants, globalants, unique_ant_genera)
```

# Bee functional Trait Database

```{r}
# Load the data
bee_traits <- read.csv("data/Sup Table 5 Trait Data Template.csv")

# Select the cols of interest
bee_traits <- bee_traits %>% dplyr::select(acceptedNameUsage, measurementValue, measurementUnit)

bees_updated <- bee_traits %>%
  mutate(
    genus = word(acceptedNameUsage, 1),
    species = word(acceptedNameUsage, 2),
    Best_guess_binomial = paste(genus, species)
  )

# Those species in the database with less than two words in their acceptednameusage
bee_traits %>%
  filter(str_count(acceptedNameUsage, "\\S+") < 2)


# Bee species without a size value in PREDICTS
bees_no_size <- all_insects_size %>%
  filter(Family %in% c("Apidae", "Halictidae", "Megachilidae", "Andrenidae", "Colletidae", "Melittidae", "Stenotritidae")) %>% 
  filter(
    !is.na(Species), # its a species
    is.na(Trait_Value),                       # Trait_Value is an NA
    Merged_diversity > 1                       # Merged_diversity is greater than 1
  )

bee_overlap <- bees_no_size[bees_no_size$Best_guess_binomial %in% bees_updated$Best_guess_binomial, ] %>% droplevels()

length(unique(bees_no_size$Best_guess_binomial)) # 695 species without size data
length(unique(bee_overlap$Best_guess_binomial)) # Looks like 243 species with bee body size data! Yay!

243/695 # 34% increase!
```

# Bee Functional Trait Database genera

```{r}
# First get a list of all the genera in PREDICTS without a size value
bee_no_genus_length <- all_insects_size %>%
  filter(Family %in% c("Apidae", "Halictidae", "Megachilidae", "Andrenidae", "Colletidae", "Melittidae", "Stenotritidae")) %>% 
  filter(is.na(genus_avg_body_length)) %>%  # Keep only rows with missing genus_avg_body_length
  group_by(Genus) %>%
  summarise(n_missing = n(), .groups = "drop") %>%
  arrange(desc(n_missing))  # Optional: sort by most missing

# Genera without a genera value
length(unique(bee_no_genus_length$Genus)) # 126 genera without a size value

# Get a list of the unique bee genera from the Bee Functional Trait Database
unique_bee_genera <- bees_updated %>% dplyr::select(genus) %>% distinct()

# Which genera are in the Bee Functional Trait Database that are also genera in my df without a size value?
bee_genera_overlap <- bee_no_genus_length[bee_no_genus_length$Genus %in% unique_bee_genera$genus, ] # 59 genera!

59/126 # 46% increase

rm(bee_genera_overlap, bee_no_genus_length, bee_overlap, bee_traits, bees_no_size, bees_updated, unique_bee_genera)
```

# Morphological trait database of saproxylic beetles

```{r}
# beetle species without a size value in PREDICTS
beetles_no_size <- all_insects_size %>%
  filter(Order == "Coleoptera") %>% 
  filter(
    !is.na(Species), # its a species
    is.na(Trait_Value),                       # Trait_Value is an NA
    Merged_diversity > 1                       # Merged_diversity is greater than 1
  ) %>% 
  droplevels()

length(unique(beetles_no_size$Best_guess_binomial)) # 1452 beetle species without size data

# Read in the data
beetle_traits <- read.csv("data/Hagge_etal_saproxylic_beetle_morphological_trait_database_20210426.csv")

# Select cols of interest
beetle_traits <- beetle_traits %>% dplyr::select(family, species, body_length)

# Recreate cols for matching
beetle_traits <- beetle_traits %>%
  separate(species, into = c("genus", "species"), sep = "_") %>% # seperate species into genus and species
  mutate(best_guess_binomial = paste(genus, species, sep = " ")) # create best_guess_binomial

# No species overlap :( - unsurprising given its saproxylic beetles
beetle_overlap <- beetles_no_size[beetles_no_size$Best_guess_binomial %in% beetle_traits$Best_guess_binomial, ] 

```

# Lets see if any of the genera match between PREDICTS beetles and the trait database

```{r}
# First get a list of all the genera in PREDICTS without a size value
beetle_no_genus_length <- all_insects_size %>%
  filter(Order == "Coleoptera") %>% 
  filter(is.na(genus_avg_body_length)) %>%  # Keep only rows with missing genus_avg_body_length
  group_by(Genus) %>%
  summarise(n_missing = n(), .groups = "drop") %>%
  arrange(desc(n_missing))  # Optional: sort by most missing

# Genera without a genera value
length(unique(beetle_no_genus_length$Genus)) # 939 genera without a size value

# Get a list of the unique beetle genera from the trait database
unique_beetle_genera <- beetle_traits %>% dplyr::select(genus) %>% distinct()

# Which genera are in the trait datatbase that are also genera in my df without a size value?
beetle_genera_overlap <- beetle_no_genus_length[beetle_no_genus_length$Genus %in% unique_beetle_genera$genus, ] # 36 genera overlap!

36/939 # 3% increase! eek!

rm(beetle_genera_overlap, beetle_no_genus_length, beetle_overlap, beetle_traits, beetles_no_size, unique_beetle_genera)
```

# Data from: A summary of eight traits of Coleoptera, Hemiptera, Orthoptera and Araneae, occurring in grasslands in Germany

```{r}
# Read in the trait data
mixed_traits <- read_delim("data/ArthropodSpeciesTraits.txt")

# Subset for Orders of interest
mixed_traits <- mixed_traits %>% filter(Order %in% c("Coleoptera", "Hemiptera", "Orthoptera"))

# select cols of interest
mixed_traits <- mixed_traits %>% dplyr::select(Family, SpeciesID, Body_Size)

# filter PREDICTS for coleopterans, hemipterans and orthopteran species without a size value
mixed_no_size <- all_insects_size %>%
  filter(Order %in% c("Coleoptera", "Hemiptera", "Orthoptera")) %>% 
  filter(
    !is.na(Species), # its a species
    is.na(Trait_Value),                       # Trait_Value is an NA
    Merged_diversity > 1                       # Merged_diversity is greater than 1
  )

mixed_overlap <- mixed_no_size[mixed_no_size$Best_guess_binomial %in% mixed_traits$SpeciesID, ] 

unique_counts <- mixed_overlap %>%
  group_by(Order) %>%
  summarise(n_unique_binomials = n_distinct(Best_guess_binomial)) %>%
  arrange(n_unique_binomials)  # Optional: sort by count

# match?
# 14 species of Hemiptera
# 18 species of Coleoptera

```

# Lets see if any of the genera match 

```{r}
# First get a list of all the genera in PREDICTS without a size value
mixed_no_genus_length <- all_insects_size %>%
  filter(Order %in% c("Coleoptera", "Hemiptera", "Orthoptera")) %>% 
  filter(is.na(genus_avg_body_length)) %>%  # Keep only rows with missing genus_avg_body_length
  group_by(Genus) %>%
  summarise(n_missing = n(), .groups = "drop") %>%
  arrange(desc(n_missing))  # Optional: sort by most missing

# Genera without a genera value
length(unique(mixed_no_genus_length$Genus)) # 1591 genera without a size value

# Recreate cols for matching
mixed_traits <- mixed_traits %>%
  separate(SpeciesID, into = c("genus", "species"), sep = " ") %>% # seperate species into genus and species
  mutate(best_guess_binomial = paste(genus, species, sep = " ")) # create best_guess_binomial

# Get a list of the unique beetle genera from the trait database
unique_mixed_genera <- mixed_traits %>% dplyr::select(genus) %>% distinct()

# Which genera are in the trait datatbase that are also genera in my df without a size value?
mixed_genera_overlap <- mixed_no_genus_length[mixed_no_genus_length$Genus %in% unique_mixed_genera$genus, ] # 18 genera overlap!

18/1591 # 1% increase! eek!

rm(mixed_genera_overlap, mixed_no_genus_length, mixed_no_size, mixed_overlap, mixed_traits, unique_counts, unique_mixed_genera)
```

# PREDICTS butterfly trait data from Cleary

```{r}
# read in the data
butterfly_traits <- read.csv("data/butterfly_traits_full.csv")

# Manipulate the columns so we can create best_guess_binomial and extract the body size value for each species
butterfly_traits <- butterfly_traits %>%
  mutate(
    Best_guess_binomial = sub("^((\\S+)\\s+(\\S+)).*", "\\1", Species.Family.Subfamily.Tribe.Genus.AuthorSp.Range.Group.Geo.Gen.MeanWing.SexDi.Conspicuousness.Migratory),
    body_size = sapply(strsplit(X, "\\s+"), function(x) as.numeric(x[7]))
  )

# filter PREDICTS for Lep species without a size value
leps_no_size <- all_insects_size %>%
  filter(Order == "Lepidoptera") %>% 
  filter(
    !is.na(Species), # its a species
    is.na(Trait_Value),                       # Trait_Value is an NA
    Merged_diversity > 1                       # Merged_diversity is greater than 1
  )

leps_overlap <- leps_no_size[leps_no_size$Best_guess_binomial %in% butterfly_traits$Best_guess_binomial, ] 

length(unique(leps_no_size$Best_guess_binomial)) # 1846 species without size data
length(unique(leps_overlap$Best_guess_binomial)) # Looks like 351 species with body size data! Yay!

351/1846 # 19% increase
```

# Match lep genera

```{r}
# First get a list of all the genera in PREDICTS without a size value
leps_no_genus_length <- all_insects_size %>%
  filter(Order == "Lepidoptera") %>% 
  filter(is.na(genus_avg_body_length)) %>%  # Keep only rows with missing genus_avg_body_length
  group_by(Genus) %>%
  summarise(n_missing = n(), .groups = "drop") %>%
  arrange(desc(n_missing))  # Optional: sort by most missing

# Genera without a genera value
length(unique(leps_no_genus_length$Genus)) # 902 genera without a size value

# Recreate cols for matching
butterfly_traits <- butterfly_traits %>%
  separate(Best_guess_binomial, into = c("genus", "species"), sep = " ") # seperate BGB into genus and species

# Get a list of the unique lepidoptera genera from the trait database
unique_butterfly_genera <- butterfly_traits %>% dplyr::select(genus) %>% distinct()

# Which genera are in the trait datatbase that are also genera in my df without a size value?
butterfly_genera_overlap <- leps_no_genus_length[leps_no_genus_length$Genus %in% unique_butterfly_genera$genus, ] # 165 genera overlap!

165/902 # 18% increase! eek!

rm(butterfly_genera_overlap, butterfly_traits, leps_no_genus_length, leps_no_size, leps_overlap, unique_butterfly_genera)
```

# Data from: Refining estimates of dry body weight from linear measurements in adult moths and butterflies (Lepidoptera)

```{r}
# read in the data
enrique_traits <- read.csv("data/LEPSIZE2024.csv")

# Select cols of interest
enrique_traits <- enrique_traits %>% dplyr::select(FAMILY, SPECIES, DBw)

# Clean and tidy the columns
enrique_traits <- enrique_traits %>%
  mutate(
    # Remove quotes and extract words
    clean_species = str_remove_all(SPECIES, '"'),
    genus = word(clean_species, 1),
    species = word(clean_species, 2),
    Best_guess_binomial = paste(genus, species)
  )

# filter PREDICTS for Lep species without a size value
leps_no_size <- all_insects_size %>%
  filter(Order == "Lepidoptera") %>% 
  filter(
    !is.na(Species), # its a species
    is.na(Trait_Value),                       # Trait_Value is an NA
    Merged_diversity > 1                       # Merged_diversity is greater than 1
  )

leps_overlap <- leps_no_size[leps_no_size$Best_guess_binomial %in% enrique_traits$Best_guess_binomial, ] 

length(unique(leps_no_size$Best_guess_binomial)) # 1846 species without size data
length(unique(leps_overlap$Best_guess_binomial)) # Looks like 119 species with body size data! Yay!
```

# Match lep genera

```{r}
# First get a list of all the genera in PREDICTS without a size value
leps_no_genus_length <- all_insects_size %>%
  filter(Order == "Lepidoptera") %>% 
  filter(is.na(genus_avg_body_length)) %>%  # Keep only rows with missing genus_avg_body_length
  group_by(Genus) %>%
  summarise(n_missing = n(), .groups = "drop") %>%
  arrange(desc(n_missing))  # Optional: sort by most missing

# Genera without a genera value
length(unique(leps_no_genus_length$Genus)) # 902 genera without a size value

# Get a list of the unique lepidoptera genera from the trait database
unique_butterfly_genera <- enrique_traits %>% dplyr::select(genus) %>% distinct()

# Which genera are in the trait datatbase that are also genera in my df without a size value?
butterfly_genera_overlap <- leps_no_genus_length[leps_no_genus_length$Genus %in% unique_butterfly_genera$genus, ] # 250 genera overlap!

250/902 # 28% increase! eek!

rm(butterfly_genera_overlap, enrique_traits, leps_no_genus_length, leps_no_size, leps_overlap, unique_butterfly_genera)
```


# What's the variance in the abundance weighted average body size among sites compared to among studies

```{r}
# merge weighted_mean_size back into original df
variation_cwms <- left_join(taxon_insects_updated, weighted_mean_size_sites, by = "SSBS")

summary(aov(weighted_mean_size ~ SS, data = variation_cwms)) # 10% of the variation is explained by variation among sites!

rm(variation_cwms)
```


# Read in overall trait dataset
```{r}
# Read in trait data
traits <- read_excel("data/All_FamilyandOrder_Traits_for_JustinIsip.xlsx") 

# As data frame 
traits <- as.data.frame(traits)

# Combine the first three rows into a single character vector (column names)
combined_names <- sapply(traits[1:3, ], function(x) paste(na.omit(x), collapse = " | "))

# Assign the new names to the columns
colnames(traits) <- combined_names

# Drop the first three rows now that they're in the header
traits <- traits[-c(1:3), ]

# Extract the first row (which contains new column names for the first 8 columns)
new_names_part <- as.character(unlist(traits[1, 1:8]))

# Replace the first 8 column names
colnames(traits)[1:8] <- new_names_part

# Remove the first row now that it's used as header
traits <- traits[-1, ]

# Reset rownames (just for cleanliness as they don't start from 1)
rownames(traits) <- NULL

# Lets have a look at a particular trait size category and its coverage across families
traits %>% dplyr::select("Family", "Taxon Name", "Minimum size | < 0.35 | mm") %>% View()

# Identify all traits present for a group
traits %>%
  pivot_longer(cols = 9:ncol(.), # all trait columns
               names_to = "Trait", # new column for trait names
               values_to = "Value") %>%# new column for the values
  filter(!is.na(Value) & Value != 0) %>%  # keep only traits that are present, i.e are not NA or 0
View()
  
# Pivot longer traits
pivot_trait <- traits %>%
  pivot_longer(
    cols = 9:ncol(.), # first 8 cols are meta data
    names_to = "Trait",
    values_to = "Value"
  ) %>%
  filter(!is.na(Value) & Value != "0") %>% # if the trait value is not an NA and is not equal to 0
  separate(Trait, into = c("Trait", "Modality", "Unit"), sep = "\\|", fill = "right", extra = "merge") %>% 
  mutate(Mode = ifelse(is.na(Unit), Modality, paste(Modality, Unit))) %>% # what is the trait category
  dplyr::select(-Modality, -Unit) %>% # remove redundant cols
  dplyr::select(-Value, Value)  # moves 'Value' to the end

# Strings 2 factor
pivot_trait <- strings2factors(pivot_trait)

# Pivot wider again
trait_matrix <- pivot_trait %>%
  group_by(Trait) %>% 
  mutate(row = row_number()) %>% 
  pivot_wider(
    names_from = Trait,
    values_from = Mode,
    values_fill = NA
  ) %>% 
  dplyr::select(-row)

# Remove trailing space after column names
names(trait_matrix) <- str_remove(names(trait_matrix), "\\s+$")

sum(model_data_ab$Family == "") # 938 sites without family data

model_data_ab %>% dplyr::select(SS, SSBST, Order, Family) %>% unique() %>% View()

merged_data <- left_join(model_data_ab, trait_matrix, by = "Family")

pivot_trait %>% filter(Trait == "Dispersal ability ") %>% View()
```
# Code up just for size

```{r}
# Filter trait database for meta data and just minimum and maximum size cols
size <- traits %>% dplyr::select(1:32)

# Rename the size columns so each one is a range
size <- size %>%
  rename(
    "Minimum size 0.17 - 0.35mm" = "Minimum size | < 0.35 | mm",
    "Minimum size 0.35 - 0.71mm" = "Minimum size | < 0.71 | mm",
    "Minimum size 0.71 - 1.41mm" = "Minimum size | < 1.41 | mm",
    "Minimum size 1.41 - 2.82mm" = "Minimum size | < 2.82 | mm",
    "Minimum size 2.82 - 5.62mm" = "Minimum size | < 5.62 | mm",
    "Minimum size 5.62 - 11.22mm" = "Minimum size | < 11.22 | mm",
    "Minimum size 11.22 - 22.39mm" = "Minimum size | < 22.39 | mm",
    "Minimum size 22.39 - 44.67mm" = "Minimum size | < 44.67 | mm",
    "Minimum size 44.67 - 89.13mm" = "Minimum size | < 89.13 | mm",
    "Minimum size 89.13 - 177.83mm" = "Minimum size | < 177.83 | mm",
    "Minimum size 177.83 - 354.81mm" = "Minimum size | < 354.81 | mm",
    "Minimum size 354.81 - 707.95mm" = "Minimum size | < 707.95 | mm",
    "Maximum size 0.17 - 0.35mm" = "Maximum size | < 0.35 | mm",
    "Maximum size 0.35 - 0.71mm" = "Maximum size | < 0.71 | mm",
    "Maximum size 0.71 - 1.41mm" = "Maximum size | < 1.41 | mm",
    "Maximum size 1.41 - 2.82mm" = "Maximum size | < 2.82 | mm",
    "Maximum size 2.82 - 5.62mm" = "Maximum size | < 5.62 | mm",
    "Maximum size 5.62 - 11.22mm" = "Maximum size | < 11.22 | mm",
    "Maximum size 11.22 - 22.39mm" = "Maximum size | < 22.39 | mm",
    "Maximum size 22.39 - 44.67mm" = "Maximum size | < 44.67 | mm",
    "Maximum size 44.67 - 89.13mm" = "Maximum size | < 89.13 | mm",
    "Maximum size 89.13 - 177.83mm" = "Maximum size | < 177.83 | mm",
    "Maximum size 177.83 - 354.81mm" = "Maximum size | < 354.81 | mm",
    "Maximum size 354.81 - 707.95mm" = "Maximum size | < 707.95 | mm",
  )

# Calculate the logs
(log(0.17) + log(0.35)) / 2 # -1.410889
(log(0.35) + log(0.71)) / 2 # -0.6961562
(log(0.71) + log(1.41)) / 2 # 0.0005496977
(log(1.41) + log(2.82)) / 2 # 0.6901633
(log(2.82) + log(5.62)) / 2 # 1.381534
(log(5.62) + log(11.22)) / 2 # 2.072015
(log(11.22) + log(22.39)) / 2 # 2.763156
(log(22.39) + log(44.67)) / 2 # 3.453958
(log(44.67) + log(89.13)) / 2 # 4.144699
(log(89.13) + log(177.83)) / 2 # 4.835462
(log(177.83) + log(354.81)) / 2 # 5.526205
(log(354.81) + log(707.95)) / 2 # 6.216978

# Rename the size columns so each one is log value now
size <- size %>%
  rename(
  "Log minimum size -1.410889" = "Minimum size 0.17 - 0.35mm",
    "Log minimum size -0.6961562" = "Minimum size 0.35 - 0.71mm",
    "Log minimum size 0.0005496977" = "Minimum size 0.71 - 1.41mm",
    "Log minimum size 0.6901633" = "Minimum size 1.41 - 2.82mm",
    "Log minimum size 1.381534" = "Minimum size 2.82 - 5.62mm",
    "Log minimum size 2.072015" = "Minimum size 5.62 - 11.22mm",
    "Log minimum size 2.763156" = "Minimum size 11.22 - 22.39mm",
    "Log minimum size 3.453958" = "Minimum size 22.39 - 44.67mm",
    "Log minimum size 4.144699"= "Minimum size 44.67 - 89.13mm",
    "Log minimum size 4.835462" = "Minimum size 89.13 - 177.83mm",
   "Log minimum size 5.526205"  = "Minimum size 177.83 - 354.81mm",
    "Log minimum size 6.216978" = "Minimum size 354.81 - 707.95mm",
    "Log maximum size -1.410889" = "Maximum size 0.17 - 0.35mm",
    "Log maximum size -0.6961562" = "Maximum size 0.35 - 0.71mm",
    "Log maximum size 0.0005496977" = "Maximum size 0.71 - 1.41mm",
    "Log maximum size 0.6901633" = "Maximum size 1.41 - 2.82mm",
    "Log maximum size 1.381534" = "Maximum size 2.82 - 5.62mm",
    "Log maximum size 2.072015" = "Maximum size 5.62 - 11.22mm",
    "Log maximum size 2.763156" = "Maximum size 11.22 - 22.39mm",
    "Log maximum size 3.453958" = "Maximum size 22.39 - 44.67mm",
    "Log maximum size 4.144699" = "Maximum size 44.67 - 89.13mm",
    "Log maximum size 4.835462" = "Maximum size 89.13 - 177.83mm",
    "Log maximum size 5.526205" = "Maximum size 177.83 - 354.81mm",
    "Log maximum size 6.216978" = "Maximum size 354.81 - 707.95mm",
  )


# Step 1: Identify columns
min_cols <- grep("^Log minimum size", names(size), value = TRUE)
max_cols <- grep("^Log maximum size", names(size), value = TRUE)

# Step 2: Extract weights (from column names)
min_weights <- as.numeric(str_extract(min_cols, "-?\\d+\\.?\\d*"))
max_weights <- as.numeric(str_extract(max_cols, "-?\\d+\\.?\\d*"))

# Step 3: Ensure columns are numeric
size[min_cols] <- lapply(size[min_cols], function(x) as.numeric(as.character(x)))
size[max_cols] <- lapply(size[max_cols], function(x) as.numeric(as.character(x)))

# Step 3: Calculate weighted sums
size <- size %>%
  rowwise() %>% # makes mutate() operate per row instead of per column
  mutate(
    min_weighted = sum(c_across(all_of(min_cols)) * min_weights, na.rm = TRUE), # Added na.rm = TRUE to skip any NA values during sum
    max_weighted = sum(c_across(all_of(max_cols)) * max_weights, na.rm = TRUE)
  ) %>%
  mutate(avg_weight = (min_weighted + max_weighted) / 2) %>% 
  ungroup()

# c_across() grabs all the columns listed in min_cols for the current row.
# all_of(min_cols) safely ensures only the columns you've selected are used (and errors if a name is missing — helpful for debugging).
unique(size$`Log minimum size -1.410889`)
unique(size$`Log minimum size 0.0005496977`)

rm(combined_names, max_cols, max_weights, min_cols, min_weights, new_names_part)
```

# Try with a range trait that has fuzzy coding
```{r}
# Lets go with lifelong fecundity

# Filter trait database for meta data and lifelong fecundity cols
fecund <- traits %>% dplyr::select(c(1:8,77:80))

# Calculate the logs
(log(10) + log(100)) / 2 # 3.453878
(log(100) + log(1000)) / 2 # 5.756463
(log(1000) + log(3000)) / 2 # 7.457061
(log(3000) + log(8000)) / 2 # 8.496782

# Rename the fecundity columns so each one is log value now
fecund <- fecund %>%
  rename(
  "Log lifelong fecundity 3.453878 number of eggs per female" = "Lifelong fecundity | < 100 | Number of eggs per female",
    "Log lifelong fecundity 5.756463 number of eggs per female" = "Lifelong fecundity | 100-1000 | Number of eggs per female",
    "Log lifelong fecundity 7.457061 number of eggs per female" = "Lifelong fecundity | 1000-3000 | Number of eggs per female",
    "Log lifelong fecundity 8.496782 number of eggs per female" = "Lifelong fecundity | > 3000 | Number of eggs per female",
  )

# Step 1: Identify columns
fecund_cols <- grep("^Log lifelong fecundity", names(fecund), value = TRUE)

# Step 2: Extract weights (from column names)
fecund_weights <- as.numeric(str_extract(fecund_cols, "-?\\d+\\.?\\d*"))

# Step 3: Ensure columns are numeric
fecund[fecund_cols] <- lapply(fecund[fecund_cols], function(x) as.numeric(as.character(x)))

# Step 3: Calculate weighted sums
fecund_test <- fecund %>%
  rowwise() %>% # makes mutate() operate per row instead of per column
  mutate(
    weighted_average = sum(c_across(all_of(fecund_cols)) * fecund_weights, na.rm = TRUE), # Added na.rm = TRUE to skip any NA values during sum
  ) %>%
  ungroup()
```

# Try with dispersal ability
```{r}
dispersal <- traits %>% dplyr::select(c(1:8, 196:200))

dispersal_weights <- c(1,2,3,4,5)

dispersal_cols <- grep("^Dispersal ability", names(dispersal), value = TRUE)

# Step 3: Ensure columns are numeric
dispersal[dispersal_cols] <- lapply(dispersal[dispersal_cols], function(x) as.numeric(as.character(x)))

# Step 3: Calculate weighted sums
dispersal_test <- dispersal %>%
  rowwise() %>% # makes mutate() operate per row instead of per column
  mutate(
    weighted_average = sum(c_across(all_of(dispersal_cols)) * dispersal_weights, na.rm = TRUE), # Added na.rm = TRUE to skip any NA values during sum
  ) %>%
  ungroup()

```

# Try with egg volume

```{r}
egg <- traits %>% dplyr::select(c(1:8,166:175))

# Calculate the midrange as they are already in log10 space
(-6.15 + -3.43) / 2 # -4.79

model_data_ab %>% filter(SSBST == "CC1_2012__Waite 1 3 14 Coleoptera") %>% View()
```


# SIZE - Merging Johns size trait to my model_data

```{r}
# Get taxon info for all insects
taxon_insects <- insects %>% dplyr::select(SS, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Parsed_name, Taxon, Best_guess_binomial, Effort_corrected_measurement, Merged_diversity)

# Get taxon info from Johns size dataset
john_size <- size %>% dplyr::select(Family, "Taxon Name", min_weighted, max_weighted, avg_weight)

# Get rid of missing size information, i.e. avg_weight == 0 or == NA
john_size <- john_size %>% filter(!is.na(avg_weight) & avg_weight != 0)

# Get rid of rows which aren't at family level, i.e. at superfamily or higher
john_size <- john_size %>% filter(!is.na(Family))

# merge the two
merged_size <- inner_join(taxon_insects, john_size, by = "Family")

# Which rows were excluded?
excluded <- anti_join(taxon_model, john_size, by = "Family")

rm(taxon_model, john_size, excluded)
```

# Test dataset

```{r}
test <- all_insects_size %>%
  group_by(SSBS) %>%
  filter(n_distinct(Order) > 2,
         n_distinct(Best_guess_binomial) >= 6,
         n_distinct(Best_guess_binomial) <= 8,
    mean(Merged_diversity != 0, na.rm = TRUE) > 0.4
         ) %>%
  ungroup()

#test_data <- test %>% filter(SSBS == "AD1_2009__Vergara 1  6")

#write.csv(test_data, file = "test_site_1.csv")

```

# Summary statistics 

```{r}
# Species-level data
n_species <- all_insects_size %>%
  filter(!is.na(Trait_Value)) %>%
  distinct(Best_guess_binomial) %>%
  nrow()

length(unique(all_insects_size$Best_guess_binomial))

# Genus-level data
n_genera <- all_insects_size %>%
  filter(!is.na(genus_avg_body_length)) %>%
  distinct(Genus) %>%
  nrow()

# Family-level data
n_families <- all_insects_size %>%
  filter(!is.na(family_avg_body_length)) %>%
  distinct(Family) %>%
  nrow()

# Show the results
cat("Species with size data:", n_species, "\n")
cat("Genera with size data:", n_genera, "\n")
cat("Families with size data:", n_families, "\n")


# Step 1: Add a column that identifies the source of body size
test <- all_insects_size %>%
  mutate(
    data_source = case_when(
      !is.na(Trait_Value) ~ "species",
      is.na(Trait_Value) & !is.na(genus_avg_body_length) ~ "genus",
      is.na(Trait_Value) & is.na(genus_avg_body_length) & !is.na(family_avg_body_length) ~ "family",
      TRUE ~ "missing"
    )
  )

# Step 2: Summarize by Family and Genus
group_summary <- test %>%
  group_by(Family, Genus) %>%
  summarise(
    n_species = n(),
    n_species_level = sum(data_source == "species"),
    n_genus_level = sum(data_source == "genus"),
    n_family_level = sum(data_source == "family"),
    n_missing = sum(data_source == "missing"),
    .groups = "drop"
  ) %>%
  mutate(
    percent_species_level = 100 * n_species_level / n_species,
    percent_genus_level = 100 * n_genus_level / n_species,
    percent_family_level = 100 * n_family_level / n_species
  )

# Step 3: Identify groups with low species-level data
low_species_data <- group_summary %>%
  filter(percent_species_level < 50) %>%  # threshold is flexible
  arrange(percent_species_level)

# Optional: View top 10 groups with lowest species-level data
head(low_species_data, 10)

```