---
title: " Matched comparisons highlight divergent impacts of land-use change on the dominant orders of insects"
author: "Justin Ephraim Isip"
date: "2026-01-13"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Overview
This document reruns Chapter 2 pairwise comparisons among focal insect orders using `glmmTMB`.
Abundance and species richness are analysed separately using negative-binomial mixed models, with zero-inflation included where warranted by diagnostics.

# Load packages

```{r load-packages}
library(tidyverse)
library(glmmTMB)
library(DHARMa)
library(car)
library(MuMIn)
library(emmeans)
library(ggpubr)
library(here)
library(performance)
```

# Load data

```{r load-data}
model_data_ab <- readRDS(here::here("data", "model_data_ab.rds"))
model_data_sr <- readRDS(here::here("data", "model_data_sr.rds"))
```

# Convert to factors 
```{r include = FALSE}
str(model_data_ab)
model_data_ab <- model_data_ab %>%
  mutate(across(c(SS, SSB, SSBS), as.factor))

str(model_data_sr)
model_data_sr <- model_data_sr %>%
  mutate(across(c(SS, SSB, SSBS), as.factor))

length(unique(model_data_ab$SS))
length(unique(model_data_sr$SS))
length(unique(model_data_ab$SSBS))
length(unique(model_data_sr$SSBS))
```

#### Create land use categories and colours for plotting

```{r}
# Land use levels
comb_plot_limits <- c("Primary", "Secondary vegetation", "Plantation forest", "Pasture", "Cropland", "Urban") 

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#197BBD", Coleoptera = "#581845") 
```

# First lets deal with the non-integer count data in the model data 

```{r}
# Are there non count data in model_data_ab?
unique(levels(model_data_ab$Diversity_metric_unit)) # yes

# First lets filter for integer only data
model_data_ab_integer_only <- model_data_ab %>% filter(Diversity_metric_unit == "individuals") %>% droplevels()

# Check again
unique(levels(model_data_ab_integer_only$Diversity_metric_unit))

# Are there non count data in model_data_sr?
unique(levels(model_data_sr$Diversity_metric_unit)) # yes

# First lets filter for integer only data
model_data_sr_integer_only <- model_data_sr %>% filter(Diversity_metric_unit == "individuals") %>% droplevels()

# Check again
unique(levels(model_data_sr_integer_only$Diversity_metric_unit))
```

# Great now both the abundance and species richness model data frames are for count data only
# But are there still non-integers despite the fact we've filtered only for individuals?
```{r}
# If you subtract each total abundance value in my data frame from the rounded version of the same value, is that value less than 0.000000001? (across all values)
all(abs(model_data_ab_integer_only$TotalAbundance -
        round(model_data_ab_integer_only$TotalAbundance)) < 1e-8,
    na.rm = TRUE) # FALSE

all(abs(model_data_sr_integer_only$SpeciesRichness -
        round(model_data_sr_integer_only$SpeciesRichness)) < 1e-8,
    na.rm = TRUE) # TRUE

# There are still non-integers in the abundance data, lets filter for them

# Check which values are non-integers
still_non_integer_abundance <- model_data_ab_integer_only %>%
  dplyr::filter(abs(TotalAbundance - round(TotalAbundance)) > 1e-8) %>% dplyr::select(SS, SSBS, SSBST, Best_guess_binomial, Diversity_metric, Diversity_metric_unit, Diversity_metric_type,Effort_corrected_measurement, Corrected_sampling_effort,  merged_diversity, TotalAbundance, SpeciesRichness, to_add, logAbundance) %>%  droplevels()

# Which studies?
unique(still_non_integer_abundance$SS)

# Ok lets round all of the Total Abundance values to the nearest integer, any value greater than 0.5 should become 1!
still_non_integer_abundance <- still_non_integer_abundance %>%
  mutate(
    RoundedTotalAbundance = if_else(
      TotalAbundance >= 0.5,
      ceiling(TotalAbundance),
      0
    )
  )

# Ok lets merge these rounded values back into the overall df for those SSBSTs
rounded_abundances <- still_non_integer_abundance %>%  dplyr::select(SSBST, RoundedTotalAbundance) # 407 SSBSTs have had their abundances rounded

# Merge into original data frame
model_data_ab_integer_only <- model_data_ab_integer_only %>% 
    left_join(rounded_abundances, by = ("SSBST")) %>%  # join the values in by their ssbst
  mutate(
    TotalAbundance = coalesce(RoundedTotalAbundance, TotalAbundance) # if a rounded values exists, assign it to TotalAbundance otherwise keep the original value
  ) %>% 
  dplyr::select(-RoundedTotalAbundance)

# Ok now all the abundance values should be integers now that the non-integers have been rounded
all(abs(model_data_ab_integer_only$TotalAbundance -
        round(model_data_ab_integer_only$TotalAbundance)) < 1e-8,
    na.rm = TRUE) # TRUE - all the values in TotalAbundance are integers

# Remove old data frames
rm(model_data_ab, model_data_sr, still_non_integer_abundance, rounded_abundances)
```

# Are the models errors distributed the same way for each land use?
If the model is appropriate the residuals should be uniformly spread between 0 and 1 within each LU category
residual spread should be similar across LUs, i.e. does the model behave equally well in primary versus cropland versus urban etc?

# How to read the plot
The boxes are one per land use, each box shows the distribution of DHARMa residuals for each land use.
If the model fits well, each box should be centered near 0.5, boxes should have similar heights, no strong skew toward 0 or 1.

# What DHARMa is telling me? (red text)
`within-group deviations from uniformity significant` this means that within each land-use category, residuals are not uniform, the model is systematically biased within each LU. `Levene test for homogeneity of variance significant` this means that different land use categories have different residual spread, in simple terms - abundance varies a lot in some LUs, ie plantation, much less in others (e.g. urban) - but my model assumes that they all have the same variance - that assumption is violated. 

# What can we see visually?
Urban has very tight residuals, many near the bottom.
Plantation has a wide spread, lots of large residuals.
So, one "average" error structure does not fit all land-use types.

# Why this happens ecologically?
Different land uses differ in habitat quality, detectibility and how often abundance is zero.
My model currently assumes: "abundance behaves the same way everywhere, just with different means"
The data says "abundance behaves fundamentally different across LUs"


# Ok now all the abundance and species richness values for both model datasets are integers we can start modelling..

# Hymenoptera versus Coleoptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
hym_col_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hymenoptera_coleoptera <- model_data_ab_integer_only %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% droplevels()

# Check the breakdown of ants, bees and wasps based on family
# everything which isn't Formicidae (ants) or the 9 families of bee, is therefore a wasp
within_hymenoptera <-
  hymenoptera_coleoptera %>%
  filter(Higher_taxon == "Hymenoptera") %>% # Subset abundance data for hymenoptera
  mutate(Which_group = 
           case_when(
             Family == "Formicidae" ~ "Ants",
             Family == "Andrenidae" | Family == "Apidae" | Family == "Colletidae" | Family == "Dasypodaidae" |
               Family == "Halictidae" | Family == "Megachilidae" | Family == "Meganomiidae" | Family == "Melittidae"|
               Family == "Stenotritidae" ~ "Bees",   
             Family != "Formicidae" | Family != "Andrenidae" | Family != "Apidae" | Family != "Colletidae" |
               Family != "Dasypodaidae" | Family != "Halictidae" | Family != "Megachilidae" | Family != "Meganomiidae" |
               Family != "Melittidae" | Family != "Stenotritidae" ~ "Wasps")) %>%
  droplevels()

plot_data <- within_hymenoptera %>%
  count(Which_group) %>%
  mutate(
    prop = n / sum(n),
    label = paste0(Which_group, "\n", n, " (", round(prop * 100, 1), "%)")
  )

ggplot(plot_data, aes(x = "", y = n, fill = Which_group)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3.5) +
  labs(
    title = paste0("Coleoptera versus Hymenoptera (n = ", nrow(within_hymenoptera), ")"),
    fill = "Which group"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

coleoptera_only <- hymenoptera_coleoptera %>% filter(Order == "Coleoptera")

table(coleoptera_only$Sampling_method, coleoptera_only$Predominant_land_use)

rm(plot_data, within_hymenoptera, hymenoptera_coleoptera)
```

# Modelling

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# NB with simple ZI
m_zinb1 <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ 1,
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# NB with ZI by LU
m_zinb2 <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ Predominant_land_use,
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# NB with ZI by higher taxon
m_zinb3 <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ Higher_taxon,
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# NB wih complex ZI
m_zinb_full <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ Predominant_land_use * Higher_taxon,
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

summary(m_nb) ; summary(m_zinb1) ; summary(m_zinb_full)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # p < 0.05
testDispersion(res_nb)
plot(res_nb)

# Does adding simple ZI fix the zero problem?
res_zinb1 <- simulateResiduals(m_zinb1, n = 1000)
testZeroInflation(res_zinb1)
plot(res_zinb1)

# ZI by LU
res_zinb2 <- simulateResiduals(m_zinb2, n = 1000)
testZeroInflation(res_zinb2)
plot(res_zinb2)

summary(m_zinb2)

#  ZI by higher taxon
res_zinb3 <- simulateResiduals(m_zinb3, n = 1000)
testZeroInflation(res_zinb3)
plot(res_zinb3)

summary(m_zinb3)

# Does adding complex ZI fix the zero problem?
res_zinbfull <- simulateResiduals(m_zinb_full, n = 1000)
testZeroInflation(res_zinbfull)
plot(res_nb) ; plot(res_zinb1) ; plot(res_zinbfull) 


AIC(m_nb, m_zinb1, m_zinb_full) # zero inflation really depends on relationship between LU * Order
BIC(m_nb, m_zinb1, m_zinb_full)


glmmTMB::diagnose(m_zinb1)
glmmTMB::diagnose(m_zinb_full)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)

hymenoptera_coleoptera <- droplevels(hymenoptera_coleoptera)

# glmm.hp does not like the * syntax in the model
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon + Predominant_land_use:Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# Check the breakdown of the df
with(hymenoptera_coleoptera, addmargins(table(Predominant_land_use, Higher_taxon)))
mf <- model.frame(m_nb)
mf %>% filter(Predominant_land_use == "Urban") %>% filter(Higher_taxon == "Coleoptera") %>% View()

# Run the hierarchical partitioning
glmm.hp(mod = m_nb)

# glmm.hp() refits lots of submodels.
# Some of your submodels can’t be estimated (rank deficiency, singular convergence).
# That makes model matrices inconsistent → the function crashes.
# The fix is to simplify the model for HP, especially the ZI formula and/or remove the interaction.

rm(m_zinb_full, m_zinb1, res_nb, res_zinb1, res_zinbfull)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# "Based on everything I've estimated, what is the predicted abundance for each land-use type within each higher taxon?"
emm <- emmeans(
  m_nb,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(hymenoptera_coleoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
hymenoptera_coleoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_nb)
```



# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
hym_col_ss <-  model_data_sr_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hymenoptera_coleoptera <- model_data_sr_integer_only %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% droplevels()


```

# Inspect the richness data
```{r}
sr <- hymenoptera_coleoptera$SpeciesRichness

summary(sr) 
sum(is.na(sr)) # number of NAs
sum(sr == 0, na.rm = TRUE) # number of zeros 889/2655 = 33%

# What proportion of my sr values are not whole numbers?
mean(abs(sr - round(sr)) > 1e-8, na.rm = TRUE)        
# round(sr) - rounds each sr value to the nearest integer, integers stay the same, non-integers get changed
# sr - round(sr) - subtract the rounded value from the original sr value element by element, if sr was already an integer -> 0, if x was not an integer -> non-zero value
# abs() takes the absolute value (removes minus sign)
# > 1e8 is the same as > 0.000000001

mean(sr, na.rm = TRUE) # 2.89
var(sr, na.rm = TRUE) # 146.6804

# var > mean, overdispersion - poisson will struggle, use negative binomial.

# Fraction of zeros?
mean(hymenoptera_coleoptera$SpeciesRichness == 0, na.rm = TRUE) # 33%
```


# Start the modelling
```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

glmmTMB::diagnose(m_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(hymenoptera_coleoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
hymenoptera_coleoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
col_hym <- ggarrange(hymenoptera_coleoptera_plot_ab, hymenoptera_coleoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("../figures/C2_latest_figs/col_hym.png", col_hym, height = 10, width = 10)
```



# Hymenoptera versus Diptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
dip_hym_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_hymenoptera <- model_data_ab_integer_only %>% filter(SS %in% dip_hym_ss) %>% filter(Order %in% c("Diptera", "Hymenoptera")) %>% droplevels()

# Check the breakdown of ants, bees and wasps based on family
# everything which isn't Formicidae (ants) or the 9 families of bee, is therefore a wasp
within_hymenoptera <-
  diptera_hymenoptera %>%
  filter(Higher_taxon == "Hymenoptera") %>% # Subset abundance data for hymenoptera
  mutate(Which_group = 
           case_when(
             Family == "Formicidae" ~ "Ants",
             Family == "Andrenidae" | Family == "Apidae" | Family == "Colletidae" | Family == "Dasypodaidae" |
               Family == "Halictidae" | Family == "Megachilidae" | Family == "Meganomiidae" | Family == "Melittidae"|
               Family == "Stenotritidae" ~ "Bees",   
             Family != "Formicidae" | Family != "Andrenidae" | Family != "Apidae" | Family != "Colletidae" |
               Family != "Dasypodaidae" | Family != "Halictidae" | Family != "Megachilidae" | Family != "Meganomiidae" |
               Family != "Melittidae" | Family != "Stenotritidae" ~ "Wasps")) %>%
  droplevels()

plot_data <- within_hymenoptera %>%
  count(Which_group) %>%
  mutate(
    prop = n / sum(n),
    label = paste0(Which_group, "\n", n, " (", round(prop * 100, 1), "%)")
  )

ggplot(plot_data, aes(x = "", y = n, fill = Which_group)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3.5) +
  labs(
    title = paste0("Diptera versus Hymenoptera (n = ", nrow(within_hymenoptera), ")"),
    fill = "Which group"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

rm(diptera_hymenoptera, within_hymenoptera, plot_data)
```

# Modelling

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # p < 0.05
testDispersion(res_nb) # dispersion ok
plot(res_nb)

# NB with simple ZI
m_zinb1 <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ 1,
  family = nbinom2(),
  data = diptera_hymenoptera
)

# Does adding simple ZI fix the zero problem?
res_zinb1 <- simulateResiduals(m_zinb1, n = 1000)
testZeroInflation(res_zinb1) # zero inflation much better now
testDispersion(res_zinb1)
plot(res_zinb1)

AIC(m_nb, m_zinb1) # m_zinb1 much better
BIC(m_nb, m_zinb1) # m_zinb1 much better

# Extract significance values
Anova(m_zinb1, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_zinb1) # R2m and R2c

rm(m_nb, m_sr, m_zinb2, res_nb, res_zinb1, res_zinb2)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_zinb1,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(diptera_hymenoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.

```{r}
diptera_hymenoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_zinb1)
```


# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
dip_hym_ss <-  model_data_sr_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_hymenoptera <- model_data_sr_integer_only %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Diptera", "Hymenoptera")) %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

 # generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(diptera_hymenoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
diptera_hymenoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
dip_hym <- ggarrange(diptera_hymenoptera_plot_ab, diptera_hymenoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/dip_hym.png", dip_hym, height = 10, width = 10)

rm(dip_hym, diptera_hymenoptera, diptera_hymenoptera_plot_ab, diptera_hymenoptera_plot_sr)
```



# Hemiptera versus Hymenoptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
hemi_hym_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hemiptera_hymenoptera <- model_data_ab_integer_only %>% filter(SS %in% hemi_hym_ss) %>% filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% droplevels()

# There are only 4 observations for each order in pasture
table(hemiptera_hymenoptera$Higher_taxon, hemiptera_hymenoptera$Predominant_land_use) 

# Lets remove pasture as there are too few rows to permit robust modelling
hemiptera_hymenoptera <- hemiptera_hymenoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Confirm "Pasture" is gone
levels(hemiptera_hymenoptera$Predominant_land_use)
table(hemiptera_hymenoptera$Predominant_land_use)

# Check the breakdown of ants, bees and wasps based on family
# everything which isn't Formicidae (ants) or the 9 families of bee, is therefore a wasp
within_hymenoptera <-
  hemiptera_hymenoptera %>%
  filter(Higher_taxon == "Hymenoptera") %>% # Subset abundance data for hymenoptera
  mutate(Which_group = 
           case_when(
             Family == "Formicidae" ~ "Ants",
             Family == "Andrenidae" | Family == "Apidae" | Family == "Colletidae" | Family == "Dasypodaidae" |
               Family == "Halictidae" | Family == "Megachilidae" | Family == "Meganomiidae" | Family == "Melittidae"|
               Family == "Stenotritidae" ~ "Bees",   
             Family != "Formicidae" | Family != "Andrenidae" | Family != "Apidae" | Family != "Colletidae" |
               Family != "Dasypodaidae" | Family != "Halictidae" | Family != "Megachilidae" | Family != "Meganomiidae" |
               Family != "Melittidae" | Family != "Stenotritidae" ~ "Wasps")) %>%
  droplevels()

group_cols <- c(
  "Ants"        = "#E41A1C",
  "Wasps"        = "#377EB8",
  "Bees"       = "#4DAF4A"
)

plot_data <- within_hymenoptera %>%
  count(Which_group) %>%
  mutate(
    prop = n / sum(n),
    label = paste0(Which_group, "\n", n, " (", round(prop * 100, 1), "%)")
  )

ggplot(plot_data, aes(x = "", y = n, fill = Which_group)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3.5) +
  labs(
    title = paste0("Hemiptera versus Hymenoptera (n = ", nrow(within_hymenoptera), ")"),
    fill = "Which group"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  ) +
   scale_fill_manual(values = group_cols, drop = FALSE)

rm(hemiptera_hymenoptera, within_hymenoptera, plot_data)
```

# Modelling 

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)

rm(res_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_nb,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(hemiptera_hymenoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.

```{r}
hemiptera_hymenoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows)
```


# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
#Extract the studies and create the model data
hemi_hym_ss <-  model_data_sr_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hemiptera_hymenoptera <- model_data_sr_integer_only %>% filter(SS %in% hemi_hym_ss) %>% filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% droplevels()

# Lets remove pasture as there are too few rows to permit robust modelling
hemiptera_hymenoptera <- hemiptera_hymenoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Confirm "Pasture" is gone
levels(hemiptera_hymenoptera$Predominant_land_use)
table(hemiptera_hymenoptera$Predominant_land_use)
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

 # generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(hemiptera_hymenoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
hemiptera_hymenoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
hemi_hym <- ggarrange(hemiptera_hymenoptera_plot_ab, hemiptera_hymenoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
ggsave("../figures/C2_latest_figs/hemi_hym.png", hemi_hym, height = 10, width = 10)

rm(hemi_hym, hemiptera_hymenoptera, hemiptera_hymenoptera_plot_ab, hemiptera_hymenoptera_plot_sr, m_nb)
```


# Lepidoptera versus Hymenoptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
lep_hym_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Lepidoptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
lepidoptera_hymenoptera <- model_data_ab_integer_only %>% filter(SS %in% lep_hym_ss) %>% filter(Order %in% c("Lepidoptera", "Hymenoptera")) %>% droplevels()

# There are only 1 observations for each order in plantation, 7 in urban might be ok
table(lepidoptera_hymenoptera$Higher_taxon, lepidoptera_hymenoptera$Predominant_land_use) 

# Lets remove plantation as there are too few rows to permit robust modelling
lepidoptera_hymenoptera <- lepidoptera_hymenoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()

# Check the breakdown of ants, bees and wasps based on family
# everything which isn't Formicidae (ants) or the 9 families of bee, is therefore a wasp
within_hymenoptera <-
  lepidoptera_hymenoptera %>%
  filter(Higher_taxon == "Hymenoptera") %>% # Subset abundance data for hymenoptera
  mutate(Which_group = 
           case_when(
             Family == "Formicidae" ~ "Ants",
             Family == "Andrenidae" | Family == "Apidae" | Family == "Colletidae" | Family == "Dasypodaidae" |
               Family == "Halictidae" | Family == "Megachilidae" | Family == "Meganomiidae" | Family == "Melittidae"|
               Family == "Stenotritidae" ~ "Bees",   
             Family != "Formicidae" | Family != "Andrenidae" | Family != "Apidae" | Family != "Colletidae" |
               Family != "Dasypodaidae" | Family != "Halictidae" | Family != "Megachilidae" | Family != "Meganomiidae" |
               Family != "Melittidae" | Family != "Stenotritidae" ~ "Wasps")) %>%
  droplevels()

plot_data <- within_hymenoptera %>%
  count(Sampling_method) %>%
  mutate(
    prop = n / sum(n),
    label = paste0(Sampling_method, "\n", n, " (", round(prop * 100, 1), "%)")
  )

ggplot(plot_data, aes(x = "", y = n, fill = Which_group)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3.5) +
  labs(
    title = paste0("Lepidoptera versus Hymenoptera (n = ", nrow(within_hymenoptera), ")"),
    fill = "Which group"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

table(within_hymenoptera$Which_group, within_hymenoptera$Predominant_land_use)

rm(lepidoptera_hymenoptera, within_hymenoptera, plot_data)
```

# Modelling 

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)

rm(res_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_nb,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(lepidoptera_hymenoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.

```{r}
lepidoptera_hymenoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows)
```


# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
# Model data
lepidoptera_hymenoptera <- model_data_sr_integer_only %>% filter(SS %in% lep_hym_ss) %>% filter(Order %in% c("Lepidoptera", "Hymenoptera")) %>% droplevels()

# Lets remove pasture as there are too few rows to permit robust modelling
lepidoptera_hymenoptera <- lepidoptera_hymenoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

 # generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(lepidoptera_hymenoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
lepidoptera_hymenoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
lep_hym <- ggarrange(lepidoptera_hymenoptera_plot_ab, lepidoptera_hymenoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/lep_hym.png", lep_hym, height = 10, width = 10)

rm(lep_hym, lepidoptera_hymenoptera, lepidoptera_hymenoptera_plot_ab, lepidoptera_hymenoptera_plot_sr)
```


# Coleoptera versus Diptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
col_dip_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Diptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
coleoptera_diptera <- model_data_ab_integer_only %>% filter(SS %in% col_dip_ss) %>% filter(Order %in% c("Coleoptera", "Diptera")) %>% droplevels()

# sample design looks good
table(coleoptera_diptera$Higher_taxon, coleoptera_diptera$Predominant_land_use)


plot_sampling <- coleoptera_diptera %>%
  count(Sampling_method) %>%
  mutate(
    prop = n / sum(n),
    label = paste0(Sampling_method, "\n", n, " (", round(prop * 100, 1), "%)")
  )

ggplot(plot_sampling, aes(x = "", y = n, fill = Sampling_method)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3.5) +
  labs(
    title = "Distribution of Sampling Methods",
    subtitle = paste0("Total observations (n = ", nrow(within_hymenoptera), ")"),
    fill = "Sampling method"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5)
  )


table(coleoptera_diptera$Sampling_method, coleoptera_diptera$Predominant_land_use)
```


```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_nb,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(coleoptera_diptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.

```{r}
coleoptera_diptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows)
```


# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
# Model data
coleoptera_diptera <- model_data_sr_integer_only %>% filter(SS %in% col_dip_ss) %>% filter(Order %in% c("Coleoptera", "Diptera")) %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

 # generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(coleoptera_diptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
coleoptera_diptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
col_dip <- ggarrange(coleoptera_diptera_plot_ab, coleoptera_diptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/col_dip.png", col_dip, height = 10, width = 10)

rm(col_dip, coleoptera_diptera, coleoptera_diptera_plot_ab, coleoptera_diptera_plot_sr, m_nb, res_nb)
```


# Coleoptera versus Hemiptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
col_hemi_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Hemiptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
coleoptera_hemiptera <- model_data_ab_integer_only %>% filter(SS %in% col_hemi_ss) %>% filter(Order %in% c("Coleoptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data (n = 4)
table(coleoptera_hemiptera$Higher_taxon, coleoptera_hemiptera$Predominant_land_use) 
coleoptera_hemiptera <- coleoptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_nb,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(coleoptera_hemiptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.

```{r}
coleoptera_hemiptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows)
```


# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
# Model data
coleoptera_hemiptera <- model_data_sr_integer_only %>% filter(SS %in% col_hemi_ss) %>% filter(Order %in% c("Coleoptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data (n = 4)
coleoptera_hemiptera <- coleoptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

 # generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(coleoptera_hemiptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
coleoptera_hemiptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
col_hemi <- ggarrange(coleoptera_hemiptera_plot_ab, coleoptera_hemiptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/col_hemi.png", col_hemi, height = 10, width = 10)

rm(col_hemi, coleoptera_hemiptera, coleoptera_hemiptera_plot_ab, coleoptera_hemiptera_plot_sr, m_nb, res_nb)
```




# Coleoptera versus Lepidoptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
col_lep_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
coleoptera_lepidoptera <- model_data_ab_integer_only %>% filter(SS %in% col_lep_ss) %>% filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% droplevels()

# Remove pasture as has too few data (n = 1)
table(coleoptera_lepidoptera$Higher_taxon, coleoptera_lepidoptera$Predominant_land_use) 
coleoptera_lepidoptera <- coleoptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()

# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_nb,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(coleoptera_lepidoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.

```{r}
coleoptera_lepidoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows)
```


# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
# Model data
coleoptera_lepidoptera <- model_data_sr_integer_only %>% filter(SS %in% col_lep_ss) %>% filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data
coleoptera_lepidoptera <- coleoptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

 # generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(coleoptera_lepidoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
coleoptera_lepidoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
col_lep <- ggarrange(coleoptera_lepidoptera_plot_ab, coleoptera_lepidoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/col_lep.png", col_lep, height = 10, width = 10)

rm(col_lep, coleoptera_lepidoptera, coleoptera_lepidoptera_plot_ab, coleoptera_lepidoptera_plot_sr, m_nb, res_nb)
```



# Diptera versus Hemiptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
dip_hemi_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hemiptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_hemiptera <- model_data_ab_integer_only %>% filter(SS %in% dip_hemi_ss) %>% filter(Order %in% c("Diptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data (n = 1)
table(diptera_hemiptera$Higher_taxon, diptera_hemiptera$Predominant_land_use) 
diptera_hemiptera <- diptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_nb,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(diptera_hemiptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.

```{r}
diptera_hemiptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows)
```


# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
# Model data
diptera_hemiptera <- model_data_sr_integer_only %>% filter(SS %in% dip_hemi_ss) %>% filter(Order %in% c("Diptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data
diptera_hemiptera <- diptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

 # generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(diptera_hemiptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
diptera_hemiptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
dip_hemi <- ggarrange(diptera_hemiptera_plot_ab, diptera_hemiptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/dip_hemi.png", dip_hemi, height = 10, width = 10)

rm(dip_hemi, diptera_hemiptera, diptera_hemiptera_plot_ab, diptera_hemiptera_plot_sr, m_nb, res_nb)
```



# Diptera versus Lepidoptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
dip_lep_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_lepidoptera <- model_data_ab_integer_only %>% filter(SS %in% dip_lep_ss) %>% filter(Order %in% c("Diptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data (n = 1)
table(diptera_lepidoptera$Higher_taxon, diptera_lepidoptera$Predominant_land_use) 
diptera_lepidoptera <- diptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()

# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +  (1 | SS), # had to remove Source_ID as model wouldn't converge
  family = nbinom2(),
  data = diptera_lepidoptera
)

summary(m_nb)

 # Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_nb,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(diptera_lepidoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.

```{r}
diptera_lepidoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows)
```


# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
# Model data
diptera_lepidoptera <- model_data_sr_integer_only %>% filter(SS %in% dip_lep_ss) %>% filter(Order %in% c("Diptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data
diptera_lepidoptera <- diptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_lepidoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

 # generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(diptera_lepidoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
diptera_lepidoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
dip_lep <- ggarrange(diptera_lepidoptera_plot_ab, diptera_lepidoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/dip_lep.png", dip_lep, height = 10, width = 10)

rm(dip_lep, diptera_lepidoptera, diptera_lepidoptera_plot_ab, diptera_lepidoptera_plot_sr, m_nb, res_nb)
```


# Hemiptera versus Lepidoptera

# Lets run the abundance model first 
```{r}
#Extract the studies and create the model data
hemi_lep_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hemiptera_lepidoptera <- model_data_ab_integer_only %>% filter(SS %in% hemi_lep_ss) %>% filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data (n = 1)
table(hemiptera_lepidoptera$Higher_taxon, hemiptera_lepidoptera$Predominant_land_use) 
hemiptera_lepidoptera <- hemiptera_lepidoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +  (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

summary(m_nb)

 # Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_nb,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

# generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(hemiptera_lepidoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.

```{r}
hemiptera_lepidoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, comparisons, comparisons_df, emm, primary_rows)
```


# Now lets run the species richness models

# Extract the studies and create the model data
```{r}
# Model data
hemiptera_lepidoptera <- model_data_sr_integer_only %>% filter(SS %in% hemi_lep_ss) %>% filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% droplevels()

# Remove pasture as has too few data
hemiptera_lepidoptera <- hemiptera_lepidoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
```

#### Simulate and prepare the fitted values for each Order 

```{r}
# generate predictions
emm <- emmeans(
  m_sr,
  ~ Predominant_land_use | Higher_taxon,
  type = "link"   # log scale for NB (link scale)
)

 # generate comparisons relative to primary
comparisons <- contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "Primary"
)

# Turn it into a df
comparisons_df <- as.data.frame(summary(comparisons, infer = c(TRUE, TRUE), level = 0.95))  # adds 95% CI

# Extract land-use name from contrast column
comparisons_df <- comparisons_df %>%
  mutate(
    Predominant_land_use = str_remove(contrast, " - Primary")) %>% 
  rename(
    lower.CL = asymp.LCL, # rename for merging
    upper.CL = asymp.UCL
  )


# Create a row for primary with the estimates set to zero as it is the baseline
primary_rows <- expand.grid(
  Predominant_land_use = "Primary",
  Higher_taxon = levels(hemiptera_lepidoptera$Higher_taxon),
  stringsAsFactors = FALSE
) %>%
  mutate(
    contrast = "Primary - Primary",
    estimate = 0,
    SE = NA_real_,
    df = Inf,
    lower.CL = 0,
    upper.CL = 0,
    z.ratio = NA_real_,
    p.value = NA_real_
  )

# merge for plotting
combined <- bind_rows(
  primary_rows,
  comparisons_df
) %>%
  arrange(Higher_taxon, Predominant_land_use)
```

#### Create the plot, with the fits on the log scale and setting primary to zero and plotting the coefficients for each of the land uses as departures from primary.
```{r}
hemiptera_lepidoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=comb_plot_limits)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

#### Arrrange two plots so they abbut and save them
```{r}
hemi_lep <- ggarrange(hemiptera_lepidoptera_plot_ab, hemiptera_lepidoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/hemi_lep.png", hemi_lep, height = 10, width = 10)
rm(hemi_lep, hemiptera_lepidoptera, hemiptera_lepidoptera_plot_ab, hemiptera_lepidoptera_plot_sr, m_nb, res_nb)
```

# Comparing the integer to non integer data

```{r}
model_data_ab %>% dplyr::select(SS, SSBS, SSBST, Diversity_metric, Diversity_metric_unit, Diversity_metric_type, merged_diversity, TotalAbundance, logAbundance) %>% View()

# only individuals are count data type

integer_abundance <- model_data_ab %>% filter(Diversity_metric_unit == "individuals") %>% droplevels()
non_integer_abundance <- model_data_ab %>% filter(Diversity_metric_unit != "individuals") %>% droplevels()

non_integer_abundance %>% dplyr::select(SS, SSBS, SSBST, Diversity_metric, Diversity_metric_unit, Diversity_metric_type, merged_diversity, TotalAbundance, logAbundance) %>% View()

# Each row is an SSBST, therefore each observation/record is an SSBST

length(unique(non_integer_abundance$SSBST)) # 338 SSBSTs are non-integer
length(unique(model_data_ab$SSBST)) 

132/2551 # 5% of sites are non count
338/5381 # 6% of observations are non count

individuals_per_plot <- model_data_ab %>% dplyr::filter(Diversity_metric_unit == "individuals/plot") %>% dplyr::select(SS, SSBS, SSBST, Diversity_metric_unit, merged_diversity, TotalAbundance, logAbundance) # non-integer

length(unique(model_data_ab$SS)) # 98 studies
length(unique(model_data_ab$SSBS)) # 2551 sites overall
length(unique(model_data_ab$SSBST)) # 5381 ssbst overall

length(unique(non_integer_abundance$SS)) # 10 studies
length(unique(non_integer_abundance$SSBS)) # 132 sites
length(unique(non_integer_abundance$SSBST)) # 338 ssbst

length(unique(integer_abundance$SS)) # 88 studies
length(unique(integer_abundance$SSBS)) # 2419 sites
length(unique(integer_abundance$SSBST)) # 5043 ssbst

# 5381 - 5043 = 338
# 2551 - 132 = 2419
# 98 - 10 = 88
```

# Create a function that returns the number of unique studies and sites for each pairwise comparison for the integer only abundance data
```{r}
filter_pairwise_comparisons <- function(integer_abundance, taxon1, taxon2) {
  
  # Find studies that contain both taxa
  which_ss <- integer_abundance %>% 
    dplyr::select(SS, Higher_taxon) %>% 
    filter(Higher_taxon %in% c(taxon1, taxon2)) %>% 
    distinct() %>% 
    group_by(SS) %>% 
    tally() %>% 
    filter(n == 2) %>% 
    pull(SS) %>% 
    droplevels()
  
  # Filter original data to those studies and taxa
  filtered_data <- integer_abundance %>% 
    filter(
      SS %in% which_ss,
      Higher_taxon %in% c(taxon1, taxon2)
    ) %>% 
    droplevels()
  
  # Return counts
  list(
    n_SS = length(unique(filtered_data$SS)),
    n_SSBST = length(unique(filtered_data$SSBST))
  )
}
```

# Lets calculate the number of SSBSTs and SS for each pairwise comparison now that the non-integer data have been removed 

```{r}
# Coleoptera vs Hymenoptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Coleoptera", taxon2 = "Hymenoptera")

# Diptera vs Hymenoptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Diptera", taxon2 = "Hymenoptera")

# Hemiptera vs Hymenoptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Hemiptera", taxon2 = "Hymenoptera")

# Hymenoptera vs Lepidoptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Hymenoptera", taxon2 = "Lepidoptera")

# Coleoptera vs Diptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Coleoptera", taxon2 = "Diptera")

# Coleoptera vs Hemiptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Coleoptera", taxon2 = "Hemiptera")

# Coleoptera vs Lepidoptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Coleoptera", taxon2 = "Lepidoptera")

# Diptera vs Hemiptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Diptera", taxon2 = "Hemiptera")

# Diptera vs Lepidoptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Diptera", taxon2 = "Lepidoptera")

# Hemiptera vs Lepidoptera
filter_pairwise_comparisons(integer_abundance, taxon1= "Hemiptera", taxon2 = "Lepidoptera")

```

`🔴 “Within-group deviations from uniformity significant”` - this means that within each order, the residuals are not behaving as the model expects
In plain terms:
* the Gaussian model is not fully capturing the distribution of logAbundance, especially within each group
This is common when:
- the data are zero-heavy
- the variance structure differs between groups
- the transformation only partially fixes skewness

`🔴 “Levene Test for homogeneity of variance significant”` - this means that the spread of residuals differs between Coleoptera and Hymenoptera
In other words
- one taxon has more variation than the other
- the model assumes equal residual variance and that assumption is violated.

`dispformula = ~ 0 + Predominant_land_use` means allows different residual variance by land use, ie urban sites are allowed to be less variable than cropland
You could switch to a distribution that naturally allows unequal variance, for abundance you would usually go with negative binomial (if counts) or tweedie (if continuous + zeros). These families allow variance to increase with the mean. My abundance data are strongly zero-inflated, are non-integer values with heteroskedasticity among land uses.
The outlier test tells you if the residuals are driven by a few extreme sites. 

Given the non-integer abundance values and the strong zero inflation: ChatGPT is recommending that I adopt a two-part hurdle model
I can't use a negative binomial/poisson as the data are not integers. Under a binomial distribution, the zeros are modelled explicitly, then positive abundance values have their own distribution and land use can affect both processes differently. 
Step 1: model presence/absence, does this species occur at this site? (binomial)
Does land use affect whether insects are present at all?
Many sites have zero individuals.
Step 2: If the species is present, how many are there? (model abundance conditional on presence)
Given that insects are present, how abundant are they?

My diagnostics are telling me 
"Abundance values in my sites are driven by presence/absence first, then magnitude"

`KS test p = 0, deviation significant` tells me that the residuals are not behaving as they should under the fitted model. 
`ziformula = ~ 1` zero inflation constant across all observations
`ziformula = ~ Predominant_land_use * Higher_taxon` zero inflation depends on predictors

`LogAbundance` is a continuous log-transformed response, I could fit a zero-inflated Gaussian model, but..
A zero-inflated model is meant for responses that are zero (like counts), my `LogAbundance` values are not true zeros in abundance anymore, they're zeros on the log scale created by add a study specific constant (adding `to_add`) and taking the `log()` so "zero inflation" in logAbundance is not the same concept, the extra zeros that are in the un-transformed abundance data are piled up at `log(to_add)` before transformation, that is not the same as a zero-inflated count data (i.e. species richness). 

`LogAbundance` does not have a meaningful exact zero value and so does not treat zeros as real zeros>?

ChatGPT recommend that I account for the "extra zeros" on the un-transformed scale?
I could model zeros on the original scale, not on the log scale.
For integer counts: zero-inflated negative binomial.
For continuous with zeros: zero-inflated Tweedie or two-part hurdle model.

I tried to model unequal variance using `dispformula = ~ 0 + Predominant_land_use` since residual plots suggested heteroskedasticity (different spreads across LUs) - which helped but still encountering `KS test p = 0, deviation significant`.

What `ziformaula` is in plain language?
In glmmTMB, a zero-inflated model assumes there are two processes generating the response:
1. A zero-inflation process -> decides whether an observation is a structural zero
2. A conditional (main) model -> generates values when the observation is not a structural zero

A zero-inflated model assumes two ways to get a zero:
1. A structural zero (e.g. truly absent/unsuitable habitat)
2. A sampling zero from the main abundance distribution (NB, Poisson, Tweedie, etc)

What `ziformula = ~ 1` specifically means?
The probability of being a structural zero is the same for every observation.

When `ziformula = ~ 1` makes sense?
This is appropriate when:
The response has many exact zeros
Zeros are structural (true absence)
You don't yet know (or don't want to model) what predicts zeros.
Classic examples:
Insect counts with many traps catching nothing!

Why its not appropriate for `LogAbundance`?
`LogAbundance` is not on the original scale, zeros in abundance were transformed into `log(0 + to add)`

With logAbundance (a continuous, log-transformed response), you can fit a zero-inflated Gaussian model in glmmTMB, but there’s an important catch:
A zero-inflated model is meant for responses that can literally be 0 (like counts).
Your `logAbundance` values are not “true zeros in abundance” anymore — they’re zeros (or near-zeros) on the log scale created by adding `to_add` and taking `log()`.

So “zero inflation” in `logAbundance` is usually not the right concept. The extra zeros you have in the original abundance become a pile-up at log(to_add) after transformation — that’s not the same as a zero-inflated count process.

# Comparing lme4 and glmmTMB (species richness)

```{r}
hc_m2 <-lme4::lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

m_nb1 <- glmmTMB::glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)


# Compare lmer coefficients to zi 
coef(summary(hc_m2)) ; coef(summary(m_nb1))

# Simulate residuals for both models
nb1_sim <- simulateResiduals(m_nb1, n = 1000)
hc2_sim <- simulateResiduals(hc_m2, n = 1000)
  
plot(nb1_sim) ; plot(hc2_sim) # nb looks to have better distributed residuals

testDispersion(nb1_sim) ; testDispersion(hc2_sim) # dispersion ok for both
testZeroInflation(nb1_sim) ; testZeroInflation(hc2_sim) # zero inflation ok for negative binomial, serious issue for normal
testOutliers(nb1_sim, type = "bootstrap") # outliers ok for negative binomial, p = 0.12
testOutliers(hc2_sim, type = "bootstrap") # outliers ok for normal, p = 0.08

DHARMa::plotResiduals(nb1_sim, form = hymenoptera_coleoptera$Predominant_land_use) ; DHARMa::plotResiduals(hc_m2, form = hymenoptera_coleoptera$Predominant_land_use) 
DHARMa::plotResiduals(nb1_sim, form = hymenoptera_coleoptera$Higher_taxon) ; DHARMa::plotResiduals(hc_m2, form = hymenoptera_coleoptera$Higher_taxon) 

r.squaredGLMM(hc_m2)
r.squaredGLMM(m_nb1)

# Compare relevelled model

# relevel model with Hymenoptera as reference level
hymenoptera_coleoptera <- within(hymenoptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Coleoptera")) %>% droplevels()

hc_m2 <-lme4::lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

m_nb1 <- glmmTMB::glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

summary(hc_m2) ; summary(m_nb1)

# Compare coefficients
plot_models(
  hc_m2,
  m_nb1,
  transform = NULL,  # keep on model scale
  show.values = TRUE,
  show.p = FALSE,
  legend.title = "Model"
)


```

# Assessing the distribution of the non-zero values

```{r}
# Remove zero values
sr_nonzero <- hymenoptera_coleoptera$SpeciesRichness[hymenoptera_coleoptera$SpeciesRichness > 0]
ab_nonzero <- hymenoptera_coleoptera$logAbundance[hymenoptera_coleoptera$logAbundance > 0]

hist(sr_nonzero)
hist(ab_nonzero)

qqnorm(sr_nonzero, main = "Q-Q plot: Species richness (non-zero)")
qqline(sr_nonzero, col = "red")

qqnorm(ab_nonzero, main = "Q-Q plot: Log abundance (non-zero)")
qqline(ab_nonzero, col = "red")

# Non-zero values are still quite right skewed/heavy tails?
# Gaussian assumptions are violated even after removing zeros?

# Removing zeros and checking the remaining distribution helps you see whether non-normality is caused only by zeros or is inherent in the data
```


“DHARMa issued a standard warning that the default binomial outlier test can be overly sensitive for integer-valued count data. A bootstrap-based outlier test confirmed no substantial deviation.”“Residual diagnostics (DHARMa) indicated no remaining overdispersion, zero inflation, or influential outliers.”

Should I be worrried about `KS test: p = 0, deviation significant`
Short answer: no
DHARMa converts residuals into numbers between 0 and 1
If the model fit the data perfectly, these numbers would be uniformly distributed
The KS test checks:
"Do these residuals look like they came from a perfectly uniform distribution?
If p < 0.05, it says:
"I can detect a difference from perfect uniformity"

WHy my KS test is p = 0, even though the model is OK!
Reason 1: very large sample size (n = 2,891)
Reason 2: discrete (integer) data
Even though DHARMa smooths residuals via simulaation:
* integer jumps create slight "steps" in the distrubtion
* these tiny steps are enough to trigger the KS data
This is expected for count data!
Reason 3: heterogeneity in the data
My data comes from many studies, sites and LUs!

How to tell when the KS test IS a problem?
You should worry if:
KS test is significant AND residuals vs predicted show strong patterns AND dispersion or zero-inflation tests fail AND group-wise residiual plots look bad!
I do not need to reject the model because of the KS test alone!

“Although the DHARMa KS test indicated a deviation from perfect uniformity, other diagnostics showed no overdispersion, zero inflation, or influential outliers, suggesting the negative binomial model adequately captured the data structure.”

KS p = 0 does NOT mean the model is invalid

You’ve done exactly what a careful analyst should do:
checked assumptions, understood warnings, and interpreted them in context.

# Comparing lme4 and glmmTMB (log abundance)

```{r}
mean(hymenoptera_coleoptera$logAbundance == 0, na.rm = TRUE) # 30%!

hc_ab <-lme4::lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

hc_m1 <- glmmTMB(
  logAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  data = hymenoptera_coleoptera,
  family = gaussian()
)

summary(hc_ab) ; summary(hc_m1) # these two are the same which makes sense

# DHARMa residual diagnostics (works well with glmmTMB)
hc_ab_sim <- simulateResiduals(hc_ab, n = 1000)
hc_m1_sim <- simulateResiduals(hc_m1, n = 1000)

plot(hc_ab_sim) ; plot(hc_m1_sim) # they're the same
testDispersion(hc_ab_sim) ; testDispersion(hc_m1_sim) # overdispersion ok - they're the same
testZeroInflation(hc_ab_sim) # zero inflation a problem

# Try to address zero inflation
hc_zi1 <- glmmTMB(
  logAbundance ~ Predominant_land_use * Higher_taxon +
    (1 | Source_ID) + (1 | SS),
  ziformula = ~ Predominant_land_use * Higher_taxon, # zero inflation depends on both predictors
  family = gaussian(),
  data = hymenoptera_coleoptera
)

# Compare lmer coefficients to zi 
coef(summary(hc_ab)) ; summary(hc_zi1)$coefficients

# Compare coefficients
plot_models(
  hc_ab,
  hc_zi1,
  transform = NULL,  # keep on model scale
  show.values = TRUE,
  show.p = FALSE,
  legend.title = "Model"
)

# Simulate zi model
hc_zi1_sim <- simulateResiduals(hc_zi1, n = 1000)
plot(hc_ab_sim) ; plot(hc_zi1_sim) # zi model has better distributed residuals
testDispersion(hc_ab_sim) ; testDispersion(hc_zi1_sim) # overdispersion ok 
testZeroInflation(hc_ab_sim) ; testZeroInflation(hc_zi1_sim) # zero inflation STILL a problem

# Compare residuals between lme4 and zi glmmTMB
DHARMa::plotResiduals(hc_ab_sim, form = hymenoptera_coleoptera$Predominant_land_use) ; DHARMa::plotResiduals(hc_zi1_sim, form = hymenoptera_coleoptera$Predominant_land_use)  
# a lot of outliers in urban in lme4

DHARMa::plotResiduals(hc_ab_sim, form = hymenoptera_coleoptera$logAbundance) ; DHARMa::plotResiduals(hc_zi1_sim, form = hymenoptera_coleoptera$logAbundance)  

DHARMa::plotResiduals(hc_ab_sim, form = hymenoptera_coleoptera$Higher_taxon) ; DHARMa::plotResiduals(hc_zi1_sim, form = hymenoptera_coleoptera$Higher_taxon) 

# Gaussian mixed model is struggling, residuals are non-uniform, variance differes between taxa - assumptions violated
# Residual variance improved for zero inflated model - allowing a dispersion parameter helped with unequal varaiance..
```