---
title: " Matched comparisons highlight divergent impacts of land-use change on the dominant orders of insects"
author: "Justin Ephraim Isip"
date: "2026-01-13"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Overview
This document reruns Chapter 2 pairwise comparisons among focal insect orders using `glmmTMB`.
Abundance and species richness are analysed separately using negative-binomial mixed models, with zero-inflation included where warranted by diagnostics.

# Load packages

```{r load-packages}
library(tidyverse)
library(glmmTMB)
library(DHARMa)
library(car)
library(MuMIn)
library(emmeans)
library(ggpubr)
library(here)
library(performance)
library(glmm.hp)
```

# Load data

```{r load-data}
model_data_ab <- readRDS(here::here("data", "model_data_ab.rds"))
model_data_sr <- readRDS(here::here("data", "model_data_sr.rds"))
```

# Prepare data

```{r include = FALSE}
model_data_ab <- model_data_ab %>%
  mutate(across(c(SS, SSB, SSBS), as.factor))

model_data_sr <- model_data_sr %>%
  mutate(across(c(SS, SSB, SSBS), as.factor))

# Land use levels
land_use_levels <- c("Primary", "Secondary vegetation", "Plantation forest", "Pasture", "Cropland", "Urban") 

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#197BBD", Coleoptera = "#581845") 

bee_families <- c(
  "Andrenidae", "Apidae", "Colletidae", "Dasypodaidae",
  "Halictidae", "Megachilidae", "Meganomiidae",
  "Melittidae", "Stenotritidae"
)
```

# Handling non-integer count data 

```{r}
# Are there non count data in model_data_ab?
unique(levels(model_data_ab$Diversity_metric_unit)) # yes

# First lets filter for integer only data
model_data_ab_integer_only <- model_data_ab %>% filter(Diversity_metric_unit == "individuals") %>% droplevels()

# Check again
unique(levels(model_data_ab_integer_only$Diversity_metric_unit))

# Are there non count data in model_data_sr?
unique(levels(model_data_sr$Diversity_metric_unit)) # yes

# First lets filter for integer only data
model_data_sr_integer_only <- model_data_sr %>% filter(Diversity_metric_unit == "individuals") %>% droplevels()

# Check again
unique(levels(model_data_sr_integer_only$Diversity_metric_unit))

# Great now both the abundance and species richness model data frames are for count data only
# But are there still non-integers despite the fact we've filtered only for individuals?

# If you subtract each total abundance value in my data frame from the rounded version of the same value, is that value less than 0.000000001? (across all values)
all(abs(model_data_ab_integer_only$TotalAbundance -
        round(model_data_ab_integer_only$TotalAbundance)) < 1e-8,
    na.rm = TRUE) # FALSE

all(abs(model_data_sr_integer_only$SpeciesRichness -
        round(model_data_sr_integer_only$SpeciesRichness)) < 1e-8,
    na.rm = TRUE) # TRUE

# There are still non-integers in the abundance data, lets filter for them

# Check which values are non-integers
still_non_integer_abundance <- model_data_ab_integer_only %>%
  dplyr::filter(abs(TotalAbundance - round(TotalAbundance)) > 1e-8) %>% dplyr::select(SS, SSBS, SSBST, Best_guess_binomial, Diversity_metric, Diversity_metric_unit, Diversity_metric_type,Effort_corrected_measurement, Corrected_sampling_effort,  merged_diversity, TotalAbundance, SpeciesRichness, to_add, logAbundance) %>%  droplevels()

# Which studies?
unique(still_non_integer_abundance$SS)

# Ok lets round all of the Total Abundance values to the nearest integer, any value greater than 0.5 should become 1!
still_non_integer_abundance <- still_non_integer_abundance %>%
  mutate(
    RoundedTotalAbundance = if_else(
      TotalAbundance >= 0.5,
      ceiling(TotalAbundance),
      0
    )
  )

# Ok lets merge these rounded values back into the overall df for those SSBSTs
rounded_abundances <- still_non_integer_abundance %>%  dplyr::select(SSBST, RoundedTotalAbundance) # 407 SSBSTs have had their abundances rounded

# Merge into original data frame
model_data_ab_integer_only <- model_data_ab_integer_only %>% 
    left_join(rounded_abundances, by = ("SSBST")) %>%  # join the values in by their ssbst
  mutate(
    TotalAbundance = coalesce(RoundedTotalAbundance, TotalAbundance) # if a rounded values exists, assign it to TotalAbundance otherwise keep the original value
  ) %>% 
  dplyr::select(-RoundedTotalAbundance)

# Ok now all the abundance values should be integers now that the non-integers have been rounded
all(abs(model_data_ab_integer_only$TotalAbundance -
        round(model_data_ab_integer_only$TotalAbundance)) < 1e-8,
    na.rm = TRUE) # TRUE - all the values in TotalAbundance are integers

# Remove old data frames
rm(model_data_ab, model_data_sr, still_non_integer_abundance, rounded_abundances)
```

# Helper function

```{r}
extract_emmeans_contrasts <- function(model, data) {
  emm <- emmeans(
    model,
    ~ Predominant_land_use | Higher_taxon,
    type = "link"
  )

  comparisons <- contrast(
    emm,
    method = "trt.vs.ctrl",
    ref = "Primary"
  )

  comparisons_df <- as.data.frame(
    summary(comparisons, infer = c(TRUE, TRUE), level = 0.95)
  ) %>%
    mutate(
      Predominant_land_use = stringr::str_remove(contrast, " - Primary")
    ) %>%
    rename(
      lower.CL = asymp.LCL,
      upper.CL = asymp.UCL
    )

  primary_rows <- expand.grid(
    Predominant_land_use = "Primary",
    Higher_taxon = levels(data$Higher_taxon),
    stringsAsFactors = FALSE
  ) %>%
    mutate(
      contrast = "Primary - Primary",
      estimate = 0,
      SE = NA_real_,
      df = Inf,
      lower.CL = 0,
      upper.CL = 0,
      z.ratio = NA_real_,
      p.value = NA_real_
    )

  bind_rows(primary_rows, comparisons_df) %>%
    arrange(Higher_taxon, Predominant_land_use)
}

```


# Hymenoptera versus Coleoptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
hym_col_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hymenoptera_coleoptera <- model_data_ab_integer_only %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% droplevels()
```

# Inspect Hymenoptera composition

```{r}
# Check the breakdown of ants, bees and wasps based on family
# everything which isn't Formicidae (ants) or the 9 families of bee, is therefore a wasp
within_hymenoptera <- hymenoptera_coleoptera %>%
  filter(Higher_taxon == "Hymenoptera") %>%
  mutate(
    Which_group = case_when(
      Family == "Formicidae" ~ "Ants",
      Family %in% bee_families ~ "Bees",
      TRUE ~ "Wasps"
    )
  ) %>%
  droplevels()

plot_data <- within_hymenoptera %>%
  count(Which_group) %>%
  mutate(
    prop = n / sum(n),
    label = paste0(Which_group, "\n", n, " (", round(prop * 100, 1), "%)")
  )

ggplot(plot_data, aes(x = "", y = n, fill = Which_group)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3.5) +
  labs(
    title = paste0("Coleoptera versus Hymenoptera (n = ", nrow(within_hymenoptera), ")"),
    fill = "Which group"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

rm(within_hymenoptera, plot_data)
```

# Abundance models

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# NB with simple ZI
m_zinb1 <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ 1,
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# NB with ZI by LU
m_zinb2 <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ Predominant_land_use,
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# NB with ZI by higher taxon
m_zinb3 <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ Higher_taxon,
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# NB wih complex ZI
m_zinb_full <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ Predominant_land_use * Higher_taxon,
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

summary(m_nb) ; summary(m_zinb1) ; summary(m_zinb_full)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # p < 0.05
testDispersion(res_nb)
plot(res_nb)

# Does adding simple ZI fix the zero problem?
res_zinb1 <- simulateResiduals(m_zinb1, n = 1000)
testZeroInflation(res_zinb1)
plot(res_zinb1)

# ZI by LU
res_zinb2 <- simulateResiduals(m_zinb2, n = 1000)
testZeroInflation(res_zinb2)
plot(res_zinb2)

summary(m_zinb2)

#  ZI by higher taxon
res_zinb3 <- simulateResiduals(m_zinb3, n = 1000)
testZeroInflation(res_zinb3)
plot(res_zinb3)

summary(m_zinb3)

# Does adding complex ZI fix the zero problem?
res_zinbfull <- simulateResiduals(m_zinb_full, n = 1000)
testZeroInflation(res_zinbfull)
plot(res_nb) ; plot(res_zinb1) ; plot(res_zinbfull) 


AIC(m_nb, m_zinb1, m_zinb_full) # zero inflation really depends on relationship between LU * Order
BIC(m_nb, m_zinb1, m_zinb_full)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)

hymenoptera_coleoptera <- droplevels(hymenoptera_coleoptera)

# glmm.hp does not like the * syntax in the model
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use + Higher_taxon + Predominant_land_use:Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

# Run the hierarchical partitioning
glmm.hp(mod = m_nb)

rm(m_zinb2, m_zinb3, m_zinb_full, m_zinb1, res_nb, res_zinb1, res_zinbfull, res_zinb2, res_zinb3)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  hymenoptera_coleoptera
)
```

# Create the plot

```{r}
hymenoptera_coleoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, m_nb)
```


# Species-richness models

# # Create pairwise data

```{r}
hym_col_ss <-  model_data_sr_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hymenoptera_coleoptera <- model_data_sr_integer_only %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Hymenoptera", "Coleoptera")) %>% droplevels()
```

# Inspect the richness data

```{r}
sr <- hymenoptera_coleoptera$SpeciesRichness

summary(sr) 
sum(is.na(sr)) # number of NAs
sum(sr == 0, na.rm = TRUE) # number of zeros 889/2655 = 33%

# What proportion of my sr values are not whole numbers?
mean(abs(sr - round(sr)) > 1e-8, na.rm = TRUE)        
# round(sr) - rounds each sr value to the nearest integer, integers stay the same, non-integers get changed
# sr - round(sr) - subtract the rounded value from the original sr value element by element, if sr was already an integer -> 0, if x was not an integer -> non-zero value
# abs() takes the absolute value (removes minus sign)
# > 1e8 is the same as > 0.000000001

mean(sr, na.rm = TRUE) # 2.89
var(sr, na.rm = TRUE) # 146.6804

# var > mean, overdispersion - poisson will struggle, use negative binomial.

# Fraction of zeros?
mean(hymenoptera_coleoptera$SpeciesRichness == 0, na.rm = TRUE) # 33%
```

# Start the modelling

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hymenoptera_coleoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

glmmTMB::diagnose(m_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  hymenoptera_coleoptera
)
```

# Create the plot

```{r}
hymenoptera_coleoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, m_sr)
```

# Arrange plots and save them 

```{r}
col_hym <- ggarrange(hymenoptera_coleoptera_plot_ab, hymenoptera_coleoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/col_hym.png", col_hym, height = 10, width = 10)

rm(col_hym, hymenoptera_coleoptera_plot_ab, hymenoptera_coleoptera_plot_sr, hymenoptera_coleoptera)
```



# Hymenoptera versus Diptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
dip_hym_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_hymenoptera <- model_data_ab_integer_only %>% filter(SS %in% dip_hym_ss) %>% filter(Order %in% c("Diptera", "Hymenoptera")) %>% droplevels()
```

# Inspect Hymenoptera composition

```{r}
# Check the breakdown of ants, bees and wasps based on family
# everything which isn't Formicidae (ants) or the 9 families of bee, is therefore a wasp
within_hymenoptera <- diptera_hymenoptera %>%
  filter(Higher_taxon == "Hymenoptera") %>%
  mutate(
    Which_group = case_when(
      Family == "Formicidae" ~ "Ants",
      Family %in% bee_families ~ "Bees",
      TRUE ~ "Wasps"
    )
  ) %>%
  droplevels()

plot_data <- within_hymenoptera %>%
  count(Which_group) %>%
  mutate(
    prop = n / sum(n),
    label = paste0(Which_group, "\n", n, " (", round(prop * 100, 1), "%)")
  )

ggplot(plot_data, aes(x = "", y = n, fill = Which_group)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3.5) +
  labs(
    title = paste0("Diptera versus Hymenoptera (n = ", nrow(within_hymenoptera), ")"),
    fill = "Which group"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

rm(within_hymenoptera, plot_data)
```

# Abundance models

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # p < 0.05
testDispersion(res_nb) # dispersion ok
plot(res_nb)

# NB with simple ZI
m_zinb1 <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  ziformula = ~ 1,
  family = nbinom2(),
  data = diptera_hymenoptera
)

# Does adding simple ZI fix the zero problem?
res_zinb1 <- simulateResiduals(m_zinb1, n = 1000)
testZeroInflation(res_zinb1) # zero inflation much better now
testDispersion(res_zinb1)
plot(res_zinb1)

AIC(m_nb, m_zinb1) # m_zinb1 much better
BIC(m_nb, m_zinb1) # m_zinb1 much better

# Extract significance values
Anova(m_zinb1, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_zinb1) # R2m and R2c

rm(m_nb, m_sr, m_zinb2, res_nb, res_zinb1, res_zinb2)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  diptera_hymenoptera
)
```

# Create the plot

```{r}
diptera_hymenoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, m_nb, m_zinb1)
```


# Species-richness models

# Create pairwise data

```{r}
dip_hym_ss <-  model_data_sr_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_hymenoptera <- model_data_sr_integer_only %>% filter(SS %in% hym_col_ss) %>% filter(Order %in% c("Diptera", "Hymenoptera")) %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hymenoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  diptera_hymenoptera
)
```

# Create the plot
```{r}
diptera_hymenoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, m_sr)
```

# Arrange plots and save them 
```{r}
dip_hym <- ggarrange(diptera_hymenoptera_plot_ab, diptera_hymenoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/dip_hym.png", dip_hym, height = 10, width = 10)

rm(dip_hym, diptera_hymenoptera, diptera_hymenoptera_plot_ab, diptera_hymenoptera_plot_sr)
```



# Hemiptera versus Hymenoptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
hemi_hym_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hemiptera_hymenoptera <- model_data_ab_integer_only %>% filter(SS %in% hemi_hym_ss) %>% filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% droplevels()

# There are only 4 observations for each order in pasture
table(hemiptera_hymenoptera$Higher_taxon, hemiptera_hymenoptera$Predominant_land_use) 

# Lets remove pasture as there are too few rows to permit robust modelling
hemiptera_hymenoptera <- hemiptera_hymenoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Confirm "Pasture" is gone
levels(hemiptera_hymenoptera$Predominant_land_use)
table(hemiptera_hymenoptera$Predominant_land_use)
```

# Inspect Hymenoptera composition

```{r}
# Check the breakdown of ants, bees and wasps based on family
# everything which isn't Formicidae (ants) or the 9 families of bee, is therefore a wasp
within_hymenoptera <- hemiptera_hymenoptera %>%
  filter(Higher_taxon == "Hymenoptera") %>%
  mutate(
    Which_group = case_when(
      Family == "Formicidae" ~ "Ants",
      Family %in% bee_families ~ "Bees",
      TRUE ~ "Wasps"
    )
  ) %>%
  droplevels()

group_cols <- c(
  "Ants"        = "#E41A1C",
  "Wasps"        = "#377EB8",
  "Bees"       = "#4DAF4A"
)

plot_data <- within_hymenoptera %>%
  count(Which_group) %>%
  mutate(
    prop = n / sum(n),
    label = paste0(Which_group, "\n", n, " (", round(prop * 100, 1), "%)")
  )

ggplot(plot_data, aes(x = "", y = n, fill = Which_group)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3.5) +
  labs(
    title = paste0("Hemiptera versus Hymenoptera (n = ", nrow(within_hymenoptera), ")"),
    fill = "Which group"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  ) +
   scale_fill_manual(values = group_cols, drop = FALSE)

rm(within_hymenoptera, plot_data)
```

# Abundance models 

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)

rm(res_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  hemiptera_hymenoptera
)
```

# Create the plot

```{r}
hemiptera_hymenoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, m_nb)
```


# Species-richness models

# Create pairwise data

```{r}
#Extract the studies and create the model data
hemi_hym_ss <-  model_data_sr_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hemiptera_hymenoptera <- model_data_sr_integer_only %>% filter(SS %in% hemi_hym_ss) %>% filter(Order %in% c("Hemiptera", "Hymenoptera")) %>% droplevels()

# Lets remove pasture as there are too few rows to permit robust modelling
hemiptera_hymenoptera <- hemiptera_hymenoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()

# Confirm "Pasture" is gone
levels(hemiptera_hymenoptera$Predominant_land_use)
table(hemiptera_hymenoptera$Predominant_land_use)
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_hymenoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  hemiptera_hymenoptera
)
```

# Create the plot

```{r}
hemiptera_hymenoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, m_sr)
```

# Arrange plots and save them 
```{r}
hemi_hym <- ggarrange(hemiptera_hymenoptera_plot_ab, hemiptera_hymenoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
# ggsave("../figures/C2_latest_figs/hemi_hym.png", hemi_hym, height = 10, width = 10)

rm(hemi_hym, hemiptera_hymenoptera, hemiptera_hymenoptera_plot_ab, hemiptera_hymenoptera_plot_sr)
```


# Lepidoptera versus Hymenoptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
lep_hym_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Lepidoptera", "Hymenoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
lepidoptera_hymenoptera <- model_data_ab_integer_only %>% filter(SS %in% lep_hym_ss) %>% filter(Order %in% c("Lepidoptera", "Hymenoptera")) %>% droplevels()

# There are only 1 observations for each order in plantation, 7 in urban might be ok
table(lepidoptera_hymenoptera$Higher_taxon, lepidoptera_hymenoptera$Predominant_land_use) 

# Lets remove plantation as there are too few rows to permit robust modelling
lepidoptera_hymenoptera <- lepidoptera_hymenoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()
```

# Inspect Hymenoptera composition

```{r}
# Check the breakdown of ants, bees and wasps based on family
# everything which isn't Formicidae (ants) or the 9 families of bee, is therefore a wasp
within_hymenoptera <- lepidoptera_hymenoptera %>%
  filter(Higher_taxon == "Hymenoptera") %>%
  mutate(
    Which_group = case_when(
      Family == "Formicidae" ~ "Ants",
      Family %in% bee_families ~ "Bees",
      TRUE ~ "Wasps"
    )
  ) %>%
  droplevels()

plot_data <- within_hymenoptera %>%
  count(Which_group) %>%
  mutate(
    prop = n / sum(n),
    label = paste0(Which_group, "\n", n, " (", round(prop * 100, 1), "%)")
  )

ggplot(plot_data, aes(x = "", y = n, fill = Which_group)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 3.5) +
  labs(
    title = paste0("Lepidoptera versus Hymenoptera (n = ", nrow(within_hymenoptera), ")"),
    fill = "Which group"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

table(within_hymenoptera$Which_group, within_hymenoptera$Predominant_land_use)

rm(within_hymenoptera, plot_data)
```

# Abundance models 

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)

rm(res_nb)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  lepidoptera_hymenoptera
)
```

# Create the plot

```{r}
lepidoptera_hymenoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, m_nb)
```


# Species-richness models

# Create pairwise data

```{r}
# Model data
lepidoptera_hymenoptera <- model_data_sr_integer_only %>% filter(SS %in% lep_hym_ss) %>% filter(Order %in% c("Lepidoptera", "Hymenoptera")) %>% droplevels()

# Lets remove plantation forest as there are too few rows to permit robust modelling
lepidoptera_hymenoptera <- lepidoptera_hymenoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = lepidoptera_hymenoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  lepidoptera_hymenoptera
)
```

# Create the plot

```{r}
lepidoptera_hymenoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, m_sr)
```

#### Arrange plots and save them 
```{r}
lep_hym <- ggarrange(lepidoptera_hymenoptera_plot_ab, lepidoptera_hymenoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/lep_hym.png", lep_hym, height = 10, width = 10)

rm(lep_hym, lepidoptera_hymenoptera, lepidoptera_hymenoptera_plot_ab, lepidoptera_hymenoptera_plot_sr)
```




# Coleoptera versus Diptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
col_dip_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Diptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
coleoptera_diptera <- model_data_ab_integer_only %>% filter(SS %in% col_dip_ss) %>% filter(Order %in% c("Coleoptera", "Diptera")) %>% droplevels()

# sample design looks good
table(coleoptera_diptera$Higher_taxon, coleoptera_diptera$Predominant_land_use)
```

# Abundance models

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

#### Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  coleoptera_diptera
)
```

# Create the plot

```{r}
coleoptera_diptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, m_nb)
```


# Species-richness models

# Create pairwise data

```{r}
# Model data
coleoptera_diptera <- model_data_sr_integer_only %>% filter(SS %in% col_dip_ss) %>% filter(Order %in% c("Coleoptera", "Diptera")) %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_diptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  coleoptera_diptera
)
```

# Create the plot

```{r}
coleoptera_diptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, m_sr)
```

# Arrange plots and save them 

```{r}
col_dip <- ggarrange(coleoptera_diptera_plot_ab, coleoptera_diptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/col_dip.png", col_dip, height = 10, width = 10)

rm(col_dip, coleoptera_diptera, coleoptera_diptera_plot_ab, coleoptera_diptera_plot_sr)
```


# Coleoptera versus Hemiptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
col_hemi_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Hemiptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
coleoptera_hemiptera <- model_data_ab_integer_only %>% filter(SS %in% col_hemi_ss) %>% filter(Order %in% c("Coleoptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data (n = 4)
table(coleoptera_hemiptera$Higher_taxon, coleoptera_hemiptera$Predominant_land_use) 
coleoptera_hemiptera <- coleoptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()
```

# Abundance models

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  coleoptera_hemiptera
)
```

# Create the plot

```{r}
coleoptera_hemiptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, m_nb)
```


# Species-richness models

# Create pairwise data

```{r}
# Model data
coleoptera_hemiptera <- model_data_sr_integer_only %>% filter(SS %in% col_hemi_ss) %>% filter(Order %in% c("Coleoptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data (n = 4)
coleoptera_hemiptera <- coleoptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_hemiptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  coleoptera_hemiptera
)
```

# Create the plot
```{r}
coleoptera_hemiptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, m_sr)
```

# Arrange plots and save them 
```{r}
col_hemi <- ggarrange(coleoptera_hemiptera_plot_ab, coleoptera_hemiptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/col_hemi.png", col_hemi, height = 10, width = 10)

rm(col_hemi, coleoptera_hemiptera, coleoptera_hemiptera_plot_ab, coleoptera_hemiptera_plot_sr, res_nb)
```




# Coleoptera versus Lepidoptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
col_lep_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
coleoptera_lepidoptera <- model_data_ab_integer_only %>% filter(SS %in% col_lep_ss) %>% filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% droplevels()

# Remove pasture as has too few data (n = 1)
table(coleoptera_lepidoptera$Higher_taxon, coleoptera_lepidoptera$Predominant_land_use) 
coleoptera_lepidoptera <- coleoptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()
```

# Abundance model

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  coleoptera_lepidoptera
)
```

# Create the plot

```{r}
coleoptera_lepidoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined, m_nb)
```


# Species-richness models

# Create pairwise data

```{r}
# Model data
coleoptera_lepidoptera <- model_data_sr_integer_only %>% filter(SS %in% col_lep_ss) %>% filter(Order %in% c("Coleoptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data
coleoptera_lepidoptera <- coleoptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = coleoptera_lepidoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
performance::r2(m_sr)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  coleoptera_lepidoptera
)
```

# Create the plot

```{r}
coleoptera_lepidoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, m_sr)
```

# Arrange plots and save them 
```{r}
col_lep <- ggarrange(coleoptera_lepidoptera_plot_ab, coleoptera_lepidoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/col_lep.png", col_lep, height = 10, width = 10)

rm(col_lep, coleoptera_lepidoptera, coleoptera_lepidoptera_plot_ab, coleoptera_lepidoptera_plot_sr, res_nb)
```



# Diptera versus Hemiptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
dip_hemi_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Hemiptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_hemiptera <- model_data_ab_integer_only %>% filter(SS %in% dip_hemi_ss) %>% filter(Order %in% c("Diptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data (n = 1)
table(diptera_hemiptera$Higher_taxon, diptera_hemiptera$Predominant_land_use) 
diptera_hemiptera <- diptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()
```

# Abundance models

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

# Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  diptera_hemiptera
)
```

# Create the plot

```{r}
diptera_hemiptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined)
```


# Species-richness models

# Create pairwise data

```{r}
# Model data
diptera_hemiptera <- model_data_sr_integer_only %>% filter(SS %in% dip_hemi_ss) %>% filter(Order %in% c("Diptera", "Hemiptera")) %>% droplevels()

# Remove pasture as has too few data
diptera_hemiptera <- diptera_hemiptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_hemiptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  diptera_hemiptera
)
```

# Create the plot

```{r}
diptera_hemiptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, m_sr, res_sr)
```

# Arrrange two plots so they abbut and save them

```{r}
dip_hemi <- ggarrange(diptera_hemiptera_plot_ab, diptera_hemiptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/dip_hemi.png", dip_hemi, height = 10, width = 10)

rm(dip_hemi, diptera_hemiptera, diptera_hemiptera_plot_ab, diptera_hemiptera_plot_sr, m_nb, res_nb)
```



# Diptera versus Lepidoptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
dip_lep_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Diptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
diptera_lepidoptera <- model_data_ab_integer_only %>% filter(SS %in% dip_lep_ss) %>% filter(Order %in% c("Diptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data (n = 1)
table(diptera_lepidoptera$Higher_taxon, diptera_lepidoptera$Predominant_land_use) 
diptera_lepidoptera <- diptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()
```

# Abundance models 

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +  (1 | SS), # had to remove Source_ID as model wouldn't converge
  family = nbinom2(),
  data = diptera_lepidoptera
)

summary(m_nb)

 # Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  diptera_lepidoptera
)
```

# Create the plot

```{r}
diptera_lepidoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined)
```


# Species-richness models

# Create pairwise data

```{r}
# Model data
diptera_lepidoptera <- model_data_sr_integer_only %>% filter(SS %in% dip_lep_ss) %>% filter(Order %in% c("Diptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data
diptera_lepidoptera <- diptera_lepidoptera %>% filter(Predominant_land_use != "Plantation forest") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = diptera_lepidoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  diptera_lepidoptera
)
```

# Create the plot

```{r}
diptera_lepidoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, comparisons, comparisons_df, emm, primary_rows, m_sr, res_sr)
```

# Arrange plots and save them 
```{r}
dip_lep <- ggarrange(diptera_lepidoptera_plot_ab, diptera_lepidoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/dip_lep.png", dip_lep, height = 10, width = 10)

rm(dip_lep, diptera_lepidoptera, diptera_lepidoptera_plot_ab, diptera_lepidoptera_plot_sr, m_nb, res_nb)
```


# Hemiptera versus Lepidoptera

# Create pairwise data

```{r}
#Extract the studies and create the model data
hemi_lep_ss <-  model_data_ab_integer_only %>% 
  dplyr::select(SS, Order) %>% 
  filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% 
  unique() %>% 
  group_by(SS) %>% 
  tally() %>% 
  filter(n == 2) %>% 
  pull(SS) %>% 
  unique() %>% 
  droplevels()

# Model data
hemiptera_lepidoptera <- model_data_ab_integer_only %>% filter(SS %in% hemi_lep_ss) %>% filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% droplevels()

# Remove plantation as has too few data (n = 1)
table(hemiptera_lepidoptera$Higher_taxon, hemiptera_lepidoptera$Predominant_land_use) 
hemiptera_lepidoptera <- hemiptera_lepidoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()
```

# Abundance models 

```{r}
# NB (no ZI)
m_nb <- glmmTMB(
  TotalAbundance ~ Predominant_land_use * Higher_taxon +  (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

summary(m_nb)

 # Are zeros a problem in the nb model?
res_nb <- simulateResiduals(m_nb, n = 1000)
testZeroInflation(res_nb) # not a problem
testDispersion(res_nb) # not a problem 
plot(res_nb)

# Extract significance values
Anova(m_nb, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_nb) # R2m and R2c
performance::r2(m_nb)
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_nb,
  hemiptera_lepidoptera
)
```

# Create the plot

```{r}
hemiptera_lepidoptera_plot_ab <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 6, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
                 position = position_dodge(width = 0.5), linewidth = 2) +
  theme_classic() +
  scale_x_discrete(limits = land_use_levels) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  ylab("Log (Total abundance difference)") +
  scale_colour_manual(values = group_colours, name = "Order") +
  guides(color = guide_legend(
    override.aes = list(shape = 16, size = 10, linetype = 0)
  )) +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.title = element_text(face = "bold", size = 16),
    legend.text  = element_text(size = 16)
  )

rm(combined)
```


# Species-richness models

# Create pairwise data

```{r}
# Model data
hemiptera_lepidoptera <- model_data_sr_integer_only %>% filter(SS %in% hemi_lep_ss) %>% filter(Order %in% c("Hemiptera", "Lepidoptera")) %>% droplevels()

# Remove pasture as has too few data
hemiptera_lepidoptera <- hemiptera_lepidoptera %>% filter(Predominant_land_use != "Pasture") %>% droplevels()
```

# Run the richness model

```{r}
# NB (no ZI)
m_sr <- glmmTMB(
  SpeciesRichness ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
  family = nbinom2(),
  data = hemiptera_lepidoptera
)

summary(m_sr)

res_sr <- simulateResiduals(m_sr, n = 1000)
testDispersion(res_sr) # dispersion not a problem
testZeroInflation(res_sr) #zero inflation not a problem
plot(res_sr)
testOutliers(res_sr)
testUniformity(res_sr)

# Extract significance values
Anova(m_sr, type = "II") # This produces a Type II Wald χ² (chi-square) test table

# This is the GLMM equivalent of the F-tests I used to report for lmer.
# “There was a significant interaction between land use and taxon (Wald χ²₅ = 77.67, p < 0.001).”

r.squaredGLMM(m_sr) # R2m and R2c
```

# Simulate and prepare the fitted values for each Order 

```{r}
combined <- extract_emmeans_contrasts(
  m_sr,
  hemiptera_lepidoptera
)
```

# Create the plot

```{r}
hemiptera_lepidoptera_plot_sr <- combined %>%
    ggplot()+
    aes(x = Predominant_land_use, y = estimate, colour = Higher_taxon, group = Higher_taxon)+
    geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
    geom_point(size = 6, position = position_dodge(width = 0.5))+
    geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), position = position_dodge(width = 0.5), linewidth = 2)+
    theme_classic()+
    scale_x_discrete(limits=land_use_levels)+
    geom_vline(xintercept = 1.5, linetype = 2)+
    geom_vline(xintercept = 2.5, linetype = 2)+
    geom_vline(xintercept = 3.5, linetype = 2)+
    geom_vline(xintercept = 4.5, linetype = 2)+
    geom_vline(xintercept = 5.5, linetype = 2)+
    theme(axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 16),  
          axis.title.x = element_blank(),
          panel.border = element_rect(colour = "black",  fill=NA))+
    xlab("Land use intensity class")+
    ylab("Log (Species richness difference)") +
    scale_colour_manual(values=group_colours, name = "Order") + # this overrides the colours for the groups as above
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 10, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = )) +
  theme(legend.title = element_text(face = "bold", size = 16),
        legend.text = element_text(size = 16)) 

rm(combined, m_sr, res_sr)
```

# Arrange plots and save them 
```{r}
hemi_lep <- ggarrange(hemiptera_lepidoptera_plot_ab, hemiptera_lepidoptera_plot_sr, nrow = 2, common.legend =  TRUE, legend = "bottom") 
#ggsave("../figures/C2_latest_figs/hemi_lep.png", hemi_lep, height = 10, width = 10)
rm(hemi_lep, hemiptera_lepidoptera, hemiptera_lepidoptera_plot_ab, hemiptera_lepidoptera_plot_sr, m_nb, res_nb)
```


