---
title: "among-orders"
author: "Justin Isip"
date: "2023-02-13"
output: html_document
---

### Clear the workspace
```{r}
rm(list=ls())
```

### Load required packages
```{r}
library(tidyverse) # data processing
library(lme4) # for mixed effects models
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(sjPlot) # for visualising results
library(effects) # for extracting model effects
library(merTools) # useful for a few things, but we're using it for extracting estimates for plotting
library(emmeans) # testing multiple comparisons (but also useful for plotting)
library(coefplot) # visualising model coefficients
library(cli)
library(ggeffects)
library(ggpubr)
```

### Load in model plot function
```{r}

## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(2,2))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}
```

### Load add_order_fold function to split the diversity data into exclusive insect orders 
```{r}

add_order_fold <- function(data, folds = "default", verbose = TRUE){
  #
  # By default (and nothing else is yet coded up), this puts each row into
  # an insect order that each have a decent, number of rows of data. 
  # This allows site-level values to be obtained within
  # each of the order folds in turn.
  
  # Set up order_fold, a factor that holds the fold identity.
  data$order_fold <- rep(NA, nrow(data))
  
  # Rows are assigned to folds based on the higher taxonomy information.
  if (folds == "default"){
    data$order_fold[data$Higher_taxon == "Diptera"] <- "Diptera"
    data$order_fold[data$Higher_taxon == "Coleoptera"] <- "Coleoptera"
    data$order_fold[data$Higher_taxon == "Lepidoptera"] <- "Lepidoptera"
    data$order_fold[data$Higher_taxon == "Hymenoptera"] <- "Hymenoptera"
    data$order_fold[data$Higher_taxon == "Hemiptera"] <- "Hemiptera"
    data$order_fold[data$Higher_taxon == "Arachnida"] <- "Arachnida"
    data$order_fold[data$Higher_taxon == "Odonata" |
                    data$Higher_taxon == "Trichoptera" |
                    data$Higher_taxon == "Ephemeroptera" |
                    data$Higher_taxon == "Plecoptera" ] <- "Aquatics"
    data$order_fold[data$Phylum == "Arthropoda" & 
                      data$Higher_taxon != "Diptera" & 
                      data$Higher_taxon != "Coleoptera" &
                      data$Higher_taxon != "Lepidoptera" &
                      data$Higher_taxon != "Hymenoptera" &
                      data$Higher_taxon != "Hemiptera" &
                      data$Higher_taxon != "Arachnida" &
                      data$Higher_taxon != "Odonata" &
                      data$Higher_taxon != "Trichoptera" &
                      data$Higher_taxon != "Ephemeroptera" &
                      data$Higher_taxon != "Plecoptera"] <- "OtherArthropods"
  }else{
    stop("Only the default folds have been coded up so far!")
  }
  
  data$order_fold <- as.factor(data$order_fold)
  
  if (verbose == TRUE){
    # By default, report on the breakdown of records per fold.
    cat("Numbers of rows within each order fold are as follows:\n\n")
    print(table(data$order_fold))
  }
  
  return(data)
}

```

### Load function to calculate total abundance for each site x order_fold combination
```{r}
# Calculate total abundance for each site --------------------------------------
# Modified to clarify reporting is on site x fold combinations
get_site_abundances_order <- function(data, 
                                order_folds = FALSE,
                                verbose = TRUE){
  # This calculates site metrics for each site or, if desired, each order fold
  # within each site.
  
  # Make it a data frame to avoid having to deal with a tibble.
  data <- as.data.frame(data)
  
  # Set up a variable for uniquely identifying the site x order_fold combination
  # or, if order_folds == FALSE, the site. This is what grouping is then done
  # by.
  print(table(data$order_fold))
  
  if (order_folds == TRUE){
    data$the_order_folds <- data$order_fold
  }else{
    data$the_order_folds <- ""
  }
  
  # Just for debugging
  print(table(data$the_order_folds))
  
  # Start to calculate site-level metrics.
  sites <- data %>%
    
    # pull out only the merged diversity data
    distinct(merge_ID, .keep_all = TRUE) %>%
    
    # Re-make SSB and SSBS values since we've now dropped a bunch of values; and
    # also add SSBST which is that order x site combo. The trimws is because the
    # SSBST would otherwise end in a space if order_folds == FALSE.
    mutate(SS = paste(Source_ID, Study_number),
           SSB = paste(SS, Block),
           SSBS = paste(SSB, Site_number),
           SSBST = trimws(paste(SSBS, the_order_folds))) %>%
    
    # group by SSBST (each unique value corresponds to a unique site x fold)
    group_by(SSBST) %>%
    
    # Now add up all the abundance measurements within each site and also 
    # estimate coverage.
    mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                   sum(merged_diversity, na.rm = TRUE),
                                   # If the diversity metric type isn't 
                                   # Abundance, then leave the TotalAbundance 
                                   # measurement as NA.
                                   NA),
           
           SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                    merged_diversity,
                                    # for abundance and occurrence measurements,
                                    # count the number of unique species names 
                                    # that are present at the site 
                                    n_distinct(Best_guess_binomial[merged_diversity > 0])),
           
           coverage = get.SC2(merged_diversity)) %>%
    
    # ungroup
    ungroup() %>%
    
    # pull out unique site/fold combinations.
    distinct(SSBST, .keep_all = TRUE) %>%
    
    # now group by Study ID and fold.
    group_by(SS, the_order_folds) %>%
    
        # pull out the maximum abundance for each study/fold combination.
    mutate(MaxAbundance = max(TotalAbundance),
           MinNonZero = min(TotalAbundance[TotalAbundance > 0])) %>%
    
    # ungroup.
    ungroup() %>%
    
    # now rescale total abundance, so that within each study, the maximum is 1.
    mutate(RescaledAbundance = TotalAbundance/MaxAbundance) %>%
    
    mutate(sqrtRescaledAbundance = sqrt(RescaledAbundance)) 
  
    # If a study reports abundance as a count, add 1 to all values; otherwise
  # add 1/2 the minimum non-zero value; then log-transform to get a useful
  # logAbundance value that may be more symmetric than sqrtRescaledAbundance.
  sites$to_add <- ifelse(sites$Diversity_metric_unit %in%
                           c("individuals", "times observed"), 1,
                         sites$MinNonZero/2)
  sites$logAbundance <- log(sites$TotalAbundance + sites$to_add)
  
  
  # Remove NA values and drop levels
  sites <- as.data.frame(sites)
  sites <- droplevels(subset(sites, !is.na(sqrtRescaledAbundance)))
  
  if (verbose == TRUE){
    # Summarise structure of site-level data frame
    cat("\nThere are abundance data for:\n")
    cat(paste("   ", sum(!is.na(sites$sqrtRescaledAbundance)), "sites, from\n"))
    cat(paste("   ", 
              length(unique(sites$SS[!is.na(sites$sqrtRescaledAbundance)])), 
              "studies and\n"))
    cat(paste("   ", 
              length(unique(sites$Source_ID[!is.na(sites$sqrtRescaledAbundance)])),
              "sources.\n"))
  }
  
  # Set provenance attributes
  attr(sites, which = "diversity_extract") <- 
    attr(data, which = "diversity_extract")
  attr(sites, which = "is_merged") <- attr(data, which = "is_merged")
  attr(sites, which = "order_folds") <- order_folds 
  attr(sites, which = "when_Created") <- Sys.time()
  
  return(sites)
}

```

### Load all other functions
```{r}
source("/Applications/PhD/Andy/data_prep_functions.R")
source("/Applications/PhD/Andy/site_comparison_functions.R")
source("https://highstat.com/Books/Book2/HighstatLibV10.R") # collinearity
source("/Applications/PhD/Projects/multi-groups/scripts/combine_and_plot_function.R")
source("/Applications/PhD/Projects/multi-groups/scripts/predict_this_model.R")
```

### Read in the data
```{r}
diversity <- readRDS("../data/diversity-2022-04-13-02-33-10.rds")
```

### Correct for sampling effort
```{r}
if(check_sampling_effort_nas(diversity) ==
   FALSE) stop("Some studies have NA for sampling effort in only some sites!")

diversity <- correct_for_sampling_effort(diversity)
print(summarise_diversity(diversity))
```

### Merge sites
```{r}
merged <- merge_sites(diversity, verbose = TRUE) 
```

### Filter merged sites for groups of interest
```{r}
arthropods <-
  merged %>%
  filter(Phylum == "Arthropoda") %>% # filter only for arthropods, so we can include Arachnida
  group_by(SS) %>% # group by SS
  mutate(
    Multiple_groups = 
      case_when(
        (length(unique(Higher_taxon))) > 1 ~ "YES",
        (length(unique(Higher_taxon))) == 1 ~ "NO")) %>% # for each study create a new column called multiple groups = YES/NO - 
  # YES if there's more than one unique higher taxon within that study, NO if not
  # Filter the data for studies with multiple groups
  filter(Multiple_groups == "YES") %>% # filter for only the rows for studies (YES) with multiple groups
  ungroup() %>% 
  droplevels() 
```

### Add_order_fold to allow taxonomic cross-validation and split arthropods into exclusive groups
```{r}
# split arthropods into Aquatics, Arachnida, Coleoptera, Diptera, Hemiptera, Hymenoptera, Lepidoptera and OtherArthropods
arthropods <- add_order_fold(arthropods, verbose = TRUE) # I've sense checked we don't lose rows 
```

### Calculate site level diversity metrics using get_site_abundances_order and order_fold for each site x order combination
```{r}
arthropods_ssbst <- get_site_abundances_order(arthropods, 
                                  order_folds = TRUE, verbose = TRUE)

# There are 13303 sites because there are 13303 rows and each row is a site x order combination 

table(arthropods_ssbst$order_fold) # number of rows for each site x order combination

```

### Subset our df of SSBST (site x order combination) only for the groups of interest
```{r}
arthropods_six_groups <- 
    arthropods_ssbst %>%
    filter(Higher_taxon != "NA") %>% # some rows have unknown species of insects/arthropods
    filter(Higher_taxon %in% c("Arachnida", "Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera")) %>%
    filter(Predominant_habitat != "Cannot decide") # filter out habitats that aren't defined
```

### Fix up your explanatory variables 
```{r}

# Rename predominant habitat since its not really habitat we're looking at
arthropods_six_groups <- rename(arthropods_six_groups,
                Predominant_land_use = Predominant_habitat)


# Relevel the land use and use intensity classes
arthropods_six_groups <- arthropods_six_groups %>%
  
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    
    # indeterminate secondary veg and cannot decide get NA
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(arthropods_six_groups$Predominant_land_use, arthropods_six_groups$Use_intensity)

# Not many MSV intense use sites, primary intense use sites and young secondary light use, might have to collapse light and intense use into one
# But luckily can still keep all 3 secondary veg categories

```

### Complete cases
```{r}
model_data_ab <- drop_na(arthropods_six_groups, 
                         RescaledAbundance, Predominant_land_use,
                         Use_intensity)

model_data_sr <- drop_na(arthropods_six_groups, 
                         SpeciesRichness, Predominant_land_use,
                         Use_intensity)

```

### Prepare the six groups abundance dataframe
```{r}

# Log transform rescaled abundance
model_data_ab$logAbundance <- log(model_data_ab$RescaledAbundance + 1)

# Check the distributions
hist(model_data_ab$RescaledAbundance) ; hist(model_data_ab$logAbundance)

# Have a look at sample sizes
table(model_data_ab$Predominant_land_use, model_data_ab$Use_intensity)

# Collapse intense and light land use intensities to have a more even spread of the data
model_data_ab <- 
  model_data_ab %>%
  mutate(
    Use_intensity = recode_factor(Use_intensity,
                                  "Intense use" = "High_use",
                                  "Light use" = "High_use"),
# reset reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )
# Have a look at sample sizes
table(model_data_ab$Predominant_land_use, model_data_ab$Use_intensity)
```



### Order comparisons for abundance


### Diptera


# diptera_hymenoptera abundance
```{r}
# Remove chubby datasets
rm(arthropods, merged, diversity, arthropods_ssbst)

diptera_hymenoptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Diptera", "Hymenoptera"))

# maximal model
dhy_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hymenoptera)

# remove block
dhy_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# remove study
dhy_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_hymenoptera)
  
# Compare models 
AIC(dhy_m1, dhy_m2, dhy_m3) # dhy_m2 best model

rm(dhy_m1, dhy_m3)

Anova(dhy_m2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dhy_m4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just land use as predictor
dhy_m5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just Higher_taxon as the predictor
dhy_m6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Make a null, intercept only model
dhy_m7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Compare the models
AIC(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7)
model.sel(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7)

# dhy_m2 still the best

# model estimates
summary(dhy_m2)
```
### Simulate diptera_hymenoptera effects
```{r}
# Predict Diptera effects
dip_eff <- predict_this_model(dhy_m2, order = "Diptera")

# relevel model with hymenoptera as reference level
diptera_hymenoptera <- within(diptera_hymenoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model
hym_1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# Predict hymenoptera effects
hym_eff <- predict_this_model(hym_1, order = "Hymenoptera")

# Combine effects and plot them
diptera_hymenoptera_plot_AB <- combine_and_plot(dip_eff, hym_eff, diversity = "Abundance")

rm(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7, hym_1, hym_eff, diptera_hymenoptera, dip_eff)
```

# diptera_coleoptera abundance
```{r}
### Subset our df of SSBST (site x order combination) only for the groups of interest
diptera_coleoptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Diptera", "Coleoptera"))

# maximal model
dc_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_coleoptera)

# remove block
dc_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_coleoptera)

# remove study
dc_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_coleoptera)

# Compare models 
AIC(dc_m1, dc_m2, dc_m3) # dc_m1 best model

rm(dc_m2, dc_m3)

# model estimates
summary(dc_m1)
Anova(dc_m1)
```
### Simulate diptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(dc_m1, order = "Coleoptera")

# relevel model with Diptera as reference level
diptera_coleoptera <- within(diptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Diptera"))

# maximal model
dip_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
    data = diptera_coleoptera)

model_plot(dip_m1)
summary(dip_m1)
Anova(dip_m1)

# Predict Diptera effects - remember that this overwrites previous effects of Diptera model
dip_eff <- predict_this_model(dip_m1, order = "Diptera")

# Combine effects and relevel Diptera as reference
diptera_coleoptera_plot_AB <- combine_and_plot(dip_eff, col_eff, relevel = "Diptera", diversity = "Abundance")

rm(dip_m1, dc_m1, col_eff, diptera_coleoptera, dip_eff)

```

# diptera_hemiptera abundance
```{r}
diptera_hemiptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Diptera", "Hemiptera")) %>%
  droplevels()

# maximal model
dh_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hemiptera)

# remove block
dh_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hemiptera)

# remove study
dh_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_hemiptera)

# Compare models 
AIC(dh_m1, dh_m2, dh_m3) # dh_m2 best model

rm(dh_m1, dh_m3)

# model estimates
summary(dh_m2)
Anova(dh_m2)
```
### Simulate diptera_hemiptera effects
```{r}
# Predict Diptera effects 
dip_eff <- predict_this_model(dh_m2, order = "Diptera")

# relevel model with Hemiptera as reference level
diptera_hemiptera <- within(diptera_hemiptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model with Hemiptera as ref
hemi_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hemiptera)

model_plot(hemi_m1)
summary(hemi_m1)
Anova(hemi_m1)

# Predict Hemiptera effects 
hemi_eff <- predict_this_model(hemi_m1, order = "Hemiptera")

# Combine effects and plot
diptera_hemiptera_plot_AB <- combine_and_plot(dip_eff, hemi_eff, diversity = "Abundance") 

rm(dip_eff, hemi_eff, hemi_m1, dh_m2, diptera_hemiptera)
```

# diptera_lepidoptera abundance
```{r}
diptera_lepidoptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Diptera", "Lepidoptera")) %>%
  droplevels()

# maximal model
dl_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_lepidoptera)

# remove block
dl_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_lepidoptera)

# remove study
dl_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_lepidoptera)

# Compare models 
AIC(dl_m1, dl_m2, dl_m3) # dl_m2 best model

rm(dl_m1, dl_m3)

# model estimates
summary(dl_m2)
Anova(dl_m2)
```
### Simulate diptera_lepidoptera effects
```{r}
# Predict Diptera effects 
dip_eff <- predict_this_model(dl_m2, order = "Diptera")

# relevel model with Lepidoptera as reference level
diptera_lepidoptera <- within(diptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_lepidoptera)

model_plot(lep_m1)
summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
diptera_lepidoptera_plot_AB <- combine_and_plot(dip_eff, lep_eff, diversity = "Abundance")

rm(dip_eff, lep_eff, lep_m1, dl_m2, diptera_lepidoptera)
```


## Hymenoptera

# hymenoptera_coleoptera abundance
```{r}
hymenoptera_coleoptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Hymenoptera", "Coleoptera")) %>%
  droplevels()

# maximal model
hc_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_coleoptera)

# remove block
hc_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

# remove study
hc_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_coleoptera)

# Compare models 
AIC(hc_m1, hc_m2, hc_m3) # hc_m1 best model

rm(hc_m2, hc_m3)

# model estimates
summary(hc_m1)
Anova(hc_m1)
```
### Simulate hymenoptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(hc_m1, order = "Coleoptera")

# relevel model with Hymenoptera as reference level
hymenoptera_coleoptera <- within(hymenoptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model with Hymenoptera as ref
hym_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + + (1 | SSB),
    data = hymenoptera_coleoptera)

model_plot(hym_m1)
summary(hym_m1)
Anova(hym_m1)

# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hym_m1, order = "Hymenoptera")

# Combine effects and plot
hymenoptera_coleoptera_plot_AB <- combine_and_plot(col_eff, hym_eff, diversity = "Abundance")

rm(col_eff, hym_eff, hym_m1, hc_m1, hymenoptera_coleoptera)
```

# hymenoptera_hemiptera abundance
```{r}
hymenoptera_hemiptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Hymenoptera", "Hemiptera")) %>%
  droplevels()

# maximal model
hh_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_hemiptera)

# remove block
hh_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_hemiptera)

# remove study
hh_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_hemiptera)

# Compare models 
AIC(hh_m1, hh_m2, hh_m3) # hh_m2 best model

rm(hh_m1, hh_m3)

# model estimates
summary(hh_m2)
Anova(hh_m2)
```
### Simulate hymenoptera_hemiptera effects
```{r}
# Predict Hemiptera effects 
hem_eff <- predict_this_model(hh_m2, order = "Hemiptera")

# relevel model with Hymenoptera as reference level
hymenoptera_hemiptera <- within(hymenoptera_hemiptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model with Hymenoptera as ref
hym_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_hemiptera)

model_plot(hym_m1)
summary(hym_m1)
Anova(hym_m1)

# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hym_m1, order = "Hymenoptera")

# Combine effects and plot
hymenoptera_hemiptera_plot_AB <- combine_and_plot(hem_eff, hym_eff, diversity = "Abundance")

rm(hem_eff, hym_eff, hym_m1, hh_m2, hymenoptera_hemiptera)
```

# hymenoptera_lepidoptera abundance
```{r}
hymenoptera_lepidoptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Hymenoptera", "Lepidoptera")) %>%
  droplevels()

# maximal model
hl_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_lepidoptera)

# remove block
hl_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_lepidoptera)

# remove study
hl_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_lepidoptera)

# Compare models 
AIC(hl_m1, hl_m2, hl_m3) # hl_m2 best model

rm(hl_m1, hl_m3)

# model estimates
summary(hl_m2)
Anova(hl_m2)
```
### Simulate hymenoptera_lepidoptera effects
```{r}
# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hl_m2, order = "Hymenoptera")

# relevel model with Lepidoptera as reference level
hymenoptera_lepidoptera <- within(hymenoptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_lepidoptera)

summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
hymenoptera_lepidoptera_plot_AB <- combine_and_plot(hym_eff, lep_eff, diversity = "Abundance")

rm(hym_eff, lep_eff, lep_m1, hl_m2, hymenoptera_lepidoptera)
```


## Hemiptera

# hemiptera_lepidoptera abundance
```{r}
hemiptera_lepidoptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Hemiptera", "Lepidoptera")) %>%
  droplevels()

# maximal model
hl_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hemiptera_lepidoptera)

# remove block
hl_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_lepidoptera)

# remove study
hl_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_lepidoptera)

# Compare models 
AIC(hl_m1, hl_m2, hl_m3) # hl_m2 best model

rm(hl_m1, hl_m3)

# model estimates
summary(hl_m2)
Anova(hl_m2)
```
### Simulate hemiptera_lepidoptera effects
```{r}
# Predict Hemiptera effects 
hem_eff <- predict_this_model(hl_m2, order = "Hemiptera")

# relevel model with Lepidoptera as reference level
hemiptera_lepidoptera <- within(hemiptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_lepidoptera)

summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
hemiptera_lepidoptera_plot_AB <- combine_and_plot(hem_eff, lep_eff, diversity = "Abundance")

rm(hem_eff, lep_eff, lep_m1, hl_m2, hemiptera_lepidoptera)
```

# hemiptera_coleoptera abundance
```{r}
hemiptera_coleoptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Hemiptera", "Coleoptera")) %>%
  droplevels()

# maximal model
hc_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hemiptera_coleoptera)

# remove block
hc_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_coleoptera)

# remove study
hc_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_coleoptera)

# Compare models 
AIC(hc_m1, hc_m2, hc_m3) # hc_m1 best model

rm(hc_m2, hc_m3)

# model estimates
summary(hc_m1)
Anova(hc_m1)
```
### Simulate hemiptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(hc_m1, order = "Coleoptera")

# relevel model with Hemiptera as reference level
hemiptera_coleoptera <- within(hemiptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model with Hemiptera as ref
hem_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
    data = hemiptera_coleoptera)

summary(hem_m1)
Anova(hem_m1)

# Predict Hemiptera effects 
hem_eff <- predict_this_model(hem_m1, order = "Hemiptera")

# Combine effects and plot
hemiptera_coleoptera_plot_AB <- combine_and_plot(col_eff, hem_eff, diversity = "Abundance")

rm(hem_eff, col_eff, hem_m1, hc_m1, hemiptera_coleoptera)
```


## Coleoptera

### coleoptera_lepidoptera abundance
```{r}
coleoptera_lepidoptera <-
  model_data_ab %>%
  filter(Higher_taxon %in% c("Coleoptera", "Lepidoptera")) %>%
  droplevels()

# maximal model
cl_m1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = coleoptera_lepidoptera)

# remove block
cl_m2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = coleoptera_lepidoptera)

# remove study
cl_m3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = coleoptera_lepidoptera)

# Compare models 
AIC(cl_m1, cl_m2, cl_m3) # cl_m1 best model

rm(cl_m2, cl_m3)

# model estimates
summary(cl_m1)
Anova(cl_m1)
```
### Simulate coleoptera_lepidoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(cl_m1, order = "Coleoptera")

# relevel model with Lepidoptera as reference level
coleoptera_lepidoptera <- within(coleoptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
    data = coleoptera_lepidoptera)

summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
coleoptera_lepidoptera_plot_AB <- combine_and_plot(col_eff, lep_eff, diversity = "Abundance")

rm(lep_eff, col_eff, lep_m1, cl_m1, coleoptera_lepidoptera)
```



### Order comparisons for Species Richness

### Prepare the model data
```{r}
# Log transform species richness
model_data_sr$logSR <- log(model_data_sr$SpeciesRichness + 1)

# Check the distributions
hist(model_data_sr$SpeciesRichness) ; hist(model_data_sr$logSR) 


# Have a look at sample sizes
table(model_data_sr$Predominant_land_use, model_data_sr$Use_intensity)

# Collapse intense and light land use intensities to have a more even spread of the data
model_data_sr <- 
  model_data_sr %>%
  mutate(
    Use_intensity = recode_factor(Use_intensity,
                                  "Intense use" = "High_use",
                                  "Light use" = "High_use"),
# reset reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )
# Have a look at sample sizes
table(model_data_sr$Predominant_land_use, model_data_sr$Use_intensity)
```


## Diptera

# diptera_hymenoptera species richness
```{r}
diptera_hymenoptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Diptera", "Hymenoptera"))

# maximal model
dhy_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hymenoptera)

# remove block
dhy_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# remove study
dhy_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_hymenoptera)

# Compare models 
AIC(dhy_m1, dhy_m2, dhy_m3) # dhy_m2 best model

rm(dhy_m1, dhy_m3)

Anova(dhy_m2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
dhy_m4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just land use as predictor
dhy_m5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Just Higher_taxon as the predictor
dhy_m6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Make a null, intercept only model
dhy_m7 <- lmer(logSR ~ (1 | Source_ID) + (1|SS), data = diptera_hymenoptera)

## Compare the models
AIC(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7)
model.sel(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7)

# dhy_m2 still the best

# model estimates
summary(dhy_m2)
```
### Simulate diptera_hymenoptera effects
```{r}
# Predict Diptera effects
dip_eff <- predict_this_model(dhy_m2, order = "Diptera")

# relevel model with hymenoptera as reference level
diptera_hymenoptera <- within(diptera_hymenoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model
hym_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hymenoptera)

# Predict hymenoptera effects
hym_eff <- predict_this_model(hym_m1, order = "Hymenoptera")

# Combine effects and plot them
diptera_hymenoptera_plot_SR <- combine_and_plot(dip_eff, hym_eff, diversity = "Species_richness")

rm(dhy_m2, dhy_m4, dhy_m5, dhy_m6, dhy_m7, hym_m1, hym_eff, diptera_hymenoptera, dip_eff)
```


# diptera_coleoptera species richness
```{r}
### Subset our df of SSBST (site x order combination) only for the groups of interest
diptera_coleoptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Diptera", "Coleoptera"))

# maximal model
dc_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_coleoptera)

# remove block
dc_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_coleoptera)

# remove study
dc_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_coleoptera)

# Compare models 
AIC(dc_m1, dc_m2, dc_m3) # dc_m1 best model

rm(dc_m2, dc_m3)

# model estimates
summary(dc_m1)
Anova(dc_m1)
```
### Simulate diptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(dc_m1, order = "Coleoptera")

# relevel model with Diptera as reference level
diptera_coleoptera <- within(diptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Diptera"))

# maximal model
dip_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
    data = diptera_coleoptera)

model_plot(dip_m1)
summary(dip_m1)
Anova(dip_m1)

# Predict Diptera effects - remember that this overwrites previous effects of Diptera model
dip_eff <- predict_this_model(dip_m1, order = "Diptera")

# Combine effects and relevel Diptera as reference
diptera_coleoptera_plot_SR <- combine_and_plot(dip_eff, col_eff, relevel = "Diptera", diversity = "Species_richness")

rm(dip_m1, dc_m1, col_eff, diptera_coleoptera, dip_eff)
```

# diptera_hemiptera species richness
```{r}
diptera_hemiptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Diptera", "Hemiptera")) %>%
  droplevels()

# maximal model
dh_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_hemiptera)

# remove block
dh_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hemiptera)

# remove study
dh_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_hemiptera)

# Compare models 
AIC(dh_m1, dh_m2, dh_m3) # dh_m2 best model

rm(dh_m1, dh_m3)

# model estimates
summary(dh_m2)
Anova(dh_m2)
```
### Simulate diptera_hemiptera effects
```{r}
# Predict Diptera effects 
dip_eff <- predict_this_model(dh_m2, order = "Diptera")

# relevel model with Hemiptera as reference level
diptera_hemiptera <- within(diptera_hemiptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model with Hemiptera as ref
hemi_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_hemiptera)

model_plot(hemi_m1)
summary(hemi_m1)
Anova(hemi_m1)

# Predict Hemiptera effects 
hemi_eff <- predict_this_model(hemi_m1, order = "Hemiptera")

# Combine effects and plot
diptera_hemiptera_plot_SR <- combine_and_plot(dip_eff, hemi_eff, diversity = "Species_richness") 

rm(dip_eff, hemi_eff, hemi_m1, dh_m2, diptera_hemiptera)
```

# diptera_lepidoptera species richness
```{r}
diptera_lepidoptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Diptera", "Lepidoptera")) %>%
  droplevels()

# maximal model
dl_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = diptera_lepidoptera)

# remove block
dl_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_lepidoptera)

# remove study
dl_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = diptera_lepidoptera)

# Compare models 
AIC(dl_m1, dl_m2, dl_m3) # dl_m1 best model

rm(dl_m2, dl_m3)

# model estimates
summary(dl_m1)
Anova(dl_m1)
```
### Simulate diptera_lepidoptera effects
```{r}
# Predict Diptera effects 
dip_eff <- predict_this_model(dl_m1, order = "Diptera")

# relevel model with Lepidoptera as reference level
diptera_lepidoptera <- within(diptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = diptera_lepidoptera)

model_plot(lep_m1)
summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
diptera_lepidoptera_plot_SR <- combine_and_plot(dip_eff, lep_eff, diversity = "Species_richness")

rm(dip_eff, lep_eff, lep_m1, dl_m1, diptera_lepidoptera)
```


## Hymenoptera

# hymenoptera_coleoptera species richness
```{r}
hymenoptera_coleoptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Hymenoptera", "Coleoptera")) %>%
  droplevels()

# maximal model
hc_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_coleoptera)

# remove block
hc_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

# remove study
hc_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_coleoptera)

# Compare models 
AIC(hc_m1, hc_m2, hc_m3) # hc_m2 best model

rm(hc_m1, hc_m3)

# model estimates
summary(hc_m2)
Anova(hc_m2)
```
### Simulate hymenoptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(hc_m2, order = "Coleoptera")

# relevel model with Hymenoptera as reference level
hymenoptera_coleoptera <- within(hymenoptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model with Hymenoptera as ref
hym_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_coleoptera)

model_plot(hym_m1)
summary(hym_m1)
Anova(hym_m1)

# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hym_m1, order = "Hymenoptera")

# Combine effects and plot
hymenoptera_coleoptera_plot_SR <- combine_and_plot(col_eff, hym_eff, diversity = "Species_richness")

rm(col_eff, hym_eff, hym_m1, hc_m2, hymenoptera_coleoptera)
```

# hymenoptera_hemiptera species richness
```{r}
hymenoptera_hemiptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Hymenoptera", "Hemiptera")) %>%
  droplevels()

# maximal model
hh_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_hemiptera)

# remove block
hh_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_hemiptera)

# remove study
hh_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_hemiptera)

# Compare models 
AIC(hh_m1, hh_m2, hh_m3) # hh_m2 best model

rm(hh_m1, hh_m3)

# model estimates
summary(hh_m2)
Anova(hh_m2)
```
### Simulate hymenoptera_hemiptera effects
```{r}
# Predict Hemiptera effects 
hem_eff <- predict_this_model(hh_m2, order = "Hemiptera")

# relevel model with Hymenoptera as reference level
hymenoptera_hemiptera <- within(hymenoptera_hemiptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model with Hymenoptera as ref
hym_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_hemiptera)

model_plot(hym_m1)
summary(hym_m1)
Anova(hym_m1)

# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hym_m1, order = "Hymenoptera")

# Combine effects and plot
hymenoptera_hemiptera_plot_SR <- combine_and_plot(hem_eff, hym_eff, diversity = "Species_richness")

rm(hem_eff, hym_eff, hym_m1, hh_m2, hymenoptera_hemiptera)
```

# hymenoptera_lepidoptera species_richness
```{r}
hymenoptera_lepidoptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Hymenoptera", "Lepidoptera")) %>%
  droplevels()

# maximal model
hl_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hymenoptera_lepidoptera)

# remove block
hl_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_lepidoptera)

# remove study
hl_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hymenoptera_lepidoptera)

# Compare models 
AIC(hl_m1, hl_m2, hl_m3) # hl_m2 best model

rm(hl_m1, hl_m3)

# model estimates
summary(hl_m2)
Anova(hl_m2)
```
### Simulate hymenoptera_lepidoptera effects
```{r}
# Predict Hymenoptera effects 
hym_eff <- predict_this_model(hl_m2, order = "Hymenoptera")

# relevel model with Lepidoptera as reference level
hymenoptera_lepidoptera <- within(hymenoptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hymenoptera_lepidoptera)

summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
hymenoptera_lepidoptera_plot_SR <- combine_and_plot(hym_eff, lep_eff, diversity = "Species_richness")

rm(hym_eff, lep_eff, lep_m1, hl_m2, hymenoptera_lepidoptera)
```

## Hemiptera

# hemiptera_lepidoptera species richness
```{r}
hemiptera_lepidoptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Hemiptera", "Lepidoptera")) %>%
  droplevels()

# maximal model
hl_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hemiptera_lepidoptera)

# remove block
hl_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_lepidoptera)

# remove study
hl_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_lepidoptera)

# Compare models 
AIC(hl_m1, hl_m2, hl_m3) # hl_m1 best model

rm(hl_m2, hl_m3)

# model estimates
summary(hl_m1)
Anova(hl_m1)
```
### Simulate hemiptera_lepidoptera effects
```{r}
# Predict Hemiptera effects 
hem_eff <- predict_this_model(hl_m1, order = "Hemiptera")

# relevel model with Lepidoptera as reference level
hemiptera_lepidoptera <- within(hemiptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_lepidoptera)

summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
hemiptera_lepidoptera_plot_SR <- combine_and_plot(hem_eff, lep_eff, diversity = "Species_richness")

rm(hem_eff, lep_eff, lep_m1, hl_m1, hemiptera_lepidoptera)
```

# hemiptera_coleoptera species_richness
```{r}
hemiptera_coleoptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Hemiptera", "Coleoptera")) %>%
  droplevels()

# maximal model
hc_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hemiptera_coleoptera)

# remove block
hc_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = hemiptera_coleoptera)

# remove study
hc_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = hemiptera_coleoptera)

# Compare models 
AIC(hc_m1, hc_m2, hc_m3) # hc_m1 best model

rm(hc_m2, hc_m3)

# model estimates
summary(hc_m1)
Anova(hc_m1)
```
### Simulate hemiptera_coleoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(hc_m1, order = "Coleoptera")

# relevel model with Hemiptera as reference level
hemiptera_coleoptera <- within(hemiptera_coleoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model with Hemiptera as ref
hem_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
    data = hemiptera_coleoptera)

summary(hem_m1)
Anova(hem_m1)

# Predict Hemiptera effects 
hem_eff <- predict_this_model(hem_m1, order = "Hemiptera")

# Combine effects and plot
hemiptera_coleoptera_plot_SR <- combine_and_plot(col_eff, hem_eff, diversity = "Species_richness")

rm(hem_eff, col_eff, hem_m1, hc_m1, hemiptera_coleoptera)
```


## Coleoptera

# coleoptera_lepidoptera species richness
```{r}
coleoptera_lepidoptera <-
  model_data_sr %>%
  filter(Higher_taxon %in% c("Coleoptera", "Lepidoptera")) %>%
  droplevels()

# maximal model
cl_m1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = coleoptera_lepidoptera)

# remove block
cl_m2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
    data = coleoptera_lepidoptera)

# remove study
cl_m3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
    data = coleoptera_lepidoptera)

# Compare models 
AIC(cl_m1, cl_m2, cl_m3) # cl_m2 best model

rm(cl_m1, cl_m3)

# model estimates
summary(cl_m2)
Anova(cl_m2)
```
### Simulate coleoptera_lepidoptera effects
```{r}
# Predict Coleoptera effects 
col_eff <- predict_this_model(cl_m2, order = "Coleoptera")

# relevel model with Lepidoptera as reference level
coleoptera_lepidoptera <- within(coleoptera_lepidoptera, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model with Lepidoptera as ref
lep_m1 <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
    data = coleoptera_lepidoptera)

summary(lep_m1)
Anova(lep_m1)

# Predict Lepidoptera effects 
lep_eff <- predict_this_model(lep_m1, order = "Lepidoptera")

# Combine effects and plot
coleoptera_lepidoptera_plot_SR <- combine_and_plot(col_eff, lep_eff, diversity = "Species_richness")

rm(lep_eff, col_eff, lep_m1, cl_m2, coleoptera_lepidoptera)
```

# Calculate sample size for each comparison of orders
```{r}

# Calculate number of rows for either abundance or SR model data depending on the two orders being compared
  model_data_ab %>%
  filter(Higher_taxon %in% c("Coleoptera", "Lepidoptera")) %>%
  nrow()

  model_data_sr %>%
  filter(Higher_taxon %in% c("Coleoptera", "Lepidoptera")) %>%
  nrow()
```

## GGSAVE
```{r}

dip_col <- ggarrange(diptera_coleoptera_plot_AB, diptera_coleoptera_plot_SR, nrow = 2)
ggsave("../figures/dip_col.png", dip_col, height = 6, width = 6)

dip_hemi <- ggarrange(diptera_hemiptera_plot_AB,  diptera_hemiptera_plot_SR, nrow = 2)
ggsave("../figures/dip_hemi.png", dip_hemi, height = 6, width = 6)

dip_hym <- ggarrange(diptera_hymenoptera_plot_AB, diptera_hymenoptera_plot_SR, nrow = 2)
ggsave("../figures/dip_hym.png", dip_hym, height = 6, width = 6)
     
dip_lep <- ggarrange(diptera_lepidoptera_plot_AB, diptera_lepidoptera_plot_SR, nrow = 2)
ggsave("../figures/dip_lep.png", dip_lep, height = 6, width = 6)         

hym_col <- ggarrange(hymenoptera_coleoptera_plot_AB, hymenoptera_coleoptera_plot_SR, nrow = 2)
ggsave("../figures/hym_col.png", hym_col, height = 6, width = 6)    

hym_hemi <- ggarrange(hymenoptera_hemiptera_plot_AB, hymenoptera_hemiptera_plot_SR, nrow = 2)
ggsave("../figures/hym_hemi.png", hym_hemi, height = 6, width = 6)    

hym_lep <- ggarrange(hymenoptera_lepidoptera_plot_AB, hymenoptera_lepidoptera_plot_SR, nrow = 2)
ggsave("../figures/hym_lep.png", hym_lep, height = 6, width = 6)    
        
hemi_col <- ggarrange(hemiptera_coleoptera_plot_AB, hemiptera_coleoptera_plot_SR, nrow = 2)
ggsave("../figures/hemi_col.png", hemi_col, height = 6, width = 6)    
  
hemi_lep <- ggarrange(hemiptera_lepidoptera_plot_AB, hemiptera_lepidoptera_plot_SR, nrow = 2)
ggsave("../figures/hemi_lep.png", hemi_lep, height = 6, width = 6)  

col_lep <- ggarrange(coleoptera_lepidoptera_plot_AB, coleoptera_lepidoptera_plot_SR, nrow = 2)
ggsave("../figures/col_lep.png", col_lep, height = 6, width = 6)  
```
