---
title: "analyses-five-orders"
author: "Justin Isip"
date: "2023-07-14"
output: html_document
---

# This Rmd analyses all of the data for five orders, as opposed to analyses-exactly-five-orders.Rmd which analyses only those studies which contain all five orders. 

### Clear the workspace
```{r}
rm(list=ls())
```

### Load required packages
```{r}
library(tidyverse) # data processing
library(lme4) # for mixed effects models
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(sjPlot) # for visualising results
library(effects) # for extracting model effects
library(merTools) # useful for a few things, but we're using it for extracting estimates for plotting
library(emmeans) # testing multiple comparisons (but also useful for plotting)
library(coefplot) # visualising model coefficients
library(cli)
library(ggeffects)
library(ggpubr)
library(lmerTest)
```

### Load in model plot function
```{r}

## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(2,2))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}
```

### Load add_order_fold function to split the diversity data into exclusive insect orders 
```{r}

add_order_fold <- function(data, folds = "default", verbose = TRUE){
  
  # By default (and nothing else is yet coded up), this puts each row into
  # an insect order that each have a decent, number of rows of data. 
  # This allows site-level values to be obtained within
  # each of the order folds in turn.
  
  # Set up order_fold, a factor that holds the fold identity.
  data$order_fold <- rep(NA, nrow(data))
  
  # Rows are assigned to folds based on the higher taxonomy information.
  if (folds == "default"){
    data$order_fold[data$Higher_taxon == "Diptera"] <- "Diptera"
    data$order_fold[data$Higher_taxon == "Coleoptera"] <- "Coleoptera"
    data$order_fold[data$Higher_taxon == "Lepidoptera"] <- "Lepidoptera"
    data$order_fold[data$Higher_taxon == "Hymenoptera"] <- "Hymenoptera"
    data$order_fold[data$Higher_taxon == "Hemiptera"] <- "Hemiptera"
    data$order_fold[data$Higher_taxon == "Arachnida"] <- "Arachnida"
    data$order_fold[data$Higher_taxon == "Odonata" |
                    data$Higher_taxon == "Trichoptera" |
                    data$Higher_taxon == "Ephemeroptera" |
                    data$Higher_taxon == "Plecoptera" ] <- "Aquatics"
    data$order_fold[data$Phylum == "Arthropoda" & 
                      data$Higher_taxon != "Diptera" & 
                      data$Higher_taxon != "Coleoptera" &
                      data$Higher_taxon != "Lepidoptera" &
                      data$Higher_taxon != "Hymenoptera" &
                      data$Higher_taxon != "Hemiptera" &
                      data$Higher_taxon != "Arachnida" &
                      data$Higher_taxon != "Odonata" &
                      data$Higher_taxon != "Trichoptera" &
                      data$Higher_taxon != "Ephemeroptera" &
                      data$Higher_taxon != "Plecoptera"] <- "OtherArthropods"
  }else{
    stop("Only the default folds have been coded up so far!")
  }
  
  data$order_fold <- as.factor(data$order_fold)
  
  if (verbose == TRUE){
    # By default, report on the breakdown of records per fold.
    cat("Numbers of rows within each order fold are as follows:\n\n")
    print(table(data$order_fold))
  }
  
  return(data)
}

```

### Load get_site_abundances_order function to calculate total abundance for each site x order_fold combination
```{r}
# Calculate total abundance for each site --------------------------------------
# Modified to clarify reporting is on site x fold combinations
get_site_abundances_order <- function(data, 
                                order_folds = FALSE,
                                verbose = TRUE){
  # This calculates site metrics for each site or, if desired, each order fold
  # within each site.
  
  # Make it a data frame to avoid having to deal with a tibble.
  data <- as.data.frame(data)
  
  # Set up a variable for uniquely identifying the site x order_fold combination
  # or, if order_folds == FALSE, the site. This is what grouping is then done
  # by.
  print(table(data$order_fold))
  
  if (order_folds == TRUE){
    data$the_order_folds <- data$order_fold
  }else{
    data$the_order_folds <- ""
  }
  
  # Just for debugging
  print(table(data$the_order_folds))
  
  # Start to calculate site-level metrics.
  sites <- data %>%
    
    # pull out only the merged diversity data
    distinct(merge_ID, .keep_all = TRUE) %>%
    
    # Re-make SSB and SSBS values since we've now dropped a bunch of values; and
    # also add SSBST which is that order x site combo. The trimws is because the
    # SSBST would otherwise end in a space if order_folds == FALSE.
    mutate(SS = paste(Source_ID, Study_number),
           SSB = paste(SS, Block),
           SSBS = paste(SSB, Site_number),
           SSBST = trimws(paste(SSBS, the_order_folds))) %>%
    
    # group by SSBST (each unique value corresponds to a unique site x fold)
    group_by(SSBST) %>%
    
    # Now add up all the abundance measurements within each site and also 
    # estimate coverage.
    mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                   sum(merged_diversity, na.rm = TRUE),
                                   # If the diversity metric type isn't 
                                   # Abundance, then leave the TotalAbundance 
                                   # measurement as NA.
                                   NA),
           
           SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                    merged_diversity,
                                    # for abundance and occurrence measurements,
                                    # count the number of unique species names 
                                    # that are present at the site 
                                    n_distinct(Best_guess_binomial[merged_diversity > 0])),
           
           coverage = get.SC2(merged_diversity)) %>%
    
    # ungroup
    ungroup() %>%
    
    # pull out unique site/fold combinations.
    distinct(SSBST, .keep_all = TRUE) %>%
    
    # now group by Study ID and fold.
    group_by(SS, the_order_folds) %>%
    
        # pull out the maximum abundance for each study/fold combination.
    mutate(MaxAbundance = max(TotalAbundance),
           MinNonZero = min(TotalAbundance[TotalAbundance > 0])) %>%
    
    # ungroup.
    ungroup() %>%
    
    # now rescale total abundance, so that within each study, the maximum is 1.
    mutate(RescaledAbundance = TotalAbundance/MaxAbundance) %>%
    
    mutate(sqrtRescaledAbundance = sqrt(RescaledAbundance)) 
  
    # If a study reports abundance as a count, add 1 to all values; otherwise
  # add 1/2 the minimum non-zero value; then log-transform to get a useful
  # logAbundance value that may be more symmetric than sqrtRescaledAbundance.
  sites$to_add <- ifelse(sites$Diversity_metric_unit %in%
                           c("individuals", "times observed"), 1,
                         sites$MinNonZero/2)
  sites$logAbundance <- log(sites$TotalAbundance + sites$to_add)
  
  
  # Remove NA values and drop levels
  sites <- as.data.frame(sites)
  sites <- droplevels(subset(sites, !is.na(sqrtRescaledAbundance)))
  
  if (verbose == TRUE){
    # Summarise structure of site-level data frame
    cat("\nThere are abundance data for:\n")
    cat(paste("   ", sum(!is.na(sites$sqrtRescaledAbundance)), "sites, from\n"))
    cat(paste("   ", 
              length(unique(sites$SS[!is.na(sites$sqrtRescaledAbundance)])), 
              "studies and\n"))
    cat(paste("   ", 
              length(unique(sites$Source_ID[!is.na(sites$sqrtRescaledAbundance)])),
              "sources.\n"))
  }
  
  # Set provenance attributes
  attr(sites, which = "diversity_extract") <- 
    attr(data, which = "diversity_extract")
  attr(sites, which = "is_merged") <- attr(data, which = "is_merged")
  attr(sites, which = "order_folds") <- order_folds 
  attr(sites, which = "when_Created") <- Sys.time()
  
  return(sites)
}

```

### Load all other functions
```{r}
source("/Applications/PhD/Andy/data_prep_functions.R")
source("/Applications/PhD/Andy/site_comparison_functions.R")
source("https://highstat.com/Books/Book2/HighstatLibV10.R") # collinearity
source("/Applications/PhD/Projects/multi-groups/scripts/combine_and_plot_function.R")
source("/Applications/PhD/Projects/multi-groups/scripts/predict_this_model.R")
```

### Read in the data
```{r}
diversity <- readRDS("../data/diversity-2022-04-13-02-33-10.rds")

```

### Correct for sampling effort
```{r}
if(check_sampling_effort_nas(diversity) ==
   FALSE) stop("Some studies have NA for sampling effort in only some sites!")

diversity <- correct_for_sampling_effort(diversity)
print(summarise_diversity(diversity))
```

### Merge sites
```{r}
merged <- merge_sites(diversity, verbose = TRUE) 
```

# Filter merged for five orders of interest and then subset for studies that look at more than one of those orders
```{r}
# Start from the merged data not the model data
five_orders <-
  merged %>%
  filter(Higher_taxon %in% c("Diptera", "Coleoptera", "Lepidoptera", "Hymenoptera", "Hemiptera")) %>% 
  # filter here for the orders of interest so that when we group by study, we are only looking at studies that observe those five orders
  group_by(SS) %>% # group by SS
  mutate(
    Multiple_groups = 
      case_when(
        (length(unique(Higher_taxon))) > 1 ~ "YES",
        (length(unique(Higher_taxon))) == 1 ~ "NO")) %>%  
  filter(Multiple_groups == "YES") %>% 
  droplevels()

# five_orders only contains studies that look at more than one of those five orders

levels(five_orders$Higher_taxon)
table(five_orders$Higher_taxon)
```

### Add_order_fold to allow taxonomic cross-validation 
```{r}
# Just adds order_fold column
five_orders <- add_order_fold(five_orders, verbose = TRUE)
levels(five_orders$order_fold)
```

### Calculate site level diversity metrics using get_site_abundances_order and order_fold for each site x order combination
```{r}

five_orders <- get_site_abundances_order(five_orders, 
                                  order_folds = TRUE, verbose = TRUE)

table(five_orders$order_fold) # number of rows for each site x order combination

# We have 8065 sites for the five orders in multi group studies

```

### Fix up your explanatory variables 
```{r}

# Rename predominant habitat since its not really habitat we're looking at
five_orders <- rename(five_orders,
                Predominant_land_use = Predominant_habitat)


# Relevel the land use and use intensity classes
five_orders <- five_orders %>%
  
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    
    # indeterminate secondary veg and cannot decide get NA
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(five_orders$Predominant_land_use, five_orders$Use_intensity)
```

# Collinearity
```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(five_orders[ , c("Predominant_land_use", "Use_intensity")])
```

### Complete cases
```{r}

# This will drop all the rows with NAs for abundance, land use or use intensity so we have complete data for modelling
model_data_ab <- drop_na(five_orders, 
                         RescaledAbundance, Predominant_land_use,
                         Use_intensity)

model_data_sr <- drop_na(five_orders, 
                         SpeciesRichness, Predominant_land_use,
                         Use_intensity)

#sum(is.na(five_orders$RescaledAbundance))
#sum(is.na(five_orders$SpeciesRichness))
#sum(is.na(five_orders$Predominant_land_use))
# There are no NAs for either of our explanatory variables so model_data_ab and model_data_sr are the same dataframe
```

### Prepare the abundance model dataframe
```{r}

# No longer needed as lme4 has improved, can use logAbundance provided in get_site_abundances()
# model_data_ab$logRescaledAbundance <- log(model_data_ab$RescaledAbundance + 1)
# sum(model_data_ab$RescaledAbundance == 1)

# Check the distributions
hist(model_data_ab$RescaledAbundance) ; hist(model_data_ab$logAbundance) ; hist(model_data_ab$sqrtRescaledAbundance)

plot(density(model_data_ab$logAbundance)) 

# Have a look at sample sizes
table(model_data_ab$Predominant_land_use, model_data_ab$Use_intensity)

# Collapse intense and light land use intensities to have a more even spread of the data
model_data_ab <- 
  model_data_ab %>%
  mutate(
    Use_intensity = recode_factor(Use_intensity,
                                  "Intense use" = "High_use",
                                  "Light use" = "High_use"),
# reset reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )
# Have a look at sample sizes
table(model_data_ab$Predominant_land_use, model_data_ab$Use_intensity)
```

# Abundance model
```{r}
# maximal model
ab1 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = model_data_ab)
# remove block
ab2 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
              data = model_data_ab)

# remove study
ab3 <-lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
              data = model_data_ab)

# Compare models 
AIC(ab1, ab2, ab3) # ab1 slightly better

Anova(ab1)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
ab4 <- lmer(logAbundance ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = model_data_ab)

## Just land use as predictor
ab5 <- lmer(logAbundance ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = model_data_ab)

## Just Higher_taxon as the predictor
ab6 <- lmer(logAbundance ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = model_data_ab)

## Make a null, intercept only model
ab7 <- lmer(logAbundance ~ (1 | Source_ID) + (1|SS), data = model_data_ab)

## Compare the models
AIC(ab1, ab4, ab5, ab6, ab7)
model.sel(ab1, ab4, ab5, ab6, ab7)

# ab1 still the best

# model estimates
summary(ab1)
anova(ab1)

```

# Predict effects for all of the orders
```{r}
# Predict effects for coleoptera first as they are the reference order
col <- predict_this_model(ab1, order = "Coleoptera")

# Relevel model with Diptera as reference level
dip_ref <- within(model_data_ab, Higher_taxon <- relevel(Higher_taxon, ref = "Diptera"))

# Run the diptera model
dip_m <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = dip_ref)

summary(dip_m) 

# Predict dip effects
dip <- predict_this_model(dip_m, order = "Diptera")

# relevel model with Hemiptera as reference level
hemi_ref <- within(model_data_ab, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model
hemi_m <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = hemi_ref)

summary(hemi_m) 

# Predict effects
hemi <- predict_this_model(hemi_m, order = "Hemiptera")

# relevel model with Hymenoptera as reference level
hym_ref <- within(model_data_ab, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model
hym_m <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
               data = hym_ref)

summary(hym_m) 

# Predict effects
hym <- predict_this_model(hym_m, order = "Hymenoptera")

# relevel model with Lep as reference level
lep_ref <- within(model_data_ab, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model
lep_m <- lmer(logAbundance ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
              data = lep_ref)

summary(lep_m) 

# Predict effects
lep <- predict_this_model(lep_m, order = "Lepidoptera")
```

# Combine the effects and plot them
```{r}
combined <- rbind(col, dip, hemi, hym, lep)

# Add land-use factors for plotting
comb_plot_limits <- c("Primary", "YSV", "ISV", "MSV", "Plantation", "Pasture", "Cropland", "Urban") 

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845")

combined %>%
  ggplot()+
  aes(x = Predominant_land_use, y =  median, colour = Order, group = Order)+
  geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
  geom_point(size = 3, position = position_dodge(width = 0.5))+
  geom_linerange(aes(ymin = Lower_ci, ymax = Upper_ci), position = position_dodge(width = 0.5), linewidth = 1)+
  theme_classic()+
  scale_x_discrete(limits=comb_plot_limits)+
  geom_vline(xintercept = 1.5, linetype = 2)+
  geom_vline(xintercept = 2.5, linetype = 2)+
  geom_vline(xintercept = 3.5, linetype = 2)+
  geom_vline(xintercept = 4.5, linetype = 2)+
  geom_vline(xintercept = 5.5, linetype = 2)+
  geom_vline(xintercept = 6.5, linetype = 2)+
  geom_vline(xintercept = 7.5, linetype = 2)+
  theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        panel.border = element_rect(colour = "black",  fill=NA))+
  xlab("Land use intensity class")+
  ylab("Log abundance") +
  scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
  guides(color = guide_legend(
    override.aes=list(shape = 15, size = 8)))

```

# Correlations
```{r}

# For each order extract the coefficients from ysv - urban 
beetles <- as.matrix(summary(ab1)$coefficients)[2:8,1]
flies <- as.matrix(summary(dip_m)$coefficients)[2:8,1]
bugs <- as.matrix(summary(hemi_m)$coefficients)[2:8,1]
bees <- as.matrix(summary(hym_m)$coefficients)[2:8,1]
moths <- as.matrix(summary(lep_m)$coefficients)[2:8,1]

# Combine them
five <- cbind(beetles, flies, bugs, bees, moths)

# Correlate
pairs(five)
cor(five)
cor.test(beetles, flies)

# Extract correlation coefficients just for one model 
# Wrong way to do it 
#five_corr <- as.matrix(summary(ab1)$coefficients)
# Split the coefficients 
# beetles <- five_corr[2:8,1]
# flies <- five_corr[13:19,1]
# bugs <- five_corr[20:26,1]
# bees <- five_corr[27:33,1]
# moths <- five_corr[34:nrow(five_corr), 1] 
# five <- cbind(beetles, flies, bugs, bees, moths)
# pairs(five)
# cor(five)

 
pairs(five,                     # Data frame of variables
      labels = colnames(five),  # Variable names
      pch = 21,                 # Pch symbol
      main = "Correlations between five orders",    # Title of the plot
      row1attop = TRUE,         # If FALSE, changes the direction of the diagonal
      gap = 1,                  # Distance between subplots
      cex.labels = NULL,        # Size of the diagonal text
      font.labels = 1)          # Font style of the diagonal text


library(corrplot)

corrplot(cor(five),        # Correlation matrix
         method = "circle", # Correlation plot method
         type = "full",    # Correlation plot style (also "upper" and "lower")
         diag = TRUE,      # If TRUE (default), adds the diagonal
         tl.col = "black", # Labels color
         bg = "white",     # Background color
         title = "",       # Main title
         col = NULL)       # Color palette

```

### Prepare the species richness model dataframe
```{r}
# Log transform species richness
model_data_sr$logSR <- log(model_data_sr$SpeciesRichness + 1)

# Check the distributions
hist(model_data_sr$SpeciesRichness) ; hist(model_data_sr$logSR) ; hist(model_data_ab$logAbundance) ; hist(log(model_data_ab$TotalAbundance + 1))

# Have a look at sample sizes
table(model_data_sr$Predominant_land_use, model_data_sr$Use_intensity)

# Collapse intense and light land use intensities to have a more even spread of the data
model_data_sr <- 
  model_data_sr %>%
  mutate(
    Use_intensity = recode_factor(Use_intensity,
                                  "Intense use" = "High_use",
                                  "Light use" = "High_use"),
    # reset reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )
# Have a look at sample sizes
table(model_data_sr$Predominant_land_use, model_data_sr$Use_intensity)
```

# Species Richness models
```{r}
# maximal model
sr1 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS) + (1 | SSB),
          data = model_data_sr)

# remove block
sr2 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
          data = model_data_sr)

# remove study
sr3 <-lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID),
          data = model_data_sr)

# Compare models 
AIC(sr1, sr2, sr3) # sr2 slightly better

Anova(sr2)

## Backwards stepwise selection of fixed effects 

## Model with the interaction removed but both predictors still present
sr4 <- lmer(logSR ~ Predominant_land_use + Higher_taxon + (1 | Source_ID) + (1|SS), data = model_data_sr)

## Just land use as predictor
sr5 <- lmer(logSR ~ Predominant_land_use + (1 | Source_ID) + (1|SS), data = model_data_sr)

## Just Higher_taxon as the predictor
sr6 <- lmer(logSR ~ Higher_taxon + (1 | Source_ID) + (1|SS), data = model_data_sr)

## Make a null, intercept only model
sr7 <- lmer(logSR ~ (1 | Source_ID) + (1|SS), data = model_data_sr)

## Compare the models
AIC(sr1, sr2, sr3, sr4, sr5, sr6, sr7)
model.sel(sr1, sr2, sr3, sr4, sr5, sr6, sr7)

# sr2 just slightly more optimal than sr1

# model estimates
summary(sr2)
anova(sr2)

rm(sr1,sr3, sr4, sr5, sr6, sr7, sr8)

```

# Predict effects for all of the orders
```{r}
# Predict effects for coleoptera first as they are the reference order
col <- predict_this_model(sr2, order = "Coleoptera")

# First, relevel model with Diptera as reference level
dip_ref <- within(model_data_sr, Higher_taxon <- relevel(Higher_taxon, ref = "Diptera"))

# dip model
dip_m <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
              data = dip_ref)

summary(dip_m) 

# Predict dip effects
dip <- predict_this_model(dip_m, order = "Diptera")

# relevel model with Hemiptera as reference level
hemi_ref <- within(model_data_sr, Higher_taxon <- relevel(Higher_taxon, ref = "Hemiptera"))

# maximal model
hemi_m <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
               data = hemi_ref)

summary(hemi_m) 

# Predict effects
hemi <- predict_this_model(hemi_m, order = "Hemiptera")

# relevel model with Hymenoptera as reference level
hym_ref <- within(model_data_sr, Higher_taxon <- relevel(Higher_taxon, ref = "Hymenoptera"))

# maximal model
hym_m <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
              data = hym_ref)

summary(hym_m) 

# Predict effects
hym <- predict_this_model(hym_m, order = "Hymenoptera")

# relevel model with Lep as reference level
lep_ref <- within(model_data_sr, Higher_taxon <- relevel(Higher_taxon, ref = "Lepidoptera"))

# maximal model
lep_m <- lmer(logSR ~ Predominant_land_use * Higher_taxon + (1 | Source_ID) + (1 | SS),
              data = lep_ref)

summary(lep_m) 

# Predict effects
lep <- predict_this_model(lep_m, order = "Lepidoptera")
```

# Combine the effects and plot them
```{r}
combined <- rbind(col, dip, hemi, hym, lep)

# Add land-use factors for plotting
comb_plot_limits <- c("Primary", "YSV", "ISV", "MSV", "Plantation", "Pasture", "Cropland", "Urban") 

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845")

combined %>%
  ggplot()+
  aes(x = Predominant_land_use, y =  median, colour = Order, group = Order)+
  geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
  geom_point(size = 3, position = position_dodge(width = 0.5))+
  geom_linerange(aes(ymin = Lower_ci, ymax = Upper_ci), position = position_dodge(width = 0.5), linewidth = 1)+
  theme_classic()+
  scale_x_discrete(limits=comb_plot_limits)+
  geom_vline(xintercept = 1.5, linetype = 2)+
  geom_vline(xintercept = 2.5, linetype = 2)+
  geom_vline(xintercept = 3.5, linetype = 2)+
  geom_vline(xintercept = 4.5, linetype = 2)+
  geom_vline(xintercept = 5.5, linetype = 2)+
  geom_vline(xintercept = 6.5, linetype = 2)+
  geom_vline(xintercept = 7.5, linetype = 2)+
  theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        panel.border = element_rect(colour = "black",  fill=NA))+
  xlab("Land use intensity class")+
  ylab("Log SR") +
  scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
  guides(color = guide_legend(
    override.aes=list(shape = 15, size = 8)))
```

# Correlations
```{r}

beetles <- as.matrix(summary(sr2)$coefficients)[2:8,1]
flies <- as.matrix(summary(dip_m)$coefficients)[2:8,1]
bugs <- as.matrix(summary(hemi_m)$coefficients)[2:8,1]
bees <- as.matrix(summary(hym_m)$coefficients)[2:8,1]
moths <- as.matrix(summary(lep_m)$coefficients)[2:8,1]

five <- cbind(beetles, flies, bugs, bees, moths)

pairs(five)
cor(five)

corrplot(cor(five),        # Correlation matrix
         method = "circle", # Correlation plot method
         type = "full",    # Correlation plot style (also "upper" and "lower")
         diag = TRUE,      # If TRUE (default), adds the diagonal
         tl.col = "black", # Labels color
         bg = "white",     # Background color
         title = "",       # Main title
         col = NULL)       # Color palette

```

# Junk code for now 
```{r}
# BE CLEAR TO SIMULATE THE EFFECTS OF EITHER AB OR SR BELOW!

# gather the effects and confidence intervals using simulation. Simulate 1000 times
effects <- FEsim(ab1, n.sims = 1000)

# make the term column a factor so it can be recoded
effects$term <- as.factor(effects$term)


col <- effects %>% slice(1:8)
dip <- effects %>% slice(9,13:19)
hemi<- effects %>% slice(10,20:26)
hym <- effects %>% slice(11,27:33)
lep <- effects %>% slice(12,34:40)


# rename factors for ease of visualisation
col$term <- recode_factor(col$term, "(Intercept)" = "Primary",
                          "Predominant_land_useYoung secondary vegetation" = "YSV",
                          "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                          "Predominant_land_useMature secondary vegetation" = "MSV",
                          "Predominant_land_usePlantation forest" = "Plantation",
                          "Predominant_land_usePasture" = "Pasture",
                          "Predominant_land_useCropland" = "Cropland",
                          "Predominant_land_useUrban" = "Urban")

# rename factors for ease of visualisation
dip$term <- recode_factor(dip$term, "Higher_taxonDiptera" = "Primary",
                          "Predominant_land_useYoung secondary vegetation:Higher_taxonDiptera" = "YSV",
                          "Predominant_land_useIntermediate secondary vegetation:Higher_taxonDiptera" = "ISV",
                          "Predominant_land_useMature secondary vegetation:Higher_taxonDiptera" = "MSV",
                          "Predominant_land_usePlantation forest:Higher_taxonDiptera" = "Plantation",
                          "Predominant_land_usePasture:Higher_taxonDiptera" = "Pasture",
                          "Predominant_land_useCropland:Higher_taxonDiptera" = "Cropland",
                          "Predominant_land_useUrban:Higher_taxonDiptera" = "Urban")

# rename factors for ease of visualisation
hemi$term <- recode_factor(hemi$term, "Higher_taxonHemiptera" = "Primary",
                          "Predominant_land_useYoung secondary vegetation:Higher_taxonHemiptera" = "YSV",
                          "Predominant_land_useIntermediate secondary vegetation:Higher_taxonHemiptera" = "ISV",
                          "Predominant_land_useMature secondary vegetation:Higher_taxonHemiptera" = "MSV",
                          "Predominant_land_usePlantation forest:Higher_taxonHemiptera" = "Plantation",
                          "Predominant_land_usePasture:Higher_taxonHemiptera" = "Pasture",
                          "Predominant_land_useCropland:Higher_taxonHemiptera" = "Cropland",
                          "Predominant_land_useUrban:Higher_taxonHemiptera" = "Urban")

# rename factors for ease of visualisation
hym$term <- recode_factor(hym$term, "Higher_taxonHymenoptera" = "Primary",
                          "Predominant_land_useYoung secondary vegetation:Higher_taxonHymenoptera" = "YSV",
                          "Predominant_land_useIntermediate secondary vegetation:Higher_taxonHymenoptera" = "ISV",
                          "Predominant_land_useMature secondary vegetation:Higher_taxonHymenoptera" = "MSV",
                          "Predominant_land_usePlantation forest:Higher_taxonHymenoptera" = "Plantation",
                          "Predominant_land_usePasture:Higher_taxonHymenoptera" = "Pasture",
                          "Predominant_land_useCropland:Higher_taxonHymenoptera" = "Cropland",
                          "Predominant_land_useUrban:Higher_taxonHymenoptera" = "Urban")

# rename factors for ease of visualisation
lep$term <- recode_factor(lep$term, "Higher_taxonLepidoptera" = "Primary",
                          "Predominant_land_useYoung secondary vegetation:Higher_taxonLepidoptera" = "YSV",
                          "Predominant_land_useIntermediate secondary vegetation:Higher_taxonLepidoptera" = "ISV",
                          "Predominant_land_useMature secondary vegetation:Higher_taxonLepidoptera" = "MSV",
                          "Predominant_land_usePlantation forest:Higher_taxonLepidoptera" = "Plantation",
                          "Predominant_land_usePasture:Higher_taxonLepidoptera" = "Pasture",
                          "Predominant_land_useCropland:Higher_taxonLepidoptera" = "Cropland",
                          "Predominant_land_useUrban:Higher_taxonLepidoptera" = "Urban")

# Add in upper and lower confidence intervals 

col <- col %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

dip <- dip %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

hemi <- hemi %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

hym <- hym %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))

lep <- lep %>%
  mutate(Upper_ci = (median + 1.96*sd)) %>%
  mutate(Lower_ci = (median - 1.96*sd))


# Back transform the log abundance/species richness estimates
col <- col %>%
  mutate(
    Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
  mutate(
    Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
  mutate(
    Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)

dip <- dip %>%
  mutate(
    Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
  mutate(
    Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
  mutate(
    Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)

hemi <- hemi %>%
  mutate(
    Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
  mutate(
    Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
  mutate(
    Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)

hym <- hym %>%
  mutate(
    Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
  mutate(
    Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
  mutate(
    Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)

lep <- lep %>%
  mutate(
    Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
  mutate(
    Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
  mutate(
    Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)


# Shift the baseline down to 0

col[1,7] <- 0
col[1,8] <- 0
col[1,9] <- 0

dip[1,7] <- 0
dip[1,8] <- 0
dip[1,9] <- 0

hemi[1,7] <- 0
hemi[1,8] <- 0
hemi[1,9] <- 0

hym[1,7] <- 0
hym[1,8] <- 0
hym[1,9] <- 0

lep[1,7] <- 0
lep[1,8] <- 0
lep[1,9] <- 0


# Rename the sim eff term column name for joining
colnames(col)[1] <- "Predominant_land_use"
colnames(dip)[1] <- "Predominant_land_use"
colnames(hemi)[1] <- "Predominant_land_use"
colnames(hym)[1] <- "Predominant_land_use"
colnames(lep)[1] <- "Predominant_land_use"


# Add in an order column for grouping
col <- col %>% mutate(Order = "Coleoptera")
dip <- dip %>% mutate(Order = "Diptera")
hym <- hym %>% mutate(Order = "Hymenoptera")
hemi <- hemi %>% mutate(Order = "Hemiptera")
lep <- lep %>% mutate(Order = "Lepidoptera")


# Combine all the predictions
combined <- rbind(col, dip, hemi, hym, lep)

# Add land-use factors for plotting
comb_plot_limits <- c("Primary", "YSV", "ISV", "MSV", "Plantation", "Pasture", "Cropland", "Urban") 

# Add specific colours for each order
group_colours <- c(Diptera = "#FFC300", Lepidoptera = "#ff5733", Hymenoptera = "#c70039", Hemiptera = "#900c3f", Coleoptera = "#581845")


plot <- combined %>%
  ggplot()+
  aes(x = Predominant_land_use, y = Percent_diff, colour = Order, group = Order)+
  geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
  geom_point(size = 3, position = position_dodge(width = 0.5))+
  geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
  theme_classic()+
  scale_x_discrete(limits=comb_plot_limits)+
  geom_vline(xintercept = 1.5, linetype = 2)+
  geom_vline(xintercept = 2.5, linetype = 2)+
  geom_vline(xintercept = 3.5, linetype = 2)+
  geom_vline(xintercept = 4.5, linetype = 2)+
  geom_vline(xintercept = 5.5, linetype = 2)+
  geom_vline(xintercept = 6.5, linetype = 2)+
  geom_vline(xintercept = 7.5, linetype = 2)+
  theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        panel.border = element_rect(colour = "black",  fill=NA))+
  xlab("Land use intensity class")+
  ylab("Total species richness difference (%)") +
  scale_colour_manual(values=group_colours) + # this overrides the colours for the groups as above
  guides(color = guide_legend(
    override.aes=list(shape = 15, size = 8)))


# Investigating flies abundance
flies <- model_data_ab %>% dplyr::select(Order, Predominant_land_use, SS, logAbundance, Sampling_method) %>% filter(Order == "Diptera")

# Investigating leps in YSV and ISV for SR
leps <- model_data_sr %>% dplyr::select(Order, Predominant_land_use, SS, logSR) %>% filter(Order == "Lepidoptera")

table(leps$Predominant_land_use) 


leps %>% filter(Predominant_land_use == "Young secondary vegetation") %>% dplyr::select(logSR) %>% count(logSR == 0)

leps %>% filter(Predominant_land_use == "Intermediate secondary vegetation") %>% dplyr::select(logSR) %>% count(logSR == 0)

leps %>% filter(Predominant_land_use == "Mature secondary vegetation") %>% View()

leps %>% dplyr::select(SS, Order, Predominant_land_use, logSR) %>% 
  filter(Predominant_land_use == c("Young secondary vegetation","Intermediate secondary vegetation","Mature secondary vegetation")) %>% View()


leps %>% filter(Predominant_land_use == c("Young secondary vegetation","Intermediate secondary vegetation","Mature secondary vegetation")) %>% View()

leps %>% dplyr::select(Predominant_land_use, logSR) %>% filter(Predominant_land_use %in% c("Young secondary vegetation", 
                                                                                                   "Intermediate secondary vegetation",
                                                                                                   "Mature secondary vegetation")) %>% View()



hist(flies$logAbundance)

# Extract f ratio, df and p value
anova(dhy_m2)

# Check residuals
hist(residuals(dhy_m2))

# Check r2
r.squaredGLMM(dhy_m2)
```

# Boot strapping
```{r}

diversity_metrics <- sitemetrics(diversity)

for(i in 1:1000){
  diversity_sub <- diversity_metrics %>%
    
    #randomly sampler 1000 rows
    
    # run your predicts model for that random samply if 1000 sitees
    
    # extraction the predictions
    
    # and then calculate the median, 2.5th,m and 97.5th percentiles
    
}

```






