---
title: "bycatch_hymenoptera"
author: "Justin Isip"
date: "2023-12-21"
output: html_document
---

# This markdown will produce a list of bycatch studies for Hymenoptera within PREDICTS based on studies that sample multiple groups of insects.

# Read in the data and load packages
```{r}
diversity <- readRDS("./data/diversity-2023-09-14-02-33-21.rds")
library(tidyverse)
```

# Load functions
```{r}
source("/Applications/PhD/Andy/data_prep_functions.R")
```

# Correct for sampling effort
```{r}
if(check_sampling_effort_nas(diversity) ==
   FALSE) stop("Some studies have NA for sampling effort in only some sites!")

diversity <- correct_for_sampling_effort(diversity) # correct for sampling effort

print(summarise_diversity(diversity)) # quick summary
```

# Merge sites
```{r}
merged <- merge_sites(diversity, verbose = TRUE) 
```

# First, I will find the studies which contain data for multiple insect groups but subset at the source level, rather than the study level.

# The idea being that some sources may contain multiple separate studies with different sampling methods for different insect groups. 

# An example could be, within a source, Study 1: Malaise traps for Diptera, Study 2: Pitfall traps for Coleoptera, Study 3: Sweep nets for Lepidoptera, they will be seperate studies with different sampling methods within a source, but would have been conducted at the EXACT SAME set of sites and habitats. Therefore they are still precisely matched and suitable for my analysis.

# I may be missing data which has multiple groups at the source level but not at the study level, and this data would not be extracted if I subset ONLY at the study level. 

```{r}
source_multiple <- merged %>% 
  filter(Higher_taxon %in% c("Diptera", "Coleoptera", "Lepidoptera", "Hymenoptera", "Hemiptera")) %>% 
  group_by(Source_ID) %>%  # Subset multiple orders at the source level
  mutate(
    Multiple_groups = 
      case_when(
        (length(unique(Higher_taxon))) > 1 ~ "YES",
        (length(unique(Higher_taxon))) == 1 ~ "NO")) %>%  # when the number of unique higher taxon > 1, its a multi group study.
  filter(Multiple_groups == "YES") %>% 
  droplevels()

ss_multiple <- merged %>% 
  filter(Higher_taxon %in% c("Diptera", "Coleoptera", "Lepidoptera", "Hymenoptera", "Hemiptera")) %>% 
  group_by(SS) %>%  # Subset multiple orders at the study level
  mutate(
    Multiple_groups = 
      case_when(
        (length(unique(Higher_taxon))) > 1 ~ "YES",
        (length(unique(Higher_taxon))) == 1 ~ "NO")) %>%  
  filter(Multiple_groups == "YES") %>% 
  droplevels()


# Which data is subset at the source level but not at the study level?
source_only <- anti_join(source_multiple, ss_multiple, 
                by = c("Source_ID", "SS", "Order", "Sampling_method", "Site_name", "Habitat_as_described", "Latitude", "Longitude")) 
# matching based on EXACT SAME site_name, habitat_as_described and lat/long coordinates!

# Simplify source level df for merging
source_only <- source_only %>%
  mutate(site_habitat_combo = paste(Site_name, Habitat_as_described),# Concatenate site and habitat_as_described
         lat_long_combo = paste(Latitude, Longitude)) %>% # Concatenate lat and long
  dplyr::select(Source_ID, SS, Order, Sampling_method, site_habitat_combo, lat_long_combo) %>% # Select the columns of interest
  unique()
  
# Simplify the study level df for merging
studies_only <- ss_multiple %>%
  mutate(site_habitat_combo = paste(Site_name, Habitat_as_described),
         lat_long_combo = paste(Latitude, Longitude)) %>% 
  dplyr::select(Source_ID, SS, Order, Sampling_method, site_habitat_combo, lat_long_combo) %>%
  unique()

# Merge the studies_only df and source_only df only by rows that have an EXACT MATCH in their values for site_habitat_combo (i.e have the EXACT same sites and habitats)
same_site_habitat <- inner_join(studies_only, source_only, by = c("site_habitat_combo", "lat_long_combo")) 

# SS.x is subset at study level, SS.y is at the source level, here we can see which site x habitat combos match 
same_site_habitat %>% dplyr::select(SS.x, Order.x, Sampling_method.x, site_habitat_combo, SS.y, Order.y, Sampling_method.y) %>% unique() %>% View()

# the 10 additional studies subset at the source level 
same_site_habitat %>% dplyr::select(Source_ID.y, SS.y, Order.y, Sampling_method.y) %>% unique() %>% View()

# Check there's no missing lat/long or site/habitat data
sum(is.na(studies_only$lat_long_combo))
sum(is.na(source_only$site_habitat_combo))

# The unique studies in this column are the studies which have the EXACT same sites and habitats subset at the source level
data2add <- unique(same_site_habitat$SS.y) %>% droplevels()
print(data2add)

rm(same_site_habitat, source_only, studies_only)
```

# Merge source level with study level multi-group data

```{r}
# Filter for those additional studies subset at source level
source_multiple <- source_multiple %>% filter(SS %in% data2add) %>% droplevels() 

# Check its pulled the right studies
source_multiple %>% select(Source_ID, SS) %>% unique() %>% View()
print(data2add)

# Merge the additional 10 studies subset at the source level to the dataset subset at the study level
overall <- rbind(source_multiple, ss_multiple)

# Check
length(unique(source_multiple$SS))
length(unique(ss_multiple$SS))
length(unique(overall$SS)) # 113 studies overall

rm(source_multiple, ss_multiple, data2add, diversity)
```

# Now that I have my overall dataset (with multiple groups subset at both the source and study level)
# I will update the sampling methods for a number of studies based on checking the source publication. The idea being that some studies in my dataset have an obscure/unspecified sampling method (e.g. specimen collection, systematic searching, various, multiple, mixed trap etc).

```{r}
# First ungroup the df
overall <- overall %>% ungroup()

overall %>% select(SS, Sampling_method) %>% unique() %>%  View()  # have a look at the sampling methods for each study

# Update the sampling methods
overall <- overall %>%
  mutate(Sampling_method = case_when(
    SS == "AD1_2002__Vazquez 1" ~ "flower visitation observation",
    SS == "AD1_2009__Vergara 1" ~ "flower visitation observation",
    SS == "AD1_2011__Bates 1" ~ "Pan traps/sweep nets/hand searching", # multiple
    SS == "AD1_2011__Weiner 1" ~ "flower visitation observation",
    SS == "AD1_2013__Grass 1" ~ "flower visitation observation",
    SS == "AR1_2008__Basset 1" ~ "Malaise traps/yellow pan traps/pitfall traps", # multiple
    SS == "CC1_2005__Barratt 1" ~ "soil sample",
    SS == "CC1_2005__Barratt 2" ~ "soil sample",
    SS == "CC1_2008__Schon 1" ~ "soil cores along transects",
    SS == "CC1_2008__Schon 2" ~ "soil cores along transects",
    SS == "CC1_2010__Schon 2" ~ "soil cores along transects",
    SS == "CC1_2012__Waite 2" ~ "bole traps/flight intercept trap/foliage sampling", # multiple
    SS == "CC1_2013__Litchwark 1" ~ "flower visitation observation",
    SS == "CC1_2014__Rader 1" ~ "flight intercept trap/pan trap", # multiple
    SS == "DG1_2012__Ge 1" ~ "soil cores along transects",
    SS == "GP1_2009__Vasconcelos 2" ~ "litter collection",
    SS == "HP1_2006__Lachat 1" ~ "Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap", # multiple
    SS == "MH1_2009__Turner 1" ~ "Leaf litter sampling/ferns cut and bagged/fogging for canopy species", # multiple
    SS == "SC1_2005__Richardson 1" ~ "litter collection",
    SS == "SJ2_2019__Eckert 1" ~ "pit-fall traps", # multiple "Pitfall trapping/direct sampling (i.e. active searching) within quadrats" but didn't include direct sampling as not a PREDICTS sampling method
    SS == "VB1_2005__Eggleton 2" ~ "Hand sorting in quadrat and winkler bags", # multiple
    SS == "VB1_2012a_Carpenter 1" ~ "Soil cores/pitfall traps", # multiple
    SS == "VB1_2012__LeightonGoodall 1" ~ "Leaf litter sampling/soil sampling", # multiple
    SS == "GP1_2009__Vasconcelos 1" ~ "litter collection", # this study was subset at source level
    SS == "VB1_2012__Carpenter 5" ~ "soil sampling/litter collection/pitfall traps", # multiple - this study was subset at source level
    SS == "AD1_2006__Blanche 1" ~ "flower visitation observation",
    SS == "AD1_2006__Blanche 2" ~ "flower visitation observation",
    SS == "AD1_2007__Meyer 2" ~ "flower visitation observation",
    SS == "SH1_2011__Todd 1" ~ "flight intercept/pan trap/pitfall trap", # multiple
    SS == "VB1_2020__Burton 2" ~ "litter collection", # multiple
    TRUE ~ Sampling_method
  ))

# Recheck
overall %>% select(SS, Order, Sampling_method) %>% unique() %>%  View()

# Merge similar methods, e.g. sweep netting and sweep net.
overall <- overall %>%
mutate(Sampling_method = recode_factor(Sampling_method,
                                       "sweep netting" = "sweep net", # Combine sweep net and sweep netting
                                       "sweep_netting" = "sweep net",
                                       "soil core" = "soil sample",
                                      "soil cores along transects" = "soil sample", # Combine soil cores, soil cores along transect and soil sample
                                      "mixed traps" = "multiple", # Combine mixed traps with multiple traps
                                      "window trap" = "flight trap" # Combine window trap with flight trap as they are the same thing
                                      ))
# Check unique sampling methods
unique(overall$Sampling_method)
```

# Some studies sample larvae, speaking with Gavin Broad (Entomologist at the NHM), you would only expect to sample larvae in the soil, therefore all other instances of larvae caught with any sampling method other than soil sampling are bycatch. I will filter PREDICTS only for larvae sampled in the soil below.

```{r}
# Lets have a look at all instances of larvae
overall %>% 
  filter(grepl('larv|immatu|pupa|nymphe|nymph_', Taxon_name_entered, ignore.case = T)) %>% # different words for larvae 
  dplyr::select(SS, Higher_taxon, Taxon_name_entered, Order, Genus, Species, Sampling_method) %>% 
  filter(Higher_taxon %in% c("Diptera", "Coleoptera", "Lepidoptera", "Hymenoptera", "Hemiptera")) %>%  # orders relevant to my analysis
  unique %>% 
  View() 

# Larvae_keep contains data only on larvae where they have been sampled appropriately in the soil.
larvae_keep <- overall %>%  
  filter(grepl('larv|immatu|pupa|nymphe|nymph_', Taxon_name_entered, ignore.case = T) & Sampling_method %in% 
           c("soil sample",
             "soil sampling/litter collection/pitfall traps",
             "Soil cores/pitfall traps",
             "Leaf litter sampling/soil sampling")) %>%  
              droplevels()

# Those studies that have multiple sampling methods including soil sampling, do not have many instances of larvae so it should not be a problem to include them.
```

# Now, based on the expert spreadsheet sent over by Andrew (containing the sampling methods used by experts for their order(s) of interest). 
# I will, for each sampling method, filter for the orders which are not bycatch.

```{r}

# active visual surveys good for all five orders
active_visual_surveys <- overall %>% filter(Order %in% c("Lepidoptera", "Coleoptera", "Diptera", "Hymenoptera", "Hemiptera") 
                                            & Sampling_method %in% c("sweep net", 
                                                                     "aerial transect", 
                                                                     "flower visitation observation", 
                                                                     "visual encounter survey", 
                                                                     "systematic searching")) 
# these five PREDICTS methods constitute active visual surveys in the expert spreadsheet.


# light trap - good for all except Hemiptera
light_trap <- overall %>% filter(Order %in% c("Lepidoptera", "Coleoptera", "Diptera", "Hymenoptera") & Sampling_method == "light trap")

 # pan trap - good for all except Lepidoptera
pan_trap <- overall %>% filter(Order %in% c("Coleoptera", "Diptera", "Hymenoptera", "Hemiptera") & Sampling_method %in% c("pan traps", 
                                                                                                "Pan traps/sweep nets/hand searching",
                                                                                                "Malaise traps/yellow pan traps/pitfall traps",
                                                                                                "flight intercept trap/pan trap",
                                                                                                "flight intercept/pan trap/pitfall trap"))

# You can see for pan traps there are instances of multiple sampling methods for a given study, when I merge all of the cleaned methods I will pass a function that will remove duplicate rows, so this won't be an issue.

# flight intercept / harp trap - good for Hymenoptera and Coleoptera
flight_trap <- overall %>% filter(Order %in% c("Coleoptera", "Hymenoptera") & Sampling_method %in% 
                                    c("flight trap",
                                      "bole traps/flight intercept trap/foliage sampling",
                                      "flight intercept trap/pan trap",
                                      "Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap",
                                      "flight intercept/pan trap/pitfall trap",
                                      "harp trap"))

# Malaise trap - good for all except Lepidoptera
malaise_trap <- overall %>% filter(Order %in% c("Coleoptera", "Diptera", "Hymenoptera", "Hemiptera") & Sampling_method %in% c("malaise traps",
                                                                                                          "Malaise traps/yellow pan traps/pitfall traps",
                                                                  "Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap"))


# Winkler bags or litter collection good for beetles and ants, other orders bycatch 
winkler_beetles <- overall %>% filter(Order == "Coleoptera" & Sampling_method %in% c("litter collection", 
                                                                                     "soil sampling/litter collection/pitfall traps",
                                                                                     "Hand sorting in quadrat and winkler bags", 
                                                                                     "Leaf litter sampling/ferns cut and bagged/fogging for canopy species",
                                                                                     "Leaf litter sampling/soil sampling"))

# Have to create a seperate object for ants in order to filter for Formicidae 
winkler_ants <- overall %>% filter(Order == "Hymenoptera" & Family == "Formicidae" & Sampling_method %in% c("litter collection", 
                                                                                     "soil sampling/litter collection/pitfall traps",
                                                                                     "Hand sorting in quadrat and winkler bags", 
                                                                                     "Leaf litter sampling/ferns cut and bagged/fogging for canopy species",
                                                                                     "Leaf litter sampling/soil sampling"))

# All expert methods have now been cleaned for non-bycatch data.
```

# Some of the PREDICTS sampling methods could not be assigned to an expert sampling method.
# These include - soil sampling, trap nests, bole trap, suction samplers, live traps and baited traps.
# Where a study had multiple sampling methods, if any of those multiple sampling methods was suitable for any of the groups then it was included.
# I will clean these methods based on comms with Victoria Burton, Gavin Broad and Ben Woodcock. 
```{r}
# Soil sample only good for ants, all other groups bycatch.
soil_sampling <- overall %>% 
  filter(Order == "Hymenoptera" & Family == "Formicidae" & Sampling_method %in% c("soil sample",
                                                                                  "soil sampling/litter collection/pitfall traps",
                                                                                  "Soil cores/pitfall traps",
                                                                                  "Leaf litter sampling/soil sampling"))


# Glue trap good for Hymenoptera, Diptera and Hemiptera, other orders bycatch.
glue_trap <- overall %>% filter(Order %in% c("Hymenoptera", "Diptera", "Hemiptera") & Sampling_method == "glue traps")

# Bole trap only good for Coleoptera
bole_trap <- overall %>%  filter(Order == "Coleoptera" & Sampling_method %in% c("bole trap", 
                                                                           "bole traps/flight intercept trap/foliage sampling"))


# BW says suction samplers good for Coleoptera and Hemiptera but not Hymenoptera and Lepidoptera.
suction <- overall %>% filter(Order %in% c("Coleoptera", "Hemiptera") & Sampling_method == "suction samplers")

# Trap nests only good for Hymenoptera says VB.
trap_nests <- overall %>% filter(Order == "Hymenoptera" & Sampling_method == "trap nests")

# Pitfall traps good for only Coleoptera and ants says VB (surprisingly not good for Hemiptera).
pitfall_beetles <- overall %>% filter(Order == "Coleoptera" & 
                                        Sampling_method %in% c("Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap",
                                                               "Malaise traps/yellow pan traps/pitfall traps", 
                                                               "Soil cores/pitfall traps", 
                                                               "soil sampling/litter collection/pitfall traps", 
                                                               "pit-fall traps", 
                                                               "baited pit-fall traps", 
                                                               "flight intercept/pan trap/pitfall trap"))

pitfall_ants <- overall %>% filter(Order == "Hymenoptera" & Family == "Formicidae" 
                                   & Sampling_method %in% c("Funnel pitfall traps/ malaise trap/flight intercept trap/pyramid shaped emergence trap",
                                                               "Malaise traps/yellow pan traps/pitfall traps", 
                                                               "Soil cores/pitfall traps", 
                                                               "soil sampling/litter collection/pitfall traps", 
                                                               "pit-fall traps", 
                                                               "baited pit-fall traps", 
                                                               "flight intercept/pan trap/pitfall trap"))


# Below is code to filter the overall dataset for the levels of a given sampling method, in this example pitfall traps.
overall %>%  
  filter(grepl('pit', Sampling_method, ignore.case = T)) %>% 
  droplevels() %>% 
  select(Sampling_method) %>% 
  unique() %>% 
  View()

# 2023 Smith is the only baited-traps paper, Diptera and Hymenoptera are OK (Hymenoptera > 20% of the individuals) but Coleoptera, Hemiptera, and Lepidoptera bycatch. 
baited_trap <- overall %>% filter(Order %in% c("Diptera", "Hymenoptera") & Sampling_method == "baited traps") 

# Live traps (TN_1_2007_Bouyer 2 and TN_1_2007_Bouyer 3) all taxa are targeted - checked source paper.
live_trap <- overall %>% filter(Sampling_method == "live traps")


# All non-expert methods have now been cleaned for non-bycatch data.
```

# Now that I've removed bycatch from all of the sampling methods in PREDICTS, I will combine them all with the larvae_keep data.
```{r}
clean <- bind_rows(active_visual_surveys,
                   baited_trap,
                   bole_trap,
                   flight_trap,
                   glue_trap,
                   larvae_keep,
                   light_trap,
                   live_trap,
                   malaise_trap,
                   pan_trap,
                   pitfall_ants,
                   pitfall_beetles,
                   soil_sampling,
                   suction,
                   trap_nests,
                   winkler_ants,
                   winkler_beetles) 
# ^ each of these R objects is a sampling method in PREDICTS which has been filtered only for the orders which are NOT bycatch.

# Remove duplicate rows as some data will have been duplicated due to the studies with multiple sampling methods.
clean <- unique(clean)

# Tidy the environment
rm(active_visual_surveys,baited_trap,bole_trap,flight_trap,glue_trap,larvae_keep,light_trap,live_trap,malaise_trap,pan_trap,pitfall_ants,pitfall_beetles,
soil_sampling,suction,trap_nests,winkler_ants,winkler_beetles)
```

# Lets create a list of which studies are bycatch for Hymenoptera 
```{r}
nrow(overall) - nrow(clean) # 87401 rows of bycatch data removed
length(unique(overall$SS)) - length(unique(clean$SS)) # 7 studies removed

different_rows <- rbind(setdiff(overall, clean)) %>% droplevels() # setdiff finds all rows in overall that aren't in clean, i.e. all of the bycatch rows.

# For Hymenopterans, given me a list of the studies which are bycatch
different_rows %>% filter(Order == "Hymenoptera") %>%  select(SS, Order, Family, Sampling_method) %>% unique() %>% View()

bycatch <- different_rows %>% filter(Order == "Hymenoptera") %>%  select(SS, Order, Family, Sampling_method) %>% unique() 

# I've included Family in the above, as some of the PREDICTS sampling methods are OK for ants (i.e. pitfall traps or Winkler bags) but not for non-ant hymenopterans, therefore when you want to choose which data to remove, for those studies where family is specified, remove those which aren't Formicidae, where the Family isn't specified then you know its not sampling ants.  

# For example, the sampling method for DI1_2013__Rousseau 2 is pit-fall traps which are only good for ants and beetles, therefore all of the data except the
# ant and beetle data is bycatch. You can see the first row is Formicidae sampled in a pitfall trap which is not bycatch and the Coleopteran data is also not bycatch, the rest is bycatch. 
overall %>% filter(SS == "DI1_2013__Rousseau 2") %>% select(SS, Order, Family, Sampling_method) %>% unique() %>% View()

saveRDS(bycatch, file = "hymenoptera_bycatch.rds")

```

