---
title: "Traits analysis"
output: html_document
date: "2025-04-07"
author: "Justin Isip"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    code_folding: hide
    theme: flatly
---

# Clear the work space

```{r}
rm(list=ls())
```

# Read in the packages 
```{r}
library(lme4)
library(readxl)
library(tidyverse)
library(stringr)
library(car)
library(qs)
library(lmerTest)
library(merTools) 
library(performance)
library(DHARMa)
library(here)
```

# Load in model plot function
```{r}
## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(2,2))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}
```

# Read in Logghe et al 2025 which has body size data for ~ 4800 arthropod species and makes the bulk of the GLiTRS trait data set
# Tidy and clean it to merge with PREDICTS trait data sets
```{r}
# Upload logghe et al
logghe <- read_excel(here("data", "Body_Size_data.xlsx"))

# keep only first 8 cols
logghe <- logghe %>% dplyr::select(1:8)

# Make first row the header
colnames(logghe) <- as.character(logghe[1, ])
logghe <- logghe[-1, ] # drop first row

# Create a genus and species column
logghe <- logghe %>%
  separate(scientificName, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop")

# Combine genus and species to create Best_guess_binomial which I will use to merge 
logghe <- logghe %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Reorder columns
logghe <- logghe %>% dplyr::select(order, family, Genus, Species, Best_guess_binomial, Trait_Value) %>% 
  rename(Family = family) %>% 
  rename(Order = order)

logghe$Trait_Value <-  as.numeric(logghe$Trait_Value)
```

# Read and tidy Geoffrey Su's ant trait data
```{r}
# Load the data
trait_ants <- read_excel(here("data", "SuFormacidaeBodyLength-2014-05-23.xlsx")) %>%  strings2factors()

summary(trait_ants)

# Rename body length average column
trait_ants <- trait_ants %>% rename(avg_body_length_mm = "Body.Length..mm...worker..AVG")

# Lets remove those rows with NAs for body length
trait_ants <- trait_ants %>%
  filter(!is.na(avg_body_length_mm) & avg_body_length_mm != 0) # 559 rows with no body length data, 2 rows where BL = 0

# Combine genus and species  
trait_ants <- trait_ants %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Simplify ants
trait_ants <- trait_ants %>% dplyr::select(Genus, Best_guess_binomial, avg_body_length_mm)

# Tidy
trait_ants <- trait_ants %>%
  mutate(
    Order = "Hymenoptera",
    Family = "Formicidae"
  ) %>% 
  rename(Trait_Value = avg_body_length_mm) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop") %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)
# ants ready
```

# Read and tidy Melanie Edgars beetle trait data
```{r}
# Load Melanies beetle trait data
melanie_beetles <- read_excel(here("data", "EdgarBeetleLength-2014-01-09.xlsx"))

# Combine genus and species for merging
melanie_beetles <- melanie_beetles %>% mutate(Best_guess_binomial = str_c(GENUS, SPECIES, sep = " "))

# Calculate average b.length
melanie_beetles <- melanie_beetles %>%
  separate(B.LENGTH, into = c("b_min", "b_max"), sep = "-", fill = "right", remove = FALSE) %>% # fill = "right" allows single values to stay in b_min while b_max becomes NA.
  mutate(
    b_min = as.numeric(trimws(b_min)), # trimws() removes any leading/trailing spaces 
    b_max = as.numeric(trimws(b_max)),
    b.length_avg = if_else(
      is.na(b_max),
      b_min,  # If no range, use the min value
      (b_min + b_max) / 2  # Else if range, compute the average
    ))

# Prepare and tidy beetles
melanie_beetles <- melanie_beetles %>% 
  mutate(Order = "Coleoptera") %>% 
  dplyr::select(Order, FAMILY, GENUS, SPECIES, Best_guess_binomial, b.length_avg) %>% 
  rename(Family = FAMILY) %>% 
  rename(Genus = GENUS) %>% 
  rename(Species = SPECIES) %>% 
  rename(Trait_Value = b.length_avg)
```

# Read and tidy Rachel Tans bee trait data
```{r}
# Load Rachel bee trait data
rachel_bee <- read_excel(here("data", "BeeTraits_P2_RachelTan.xlsx"))

# Create a genus and species column
rachel_bee <- rachel_bee %>%
  separate(CorrectedName, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop")

# Filter for columns of interest
rachel_bee <- rachel_bee %>% dplyr::select(HigherTaxon, Family, Genus, Species, CorrectedName, MidRangeBodyLength) %>% drop_na() # drop those sp. with NA for body length

# Remove rows identified as "sp." or inside brackets e.g., (Dialictus) or (Psithyrus) which are subgenera and not a true species name
rachel_bee <- rachel_bee %>%
  filter(
    !str_detect(Species, "sp\\.") &         # remove rows with "sp."
    !str_detect(Species, "\\(.*\\)")        # remove rows with anything in brackets
  )

# Rename for merging
rachel_bee <- rachel_bee %>% 
  rename(Best_guess_binomial = CorrectedName) %>% 
  rename(Order = HigherTaxon) %>% 
  rename(Trait_Value = MidRangeBodyLength)

```

# Merge Logghe, Geoffrey, Melanies and Rachels trait datasets to have an overall body size trait dataset
```{r}
# merge all trait datasets
all_traits <- bind_rows(logghe, melanie_beetles, rachel_bee, trait_ants)

# There are duplicate species entries
all_traits[duplicated(all_traits["Best_guess_binomial"]) | 
                                  duplicated(all_traits["Best_guess_binomial"], fromLast = TRUE), ] %>%  View()

# Sense check - there are 634 duplicates
all_traits[duplicated(all_traits["Best_guess_binomial"]),] 

# For those species with duplicate entries, calculate an average Trait_Value which is the mean of the duplicate entries
all_traits_clean <- all_traits %>%
  group_by(Best_guess_binomial) %>%
  summarise(
    Trait_Value = mean(Trait_Value, na.rm = TRUE),  # average Trait_Value
    across(.cols = -Trait_Value, .fns = ~ first(.x)),  # for each column, take the first value
    .groups = "drop"
  ) %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)

nrow(all_traits) - nrow(all_traits_clean) # 634

rm(all_traits, logghe, melanie_beetles, rachel_bee, trait_ants)
```

# Read in PREDICTS data 
```{r}
# All insect data effort-corrected and merged
insects <-  qread("/Users/jusi/Desktop/multi-groups/data/predicts-prepared-whole-database.qs")
```

# Clean and tidy PREDICTS data then merge trait dataset based on taxon resolution
```{r}
# Get relevant taxon cols for all insects
taxon_insects <- insects %>% dplyr::select(SS, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, Merged_diversity, Diversity_metric_unit, Country)

# Get rid of rows which aren't resolved to order level
taxon_insects <- taxon_insects %>% filter(Order != "") # 1056 rows not resolved to order

# Recreate genus and species column as sometimes BGB is present in PREDICTS but no info for genus or sp.
taxon_insects <- taxon_insects %>%
  mutate(Best_guess_binomial = as.character(Best_guess_binomial)) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "merge", fill = "right")

# Get species trait values for merging
sp_value <- all_traits_clean %>% dplyr::select(Best_guess_binomial, Trait_Value) 

# First merge Logghe species level data
taxon_insects <- left_join(taxon_insects, sp_value, by = "Best_guess_binomial")

rm(sp_value)

# Calculate average body length for genera
genus_value <- all_traits_clean %>%
  group_by(Genus) %>%
  summarise(genus_avg_body_length = mean(Trait_Value, na.rm = TRUE)) # 2381 genus data

# Merge based on genus where best_guess_binomial is empty
taxon_insects <- taxon_insects %>%
  left_join(genus_value, by = "Genus", suffix = c("", "_genus")) 

rm(genus_value)

# Average body length per family based on species
family_value <- all_traits_clean %>%
  group_by(Family) %>%
  summarise(family_avg_body_length = mean(Trait_Value, na.rm = TRUE))

# Merge based on family where genus and species is empty
taxon_insects <- taxon_insects %>%
  left_join(family_value, by = "Family", suffix = c("", "_family")) 

rm(family_value)

# priority-based fallback assignment where final body length values are selected based on the first non-missing (or non-blank) column in a defined order.
taxon_insects <- taxon_insects %>%
  mutate(final_body_length = coalesce(
    Trait_Value,
    genus_avg_body_length,
    family_avg_body_length
  ))
```

# Rainford et al 2014 provides min and max body length data for 760 insect families, I will calculate the average body length between the min and the max and then merge those values onto those rows in my data frame that do not yet have a body length value. 
 
```{r}
# Upload 
rainford <- read_excel(here("data", "rainford_body_size.xlsx"))

# Make first row the header
colnames(rainford) <- as.character(rainford[1, ])
rainford <- rainford[-1, ] # drop first row

# Extract cols of interest
rainford <- rainford %>% dplyr::select("Order", "Family", "Min (mm)", "Max (mm)") %>% rename(min_bl = "Min (mm)") %>% rename(max_bl = "Max (mm)")

# Drop rows with all NAs
rainford <- rainford %>%
  filter(!if_all(everything(), is.na))

# Convert cols to numeric
rainford$min_bl <- as.numeric(rainford$min_bl) # Odonata (Pseudolestidae) has an NA for BL hence the warning message
rainford$max_bl <- as.numeric(rainford$max_bl)

# Drop rows only to Order level
rainford <- rainford %>%
  filter(!is.na(Family))

# Calculate average body length
rainford <- rainford %>%
  mutate(Rainford_family_avg = (min_bl + max_bl) / 2)
         
# Simplify for merging
rainford <- rainford %>% 
  dplyr::select(c(-min_bl, -max_bl, -Order))
```

# Merge Rainford 

```{r}
# Step 1: Filter rows in taxon_insects where there is no body size value at the species, genus or family level and therefore we can merge a family body size value from Rainford
needs_family_avg <- taxon_insects %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length))

# Step 2: Join family_avg body size values from Rainford based on Family column
needs_family_avg <- needs_family_avg %>%
  left_join(rainford, dplyr::select(Family, Rainford_family_avg), by = "Family")

# Step 3: Combine updated rows (need_family_avg) back into original data (taxon_insects)
# First, remove the rows you updated from taxon_insects
remaining_rows <- taxon_insects %>%
  filter(!(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length)))

# sense check remaining rows should be taxon_insects - needs_family_avg
nrow(taxon_insects) - nrow(needs_family_avg)

# Bind them together
taxon_insects_updated <- bind_rows(remaining_rows, needs_family_avg)

# How many rows still without a family average
sum(!is.na(taxon_insects_updated$Rainford_family_avg)) # 134120

# How many rows have data based on each level of taxonomic resolution
taxon_insects_updated %>%
  summarise(across(
    c(Trait_Value,
    genus_avg_body_length,
    family_avg_body_length,
    final_body_length,
    Rainford_family_avg),
    list(
      non_NAs = ~sum(!is.na(.))
    )
  ))

rm(needs_family_avg, remaining_rows, taxon_insects)
```

# Now that we have merged all the trait data, coalesce the family averages from Rainford into the final_body_length column (for those rows that don't yet have a body length value at the family level) - if they already do then we stick with the earlier value from Logghe and the PREDICTS datasets.

```{r}
# Replace those rows with missing values in final_body_length with the corresponding family_avg
taxon_insects_updated <- taxon_insects_updated %>%
  mutate(final_body_length = coalesce(final_body_length, Rainford_family_avg)) # coalesce(a,b) returns a if its not NA otherwise it returns b

# How many rows do we still not have a body size value for and are still NA - 71437
taxon_insects_updated %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length) & is.na(final_body_length) & is.na(Rainford_family_avg)) %>%
  nrow()

# Have a look at the rows
taxon_insects_updated %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length) & is.na(final_body_length) & is.na(Rainford_family_avg)) %>% 
  View()

# Filter out rows which still don't have a body size value
all_insects_size <- taxon_insects_updated %>%
  filter(!is.na(final_body_length)) %>% # 1195212
  droplevels()


all_insects_size %>% distinct(Family)
```

# First split sites based on their latitude to tropical or temperate

```{r}
# Filter overall predicts for sites and their latitude
latitude <- insects %>% dplyr::select(SSBS, Latitude) 

# There are duplicated sites
any(duplicated(latitude$SSBS))

# Lets get a list of distinct sites
latitude <- latitude %>% distinct(SSBS, .keep_all = TRUE)

# Merge latitude values into size data frame for all of the sites
all_insects_size <- all_insects_size %>%
  left_join(latitude, by = "SSBS")

# Add a new column classifying each site into tropical or temperate
all_insects_size$climate_zone <- ifelse(all_insects_size$Latitude >= -23.4368 & all_insects_size$Latitude <= 23.4368, 
                          "tropical", 
                          "temperate")
```

# Read in allometry coefficients
```{r}
coefficients <- read_excel(here("data", "allometry_coefficients.xlsx"))
```

# Merge allometry coefficients to size data
```{r}
# Step 1: Join by order and climate_zone
all_insects_size <- all_insects_size %>%
  left_join(coefficients, by = c("Order", "climate_zone"))

# Step 2: Assign All_insects coefficients for all other taxa without order specific equations
# First extract coefficients for All_insects
fallback <- coefficients %>%
  filter(Order == "All_insects") %>%
  dplyr::select(climate_zone, a_fallback = a, b_fallback = b)

# Step 3: Join coefficients for All_insects based on whether its a tropical or temperate site
all_insects_size <- all_insects_size %>%
  left_join(fallback, by = "climate_zone") %>%
  mutate(
    a = if_else(is.na(a), a_fallback, a), # if a is NA, assign it a for all_insects, else keep it as its order specific a
    b = if_else(is.na(b), b_fallback, b)
  ) %>%
  dplyr::select(-a_fallback, -b_fallback)

# Have a look
all_insects_size %>% dplyr::select(SSBS, Order, final_body_length, Latitude, climate_zone, a, b) %>% unique() %>%  View()
```

# Calculate mass in mg using allometric specific coefficients for each order
```{r}
all_insects_size <- all_insects_size %>%
  mutate(
    log10_mass = a + b * log10(final_body_length),
    mass_mg = 10^log10_mass
  )

summary(all_insects_size$final_body_length) # no zero values for final_body_length
any(all_insects_size$final_body_length == 0) # FALSE

rm(coefficients, fallback, latitude, rainford)
```

# Summary stats for methods

```{r}
# Number of species by order
all_insects_size %>%
  filter(!is.na(Best_guess_binomial)) %>% 
  filter(Best_guess_binomial != "") %>% 
  group_by(Order) %>%
  summarise(n_species = n_distinct(Best_guess_binomial)) %>%
  arrange(desc(n_species))  #  Optional: orders by number of species

# Number of species with species-level body length data
 all_insects_size %>%
  filter(!is.na(Trait_Value)) %>%
  distinct(Best_guess_binomial) %>%
  nrow()

# Number of genera with genus-level body length data
all_insects_size %>%
  filter(!is.na(genus_avg_body_length)) %>%
  distinct(Genus) %>%
  nrow()

# Number of families with family-level body length data
all_insects_size %>%
  filter(!is.na(family_avg_body_length)) %>%
  distinct(Family) %>%
  nrow()

# Table X highlighting availability of body length data for numbers of species, genera and families across orders.
order_summary <- all_insects_size %>%
  # Keep only rows with body length data available
  filter(!is.na(Trait_Value) | !is.na(genus_avg_body_length) | !is.na(family_avg_body_length | !is.na(Rainford_family_avg))) %>%
  group_by(Order) %>%
  summarise(
    n_species = n_distinct(Best_guess_binomial[!is.na(Trait_Value)]),
    n_genera  = n_distinct(Genus[!is.na(genus_avg_body_length)]),
    n_families = n_distinct(Family[!is.na(family_avg_body_length) | !is.na(Rainford_family_avg)]) # two columns for family body length!
  ) %>%
  arrange(desc(n_species))  # Optional: sort by most species with data

# View the result
print(order_summary)

```

# Breakdown of the data
```{r}
# Which rows are assigned to family level but have no genus or sp. info?
family_only_rows <- all_insects_size %>% 
  filter(
    !is.na(Family),
    (is.na(Genus) | Genus == ""),
    (is.na(Best_guess_binomial) | Best_guess_binomial == "")
  )

# Whats the percentage of the overall data frame that are resolved to ONLY family?
NROW(family_only_rows) / NROW(all_insects_size)

# Which rows have only a family-level body length value, but ARE taxonomically resolved to genus or species.
genus_or_species_rows <- all_insects_size %>%
  filter(
    (!is.na(family_avg_body_length) | !is.na(Rainford_family_avg)),  # contains a family value from either column
    is.na(Trait_Value),                   # no species level value
    is.na(genus_avg_body_length),         # no genus level value
    (!is.na(Genus) & Genus != "") | (!is.na(Best_guess_binomial) & Best_guess_binomial != "") # genus or species present
  )

# Whats the percentage of the overall data frame have only a family-level body length value, but ARE taxonomically resolved to genus or species.
NROW(genus_or_species_rows) / NROW(all_insects_size)

# A. The number of rows with only a family-level body length value but with genus or species assigned
genus_or_sp_with_familyBL <- genus_or_species_rows %>%
  filter(!is.na(Family)) %>%  # Ensure Family is not NA
  group_by(Family) %>%
  summarise(genus_or_species_n_rows = n())

# B. The total number of rows for each Family in the entire all_insects_size data frame.
total_family_counts <- all_insects_size %>%
  filter(!is.na(Family)) %>%
  group_by(Family) %>%
  summarise(total_n_rows = n())

# C. Join the two and calculate the proportion
percentage <- genus_or_sp_with_familyBL %>%
  inner_join(total_family_counts, by = "Family") %>%
  mutate(percentage = genus_or_species_n_rows / total_n_rows)

# Round to two decimal places
percentage <- percentage %>%
  mutate(percentage = round(percentage, 2))


all_insects_size %>% filter(Family == "Apidae") %>% View()


all_insects_size %>% filter(Genus == "Pheidole") %>% View()


# Sense check Formicidae
all_insects_size %>% filter(Family == "Formicidae",
                              is.na(Trait_Value), # no species level value
                              is.na(genus_avg_body_length), # no genus level value
                            (!is.na(Genus) & Genus != "") | (!is.na(Best_guess_binomial) & Best_guess_binomial != "") # either genus or species is present
                            ) %>% View()

# Sense check Apidae
all_insects_size %>% filter(Family == "Apidae",
                              is.na(Trait_Value), # no species level value
                              is.na(genus_avg_body_length), # no genus level value
                            (!is.na(Genus) & Genus != "") | (!is.na(Best_guess_binomial) & Best_guess_binomial != "") # either genus or species is present
                            ) %>% View()

# Bees that are resolved to genus or species and matched to genus or species
 all_insects_size %>%
  filter(
    Family == "Apidae",
    # Resolved to genus or species
    (!is.na(Genus) & Genus != "") | (!is.na(Best_guess_binomial) & Best_guess_binomial != ""),
    # Matched to genus or species body size value
    !is.na(genus_avg_body_length) | !is.na(Trait_Value)
  )


# For each family, how many genera are there without a genus level body length value?
genera_without_genus_value <- all_insects_size %>%
  filter(!is.na(Genus) & Genus != "") %>%           # must be a genus
  group_by(Family, Genus) %>%
  summarise(has_genus_value = any(!is.na(genus_avg_body_length)), .groups = "drop") %>%  # for each genus within each family, calculate whether the genus has a genus-level body length value present 
  filter(!has_genus_value) %>%                       # keep only genera that do NOT have a genus-level value
  group_by(Family) %>%
  summarise(num_genera_without_genus_value = n()) # count the number of genera without a genus level body length value

# Append num_genera_without_genus_value by joining on Family
percentage_with_genera <- percentage %>%
  left_join(genera_without_genus_value, by = "Family")

# Some genera have numbers e.g. "Hyich91" - this is Hymenoptera, this is Ichneumoidae - it is different from 90 other Ichneumonidae that I have put into other groupings. The authors didn’t know the genus and so we will never know the genus - that data is not salvageable.

# Lets remove such genera

# Get genera without genus-level value, add a column if genus has digits
genera_without_genus_value_and_digits <- all_insects_size %>%
  filter(!is.na(Genus) & Genus != "") %>% # must be a genus
  group_by(Family, Genus) %>%
  summarise(has_genus_value = any(!is.na(genus_avg_body_length)), .groups = "drop") %>% # for each genus within each family, calculate whether the genus has a genus-level body length value present 
  filter(!has_genus_value) %>% # keep only genera that do NOT have a genus-level value
  mutate(real_sounding = !str_detect(Genus, "[0-9]")) # does the genus contain digits in its name?

# Count real-sounding genera per family
num_genera_with_real_names <- genera_without_genus_value_and_digits %>%
  group_by(Family) %>%
  summarise(num_genera_without_genus_value_with_real_names = sum(real_sounding))

# Add new column
percentage_with_genera <- percentage_with_genera %>%
  left_join(num_genera_with_real_names, by = "Family")

# Next step I want to calculate the total number of rows that belong to "real" genera (i.e. genera in num_genera_without_genus_value_with_real_names)

# Filter for those genera without a genus level body length value and digits in their names (and so can actually be improved)
real_genera <- genera_without_genus_value_and_digits %>%
  filter(real_sounding) %>%
  dplyr::select(Family, Genus)

# Count rows for real genera in all_insects_size
real_genera_row_counts <- all_insects_size %>%
  inner_join(real_genera, by = c("Family", "Genus")) %>%
  group_by(Family) %>%
  summarise(num_rows_in_real_genera = n())

# merge into main data frame
percentage_with_genera <- percentage_with_genera %>%
  left_join(real_genera_row_counts, by = "Family")

# Next step, for each family, calculate the number of SSBS that the rows belonging to "real" genera come from
real_genera_sites <- all_insects_size %>% 
  inner_join(real_genera, by = c("Family", "Genus")) %>% 
  filter(!is.na(SSBS) & SSBS != "")

# count unique sites per family
site_counts <- real_genera_sites %>%
  group_by(Family) %>%
  summarise(num_SSBS_in_real_genera = n_distinct(SSBS))

# merge into main data frame
percentage_with_genera <- percentage_with_genera %>% 
  left_join(site_counts, by = "Family")

# Next step, add which Order each Family belongs to

# Select order family pairs
order_family_combo <- all_insects_size %>%
  dplyr::select(Order, Family) %>%
  distinct()

# Join and move Order to the front
percentage_with_genera <- percentage_with_genera %>%
  left_join(order_family_combo, by = "Family") %>%
  dplyr::select(Order, everything())

# There's very little benefit to refining the data when each site has only a single species of a family. So the ratio of num_rows_in_real_genera / num_SSBS_in_real_genera is also useful. High values are more appealing, other things being equal.
percentage_with_genera <- percentage_with_genera %>%
  mutate(avg_rows_per_SSBS_in_real_genera = num_rows_in_real_genera / num_SSBS_in_real_genera)


# Next step, for each family (only including those those real-sounding genera; no digits) - I will create a column that is a list of unique Country values sortd by greaatest to least frequently occurring.

# filter all_insects_size for rows in real genera
rows_in_real_genera <- all_insects_size %>%
  inner_join(real_genera, by = c("Family", "Genus")) %>% # keep all the observations in all_insects_size that have a match in real_genera
  filter(!is.na(Country), Country != "") 

# Count country frequencies per family, and arrange by frequency
countries_ordered <- rows_in_real_genera %>%
  group_by(Family, Country) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(Family, desc(n)) %>%
  group_by(Family) %>%
  summarise(
    countries_by_frequency = paste(Country, collapse = ", ")
  )

# merge into main data frame
percentage_with_genera <- percentage_with_genera %>%
  left_join(countries_ordered, by = "Family")


write.csv(percentage_with_genera, "percentage_with_generaV2.csv", row.names = FALSE)

# find all the unique genera in the family Ichneumonidae 
unique_genera_ichneumonidae <- all_insects_size %>%
  filter(Family == "Ichneumonidae", !is.na(Genus), Genus != "") %>%
  distinct(Genus) %>%
  arrange(Genus)

# find all the unique genera in the family Ichneumonidae 
unique_genera_saturniidae <- all_insects_size %>%
  filter(Family == "Saturniidae", !is.na(Genus), Genus != "") %>%
  distinct(Genus) x%>%
  arrange(Genus)

# 1. Calculate the number of unique genera per family (for rows that have a family-level body length value but are taxonomically resolved to genus or species)
genera_per_family <- genus_or_species_rows %>%
  filter(!is.na(Genus), Genus != "") %>%      # remove empty genera  
  group_by(Family) %>%
  summarise(num_genera = n_distinct(Genus)) %>%
  arrange(desc(num_genera))

mean(c(0.97, 0.92, 0.47, 0.93, 0.95, 0.76, 0.93, 0.73, 0.30, 0))
mean(c(0.76, 0.36, 0.99, 0.42, 0.61, 0.69, 0.97, 0.58, 0.30, 0.53))

mean(c(0.90, 0.93, 0.82, 0.97, 0.99, 0.95, 0.99, 0.98, 0.97, 0.91))
mean(c(0.99, 0.96, 0.99, 0.71, 0.96, 0.99, 0.97, 0.99, 0.96, 0.91))

```



# Analysis looking at ALL of the insects 

# first lets calculate one weighted mean length value per site

```{r}
# Create a separate data frame for the overall site level analysis
overall_site_size <- all_insects_size

# Calculate the biomass of each taxa
overall_site_size <- overall_site_size %>% 
  mutate(
    Biomass = mass_mg * Merged_diversity
  )

# Calculate TotalSiteBiomass, SmallestNonZeroTotalSiteBiomass for a site within a study and Log_plus_what
overall_site_size <- overall_site_size %>%
  group_by(SSBS) %>%
  mutate(TotalSiteBiomass = sum(Biomass, na.rm = TRUE)) %>%  # Calculate total site biomass
  ungroup() %>%
  group_by(SS) %>% # Now group by study
  mutate(
    SmallestNonZeroTotalSiteBiomass = min(TotalSiteBiomass[TotalSiteBiomass > 0]), # Which site has the smallest total biomass greater than 0 per study? 
    Log_plus_what = ifelse(
    Diversity_metric_unit %in% c("individuals", "times observed"), min(mass_mg, na.rm = TRUE),  
    # If abundance data are counts, assign the smallest mass_mg for an individual of any species in the study
    SmallestNonZeroTotalSiteBiomass / 2)) %>%  # else 1/2 the SmallestNonZeroTotalSiteBiomass for any site within the study 
  ungroup() %>% 
  group_by(SSBS) %>% 
  mutate(LogBiomass = log(TotalSiteBiomass + Log_plus_what)) %>% 
  ungroup()
    
# Have a look
overall_site_size %>% dplyr::select(SS, SSBS, Merged_diversity, mass_mg, Biomass, Diversity_metric_unit, TotalSiteBiomass, SmallestNonZeroTotalSiteBiomass, Log_plus_what, LogBiomass) %>% unique() %>%  View()
    
overall_site_size %>%  filter(SS == "VB1_2023__Hendy 3") %>%  dplyr::select(SS, SSBS, Merged_diversity, mass_mg, Biomass, Diversity_metric_unit, TotalSiteBiomass, SmallestNonZeroTotalSiteBiomass, Log_plus_what, LogBiomass) %>% unique() %>%  View()  

# Calculate one LogBiomass value per site
LogBiomass_summary <- overall_site_size %>%
  group_by(SSBS) %>%
  summarise(
    n_TSB = n_distinct(TotalSiteBiomass), 
    n_LPW = n_distinct(Log_plus_what),
    is_consistent = (n_TSB == 1 & n_LPW == 1), # sense check that all TotalSiteBiomass and Log_plus_what values are the same for all rows in the site
    TotalSiteBiomass = first(TotalSiteBiomass),
    Log_plus_what = first(Log_plus_what),
    LogBiomass = log(first(TotalSiteBiomass) + first(Log_plus_what)),
    .groups = "drop"
  )

# All sites are consistent
LogBiomass_summary %>% filter(is_consistent != TRUE)

inconsistent_groups <- overall_site_size %>%
  group_by(SSBS) %>%
  summarise(
    n_TSB = n_distinct(TotalSiteBiomass),
    n_LPW = n_distinct(Log_plus_what)
  ) %>%
  filter(n_TSB > 1 | n_LPW > 1) 

# No rows where LogBiomass is 0, NA, or blank
LogBiomass_summary %>%
  filter(is.na(LogBiomass) | LogBiomass == 0 | LogBiomass == "") %>% View()

rm(inconsistent_groups)
```

# OLD CODE FOR CALCULATING LogBiomass
```{r}
# Create a separate data frame for the overall site level analysis as we'll use all_insects_size to calculate order by order
overall_site_size <- all_insects_size

# Calculate abundance weighted log mass
overall_site_size <- overall_site_size %>%
  group_by(SSBS) %>%
  mutate(
    relativeAb_logMass = (log10_mass * Merged_diversity) / sum(Merged_diversity, na.rm = TRUE)
  ) %>%
  ungroup()

# Have a look at the worked exampled
overall_site_size %>% filter(SSBS == "AD1_2009__Vergara 1  6") %>% View()

# Calculate total log mass per site
logMass_summary <- overall_site_size %>%
  group_by(SSBS) %>%
  summarise(
    total_div = sum(Merged_diversity, na.rm = TRUE),
    relAb_sum = sum(relativeAb_logMass, na.rm = TRUE),
    logMass_total = ifelse(total_div > 0, relAb_sum + log(total_div), NA_real_) # this addresses -Inf values due to sites with 0 merged diversity, i.e. taking a log of 0.
  ) 

# Have a look at the worked exampled
logMass_summary %>% filter(SSBS == "AD1_2009__Vergara 1  6") %>% View()

# View rows where logMass_total is 0, NA, or blank
logMass_summary %>%
  filter(is.na(logMass_total) | logMass_total == 0 | logMass_total == "") %>% View()

rm(coefficients, fallback, latitude)
```

# Merge with PREDICTS insects
```{r}
# Use distinct() to keep only the first appearance of each site within the overall PREDICTS div extract for insects
insects_single <- insects %>%
  distinct(SSBS, .keep_all = TRUE)

# merge LogBiomass_summary values
final_size <- insects_single %>%
  left_join(LogBiomass_summary, by = "SSBS") %>% 
  droplevels()

# Have a look at sites with no LogBiomass
no_biomass <- final_size %>%
  filter(
    is.na(LogBiomass) |         # remove NA
    is.nan(LogBiomass)|        # remove NaN
    LogBiomass == ""            # remove empty strings
  )  # I think this is because these rows do not have a body size value, therefore no LogBiomass value. 

# filter out sites with no LogBiomass 
final_size <- final_size %>%
  filter(
    !is.na(LogBiomass),         # remove NA
    !is.nan(LogBiomass),        # remove NaN
    LogBiomass != ""            # remove empty strings
  )

# 14009 sites in total

rm(insects_single, no_biomass, LogBiomass_summary)
```

# Load additional explanatory variables
```{r}
# Read in the data
add_var <-  qread("/Users/jusi/Desktop/multi-groups/data/abundance-model-data-prepared 2.qs")

# Subset for cols of interest
add_var <- add_var %>% dplyr::select(SSBS, t30, Landscape_Natural_Habitat) 

# Creating log_t30_scaled
add_var$log_t30 <- log1p(add_var$t30)
add_var$log_t30_scaled <- scale(add_var$log_t30)

# Creating nat_hab_scaled
add_var$nathab_scaled <- scale(add_var$Landscape_Natural_Habitat)

# merge logMass_total values
final_size <- final_size %>%
  left_join(add_var, by = "SSBS") %>% 
  droplevels()

# have a look
final_size %>% dplyr::select(SSBS, t30, log_t30, log_t30_scaled, Landscape_Natural_Habitat, nathab_scaled) %>% View()


final_size %>%
  summarise(
    t30_non_NA = sum(!is.na(t30)),
    log_t30_non_NA = sum(!is.na(log_t30)),
    log_t30_scaled_non_NA = sum(!is.na(log_t30_scaled)),
    Landscape_Natural_Habitat_non_NA = sum(!is.na(Landscape_Natural_Habitat)),
    nathab_scaled_non_NA = sum(!is.na(nathab_scaled))
  )
# 12036/12918 for log_t30 and 12033/12918 for nathab scaled.

```

# Start the modelling of all of the insects

```{r}
# This PREDICTS data file is diversity extract of all insect data which has been merged and effort corrected.
# It has also had a set of sampling methods updated, see PreparePREDICTSData.R
# It has not had accidentally sampled larvae removed as this analysis does not focus on multiple groups.

# Rename predominant habitat since its not really habitat we're looking at
final_size <- rename(final_size,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
final_size <- final_size %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(final_size$Predominant_land_use, final_size$Use_intensity)
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
final_size <- final_size %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# Have a look at sample sizes
table(final_size$LUI)

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 61 sites, merge with light use
final_size <- final_size %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Have a look at sample sizes
table(final_size$LUI)

# Set the reference levels for LUI
final_size <- final_size %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )

# Have a look at sample sizes
table(final_size$LUI)
```

# Create NatOrHumDom

```{r}
final_size <- final_size %>%
  mutate(
    NatOrHumDom = case_when(
      is.na(Predominant_land_use) ~ NA_character_,
      Predominant_land_use %in% c("Primary", "Young_secondary_vegetation", "Intermediate secondary vegetation", "Mature secondary vegetation") ~ "natural",
      TRUE ~ "human_dominated"
    )
  )
# Have a look
final_size %>% dplyr::select(Predominant_land_use, NatOrHumDom) %>% View()

```

# Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(final_size[ , c("Predominant_land_use", "Use_intensity")])
```

# Complete cases and model data

```{r}
# Lets have a look at rows with an NA for our variables of interest
final_size %>%
  filter(if_any(c(Predominant_land_use, Use_intensity, LUI, log_t30_scaled, nathab_scaled, NatOrHumDom), is.na)) %>% 
  dplyr::select(SS, SSBS,Predominant_land_use, Use_intensity, LUI, log_t30_scaled, nathab_scaled, NatOrHumDom) %>% 
  View()

final_size %>%
  filter(!if_any(c(Predominant_land_use, Use_intensity, LUI, log_t30_scaled, nathab_scaled, NatOrHumDom), is.na)) %>% 
  dplyr::select(SS, SSBS,Predominant_land_use, Use_intensity, LUI, log_t30_scaled, nathab_scaled, NatOrHumDom) %>% 
  View()

# drop all rows with NAs for land use and use intensity and drop empty levels
model_data_biomass <- drop_na(final_size, Predominant_land_use, Use_intensity) %>% droplevels() 

# drop all rows for additional variables for modelling
#model_data_t30 <- drop_na(final_size, Predominant_land_use, Use_intensity, log_t30_scaled, nathab_scaled) %>% droplevels()

# check distributions
hist(model_data_biomass$LogBiomass)
#hist(model_data_t30$log_t30_scaled)
#hist(model_data_t30$nathab_scaled)

# Have a look at sample sizes
table(model_data_biomass$Predominant_land_use, model_data_biomass$Use_intensity)
table(model_data_biomass$LUI)

length(unique(model_data_biomass$SS)) # 444 studies 
length(unique(model_data_biomass$SSBS)) # 11516 sites

#length(unique(model_data_t30$SS)) # 402 studies 
#length(unique(model_data_t30$SSBS)) # 9773 sites


table(model_data_biomass$Region)

rm(add_var, final_size, logMass_summary, model_data_t30, overall_site_size)
```

# Have a look at temperate versus tropical split for log biomass model data

```{r}
# Filter overall predicts for sites and their latitude
latitude <- model_data_biomass %>% dplyr::select(SSBS, Latitude) 

# There are duplicated sites
any(duplicated(latitude$SSBS))

# Lets get a list of distinct sites
latitude <- latitude %>% distinct(SSBS, .keep_all = TRUE)

# Add a new column classifying each site into tropical or temperate
latitude$climate_zone <- ifelse(latitude$Latitude >= -23.4368 & latitude$Latitude <= 23.4368, 
                          "tropical", 
                          "temperate")

table(latitude$climate_zone)
```

# Modelling LogBiomass ~ LU and creating the LU plot for all insects

```{r}
# maximal model
m1 <- lmer(LogBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_biomass)

# remove block
m2 <-lmer(LogBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS),
    data = model_data_biomass)

# remove study
m3 <-lmer(LogBiomass ~ Predominant_land_use + (1 | Source_ID),
    data = model_data_biomass)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(biomass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
  #ggsave(filename = "../figures/biomass_LU.png",  plot = p, width = 6, height = 4, dpi = 300)

biomass_eff <- eff %>% mutate(Variable = "Biomass")


       
rm(eff, m1, m2, m3, p)     
```

# Modelling LogBiomass ~ LUI (LU + UI) and creating the LUI plot for all insects

```{r}
# maximal model
m1 <- lmer(LogBiomass ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_biomass)

# remove block
m2 <-lmer(LogBiomass ~ LUI + (1 | Source_ID) + (1 | SS),
    data = model_data_biomass)

# remove study
m3 <-lmer(LogBiomass ~ LUI + (1 | Source_ID),
    data = model_data_biomass)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary (minimal)",
                            "LUIPrimary_Light_use" = "Primary (light)",
                                "LUIPrimary_Intense_use" = "Primary (intense)",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV (minimal)",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV (light)",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV (intense)",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV (minimal)",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV (high)",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV (minimal)",
                            "LUIMature_secondary_vegetation_High_use" = "MSV (high)",
                            "LUIPlantation_forest_Minimal_use" = "Plantation (minimal)",
                            "LUIPlantation_forest_Light_use" = "Plantation (light)",
                            "LUIPlantation_forest_Intense_use" = "Plantation (intense)",
                            "LUIPasture_Minimal_use" = "Pasture (minimal)",
                            "LUIPasture_Light_use" = "Pasture (light)",
                            "LUIPasture_Intense_use" = "Pasture (intense)",
                            "LUICropland_Minimal_use" = "Cropland (minimal)",
                            "LUICropland_Light_use" = "Cropland (light)",
                            "LUICropland_Intense_use" = "Cropland (intense)",
                            "LUIUrban_Minimal_use" = "Urban (minimal)",
                            "LUIUrban_Light_use" = "Urban (light)",
                            "LUIUrban_Intense_use" = "Urban (intense)")
                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    p <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, colour = "purple")+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1, colour = "purple")+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(biomass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
       #ggsave(filename = "../figures/biomass_LUI.png",  plot = p, width = 8, height = 7, dpi = 300)

#biomass_LUI_eff <- eff %>% mutate(Variable = "Biomass")    
  
rm(eff, m1, m2, m3, p)     
```

# Modelling size ~ LUI + logt30 + nathab + NatOHumDom

```{r}
# nathab_scaled = amount of natural habitat in the surrounding area
# log_t30_scaled = log(number of years since the broad landscape around the site first got 30% converted to human use)
# NatOrHumDom = natural or human dominated, natural = primary or secondary, all others human dominated.


# maximal model
m1 <- lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled + log_t30_scaled:NatOrHumDom + (1 | Source_ID) + (1 | SS) + (1 | SSB), data = model_data_t30)

# remove block
m2 <- lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled + log_t30_scaled:NatOrHumDom + (1 | Source_ID) + (1 | SS), 
           data = model_data_t30)

# remove study
m3 <-lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled + log_t30_scaled:NatOrHumDom + (1 | Source_ID),
    data = model_data_t30)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary_minimal_use",
                                "LUIPrimary_High_use" = "Primary_high_use",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "Young_secondary_vegetation_Minimal_use",
                            "LUIYoung_secondary_vegetation_High_use" = "Young_secondary_vegetation_High_use",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "Intermediate_secondary_vegetation_Minimal_use",
                            "LUIIntermediate_secondary_vegetation_High_use" = "Intermediate_secondary_vegetation_High_use",
                            "LUIMature_secondary_vegetation_Minimal_use" = "Mature_secondary_vegetation_Minimal_use",
                            "LUIMature_secondary_vegetation_High_use" = "Mature_secondary_vegetation_High_use",
                            "LUIPlantation_forest_Minimal_use" = "Plantation_forest_Minimal_use",
                            "LUIPlantation_forest_High_use" = "Plantation_forest_High_use",
                            "LUIPasture_Minimal_use" = "Pasture_Minimal_use",
                            "LUIPasture_High_use" = "Pasture_High_use",
                            "LUICropland_Minimal_use" = "Cropland_Minimal_use",
                            "LUICropland_High_use" = "Cropland_High_use")
                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3)+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 12.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("Total log mass difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
rm(eff, m1, m2, m3, final_size, logMass_summary, model_data_biomass, overall_site_size)     
```

# Model simplification
```{r}
# maximal model
m1 <- lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled + log_t30_scaled:NatOrHumDom + (1 | Source_ID) + (1 | SS) + (1 | SSB), data = model_data_t30)

# Model without interaction
m2 <- lmer(logMass_total ~ LUI + nathab_scaled + log_t30_scaled +
             (1 | Source_ID) + (1 | SS) + (1 | SSB),
           data = model_data_t30)

summary(m1)

anova(m1, m2)  # Compare full model (with interaction) to simpler model

# p-value is > 0.05, the interaction is not contributing significantly → proceed with m2.

# Step 2: Remove log_t30_scaled 
m3 <- lmer(logMass_total ~ LUI + nathab_scaled +
             (1 | Source_ID) + (1 | SS) + (1 | SSB),
           data = model_data_t30)

anova(m2, m3)

# Step 3: Remove nathab_scaled 
m4 <- lmer(logMass_total ~ LUI +
             (1 | Source_ID) + (1 | SS) + (1 | SSB),
           data = model_data_t30)

anova(m3, m4) # only just not significant

AIC(m1, m2, m3, m4)

summary(m4)

# Stepwise backward selection for fixed effects
step(m1)

```

# Have a look at model residuals
```{r}

residuals_m4 <- resid(m4)

# 2. Plot residuals vs fitted values
plot(fitted(m4), residuals_m4,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red")

hist(residuals_m4, breaks = 50, main = "Histogram of Residuals",
     xlab = "Residuals", col = "lightblue")

qqnorm(residuals_m4)
qqline(residuals_m4, col = "red")

library(performance)
check_model(m4)

rm(add_var, m1, m2, m3, m4)
```


# Start with biomass ~ LU * Order analysis

```{r}
# Create a separate data frame
orders_size <- all_insects_size 

# Lets look at the number of sites per order
orders_size %>%
  group_by(Order) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Orders to keep and assign all others to "other_orders"
target_orders <- c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera", "Hemiptera")

orders_size <- orders_size %>%
  mutate(Order = case_when(
    Order %in% target_orders ~ Order,
    TRUE ~ "Other_orders"
  ))

unique(orders_size$Order)
```


# (NEW CODE) the big five orders - up to five size values per site

```{r}
# Calculate the biomass of each taxa
orders_size <- orders_size %>% 
  mutate(
    Biomass = mass_mg * Merged_diversity
  )

orders_size <- orders_size %>%
  group_by(SSBS, Order) %>% # Calculate for each site x order combination
  mutate(TotalSiteOrderBiomass = sum(Biomass, na.rm = TRUE)) %>%  # Calculate total site biomass for each order 
  ungroup() %>%
  group_by(SS) %>% # Now group by study
  mutate(
    SmallestNonZeroTotalSiteOrderBiomass = min(TotalSiteOrderBiomass[TotalSiteOrderBiomass > 0]), # Which site x order combo has the smallest total biomass greater than 0 per study? 
    Log_plus_what = ifelse(
    Diversity_metric_unit %in% c("individuals", "times observed"), min(mass_mg, na.rm = TRUE),  
    # If abundance data are counts, assign the smallest mass_mg for an individual of any species in the study
    SmallestNonZeroTotalSiteOrderBiomass / 2)) %>%  # else 1/2 the SmallestNonZeroTotalSiteOrderBiomass for any site within the study 
  ungroup() %>% 
  group_by(SSBS, Order) %>%  # Calculate for each site x order combination
  mutate(LogBiomass = log(TotalSiteOrderBiomass + Log_plus_what)) %>% 
  ungroup()
    
# Have a look
orders_size %>% dplyr::select(SS, SSBS, Order, Merged_diversity, mass_mg, Biomass, Diversity_metric_unit, TotalSiteOrderBiomass, SmallestNonZeroTotalSiteOrderBiomass, Log_plus_what, LogBiomass) %>% unique() %>%  View()
    
orders_size %>%  filter(SSBS == "AD1_2002__Vazquez 1 1 1") %>%  dplyr::select(SS, SSBS, Order, Merged_diversity, mass_mg, Biomass, Diversity_metric_unit, TotalSiteOrderBiomass, SmallestNonZeroTotalSiteOrderBiomass, Log_plus_what, LogBiomass) %>% unique() %>%  View()  

# Calculate one LogBiomass value per site x order combination
LogBiomass_summary <- orders_size %>%
  group_by(SSBS, Order) %>%
  summarise(
    n_TSOB = n_distinct(TotalSiteOrderBiomass), 
    n_LPW = n_distinct(Log_plus_what),
    is_consistent = (n_TSOB == 1 & n_LPW == 1), # sense check that all TotalSiteOrderBiomass and Log_plus_what values are the same for all rows in the site x order combination
    TotalSiteBiomass = first(TotalSiteOrderBiomass),
    Log_plus_what = first(Log_plus_what),
    LogBiomass = log(first(TotalSiteBiomass) + first(Log_plus_what)),
    .groups = "drop"
  )

# All sites are consistent
LogBiomass_summary %>% filter(is_consistent != TRUE)

inconsistent_groups <- orders_size %>%
  group_by(SSBS, Order) %>%
  summarise(
    n_TSB = n_distinct(TotalSiteOrderBiomass),
    n_LPW = n_distinct(Log_plus_what)
  ) %>%
  filter(n_TSB > 1 | n_LPW > 1) 

# No rows where LogBiomass is 0, NA, or blank
LogBiomass_summary %>%
  filter(is.na(LogBiomass) | LogBiomass == 0 | LogBiomass == "") %>% View()

rm(inconsistent_groups)
```

# (OLD CODE) the big five orders - up to five size values per site
```{r}
# Calculate abundance weighted log mass per Order
orders_size <- orders_size %>%
  group_by(SSBS, Order) %>%
  mutate(
    relativeAb_logMass = (log10_mass * Merged_diversity) / sum(Merged_diversity, na.rm = TRUE)
  ) %>%
  ungroup()

# Have a look at the worked exampled
orders_size %>% filter(SSBS == "AD1_2009__Vergara 1  6") %>% View()

# Calculate total log mass per ORDER within each site
logMass_summary <- orders_size %>%
  group_by(SSBS, Order) %>%
  summarise(
    total_div = sum(Merged_diversity, na.rm = TRUE),
    relAb_sum = sum(relativeAb_logMass, na.rm = TRUE),
    logMass_total = ifelse(total_div > 0, relAb_sum + log10(total_div), NA_real_) # this addresses -Inf values due to sites with 0 merged diversity, i.e. taking a log of 0.
  ) %>%
  dplyr::select(SSBS, Order, logMass_total)

# Have a look at the worked exampled
logMass_summary %>% filter(SSBS == "AD1_2009__Vergara 1  6") %>% View()

# View rows where logMass_total is 0, NA, or blank
logMass_summary %>%
  filter(is.na(logMass_total) | logMass_total == 0 | logMass_total == "") %>% View() # NAs are for Orders with 0 abundance across all sampled taxa for that order, so results in an NA rather than a 0

# Have a look 
orders_size %>% filter(SSBS == "AD1_2002__Vazquez 1 1 2") %>% View()
```

# Merge with PREDICTS but this time keep a row for each order within each site!
```{r}
# Create the other orders column so we filter for the correct number of sites
insects <- insects %>%
  mutate(Order = case_when(
    Order %in% target_orders ~ Order,
    TRUE ~ "Other_orders"
  ))

# Use distinct() to keep only the first appearance of each Order within each site
ssbs_by_order <- insects %>%
  distinct(SSBS, Order, .keep_all = TRUE) # 23,381 sites!

# merge LogBiomass_summary values
BiomassByOrder <- ssbs_by_order %>%
  left_join(LogBiomass_summary, by = c("SSBS", "Order")) %>% 
  droplevels()

# Have a look at sites with no LogBiomass
no_biomass <- BiomassByOrder %>%
  filter(
    is.na(LogBiomass) |         # remove NA
    is.nan(LogBiomass)|        # remove NaN
    LogBiomass == ""            # remove empty strings
  )  # I think this is because these rows do not have a body size value, therefore no LogBiomass value. 

taxon_insects_updated %>% filter(SSBS == "AD1_2002__Vazquez 1 1 1") %>% View()

# filter out sites with no LogBiomass
BiomassByOrder <- BiomassByOrder %>%
  filter(
    !is.na(LogBiomass),         # remove NA
    !is.nan(LogBiomass),        # remove NaN
    LogBiomass != ""            # remove empty strings
  )

sum(!is.na(LogBiomass_summary$LogBiomass)) # number of site x order combos with data, sensitivity check as should match number of sites in BiomassByOrder 

rm(ssbs_by_order, no_biomass, LogBiomass_summary, orders_size)
```

# Prep model data
```{r}
# Create model data
big_five <- BiomassByOrder 

# Lets look at the number of sites per order
big_five %>%
  group_by(Order) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Rename predominant habitat since its not really habitat we're looking at
big_five <- rename(big_five,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
big_five <- big_five %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   #Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
   
       # collapse secondary age categories into one and include SV (indeterminate age)
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Young secondary vegetation" = "Secondary vegetation", 
                                         "Intermediate secondary vegetation" = "Secondary vegetation",
                                         "Mature secondary vegetation" = "Secondary vegetation",
                                         "Secondary vegetation (indeterminate age)" = "Secondary vegetation"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(big_five$Order, big_five$Predominant_land_use)

```


# Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(big_five[ , c("Predominant_land_use", "Use_intensity")])
```

# Complete cases and summary stats

```{r}
# Lets have a look at rows with an NA for land use or use intensity
big_five %>%
  filter(if_any(c(Predominant_land_use, Use_intensity), is.na)) %>% 
  dplyr::select(SS, SSBS, Predominant_land_use, LogBiomass, Use_intensity) %>% 
  View()

# drop all rows with NAs for land use and use intensity and drop empty levels
big_five <- drop_na(big_five, Predominant_land_use, Use_intensity) %>% droplevels() 

# check distribution of LogBiomass
hist(big_five$LogBiomass)

# Have a look at sample sizes
table(big_five$Predominant_land_use, big_five$Use_intensity)
table(big_five$Order, big_five$Predominant_land_use)

length(unique(big_five$SS)) # 455 studies
length(unique(big_five$SSBS)) # 12232 sites

# Save the data frame for ease in the future
#saveRDS(big_five, "BigFiveBiomass.rds")
```

# Modelling biomass ~ LU * Order

```{r}
# Set order as a factor variable for modelling
big_five <- big_five %>%
  mutate(Order = as.factor(Order))

# maximal model
m1 <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = big_five)

# remove block
m2 <-lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS),
    data = big_five)

# remove study
m3 <-lmer(LogBiomass ~ Predominant_land_use + (1 | Source_ID),
    data = big_five)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)

rm(m2, m3)
```

# Coleoptera effects
```{r}
# gather the effects and confidence intervals using simulation. Simulate 1000 times
col_eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  col_eff$term <- as.factor(col_eff$term)
  
   # pull out effects
  col_eff <- col_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  col_eff$term <- recode_factor(col_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  col_eff <- col_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    col_eff <- col_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      col_eff[1,7] <- 0 #shift median
      col_eff[1,8] <- 0 #shift upper
      col_eff[1,9] <- 0 #shift lower
      
      colnames(col_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   col_eff <- col_eff %>% mutate(Order = "Coleoptera")

  # make the order column a factor so it can be recoded
   col_eff$Order <- as.factor(col_eff$Order)      
      
rm(m1)
```

# Diptera effects
```{r}
# Relevel model with Diptera as reference level
dip_ref <- within(big_five, Order <- relevel(Order, ref = "Diptera"))

# dip model
dip_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = dip_ref)

summary(dip_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
dip_eff <- FEsim(dip_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  dip_eff$term <- as.factor(dip_eff$term)
  
   # pull out effects
  dip_eff <- dip_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  dip_eff$term <- recode_factor(dip_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  dip_eff <- dip_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    dip_eff <- dip_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      dip_eff[1,7] <- 0 #shift median
      dip_eff[1,8] <- 0 #shift upper
      dip_eff[1,9] <- 0 #shift lower
      
      colnames(dip_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   dip_eff <- dip_eff %>% mutate(Order = "Diptera")

  # make the order column a factor so it can be recoded
   dip_eff$Order <- as.factor(dip_eff$Order)      
      
rm(dip_ref, dip_m)
```

# Hemiptera effects
```{r}
# Relevel model with Hemiptera as reference level
hemi_ref <- within(big_five, Order <- relevel(Order, ref = "Hemiptera"))

# dip model
hemi_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = hemi_ref)

summary(hemi_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
hemi_eff <- FEsim(hemi_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  hemi_eff$term <- as.factor(hemi_eff$term)
  
   # pull out effects
  hemi_eff <- hemi_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  hemi_eff$term <- recode_factor(hemi_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hemi_eff <- hemi_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    hemi_eff <- hemi_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      hemi_eff[1,7] <- 0 #shift median
      hemi_eff[1,8] <- 0 #shift upper
      hemi_eff[1,9] <- 0 #shift lower
      
      colnames(hemi_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   hemi_eff <- hemi_eff %>% mutate(Order = "Hemiptera")

  # make the order column a factor so it can be recoded
   hemi_eff$Order <- as.factor(hemi_eff$Order)      
      
rm(hemi_ref, hemi_m)
```

# Hymenoptera effects
```{r}
# Relevel model with Hymenoptera as reference level
hym_ref <- within(big_five, Order <- relevel(Order, ref = "Hymenoptera"))

# dip model
hym_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = hym_ref)

summary(hym_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
hym_eff <- FEsim(hym_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  hym_eff$term <- as.factor(hym_eff$term)
  
   # pull out effects
  hym_eff <- hym_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  hym_eff$term <- recode_factor(hym_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  hym_eff <- hym_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    hym_eff <- hym_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      hym_eff[1,7] <- 0 #shift median
      hym_eff[1,8] <- 0 #shift upper
      hym_eff[1,9] <- 0 #shift lower
      
      colnames(hym_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   hym_eff <- hym_eff %>% mutate(Order = "Hymenoptera")

  # make the order column a factor so it can be recoded
   hym_eff$Order <- as.factor(hym_eff$Order)     
      
rm(hym_ref, hym_m)
```

# Lepidoptera effects
```{r}
# Relevel model with Lepidoptera as reference level
lep_ref <- within(big_five, Order <- relevel(Order, ref = "Lepidoptera"))

# dip model
lep_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = lep_ref)

summary(lep_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
lep_eff <- FEsim(lep_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  lep_eff$term <- as.factor(lep_eff$term)
  
   # pull out effects
  lep_eff <- lep_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  lep_eff$term <- recode_factor(lep_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  lep_eff <- lep_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    lep_eff <- lep_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      lep_eff[1,7] <- 0 #shift median
      lep_eff[1,8] <- 0 #shift upper
      lep_eff[1,9] <- 0 #shift lower
      
      colnames(lep_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   lep_eff <- lep_eff %>% mutate(Order = "Lepidoptera")

  # make the order column a factor so it can be recoded
   lep_eff$Order <- as.factor(lep_eff$Order)     
      
rm(lep_ref, lep_m)
```

# Other_orders effects
```{r}
# Relevel model with Other_orders as reference level
other_ref <- within(big_five, Order <- relevel(Order, ref = "Other_orders"))

# other_orders model
other_m <- lmer(LogBiomass ~ Predominant_land_use * Order + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = other_ref)

summary(other_m) 

# gather the effects and confidence intervals using simulation. Simulate 1000 times
other_eff <- FEsim(other_m, n.sims = 1000)

  # make the term column a factor so it can be recoded
  other_eff$term <- as.factor(other_eff$term)
  
   # pull out effects
  other_eff <- other_eff %>%
    slice(1:6) 
  
  # rename factors for ease of visualisation
  other_eff$term <- recode_factor(other_eff$term, "(Intercept)" = "Primary",
                                #"Predominant_land_useYoung secondary vegetation" = "YSV",
                                #"Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                #"Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  other_eff <- other_eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    other_eff <- other_eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      other_eff[1,7] <- 0 #shift median
      other_eff[1,8] <- 0 #shift upper
      other_eff[1,9] <- 0 #shift lower
      
      colnames(other_eff)[1] <- "Predominant_land_use"
      
  # Add in a column of Order to make sure there is a grouping variable
   other_eff <- other_eff %>% mutate(Order = "Other_orders")

  # make the order column a factor so it can be recoded
   other_eff$Order <- as.factor(other_eff$Order)     
      
rm(other_ref, lep_m, other_m)
```

# Combine the effects and plot them
```{r}
combined <- rbind(col_eff, dip_eff, hemi_eff, hym_eff, lep_eff, other_eff)

# Add land-use factors for plotting
comb_plot_limits <- c("Primary", "Secondary vegetation", "Plantation", "Pasture", "Cropland", "Urban") 

# Add specific colours for each order
group_colours <- c(
  "#1b9e77", "#d95f02", "#7570b3",
  "#e7298a", "#66a61e", "#e6ab02"
)

combined <- combined[-c(27,33),]

  p <- combined %>%
  ggplot() +
  aes(x = Predominant_land_use, y = Percent_diff, colour = Order, group = Order) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper),
                 position = position_dodge(width = 0.5), linewidth = 1) +
  theme_classic() +
  scale_x_discrete(limits = comb_plot_limits) +
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype = 2) +
  theme(
    axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "bottom"        # 👈 Move legend
  ) +
  xlab("Land use intensity class") +
  ylab("log(biomass) difference (%)") +
  scale_colour_manual(values = group_colours) +
  guides(color = guide_legend(override.aes = list(shape = 19, size = 3))) +  # 👈 round icons
  scale_colour_manual(
  values = group_colours,
  labels = c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera",
             "Lepidoptera", "Other orders") # Change Other_orders to Other orders
) +
theme(
  legend.title = element_text(face = "bold")
) +
  scale_y_continuous(breaks = seq(-100, 150, by = 25))


 
        
#ggsave(filename = "../figures/C3 figures/big_five_LU_second.png",  plot = p, width = 8, height = 6, dpi = 300)
#saveRDS(combined, "/Users/jusi/Desktop/multi-groups/C3 data/C3-BigFive-AllEffectsCombined.rds")

rm(col_eff, combined, dip_eff, hemi_eff, hym_eff, lep_eff, p, other_eff)
```



# Average mass of an insect at a site (FOR ALL INSECTS)

```{r}
# Create a separate data frame for average weighted mean size of an insect at a site
avg_mass <- all_insects_size

# Calculate the biomass of each taxa
avg_mass <- avg_mass %>% 
  mutate(
    Biomass = mass_mg * Merged_diversity
  )

# Calculate AvgSiteBiomass
site_avg_mass <- avg_mass %>%
  group_by(SSBS) %>%
  summarise(TotalSiteBiomass = sum(Biomass, na.rm = TRUE),
         TotalSiteAbundance = sum(Merged_diversity, na.rm = TRUE),
         LogAvgSiteBiomass = log(TotalSiteBiomass / TotalSiteAbundance)
         ) %>%  
  ungroup() 

# View rows where LogAvgSiteBiomass is 0, NA, or blank - i.e. no total abundance!
site_avg_mass %>%
  filter(is.na(LogAvgSiteBiomass) | LogAvgSiteBiomass == 0 | LogAvgSiteBiomass == "") %>% View()
```

# Check if there are any studies where the average mass does not vary among sites and exclude them from the analysis
# If mass is the same in all sites, then either its a 1-species study or its a study where they're all in the same family/genus and we don't have any size data more finely resolved
# Either way, such studies cannot contribute to the model for average mass, so their inclusion would bias the results (it would flatten all the coefficients towards zero) so lets remove them

```{r}
# Get a list of the unique studies and their sites
unique_SS <- avg_mass %>% dplyr::select(SS, SSBS) %>% unique()

# First lets merge SS into out site_avg_mass data frame
site_avg_mass <- site_avg_mass %>% 
  left_join(unique_SS %>% dplyr::select(SSBS, SS), by = "SSBS") 

# Filter out those sites with NaN for LogAvgSiteBiomass due to log(0)
site_avg_mass <- site_avg_mass %>% filter(!is.nan(LogAvgSiteBiomass))

# Identify studies that contain sites with no variation in LogAvgSiteBiomass
no_variation <- site_avg_mass %>%
  group_by(SS) %>%
  summarise(n_unique = n_distinct(round(LogAvgSiteBiomass, 5))) %>% # round all values to 5 digits
  filter(n_unique == 1)

# Lets filter out those rows which come from studies with no variation in LogAvgSiteBiomass (i.e. the same average mass value across sites)
filtered_site_avg_mass <- site_avg_mass %>% filter(!SS %in% no_variation$SS)

# Sense check
any(filtered_site_avg_mass$SS %in% no_variation$SS) # are any of thise studies still present in my filtered data frame? 
```

# Calculating what % of taxa at a site have a unique mass_mg value. This captures both dimensions of the quality of the size data in that 1. If mass_mg values across taxa at a site are resolved to species level then the % of taxa with a unique mass_mg value will be high AND the taxa will all have different mass_mg values because they are resolved to species level.

```{r}
# Calculate the number of unique mass_mg values per site as a % of the overall number of rows per site
UniqueSiteMassValues <- avg_mass %>%
  group_by(SSBS) %>%
  summarise(
    unique_mass_mg = n_distinct(mass_mg),
    total_rows = n(),
    percent_unique = (unique_mass_mg / total_rows) * 100
  )

# Merge SS
UniqueSiteMassValues <- UniqueSiteMassValues %>% 
  left_join(unique_SS %>% dplyr::select(SSBS, SS), by = "SSBS") %>% 
  dplyr::select(SS, SSBS, unique_mass_mg, total_rows, percent_unique)

# Filter out those studies already identified with no variation in average mass among sites
FilteredUniqueSiteMassValues <- UniqueSiteMassValues %>% filter(!SS %in% no_variation$SS)

#all_insects_size %>% filter(SSBS == "SC1_2006__Benedick 1  1") %>% View()
#all_insects_size %>% filter(SSBS == "CC1_2007__Ewers 1 J 133") %>% View()
#avg_mass %>% filter(SS == "LH2_2009__Giangarelli 1") %>% View()
#site_avg_mass %>% filter(SS == "SS1_2011__Lange 1") %>% View()
#site_avg_mass %>% filter(SS == "SS1_2011__Lange 1") %>% distinct(round(LogAvgSiteBiomass, 5))
```

# Old calculations for average mass

```{r}
# Create a separate data frame for average weighted mean size of an insect at a site
avg_mass <- all_insects_size

# Calculate abundance weighted log10 average size/mass of an insect at a site
# So going from log10_mass -> ab_weighted_log10_mass
avg_mass <- avg_mass %>%
  group_by(SSBS) %>%
  mutate(
    relativeAb_logMass = (log10_mass * Merged_diversity) / sum(Merged_diversity, na.rm = TRUE)
  ) %>%
  ungroup()

# Now, to calculate the ab weighted average mass of an insect at a site we need to sum all of the mass values at a site and then divide by the sites total abundance, this will give us the ab weighted average mass value of an insect at a site.
site_avg_mass <- avg_mass %>%
  group_by(SSBS) %>%
  summarise(
    sum_relativeAb_logMass = sum(relativeAb_logMass, na.rm = TRUE),
    total_Merged_diversity = sum(Merged_diversity, na.rm = TRUE),
    average_relativeAb_logMass = sum_relativeAb_logMass / total_Merged_diversity
  ) 
```

# Merge with PREDICTS insects
```{r}
# Use distinct() to keep only the first appearance of each site within the overall PREDICTS div extract for insects
insects_single <- insects %>% 
  distinct(SSBS, .keep_all = TRUE)

 # Drop the extra SS column 
filtered_site_avg_mass <- filtered_site_avg_mass %>% dplyr::select(-SS)

# merge site level avg mass values
final_avg_mass <- insects_single %>%
  left_join(filtered_site_avg_mass, by = "SSBS") %>% 
  droplevels()

# Have a look at sites with no value for average biomass, either 0 total abundance or no body size value assigned
no_avg_biomass <- final_avg_mass %>%
  filter(
    is.na(LogAvgSiteBiomass) |         # remove NA
    is.nan(LogAvgSiteBiomass)|        # remove NaN
    LogAvgSiteBiomass == 0 |            # remove 0
    LogAvgSiteBiomass == ""            # remove empty strings
  ) 

# filter out sites with no avg mass value 
final_avg_mass <- final_avg_mass %>%
  filter(
    !is.na(LogAvgSiteBiomass),         # remove NA
    !is.nan(LogAvgSiteBiomass),        # remove NaN
    LogAvgSiteBiomass != 0,            # remove 0
    LogAvgSiteBiomass != ""            # remove empty strings
  )

# 10646 sites in total

# These are the studies with the same average mass value across sites
site_avg_mass %>% filter(SS %in% no_variation$SS) %>% View()

rm(insects_single, avg_mass, no_avg_biomass, site_avg_mass, unique_SS, no_variation, filtered_site_avg_mass, UniqueSiteMassValues)
```

# Tidy the LU classes

```{r}
# Rename predominant habitat since its not really habitat we're looking at
final_avg_mass <- rename(final_avg_mass,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
final_avg_mass <- final_avg_mass %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(final_avg_mass$Predominant_land_use, final_avg_mass$Use_intensity)
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
final_avg_mass <- final_avg_mass %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# Have a look at sample sizes
table(final_avg_mass$LUI)

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 61 sites, merge with light use
final_avg_mass <- final_avg_mass %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Have a look at sample sizes
table(final_avg_mass$LUI)

# Set the reference levels for LUI
final_avg_mass <- final_avg_mass %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )
```

# Create model data for average mass

```{r}
# drop all rows with NAs for land use and use intensity and drop empty levels
model_data_avg_mass <- drop_na(final_avg_mass, Predominant_land_use, Use_intensity) %>% droplevels() 

# check distributions
hist(model_data_avg_mass$LogAvgSiteBiomass)

# Have a look at sample sizes
table(model_data_avg_mass$Predominant_land_use, model_data_avg_mass$Use_intensity)
table(model_data_avg_mass$LUI)

length(unique(model_data_avg_mass$SS)) # 372 studies 
length(unique(model_data_avg_mass$SSBS)) # 8621 sites
```

# Figure which removes studies with < 10% unique size values across taxa within a site for LU

```{r}
lessthan10 <- FilteredUniqueSiteMassValues %>%
  filter(percent_unique < 10) %>%
  pull(SS) %>% 
  unique()

model10 <- model_data_avg_mass %>%
  filter(!SS %in% lessthan10)

# One for LU
reduced_model <- lmer(LogAvgSiteBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model10)

# model estimates
summary(reduced_model)
anova(reduced_model)
model_plot(reduced_model)


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(reduced_model, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(average mass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
  #ggsave(filename = "../figures/avg_mass_LU_10percent.png",  plot = p, width = 6, height = 4, dpi = 300)

mass_sensitivity_eff <- eff %>% mutate(Analysis = "Sensitivity analysis")
  
rm(eff, reduced_model, model10, lessthan10, p)     

```

# # Figure which removes studies with < 10% unique size values across taxa within a site for LUI 

```{r}
lessthan10 <- FilteredUniqueSiteMassValues %>%
  filter(percent_unique < 10) %>%
  pull(SS) %>% 
  unique()

model10 <- model_data_avg_mass %>%
  filter(!SS %in% lessthan10)

reduced_model <- lmer(LogAvgSiteBiomass ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model10)

# model estimates
summary(reduced_model)
anova(reduced_model)
      
      
      # gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(reduced_model, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
 # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary (minimal)",
                            "LUIPrimary_Light_use" = "Primary (light)",
                                "LUIPrimary_Intense_use" = "Primary (intense)",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV (minimal)",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV (light)",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV (intense)",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV (minimal)",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV (high)",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV (minimal)",
                            "LUIMature_secondary_vegetation_High_use" = "MSV (high)",
                            "LUIPlantation_forest_Minimal_use" = "Plantation (minimal)",
                            "LUIPlantation_forest_Light_use" = "Plantation (light)",
                            "LUIPlantation_forest_Intense_use" = "Plantation (intense)",
                            "LUIPasture_Minimal_use" = "Pasture (minimal)",
                            "LUIPasture_Light_use" = "Pasture (light)",
                            "LUIPasture_Intense_use" = "Pasture (intense)",
                            "LUICropland_Minimal_use" = "Cropland (minimal)",
                            "LUICropland_Light_use" = "Cropland (light)",
                            "LUICropland_Intense_use" = "Cropland (intense)",
                            "LUIUrban_Minimal_use" = "Urban (minimal)",
                            "LUIUrban_Light_use" = "Urban (light)",
                            "LUIUrban_Intense_use" = "Urban (intense)")

                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    p <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, colour = "red")+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1, colour = "red")+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(average mass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
       #ggsave(filename = "../figures/avg_mass_LUI_10percent.png",  plot = p, width = 8, height = 7, dpi = 300)
    
    mass_LUI_sensitivity_eff <- eff %>% mutate(Data = "Studies with homogenous size data excluded")
     
rm(eff, m1, m2, m3, p)    

```

# Modelling AVG_MASS ~ LU and creating the LU plot for all insects

```{r}
# maximal model
m1 <- lmer(LogAvgSiteBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_avg_mass)

# remove block
m2 <-lmer(LogAvgSiteBiomass ~ Predominant_land_use + (1 | Source_ID) + (1 | SS),
    data = model_data_avg_mass)

# remove study
m3 <-lmer(LogAvgSiteBiomass ~ Predominant_land_use + (1 | Source_ID),
    data = model_data_avg_mass)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))

# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(average mass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
  #ggsave(filename = "../figures/avg_mass_LU.png",  plot = p, width = 6, height = 4, dpi = 300)
  #avgmass_eff <- eff %>% mutate(Analysis = "Non sensitivity analysis")
  
rm(eff, m1, m2, m3, p)     
```

# Combine coefficients from non-sensitivity and sensitivity analysis of mass (LU)

```{r}
avgmass_eff <- avgmass_eff %>% mutate(Data = "All of the data")
mass_sensitivity_eff <- mass_sensitivity_eff %>% mutate(Data = "Studies with homogenous size data excluded")

combined <- rbind(mass_sensitivity_eff, avgmass_eff)

p <- combined %>%
  ggplot(aes(x = Predominant_land_use,
             y = Percent_diff,
             colour = Data,
             group = Data)) +
  
  # zero line
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  
  # confidence intervals for each Variable
  geom_linerange(aes(ymin = Percent_lower,
                     ymax = Percent_upper),
                 linewidth = 1,
                 position = position_dodge(width = 0.75)) +
  
  # points for each Variable
  geom_point(size = 3,
             position = position_dodge(width = 0.75)) +
  
  # manual colours
  scale_colour_manual(values = c(
    "All of the data" = "black",
    "Studies with homogenous size data excluded" = "red"
  )) +
  
  # aesthetics & theme
  theme_classic() +
  geom_vline(xintercept = 1.5, linetype = 2) +
  geom_vline(xintercept = 2.5, linetype = 2) +
  geom_vline(xintercept = 3.5, linetype = 2) +
  geom_vline(xintercept = 4.5, linetype = 2) +
  geom_vline(xintercept = 5.5, linetype = 2) +
  geom_vline(xintercept = 6.5, linetype = 2) +
  geom_vline(xintercept = 7.5, linetype = 2) +
  
  theme(
    axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "bottom"
  ) +
  
  xlab("Land use intensity class") +
  ylab("log(average mass) difference (%)") +
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 6, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = 8)) +
  theme(legend.title = element_text(face = "bold"),
        legend.text = element_text(size = 10)) 

  #ggsave(filename = "../figures/C3-sensitivity-mass.png",  plot = p, width = 8, height = 5, dpi = 300)
  #saveRDS(combined, "../figures/C3 figures/C3-mass-LU-sensitivity-df.rds")


```

# Bens pesticide meta analysis code
```{r}
data <- read_xlsx(file.choose())

data <- strings2factors(data)

summary(data)


invertebrates <- data %>% filter(`Primary classification of non-target organism` == "Invertebrate")

filter_inverts <- invertebrates %>% dplyr::select(c(`Non-target organism name`, `Non-target organism Latin name`, `Secondary classification of non-target organism`, `Primary classification of non-target organism`))

arthropods <- filter_inverts %>% filter(`Secondary classification of non-target organism` == "Arthropod")

unique_names <- arthropods %>% dplyr::select(`Non-target organism Latin name`) %>% distinct()

unique_names_filtered <- unique_names %>%
  filter(str_count(`Non-target organism Latin name`, "\\S+") == 2) %>% 
  filter(!str_detect(`Non-target organism Latin name`, "^\\S+\\s+(sp\\.|spp\\.)$"))


non_sp_names <- unique_names %>% filter(!`Non-target organism Latin name` %in% unique_names_filtered$`Non-target organism Latin name`) %>%
  filter(!str_detect(`Non-target organism Latin name`, "^\\S+\\s+(sp\\.|spp\\.)$"))





matching_rows <- unique_names_filtered %>%
  filter(`Non-target organism Latin name` %in% all_insects_size$Best_guess_binomial)






colnames(invertebrates)


unique(levels(data$`Primary classification of non-target organism`))

```

# Modelling AVG_MASS ~ LUI (LU + UI) and creating the LUI plot for all insects

```{r}
# maximal model
m1 <- lmer(LogAvgSiteBiomass ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_avg_mass)

# remove block
m2 <-lmer(LogAvgSiteBiomass ~ LUI + (1 | Source_ID) + (1 | SS),
    data = model_data_avg_mass)

# remove study
m3 <-lmer(LogAvgSiteBiomass ~ LUI + (1 | Source_ID),
    data = model_data_avg_mass)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary (minimal)",
                            "LUIPrimary_Light_use" = "Primary (light)",
                                "LUIPrimary_Intense_use" = "Primary (intense)",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV (minimal)",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV (light)",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV (intense)",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV (minimal)",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV (high)",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV (minimal)",
                            "LUIMature_secondary_vegetation_High_use" = "MSV (high)",
                            "LUIPlantation_forest_Minimal_use" = "Plantation (minimal)",
                            "LUIPlantation_forest_Light_use" = "Plantation (light)",
                            "LUIPlantation_forest_Intense_use" = "Plantation (intense)",
                            "LUIPasture_Minimal_use" = "Pasture (minimal)",
                            "LUIPasture_Light_use" = "Pasture (light)",
                            "LUIPasture_Intense_use" = "Pasture (intense)",
                            "LUICropland_Minimal_use" = "Cropland (minimal)",
                            "LUICropland_Light_use" = "Cropland (light)",
                            "LUICropland_Intense_use" = "Cropland (intense)",
                            "LUIUrban_Minimal_use" = "Urban (minimal)",
                            "LUIUrban_Light_use" = "Urban (light)",
                            "LUIUrban_Intense_use" = "Urban (intense)")
                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    MassLUIPlot <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, colour = "blue")+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1, colour = "blue")+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(average mass) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10))) +
      scale_y_continuous(
  breaks = seq(-25, 50, by = 25)   # add more ticks
)

     
       #ggsave(filename = "../figures/avg_mass_LUI.png",  plot = p, width = 8, height = 7, dpi = 300)
       
       mass_LUI_eff <- eff %>% mutate(Variable = "Average mass")
       mass_LUI_eff <- eff %>% mutate(Data = "All of the data")
     
rm(eff, m1, m2, m3, p)     
```

# Combine coefficients from non-sensitivity and sensitivity analysis of mass (LUI)

```{r}
avgmass_eff <- avgmass_eff %>% mutate(Data = "All of the data")
mass_sensitivity_eff <- mass_sensitivity_eff %>% mutate(Data = "Studies with homogenous size data excluded")

combined <- rbind(mass_LUI_eff, mass_LUI_sensitivity_eff)

p <- combined %>%
  ggplot(aes(x = Predominant_land_use,
             y = Percent_diff,
             colour = Data,
             group = Data)) +
  
  # zero line
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  
  # confidence intervals for each Variable
  geom_linerange(aes(ymin = Percent_lower,
                     ymax = Percent_upper),
                 linewidth = 1,
                 position = position_dodge(width = 0.75)) +
  
  # points for each Variable
  geom_point(size = 3,
             position = position_dodge(width = 0.75)) +
  
  # manual colours
  scale_colour_manual(values = c(
    "All of the data" = "black",
    "Studies with homogenous size data excluded" = "red"
  )) +
  
  # aesthetics & theme
  theme_classic() +
   geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
  
  theme(
    axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "bottom"
  ) +
  
  xlab("Land use intensity class") +
  ylab("log(average mass) difference (%)") +
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 6, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = 8)) +
  theme(legend.title = element_text(face = "bold"),
        legend.text = element_text(size = 10)) 

  #ggsave(filename = "../figures/C3 figures/C3-sensitivity-mass-LUI.png",  plot = p, width = 10, height = 6, dpi = 300)
  #saveRDS(combined, "../figures/C3 figures/C3-mass-LU-sensitivity-df.rds")
  
combined <- read_rds(file.choose())  

  combined <- combined %>% mutate(percentdiff = round(Percent_upper - Percent_lower, 2)


```


# Total abundance responses to LU

```{r}
# All-insect set of site-level log(abundance) and species-richness data
# Has filtered out accidentally sampled larvae
abundance <-  qread("/Users/jusi/Desktop/multi-groups/data/abundance-data-prepared-all-insects.qs")

# Use dplyr to reshape the abundance data
# For each SSBS: sum the total abundance values then take the mean log_plus_what (they'll all be the same - comes from the minimum non-zero value across the entire study) 
# logAllInsectAbundance = log(sumTotalAbundance + mean_log_plus_what) - an overall abundance value for all of the insects at each site!
abundance_summary <- abundance %>%
  group_by(SSBS) %>%
  summarise(
    sumTotalAbundance = sum(Total_abundance, na.rm = TRUE),
    mean_log_plus_what = mean(Log_plus_what, na.rm = TRUE)
  ) %>%
  mutate(logAllInsectAbundance = log(sumTotalAbundance + mean_log_plus_what))

# Filter PREDICTS for same set of sites as in my biomass model data
filtered_abundance <-  abundance_summary %>%
  filter(SSBS %in% model_data_biomass$SSBS) %>%
  distinct(SSBS, .keep_all = TRUE)

# Append use intensity based on SSBS
filtered_abundance <- filtered_abundance %>%
  left_join(
    model_data_biomass %>% dplyr::select(SS, SSBS, Predominant_land_use, Use_intensity, Source_ID, SSB),
    by = "SSBS"
  )
```

# Start the modelling of all of the insects (Abundance responses)

```{r}
# Relevel the land use and use intensity classes
filtered_abundance <- filtered_abundance %>%
  mutate(
    # convert back to factors and set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(filtered_abundance$Predominant_land_use, filtered_abundance$Use_intensity)

levels(filtered_abundance$Predominant_land_use)
levels(filtered_abundance$Use_intensity)
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
filtered_abundance <- filtered_abundance %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 61 sites, merge with light use
filtered_abundance <- filtered_abundance %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Set the reference levels for LUI
filtered_abundance <- filtered_abundance %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )
```

# Create model data for abundance

```{r}
# drop all rows with NAs for land use, use intensity and log_abundance_px and drop empty levels
model_data_abundance <- drop_na(filtered_abundance, Predominant_land_use, Use_intensity, logAllInsectAbundance) %>% droplevels() 

# check distributions
hist(model_data_abundance$logAllInsectAbundance)
hist(abundance$Log_abundance_pX)

# Have a look at sample sizes
table(model_data_abundance$Predominant_land_use, model_data_abundance$Use_intensity)
table(model_data_abundance$LUI)

length(unique(model_data_abundance$SS)) # 444 studies 
length(unique(model_data_abundance$SSBS)) # 11516 sites

  length(unique(model_data_biomass$SS)) # 444 studies
 length(unique(model_data_biomass$SSBS)) # 11516 sites

  length(unique(model_data_avg_mass$SS)) # 372 studies
  length(unique(model_data_avg_mass$SSBS)) # 8621 sites
  
       

rm(abundance, abundance_summary, filtered_abundance)
```

# Modelling ABUNDANCE ~ LU and creating the LU plot for all insects

```{r}
# maximal model
m1 <- lmer(logAllInsectAbundance ~ Predominant_land_use + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_abundance)

# remove block
m2 <-lmer(logAllInsectAbundance ~ Predominant_land_use + (1 | Source_ID) + (1 | SS),
    data = model_data_abundance)

# remove study
m3 <-lmer(logAllInsectAbundance ~ Predominant_land_use + (1 | Source_ID),
    data = model_data_abundance)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary",
                                "Predominant_land_useYoung secondary vegetation" = "YSV",
                                "Predominant_land_useIntermediate secondary vegetation" = "ISV",
                                "Predominant_land_useMature secondary vegetation" = "MSV",
                                "Predominant_land_useSecondary vegetation" = "Secondary vegetation",
                                "Predominant_land_usePlantation forest" = "Plantation",
                                "Predominant_land_usePasture" = "Pasture",
                                "Predominant_land_useCropland" = "Cropland",
                                "Predominant_land_useUrban" = "Urban")
  
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
     p <- eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, position = position_dodge(width = 0.5))+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), position = position_dodge(width = 0.5), linewidth = 1)+
        theme_classic()+
        geom_vline(xintercept = 1.5, linetype = 2)+
        geom_vline(xintercept = 2.5, linetype = 2)+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 4.5, linetype = 2)+
        geom_vline(xintercept = 5.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 7.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(total abundance) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10)))
     
  #ggsave(filename = "../figures/abundance_LU.png",  plot = p, width = 6, height = 4, dpi = 300)
     
     
     abundance_eff <- eff %>% mutate(Variable = "Abundance")
  
rm(eff, m1, m2, m3, p)     
```

# Combine Biomass, Mass and Abundance into one plot with three different colours

```{r}
biomass_eff <- biomass_eff %>% mutate("Response variable" = "Total biomass")
avgmass_eff <- avgmass_eff %>% mutate("Response variable" = "Average mass")
abundance_eff <- abundance_eff %>% mutate("Response variable" = "Species abundance")


combined <- rbind(biomass_eff, avgmass_eff, abundance_eff)


p <- combined %>%
  ggplot(aes(x = Predominant_land_use,
             y = Percent_diff,
             colour = Variable,
             group = Variable)) +
  
  # zero line
  geom_hline(yintercept = 0, linewidth = 0.5, color = "black") +
  
  # confidence intervals for each Variable
  geom_linerange(aes(ymin = Percent_lower,
                     ymax = Percent_upper),
                 linewidth = 1,
                 position = position_dodge(width = 0.75)) +
  
  # points for each Variable
  geom_point(size = 3,
             position = position_dodge(width = 0.75)) +
  
  # manual colours
  scale_colour_manual(values = c(
    "Total biomass" = "purple",
    "Average mass" = "blue",
    "Species abundance" = "red"
  )) +
  
  # aesthetics & theme
  theme_classic() +
  geom_vline(xintercept = 1.5, linetype = 2) +
  geom_vline(xintercept = 2.5, linetype = 2) +
  geom_vline(xintercept = 3.5, linetype = 2) +
  geom_vline(xintercept = 4.5, linetype = 2) +
  geom_vline(xintercept = 5.5, linetype = 2) +
  geom_vline(xintercept = 6.5, linetype = 2) +
  geom_vline(xintercept = 7.5, linetype = 2) +
  
  theme(
    axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    panel.border = element_rect(colour = "black", fill = NA),
    legend.position = "bottom"
  ) +
  
  xlab("Land use intensity class") +
  ylab("Difference in log response variable (%)") +
    guides(color = guide_legend(
      override.aes=list(shape = 16, size = 6, linetype = 0))) +
        scale_y_continuous(breaks = scales::pretty_breaks(n = 8)) +
  theme(legend.title = element_text(face = "bold"),
        legend.text = element_text(size = 10)) 

  #ggsave(filename = "../figures/C3 figures/C3-combined-plots.png",  plot = p, width = 8, height = 5, dpi = 300)
  #saveRDS(combined, "../figures/C3 figures/C3-combined-df.rds")

```

# Modelling ABUNDANCE ~ LUI (LU + UI) and creating the LUI plot for all insects

```{r}
# maximal model
m1 <- lmer(logAllInsectAbundance ~ LUI + (1 | Source_ID) + (1 | SS) + (1 | SSB), 
           data = model_data_abundance)

# remove block
m2 <-lmer(logAllInsectAbundance ~ LUI + (1 | Source_ID) + (1 | SS),
    data = model_data_abundance)

# remove study
m3 <-lmer(logAllInsectAbundance ~ LUI + (1 | Source_ID),
    data = model_data_abundance)

AIC(m1, m2, m3) # m1 best model

# model estimates
summary(m1)
anova(m1)
model_plot(m1)
coef(summary(m1))


# gather the effects and confidence intervals using simulation. Simulate 1000 times
eff <- FEsim(m1, n.sims = 1000)

  # make the term column a factor so it can be recoded
  eff$term <- as.factor(eff$term)
  
  # rename factors for ease of visualisation
  eff$term <- recode_factor(eff$term, "(Intercept)" = "Primary (minimal)",
                            "LUIPrimary_Light_use" = "Primary (light)",
                                "LUIPrimary_Intense_use" = "Primary (intense)",
                            "LUIYoung_secondary_vegetation_Minimal_use" = "YSV (minimal)",
                            "LUIYoung_secondary_vegetation_Light_use" = "YSV (light)",
                            "LUIYoung_secondary_vegetation_Intense_use" = "YSV (intense)",
                            "LUIIntermediate_secondary_vegetation_Minimal_use" = "ISV (minimal)",
                            "LUIIntermediate_secondary_vegetation_High_use" = "ISV (high)",
                            "LUIMature_secondary_vegetation_Minimal_use" = "MSV (minimal)",
                            "LUIMature_secondary_vegetation_High_use" = "MSV (high)",
                            "LUIPlantation_forest_Minimal_use" = "Plantation (minimal)",
                            "LUIPlantation_forest_Light_use" = "Plantation (light)",
                            "LUIPlantation_forest_Intense_use" = "Plantation (intense)",
                            "LUIPasture_Minimal_use" = "Pasture (minimal)",
                            "LUIPasture_Light_use" = "Pasture (light)",
                            "LUIPasture_Intense_use" = "Pasture (intense)",
                            "LUICropland_Minimal_use" = "Cropland (minimal)",
                            "LUICropland_Light_use" = "Cropland (light)",
                            "LUICropland_Intense_use" = "Cropland (intense)",
                            "LUIUrban_Minimal_use" = "Urban (minimal)",
                            "LUIUrban_Light_use" = "Urban (light)",
                            "LUIUrban_Intense_use" = "Urban (intense)")

                        
  # Add in upper and lower confidence intervals 
  eff <- eff %>%
    mutate(Upper_ci = (median + 1.96*sd)) %>%
    mutate(Lower_ci = (median - 1.96*sd))
  
  
    eff <- eff %>%
        mutate(
          Percent_diff = (((exp(median[1] + median) - 1) / (exp(median[1]) - 1))*100)-100) %>%
        mutate(
          Percent_upper = (((exp(median[1] + Upper_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100) %>%
        mutate(
          Percent_lower = (((exp(median[1] + Lower_ci) - 1) / (exp(median[1]) - 1)) * 100) - 100)
  
      eff[1,7] <- 0 #shift median
      eff[1,8] <- 0 #shift upper
      eff[1,9] <- 0 #shift lower
      
      colnames(eff)[1] <- "Predominant_land_use"
      
    AbundanceLUIPlot <-  eff %>%
        ggplot()+
        aes(x = Predominant_land_use, y = Percent_diff)+
        geom_hline(yintercept = 0, linewidth = 0.5, color = ("black"), linetype = 1)+
        geom_point(size = 3, colour = "red")+
        geom_linerange(aes(ymin = Percent_lower, ymax = Percent_upper), linewidth = 1, colour = "red")+
        theme_classic()+
        geom_vline(xintercept = 3.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 6.5, linetype = 2)+
        geom_vline(xintercept = 8.5, linetype = 2)+
        geom_vline(xintercept = 10.5, linetype = 2)+
        geom_vline(xintercept = 13.5, linetype = 2)+
        geom_vline(xintercept = 16.5, linetype = 2)+
        geom_vline(xintercept = 19.5, linetype = 2)+
        theme(axis.text.x = element_text(face= "bold", angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.title.y = element_text(face = "bold"),
              panel.border = element_rect(colour = "black",  fill=NA))+
        xlab("Land use intensity class")+
        ylab("log(total abundance) difference (%)") +
        guides(color = guide_legend(
          override.aes=list(shape = 15, size = 10))) 
      
     
       #ggsave(filename = "../figures/abundance_LUI.png",  plot = p, width = 8, height = 7, dpi = 300)
    
    AbundanceLUIPlot <- p
     
rm(eff, m1, m2, m3, p)     
```


# Combine all three LUI plots for biomass, mass and abundance. Stack them above one another with the x labels only appearing once

```{r}
library(patchwork)

p1 <- AbundanceLUIPlot + theme(axis.title.y = element_text(size = 14), axis.text.x = element_text(size = 14))


p1 <- AbundanceLUIPlot + theme(axis.title.y = element_text(size = 14), axis.text.x  = element_text(size = 14))
p2 <- MassLUIPlot + themeaxis.title.y  = element_text(size = 14), axis.text.x  = element_text(size = 14))
p3 <- BiomassLUIPlot + theme(axis.title.y = element_text(size = 14), axis.text.x  = element_text(size = 14))

final_plot <- (p1 / p2 / p3) &
  theme(legend.position = "bottom")

# Now blank FIRST TWO x axes again (after global theme):
final_plot[[1]] <- final_plot[[1]] + theme(axis.text.x = element_blank(),
                                           axis.title.x = element_blank())
final_plot[[2]] <- final_plot[[2]] + theme(axis.text.x = element_blank(),
                                           axis.title.x = element_blank())
  


  ggsave(filename = "../figures/C3 figures/C3-combined-LUI.png",  plot = final_plot, width = 18, height = 10, dpi = 300)
```


# Calculate taxonomic precision at a site
```{r}
# Get a taxonomic precision score for each site based on how well resolved each taxa at the site is!
# use the same machinery as calculating community weighted mean size

# Simplify df
test <- all_insects_size %>% dplyr::select(SS, SSBS, Order, Family, Best_guess_binomial, Genus, Species, Merged_diversity)

# Assign score based on taxonomy
test <- test %>%
  mutate(taxonomic_precision = case_when(
    !(is.na(Species) | Species == "") ~ 3,
    !(is.na(Genus) | Genus == "")     ~ 2,
    !(is.na(Family) | Family == "")   ~ 1,
    !(is.na(Order) | Order == "")     ~ 0,
    TRUE                              ~ NA_real_
  ))

# Get precision score per site
weighted_taxonomic_precision <- test %>%
  group_by(SSBS) %>%
  summarise(
    weighted_precision = sum(taxonomic_precision * Merged_diversity, na.rm = TRUE) / 
                         sum(Merged_diversity, na.rm = TRUE),
    .groups = "drop"
  )

# Classify each score
weighted_taxonomic_precision <- weighted_taxonomic_precision %>%
  mutate(precision_rank = case_when(
    weighted_precision < 1 ~ "Order",
    weighted_precision < 2 ~ "Family",
    weighted_precision < 3 ~ "Genus",
    weighted_precision == 3 ~ "Species",
    TRUE ~ "Unknown"
  ))

# Join this back to main df to assign score to each row
test <- test %>%
  left_join(weighted_taxonomic_precision, by = "SSBS")

# Filter for low-precision sites and count families, i.e. the number of rows resolved only to the family level of a given family
 test %>%
  filter(precision_rank %in% c("Order", "Family")) %>%
  count(Family, sort = TRUE) 
 
test %>% filter(SSBS == "AD1_2002__Kruess 1  7") %>% View()
```

```{r}
# Do the same but see which groups we do/don't have size data for
# The species we most want to get size values for are the really abundant ones in PREDICTS
# Particularly those where the abundances changes the most in response to land use

 # A list of species with > 1000 merged diversity and no size data
unique_species_high_div <- sp_no_size %>%
  filter(Merged_diversity > 10) %>%
  distinct(Best_guess_binomial, .keep_all = T) %>% 
  filter(!str_detect(Best_guess_binomial, "0"))  # Species entries with incorrect names!

sp_no_size %>%
  #filter(Merged_diversity > 100) %>%
  distinct(Best_guess_binomial, .keep_all = T) %>% 
  View()

# Family x genera combinations and number of rows without species level size data
 sp_no_size %>%
  group_by(Family, Genus) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) 
 
sp_no_size %>% filter(Best_guess_binomial == "Drosophila willistoni") %>% distinct() 

sp_no_size %>%
  filter(Merged_diversity > 100) %>%
  distinct(Best_guess_binomial, .keep_all = T)

rm(sp_no_size, weighted_taxonomic_precision, test)
```

# How much of the dataset is resolved to sp, genus and family level.

```{r}
# Count total number of rows
total_rows <- nrow(all_insects_size)

# Count how many are resolved to each level
species_count <- sum(!is.na(all_insects_size$Trait_Value))
genus_count <- sum(is.na(all_insects_size$Trait_Value) & !is.na(all_insects_size$genus_avg_body_length))
family_count <- sum(is.na(all_insects_size$Trait_Value) & is.na(all_insects_size$genus_avg_body_length) &
                    (!is.na(all_insects_size$family_avg_body_length) | !is.na(all_insects_size$Rainford_family_avg)))

# Calculate proportions
species_prop <- species_count / total_rows
genus_prop <- genus_count / total_rows
family_prop <- family_count / total_rows

# Display results
cat("Proportion resolved to species level:", species_prop, "\n")
cat("Proportion resolved to genus level:", genus_prop, "\n")
cat("Proportion resolved to family level:", family_prop, "\n")

```

# How many studies sample only bees or ants

```{r}
bee_families <- c("Apidae", "Halictidae", "Megachilidae", "Andrenidae", "Colletidae", "Melittidae", "Stenotritidae")
ant_family <- "Formicidae"

study_family_summary <- all_insects_size %>%
  group_by(SS) %>%
  summarize(unique_families = list(unique(Family)), .groups = "drop")

# Now evaluate each row's family list separately
only_bee <- study_family_summary %>%
  rowwise() %>%
  filter(all(unique_families %in% bee_families)) %>%
  ungroup() %>%
  nrow()

only_ants <- study_family_summary %>%
  rowwise() %>%
  filter(all(unique_families %in% ant_family)) %>%
  ungroup() %>%
  nrow()

cat("Number of studies sampling only bees:", only_bees, "\n")
cat("Number of studies sampling only ants:", only_ants, "\n")
```

# Extract studies for only bees or only ants

```{r}
# Get study IDs that sample only bees
only_bee_studies <- study_family_summary %>%
  rowwise() %>%
  filter(all(unique_families %in% bee_families)) %>%
  pull(SS)

# Get study IDs that sample only ants
only_ant_studies <- study_family_summary %>%
  rowwise() %>%
  filter(all(unique_families %in% ant_family)) %>%
  pull(SS)

# Filter the original data for just those studies

# Filter rows for studies that sampled only bees
bee_only_data <- all_insects_size %>%
  filter(SS %in% only_bee_studies)

# Filter rows for studies that sampled only ants
ant_only_data <- all_insects_size %>%
  filter(SS %in% only_ant_studies)

```

# Function to calculate taxonomic resolution

```{r}
calculate_resolution_proportions <- function(data) {
  total_rows <- nrow(data)
  
  species_count <- sum(!is.na(data$Trait_Value))
  genus_count <- sum(is.na(data$Trait_Value) & !is.na(data$genus_avg_body_length))
  family_count <- sum(is.na(data$Trait_Value) & is.na(data$genus_avg_body_length) &
                        (!is.na(data$family_avg_body_length) | !is.na(data$family_avg)))
  
  species_prop <- species_count / total_rows
  genus_prop <- genus_count / total_rows
  family_prop <- family_count / total_rows
  
  return(list(
    species_count = species_count,
    genus_count = genus_count,
    family_count = family_count,
    species_prop = species_prop,
    genus_prop = genus_prop,
    family_prop = family_prop
  ))
}

# For bee-only studies
bee_resolution <- calculate_resolution_proportions(bee_only_data)

# For ant-only studies
ant_resolution <- calculate_resolution_proportions(ant_only_data)

# Print results
cat("Bee-only studies:\n")
cat("Species-level:", bee_resolution$species_count, 
    " (", round(bee_resolution$species_prop * 100, 1), "%)\n", sep = "")
cat("Genus-level:", bee_resolution$genus_count, 
    " (", round(bee_resolution$genus_prop * 100, 1), "%)\n", sep = "")
cat("Family-level:", bee_resolution$family_count, 
    " (", round(bee_resolution$family_prop * 100, 1), "%)\n\n", sep = "")

cat("Ant-only studies:\n")
cat("Species-level:", ant_resolution$species_count, 
    " (", round(ant_resolution$species_prop * 100, 1), "%)\n", sep = "")
cat("Genus-level:", ant_resolution$genus_count, 
    " (", round(ant_resolution$genus_prop * 100, 1), "%)\n", sep = "")
cat("Family-level:", ant_resolution$family_count, 
    " (", round(ant_resolution$family_prop * 100, 1), "%)\n", sep = "")

```



# What's the variance in the abundance weighted average body size among sites compared to among studies

```{r}
# merge weighted_mean_size back into original df
variation_cwms <- left_join(taxon_insects_updated, weighted_mean_size_sites, by = "SSBS")

summary(aov(weighted_mean_size ~ SS, data = variation_cwms)) # 10% of the variation is explained by variation among sites!

rm(variation_cwms)
```


# Read in overall trait dataset
```{r}
# Read in trait data
traits <- read_excel("../data/All_FamilyandOrder_Traits_for_JustinIsip.xlsx") 

# As data frame 
traits <- as.data.frame(traits)

# Combine the first three rows into a single character vector (column names)
combined_names <- sapply(traits[1:3, ], function(x) paste(na.omit(x), collapse = " | "))

# Assign the new names to the columns
colnames(traits) <- combined_names

# Drop the first three rows now that they're in the header
traits <- traits[-c(1:3), ]

# Extract the first row (which contains new column names for the first 8 columns)
new_names_part <- as.character(unlist(traits[1, 1:8]))

# Replace the first 8 column names
colnames(traits)[1:8] <- new_names_part

# Remove the first row now that it's used as header
traits <- traits[-1, ]

# Reset rownames (just for cleanliness as they don't start from 1)
rownames(traits) <- NULL

# Lets have a look at a particular trait size category and its coverage across families
traits %>% dplyr::select("Family", "Taxon Name", "Minimum size | < 0.35 | mm") %>% View()

# Identify all traits present for a group
traits %>%
  pivot_longer(cols = 9:ncol(.), # all trait columns
               names_to = "Trait", # new column for trait names
               values_to = "Value") %>%# new column for the values
  filter(!is.na(Value) & Value != 0) %>%  # keep only traits that are present, i.e are not NA or 0
View()
  
# Pivot longer traits
pivot_trait <- traits %>%
  pivot_longer(
    cols = 9:ncol(.), # first 8 cols are meta data
    names_to = "Trait",
    values_to = "Value"
  ) %>%
  filter(!is.na(Value) & Value != "0") %>% # if the trait value is not an NA and is not equal to 0
  separate(Trait, into = c("Trait", "Modality", "Unit"), sep = "\\|", fill = "right", extra = "merge") %>% 
  mutate(Mode = ifelse(is.na(Unit), Modality, paste(Modality, Unit))) %>% # what is the trait category
  dplyr::select(-Modality, -Unit) %>% # remove redundant cols
  dplyr::select(-Value, Value)  # moves 'Value' to the end

# Strings 2 factor
pivot_trait <- strings2factors(pivot_trait)

# Pivot wider again
trait_matrix <- pivot_trait %>%
  group_by(Trait) %>% 
  mutate(row = row_number()) %>% 
  pivot_wider(
    names_from = Trait,
    values_from = Mode,
    values_fill = NA
  ) %>% 
  dplyr::select(-row)

# Remove trailing space after column names
names(trait_matrix) <- str_remove(names(trait_matrix), "\\s+$")

pivot_trait %>% filter(Trait == "Dispersal ability ") %>% View()
```

# Code up just for size

```{r}
# Filter trait database for meta data and just minimum and maximum size cols
size <- traits %>% dplyr::select(1:32)

# Rename the size columns so each one is a range
size <- size %>%
  rename(
    "Minimum size 0.17 - 0.35mm" = "Minimum size | < 0.35 | mm",
    "Minimum size 0.35 - 0.71mm" = "Minimum size | < 0.71 | mm",
    "Minimum size 0.71 - 1.41mm" = "Minimum size | < 1.41 | mm",
    "Minimum size 1.41 - 2.82mm" = "Minimum size | < 2.82 | mm",
    "Minimum size 2.82 - 5.62mm" = "Minimum size | < 5.62 | mm",
    "Minimum size 5.62 - 11.22mm" = "Minimum size | < 11.22 | mm",
    "Minimum size 11.22 - 22.39mm" = "Minimum size | < 22.39 | mm",
    "Minimum size 22.39 - 44.67mm" = "Minimum size | < 44.67 | mm",
    "Minimum size 44.67 - 89.13mm" = "Minimum size | < 89.13 | mm",
    "Minimum size 89.13 - 177.83mm" = "Minimum size | < 177.83 | mm",
    "Minimum size 177.83 - 354.81mm" = "Minimum size | < 354.81 | mm",
    "Minimum size 354.81 - 707.95mm" = "Minimum size | < 707.95 | mm",
    "Maximum size 0.17 - 0.35mm" = "Maximum size | < 0.35 | mm",
    "Maximum size 0.35 - 0.71mm" = "Maximum size | < 0.71 | mm",
    "Maximum size 0.71 - 1.41mm" = "Maximum size | < 1.41 | mm",
    "Maximum size 1.41 - 2.82mm" = "Maximum size | < 2.82 | mm",
    "Maximum size 2.82 - 5.62mm" = "Maximum size | < 5.62 | mm",
    "Maximum size 5.62 - 11.22mm" = "Maximum size | < 11.22 | mm",
    "Maximum size 11.22 - 22.39mm" = "Maximum size | < 22.39 | mm",
    "Maximum size 22.39 - 44.67mm" = "Maximum size | < 44.67 | mm",
    "Maximum size 44.67 - 89.13mm" = "Maximum size | < 89.13 | mm",
    "Maximum size 89.13 - 177.83mm" = "Maximum size | < 177.83 | mm",
    "Maximum size 177.83 - 354.81mm" = "Maximum size | < 354.81 | mm",
    "Maximum size 354.81 - 707.95mm" = "Maximum size | < 707.95 | mm",
  )

# Calculate the logs
(log(0.17) + log(0.35)) / 2 # -1.410889
(log(0.35) + log(0.71)) / 2 # -0.6961562
(log(0.71) + log(1.41)) / 2 # 0.0005496977
(log(1.41) + log(2.82)) / 2 # 0.6901633
(log(2.82) + log(5.62)) / 2 # 1.381534
(log(5.62) + log(11.22)) / 2 # 2.072015
(log(11.22) + log(22.39)) / 2 # 2.763156
(log(22.39) + log(44.67)) / 2 # 3.453958
(log(44.67) + log(89.13)) / 2 # 4.144699
(log(89.13) + log(177.83)) / 2 # 4.835462
(log(177.83) + log(354.81)) / 2 # 5.526205
(log(354.81) + log(707.95)) / 2 # 6.216978

# Rename the size columns so each one is log value now
size <- size %>%
  rename(
  "Log minimum size -1.410889" = "Minimum size 0.17 - 0.35mm",
    "Log minimum size -0.6961562" = "Minimum size 0.35 - 0.71mm",
    "Log minimum size 0.0005496977" = "Minimum size 0.71 - 1.41mm",
    "Log minimum size 0.6901633" = "Minimum size 1.41 - 2.82mm",
    "Log minimum size 1.381534" = "Minimum size 2.82 - 5.62mm",
    "Log minimum size 2.072015" = "Minimum size 5.62 - 11.22mm",
    "Log minimum size 2.763156" = "Minimum size 11.22 - 22.39mm",
    "Log minimum size 3.453958" = "Minimum size 22.39 - 44.67mm",
    "Log minimum size 4.144699"= "Minimum size 44.67 - 89.13mm",
    "Log minimum size 4.835462" = "Minimum size 89.13 - 177.83mm",
   "Log minimum size 5.526205"  = "Minimum size 177.83 - 354.81mm",
    "Log minimum size 6.216978" = "Minimum size 354.81 - 707.95mm",
    "Log maximum size -1.410889" = "Maximum size 0.17 - 0.35mm",
    "Log maximum size -0.6961562" = "Maximum size 0.35 - 0.71mm",
    "Log maximum size 0.0005496977" = "Maximum size 0.71 - 1.41mm",
    "Log maximum size 0.6901633" = "Maximum size 1.41 - 2.82mm",
    "Log maximum size 1.381534" = "Maximum size 2.82 - 5.62mm",
    "Log maximum size 2.072015" = "Maximum size 5.62 - 11.22mm",
    "Log maximum size 2.763156" = "Maximum size 11.22 - 22.39mm",
    "Log maximum size 3.453958" = "Maximum size 22.39 - 44.67mm",
    "Log maximum size 4.144699" = "Maximum size 44.67 - 89.13mm",
    "Log maximum size 4.835462" = "Maximum size 89.13 - 177.83mm",
    "Log maximum size 5.526205" = "Maximum size 177.83 - 354.81mm",
    "Log maximum size 6.216978" = "Maximum size 354.81 - 707.95mm",
  )


# Step 1: Identify columns
min_cols <- grep("^Log minimum size", names(size), value = TRUE)
max_cols <- grep("^Log maximum size", names(size), value = TRUE)

# Step 2: Extract weights (from column names)
min_weights <- as.numeric(str_extract(min_cols, "-?\\d+\\.?\\d*"))
max_weights <- as.numeric(str_extract(max_cols, "-?\\d+\\.?\\d*"))

# Step 3: Ensure columns are numeric
size[min_cols] <- lapply(size[min_cols], function(x) as.numeric(as.character(x)))
size[max_cols] <- lapply(size[max_cols], function(x) as.numeric(as.character(x)))

# Step 3: Calculate weighted sums
size <- size %>%
  rowwise() %>% # makes mutate() operate per row instead of per column
  mutate(
    min_weighted = sum(c_across(all_of(min_cols)) * min_weights, na.rm = TRUE), # Added na.rm = TRUE to skip any NA values during sum
    max_weighted = sum(c_across(all_of(max_cols)) * max_weights, na.rm = TRUE)
  ) %>%
  mutate(avg_weight = (min_weighted + max_weighted) / 2) %>% 
  ungroup()

# c_across() grabs all the columns listed in min_cols for the current row.
# all_of(min_cols) safely ensures only the columns you've selected are used (and errors if a name is missing — helpful for debugging).
unique(size$`Log minimum size -1.410889`)
unique(size$`Log minimum size 0.0005496977`)

rm(combined_names, max_cols, max_weights, min_cols, min_weights, new_names_part)
```

# Try with a range trait that has fuzzy coding
```{r}
# Lets go with lifelong fecundity

# Filter trait database for meta data and lifelong fecundity cols
fecund <- traits %>% dplyr::select(c(1:8,77:80))

# Calculate the logs
(log(10) + log(100)) / 2 # 3.453878
(log(100) + log(1000)) / 2 # 5.756463
(log(1000) + log(3000)) / 2 # 7.457061
(log(3000) + log(8000)) / 2 # 8.496782

# Rename the fecundity columns so each one is log value now
fecund <- fecund %>%
  rename(
  "Log lifelong fecundity 3.453878 number of eggs per female" = "Lifelong fecundity | < 100 | Number of eggs per female",
    "Log lifelong fecundity 5.756463 number of eggs per female" = "Lifelong fecundity | 100-1000 | Number of eggs per female",
    "Log lifelong fecundity 7.457061 number of eggs per female" = "Lifelong fecundity | 1000-3000 | Number of eggs per female",
    "Log lifelong fecundity 8.496782 number of eggs per female" = "Lifelong fecundity | > 3000 | Number of eggs per female",
  )

# Step 1: Identify columns
fecund_cols <- grep("^Log lifelong fecundity", names(fecund), value = TRUE)

# Step 2: Extract weights (from column names)
fecund_weights <- as.numeric(str_extract(fecund_cols, "-?\\d+\\.?\\d*"))

# Step 3: Ensure columns are numeric
fecund[fecund_cols] <- lapply(fecund[fecund_cols], function(x) as.numeric(as.character(x)))

# Step 3: Calculate weighted sums
fecund_test <- fecund %>%
  rowwise() %>% # makes mutate() operate per row instead of per column
  mutate(
    weighted_average = sum(c_across(all_of(fecund_cols)) * fecund_weights, na.rm = TRUE), # Added na.rm = TRUE to skip any NA values during sum
  ) %>%
  ungroup()
```

# Try with dispersal ability
```{r}
dispersal <- traits %>% dplyr::select(c(1:8, 196:200))

dispersal_weights <- c(1,2,3,4,5)

dispersal_cols <- grep("^Dispersal ability", names(dispersal), value = TRUE)

# Step 3: Ensure columns are numeric
dispersal[dispersal_cols] <- lapply(dispersal[dispersal_cols], function(x) as.numeric(as.character(x)))

# Step 3: Calculate weighted sums
dispersal_test <- dispersal %>%
  rowwise() %>% # makes mutate() operate per row instead of per column
  mutate(
    weighted_average = sum(c_across(all_of(dispersal_cols)) * dispersal_weights, na.rm = TRUE), # Added na.rm = TRUE to skip any NA values during sum
  ) %>%
  ungroup()

```

# Try with egg volume

```{r}
egg <- traits %>% dplyr::select(c(1:8,166:175))

# Calculate the midrange as they are already in log10 space
(-6.15 + -3.43) / 2 # -4.79

model_data_ab %>% filter(SSBST == "CC1_2012__Waite 1 3 14 Coleoptera") %>% View()
```


# SIZE - Merging Johns size trait to my model_data

```{r}
# Get taxon info for all insects
taxon_insects <- insects %>% dplyr::select(SS, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Parsed_name, Taxon, Best_guess_binomial, Effort_corrected_measurement, Merged_diversity)

# Get taxon info from Johns size dataset
john_size <- size %>% dplyr::select(Family, "Taxon Name", min_weighted, max_weighted, avg_weight)

# Get rid of missing size information, i.e. avg_weight == 0 or == NA
john_size <- john_size %>% filter(!is.na(avg_weight) & avg_weight != 0)

# Get rid of rows which aren't at family level, i.e. at superfamily or higher
john_size <- john_size %>% filter(!is.na(Family))

# merge the two
merged_size <- inner_join(taxon_insects, john_size, by = "Family")

# Which rows were excluded?
excluded <- anti_join(taxon_model, john_size, by = "Family")

rm(taxon_model, john_size, excluded)
```

# Test dataset

```{r}
test <- all_insects_size %>%
  group_by(SSBS) %>%
  filter(n_distinct(Order) > 2,
         n_distinct(Best_guess_binomial) >= 6,
         n_distinct(Best_guess_binomial) <= 8,
    mean(Merged_diversity != 0, na.rm = TRUE) > 0.4
         ) %>%
  ungroup()

#test_data <- test %>% filter(SSBS == "AD1_2009__Vergara 1  6")

#write.csv(test_data, file = "test_site_1.csv")

```

# Summary statistics 

```{r}
# Species-level data
n_species <- all_insects_size %>%
  filter(!is.na(Trait_Value)) %>%
  distinct(Best_guess_binomial) %>%
  nrow()

length(unique(all_insects_size$Best_guess_binomial))

# Genus-level data
n_genera <- all_insects_size %>%
  filter(!is.na(genus_avg_body_length)) %>%
  distinct(Genus) %>%
  nrow()

# Family-level data
n_families <- all_insects_size %>%
  filter(!is.na(family_avg_body_length)) %>%
  distinct(Family) %>%
  nrow()

# Show the results
cat("Species with size data:", n_species, "\n")
cat("Genera with size data:", n_genera, "\n")
cat("Families with size data:", n_families, "\n")


# Step 1: Add a column that identifies the source of body size
test <- all_insects_size %>%
  mutate(
    data_source = case_when(
      !is.na(Trait_Value) ~ "species",
      is.na(Trait_Value) & !is.na(genus_avg_body_length) ~ "genus",
      is.na(Trait_Value) & is.na(genus_avg_body_length) & !is.na(family_avg_body_length) ~ "family",
      TRUE ~ "missing"
    )
  )

# Step 2: Summarize by Family and Genus
group_summary <- test %>%
  group_by(Family, Genus) %>%
  summarise(
    n_species = n(),
    n_species_level = sum(data_source == "species"),
    n_genus_level = sum(data_source == "genus"),
    n_family_level = sum(data_source == "family"),
    n_missing = sum(data_source == "missing"),
    .groups = "drop"
  ) %>%
  mutate(
    percent_species_level = 100 * n_species_level / n_species,
    percent_genus_level = 100 * n_genus_level / n_species,
    percent_family_level = 100 * n_family_level / n_species
  )

# Step 3: Identify groups with low species-level data
low_species_data <- group_summary %>%
  filter(percent_species_level < 50) %>%  # threshold is flexible
  arrange(percent_species_level)

# Optional: View top 10 groups with lowest species-level data
head(low_species_data, 10)

```