---
title: "Untitled"
output: html_document
date: "2026-02-23"
---

# This markdown 23rd February 2026, is going to filter my data frame for all those rows that are currently matched to a family-level body length value BUT are resolved to genus or species level and so can therefore be improved.
# I will then check how much of that data can be matched to a number of trait databases to see how much can be improved.

# Read in the packages 
```{r}
library(here)
```

# Read in the data
```{r}
all_insects_size <- readRDS(here("data", "all_insects_size.RDS"))
```


# First lets subset my overall dataframe for rows currently matched to a family-level body length value BUT are resolved to genus or species level.

```{r}
# Which rows have only a family-level body length value, but ARE taxonomically resolved to genus or species.
genus_or_species_rows <- all_insects_size %>%
  filter(
    (!is.na(family_avg_body_length) | !is.na(Rainford_family_avg)),  # contains a family value from either column
    is.na(Trait_Value),                   # no species level value
    is.na(genus_avg_body_length),         # no genus level value
    (!is.na(Genus) & Genus != "") | (!is.na(Best_guess_binomial) & Best_guess_binomial != "") # genus or species present
  )

# Get a list of unique species
unique_species <- genus_or_species_rows %>%
  filter(!is.na(Best_guess_binomial), Best_guess_binomial != "") %>%
  distinct(Best_guess_binomial) %>%
  arrange(Best_guess_binomial)

# Get a list of unique genera
unique_genera <- genus_or_species_rows %>%
  filter(!is.na(Genus), Genus != "") %>%
  distinct(Genus) %>%
  arrange(Genus)
```

# Animal traits database

```{r}
# Load the database
animaltraits <- read_excel(here("data", "animaltraitsdatabase.xlsx"))

# Filter for insects
just_insects <- animaltraits %>% filter(class == "Insecta")

# Columns of interest
just_insects <- just_insects %>% dplyr::select(genus, species, `body mass`, `body mass - units`, `original body mass`, `original body mass - units`)

# Remove those rows without data
just_insects <- just_insects %>%
  filter(!if_all(c("body mass", "body mass - units", "original body mass", "original body mass - units"), is.na))

# Convert kg into mg
just_insects <- just_insects %>%
  rename(body_mass = `body mass`) %>% # deal with space between body and mass
  mutate(body_mass_mg = body_mass * 1e6)

# Calculate an average for duplicate species entries
just_insects <- just_insects %>%
  group_by(species) %>%
  summarise(
    avg_body_mass_mg = mean(body_mass_mg, na.rm = TRUE),
    .groups = "drop"
  )

# Recreate genus column
just_insects <- just_insects %>%
  mutate(genus = sub(" .*", "", species)) # subtract everything before the first space

# Which species are in animaltraits that are also species in my df without a size value?
just_insects[just_insects$species %in% unique_species$Best_guess_binomial, ] # 10 species match
```

# Do the same but work out if there are any genera which overlap between PREDICTS and Animal Traits

```{r}
# Which genera are in Animal Traits that are also genera in my df without a size value?
genera_overlap <- just_insects[just_insects$genus %in% unique_genera$Genus, ]

length(unique(genera_overlap$genus)) # 20 genera added from AnimalTraits

rm(animaltraits, genera_overlap, just_insects)
```


# Lets see the overlap between Global Ants and PREDICTS ant taxa with missing size data

```{r}
# Read in Global Ants trait data
globalants <- read.csv(here("data", "globalants.csv"))

# Filter for columns of interest
globalants <- globalants %>% dplyr::select(Genus, Species, Whole.body.length..mm.) %>% droplevels()

# Filter out those rows without a body length value
globalants <- globalants %>% filter(`Whole.body.length..mm.` != "Not in paper")

# Create best guess binomial
globalants <- globalants %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Convert variable to numeric for summarizing
globalants <- globalants %>%
  mutate(Whole.body.length..mm. = as.numeric(Whole.body.length..mm.))

# Calculate an average for duplicate species entries
globalants <- globalants %>%
  group_by(Best_guess_binomial) %>%
  summarise(
    avg_body_length_mm = mean(`Whole.body.length..mm.`, na.rm = TRUE),
    .groups = "drop"
  )

# Recreate genus column
globalants <- globalants %>%
  mutate(Genus = sub(" .*", "", Best_guess_binomial))  # subtract everything before the first space

# Which ant species are in globalants that are also species in my df without a size value?
ant_overlap <- globalants[globalants$Best_guess_binomial %in% unique_species$Best_guess_binomial, ] # 2 species :( 
```

# Same again but for genera

```{r}
genera_overlap <- globalants[globalants$Genus %in% unique_genera$Genus, ] # 12 genera we can add ant data for

rm(globalants, genera_overlap)
```


# Bee functional Trait Database

```{r}
# Load the data
bee_traits <- read.csv(here("data", "Sup Table 5 Trait Data Template.csv")) %>% strings2factors()

# Select the cols of interest
bee_traits <- bee_traits %>% dplyr::select(acceptedNameUsage, measurementValue, measurementUnit)

bees_updated <- bee_traits %>%
  mutate(
    genus = word(acceptedNameUsage, 1),
    species = word(acceptedNameUsage, 2),
    Best_guess_binomial = paste(genus, species)
  )

# Those species in the database with less than two words in their acceptednameusage
bee_traits %>%
  filter(str_count(acceptedNameUsage, "\\S+") < 2)

# Species overlap
bee_overlap <- unique_species[unique_species$Best_guess_binomial %in% bees_updated$Best_guess_binomial, ] %>% droplevels() # 85 species

length(unique(bee_overlap$Best_guess_binomial)) # 85 species matched 

# Which bees have only a family-level body length value, but ARE taxonomically resolved to genus or species.
bee_rows <- all_insects_size %>%
  filter(Family %in% c("Apidae", "Halictidae", "Megachilidae", "Andrenidae", "Colletidae", "Melittidae", "Stenotritidae")) %>% 
  filter(!is.na(family_avg_body_length) | !is.na(Rainford_family_avg),  # contains a family value from either column
    is.na(Trait_Value),                   # no species level value
    is.na(genus_avg_body_length),         # no genus level value
    (!is.na(Genus) & Genus != "") | (!is.na(Best_guess_binomial) & Best_guess_binomial != "")) # genus or species present

length(unique(bee_rows$Best_guess_binomial)) # 368 species currently matched at the family level but are taxonomically resolved to species

85/368 # 23% increase!
```

# Bee Functional Trait Database genera

```{r}
# Which genera are in the Bee Functional Trait Database that are also genera in my df without a size value?
bee_genera_overlap <- bees_updated[bees_updated$genus %in% unique_genera$Genus, ]

length(unique(bee_genera_overlap$genus)) # 60 genera matched!
length(unique(bee_rows$Genus)) # 125 genera currently matched at the family level

60/125 # 48% increase!

rm(bee_genera_overlap, bee_overlap, bee_rows, bee_traits, bees_updated)
```


# PREDICTS butterfly trait data from Cleary

```{r}
# read in the data
butterfly_traits <- read.csv(here("data", "butterfly_traits_full.csv"))

# Manipulate the columns so we can create best_guess_binomial and extract the body size value for each species
butterfly_traits <- butterfly_traits %>%
  mutate(
    Best_guess_binomial = sub("^((\\S+)\\s+(\\S+)).*", "\\1", Species.Family.Subfamily.Tribe.Genus.AuthorSp.Range.Group.Geo.Gen.MeanWing.SexDi.Conspicuousness.Migratory),
    meanwingspan = sapply(strsplit(X, "\\s+"), function(x) as.numeric(x[7]))
  )

leps_overlap <- unique_species[unique_species$Best_guess_binomial %in% butterfly_traits$Best_guess_binomial, ] 

length(unique(leps_overlap$Best_guess_binomial)) # 357 species matched!
```

# Match lep genera

```{r}
# Recreate cols for matching
butterfly_traits <- butterfly_traits %>%
  separate(Best_guess_binomial, into = c("genus", "species"), sep = " ") # seperate BGB into genus and species

butterfly_genera_overlap <- butterfly_traits[butterfly_traits$genus %in% unique_genera$Genus, ] 
length(unique(butterfly_genera_overlap$genus)) # 165 genera matched!

rm(butterfly_genera_overlap, butterfly_traits, leps_overlap)
```


# Data from: Refining estimates of dry body weight from linear measurements in adult moths and butterflies (Lepidoptera)

```{r}
# read in the data
enrique_traits <- read.csv(here("data", "LEPSIZE2024.csv"))

summary(enrique_traits)

# Select cols of interest
enrique_traits <- enrique_traits %>% dplyr::select(FAMILY, SPECIES, DBw)

# Clean and tidy the columns
enrique_traits <- enrique_traits %>%
  mutate(
    # Remove quotes and extract words
    clean_species = str_remove_all(SPECIES, '"'),
    genus = word(clean_species, 1),
    species = word(clean_species, 2),
    Best_guess_binomial = paste(genus, species)
  )

species_overlap <- unique_species[unique_species$Best_guess_binomial %in% enrique_traits$Best_guess_binomial, ] 
length(unique(species_overlap$Best_guess_binomial)) # 112 species matched
```

# Match lep genera

```{r}
# Which genera are in the trait datatbase that are also genera in my df without a size value?
butterfly_genera_overlap <- enrique_traits[enrique_traits$genus %in% unique_genera$Genus, ] # 250 genera overlap!
length(unique(butterfly_genera_overlap$genus)) # 250 genera matched!

rm(butterfly_genera_overlap, enrique_traits, species_overlap)
```

# Morphological trait database of saproxylic beetles

```{r}
# Read in the data
beetle_traits <- read.csv(here("data", "Hagge_etal_saproxylic_beetle_morphological_trait_database_20210426.csv"))

# Select cols of interest
beetle_traits <- beetle_traits %>% dplyr::select(family, species, body_length)

# Recreate cols for matching
beetle_traits <- beetle_traits %>%
  separate(species, into = c("genus", "species"), sep = "_") %>% # seperate species into genus and species
  mutate(best_guess_binomial = paste(genus, species, sep = " ")) # create best_guess_binomial

# No species overlap :( - unsurprising given its saproxylic beetles
beetle_overlap <- unique_species[unique_species$Best_guess_binomial %in% beetle_traits$Best_guess_binomial, ] 

```

# Lets see if any of the genera match between PREDICTS beetles and the trait database

```{r}
# Which genera are in the trait datatbase that are also genera in my df without a size value?
beetle_genera_overlap <- beetle_traits[beetle_traits$genus %in% unique_genera$Genus, ] 
length(unique(beetle_genera_overlap$genus)) # 36 genera overlap!

rm(beetle_genera_overlap, beetle_overlap, beetle_traits)
```

# Data from: A summary of eight traits of Coleoptera, Hemiptera, Orthoptera and Araneae, occurring in grasslands in Germany

```{r}
# Read in the trait data
mixed_traits <- read_delim(here("data", "ArthropodSpeciesTraits.txt"))

# Species overlap
mixed_overlap <- unique_species[unique_species$Best_guess_binomial %in% mixed_traits$SpeciesID, ] # 13 species

```

# Lets see if any of the genera match 

```{r}
# Recreate cols for matching
mixed_traits <- mixed_traits %>%
  separate(SpeciesID, into = c("genus", "species"), sep = " ") %>% # seperate species into genus and species
  mutate(best_guess_binomial = paste(genus, species, sep = " ")) # create best_guess_binomial

# Which genera are in the trait datatbase that are also genera in my df without a size value?
mixed_genera_overlap <- mixed_traits[mixed_traits$genus %in% unique_genera$Genus, ] # 26 genera overlap!

rm(mixed_genera_overlap, mixed_overlap, mixed_traits)
```


# Data from AnthropInsect

```{r}
# Read in the trait data
anthroptraits <- read_xlsx(here("data", "AnthropInsect_V1.xlsx"))

summary(anthroptraits)

# Species size
species_size <- anthroptraits %>% filter(!is.na(Body_size_adult_mean)) %>% dplyr::select(Order, Family, Species, Body_size_adult_mean) 

# Genera size
genera_size <- anthroptraits %>% filter(!is.na(Body_size_adult_mean)) %>% distinct(Genus)

# Assume your data frame is called df
summary_table <- data.frame(
  column = names(anthroptraits),
  num_NA = sapply(anthroptraits, function(x) sum(is.na(x))),
  num_blank = sapply(anthroptraits, function(x) sum(x == "" & !is.na(x)))
)

print(summary_table)

# Species overlap
species_overlap <- unique_species[unique_species$Best_guess_binomial %in% species_size$Species, ] # 81 species

# Genera overlap
genera_overlap <- unique_genera[unique_genera$Genus %in% genera_size$Genus, ] # 130 genera

```


# Check matching for Diptera from PREDICTS trait data (BeeDipteraTraits_P1_WoodFlynn)

```{r}
# Read in the data
diptera_traits <- read.csv(here("data", "WoodDiptera.csv"))

summary(diptera_traits)

# Species size
species_size <- diptera_traits %>% filter(!is.na(Adult.body.length.average)) %>% mutate(Best_guess_binomial = paste(Genus, Species)) %>% dplyr::select(Best_guess_binomial, Adult.body.length.average)

# Genera size
genera_size <- diptera_traits %>% filter(!is.na(Adult.body.length.average)) %>% distinct(Genus)

# Species overlap
species_overlap <- unique_species[unique_species$Best_guess_binomial %in% species_size$Best_guess_binomial, ] # 621 species

# Genera overlap
genera_overlap <- unique_genera[unique_genera$Genus %in% genera_size$Genus, ] # 274 genera

rm(diptera_traits, species_overlap, species_size, genera_overlap, genera_size)
```

# Check matching for MiddletonWelling

```{r}
# Read in the data
middletonwellingtraits <- read_xlsx(here("data", "MiddletonWelling-2014-07-08.xlsx")) %>% strings2factors()

summary(middletonwellingtraits)

# Species size
species_size <- middletonwellingtraits %>% filter(!is.na(`Wingspan male`)) %>% mutate(Best_guess_binomial = paste(Genus, Species)) %>% dplyr::select(Best_guess_binomial, `Wingspan male`)

# Genera size
genera_size <- middletonwellingtraits %>% filter(!is.na(`Wingspan male`)) %>% distinct(Genus)

# Species overlap
species_overlap <- unique_species[unique_species$Best_guess_binomial %in% species_size$Best_guess_binomial, ] # 392 species

# Genera overlap
genera_overlap <- unique_genera[unique_genera$Genus %in% genera_size$Genus, ] # 304 genera

rm(middletonwellingtraits, species_size, genera_size, species_overlap, genera_overlap)
```

# Check matching for Leptraits
```{r}
# Read in the data
leptraits <- read.csv(here("data", "consensus.csv")) 

summary(leptraits)

# WS_L = The lowest measurement reported for unspecified sex wingspan in centimeters.
# WS_U = The highest (or sole) measurement reported for unspecified sex wingspan in centimeters.

leptraits <- leptraits %>%
  mutate(WS_AVG = ifelse(
    is.na(WS_L) & is.na(WS_U), NA, # if both are an NA = NA
    ifelse(is.na(WS_L), WS_U, # if WS_L is an NA then give me WS_U
           ifelse(is.na(WS_U), WS_L, # if WS_U is an NA then give me WS_L
                  (WS_L + WS_U)/2))) # else give me the average beteween WS_L and WS_U
  )

# Species size
species_size <- leptraits %>% filter(!is.na(WS_AVG)) %>% dplyr::select(verbatimSpecies, WS_AVG)

# Genera size
genera_size <- leptraits %>% filter(!is.na(WS_AVG)) %>% distinct(Genus)

# Species overlap
species_overlap <- unique_species[unique_species$Best_guess_binomial %in% species_size$verbatimSpecies, ] # 829 species

# Genera overlap
genera_overlap <- unique_genera[unique_genera$Genus %in% genera_size$Genus, ] # 365 genera

rm(leptraits, species_size, genera_size, species_overlap, genera_overlap)
```

