---
title: "Untitled"
output: html_document
date: "2025-11-27"
---

## R Markdown




# Clear the work space

```{r}
rm(list=ls())
```

# Read in the packages 
```{r}
library(lme4)
library(readxl)
library(tidyverse)
library(stringr)
library(car)
library(qs)
library(lmerTest)
library(merTools) 
library(performance)
library(DHARMa)
library(here)
library(blme)
library(MuMIn)
library(parallel)
library(cowplot)
library(ggeffects)
library(patchwork)
library(tictoc)
library(glmmTMB)
```


# Read in size data
```{r}
ModelDataSizeAbundance <- readRDS("../C4 data/ModelDataSizeAbundance.rds")
ModelDataSizeOccurrence <- readRDS("../C4 data/ModelDataSizeOccurrence.rds")
```


# Modelling AbundanceWhenPresent ~ LUFinal * mass * order

```{r}
# Filter for big five
BigFiveModeDataSizeAbundance <- ModelDataSizeAbundance %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera")) %>% droplevels()

# Maximal model
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * log10_mass * Order + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = BigFiveModeDataSizeAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Check the model
model_plot(ab_m1)
summary(ab_m1)
anova(ab_m1)
Anova(ab_m1)

# Mass range based on model data
mass_range <- range(BigFiveModeDataSizeAbundance$log10_mass, na.rm = TRUE)

# Create a grid of values
NewDataAbundanceMassOrder <- expand.grid(
  log10_mass  = seq(from = mass_range[1],
                    to   = mass_range[2],
                    length.out = 100),
  LandUseFinal = levels(BigFiveModeDataSizeAbundance$LandUseFinal),
  Order        = levels(BigFiveModeDataSizeAbundance$Order)
)

# Get predictions without random effects, with SEs
AbundancePredictions <- predict(
  ab_m1,
  newdata = NewDataAbundanceMassOrder,
  re.form = NA,      # fixed-effects only (population-level)
  se.fit  = TRUE,    # get SE on log10 scale
)

# Predictions on log10 scale
NewDataAbundanceMassOrder$pred_log10 <- AbundancePredictions$fit
NewDataAbundanceMassOrder$se_log10   <- AbundancePredictions$se.fit

# 95% CI on log10 scale
value <- qnorm(0.975)
NewDataAbundanceMassOrder$log10_lwr <- NewDataAbundanceMassOrder$pred_log10 -
                                  value * NewDataAbundanceMassOrder$se_log10
NewDataAbundanceMassOrder$log10_upr <- NewDataAbundanceMassOrder$pred_log10 +
                                  value * NewDataAbundanceMassOrder$se_log10

# Back-transform all of them
NewDataAbundanceMassOrder$pred_backtransformed <- 10^(NewDataAbundanceMassOrder$pred_log10)
NewDataAbundanceMassOrder$lwr_backtransformed <- 10^(NewDataAbundanceMassOrder$log10_lwr)
NewDataAbundanceMassOrder$upr_backtransformed  <- 10^(NewDataAbundanceMassOrder$log10_upr)


# Plot
PlotBigFiveAbundanceMass <- ggplot(NewDataAbundanceMassOrder, aes(x = 10^log10_mass, y = pred_backtransformed, color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lwr_backtransformed, ymax = upr_backtransformed),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  scale_y_continuous(labels = scales::label_number())   +
  scale_x_log10(labels = scales::label_number()) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"),
  #legend.position = c(1, 0.4),  # x = horizontal, y = vertical
  #legend.justification = c(1, 1)
        ) +
  labs(
    x = "Body mass (mg)",
    y = "Abundance of present species (± 95 % CI)",
    color = "Land use",
    fill  = "Land use",
    title = ""
  )


#ggsave(filename = "../figures/c4_updated/AbundanceWhenPresentBigFiveBodyMass.png",   plot = PlotBigFiveAbundanceMass, width = 10, height = 6, dpi = 300)
```

# Modelling Occurrence ~ LUFinal * mass * order
```{r}
# Filter for the big five
BigFiveModelBodyMassOccurrence <- ModelDataSizeOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))

BigFiveModelBodyMassOccurrence <- droplevels(BigFiveModelBodyMassOccurrence) # Avoids trouble later on

# Scale the predictors
BigFiveModelBodyMassOccurrence$ScaledLog10Mass <- scale(BigFiveModelBodyMassOccurrence$log10_mass, center = TRUE, scale = TRUE)

tictoc::tic("Fitting glmmTMB")

m1 <- glmmTMB::glmmTMB(PresentOrAbsent ~
                         LandUseFinal * log10_mass * Order +
                         (1 | SS) + (1 | SpeciesFinal), 
                       data = BigFiveModelBodyMassOccurrence,
                       family = binomial(link = "logit"),
                       control = glmmTMBControl(
                         optCtrl = list(iter.max = 1e3, eval.max = 1e3)
                         )
)

# Check the model
summary(m1)
Anova(m1)

# Calculate the mass range for predictions
mass_range <- range(BigFiveModelBodyMassOccurrence$log10_mass, na.rm = TRUE)

# Create a grid of values
NewDataOccurrenceMassOrder <- expand.grid(
  log10_mass  = seq(from = mass_range[1],
                    to   = mass_range[2],
                    length.out = 100),
  LandUseFinal = levels(BigFiveModelBodyMassOccurrence$LandUseFinal),
  Order        = levels(BigFiveModelBodyMassOccurrence$Order)
)

# Get predictions on the link scale 
OccurrencePredictions <- predict(
  m1,
  newdata = NewDataOccurrenceMassOrder,
  re.form = NA,      # fixed-effects only (population-level)
  se.fit  = TRUE,    # get SE on log10 scale
  type = "link"
)

# Add predictions into data frame
NewDataOccurrenceMassOrder <- NewDataOccurrenceMassOrder %>%
  mutate(
    OccurrenceFit   = OccurrencePredictions$fit,
    OccurrenceSE    = OccurrencePredictions$se.fit,
    OccurrenceLwr = OccurrenceFit - 1.96 * OccurrenceSE,
    OccurrenceUpr = OccurrenceFit + 1.96 * OccurrenceSE,
    # Back-transform to probability scale
    OccurrenceProb   = plogis(OccurrenceFit), # predicted probability - plogis() is the inverse of the logit function.
    # It takes values on the logit (link) scale (from −∞ to +∞) and transforms them into probabilities between 0 and 1.
    LowerProb = plogis(OccurrenceLwr), # 95% CI lower bound
    UpperProb = plogis(OccurrenceUpr) # 95% CI upper bound
  )

# Plot
PlotBigFiveOccurrenceMass <- ggplot(NewDataOccurrenceMassOrder, aes(x = 10^log10_mass, y = OccurrenceProb, color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = LowerProb, ymax = UpperProb),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  scale_x_log10(labels = scales::label_number()) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold")) +
  labs(
    x = "Body mass (mg)",
    y = "Probability of occurrence (± 95 % CI)",
    color = "Land use",
    fill  = "Land use",
    title = ""
  )

#ggsave(filename = "../figures/c4_updated/BodyMassBigFiveOccurrence.png",   plot = PlotBigFiveOccurrenceMass, width = 10, height = 6, dpi = 300)
```

# If my desired outcome is the predicted abundance when accounting for probability of occurrence, it makes more sense to backtransform each prediction to its natural scale first, such that occurrence_prob = plogis(fit_link) and abund_exp = exp(fit_abund). This would lead to occurrence_prob being the probability of occurrence and abund_exp being the predicted abundance. I would then multiply these two components such that pred_abundance = occ_prob * abund_exp and use the delta method to calculate the SEs. 

```{r}

# Backtransform and calculate predicted abundance with CIs
NewDataCombined <- NewDataOccurrenceMassOrder %>%
  bind_cols(NewDataAbundanceMassOrder %>% 
              dplyr::select(
                pred_backtransformed = pred_backtransformed,
                     se_log10  = se_log10)) %>%
  mutate(
    # Probability of occurrence * predicted abundance when present = predicted abundance 
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * pred_backtransformed,  # c = a * b
    # Delta method
    occurrence_var  = (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2), # variance of plogis(fit_link)
    abundance_var = (log(10)^2) * (pred_backtransformed^2) * (se_log10^2), # variance of 10^(fit_abund)

    combined_var =
      (pred_backtransformed^2) * occurrence_var +
      (OccurrenceProb^2) * abundance_var,

    predictedcombinedvar = sqrt(combined_var),

    # 95% CIs on original (abundance) scale
    predictedcombinedlwr =
      PredictedAbundanceAccountingForOccurrenceProbability -
      1.96 * predictedcombinedvar,
    predictedcombinedupr =
      PredictedAbundanceAccountingForOccurrenceProbability +
      1.96 * predictedcombinedvar
  )
    
# Plot
FinalPlotAbundanceOccurrenceMassCombined <- ggplot(NewDataCombined,
       aes(x = 10^log10_mass, y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = predictedcombinedlwr, ymax = predictedcombinedupr),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  scale_x_log10(labels = scales::label_number()) +
  labs(x = "Body mass (mg)",
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal()

#ggsave(filename = "../figures/c4_updated/BodyMassAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceMassCombined, width = 6, height = 6, dpi = 300)
```

# Combine the predictions and create the plot
```{r}
# abundance df
abundance <- NewDataAbundanceMassOrder %>%           
  transmute(
    log10_mass, LandUseFinal, Order,
    pred = pred_backtransformed,
    lwr  = lwr_backtransformed,
    upr  = upr_backtransformed,
    model = "Abundance"
  )

# occurrence df
occurrence <- NewDataOccurrenceMassOrder %>%              
  transmute(
    log10_mass, LandUseFinal, Order,
    pred = OccurrenceProb,
    lwr  = LowerProb,
    upr  = UpperProb,
    model = "Occurrence"
  )

# combined df
combined <- NewDataCombined %>%    
  transmute(
    log10_mass, LandUseFinal, Order,
    pred = PredictedAbundanceAccountingForOccurrenceProbability,
    lwr  = predictedcombinedlwr,
    upr  = predictedcombinedupr,
    model = "Combined"
  )

allpredictions <- bind_rows(abundance, occurrence, combined)

# Relevel so occurrence appears second in the plots
allpredictions <- allpredictions %>% 
  dplyr::mutate(
    model = factor(model,
                   levels = c("Abundance", "Occurrence", "Combined")))

base_plot <- function(df, ylab) {
  ggplot(df, aes(x = 10^log10_mass, y = pred, color = LandUseFinal, fill = LandUseFinal)) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
    facet_grid(Order ~ ., scales = "free_y") +   # rows = orders
    scale_x_log10(labels = scales::label_number()) +
    labs(
      x = "Body mass (mg)",
      y = ylab,
      color = "Land use",
      fill  = "Land use"
    ) +
    theme_minimal() +
    theme(
      #strip.text.y.right = element_blank(),  # remove RHS facet labels if you like
      legend.position = "bottom" # legend position for each plot (collected later)
    )
}


# a) remove facet labels for the first two columns
p_abund <- base_plot(filter(allpredictions, model == "Abundance"),
                     "Abundance of present species (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank())

p_occ <- base_plot(filter(allpredictions, model == "Occurrence"),
                   "Probability of occurrence (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank())

# b) keep facet labels for the third column
p_comb <- base_plot(filter(allpredictions, model == "Combined"),
                    "Predicted abundance = a × b (± 95 % CI)") +
  theme(
    strip.text.y.right = element_text(size = 12, face = "bold")
  )

# Combine — collect legends
FinalPlot <- (p_abund | p_occ | p_comb) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'a') &
  theme(
    legend.position = "bottom",
    plot.tag = element_text(size = 14, face = "bold"),
    plot.tag.position = c(0, 1)  # top-right corner
  )

FinalPlot

#saveRDS(allpredictions, "../C4 data/BodyMassAllPredictions.rds")
#ggsave(filename = "../figures/C4_updated/BodyMassOrders.png",  plot = FinalPlot, width = 14, height = 8, dpi = 300)

```


# Read in egg data
```{r}
ModelDataEggAbundance <- readRDS("../C4 data/ModelDataEggAbundance.rds")
ModelDataEggOccurrence <- readRDS("../C4 data/ModelDataEggOccurrence.rds")
```

# Modelling AbundanceWhenPresent ~ LUFinal * egg size * order

```{r}
# Filter for big five
BigFiveModeDataEggAbundance <- ModelDataEggAbundance %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera")) %>% droplevels()

# Maximal model
BigFiveAb_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * LogAvgEggVolume_mm3 * Order + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = BigFiveModeDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Check the model
model_plot(BigFiveAb_m1)
summary(BigFiveAb_m1)
anova(BigFiveAb_m1)
Anova(BigFiveAb_m1)


# Calculate the egg volume range for predictions
egg_range <- range(BigFiveModeDataEggAbundance$LogAvgEggVolume_mm3, na.rm = TRUE)

# Create a grid of values
NewDataAbundanceEggOrder <- expand.grid(
  LogAvgEggVolume_mm3  = seq(from = egg_range[1],
                    to   = egg_range[2],
                    length.out = 100),
  LandUseFinal = levels(BigFiveModeDataEggAbundance$LandUseFinal),
  Order        = levels(BigFiveModeDataEggAbundance$Order)
)

# Get predictions without random effects, with SEs
AbundancePredictions <- predict(
  BigFiveAb_m1,
  newdata = NewDataAbundanceEggOrder,
  re.form = NA,      # fixed-effects only (population-level)
  se.fit  = TRUE,    # get SE
)

# Predictions 
NewDataAbundanceEggOrder$pred_log <- AbundancePredictions$fit
NewDataAbundanceEggOrder$se_log   <- AbundancePredictions$se.fit

# 95% CI 
value <- qnorm(0.975)
NewDataAbundanceEggOrder$log_lwr <- NewDataAbundanceEggOrder$pred_log -
                                  value * NewDataAbundanceEggOrder$se_log
NewDataAbundanceEggOrder$log_upr <- NewDataAbundanceEggOrder$pred_log +
                                  value * NewDataAbundanceEggOrder$se_log

# Back-transform all of them
NewDataAbundanceEggOrder$pred_backtransformed <- exp(NewDataAbundanceEggOrder$pred_log)
NewDataAbundanceEggOrder$lwr_backtransformed <- exp(NewDataAbundanceEggOrder$log_lwr)
NewDataAbundanceEggOrder$upr_backtransformed  <- exp(NewDataAbundanceEggOrder$log_upr)

# Plot
PlotBigFiveAbundanceEggSize <- ggplot(NewDataAbundanceEggOrder, aes(x = exp(LogAvgEggVolume_mm3), y = pred_backtransformed, color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = lwr_backtransformed, ymax = upr_backtransformed),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))  +
  labs(x = expression("Egg volume (mm"^3*")"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  scale_x_log10(
    labels = scales::label_number()) +
  scale_y_continuous(labels = scales::label_number()) 



  
#ggsave(filename = "../figures/AbundanceWhenPresentBigFiveEggVolume.png",   plot = PlotBigFiveAbundanceEggSize, width = 10, height = 6, dpi = 300)
```

# Modelling Occurrence ~ LUFinal * egg size * order

```{r}
# Convert to factors
ModelDataEggOccurrence$Order <- as.factor(ModelDataEggOccurrence$Order)
ModelDataEggOccurrence$Family <- as.factor(ModelDataEggOccurrence$Family)
ModelDataEggOccurrence$SpeciesFinal <- as.factor(ModelDataEggOccurrence$SpeciesFinal)
ModelDataEggOccurrence$SS <- as.factor(ModelDataEggOccurrence$SS)
ModelDataEggOccurrence$SSB <- as.factor(ModelDataEggOccurrence$SSB)

# Filter for the big five
BigFiveModelDataEggOccurrence <- ModelDataEggOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera")) %>% droplevels()

m1 <- glmmTMB::glmmTMB(PresentOrAbsent ~
                         LandUseFinal * LogAvgEggVolume_mm3 * Order +
                         (1 | SS) + (1 | SpeciesFinal), 
                       data = BigFiveModelDataEggOccurrence,
                       family = binomial(link = "logit"),
                       control = glmmTMBControl(
                         optCtrl = list(iter.max = 1e3, eval.max = 1e3)
                         )
)

# Check the model
summary(m1)
Anova(m1)

# Calculate the range for predictions
egg_range <- range(BigFiveModelDataEggOccurrence$LogAvgEggVolume_mm3)

# Create a grid of values
NewDataOccurrenceEggOrder <- expand.grid(
LogAvgEggVolume_mm3  = seq(from = egg_range[1],
                    to   = egg_range[2],
                    length.out = 100),
  LandUseFinal = levels(BigFiveModelDataEggOccurrence$LandUseFinal),
  Order        = levels(BigFiveModelDataEggOccurrence$Order)
)

# Get predictions on the link scale 
OccurrencePredictions <- predict(
  m1,
  newdata = NewDataOccurrenceEggOrder,
  re.form = NA,      # fixed-effects only (population-level)
  se.fit  = TRUE,    # get SE 
  type = "link"
)


# Add predictions into data frame
NewDataOccurrenceEggOrder <- NewDataOccurrenceEggOrder %>%
  mutate(
    OccurrenceFit   = OccurrencePredictions$fit,
    OccurrenceSE    = OccurrencePredictions$se.fit,
    OccurrenceLwr = OccurrenceFit - 1.96 * OccurrenceSE,
    OccurrenceUpr = OccurrenceFit + 1.96 * OccurrenceSE,
    # Back-transform to probability scale
    OccurrenceProb   = plogis(OccurrenceFit), # predicted probability - plogis() is the inverse of the logit function.
    # It takes values on the logit (link) scale (from −∞ to +∞) and transforms them into probabilities between 0 and 1.
    LowerProb = plogis(OccurrenceLwr), # 95% CI lower bound
    UpperProb = plogis(OccurrenceUpr) # 95% CI upper bound
  )

# Plot
PlotBigFiveOccurrenceEggVolume <- ggplot(NewDataOccurrenceEggOrder, aes(x = exp(LogAvgEggVolume_mm3), y = OccurrenceProb, color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = LowerProb, ymax = UpperProb),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold")) +
  scale_x_log10(
    labels = scales::label_number()
  ) +
  labs(
  x = expression("Egg volume (mm"^3*")"),
    y = "Probability of occurrence (± 95 % CI)",
    color = "Land use",
    fill  = "Land use",
    title = ""
  ) 

#ggsave(filename = "../figures/C4_updated/EggVolumeBigFiveOccurrence.png",   plot = PlotBigFiveOccurrenceEggVolume, width = 10, height = 6, dpi = 300)
```

# # Create the combined plot for egg size

```{r}
# Backtransform and calculate predicted abundance with CIs
NewDataCombined <- NewDataOccurrenceEggOrder %>%
  bind_cols(NewDataAbundanceEggOrder %>% 
              dplyr::select(
                pred_backtransformed = pred_backtransformed,
                     se_log  = se_log)) %>%
  mutate(
    # Probability of occurrence * predicted abundance when present = predicted abundance 
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * pred_backtransformed,  # c = a * b
    # Delta method
    occurrence_var  = (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2), # variance of plogis(fit_link)
    abundance_var = (log(10)^2) * (pred_backtransformed^2) * (se_log^2), # variance of 10^(fit_abund)

    combined_var =
      (pred_backtransformed^2) * occurrence_var +
      (OccurrenceProb^2) * abundance_var,

    predictedcombinedvar = sqrt(combined_var),

    # 95% CIs on original (abundance) scale
    predictedcombinedlwr =
      PredictedAbundanceAccountingForOccurrenceProbability -
      1.96 * predictedcombinedvar,
    predictedcombinedupr =
      PredictedAbundanceAccountingForOccurrenceProbability +
      1.96 * predictedcombinedvar
  )
    
# Plot
FinalPlotAbundanceOccurrenceEggSizeCombined <- ggplot(NewDataCombined,
       aes(x = exp(LogAvgEggVolume_mm3), y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = predictedcombinedlwr, ymax = predictedcombinedupr),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  scale_x_log10(labels = scales::label_number()) +
  labs(x = expression("Egg volume (mm"^3*")"),
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal()

FinalPlotAbundanceOccurrenceEggSizeCombined


#ggsave(filename = "../figures/c4_updated/BodyMassAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceMassCombined, width = 6, height = 6, dpi = 300)
```

# Combine the predictions and create the plot
```{r}
# abundance df
abundance <- NewDataAbundanceEggOrder %>%           
  transmute(
    LogAvgEggVolume_mm3, LandUseFinal, Order,
    pred = pred_backtransformed,
    lwr  = lwr_backtransformed,
    upr  = upr_backtransformed,
    model = "Abundance"
  )

# occurrence df
occurrence <- NewDataOccurrenceEggOrder %>%              
  transmute(
    LogAvgEggVolume_mm3, LandUseFinal, Order,
    pred = OccurrenceProb,
    lwr  = LowerProb,
    upr  = UpperProb,
    model = "Occurrence"
  )

# combined df
combined <- NewDataCombined %>%    
  transmute(
    LogAvgEggVolume_mm3, LandUseFinal, Order,
    pred = PredictedAbundanceAccountingForOccurrenceProbability,
    lwr  = predictedcombinedlwr,
    upr  = predictedcombinedupr,
    model = "Combined"
  )

allpredictions <- bind_rows(abundance, occurrence, combined)

# Relevel so occurrence appears second in the plots
allpredictions <- allpredictions %>% 
  dplyr::mutate(
    model = factor(model,
                   levels = c("Abundance", "Occurrence", "Combined")))

base_plot <- function(df, ylab) {
  ggplot(df, aes(x = exp(LogAvgEggVolume_mm3), y = pred, color = LandUseFinal, fill = LandUseFinal)) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
    facet_grid(Order ~ ., scales = "free_y") +   # rows = orders
    scale_x_log10(labels = scales::label_number()) +
    labs(
      x = expression("Egg volume (mm"^3*")"),
      y = ylab,
      color = "Land use",
      fill  = "Land use"
    ) +
    theme_minimal() +
    theme(
      #strip.text.y.right = element_blank(),  # remove RHS facet labels if you like
      legend.position = "bottom" # legend position for each plot (collected later)
    )
}


# a) remove facet labels for the first two columns
p_abund <- base_plot(filter(allpredictions, model == "Abundance"),
                     "Abundance of present species (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank())

p_occ <- base_plot(filter(allpredictions, model == "Occurrence"),
                   "Probability of occurrence (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank())

# b) keep facet labels for the third column
p_comb <- base_plot(filter(allpredictions, model == "Combined"),
                    "Predicted abundance = a × b (± 95 % CI)") +
  theme(
    strip.text.y.right = element_text(size = 12, face = "bold")
  )

# Combine — collect legends
FinalPlot <- (p_abund | p_occ | p_comb) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'a') &
  theme(
    legend.position = "bottom",
    plot.tag = element_text(size = 14, face = "bold"),
    plot.tag.position = c(0, 1)  # top-right corner
  )

FinalPlot

#saveRDS(allpredictions, "../C4 data/EggVolumeAllPredictions.rds")
#ggsave(filename = "../figures/C4_updated/EggVolumeOrders.png",  plot = FinalPlot, width = 14, height = 8, dpi = 300)
```


# Read in adult trophic level data
```{r}
ModelDataAdultDietAbundance <- readRDS("../C4 data/ModelDataAdultDietAbundance.rds")
ModelDataAdultDietOccurrence <- readRDS("../C4 data/ModelDataAdultDietOccurrence.rds")
```

# Modelling LogAbundanceWhenPresent ~ LUFinal * ADULT trophic level * order

```{r}
# Filter for big five
BigFiveModelDataAdultDietAbundance <- ModelDataAdultDietAbundance %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera")) %>% droplevels()

# Maximal model
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevelAdult * Order + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = BigFiveModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Check the model
model_plot(ab_m1)
summary(ab_m1)
anova(ab_m1)
Anova(ab_m1)

# Calculate the adult trophic level range for predictions
AdultTrophicLevelRange <- range(BigFiveModelDataAdultDietAbundance$AvgTrophicLevelAdult, na.rm = TRUE)

# Create a grid of values
NewDataAbundanceAdultTrophicOrder <- expand.grid(
  AvgTrophicLevelAdult  = seq(from = AdultTrophicLevelRange[1],
                    to   = AdultTrophicLevelRange[2],
                    length.out = 100),
  LandUseFinal = levels(BigFiveModelDataAdultDietAbundance$LandUseFinal),
  Order        = levels(BigFiveModelDataAdultDietAbundance$Order)
)

# Get predictions
AbundancePredictions <- predict(ab_m1, newdata = NewDataAbundanceAdultTrophicOrder, re.form = NA, se.fit = TRUE)

# Predictions on log scale
NewDataAbundanceAdultTrophicOrder$pred_log <- AbundancePredictions$fit
NewDataAbundanceAdultTrophicOrder$se_log   <- AbundancePredictions$se.fit

# 95% CI on log10 scale
value <- qnorm(0.975)
NewDataAbundanceAdultTrophicOrder$log_lwr <- NewDataAbundanceAdultTrophicOrder$pred_log -
                                  value * NewDataAbundanceAdultTrophicOrder$se_log
NewDataAbundanceAdultTrophicOrder$log_upr <- NewDataAbundanceAdultTrophicOrder$pred_log +
                                  value * NewDataAbundanceAdultTrophicOrder$se_log

# Back-transform all of them
NewDataAbundanceAdultTrophicOrder$pred_backtransformed <- exp(NewDataAbundanceAdultTrophicOrder$pred_log)
NewDataAbundanceAdultTrophicOrder$lwr_backtransformed <- exp(NewDataAbundanceAdultTrophicOrder$log_lwr)
NewDataAbundanceAdultTrophicOrder$upr_backtransformed  <- exp(NewDataAbundanceAdultTrophicOrder$log_upr)

# Plot
PlotAdultTrophicLevelAbundance <- ggplot(NewDataAbundanceAdultTrophicOrder, aes(x = AvgTrophicLevelAdult, y = pred_backtransformed, 
                                                                                color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lwr_backtransformed, ymax = upr_backtransformed),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  scale_y_continuous(labels = scales::label_number())   +
  scale_x_log10(labels = scales::label_number()) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"),
  #legend.position = c(1, 0.4),  # x = horizontal, y = vertical
  #legend.justification = c(1, 1)
        ) +
  labs(
    x = expression("Adult trophic level"),
    y = "Abundance of present species (± 95 % CI)",
    color = "Land use",
    fill  = "Land use",
    title = ""
  ) +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC")) 

#ggsave(filename = "../figures/AbundanceWhenPresentBigFiveAdultTrophicLevel.png",   plot = PlotAdultTrophicLevelAbundance, width = 10, height = 6, dpi = 300)

#rm(ab_m1, PlotLarvalTrophicLevelAbundance, LarvalTrophicLevelPredictions)
```

# Modelling Occurrence ~ LUFinal * adult TL * order

```{r}
# Convert to factors
ModelDataAdultDietOccurrence$Order <- as.factor(ModelDataAdultDietOccurrence$Order)
ModelDataAdultDietOccurrence$Family <- as.factor(ModelDataAdultDietOccurrence$Family)
ModelDataAdultDietOccurrence$SpeciesFinal <- as.factor(ModelDataAdultDietOccurrence$SpeciesFinal)
ModelDataAdultDietOccurrence$SS <- as.factor(ModelDataAdultDietOccurrence$SS)
ModelDataAdultDietOccurrence$SSB <- as.factor(ModelDataAdultDietOccurrence$SSB)

# Filter for the big five
BigFiveModelDataAdultDietOccurrence <- ModelDataAdultDietOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera")) %>%  droplevels()

m1 <- glmmTMB::glmmTMB(PresentOrAbsent ~
                         LandUseFinal * AvgTrophicLevelAdult * Order +
                         (1 | SS) + (1 | SpeciesFinal), 
                       data = BigFiveModelDataAdultDietOccurrence,
                       family = binomial(link = "logit"),
                       control = glmmTMBControl(
                         optCtrl = list(iter.max = 1e3, eval.max = 1e3)
                         )
)

# Check the model
summary(m1)
Anova(m1)

# Create a grid of values
NewDataOccurrenceAdultTrophicOrder <- expand.grid(
  AvgTrophicLevelAdult  = seq(from = AdultTrophicLevelRange[1],
                    to   = AdultTrophicLevelRange[2],
                    length.out = 100),
  LandUseFinal = levels(BigFiveModelDataAdultDietOccurrence$LandUseFinal),
  Order        = levels(BigFiveModelDataAdultDietOccurrence$Order)
)

# Get predictions on the link scale 
OccurrencePredictions <- predict(
  m1,
  newdata = NewDataOccurrenceAdultTrophicOrder,
  re.form = NA,      # fixed-effects only (population-level)
  se.fit  = TRUE,    # get SE 
  type = "link"
)

# Add predictions into data frame
NewDataOccurrenceAdultTrophicOrder <- NewDataOccurrenceAdultTrophicOrder %>%
  mutate(
    OccurrenceFit   = OccurrencePredictions$fit,
    OccurrenceSE    = OccurrencePredictions$se.fit,
    OccurrenceLwr = OccurrenceFit - 1.96 * OccurrenceSE,
    OccurrenceUpr = OccurrenceFit + 1.96 * OccurrenceSE,
    # Back-transform to probability scale
    OccurrenceProb   = plogis(OccurrenceFit), # predicted probability - plogis() is the inverse of the logit function.
    # It takes values on the logit (link) scale (from −∞ to +∞) and transforms them into probabilities between 0 and 1.
    LowerProb = plogis(OccurrenceLwr), # 95% CI lower bound
    UpperProb = plogis(OccurrenceUpr) # 95% CI upper bound
  )


# Plot
PlotBigFiveOccurrenceAdultTrophic <- ggplot(NewDataOccurrenceAdultTrophicOrder, aes(x = AvgTrophicLevelAdult, y = OccurrenceProb, color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = LowerProb, ymax = UpperProb, fill = LandUseFinal),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold")) +
  labs(
  x = expression("Adult trophic level"),
    y = "Probability of occurrence (± 95 % CI)",
    color = "Land use",
    fill  = "Land use",
    title = ""
  ) +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC"))

#ggsave(filename = "../figures/C4_updated/AdultTLBigFiveOccurrence.png",   plot = PlotBigFiveOccurrenceAdultTrophic, width = 10, height = 6, dpi = 300)
```

# Create the combined plot for adult trophic level

```{r}
NewDataCombined <- NewDataOccurrenceAdultTrophicOrder %>%
  bind_cols(NewDataAbundanceAdultTrophicOrder %>% 
              dplyr::select(pred_backtransformed = pred_backtransformed,
                     se_log  = se_log)) %>%
 mutate(
    # Probability of occurrence * predicted abundance when present = predicted abundance 
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * pred_backtransformed,  # c = a * b
    # Delta method
    occurrence_var  = (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2), # variance of plogis(fit_link)
    abundance_var = (log(10)^2) * (pred_backtransformed^2) * (se_log^2), # variance of 10^(fit_abund)

    combined_var =
      (pred_backtransformed^2) * occurrence_var +
      (OccurrenceProb^2) * abundance_var,

    predictedcombinedvar = sqrt(combined_var),

    # 95% CIs on original (abundance) scale
    predictedcombinedlwr =
      PredictedAbundanceAccountingForOccurrenceProbability -
      1.96 * predictedcombinedvar,
    predictedcombinedupr =
      PredictedAbundanceAccountingForOccurrenceProbability +
      1.96 * predictedcombinedvar
  )
    

# Plot
FinalPlotAbundanceOccurrenceAdultTLCombined <- ggplot(NewDataCombined,
       aes(x = AvgTrophicLevelAdult, y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = predictedcombinedlwr, ymax = predictedcombinedupr),
              alpha = 0.2, color = NA) +
    facet_wrap(~ Order, scales = "free_y") +
  labs(x = expression("Adult trophic level"),
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC")) 

#ggsave(filename = "../figures/c4_updated/AdultTrophicLevelAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceMassCombined, width = 6, height = 6, dpi = 300)
```

# Combine the predictions and create the plot
```{r}
# abundance df
abundance <- NewDataAbundanceAdultTrophicOrder %>%           
  transmute(
    AvgTrophicLevelAdult, LandUseFinal, Order,
    pred = pred_backtransformed,
    lwr  = lwr_backtransformed,
    upr  = upr_backtransformed,
    model = "Abundance"
  )

# occurrence df
occurrence <- NewDataOccurrenceAdultTrophicOrder %>%              
  transmute(
    AvgTrophicLevelAdult, LandUseFinal, Order,
    pred = OccurrenceProb,
    lwr  = LowerProb,
    upr  = UpperProb,
    model = "Occurrence"
  )

# combined df
combined <- NewDataCombined %>%    
  transmute(
    AvgTrophicLevelAdult, LandUseFinal, Order,
    pred = PredictedAbundanceAccountingForOccurrenceProbability,
    lwr  = predictedcombinedlwr,
    upr  = predictedcombinedupr,
    model = "Combined"
  )

allpredictions <- bind_rows(abundance, occurrence, combined)

# Relevel so occurrence appears second in the plots
allpredictions <- allpredictions %>% 
  dplyr::mutate(
    model = factor(model,
                   levels = c("Abundance", "Occurrence", "Combined")))

base_plot <- function(df, ylab) {
  ggplot(df, aes(x = AvgTrophicLevelAdult, y = pred, color = LandUseFinal, fill = LandUseFinal)) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
    facet_grid(Order ~ ., scales = "free_y") +   # rows = orders
    scale_x_log10(labels = scales::label_number()) +
    labs(
      x = expression("Adult trophic level"),
      y = ylab,
      color = "Land use",
      fill  = "Land use"
    ) +
    theme_minimal() +
    theme(
      #strip.text.y.right = element_blank(),  # remove RHS facet labels if you like
      legend.position = "bottom" # legend position for each plot (collected later)
    )
}


# a) remove facet labels for the first two columns
p_abund <- base_plot(filter(allpredictions, model == "Abundance"),
                     "Abundance of present species (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank())  +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC")) 

p_occ <- base_plot(filter(allpredictions, model == "Occurrence"),
                   "Probability of occurrence (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank())  +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC")) 

# b) keep facet labels for the third column
p_comb <- base_plot(filter(allpredictions, model == "Combined"),
                    "Predicted abundance = a × b (± 95 % CI)") +
  theme(
    strip.text.y.right = element_text(size = 12, face = "bold")
  )  +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC")) 

# Combine — collect legends
FinalPlot <- (p_abund | p_occ | p_comb) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'a') &
  theme(
    legend.position = "bottom",
    plot.tag = element_text(size = 14, face = "bold"),
    plot.tag.position = c(0, 1)  # top-right corner
  )

FinalPlot

#saveRDS(allpredictions, "../C4 data/AdultTLAllPredictions.rds")
#ggsave(filename = "../figures/C4_updated/AdultTLOrders.png",  plot = FinalPlot, width = 14, height = 8, dpi = 300)
```

# Read in larval TL data
```{r}
ModelDataLarvalDietAbundance <- readRDS("../C4 data/ModelDataLarvalDietAbundance.rds")
ModelDataLarvalDietOccurrence <- readRDS("../C4 data/ModelDataLarvalDietOccurrence.rds")
```

# Modelling LogAbundanceWhenPresent ~ LUFinal * larval trophic level * order

```{r}
# Filter for big five
BigFiveModelDataLarvalDietAbundance <- ModelDataLarvalDietAbundance %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera")) %>%  droplevels()

# Maximal model
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevel_larval * Order + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = BigFiveModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Check the model
model_plot(ab_m1)
summary(ab_m1)
anova(ab_m1)
Anova(ab_m1)

# Calculate the larval trophic level range for predictions
LarvalTrophicLevelRange <- range(BigFiveModelDataLarvalDietAbundance$AvgTrophicLevel_larval, na.rm = TRUE)

# Create a grid of values
NewDataAbundanceLarvalTrophicOrder <- expand.grid(
  AvgTrophicLevel_larval  = seq(from = LarvalTrophicLevelRange[1],
                    to   = LarvalTrophicLevelRange[2],
                    length.out = 100),
  LandUseFinal = levels(BigFiveModelDataLarvalDietAbundance$LandUseFinal),
  Order        = levels(BigFiveModelDataLarvalDietAbundance$Order)
)

# Get predictions
AbundancePredictions <- predict(ab_m1, newdata = NewDataAbundanceLarvalTrophicOrder, re.form = NA, se.fit = TRUE)

# Predictions on log scale
NewDataAbundanceLarvalTrophicOrder$pred_log <- AbundancePredictions$fit
NewDataAbundanceLarvalTrophicOrder$se_log   <- AbundancePredictions$se.fit

# 95% CI on log10 scale
value <- qnorm(0.975)
NewDataAbundanceLarvalTrophicOrder$log_lwr <- NewDataAbundanceLarvalTrophicOrder$pred_log -
                                  value * NewDataAbundanceLarvalTrophicOrder$se_log
NewDataAbundanceLarvalTrophicOrder$log_upr <- NewDataAbundanceLarvalTrophicOrder$pred_log +
                                  value * NewDataAbundanceLarvalTrophicOrder$se_log

# Back-transform all of them
NewDataAbundanceLarvalTrophicOrder$pred_backtransformed <- exp(NewDataAbundanceLarvalTrophicOrder$pred_log)
NewDataAbundanceLarvalTrophicOrder$lwr_backtransformed <- exp(NewDataAbundanceLarvalTrophicOrder$log_lwr)
NewDataAbundanceLarvalTrophicOrder$upr_backtransformed  <- exp(NewDataAbundanceLarvalTrophicOrder$log_upr)


# Plot
PlotLarvalTrophicLevelAbundance <- ggplot(NewDataAbundanceLarvalTrophicOrder, aes(x = AvgTrophicLevel_larval, y = pred_backtransformed, color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = lwr_backtransformed, ymax = upr_backtransformed, fill = LandUseFinal),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))  +
  labs(x = expression("Larval trophic level"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC")) 

#ggsave(filename = "../figures/AbundanceWhenPresentBigFiveLarvalTrophicLevel.png",   plot = PlotLarvalTrophicLevelAbundance, width = 10, height = 6, dpi = 300)
```


# # Modelling Occurrence ~ LUFinal * Larval TL * order


```{r}
# Convert to factors
ModelDataLarvalDietOccurrence$Order <- as.factor(ModelDataLarvalDietOccurrence$Order)
ModelDataLarvalDietOccurrence$Family <- as.factor(ModelDataLarvalDietOccurrence$Family)
ModelDataLarvalDietOccurrence$SpeciesFinal <- as.factor(ModelDataLarvalDietOccurrence$SpeciesFinal)
ModelDataLarvalDietOccurrence$SS <- as.factor(ModelDataLarvalDietOccurrence$SS)
ModelDataLarvalDietOccurrence$SSB <- as.factor(ModelDataLarvalDietOccurrence$SSB)

# Filter for the big five
BigFiveModelDataLarvalDietOccurrence <- ModelDataLarvalDietOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera")) %>% droplevels()

m1 <- glmmTMB::glmmTMB(PresentOrAbsent ~
                         LandUseFinal * AvgTrophicLevel_larval * Order +
                         (1 | SS) + (1 | SpeciesFinal), 
                       data = BigFiveModelDataLarvalDietOccurrence,
                       family = binomial(link = "logit"),
                       control = glmmTMBControl(
                         optCtrl = list(iter.max = 1e3, eval.max = 1e3)
                         )
)
# Check the model
summary(m1)
Anova(m1)

# Create a grid of values
NewDataOccurrenceLarvalTrophicOrder <- expand.grid(
  AvgTrophicLevel_larval  = seq(from = LarvalTrophicLevelRange[1],
                    to   = LarvalTrophicLevelRange[2],
                    length.out = 100),
  LandUseFinal = levels(BigFiveModelDataLarvalDietOccurrence$LandUseFinal),
  Order        = levels(BigFiveModelDataLarvalDietOccurrence$Order)
)

# Get predictions on the link scale 
OccurrencePredictions <- predict(
  m1,
  newdata = NewDataOccurrenceLarvalTrophicOrder,
  re.form = NA,      # fixed-effects only (population-level)
  se.fit  = TRUE,    # get SE 
  type = "link"
)

# Add predictions into data frame
NewDataOccurrenceLarvalTrophicOrder <- NewDataOccurrenceLarvalTrophicOrder %>%
  mutate(
    OccurrenceFit   = OccurrencePredictions$fit,
    OccurrenceSE    = OccurrencePredictions$se.fit,
    OccurrenceLwr = OccurrenceFit - 1.96 * OccurrenceSE,
    OccurrenceUpr = OccurrenceFit + 1.96 * OccurrenceSE,
    # Back-transform to probability scale
    OccurrenceProb   = plogis(OccurrenceFit), # predicted probability - plogis() is the inverse of the logit function.
    # It takes values on the logit (link) scale (from −∞ to +∞) and transforms them into probabilities between 0 and 1.
    LowerProb = plogis(OccurrenceLwr), # 95% CI lower bound
    UpperProb = plogis(OccurrenceUpr) # 95% CI upper bound
  )



# Plot
PlotBigFiveOccurrenceLarvalTrophic <- ggplot(NewDataOccurrenceLarvalTrophicOrder, aes(x = AvgTrophicLevel_larval, y = OccurrenceProb, color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = LowerProb, ymax = UpperProb, fill = LandUseFinal),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold")) +
  labs(
  x = expression("Larval trophic level"),
    y = "Probability of occurrence (± 95 % CI)",
    color = "Land use",
    fill  = "Land use",
    title = ""
  ) +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC"))

#ggsave(filename = "../figures/C4_updated/LarvalTLBigFiveOccurrence.png",   plot = PlotBigFiveOccurrenceLarvalTrophic, width = 10, height = 6, dpi = 300)
```

# # Create the combined plot for larval trophic level

```{r}

NewDataCombined <- NewDataOccurrenceLarvalTrophicOrder %>%
  bind_cols(NewDataAbundanceLarvalTrophicOrder %>% 
              dplyr::select(pred_backtransformed = pred_backtransformed,
                     se_log  = se_log)) %>%
 mutate(
    # Probability of occurrence * predicted abundance when present = predicted abundance 
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * pred_backtransformed,  # c = a * b
    # Delta method
    occurrence_var  = (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2), # variance of plogis(fit_link)
    abundance_var = (log(10)^2) * (pred_backtransformed^2) * (se_log^2), # variance of 10^(fit_abund)

    combined_var =
      (pred_backtransformed^2) * occurrence_var +
      (OccurrenceProb^2) * abundance_var,

    predictedcombinedvar = sqrt(combined_var),

    # 95% CIs on original (abundance) scale
    predictedcombinedlwr =
      PredictedAbundanceAccountingForOccurrenceProbability -
      1.96 * predictedcombinedvar,
    predictedcombinedupr =
      PredictedAbundanceAccountingForOccurrenceProbability +
      1.96 * predictedcombinedvar
  )
    

# Plot
FinalPlotAbundanceOccurrenceLarvalTLCombined <- ggplot(NewDataCombined,
       aes(x = AvgTrophicLevel_larval, y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = predictedcombinedlwr, ymax = predictedcombinedupr),
              alpha = 0.2, color = NA) +
  facet_wrap(~ Order, scales = "free_y") +
  labs(x = "Larval trophic level",
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC"))


#ggsave(filename = "../figures/c4_updated/LarvalTrophicLevelAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceLarvalTLCombined, width = 6, height = 6, dpi = 300)
```

# Combine the predictions and create the plot
```{r}
# abundance df
abundance <- NewDataAbundanceLarvalTrophicOrder %>%           
  transmute(
    AvgTrophicLevel_larval, LandUseFinal, Order,
    pred = pred_backtransformed,
    lwr  = lwr_backtransformed,
    upr  = upr_backtransformed,
    model = "Abundance"
  )

# occurrence df
occurrence <- NewDataOccurrenceLarvalTrophicOrder %>%              
  transmute(
    AvgTrophicLevel_larval, LandUseFinal, Order,
    pred = OccurrenceProb,
    lwr  = LowerProb,
    upr  = UpperProb,
    model = "Occurrence"
  )

# combined df
combined <- NewDataCombined %>%    
  transmute(
    AvgTrophicLevel_larval, LandUseFinal, Order,
    pred = PredictedAbundanceAccountingForOccurrenceProbability,
    lwr  = predictedcombinedlwr,
    upr  = predictedcombinedupr,
    model = "Combined"
  )

allpredictions <- bind_rows(abundance, occurrence, combined)

# Relevel so occurrence appears second in the plots
allpredictions <- allpredictions %>% 
  dplyr::mutate(
    model = factor(model,
                   levels = c("Abundance", "Occurrence", "Combined")))

base_plot <- function(df, ylab) {
  ggplot(df, aes(x = AvgTrophicLevel_larval, y = pred, color = LandUseFinal, fill = LandUseFinal)) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
    facet_grid(Order ~ ., scales = "free_y") +   # rows = orders
    labs(
      x = "Larval trophic level",
      y = ylab,
      color = "Land use",
      fill  = "Land use"
    ) +
    theme_minimal() +
    theme(
      #strip.text.y.right = element_blank(),  # remove RHS facet labels if you like
      legend.position = "bottom" # legend position for each plot (collected later)
    ) +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC"))
}

# a) remove facet labels for the first two columns
p_abund <- base_plot(filter(allpredictions, model == "Abundance"),
                     "Abundance of present species (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank()) 

p_occ <- base_plot(filter(allpredictions, model == "Occurrence"),
                   "Probability of occurrence (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank())

# b) keep facet labels for the third column
p_comb <- base_plot(filter(allpredictions, model == "Combined"),
                    "Predicted abundance = a × b (± 95 % CI)") +
  theme(
    strip.text.y.right = element_text(size = 12, face = "bold")
  )

# Combine — collect legends
FinalPlot <- (p_abund | p_occ | p_comb) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'a') &
  theme(
    legend.position = "bottom",
    plot.tag = element_text(size = 14, face = "bold"),
    plot.tag.position = c(0, 1)  # top-right corner
  )

FinalPlot

#saveRDS(allpredictions, "../C4 data/LarvalTLAllPredictions.rds")
#ggsave(filename = "../figures/C4_updated/LarvalTLOrders.png",  plot = FinalPlot, width = 14, height = 8, dpi = 300)
```

# Read in dispersal data (have not updated dispersal plot yet)

```{r}
ModelDataDispersalAbundance <- readRDS("../C4 data/ModelDataDispersalAbundance.rds")
ModelDataDispersalOccurrence <- readRDS("../C4 data/ModelDataDispersalOccurrence.rds")
```

# Modelling LogAbundanceWhenPresent ~ LUFinal * dispersal ability * order

```{r}
# Filter for big five
BigFiveModelDataDispersalAbundance <- ModelDataDispersalAbundance %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))

# Maximal model
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * avg_dispersal * Order + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = BigFiveModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Check the model
model_plot(ab_m1)
summary(ab_m1)
anova(ab_m1)
Anova(ab_m1)

# Calculate the dispersal range for predictions
DispersalRange <- range(BigFiveModelDataDispersalAbundance$avg_dispersal, na.rm = TRUE)

# Generate predictions
# How does abundance scale with dispersal ability for each LU * Order combo
AbundancePredictions <- ggpredict(
  ab_m1,
  terms = c(
    sprintf("avg_dispersal [%.2f:%.2f]", DispersalRange[1], DispersalRange[2]), # Generate predictions using that range
    "LandUseFinal",
    "Order"
  ))

# Plot
PlotDispersalAbilityAbundance <- ggplot(AbundancePredictions, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.15, color = NA) +
  facet_wrap(~ facet, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))  +
  labs(x = expression("Dispersal ability"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  scale_x_continuous(breaks = 0:4, labels = c("very low", "low", "medium", "high", "very high")) +
  scale_y_continuous(
    labels = function(x) round(exp(x), 2)   # back-transform tick labels
  ) 


#ggsave(filename = "../figures/AbundanceWhenPresentBigFiveDispersalAbility.png",   plot = PlotDispersalAbilityAbundance, width = 10, height = 6, dpi = 300)
```

# Modelling Occurrence ~ LUFinal * Dispersal Ability * order

```{r}
# Filter for the big five
BigFiveModelDataDispersalOccurrence <- ModelDataDispersalOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))


m1 <- glmmTMB::glmmTMB(PresentOrAbsent ~
                         LandUseFinal * avg_dispersal * Order +
                         (1 | SS) + (1 | Genus) + (1 | SpeciesFinal), 
                       data = BigFiveModelDataDispersalOccurrence,
                       family = binomial(link = "logit"),
                       control = glmmTMBControl(
                         optCtrl = list(iter.max = 1e3, eval.max = 1e3)
                         )
)

# Check the model
summary(m1)
Anova(m1)

# Calculate the range for predictions
dispersal_range <- range(BigFiveModelDataDispersalOccurrence$avg_dispersal, na.rm = TRUE)

# Generate predictions
BigFiveOccurrencePredictions <- ggpredict(
  m1,
  terms = c(
    sprintf("avg_dispersal [%.2f:%.2f]", dispersal_range[1], dispersal_range[2]), # Generate predictions using that range
    "LandUseFinal",
    "Order"
  ))

# Plot
PlotBigFiveOccurrenceDispersal <- ggplot(BigFiveOccurrencePredictions, aes(x = x, y = predicted, color = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.15, color = NA) +
  facet_wrap(~ facet, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold")) +
  labs(
  x = expression("Dispersal ability"),
    y = "Probability of occurrence (± 95 % CI)",
    color = "Land use",
    fill  = "Land use",
    title = ""
  ) +
  scale_x_continuous(breaks = 0:4, labels = c("very low", "low", "medium", "high", "very high"))

#ggsave(filename = "../figures/C4_updated/DispersalBigFiveOccurrence.png",   plot = PlotBigFiveOccurrenceDispersal, width = 10, height = 6, dpi = 300)
```

# Combine the predictions

```{r}
AbundancePredictions$ExpAbundFit <- exp(AbundancePredictions$predicted)  # back-transform
AbundancePredictions$AbundLwr <- exp(AbundancePredictions$conf.low) # exponentiate log abundance values 
AbundancePredictions$AbundUpr <- exp(AbundancePredictions$conf.high)

AbundancePredictions <- AbundancePredictions %>% rename(AbundSE = std.error)

BigFiveOccurrencePredictions <- BigFiveOccurrencePredictions %>% rename(OccurrenceProb = predicted, OccurrenceSE = std.error)

# Turn ggeffects objects into plain data frames first
occurrence <- BigFiveOccurrencePredictions  |> as.data.frame()
abundance <- AbundancePredictions      |> as.data.frame()

NewDataCombined <- occurrence %>%
  bind_cols(abundance %>% 
              dplyr::select(ExpAbundFit = ExpAbundFit,
                     AbundSE  = AbundSE)) %>%
  mutate(
    # Probability of occurrence * predicted abundance when present 
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * ExpAbundFit,
    
    # Delta method for SE of product on natural scale
    PredictedAbundanceSE = sqrt(
      (ExpAbundFit^2) * (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2) +  # variance of plogis(fit_link)
      (OccurrenceProb^2) * (ExpAbundFit^2) * (AbundSE^2)                      # variance of exp(fit_abund)
    ),
    
    # 95% confidence intervals
    PredictedAbundanceLwr = PredictedAbundanceAccountingForOccurrenceProbability - 1.96 * PredictedAbundanceSE,
    PredictedAbundanceUpr = PredictedAbundanceAccountingForOccurrenceProbability + 1.96 * PredictedAbundanceSE
  )

# Plot
FinalPlotAbundanceOccurrenceDispersalCombined <- ggplot(NewDataCombined,
       aes(x = x, y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = group, fill = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = PredictedAbundanceLwr, ymax = PredictedAbundanceUpr),
              alpha = 0.2, color = NA) +
  facet_wrap(~ facet, scales = "free_y") +
  labs(x = expression("Dispersal ability"),
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = 0:4, labels = c("very low", "low", "medium", "high", "very high")) 

#ggsave(filename = "../figures/c4_updated/DispersalAbilityAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceDispersalCombined, width = 6, height = 6, dpi = 300)
```

# Combine the predictions and create the plot
```{r}
plot_dat_abund <- abundance %>%           # <-- abundance df
  transmute(
    x, group, facet,
    pred = ExpAbundFit,
    lwr  = AbundLwr,
    upr  = AbundUpr,
    model = "Abundance"
  )

plot_dat_occ <- occurrence %>%              # <-- occurrence df
  transmute(
    x, group, facet,
    pred = OccurrenceProb,
    lwr  = conf.low,
    upr  = conf.high,
    model = "Occurrence"
  )

plot_dat_comb <- NewDataCombined %>%    # combined df
  transmute(
    x, group, facet,
    pred = PredictedAbundanceAccountingForOccurrenceProbability,
    lwr  = PredictedAbundanceLwr,
    upr  = PredictedAbundanceUpr,
    model = "Combined"
  )

allpredictions <- bind_rows(plot_dat_abund, plot_dat_occ, plot_dat_comb)

allpredictions <- allpredictions %>% 
  dplyr::mutate(
    model = factor(model,
                   levels = c("Abundance", "Occurrence", "Combined")))

base_plot <- function(df, ylab) {
  ggplot(df, aes(x = x, y = pred, color = group, fill = group)) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
    facet_grid(facet ~ ., scales = "free_y") +   # rows = orders
  scale_x_continuous(breaks = 0:4, labels = c("very low", "low", "medium", "high", "very high"))+
    labs(
      x = "Dispersal ability",
      y = ylab,
      color = "Land use",
      fill  = "Land use"
    ) +
    theme_minimal() +
    theme(
      #strip.text.y.right = element_blank(),  # remove RHS facet labels if you like
      legend.position = "bottom" # legend position for each plot (collected later)
    )
}

# a) remove facet labels for the first two columns
p_abund <- base_plot(filter(allpredictions, model == "Abundance"),
                     "Abundance of present species (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank())

p_occ <- base_plot(filter(allpredictions, model == "Occurrence"),
                   "Probability of occurrence (± 95 % CI)") +
  theme(strip.text.y = element_blank(),
        strip.background.y = element_blank())

# b) keep facet labels for the third column
p_comb <- base_plot(filter(allpredictions, model == "Combined"),
                    "Predicted abundance = a × b (± 95 % CI)") +
  theme(
    strip.text.y.right = element_text(size = 12, face = "bold")
  )


# Combine — collect legends
FinalPlot <- (p_abund | p_occ | p_comb) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'a') &
  theme(
    legend.position = "bottom",
    plot.tag = element_text(size = 14, face = "bold"),
    plot.tag.position = c(0, 1)  # top-right corner
  )

FinalPlot

#saveRDS(allpredictions, "../C4 data/DispersalAllPredictions.rds")
#ggsave(filename = "../figures/C4_updated/DispersalAbilityOrders.png",  plot = FinalPlot, width = 14, height = 8, dpi = 300)

```

```{r summary}
best <- m1
summary(best)
performance::r2(best)
car::Anova(best)
```

```{r diagnostics}
sjPlot::plot_model(best,
                   type = "diag")
glmmTMB::diagnose(best)
sim <- DHARMa::simulateResiduals(best, nSim = 250)
plot(sim)
DHARMa::testZeroInflation(sim)
DHARMa::testDispersion(sim)
DHARMa::plotResiduals(sim, form = BigFiveModelBodyMassOccurrence$Order)
DHARMa::plotResiduals(sim, form = BigFiveModelBodyMassOccurrence$LandUseFinal)
DHARMa::plotResiduals(sim, form = BigFiveModelBodyMassOccurrence$log10_mass)
```
