---
title: "Untitled"
output: html_document
date: "2025-08-15"
---

## R Markdown




# Clear the work space

```{r}
rm(list=ls())
```

# Read in the packages 
```{r}
library(lme4)
library(readxl)
library(tidyverse)
library(stringr)
library(car)
library(qs)
library(lmerTest)
library(merTools) 
library(performance)
library(DHARMa)
library(here)
library(blme)
library(MuMIn)
library(parallel)
library(cowplot)
library(ggeffects)
library(patchwork)
```


# Read in Logghe et al 2025 which has body size data for ~ 4800 arthropod species and makes up the bulk of the GLiTRS trait data set on body size
# Tidy and clean this data frame to prepare it to merge with the PREDICTS trait data sets
```{r}
# Upload logghe et al
logghe <- read_excel(here("data", "Body_Size_data.xlsx"))

# keep only first 8 cols
logghe <- logghe %>% dplyr::select(1:8)

# Make first row the header
colnames(logghe) <- as.character(logghe[1, ])

# drop first row
logghe <- logghe[-1, ] 

# Create a genus and species column
logghe <- logghe %>%
  separate(scientificName, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop")

# Combine genus and species to create Best_guess_binomial which I will use to merge 
logghe <- logghe %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Reorder columns
logghe <- logghe %>% dplyr::select(order, family, Genus, Species, Best_guess_binomial, Trait_Value) %>% 
  rename(Family = family) %>% 
  rename(Order = order)

# Convert to numeric 
logghe$Trait_Value <-  as.numeric(logghe$Trait_Value)
```

# Read and tidy Geoffrey Su's ant trait data
```{r}
# Load the data
trait_ants <- read_excel(here("data", "SuFormacidaeBodyLength-2014-05-23.xlsx"))

# Rename body length average column into something more readable
trait_ants <- trait_ants %>% rename(avg_body_length_mm = "Body.Length..mm...worker..AVG")

# Lets remove those rows with NAs for body length
trait_ants <- trait_ants %>%
  filter(!is.na(avg_body_length_mm) & avg_body_length_mm != 0) # 559 rows with no body length data, 2 rows where BL = 0

# Create best_guess_binomial by combining genus and species  
trait_ants <- trait_ants %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " "))

# Filter for the cols of interest for merging
trait_ants <- trait_ants %>% dplyr::select(Genus, Best_guess_binomial, avg_body_length_mm)

# Tidy
trait_ants <- trait_ants %>%
  mutate(
    Order = "Hymenoptera",
    Family = "Formicidae"
  ) %>% 
  rename(Trait_Value = avg_body_length_mm) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop") %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)
# ants ready
```

# Read and tidy Melanie Edgars beetle trait data
```{r}
# Load Melanies beetle trait data
melanie_beetles <- read_excel(here("data", "EdgarBeetleLength-2014-01-09.xlsx"))

# Combine genus and species for merging
melanie_beetles <- melanie_beetles %>% mutate(Best_guess_binomial = str_c(GENUS, SPECIES, sep = " "))

# Calculate average b.length
melanie_beetles <- melanie_beetles %>%
  separate(B.LENGTH, into = c("b_min", "b_max"), sep = "-", fill = "right", remove = FALSE) %>% # fill = "right" allows single values to stay in b_min while b_max becomes NA.
  mutate(
    b_min = as.numeric(trimws(b_min)), # trimws() removes any leading/trailing spaces 
    b_max = as.numeric(trimws(b_max)),
    b.length_avg = if_else(
      is.na(b_max),
      b_min,  # If no range, use the min value
      (b_min + b_max) / 2  # Else if range, compute the average
    ))

# Prepare and tidy beetles
melanie_beetles <- melanie_beetles %>% 
  mutate(Order = "Coleoptera") %>% 
  dplyr::select(Order, FAMILY, GENUS, SPECIES, Best_guess_binomial, b.length_avg) %>% 
  rename(Family = FAMILY) %>% 
  rename(Genus = GENUS) %>% 
  rename(Species = SPECIES) %>% 
  rename(Trait_Value = b.length_avg)
```

# Read and tidy Rachel Tans bee trait data
```{r}
# Load Rachel bee trait data
rachel_bee <- read_excel(here("data", "BeeTraits_P2_RachelTan.xlsx"))

# Create a genus and species column
rachel_bee <- rachel_bee %>%
  separate(CorrectedName, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "drop")

# Filter for columns of interest
rachel_bee <- rachel_bee %>% dplyr::select(HigherTaxon, Family, Genus, Species, CorrectedName, MidRangeBodyLength) %>% drop_na() # drop those sp. with NA for body length

# Remove rows identified as "sp." or inside brackets e.g., (Dialictus) or (Psithyrus) which are subgenera and not a true species name
rachel_bee <- rachel_bee %>%
  filter(
    !str_detect(Species, "sp\\.") &         # remove rows with "sp."
    !str_detect(Species, "\\(.*\\)")        # remove rows with anything in brackets
  )

# Rename for merging
rachel_bee <- rachel_bee %>% 
  rename(Best_guess_binomial = CorrectedName) %>% 
  rename(Order = HigherTaxon) %>% 
  rename(Trait_Value = MidRangeBodyLength)

```

# Merge Logghe, Geoffrey, Melanies and Rachels trait datasets to have an overall body size trait dataset
```{r}
# merge all trait datasets
all_traits <- bind_rows(logghe, melanie_beetles, rachel_bee, trait_ants)

# There are duplicate species entries
all_traits[duplicated(all_traits["Best_guess_binomial"]) | 
                                  duplicated(all_traits["Best_guess_binomial"], fromLast = TRUE), ] %>%  View()

# Sense check - there are 634 duplicates
all_traits[duplicated(all_traits["Best_guess_binomial"]),] 

# For those species with duplicate entries, calculate an average Trait_Value which is the mean of the duplicate entries
all_traits_clean <- all_traits %>%
  group_by(Best_guess_binomial) %>%
  summarise(
    Trait_Value = mean(Trait_Value, na.rm = TRUE),  # average Trait_Value
    across(.cols = -Trait_Value, .fns = ~ first(.x)),  # for each column, take the first value
    .groups = "drop"
  ) %>% 
  dplyr::select(Order, Family, Genus, Species, Best_guess_binomial, Trait_Value)

nrow(all_traits) - nrow(all_traits_clean) # 634

rm(all_traits, logghe, melanie_beetles, rachel_bee, trait_ants)
```

# Read in PREDICTS data 
```{r}
# All insect data effort-corrected and merged
taxon_insects <-  qread("/Users/jusi/Desktop/multi-groups/data/predicts-prepared-whole-database.qs")

# Get rid of rows which aren't resolved to order level
taxon_insects <- taxon_insects %>% filter(Order != "") # 1056 rows not resolved to order

# Recreate genus and species column as sometimes the best_guess_binomial column has a name in PREDICTS but there is no entry in the genus or species column
taxon_insects <- taxon_insects %>%
  mutate(Best_guess_binomial = as.character(Best_guess_binomial)) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "merge", fill = "right")

length(unique(taxon_insects$SS))
```

# Clean and tidy PREDICTS data then merge trait dataset based on taxon resolution
```{r}
# There are entries in the column Taxon_name_entered such as "Lasioglossum sp" - such rows tend not to have an entry in the genus column (i.e. Lasioglossum) due to issues with the col API.
# I realised that these rows were therefore being assigned an average mass value at the family level, rather than the genus level - despite the genus information being available. 
# I will extract the genus information from the Taxon_name_entered column for those rows that do not yet have an entry in genus.
# This will improve the precision of the matched mass data, such that these rows will now be matched at the genus level instead of the family level (as they were beforehand)
taxon_insects <- taxon_insects %>%
  mutate(
    Genus = case_when(
      is.na(Genus) | Genus == "" ~ word(Taxon_name_entered, 1), # Check if Genus is empty or missing, if so, extract the first word from Taxon_name_entered and assign it to genus.
      TRUE ~ Genus # If genus is not empty, leave it as the pre-existing genus entry. 
    ))

# Have a look
taxon_insects %>% dplyr::select(SS, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, Precision_score) %>% View()

# Remove those entries which have a "?" before or after the name of the genus e.g. "?Abalabus" or "Abalabus?"
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove_all(Genus, "\\?"))

# Remove leading and trailing double quotation marks
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_replace_all(Genus, '^"|"$', ""))

# Remove any numbers 
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove_all(Genus, "[0-9]"))

# Remove entries with sp at the end
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove(Genus, "(_|\\s)?sp\\.?$"))

# Get species trait values for merging
sp_value <- all_traits_clean %>% dplyr::select(Best_guess_binomial, Trait_Value) 

# First merge Logghe species level data
taxon_insects <- left_join(taxon_insects, sp_value, by = "Best_guess_binomial")

rm(sp_value)

# Calculate average body length for genera
genus_value <- all_traits_clean %>%
  group_by(Genus) %>%
  summarise(genus_avg_body_length = mean(Trait_Value, na.rm = TRUE)) # 2381 genus data

# Merge based on genus where best_guess_binomial is empty
taxon_insects <- taxon_insects %>%
  left_join(genus_value, by = "Genus", suffix = c("", "_genus")) 

rm(genus_value)

# Average body length per family based on species
family_value <- all_traits_clean %>%
  group_by(Family) %>%
  summarise(family_avg_body_length = mean(Trait_Value, na.rm = TRUE))

# Merge based on family where genus and species is empty
taxon_insects <- taxon_insects %>%
  left_join(family_value, by = "Family", suffix = c("", "_family")) 

rm(family_value)

# priority-based fallback assignment where final body length values are selected based on the first non-missing (or non-blank) column in a defined order.
taxon_insects <- taxon_insects %>%
  mutate(final_body_length = coalesce(
    Trait_Value,
    genus_avg_body_length,
    family_avg_body_length
  ))

# How many rows have data based on each level of taxonomic resolution
taxon_insects %>%
  summarise(across(
    c(Trait_Value,
    genus_avg_body_length,
    family_avg_body_length,
    final_body_length),
    list(
      non_NAs = ~sum(!is.na(.))
    )
  ))
```

# Rainford et al 2014 provides min and max body length data for 760 insect families, I will calculate the average body length between the min and the max and then merge those values onto those rows in my data frame that do not yet have a body length value. 

# First lets tidy the data frame 

```{r}
# Upload 
rainford <- read_excel(here("data", "rainford_body_size.xlsx"))

# Make first row the header
colnames(rainford) <- as.character(rainford[1, ])
rainford <- rainford[-1, ] # drop first row

# Extract cols of interest
rainford <- rainford %>% dplyr::select("Order", "Family", "Min (mm)", "Max (mm)") %>% rename(min_bl = "Min (mm)") %>% rename(max_bl = "Max (mm)")

# Drop rows with all NAs
rainford <- rainford %>%
  filter(!if_all(everything(), is.na))

# Convert cols to numeric
rainford$min_bl <- as.numeric(rainford$min_bl) # Odonata (Pseudolestidae) has an NA for BL hence the warning message
rainford$max_bl <- as.numeric(rainford$max_bl)

# Drop rows only to Order level
rainford <- rainford %>%
  filter(!is.na(Family))

# Calculate average body length
rainford <- rainford %>%
  mutate(Rainford_family_avg = (min_bl + max_bl) / 2)
         
# Simplify for merging
rainford <- rainford %>% 
  dplyr::select(c(-min_bl, -max_bl, -Order))
```

# Merge family average body length from Rainford for those rows in my data frame that do not yet have a body length value

```{r}
# Step 1: Filter rows in taxon_insects where there is no body size value at the species, genus or family level and therefore we can merge a family body size value from Rainford
needs_family_avg <- taxon_insects %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length))

# Step 2: Join family_avg body size values from Rainford based on Family column
needs_family_avg <- needs_family_avg %>%
  left_join(rainford, dplyr::select(Family, Rainford_family_avg), by = "Family")

# Step 3: Combine updated rows (need_family_avg) back into original data (taxon_insects)

# First, remove the rows you updated from taxon_insects
remaining_rows <- taxon_insects %>%
  filter(!(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length)))

# sense check remaining rows should be taxon_insects - needs_family_avg
nrow(taxon_insects) - nrow(needs_family_avg)

# Bind the updated rows with the remaining rows
taxon_insects_updated <- bind_rows(remaining_rows, needs_family_avg)

# How many rows still without a family average
sum(!is.na(taxon_insects_updated$Rainford_family_avg)) # 134120

# How many rows have data based on each level of taxonomic resolution
taxon_insects_updated %>%
  summarise(across(
    c(Trait_Value,
    genus_avg_body_length,
    family_avg_body_length,
    final_body_length,
    Rainford_family_avg),
    list(
      non_NAs = ~sum(!is.na(.))
    )
  ))

rm(needs_family_avg, remaining_rows, taxon_insects)
```

# Now that we have merged all the trait data,  I will coalesce the family average body length values from Rainford into the final_body_length column (for those rows that didn't previously have a body length value at the family level) - if they had already been assigned a body length value at the family level from Logghe or any of the PREDICTS trait datasets - then we will stick with this value.

```{r}
# Coalesce those rows with missing values in final_body_length with the corresponding family_avg
taxon_insects_updated <- taxon_insects_updated %>%
  mutate(final_body_length = coalesce(final_body_length, Rainford_family_avg)) # coalesce(a,b) returns a if its not NA otherwise it returns b

# How many rows do we still not have a body size value for and are still NA - 69814
taxon_insects_updated %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length) & is.na(final_body_length) & is.na(Rainford_family_avg)) %>%
  nrow()

# Have a look at the rows
taxon_insects_updated %>%
  filter(is.na(Trait_Value) & is.na(genus_avg_body_length) & is.na(family_avg_body_length) & is.na(final_body_length) & is.na(Rainford_family_avg)) %>% 
  View()

# Filter out rows which still don't have a body size value
all_insects_size <- taxon_insects_updated %>%
  filter(!is.na(final_body_length)) %>% # 1195212
  droplevels()

```

# Sohlstrom et al 2018 has tropical and temperate allometries for different insect orders since the scaling between mass and length differs significantly between either realm. We will therefore split sites in PREDICTS based on their latitude into either tropical or temperate realms.

```{r}
# Add a new column classifying each site into tropical or temperate
all_insects_size$climate_zone <- ifelse(all_insects_size$Latitude >= -23.4368 & all_insects_size$Latitude <= 23.4368, 
                          "tropical", 
                          "temperate")
```

# Read in allometry coefficients
```{r}
coefficients <- read_excel(here("data", "allometry_coefficients.xlsx"))
```

# Merge allometry coefficients to size data
```{r}
# Step 1: Join by order and climate_zone
all_insects_size <- all_insects_size %>%
  left_join(coefficients, by = c("Order", "climate_zone"))

# Step 2: For those taxa in PREDICTS which are not in the big five orders, we will assign them the allometries for all insects
fallback <- coefficients %>%
  filter(Order == "All_insects") %>%
  dplyr::select(climate_zone, a_fallback = a, b_fallback = b)

# Step 3: Join allometries for All_insects based on whether its a tropical or temperate site
all_insects_size <- all_insects_size %>%
  left_join(fallback, by = "climate_zone") %>%
  mutate(
    a = if_else(is.na(a), a_fallback, a), # if a is NA, assign it a for all_insects, else keep it as its order specific a
    b = if_else(is.na(b), b_fallback, b)
  ) %>%
  dplyr::select(-a_fallback, -b_fallback)

# Have a look
all_insects_size %>% dplyr::select(SSBS, Order, final_body_length, Latitude, climate_zone, a, b) %>% unique() %>%  View()
```

# Calculate mass in mg using allometric specific coefficients for each order
```{r}
AllInsectsSize <- all_insects_size %>%
  mutate(
    log10_mass = a + b * log10(final_body_length),
    mass_mg = 10^log10_mass
  )

summary(all_insects_size$final_body_length) # no zero values for final_body_length as min is 0.6
any(all_insects_size$final_body_length == 0) # FALSE

rm(coefficients, fallback, rainford, all_insects_size, all_traits_clean, taxon_insects_updated)
```


# Now we have a data frame with a mass value for all the insect records in PREDICTS. There are many more species level entities in PREDICTS than just filtering based on Best_guess_binomial or the Species column, this is because as mentioned earlier there are many rows in the Taxon_name_entered column such as "Lasioglossum sp" which are species level entities but would be dropped out if I filtered for species level entities using Best_guess_binomial or the Species column (as I did previously). AP has sent an updated taxon resolution file with all of the species level entities in PREDICTS, I will use this file to filter my data frame for all species level entities as my analysis will be at the species level. 

```{r}
# Lets have a look at those rows in Taxon_name_entered that are likely a species level entity
AllInsectsSize %>%
  filter(str_detect(Taxon_name_entered, regex("sp\\.|cf|aff|[0-9]", ignore_case = TRUE))) %>% 
dplyr::select(SS, SSBS, Taxon_name_entered, Best_guess_binomial, Family, Genus, Species, Trait_Value, genus_avg_body_length, family_avg_body_length, final_body_length, log10_mass) %>% View()

# Filter for columns of interest
AllInsectsSizeFiltered <- AllInsectsSize %>% dplyr::select(Source_ID, SS, SSB, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, log10_mass, Diversity_metric, Diversity_metric_unit, Merged_diversity, Predominant_habitat, Use_intensity)

# Filter for diversity metrics of interest - abundance and occurrence data can be used
AllInsectsSizeFiltered <- AllInsectsSizeFiltered %>% filter(Diversity_metric %in% c("abundance", "effort-corrected abundance", "density", "group abundance", "occurrence", "occurrence frequency", "relative abundance"))

# Load taxon resolution file
TaxonResolution <- readRDS(here("data", "CuratedTaxonResolution2.rds"))

# Step 1: Filter TaxonResolution for species-level entries - i.e. those entries with a precision score of 4!
TaxonResolutionSp <- TaxonResolution %>%
  filter(PrecisionScore == 4) %>%
  dplyr::select(SS, Taxon_name_entered) # select cols for merging

# Step 2: Keep only matching rows in AllInsectsSizeFiltered
OnlySpecies <- AllInsectsSizeFiltered %>%
  semi_join(TaxonResolutionSp, by = c("SS", "Taxon_name_entered")) # keeps only rows in AllInsectsSizeFiltered that have a match in TaxonResolutionSp. based on SS and Taxon_name_entered

# OnlySpecies now contains only the rows from AllInsectsSizeFiltered that correspond to species-level entries in TaxonResolutionSp. 

# Lets have a look at those rows which have an entry for Taxon_name_entered but no entry for Best_guess_binomial!
OnlySpecies %>%
  filter(
    !is.na(Taxon_name_entered) & Taxon_name_entered != "",
    is.na(Best_guess_binomial) | Best_guess_binomial == ""
  ) %>% View()

# Create a present or absent column - its present if Merged_diversity > 0 and absent if its 0
OnlySpecies <- OnlySpecies %>%
  mutate(PresentOrAbsent = ifelse(Merged_diversity > 0, 1, 0))

# Create a log(abundance when present) column
OnlySpecies <- OnlySpecies %>% 
  mutate(
    LogAbundanceWhenPresent = ifelse(Merged_diversity > 0,
                                    log(Merged_diversity),
                                    NA))

# Have a look at how many rows are present versus absent
table(OnlySpecies$PresentOrAbsent)

OnlySpecies %>% dplyr::select(SS, SSBS, Taxon_name_entered, Best_guess_binomial, PresentOrAbsent, Merged_diversity, LogAbundanceWhenPresent) %>% View()


# Rename predominant habitat since its not really habitat we're looking at
OnlySpecies <- rename(OnlySpecies,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
OnlySpecies <- OnlySpecies %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(OnlySpecies$Predominant_land_use)
table(OnlySpecies$Predominant_land_use, OnlySpecies$Use_intensity)
```

# Following Charlie and Tims analysis, I will collapse LU categories into Primary, Secondary vegetation, Low-intensity agriculture and high-intensity agriculture so that we have a manageable number of coefficients to analyse. 

```{r}

OnlySpecies <- OnlySpecies %>%
  mutate(
    LandUseFinal = case_when(
      # Primary stays as is
      Predominant_land_use == "Primary" ~ "Primary",

      # Any secondary vegetation type -> Secondary Vegetation
      Predominant_land_use %in% c("Young secondary vegetation",
                                  "Intermediate secondary vegetation",
                                  "Mature secondary vegetation") ~ "Secondary vegetation",

      # Plantation forest depends on intensity
      Predominant_land_use == "Plantation forest" & Use_intensity == "Minimal use" ~ "Low-intensity agriculture",
      Predominant_land_use == "Plantation forest" & Use_intensity %in% c("Light use", "Intense use") ~ "High-intensity agriculture",

      # Pasture
      Predominant_land_use == "Pasture" & Use_intensity %in% c("Minimal use", "Light use") ~ "Low-intensity agriculture",
      Predominant_land_use == "Pasture" & Use_intensity == "Intense use" ~ "High-intensity agriculture",

      # Cropland
      Predominant_land_use == "Cropland" & Use_intensity %in% c("Minimal use", "Light use") ~ "Low-intensity agriculture",
      Predominant_land_use == "Cropland" & Use_intensity == "Intense use" ~ "High-intensity agriculture",

      # Drop Urban by assigning NA -these will get dropped when we create the model data
      Predominant_land_use == "Urban" ~ NA_character_
    )
  ) %>%
  # Turn into factor if you want ordered categories
  mutate(LandUseFinal = factor(LandUseFinal,
                               levels = c("Primary", "Secondary vegetation",
                                          "Low-intensity agriculture", "High-intensity agriculture")))

table(OnlySpecies$LandUseFinal)
```

# Create random intercept for Species

```{r}
OnlySpecies <- OnlySpecies %>%
  mutate(
    SpeciesFinal = case_when(
      Species != "" & !is.na(Species) ~ Species, # If species is present, use that as the random intercept
      (Species == "" | is.na(Species)) & Best_guess_binomial != "" & !is.na(Best_guess_binomial) ~ Best_guess_binomial, # If species is blank, use BGB as the random intercept
      (Species == "" | is.na(Species)) & (Best_guess_binomial == "" | is.na(Best_guess_binomial)) ~ paste(SS, Taxon_name_entered, sep = " "),
      # If Species and BGB is blank, use SS and Taxon_name_entered combined as the random intercept
      TRUE ~ NA_character_
    )
  )

# Lets have a look 
OnlySpecies %>% dplyr::select(SS, SSBS, Order, Family, Genus, Best_guess_binomial, Species, Taxon_name_entered, SpeciesFinal) %>% View()


sum(is.na(OnlySpecies$SpeciesFinal))
OnlySpecies %>% filter(SpeciesFinal == "")
sum(is.na(OnlySpecies$SpeciesFinal) | OnlySpecies$SpeciesFinal == "")


length(unique(OnlySpecies$SpeciesFinal))
length(unique(OnlySpecies$Taxon_name_entered))
length(unique(OnlySpecies$Best_guess_binomial))
length(unique(OnlySpecies$Species))
```

# Create model data

```{r}
# drop all rows with NAs for variables of interest
ModelDataSizeOccurrence <- drop_na(OnlySpecies, LandUseFinal, Use_intensity, PresentOrAbsent) %>% droplevels()
ModelDataSizeAbundance <- drop_na(OnlySpecies, LandUseFinal, Use_intensity, LogAbundanceWhenPresent) %>% droplevels()

length(unique(ModelDataSizeAbundance$SS)) # 389 studies 
length(unique(ModelDataSizeAbundance$SSBS)) # 8896 sites

length(unique(ModelDataSizeOccurrence$SS)) # 389 studies 
length(unique(ModelDataSizeOccurrence$SSBS)) # 9658 sites

rm(TaxonResolution, TaxonResolutionSp)
```

# The number of species matched to species-, genus- and family- level trait data for body mass

```{r}
length(unique(OnlySpecies$Best_guess_binomial[!is.na(OnlySpecies$Trait_Value)]))
length(unique(OnlySpecies$Best_guess_binomial[!is.na(OnlySpecies$genus_avg_body_length)]))
length(unique(OnlySpecies$Best_guess_binomial[!is.na(OnlySpecies$family_avg_body_length)]))
```

#### Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(ModelDataSizeAbundance[ , c("LandUseFinal", "Use_intensity", "log10_mass", "LogAbundanceWhenPresent")])
corvif(ModelDataSizeOccurrence[ , c("LandUseFinal", "Use_intensity","log10_mass", "PresentOrAbsent")])
```

#### Load in model plot function

```{r}
## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(2,2))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}
```

# Modelling AbundanceWhenPresent ~ LUFinal * Mass

```{r}
# basic data diagnostics
str(ModelDataSizeAbundance)
table(ModelDataSizeAbundance$LandUseFinal, useNA="ifany")
n_distinct(ModelDataSizeAbundance$Source_ID)
n_distinct(ModelDataSizeAbundance$SS)
n_distinct(ModelDataSizeAbundance$SpeciesFinal)

# Convert to factors
ModelDataSizeAbundance$Order <- as.factor(ModelDataSizeAbundance$Order)
ModelDataSizeAbundance$Family <- as.factor(ModelDataSizeAbundance$Family)
ModelDataSizeAbundance$SpeciesFinal <- as.factor(ModelDataSizeAbundance$SpeciesFinal)
ModelDataSizeAbundance$SS <- as.factor(ModelDataSizeAbundance$SS)
ModelDataSizeAbundance$SSB <- as.factor(ModelDataSizeAbundance$SSB)
```


# Backwards stepwise selection of random effects

```{r}
# Maximal random effects structure
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * log10_mass + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = ModelDataSizeAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop block
ab_m2 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * log10_mass + (1 | Source_ID/SS) + (1| Family/SpeciesFinal), 
               data = ModelDataSizeAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop SS
ab_m3 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * log10_mass + (1 | Source_ID) + (1| Family/SpeciesFinal), 
               data = ModelDataSizeAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


# Drop Family
ab_m4 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * log10_mass + (1 | Source_ID) + (1 | SpeciesFinal), 
               data = ModelDataSizeAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Compare models
anova(ab_m1, ab_m2, ab_m3, ab_m4)  # Likelihood ratio test
AIC(ab_m1, ab_m2, ab_m3, ab_m4)


# ab_m1 best!
```

# Backwards stepwise selection of fixed effects

```{r}
# Additive model with the interaction removed but both predictors still present
ab_m5 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + log10_mass + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataSizeAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


# Just land use as predictor
ab_m6 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataSizeAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Just Mass as the predictor
ab_m7 <- blmer(LogAbundanceWhenPresent ~ log10_mass + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataSizeAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Make a null, intercept only model
ab_m8 <- blmer(LogAbundanceWhenPresent ~ (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataSizeAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

## Compare the models
AIC(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m7, ab_m5, ab_m1)
```

# Ok so the interaction is not supported, the additive model model is the best fit - LU and body mass affect AbundanceWhenPresent independently and both matter but no evidence that the impact of body mass changes depending on LU class. We will still go ahead and do the predictions/create the plot.

```{r}
# Model checks
model_plot(ab_m1)
summary(ab_m1)

# Create a grid of values
NewDataAbundanceMass <- expand.grid( 
  log10_mass = seq(from = min(ModelDataSizeAbundance$log10_mass),
                   to = max(ModelDataSizeAbundance$log10_mass),
                   length.out = 100),
  LandUseFinal = levels(ModelDataSizeAbundance$LandUseFinal))
  
# Get predictions without random effects, with SEs
AbundancePredictions <- predict(ab_m1, newdata = NewDataAbundanceMass, re.form = NA, se.fit = TRUE)

# Predictions on log10 scale
NewDataAbundanceMass$pred_log10 <- AbundancePredictions$fit
NewDataAbundanceMass$se_log10   <- AbundancePredictions$se.fit

# 95% CI on log10 scale
value <- qnorm(0.975)
NewDataAbundanceMass$log10_lwr <- NewDataAbundanceMass$pred_log10 -
                                  value * NewDataAbundanceMass$se_log10
NewDataAbundanceMass$log10_upr <- NewDataAbundanceMass$pred_log10 +
                                  value * NewDataAbundanceMass$se_log10

# Back-transform all of them
NewDataAbundanceMass$pred_backtransformed <- 10^(NewDataAbundanceMass$pred_log10)
NewDataAbundanceMass$lwr_backtransformed <- 10^(NewDataAbundanceMass$log10_lwr)
NewDataAbundanceMass$upr_backtransformed  <- 10^(NewDataAbundanceMass$log10_upr)


# Extract coefficient in primary
fixef(ab_m1)["log10_mass"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p(parameters::p_value(ab_m1, method = "eti"))

# Extract mass x land use F ratio
anova(ab_m1)
Anova(ab_m1)

# Plot
PlotAbundanceMass <- ggplot(
  NewDataAbundanceMass,
  aes(x = 10^log10_mass,   # back-transform mass for plotting
      y = pred_backtransformed, 
      color = LandUseFinal, 
      fill  = LandUseFinal)
) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(ymin = lwr_backtransformed, ymax = upr_backtransformed),
    alpha = 0.2,
    color = NA
  ) +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_continuous(labels = scales::label_number()) +
  labs(
    x = "Body mass (mg)",
    y = "Abundance of present species (± 95% CI)",
    color = "Land use"
  ) +
  guides(fill = "none") +
  theme_minimal()


FinalPlotAbundanceMass <- ggdraw(PlotAbundanceMass) +
  draw_label("Coefficient in primary: 
  - 0.09 [95% CI: -0.12, -0.06], p < 0.001
              Mass x LU interaction: 
  F-ratio = 0.145, p = 0.93", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

PlotAbundanceMassNoLegend <-  PlotAbundanceMass + theme(legend.position = "none")
  
 FinalPlotAbundanceMassNoLegend  <- ggdraw(PlotAbundanceMassNoLegend) +
  draw_label("Coefficient in primary: 
  - 0.09 [95% CI: -0.12, -0.06], p < 0.001
              Mass x LU interaction: 
  F-ratio = 0.145, p = 0.93", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)


      #ggsave(filename = "./figures/c4_updated/BodyMassAbundanceWhenPresent.png",   plot = FinalPlotAbundanceMass, width = 6, height = 6, dpi = 300)
```

# Modelling Occurrence ~ LUFinal * Mass 

```{r}
# basic data diagnostics
str(ModelDataSizeOccurrence)
table(ModelDataSizeOccurrence$PresentOrAbsent)        # balance of 0/1
table(ModelDataSizeOccurrence$LandUseFinal, useNA="ifany")
n_distinct(ModelDataSizeOccurrence$Source_ID)
n_distinct(ModelDataSizeOccurrence$SS)
n_distinct(ModelDataSizeOccurrence$SpeciesFinal)

# Convert to factors
ModelDataSizeOccurrence$Order <- as.factor(ModelDataSizeOccurrence$Order)
ModelDataSizeOccurrence$Family <- as.factor(ModelDataSizeOccurrence$Family)
ModelDataSizeOccurrence$SpeciesFinal <- as.factor(ModelDataSizeOccurrence$SpeciesFinal)
ModelDataSizeOccurrence$SS <- as.factor(ModelDataSizeOccurrence$SS)
ModelDataSizeOccurrence$SSB <- as.factor(ModelDataSizeOccurrence$SSB)
```

# Backwards stepwise selection of random effects

```{r}
# Just run oc_m1
 oc_m1 <- bglmer(PresentOrAbsent ~ LandUseFinal * log10_mass +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = ModelDataSizeOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "bobyqa",
             optCtrl = list(maxfun = 200000)))

# Put model formulas in a list
model_calls <- list(
  oc_m1 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * log10_mass +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = ModelDataSizeOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "bobyqa",
             optCtrl = list(maxfun = 200000)
           ))
  ),
  oc_m2 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * log10_mass +
             (1 | Source_ID), # Drop species
           data = ModelDataSizeOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "bobyqa",
             optCtrl = list(maxfun = 200000)
           ))
  )
)

# Run models in parallel
results <- mclapply(model_calls, function(expr) eval(expr),
                    mc.cores = 2)  # adjust cores


oc_m1 <- results[[1]]
oc_m2 <- results[[2]]

summary(oc_m1)
summary(oc_m2)

# Compare models
anova(oc_m1, oc_m2)  # Likelihood ratio test
AIC(oc_m1, oc_m2)
```

## Backwards stepwise selection of fixed effects 

```{r}
# Put model formulas in a list

# Additive model
model_calls <- list(
  oc_m3 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal + log10_mass +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = ModelDataSizeOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "bobyqa",
             optCtrl = list(maxfun = 200000)
           ))
  ),
# just LU as predictor
  oc_m4 = quote(
  bglmer(PresentOrAbsent ~ LandUseFinal + 
           (1 | Source_ID) + (1|SpeciesFinal),
         data = ModelDataSizeOccurrence,
         family = binomial(link = "logit"),
         control = lme4::glmerControl(
           optimizer = "bobyqa",
           optCtrl = list(maxfun = 200000)
           ))
  ),
# just Mass as predictor
oc_m5 <- quote(
  bglmer(PresentOrAbsent ~ log10_mass + 
           (1 | Source_ID) + (1|SpeciesFinal),
         data = ModelDataSizeOccurrence,
         family = binomial(link = "logit"),
         control = lme4::glmerControl(
           optimizer = "bobyqa",
           optCtrl = list(maxfun = 200000)
          ))
  ),
# Make a null, intercept only model
oc_m6 <- quote(
  bglmer(PresentOrAbsent ~
           (1 | Source_ID) + (1|SpeciesFinal),
         data = ModelDataSizeOccurrence,
         family = binomial(link = "logit"),
         control = lme4::glmerControl(
           optimizer = "bobyqa",
           optCtrl = list(maxfun = 200000)
         ))
))

# Run models in parallel
results <- mclapply(model_calls, function(expr) eval(expr),
                    mc.cores = 4)  # adjust cores


oc_m3 <- results[[1]]
oc_m4 <- results[[2]]
oc_m5 <- results[[3]]
oc_m6 <- results[[4]]

summary(oc_m3) # additive model - converged
summary(oc_m4) # just LU as predictor - failed to converge
summary(oc_m5) # just mass as predictor - converged
summary(oc_m6) # null intercept only - converged

## Compare the models
AIC(oc_m1, oc_m3, oc_m4, oc_m6)
anova(oc_m1, oc_m3, oc_m4, oc_m6)
anova(oc_m4, oc_m1)
model.sel(oc_m1, oc_m3, oc_m4, oc_m6)

# oc_m1 the best
```

# The interactive model is the best fit, go ahead and generate predictions and create the plot. 

```{r}
# Interaction more significant that just additive effect of mass
anova(oc_m1)

# Create new data frame for predictions
NewDataOccurrenceMass <- expand.grid(
  log10_mass = seq(from = min(ModelDataSizeOccurrence$log10_mass, na.rm = TRUE),
                to   = max(ModelDataSizeOccurrence$log10_mass, na.rm = TRUE),
                length.out = 100),
  LandUseFinal = levels(ModelDataSizeOccurrence$LandUseFinal))

# Copy the factor levels from the model data
NewDataOccurrenceMass$LandUseFinal <- factor(NewDataOccurrenceMass$LandUseFinal, 
                              levels = levels(ModelDataSizeOccurrence$LandUseFinal))

# Predicted probabilities on link scale (logit) with SE
OccurrencePredictions <- predict(oc_m1, newdata = NewDataOccurrenceMass, type = "link", se.fit = TRUE, re.form = NA)

# Add predictions into data frame
NewDataOccurrenceMass <- NewDataOccurrenceMass %>%
  mutate(
    OccurrenceFit   = OccurrencePredictions$fit,
    OccurrenceSE    = OccurrencePredictions$se.fit,
    OccurrenceLwr = OccurrenceFit - 1.96 * OccurrenceSE,
    OccurrenceUpr = OccurrenceFit + 1.96 * OccurrenceSE,
    # Back-transform to probability scale
    OccurrenceProb   = plogis(OccurrenceFit), # predicted probability - plogis() is the inverse of the logit function.
    # It takes values on the logit (link) scale (from −∞ to +∞) and transforms them into probabilities between 0 and 1.
    LowerProb = plogis(OccurrenceLwr), # 95% CI lower bound
    UpperProb = plogis(OccurrenceUpr) # 95% CI upper bound
  ) 

# Extract coefficient in primary
fixef(oc_m1)["log10_mass"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p(parameters::p_value(oc_m1, method = "eti"))

# Extract mass x land use F ratio
anova(oc_m1)
Anova(oc_m1)

# Plot
PlotOccurrenceMass <- ggplot(NewDataOccurrenceMass,
       aes(x = 10^log10_mass, y = OccurrenceProb,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = LowerProb, ymax = UpperProb),
              alpha = 0.2, color = NA) +
  scale_x_log10(labels = scales::label_number()) +
  labs(x = "Body mass (mg)",
       y = "Probability of occurrence (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() 

PlotOccurrenceMassNoLegend <-  PlotOccurrenceMass + theme(legend.position = "none")

FinalPlotOccurrenceMassNoLegend <- ggdraw(PlotOccurrenceMassNoLegend) +
  draw_label("Coefficient in primary: 
  - 0.03 [95% CI: -0.06, -0.00], p = 0.021
              Mass x LU interaction: 
  F-ratio = 10.33, p  < 0.0001 ", x = 0.97, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10) 

# Add the coefficient and interaction text
FinalPlotOccurrenceMass <- ggdraw(PlotOccurrenceMass) +
  draw_label("Coefficient in primary: 
  - 0.03 [95% CI: -0.06, -0.00], p = 0.021
              Mass x LU interaction: 
  F-ratio = 10.33, p < 0.0001 ", x = 0.97, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10) 

       #ggsave(filename = "./figures/c4_updated/BodyMassOccurrence.png",  plot = FinalPlotOccurrenceMass, width = 6, height = 6, dpi = 300)
```

# The occurrence plots show us how likely is the taxa to be present and the abundance plots show us the predicted abundance when the taxa is present. If I multiply the predictions from both plots together - that gives me the predicted abundance of a taxa without knowing in advance whether it is present or not. I will multiply the predictions and calculate the combined standard errors to achieve this. 

# I tried two methods for combining the predictions/standard errors from the occurrence and abundance models. 

# (OLD CODE) - The first was to multiply the fits and use a formula to calculate the standard errors, see below. However the predictions are on different scales. The occcurrence predictions are on the logit link scale, which requires plogis() to backtransform them to get probabilities and the abundance predictions are on the log scale which requires exp() to get abundance counts. Multiply them together would be fitofboth = fit_link * fit_abundance OR logit(probability) * log(abundance) - there's no standard inverse for this and the multiplication of different link scales is not equivalent to multiplying probabilities with counts. See the next chunk for the correct method.

```{r}
# Combine data frames and calculate product and SE
NewDataCombined <- NewDataOccurrenceMass %>%
  bind_cols(NewDataAbundanceMass %>% 
              dplyr::select(fit_abund = fit,
                     se_abund  = se)) %>%
  mutate(
    # product of predictions
    FitOfBoth = fit_link * fit_abund,
    
    # standard error of product (assuming independence)
    SEofBoth  = sqrt((fit_link^2) * (se_abund^2) +
                       (fit_abund^2) * (se_link^2) +
                       (se_link^2) * (se_abund^2)),
    # 95% confidence intervals
    LowerOfBoth = FitOfBoth - 1.96 * SEofBoth,
    UpperOfBoth = FitOfBoth + 1.96 * SEofBoth
  )
  
```


# If my desired outcome is the predicted abundance when accounting for probability of occurrence, it makes more sense to backtransform each prediction to its natural scale first, such that occurrence_prob = plogis(fit_link) and abund_exp = exp(fit_abund). This would lead to occurrence_prob being the probability of occurrence and abund_exp being the predicted abundance. I would then multiply these two components such that pred_abundance = occ_prob * abund_exp and use the delta method to calculate the SEs. 

```{r}
# Backtransform and calculate predicted abundance with CIs
NewDataCombined <- NewDataOccurrenceMass %>%
  bind_cols(NewDataAbundanceMass %>% 
              dplyr::select(
                pred_backtransformed = pred_backtransformed,
                     se_log10  = se_log10)) %>%
  mutate(
    # Probability of occurrence * predicted abundance when present = predicted abundance 
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * pred_backtransformed,  # c = a * b
    # Delta method
    occurrence_var  = (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2), # variance of plogis(fit_link)
    abundance_var = (log(10)^2) * (pred_backtransformed^2) * (se_log10^2), # variance of 10^(fit_abund)

    combined_var =
      (pred_backtransformed^2) * occurrence_var +
      (OccurrenceProb^2) * abundance_var,

    predictedcombinedvar = sqrt(combined_var),

    # 95% CIs on original (abundance) scale
    predictedcombinedlwr =
      PredictedAbundanceAccountingForOccurrenceProbability -
      1.96 * predictedcombinedvar,
    predictedcombinedupr =
      PredictedAbundanceAccountingForOccurrenceProbability +
      1.96 * predictedcombinedvar
  )
    
  
# Plot
FinalPlotAbundanceOccurrenceMassCombined <- ggplot(NewDataCombined,
       aes(x = 10^log10_mass, y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = predictedcombinedlwr, ymax = predictedcombinedupr),
              alpha = 0.2, color = NA) +
  scale_x_log10(labels = scales::label_number()) +
  labs(x = "Body mass (mg)",
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal()


#ggsave(filename = "./figures/c4_updated/BodyMassAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceMassCombined, width = 6, height = 6, dpi = 300)
```

# Combine all three plots 

```{r}
# Combine with shared legend and auto subplot labels
AllThreeTogether <- (FinalPlotAbundanceMassNoLegend | FinalPlotOccurrenceMassNoLegend | FinalPlotAbundanceOccurrenceMassCombined) +
  plot_layout(guides = "collect") &    # combine legends
  plot_annotation(tag_levels = 'a') &  # adds a, b, c tags
  theme(legend.position = "bottom",
        plot.tag = element_text(size = 18, face = "bold"),
                 plot.tag.position = c(0, 1))    # shared legend position

ggsave(filename = "./figures/c4_updated/BodyMassAllThreeCombined.png",  plot = AllThreeTogether, width = 14, height = 7, dpi = 300)


```

# Read in egg data

```{r}
egg_data <- read_xlsx("/Users/jusi/Desktop/multi-groups/data/egg_data.xlsx")

# Create a genus and species column
egg_data <- egg_data %>%
  separate(Binomial_Name, into = c("Genus", "Species"), sep = "_", remove = FALSE, fill = "right")

# Combine genus and species to create Best_guess_binomial which I will use to merge to PREDICTS
egg_data <- egg_data %>% mutate(Best_guess_binomial = str_c(Genus, Species, sep = " ")) %>% dplyr::select(Order, Family, Binomial_Name, Best_guess_binomial, Genus, Species, Volume_mm3, log10Vol)
```

# Read in PREDICTS data and ensure that species level entities such as "Lasioglossum sp." in the Taxon_name_entered column are assigned a genus level egg size value. 

```{r}
# All insect data effort-corrected and merged
taxon_insects <-  qread("/Users/jusi/Desktop/multi-groups/data/predicts-prepared-whole-database.qs")

# Get rid of rows which aren't resolved to order level
taxon_insects <- taxon_insects %>% filter(Order != "") # 1056 rows not resolved to order

# Recreate genus and species column as sometimes BGB is present in PREDICTS but no info for genus or sp.
taxon_insects <- taxon_insects %>%
  mutate(Best_guess_binomial = as.character(Best_guess_binomial)) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "merge", fill = "right")

# There are entries in PREDICTS such as "Lasioglossum sp" - rows with such names tend to not have an entry in the genus column (due to issues with the COL API)
# Previously these taxa were getting family level values in the average mass analysis, despite their genus information being available.
# I will extract the genus information from the Taxon_name_entered column for those rows that do not yet have an entry in genus.
# This will improve the precision of the matching, such that these rows will now be matched at the genus level instead of the family level (as they were beforehand)
taxon_insects <- taxon_insects %>%
  mutate(
    Genus = case_when(
      is.na(Genus) | Genus == "" ~ word(Taxon_name_entered, 1), # Check if Genus is empty or missing, if so, extract the first word from Taxon_name_entered and assign it to genus.
      TRUE ~ Genus # If genus is not empty, leave it as the pre-existing genus entry. 
    ))

# Have a look
taxon_insects %>% dplyr::select(SS, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, Precision_score) %>% View()

# Remove those entries which have a "?" before or after the name of the genus e.g. "?Abalabus" or "Abalabus?"
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove_all(Genus, "\\?"))

# Remove leading and trailing double quotation marks
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_replace_all(Genus, '^"|"$', ""))

# Remove any numbers 
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove_all(Genus, "[0-9]"))

# Remove entries with sp at the end
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove(Genus, "(_|\\s)?sp\\.?$"))

# Filter for columns of interest
taxon_insects <- taxon_insects %>% dplyr::select(Source_ID, SS, SSB, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, Diversity_metric, Diversity_metric_unit, Merged_diversity, Predominant_habitat, Use_intensity)
```

# Start merging egg volume data with PREDICTS

```{r}
# helper: returns the value if all the same, otherwise NA - this sets an NA for log10Vol for those rows with duplicate species entries
all_same_or_na <- function(x) {
  if (length(unique(na.omit(x))) == 1) unique(na.omit(x)) else NA
}

# Calculate mean egg volume for those taxa with duplicate entries
EggDataNoDuplicates <- egg_data %>%
  group_by(Best_guess_binomial, drop = FALSE) %>% # don't drop rows with NA for BGB, i.e. those rows with genus egg size values
  summarise(
    Volume_mm3 = mean(Volume_mm3, na.rm = TRUE),
    across(-Volume_mm3, all_same_or_na),
    .groups = "drop") 

# Get species trait values for merging
sp_value <- EggDataNoDuplicates %>% dplyr::select(Best_guess_binomial, Volume_mm3) 

# First merge species level data
taxon_insects <- left_join(taxon_insects, sp_value, by = "Best_guess_binomial")

rm(sp_value)

# Number of species matched
taxon_insects %>%
  filter(!is.na(Volume_mm3)) %>% 
  group_by(Order) %>%
  summarise(n_species = n_distinct(Best_guess_binomial)) %>%
  arrange(desc(n_species))  

# Calculate average egg volume for genera
genus_value <- EggDataNoDuplicates %>%
  group_by(Genus) %>%
  summarise(GenusAvgEggVolume_mm3 = mean(Volume_mm3, na.rm = TRUE)) # 3322 genus data

# Merge based on genus where best_guess_binomial is empty
taxon_insects <- taxon_insects %>%
  left_join(genus_value, by = "Genus", suffix = c("", "_genus")) 

# Average egg volume per family based on species values
family_value <- EggDataNoDuplicates %>%
  group_by(Family) %>%
  summarise(FamilyAvgEggVolume_mm3 = mean(Volume_mm3, na.rm = TRUE))

# Merge based on family where genus and species is empty
taxon_insects <- taxon_insects %>%
  left_join(family_value, by = "Family", suffix = c("", "_family")) 

# priority-based fallback assignment where final egg volume values are selected based on the first non-missing (or non-blank) column in a defined order.
taxon_insects <- taxon_insects %>%
  mutate(FinalAvgEggVolume_mm3 = coalesce(
    Volume_mm3,
    GenusAvgEggVolume_mm3,
    FamilyAvgEggVolume_mm3
  ))

# How many rows have data based on each level of taxonomic resolution
taxon_insects %>%
  summarise(across(
    c(Volume_mm3,
    GenusAvgEggVolume_mm3,
    FamilyAvgEggVolume_mm3,
    FinalAvgEggVolume_mm3),
    list(
      non_NAs = ~sum(!is.na(.))
    )
  ))
```

# There are some rows in the egg volume dataset which give an egg volume value for the name of a genus rather than a species (i.e. a full binomial). I have already matched egg volume values using the species values, the average of species within the same genus and the average of genera within the same family. I will now extract the egg volume values for those rows in the dataset which are ONLY for genera. I will then merge the egg volume values for those genera to taxa in PREDICTS which do not currently have a genus level egg volume value, I will then do the same for those rows resolved to family (i.e. calculate the average egg volume for the family based on the egg volume data JUST for genera)

# First lets have a look at those rows which are just values for genera..

```{r}
# Which rows are egg volume values for taxa with both a genus and a species name, i.e. a full binomial
with_binomial <- egg_data %>%
  filter(!is.na(Best_guess_binomial)) %>%
  group_by(Best_guess_binomial) %>%
  summarise(
    Volume_mm3 = mean(Volume_mm3, na.rm = TRUE),
    across(-Volume_mm3, all_same_or_na),
    .groups = "drop") %>% 
  dplyr::select(Order, Family, Binomial_Name, Best_guess_binomial, Genus, Species, Volume_mm3, log10Vol)

# Filter for rows that do not have a species name and are therefore egg volume values for genera
without_binomial <- egg_data %>%
  filter(is.na(Best_guess_binomial))

# Which genera have I not already used in the previous chunk to match to species in predicts (i.e. are not in the with_binomial df) and therefore will be useful to match to taxa in PREDICTS without an egg volume value already..
UsefulGenera <- without_binomial %>%
  anti_join(with_binomial, by = "Genus") 
```

# I will now match those egg volume values for genera which I have not yet matched to taxa in PREDICTS (because the value was only for the name of a genus rather than a full latin binomial)
# I will do this only for those taxa in PREDICTS which do not yet have an egg volume value at the species, genus or family level (that is if they already have an egg volume value at the species level, there would be no point assigning a genus or family level value)..

```{r}
# First lets calculate the mean egg volume for duplicate genus entries 
UsefulGeneraNoDuplicates <- UsefulGenera %>%
  group_by(Genus, drop = FALSE) %>% # don't drop NAs, i.e. those rows with genus egg size values
  summarise(
    GenusAvgEggVolume_mm3 = mean(Volume_mm3, na.rm = TRUE),
    across(-Volume_mm3, all_same_or_na),
    .groups = "drop") %>% 
  dplyr::select(Order, Family, Binomial_Name, Best_guess_binomial, Genus, Species, GenusAvgEggVolume_mm3, log10Vol)

# Simplify for merging
DiffGenusAvg <- UsefulGeneraNoDuplicates %>% 
  dplyr::select(Genus, GenusAvgEggVolume_mm3)

# Step 1: Filter rows in test without an egg volume value at the species, genus or family level and therefore require an egg volume value
NeedsValue <- taxon_insects %>%
  filter(is.na(Volume_mm3) & is.na(GenusAvgEggVolume_mm3) & is.na(FamilyAvgEggVolume_mm3))

# Step 2: Join genus egg volume values for those genera we have not captured yet
NeedsValue <- NeedsValue %>%
  dplyr::select(-GenusAvgEggVolume_mm3) %>%   # drop old column to avoid duplicate
  left_join(DiffGenusAvg, by = "Genus")

# Average egg volume for families
DiffFamilyAvg <- UsefulGeneraNoDuplicates %>%
  group_by(Family) %>%
  summarise(FamilyAvgEggVolume_mm3 = mean(GenusAvgEggVolume_mm3, na.rm = TRUE))

# Merge based on family where genus and species is empty
NeedsValue <- NeedsValue %>%
  dplyr::select(-FamilyAvgEggVolume_mm3) %>% 
  left_join(DiffFamilyAvg, by = "Family", suffix = c("", "_family")) 

# Combine updated rows (NeedsValue) back into original data (taxon_insects)
# First, remove the rows you updated from test
remaining_rows <- taxon_insects %>%
  filter(!(is.na(Volume_mm3) & is.na(GenusAvgEggVolume_mm3) & is.na(FamilyAvgEggVolume_mm3)))

# sense check remaining_rows should be taxon_insects - NeedsValue
nrow(taxon_insects) - nrow(NeedsValue)

# Bind them together
taxon_insects_updated <- bind_rows(remaining_rows, NeedsValue)

# How many rows still without a family average
sum(is.na(taxon_insects_updated$FamilyAvgEggVolume_mm3)) # 167353
sum(is.na(taxon_insects$FamilyAvgEggVolume_mm3)) # 178615

# 178615 - 167353 = 11262 rows which got a family value from the additional genera
```

# Now that we have merged all the egg volume data, coalesce the additional genus and family averages into the FinalAvgEggVolume_mm3 column

```{r}
# Replace those rows with missing values in FinalAvgEggVolume_mm3 with the corresponding FamilyAvgEggVolume_mm3
taxon_insects_updated <- taxon_insects_updated %>%
  mutate(FinalAvgEggVolume_mm3 = coalesce(FinalAvgEggVolume_mm3, FamilyAvgEggVolume_mm3)) # coalesce(a,b) returns a if its not NA otherwise it returns b

# How many rows do we still not have a egg volume value for and are still NA - 153284
taxon_insects_updated %>%
  filter(is.na(Volume_mm3) & is.na(GenusAvgEggVolume_mm3) & is.na(FamilyAvgEggVolume_mm3) & is.na(FinalAvgEggVolume_mm3)) %>%
  nrow()

# Have a look at those rows
taxon_insects_updated %>%
  filter(is.na(Volume_mm3) & is.na(GenusAvgEggVolume_mm3) & is.na(FamilyAvgEggVolume_mm3) & is.na(FinalAvgEggVolume_mm3)) %>%
  View()

# Filter out rows which still don't have a egg volume value
AllInsectsEggSize <- taxon_insects_updated %>%
  filter(!is.na(FinalAvgEggVolume_mm3)) %>% 
  droplevels()

rm(DiffFamilyAvg, DiffGenusAvg, family_value, genus_value, NeedsValue, remaining_rows, UsefulGenera, UsefulGeneraNoDuplicates, with_binomial, without_binomial,
   EggDataNoDuplicates, egg_data, taxon_insects_updated, taxon_insects)
```

# Now that I have an overall data frame for egg size for all insect taxa in PREDICTS, I will now filter it for species level entities using APs taxon resolution file and then prepare the data frame for analysis.

```{r}
# Filter for diversity metrics of interest - abundance and occurrence data can be used (i.e. no species richness data)
AllInsectsEggSize <- AllInsectsEggSize %>% filter(Diversity_metric %in% c("abundance", "effort-corrected abundance", "density", "group abundance", "occurrence", "occurrence frequency", "relative abundance"))

# Load taxon resolution file
TaxonResolution <- readRDS(here("data", "CuratedTaxonResolution2.rds"))

# Step 1: Filter TaxonResolution for species-level entries - i.e. those entries with a precision score of 4!
TaxonResolutionSp <- TaxonResolution %>%
  filter(PrecisionScore == 4) %>%
  dplyr::select(SS, Taxon_name_entered) # select cols for merging

# Step 2: Keep only matching rows in AllInsectsEggSize
AllInsectsEggSizeSpeciesOnly <- AllInsectsEggSize %>%
  semi_join(TaxonResolutionSp, by = c("SS", "Taxon_name_entered")) # keeps only rows in AllInsectsEggSize that have a match in TaxonResolutionSp. based on SS and Taxon_name_entered

# AllInsectsEggSizeSpeciesOnly now contains only the rows from AllInsectsEggSize that correspond to species-level entries in TaxonResolutionSp. 

# Lets have a look at those rows which have an entry for Taxon_name_entered but no entry for Best_guess_binomial!
AllInsectsEggSizeSpeciesOnly %>%
  filter(
    !is.na(Taxon_name_entered) & Taxon_name_entered != "",
    is.na(Best_guess_binomial) | Best_guess_binomial == ""
  ) %>% View()

# Create a present or absent column - its present if Merged_diversity > 0 and absent if its 0
AllInsectsEggSizeSpeciesOnly <- AllInsectsEggSizeSpeciesOnly %>%
  mutate(PresentOrAbsent = ifelse(Merged_diversity > 0, 1, 0))

# Create a log(abundance when present) column
AllInsectsEggSizeSpeciesOnly <- AllInsectsEggSizeSpeciesOnly %>% 
  mutate(
    LogAbundanceWhenPresent = ifelse(Merged_diversity > 0,
                                    log(Merged_diversity),
                                    NA))

# Create a log(egg size) column
AllInsectsEggSizeSpeciesOnly <- AllInsectsEggSizeSpeciesOnly %>% 
  mutate(
    LogAvgEggVolume_mm3 = log(FinalAvgEggVolume_mm3))


# Have a look at how many rows are present versus absent
table(AllInsectsEggSizeSpeciesOnly$PresentOrAbsent)

# Rename predominant habitat since its not really habitat we're looking at
AllInsectsEggSizeSpeciesOnly <- rename(AllInsectsEggSizeSpeciesOnly,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
AllInsectsEggSizeSpeciesOnly <- AllInsectsEggSizeSpeciesOnly %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(AllInsectsEggSizeSpeciesOnly$Predominant_land_use)
table(AllInsectsEggSizeSpeciesOnly$Predominant_land_use, AllInsectsEggSizeSpeciesOnly$Use_intensity)
```

# Summary statistics for egg volume

```{r}
# How many species matched to species-level data
AllInsectsEggSizeSpeciesOnly %>%
  filter(!is.na(Volume_mm3)) %>%
  distinct(Best_guess_binomial) %>%
  nrow()

# Genus-level data
AllInsectsEggSizeSpeciesOnly %>%
  filter(!is.na(GenusAvgEggVolume_mm3)) %>%
  distinct(Genus) %>%
  nrow()

# How many families matched to family level data
AllInsectsEggSizeSpeciesOnly %>%
  filter(!is.na(FamilyAvgEggVolume_mm3)) %>%
  distinct(Family) %>%
  nrow()


# The number of species matched to species-, genus- and family- level trait data
length(unique(AllInsectsEggSizeSpeciesOnly$Best_guess_binomial[!is.na(AllInsectsEggSizeSpeciesOnly$Volume_mm3)]))
length(unique(AllInsectsEggSizeSpeciesOnly$Best_guess_binomial[!is.na(AllInsectsEggSizeSpeciesOnly$GenusAvgEggVolume_mm3)]))
length(unique(AllInsectsEggSizeSpeciesOnly$Best_guess_binomial[!is.na(AllInsectsEggSizeSpeciesOnly$FamilyAvgEggVolume_mm3)]))
length(unique(AllInsectsEggSizeSpeciesOnly$Best_guess_binomial))
length(unique(AllInsectsEggSizeSpeciesOnly$Family))


AllInsectsEggSizeSpeciesOnly %>% dplyr::select(Best_guess_binomial, Genus, Family, Volume_mm3, GenusAvgEggVolume_mm3, FamilyAvgEggVolume_mm3) %>% distinct() %>% View()
```

# Recode the land use classes to have Primary, Secondary vegetation, Low-intensity agriculture and high-intensity agriculture

```{r}

AllInsectsEggSizeSpeciesOnly <- AllInsectsEggSizeSpeciesOnly %>%
  mutate(
    LandUseFinal = case_when(
      # Primary stays as is
      Predominant_land_use == "Primary" ~ "Primary",

      # Any secondary vegetation type -> Secondary Vegetation
      Predominant_land_use %in% c("Young secondary vegetation",
                                  "Intermediate secondary vegetation",
                                  "Mature secondary vegetation") ~ "Secondary vegetation",

      # Plantation forest depends on intensity
      Predominant_land_use == "Plantation forest" & Use_intensity == "Minimal use" ~ "Low-intensity agriculture",
      Predominant_land_use == "Plantation forest" & Use_intensity %in% c("Light use", "Intense use") ~ "High-intensity agriculture",

      # Pasture
      Predominant_land_use == "Pasture" & Use_intensity %in% c("Minimal use", "Light use") ~ "Low-intensity agriculture",
      Predominant_land_use == "Pasture" & Use_intensity == "Intense use" ~ "High-intensity agriculture",

      # Cropland
      Predominant_land_use == "Cropland" & Use_intensity %in% c("Minimal use", "Light use") ~ "Low-intensity agriculture",
      Predominant_land_use == "Cropland" & Use_intensity == "Intense use" ~ "High-intensity agriculture",

      # Drop Urban by assigning NA -these will get dropped when we create the model data
      Predominant_land_use == "Urban" ~ NA_character_
    )
  ) %>%
  # Turn into factor if you want ordered categories
  mutate(LandUseFinal = factor(LandUseFinal,
                               levels = c("Primary", "Secondary vegetation",
                                          "Low-intensity agriculture", "High-intensity agriculture")))

table(AllInsectsEggSizeSpeciesOnly$LandUseFinal)
```

# Create random intercept for Species

```{r}
AllInsectsEggSizeSpeciesOnly <- AllInsectsEggSizeSpeciesOnly %>%
  mutate(
    SpeciesFinal = case_when(
      Species != "" & !is.na(Species) ~ Species,
      (Species == "" | is.na(Species)) & Best_guess_binomial != "" & !is.na(Best_guess_binomial) ~ Best_guess_binomial,
      (Species == "" | is.na(Species)) & (Best_guess_binomial == "" | is.na(Best_guess_binomial)) ~ paste(SS, Taxon_name_entered, sep = " "),
      TRUE ~ NA_character_
    )
  )

sum(is.na(AllInsectsEggSizeSpeciesOnly$SpeciesFinal))
AllInsectsEggSizeSpeciesOnly %>% filter(SpeciesFinal == "")
sum(is.na(AllInsectsEggSizeSpeciesOnly$SpeciesFinal) | AllInsectsEggSizeSpeciesOnly$SpeciesFinal == "")


length(unique(AllInsectsEggSizeSpeciesOnly$SpeciesFinal))
length(unique(AllInsectsEggSizeSpeciesOnly$Taxon_name_entered))
length(unique(AllInsectsEggSizeSpeciesOnly$Best_guess_binomial))
length(unique(AllInsectsEggSizeSpeciesOnly$Species))


AllInsectsEggSizeSpeciesOnly %>% dplyr::select(SS, Order, Family, Genus, Species, Best_guess_binomial, Taxon_name_entered, SpeciesFinal) %>% View()
```

# Create model data

```{r}
# drop all rows with NAs for variables of interest
ModelDataEggOccurrence <- drop_na(AllInsectsEggSizeSpeciesOnly, LandUseFinal, Use_intensity, PresentOrAbsent) %>% droplevels()
ModelDataEggAbundance <- drop_na(AllInsectsEggSizeSpeciesOnly, LandUseFinal, Use_intensity, LogAbundanceWhenPresent) %>% droplevels()

length(unique(ModelDataEggAbundance$SS)) # 392 studies 
length(unique(ModelDataEggAbundance$SSBS)) # 8886 sites

length(unique(ModelDataEggOccurrence$SS)) # 392 studies 
length(unique(ModelDataEggOccurrence$SSBS)) # 9670 sites


#saveRDS(ModelDataEggOccurrence, "./C4 data/ModelDataEggOccurrence.rds")
#saveRDS(ModelDataEggAbundance, "./C4 data/ModelDataEggAbundance.rds")
rm(TaxonResolution, TaxonResolutionSp)
```

#### Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(ModelDataEggAbundance[ , c("LandUseFinal", "Use_intensity", "LogAvgEggVolume_mm3", "LogAbundanceWhenPresent")])
corvif(ModelDataEggOccurrence[ , c("LandUseFinal", "Use_intensity","LogAvgEggVolume_mm3", "PresentOrAbsent")])
```

# Modelling log(abundance of present species) ~ LUFinal * Egg Size

```{r}
# basic data diagnostics
str(ModelDataEggAbundance)
table(ModelDataEggAbundance$LandUseFinal, useNA="ifany")
n_distinct(ModelDataEggAbundance$Source_ID)
n_distinct(ModelDataEggAbundance$SS)
n_distinct(ModelDataEggAbundance$SpeciesFinal)

# Convert to factors
ModelDataEggAbundance$Order <- as.factor(ModelDataEggAbundance$Order)
ModelDataEggAbundance$Family <- as.factor(ModelDataEggAbundance$Family)
ModelDataEggAbundance$SpeciesFinal <- as.factor(ModelDataEggAbundance$SpeciesFinal)
ModelDataEggAbundance$SS <- as.factor(ModelDataEggAbundance$SS)
ModelDataEggAbundance$SSB <- as.factor(ModelDataEggAbundance$SSB)
```

# Maximal random effects structure

```{r}
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * LogAvgEggVolume_mm3 + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = ModelDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop block
ab_m2 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * LogAvgEggVolume_mm3 + (1 | Source_ID/SS) + (1| Family/SpeciesFinal), 
               data = ModelDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop SS
ab_m3 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * LogAvgEggVolume_mm3 + (1 | Source_ID) + (1| Family/SpeciesFinal), 
               data = ModelDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


# Drop Family
ab_m4 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * LogAvgEggVolume_mm3 + (1 | Source_ID) + (1 | SpeciesFinal), 
               data = ModelDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Compare models
anova(ab_m1, ab_m2, ab_m3, ab_m4)  # Likelihood ratio test
AIC(ab_m1, ab_m2, ab_m3, ab_m4)


# ab_m1 best!
```

# Backwards stepwise selection of fixed effects

```{r}
# Additive model with the interaction removed but both predictors still present
ab_m5 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + LogAvgEggVolume_mm3 + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


# Just land use as predictor
ab_m6 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Just egg as the predictor
ab_m7 <- blmer(LogAbundanceWhenPresent ~ LogAvgEggVolume_mm3 + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Make a null, intercept only model
ab_m8 <- blmer(LogAbundanceWhenPresent ~ (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


## Compare the models
AIC(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m7, ab_m5, ab_m1)
```

# Interaction model only marginally better (p = 0.046) than additive model. AIC difference small, only 2 points.  
# Generate predictions from the model

```{r}
# Create a grid of values
NewDataAbundanceEggSize <- expand.grid(
  LogAvgEggVolume_mm3 = seq(from = min(ModelDataEggAbundance$LogAvgEggVolume_mm3),
                   to = max(ModelDataEggAbundance$LogAvgEggVolume_mm3),
                   length.out = 100),
  LandUseFinal = levels(ModelDataEggAbundance$LandUseFinal))
  

# Get predictions without random effects, with SEs
AbundancePredictions <- predict(ab_m1, newdata = NewDataAbundanceEggSize, re.form = NA, se.fit = TRUE)

# Predictions on log scale
NewDataAbundanceEggSize$pred_log <- AbundancePredictions$fit
NewDataAbundanceEggSize$se_log   <- AbundancePredictions$se.fit

# 95% CI on log scale
value <- qnorm(0.975)
NewDataAbundanceEggSize$log_lwr <- NewDataAbundanceEggSize$pred_log -
                                  value * NewDataAbundanceEggSize$se_log
NewDataAbundanceEggSize$log_upr <- NewDataAbundanceEggSize$pred_log +
                                  value * NewDataAbundanceEggSize$se_log

# Back-transform all of them
NewDataAbundanceEggSize$pred_backtransformed <- exp(NewDataAbundanceEggSize$pred_log)
NewDataAbundanceEggSize$lwr_backtransformed <- exp(NewDataAbundanceEggSize$log_lwr)
NewDataAbundanceEggSize$upr_backtransformed  <- exp(NewDataAbundanceEggSize$log_upr)


# Extract coefficient in primary
fixef(ab_m1)["LogAvgEggVolume_mm3"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p(parameters::p_value(ab_m1, method = "eti"))

# Extract mass x land use F ratio
anova(ab_m1)
Anova(ab_m1)

# Create a back-transformed egg volume column for plotting
NewDataAbundanceEggSize <- NewDataAbundanceEggSize %>%
  mutate(EggVolume_mm3 = exp(LogAvgEggVolume_mm3))

# Plot
PlotAbundanceEggSize <- ggplot(NewDataAbundanceEggSize,
       aes(x = EggVolume_mm3, y = pred_backtransformed,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lwr_backtransformed, ymax = upr_backtransformed),
              alpha = 0.2, color = NA) +
  scale_x_log10(
    labels = scales::label_number()) +
  scale_y_continuous(labels = scales::label_number()) +
  labs(x = expression("Egg volume (mm"^3*")"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal()



# Remove legend
PlotAbundanceEggSizeNoLegend <- PlotAbundanceEggSize + theme(legend.position = "none")

# Add text box
FinalPlotAbundanceEggSizeNoLegend <- ggdraw(PlotAbundanceEggSizeNoLegend) +
  draw_label("Coefficient in primary: 
  - 0.02 [95% CI: -0.03, -0.01], p = 0.005
              Egg volume x LU interaction: 
  F-ratio = 2.67, p = 0.05", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)


FinalPlotAbundanceEggSize <- ggdraw(PlotAbundanceEggSize) +
  draw_label("Coefficient in primary: 
  - 0.02 [95% CI: -0.03, -0.01], p = 0.005
              Egg volume x LU interaction: 
  F-ratio = 2.67, p = 0.05", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

       #ggsave(filename = "./figures/c4_updated/EggSizeAbundanceWhenPresent.png",  plot = FinalPlotAbundanceEggSize, width = 6, height = 6, dpi = 300)
```

# Modelling AbundanceWhenPresent ~ LUFinal * egg size * order

```{r}
# Filter for big five
BigFiveModeDataEggAbundance <- ModelDataEggAbundance %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))

# Maximal model
BigFiveAb_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * LogAvgEggVolume_mm3 * Order + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = BigFiveModeDataEggAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Check the model
model_plot(BigFiveAb_m1)
summary(BigFiveAb_m1)
anova(BigFiveAb_m1)
Anova(BigFiveAb_m1)


# Calculate the egg volume range for predictions
egg_range <- range(BigFiveModeDataEggAbundance$LogAvgEggVolume_mm3, na.rm = TRUE)

# Generate predictions
# How does abundance scale with mass for each LU * Order combo
BigFiveAbundancePredictions <- ggpredict(
  BigFiveAb_m1,
  terms = c(
    sprintf("LogAvgEggVolume_mm3 [%.2f:%.2f]", egg_range[1], egg_range[2]), # Generate predictions using that range
    "LandUseFinal",
    "Order"
  ))

# Plot
PlotBigFiveAbundanceEggVolume <- ggplot(BigFiveAbundancePredictions, aes(x = exp(x), y = predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.15, color = NA) +
  facet_wrap(~ facet, scales = "free_y") +
  scale_x_continuous(trans = "log", labels = scales::label_number(accuracy = 0.01)) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))  +
  labs(x = expression("Egg volume (mm"^3*")"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") 


#ggsave(filename = "./figures/AbundanceWhenPresentBigFiveEggVolume.png",   plot = PlotBigFiveAbundanceEggVolume, width = 10, height = 6, dpi = 300)

rm(BigFiveAb_m1)
```

# Modelling PresenceOrAbsent ~ LUFinal * Egg Size 

```{r}
# basic data diagnostics
str(ModelDataEggOccurrence)
table(ModelDataEggOccurrence$PresentOrAbsent)        # balance of 0/1
table(ModelDataEggOccurrence$LandUseFinal, useNA="ifany")
n_distinct(ModelDataEggOccurrence$Source_ID)
n_distinct(ModelDataEggOccurrence$SS)
n_distinct(ModelDataEggOccurrence$SpeciesFinal)

# Convert to factors
ModelDataEggOccurrence$Order <- as.factor(ModelDataEggOccurrence$Order)
ModelDataEggOccurrence$Family <- as.factor(ModelDataEggOccurrence$Family)
ModelDataEggOccurrence$SpeciesFinal <- as.factor(ModelDataEggOccurrence$SpeciesFinal)
ModelDataEggOccurrence$SS <- as.factor(ModelDataEggOccurrence$SS)
ModelDataEggOccurrence$SSB <- as.factor(ModelDataEggOccurrence$SSB)

```

# Scale egg volume to help with convergence
```{r}
ModelDataEggOccurrence$ScaledLogAvgEggVolume_mm3 <- scale(ModelDataEggOccurrence$LogAvgEggVolume_mm3, center = TRUE, scale = TRUE)

hist(ModelDataEggOccurrence$ScaledLogAvgEggVolume_mm3) ; hist(ModelDataEggOccurrence$LogAvgEggVolume_mm3)
```

# Backwards stepwise selection of random effects

```{r}
# Just run the model
  oc_m1 <- bglmer(PresentOrAbsent ~ LandUseFinal * ScaledLogAvgEggVolume_mm3 +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = ModelDataEggOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "Nelder_Mead",
             optCtrl = list(maxfun = 2e5)))

# Save the model
  #saveRDS(oc_m1, file = "EggVolumeOccurrenceModel.rds")

# Put model formulas in a list
RandomEffectsModelCalls <- list(
  oc_m1 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledLogAvgEggVolume_mm3 +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = ModelDataEggOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "Nelder_Mead",
             optCtrl = list(maxfun = 2e5)
           ))
  ),
  oc_m2 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledLogAvgEggVolume_mm3 +
             (1 | Source_ID), # Drop species
           data = ModelDataEggOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "Nelder_Mead",
             optCtrl = list(maxfun = 2e5)
           ))
  )
)

# Run models in parallel
RandomResults <- mclapply(RandomEffectsModelCalls, function(expr) eval(expr),
                    mc.cores = 2)  # adjust cores


oc_m1 <- RandomResults[[1]]
oc_m2 <- RandomResults[[2]]

summary(oc_m1)
summary(oc_m2)

# Compare models
anova(oc_m1, oc_m2)  # Likelihood ratio test
AIC(oc_m1, oc_m2)


```

## Backwards stepwise selection of fixed effects 

```{r}
# Put model formulas in a list

# Additive model
FixedEffectsModelCalls <- list(
  oc_m3 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal + ScaledLogAvgEggVolume_mm3 +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = ModelDataEggOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "Nelder_Mead",
             optCtrl = list(maxfun = 2e5)
           ))
  ),
# just LU as predictor
  oc_m4 = quote(
  bglmer(PresentOrAbsent ~ LandUseFinal + 
           (1 | Source_ID) + (1|SpeciesFinal),
         data = ModelDataEggOccurrence,
         family = binomial(link = "logit"),
         control = lme4::glmerControl(
           optimizer = "Nelder_Mead",
           optCtrl = list(maxfun = 2e5)
           ))
  ),
# just egg size as predictor
oc_m5 <- quote(
  bglmer(PresentOrAbsent ~ ScaledLogAvgEggVolume_mm3 + 
           (1 | Source_ID) + (1|SpeciesFinal),
         data = ModelDataEggOccurrence,
         family = binomial(link = "logit"),
         control = lme4::glmerControl(
           optimizer = "Nelder_Mead",
           optCtrl = list(maxfun = 2e5)
          ))
  ),
# Make a null, intercept only model
oc_m6 <- quote(
  bglmer(PresentOrAbsent ~
           (1 | Source_ID) + (1|SpeciesFinal),
         data = ModelDataEggOccurrence,
         family = binomial(link = "logit"),
         control = lme4::glmerControl(
           optimizer = "Nelder_Mead",
           optCtrl = list(maxfun = 2e5)
         ))
))

# Run models in parallel
FixedResults <- mclapply(FixedEffectsModelCalls, function(expr) eval(expr),
                    mc.cores = 4)  # adjust cores


oc_m3 <- FixedResults[[1]]
oc_m4 <- FixedResults[[2]]
oc_m5 <- FixedResults[[3]]
oc_m6 <- FixedResults[[4]]
```

# Check model convergence and compare
```{r}
summary(oc_m3) # additive model - converged
summary(oc_m4) # just LU as predictor - converged
summary(oc_m5) # just egg size as predictor - converged
summary(oc_m6) # null intercept only - converged

## Compare the models
AIC(oc_m1, oc_m3, oc_m4, oc_m5, oc_m6)
anova(oc_m1, oc_m3, oc_m4, oc_m5, oc_m6)

# oc_m1 the best
```

# The interactive model is the best fit, go ahead and generate predictions and create the plot. 

```{r}
# Interaction more significant that just additive effect of mass
anova(oc_m1)

# Create new data frame for predictions
NewDataOccurrenceEggSize <- expand.grid(
  ScaledLogAvgEggVolume_mm3 = seq(from = min(ModelDataEggOccurrence$ScaledLogAvgEggVolume_mm3, na.rm = TRUE),
                to   = max(ModelDataEggOccurrence$ScaledLogAvgEggVolume_mm3, na.rm = TRUE),
                length.out = 100),
  LandUseFinal = levels(ModelDataEggAbundance$LandUseFinal))

# Copy the factor levels from the model data
NewDataOccurrenceEggSize$LandUseFinal <- factor(NewDataOccurrenceEggSize$LandUseFinal, 
                              levels = levels(ModelDataEggOccurrence$LandUseFinal))

# Predicted probabilities on link scale (logit) with SE
OccurrencePredictions <- predict(oc_m1, newdata = NewDataOccurrenceEggSize, type = "link", se.fit = TRUE, re.form = NA)
       
# Add predictions into data frame
NewDataOccurrenceEggSize <- NewDataOccurrenceEggSize %>%
  mutate(
    OccurrenceFit   = OccurrencePredictions$fit,
    OccurrenceSE    = OccurrencePredictions$se.fit,
    OccurrenceLwr = OccurrenceFit - 1.96 * OccurrenceSE,
    OccurrenceUpr = OccurrenceFit + 1.96 * OccurrenceSE,
    # Back-transform to probability scale
    OccurrenceProb   = plogis(OccurrenceFit), # predicted probability - plogis() is the inverse of the logit function.
    # It takes values on the logit (link) scale (from −∞ to +∞) and transforms them into probabilities between 0 and 1.
    LowerProb = plogis(OccurrenceLwr), # 95% CI lower bound
    UpperProb = plogis(OccurrenceUpr) # 95% CI upper bound
  )

# Extract coefficient in primary
fixef(oc_m1)["ScaledLogAvgEggVolume_mm3"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p(parameters::p_value(oc_m1, method = "eti"))

# Extract mass x land use F ratio
anova(oc_m1)
Anova(oc_m1)

# Extract mean and sd from scaled variable to backtransform
log_mean <- attr(ModelDataEggOccurrence$ScaledLogAvgEggVolume_mm3, "scaled:center")
log_sd   <- attr(ModelDataEggOccurrence$ScaledLogAvgEggVolume_mm3, "scaled:scale")

# Add a new column with the back transformed values:
NewDataOccurrenceEggSize$BackTransformedEggVolume_mm3 <- exp(NewDataOccurrenceEggSize$ScaledLogAvgEggVolume_mm3 * log_sd + log_mean)

# Plot
PlotOccurrenceEggSize <- ggplot(NewDataOccurrenceEggSize,
       aes(x = BackTransformedEggVolume_mm3, y = OccurrenceProb,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = LowerProb, ymax = UpperProb),
              alpha = 0.2, color = NA) +
  scale_x_log10(
    labels = scales::label_number()
  ) +
  labs(x = expression("Egg volume (mm"^3*")"),
       y = "Probability of occurrence (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal()

# Remove legend
PlotOccurrenceEggSizeNoLegend <- PlotOccurrenceEggSize + theme(legend.position = "none")

# Add text box
FinalPlotOccurrenceEggSizeNoLegend <- ggdraw(PlotOccurrenceEggSizeNoLegend) +
  draw_label("Coefficient in primary: 
   0.06 [95% CI: 0.04, 0.09], p < 0.001
              Egg volume x LU interaction: 
  F-ratio = 31.53, p < 0.0001", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

# Plot just for saving on its own
FinalPlotOccurrenceEggSize <- ggdraw(PlotOccurrenceEggSize) +
  draw_label("Coefficient in primary: 
   0.06 [95% CI: 0.04, 0.09], p < 0.001
              Egg volume x LU interaction:
  F-ratio = 31.53, p < 0.0001", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

       #ggsave(filename = "./figures/c4_updated/EggSizeOccurrence.png",  plot = FinalPlotOccurrenceEggSize, width = 6, height = 6, dpi = 300)

FinalPlotAbundanceEggSize
```

# If my desired outcome is the predicted abundance when accounting for probability of occurrence, it makes more sense to backtransform each prediction to its natural scale first, such that occurrence_prob = plogis(fit_link) and abund_exp = exp(fit_abund). This would lead to occurrence_prob being the probability of occurrence and abund_exp being the predicted abundance. I would then multiply these two components such that pred_abundance = occ_prob * abund_exp and use the delta method to calculate the SEs. 

```{r}
# Backtransform and calculate predicted abundance with CIs
NewDataCombined <- NewDataOccurrenceEggSize %>%
  bind_cols(NewDataAbundanceEggSize %>% 
              dplyr::select(
                pred_backtransformed = pred_backtransformed,
                     se_log  = se_log)) %>%
  mutate(
    # Probability of occurrence * predicted abundance when present = predicted abundance 
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * pred_backtransformed,  # c = a * b
    # Delta method
    occurrence_var  = (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2), # variance of plogis(fit_link)
    abundance_var = (pred_backtransformed^2) * (se_log^2), # variance of exp(fit_abund)

    combined_var =
      (pred_backtransformed^2) * occurrence_var +
      (OccurrenceProb^2) * abundance_var,

    predictedcombinedvar = sqrt(combined_var),

    # 95% CIs on original (abundance) scale
    predictedcombinedlwr =
      PredictedAbundanceAccountingForOccurrenceProbability -
      1.96 * predictedcombinedvar,
    predictedcombinedupr =
      PredictedAbundanceAccountingForOccurrenceProbability +
      1.96 * predictedcombinedvar
  )
    
  
# Plot
FinalPlotAbundanceOccurrenceEggSizeCombined <- ggplot(NewDataCombined,
       aes(x = BackTransformedEggVolume_mm3, y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = predictedcombinedlwr, ymax = predictedcombinedupr),
              alpha = 0.2, color = NA) +
  scale_x_log10(labels = scales::label_number()) +
  labs(x = expression("Egg volume (mm"^3*")"),
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal()

#ggsave(filename = "./figures/c4_updated/EggSizeAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceEggSizeCombined, width = 6, height = 6, dpi = 300)
```

# Combine all three plots 

```{r}
# Combine with shared legend and auto subplot labels
AllThreeTogether <- (FinalPlotAbundanceEggSizeNoLegend | FinalPlotOccurrenceEggSizeNoLegend | FinalPlotAbundanceOccurrenceEggSizeCombined) +
  plot_layout(guides = "collect") &    # combine legends
  plot_annotation(tag_levels = 'a') &  # adds a, b, c tags
  theme(legend.position = "bottom",
        plot.tag = element_text(size = 18, face = "bold"),
                 plot.tag.position = c(0, 1))    # shared legend position

ggsave(filename = "./figures/c4_updated/EggSizeAllThreeCombined.png",  plot = AllThreeTogether, width = 14, height = 7, dpi = 300)
```

# Read and tidy overall trait dataset
```{r}
# Read in trait data
traits <- read_excel("data/All_FamilyandOrder_Traits_for_JustinIsip.xlsx") 

# As data frame 
traits <- as.data.frame(traits)

# Combine the first three rows into a single character vector (column names)
combined_names <- sapply(traits[1:3, ], function(x) paste(na.omit(x), collapse = " | "))

# Assign the new names to the columns
colnames(traits) <- combined_names

# Drop the first three rows now that they're in the header
traits <- traits[-c(1:3), ]

# Extract the first row (which contains new column names for the first 8 columns)
new_names_part <- as.character(unlist(traits[1, 1:8]))

# Replace the first 8 column names
colnames(traits)[1:8] <- new_names_part

# Remove the first row now that it's used as header
traits <- traits[-1, ]

# Reset rownames (just for cleanliness as they don't start from 1)
rownames(traits) <- NULL
```

# Calculate average trophic level based on LARVAL diet

```{r}
larvae_diet <- traits %>% dplyr::select(c(1:8, 33:40))

# define diet columns for filtering
LarvalDiets <- c("Larval diet | Fungivory",                                     
                    "Larval diet | Detritivory",                                  
                    "Larval diet | Phytophagy",                                    
                    "Larval diet | Predation",                                    
                    "Larval diet | Parasitoidism",                                 
                    "Larval diet | Ecto-parasitism",                              
                    "Larval diet | Non-feeding",
                    "Larval diet | Liquid feeding and nectivory")

# filter for taxa where the larval diet is not an NA and has a fuzzy coding > 0 - i.e. that diet is present!
larvae_diet <- larvae_diet %>%
  filter(if_any(all_of(LarvalDiets), ~ !is.na(.) & . > 0)) 

# Filter for big five
BigFiveLarvalDiet <- larvae_diet %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))

# Map trophic levels onto larval diet categories
trophic_map <- c(
  "Larval diet | Fungivory" = 2,
  "Larval diet | Detritivory" = 2,
  "Larval diet | Phytophagy" = 2,
  "Larval diet | Predation" = 3,
  "Larval diet | Parasitoidism" = 3,
  "Larval diet | Ecto-parasitism" = 3,
  "Larval diet | Non-feeding" = NA,   # But there are no larval diets which are non feeding
  "Larval diet | Liquid feeding and nectivory" = 2.5
)

# Pivot longer to have each row a larval diet category across families
BigFiveLarvalDiet_long <- BigFiveLarvalDiet %>%
  pivot_longer(
    cols = starts_with("Larval diet"),
    names_to = "Diet",
    values_to = "Value"
  ) %>%
  mutate(TrophicLevel = trophic_map[Diet])

# Weighted mean trophic level per family
AverageLarvalTrophicLevelFamily <- BigFiveLarvalDiet_long %>%
  filter(!is.na(TrophicLevel)) %>%
  mutate(
    Value = as.numeric(Value),
    TrophicLevel = as.numeric(TrophicLevel)
  ) %>%
  group_by(Family) %>%
  summarise(
    AvgTrophicLevel = weighted.mean(TrophicLevel, Value, na.rm = TRUE),
    .groups = "drop"
  )
```

# Calculate average trophic level based on ADULT diet

```{r}
# Select cols of interest
adult_diet <- traits %>% dplyr::select(c(1:8, 41:48))

# define diet columns for filtering
AdultDiets <- c("Adult diet | Fungivory",                                     
                    "Adult diet | Detritivory",                                  
                    "Adult diet | Phytophagy",                                    
                    "Adult diet | Predation",                                    
                    "Adult diet | Parasitoidism",                                 
                    "Adult diet | Ecto-parasitism",                              
                    "Adult diet | Non-feeding",
                    "Adult diet | Liquid feeding and nectivory")

# filter for taxa where the adult diet is not an NA and has a fuzzy coding > 0 - i.e. that diet is present!
adult_diet <- adult_diet %>%
  filter(if_any(all_of(AdultDiets), ~ !is.na(.) & . > 0)) 

# Filter for big five
BigFiveAdultDiet <- adult_diet %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))

# Map trophic levels onto adult diet categories
trophic_map <- c(
  "Adult diet | Fungivory" = 2,
  "Adult diet | Detritivory" = 2,
  "Adult diet | Phytophagy" = 2,
  "Adult diet | Predation" = 3,
  "Adult diet | Parasitoidism" = 3,
  "Adult diet | Ecto-parasitism" = 3,
  "Adult diet | Non-feeding" = NA,   # no trophic level
  "Adult diet | Liquid feeding and nectivory" = 2.5
)

# Pivot longer to have each row an adult diet category across families
BigFiveAdultDiet_long <- BigFiveAdultDiet %>%
  pivot_longer(
    cols = starts_with("Adult diet"),
    names_to = "Diet",
    values_to = "Value"
  ) %>%
  mutate(TrophicLevel = trophic_map[Diet])

# Weighted mean trophic level per family
AverageAdultTrophicLevelFamily <- BigFiveAdultDiet_long %>%
  filter(!is.na(TrophicLevel)) %>%
  mutate(
    Value = as.numeric(Value),
    TrophicLevel = as.numeric(TrophicLevel)
  ) %>%
  group_by(Family) %>%
  summarise(
    AvgTrophicLevel = weighted.mean(TrophicLevel, Value, na.rm = TRUE),
    .groups = "drop"
  )
```

# For those families that have non-feeding as an adult diet category, I will assign them the same trophic level based on their diet category as a larvae.

```{r}

AverageAdultLarvaeTrophicLevelFamily <- AverageAdultTrophicLevelFamily %>%
  left_join(
    AverageLarvalTrophicLevelFamily, # join larval values into the adult data frame by Family
    by = "Family",
    suffix = c("_adult", "_larval")
  ) %>%
  mutate(
    AvgTrophicLevelAdult = ifelse(
      is.na(AvgTrophicLevel_adult), # If the adult diet is NA, replace it with the larval one
      AvgTrophicLevel_larval,
      AvgTrophicLevel_adult
    )
  ) %>%
  dplyr::select(Family, AvgTrophicLevel_larval, AvgTrophicLevelAdult) 

  hist(AverageLarvalTrophicLevelFamily$AvgTrophicLevel)

rm(adult_diet, BigFiveAdultDiet, BigFiveAdultDiet_long, BigFiveLarvalDiet, BigFiveLarvalDiet_long, larvae_diet, AverageLarvalTrophicLevelFamily, AverageAdultTrophicLevelFamily)
```

# Read in PREDICTS data and ensure that species level entities such as "Lasioglossum sp." are included for merging average trophic level values using the Taxon_name_entered column

```{r}
# All insect data effort-corrected and merged
taxon_insects <-  qread("/Users/jusi/Desktop/multi-groups/data/predicts-prepared-whole-database.qs")

# Get rid of rows which aren't resolved to order level
taxon_insects <- taxon_insects %>% filter(Order != "") # 1056 rows not resolved to order

# Recreate genus and species column as sometimes BGB is present in PREDICTS but no info for genus or sp.
taxon_insects <- taxon_insects %>%
  mutate(Best_guess_binomial = as.character(Best_guess_binomial)) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "merge", fill = "right")

# There are entries in PREDICTS such as "Lasioglossum sp" - rows with such names tend to not have an entry in the genus column (due to issues with the COL API)
# Previously these taxa were getting family level values in the average mass analysis, despite their genus information being available.
# I will extract the genus information from the Taxon_name_entered column for those rows that do not yet have an entry in genus.
# This will improve the precision of the matching, such that these rows will now be matched at the genus level instead of the family level (as they were beforehand)
taxon_insects <- taxon_insects %>%
  mutate(
    Genus = case_when(
      is.na(Genus) | Genus == "" ~ word(Taxon_name_entered, 1), # Check if Genus is empty or missing, if so, extract the first word from Taxon_name_entered and assign it to genus.
      TRUE ~ Genus # If genus is not empty, leave it as the pre-existing genus entry. 
    ))

# Have a look
taxon_insects %>% dplyr::select(SS, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, Precision_score) %>% View()

# Remove those entries which have a "?" before or after the name of the genus e.g. "?Abalabus" or "Abalabus?"
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove_all(Genus, "\\?"))

# Remove leading and trailing double quotation marks
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_replace_all(Genus, '^"|"$', ""))

# Remove any numbers 
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove_all(Genus, "[0-9]"))

# Remove entries with sp at the end
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove(Genus, "(_|\\s)?sp\\.?$"))

# Filter for columns of interest
taxon_insects <- taxon_insects %>% dplyr::select(Source_ID, SS, SSB, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, Diversity_metric, Diversity_metric_unit, Merged_diversity, Predominant_habitat, Use_intensity)
```

# Merge adult and larvae trophic levels into my overall data frame based on Family

```{r}
AllInsectsTrophicLevel <- taxon_insects %>%
  left_join(
    AverageAdultLarvaeTrophicLevelFamily,
    by = "Family"
  )

# Have a look at an example family where we have the average trophic level for the adult but not for the larvae
AllInsectsTrophicLevel %>% dplyr::select(SS, SSBS, Family, Genus, Species, Taxon_name_entered, AvgTrophicLevel_larval, AvgTrophicLevelAdult) %>% 
  filter(Family == "Acartophthalmidae") %>% View()

rm(AverageAdultLarvaeTrophicLevelFamily, taxon_insects)
```

# Now that I have an overall data frame for average trophic level for all insect taxa in PREDICTS, I will now filter it for species level entities using APs taxon resolution file and then prepare the data frame for analysis.

```{r}
# Filter for diversity metrics of interest - abundance and occurrence data can be used (i.e. no species richness data)
AllInsectsTrophicLevel <- AllInsectsTrophicLevel %>% filter(Diversity_metric %in% c("abundance", "effort-corrected abundance", "density", "group abundance", "occurrence", "occurrence frequency", "relative abundance"))

# Load taxon resolution file
TaxonResolution <- readRDS(here("data", "CuratedTaxonResolution2.rds"))

# Step 1: Filter TaxonResolution for species-level entries - i.e. those entries with a precision score of 4!
TaxonResolutionSp <- TaxonResolution %>%
  filter(PrecisionScore == 4) %>%
  dplyr::select(SS, Taxon_name_entered) # select cols for merging

# Step 2: Keep only matching rows in AllInsectsTrophicLevel
AllInsectsTrophicLevelSpeciesOnly <- AllInsectsTrophicLevel %>%
  semi_join(TaxonResolutionSp, by = c("SS", "Taxon_name_entered")) # keeps only rows in AllInsectsTrophicLevel that have a match in TaxonResolutionSp. based on SS and Taxon_name_entered

# AllInsectsTrophicLevelSpeciesOnly now contains only the rows from AllInsectsTrophicLevel that correspond to species-level entries in TaxonResolutionSp. 

# Create a present or absent column - its present if Merged_diversity > 0 and absent if its 0
AllInsectsTrophicLevelSpeciesOnly <- AllInsectsTrophicLevelSpeciesOnly %>%
  mutate(PresentOrAbsent = ifelse(Merged_diversity > 0, 1, 0))

# Create a log(abundance when present) column
AllInsectsTrophicLevelSpeciesOnly <- AllInsectsTrophicLevelSpeciesOnly %>% 
  mutate(
    LogAbundanceWhenPresent = ifelse(Merged_diversity > 0,
                                    log(Merged_diversity),
                                    NA))

# Have a look at how many rows are present versus absent
table(AllInsectsTrophicLevelSpeciesOnly$PresentOrAbsent)

# Rename predominant habitat since its not really habitat we're looking at
AllInsectsTrophicLevelSpeciesOnly <- rename(AllInsectsTrophicLevelSpeciesOnly,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
AllInsectsTrophicLevelSpeciesOnly <- AllInsectsTrophicLevelSpeciesOnly %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(AllInsectsTrophicLevelSpeciesOnly$Predominant_land_use)
table(AllInsectsTrophicLevelSpeciesOnly$Predominant_land_use, AllInsectsTrophicLevelSpeciesOnly$Use_intensity)
```

# Recode the land use classes to have Primary, Secondary vegetation, Low-intensity agriculture and high-intensity agriculture

```{r}

AllInsectsTrophicLevelSpeciesOnly <- AllInsectsTrophicLevelSpeciesOnly %>%
  mutate(
    LandUseFinal = case_when(
      # Primary stays as is
      Predominant_land_use == "Primary" ~ "Primary",

      # Any secondary vegetation type -> Secondary Vegetation
      Predominant_land_use %in% c("Young secondary vegetation",
                                  "Intermediate secondary vegetation",
                                  "Mature secondary vegetation") ~ "Secondary vegetation",

      # Plantation forest depends on intensity
      Predominant_land_use == "Plantation forest" & Use_intensity == "Minimal use" ~ "Low-intensity agriculture",
      Predominant_land_use == "Plantation forest" & Use_intensity %in% c("Light use", "Intense use") ~ "High-intensity agriculture",

      # Pasture
      Predominant_land_use == "Pasture" & Use_intensity %in% c("Minimal use", "Light use") ~ "Low-intensity agriculture",
      Predominant_land_use == "Pasture" & Use_intensity == "Intense use" ~ "High-intensity agriculture",

      # Cropland
      Predominant_land_use == "Cropland" & Use_intensity %in% c("Minimal use", "Light use") ~ "Low-intensity agriculture",
      Predominant_land_use == "Cropland" & Use_intensity == "Intense use" ~ "High-intensity agriculture",

      # Drop Urban by assigning NA -these will get dropped when we create the model data
      Predominant_land_use == "Urban" ~ NA_character_
    )
  ) %>%
  # Turn into factor if you want ordered categories
  mutate(LandUseFinal = factor(LandUseFinal,
                               levels = c("Primary", "Secondary vegetation",
                                          "Low-intensity agriculture", "High-intensity agriculture")))

table(AllInsectsTrophicLevelSpeciesOnly$LandUseFinal)
```

# Create random intercept for Species

```{r}
AllInsectsTrophicLevelSpeciesOnly <- AllInsectsTrophicLevelSpeciesOnly %>%
  mutate(
    SpeciesFinal = case_when(
      Species != "" & !is.na(Species) ~ Species,
      (Species == "" | is.na(Species)) & Best_guess_binomial != "" & !is.na(Best_guess_binomial) ~ Best_guess_binomial,
      (Species == "" | is.na(Species)) & (Best_guess_binomial == "" | is.na(Best_guess_binomial)) ~ paste(SS, Taxon_name_entered, sep = " "),
      TRUE ~ NA_character_
    )
  )

sum(is.na(AllInsectsTrophicLevelSpeciesOnly$SpeciesFinal))
AllInsectsTrophicLevelSpeciesOnly %>% filter(SpeciesFinal == "")
sum(is.na(AllInsectsTrophicLevelSpeciesOnly$SpeciesFinal) | AllInsectsTrophicLevelSpeciesOnly$SpeciesFinal == "")


length(unique(AllInsectsTrophicLevelSpeciesOnly$SpeciesFinal))
length(unique(AllInsectsTrophicLevelSpeciesOnly$Taxon_name_entered))
length(unique(AllInsectsTrophicLevelSpeciesOnly$Best_guess_binomial))
length(unique(AllInsectsTrophicLevelSpeciesOnly$Species))
```

# Create model data

```{r}
# drop all rows with NAs for variables of interest
ModelDataLarvalDietOccurrence <- drop_na(AllInsectsTrophicLevelSpeciesOnly, LandUseFinal, Use_intensity, PresentOrAbsent, AvgTrophicLevel_larval) %>% droplevels()
ModelDataLarvalDietAbundance <- drop_na(AllInsectsTrophicLevelSpeciesOnly, LandUseFinal, Use_intensity, LogAbundanceWhenPresent, AvgTrophicLevel_larval) %>% droplevels()
ModelDataAdultDietOccurrence <- drop_na(AllInsectsTrophicLevelSpeciesOnly, LandUseFinal, Use_intensity, PresentOrAbsent, AvgTrophicLevelAdult) %>% droplevels()
ModelDataAdultDietAbundance <- drop_na(AllInsectsTrophicLevelSpeciesOnly, LandUseFinal, Use_intensity, LogAbundanceWhenPresent, AvgTrophicLevelAdult) %>% droplevels()

rm(TaxonResolution, TaxonResolutionSp, AllInsectsTrophicLevel, AllInsectsTrophicLevelSpeciesOnly)
```


# Summary statistics

```{r}
# How many species matched to species level data 
# Check adult
length(unique(AllInsectsTrophicLevelSpeciesOnly$Best_guess_binomial[!is.na(AllInsectsTrophicLevelSpeciesOnly$AvgTrophicLevelAdult)]))
# Check larval
length(unique(AllInsectsTrophicLevelSpeciesOnly$Best_guess_binomial[!is.na(AllInsectsTrophicLevelSpeciesOnly$AvgTrophicLevel_larval)]))
#How many species in total
length(unique(AllInsectsTrophicLevelSpeciesOnly$Best_guess_binomial))
# How many families in total
length(unique(AllInsectsTrophicLevelSpeciesOnly$Family))
```

#### Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(ModelDataLarvalDietOccurrence[ , c("LandUseFinal", "Use_intensity", "AvgTrophicLevel_larval", "PresentOrAbsent")])
corvif(ModelDataLarvalDietAbundance[ , c("LandUseFinal", "Use_intensity","AvgTrophicLevel_larval", "LogAbundanceWhenPresent")])
corvif(ModelDataAdultDietOccurrence[ , c("LandUseFinal", "Use_intensity", "AvgTrophicLevelAdult", "PresentOrAbsent")])
corvif(ModelDataAdultDietAbundance[ , c("LandUseFinal", "Use_intensity","AvgTrophicLevelAdult", "LogAbundanceWhenPresent")])
```

# Convert to factors

```{r}
# ModelDataLarvalDietOccurrence
ModelDataLarvalDietOccurrence$Order <- as.factor(ModelDataLarvalDietOccurrence$Order)
ModelDataLarvalDietOccurrence$Family <- as.factor(ModelDataLarvalDietOccurrence$Family)
ModelDataLarvalDietOccurrence$SpeciesFinal <- as.factor(ModelDataLarvalDietOccurrence$SpeciesFinal)
ModelDataLarvalDietOccurrence$SS <- as.factor(ModelDataLarvalDietOccurrence$SS)
ModelDataLarvalDietOccurrence$SSB <- as.factor(ModelDataLarvalDietOccurrence$SSB)

# ModelDataLarvalDietAbundance
ModelDataLarvalDietAbundance$Order <- as.factor(ModelDataLarvalDietAbundance$Order)
ModelDataLarvalDietAbundance$Family <- as.factor(ModelDataLarvalDietAbundance$Family)
ModelDataLarvalDietAbundance$SpeciesFinal <- as.factor(ModelDataLarvalDietAbundance$SpeciesFinal)
ModelDataLarvalDietAbundance$SS <- as.factor(ModelDataLarvalDietAbundance$SS)
ModelDataLarvalDietAbundance$SSB <- as.factor(ModelDataLarvalDietAbundance$SSB)

# ModelDataAdultDietOccurrence
ModelDataAdultDietOccurrence$Order <- as.factor(ModelDataAdultDietOccurrence$Order)
ModelDataAdultDietOccurrence$Family <- as.factor(ModelDataAdultDietOccurrence$Family)
ModelDataAdultDietOccurrence$SpeciesFinal <- as.factor(ModelDataAdultDietOccurrence$SpeciesFinal)
ModelDataAdultDietOccurrence$SS <- as.factor(ModelDataAdultDietOccurrence$SS)
ModelDataAdultDietOccurrence$SSB <- as.factor(ModelDataAdultDietOccurrence$SSB)

# ModelDataAdultDietAbundance
ModelDataAdultDietAbundance$Order <- as.factor(ModelDataAdultDietAbundance$Order)
ModelDataAdultDietAbundance$Family <- as.factor(ModelDataAdultDietAbundance$Family)
ModelDataAdultDietAbundance$SpeciesFinal <- as.factor(ModelDataAdultDietAbundance$SpeciesFinal)
ModelDataAdultDietAbundance$SS <- as.factor(ModelDataAdultDietAbundance$SS)
ModelDataAdultDietAbundance$SSB <- as.factor(ModelDataAdultDietAbundance$SSB)

#saveRDS(ModelDataAdultDietAbundance, "./C4 data/ModelDataAdultDietAbundance.rds")
#saveRDS(ModelDataAdultDietOccurrence, "./C4 data/ModelDataAdultDietOccurrence.rds")
#saveRDS(ModelDataLarvalDietAbundance, "./C4 data/ModelDataLarvalDietAbundance.rds")
#saveRDS(ModelDataLarvalDietOccurrence, "./C4 data/ModelDataLarvalDietOccurrence.rds")
```

# Modelling log(abundance of present species) ~ LUFinal * average LARVAL trophic level

```{r}
# Maximal random effects structure
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevel_larval + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = ModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop block
ab_m2 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevel_larval + (1 | Source_ID/SS) + (1| Family/SpeciesFinal), 
               data = ModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop SS
ab_m3 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevel_larval + (1 | Source_ID) + (1| Family/SpeciesFinal), 
               data = ModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop Family
ab_m4 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevel_larval + (1 | Source_ID) + (1 | SpeciesFinal), 
               data = ModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Compare models
anova(ab_m1, ab_m2, ab_m3, ab_m4)  # Likelihood ratio test
AIC(ab_m1, ab_m2, ab_m3, ab_m4)


# ab_m1 best!
```

# Backwards stepwise selection of fixed effects

```{r}
# Additive model with the interaction removed but both predictors still present
ab_m5 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + AvgTrophicLevel_larval + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


# Just land use as predictor
ab_m6 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Just larval trophic level as the predictor
ab_m7 <- blmer(LogAbundanceWhenPresent ~ AvgTrophicLevel_larval + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Make a null, intercept only model
ab_m8 <- blmer(LogAbundanceWhenPresent ~ (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


## Compare the models
AIC(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m5, ab_m1)
```

# Check the model
```{r}
model_plot(ab_m1)
summary(ab_m1)

rm(ab_m2, ab_m3, ab_m4, ab_m5, ab_m6, ab_m7, ab_m8)
```

# Generate predictions from the model

```{r}
# Create a grid of values
NewDataAbundanceTrophicLarvae <- expand.grid(
  AvgTrophicLevel_larval = seq(from = min(ModelDataLarvalDietAbundance$AvgTrophicLevel_larval),
                   to = max(ModelDataLarvalDietAbundance$AvgTrophicLevel_larval),
                   length.out = 100),
  LandUseFinal = levels(ModelDataLarvalDietAbundance$LandUseFinal))
  

# Get predictions without random effects, with SEs
AbundancePredictions <- predict(ab_m1, newdata = NewDataAbundanceTrophicLarvae, re.form = NA, se.fit = TRUE)

# Add predictions and 95% CIs
NewDataAbundanceTrophicLarvae$AbundFit <- AbundancePredictions$fit
NewDataAbundanceTrophicLarvae$AbundSE  <- AbundancePredictions$se.fit
NewDataAbundanceTrophicLarvae$ExpAbundFit <- exp(NewDataAbundanceTrophicLarvae$AbundFit)  # back-transform
NewDataAbundanceTrophicLarvae$AbundLwr <- exp(NewDataAbundanceTrophicLarvae$AbundFit - 1.96 * NewDataAbundanceTrophicLarvae$AbundSE) # exponentiate log abundance values 
NewDataAbundanceTrophicLarvae$AbundUpr <- exp(NewDataAbundanceTrophicLarvae$AbundFit + 1.96 * NewDataAbundanceTrophicLarvae$AbundSE)

summary(ab_m1)

# Extract coefficient in primary
fixef(ab_m1)["AvgTrophicLevel_larval"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p( parameters::p_value(ab_m1, method = "eti"))

# Extract mass x land use F ratio
anova(ab_m1)
Anova(ab_m1)

# Plot
PlotAbundanceTrophicLarvae <- ggplot(NewDataAbundanceTrophicLarvae, aes(x = AvgTrophicLevel_larval, y = ExpAbundFit,
                   color = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = AbundLwr, ymax = AbundUpr, fill = LandUseFinal), alpha = 0.2, color = NA, show.legend = F) +
  labs(x = expression("Larval trophic level"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = c(2,3), labels = c("Primary consumer", "Secondary consumer")) 


# Remove legend for plot for combining later
PlotAbundanceTrophicLarvaeNoLegend <- PlotAbundanceTrophicLarvae + theme(legend.position = "none") +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC"))

# Add text
FinalPlotAbundanceTrophicLarvaeNoLegend <-  ggdraw(PlotAbundanceTrophicLarvaeNoLegend) +
  draw_label("Coefficient in primary: 
  - 0.13 [95% CI: -0.24, -0.02], p = 0.02
              Larval TL x LU interaction: 
  F-ratio = 5.04, p = 0.002", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

# Singular plot to save on its own (not combining later)
FinalPlotAbundanceTrophicLarvae <- ggdraw(PlotAbundanceTrophicLarvae) +
  draw_label("Coefficient in primary: 
  - 0.13 [95% CI: -0.24, -0.02], p = 0.02
              Larval TL x LU interaction: 
  F-ratio = 5.04, p = 0.002", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

#ggsave(filename = "./figures/c4_updated/LarvalTrophicLevelAbundanceWhenPresent.png",  plot = FinalPlotAbundanceTrophicLarvae, width = 6, height = 6, dpi = 300)
```

# Modelling LogAbundanceWhenPresent ~ LUFinal * larval trophic level * order

```{r}
# Filter for big five
BigFiveModelDataLarvalDietAbundance <- ModelDataLarvalDietAbundance %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))

# Maximal model
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevel_larval * Order + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = BigFiveModelDataLarvalDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Check the model
model_plot(ab_m1)
summary(ab_m1)
anova(ab_m1)
Anova(ab_m1)

# Calculate the larval trophic level range for predictions
LarvalTrophicLevelRange <- range(BigFiveModelDataLarvalDietAbundance$AvgTrophicLevel_larval, na.rm = TRUE)

# Generate predictions
# How does abundance scale with larval trophic level for each LU * Order combo
LarvalTrophicLevelPredictions <- ggpredict(
  ab_m1,
  terms = c(
    sprintf("AvgTrophicLevel_larval [%.2f:%.2f]", LarvalTrophicLevelRange[1], LarvalTrophicLevelRange[2]), # Generate predictions using that range
    "LandUseFinal",
    "Order"
  ))

# Plot
PlotLarvalTrophicLevelAbundance <- ggplot(LarvalTrophicLevelPredictions, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.15, color = NA) +
  facet_wrap(~ facet, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))  +
  labs(x = expression("Larval trophic level"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC"))


ggsave(filename = "./figures/AbundanceWhenPresentBigFiveLarvalTrophicLevel.png",   plot = PlotLarvalTrophicLevelAbundance, width = 10, height = 6, dpi = 300)

rm(ab_m1, PlotLarvalTrophicLevelAbundance, LarvalTrophicLevelPredictions)
```

# Modelling PresenceOrAbsent ~ LUFinal * average LARVAL trophic level 

```{r}
# basic data diagnostics
str(ModelDataLarvalDietOccurrence)
table(ModelDataLarvalDietOccurrence$PresentOrAbsent)        # balance of 0/1
table(ModelDataLarvalDietOccurrence$LandUseFinal, useNA="ifany")
n_distinct(ModelDataLarvalDietOccurrence$Source_ID)
n_distinct(ModelDataLarvalDietOccurrence$SS)
n_distinct(ModelDataLarvalDietOccurrence$SpeciesFinal)

# Convert to factors
ModelDataLarvalDietOccurrence$Order <- as.factor(ModelDataLarvalDietOccurrence$Order)
ModelDataLarvalDietOccurrence$Family <- as.factor(ModelDataLarvalDietOccurrence$Family)
ModelDataLarvalDietOccurrence$SpeciesFinal <- as.factor(ModelDataLarvalDietOccurrence$SpeciesFinal)
ModelDataLarvalDietOccurrence$SS <- as.factor(ModelDataLarvalDietOccurrence$SS)
ModelDataLarvalDietOccurrence$SSB <- as.factor(ModelDataLarvalDietOccurrence$SSB)
```

# Scale average larval trophic level to help with convergence
```{r}
ModelDataLarvalDietOccurrence$ScaledAvgTrophicLevel_larval <- scale(ModelDataLarvalDietOccurrence$AvgTrophicLevel_larval, center = TRUE, scale = TRUE)

hist(ModelDataLarvalDietOccurrence$ScaledAvgTrophicLevel_larval) ; hist(ModelDataLarvalDietOccurrence$AvgTrophicLevel_larval)
```

# Random effects structure
```{r}
# Just run the model on its own
  oc_m1 <- bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgTrophicLevel_larval + (1 | SS) + (1| SpeciesFinal), 
               data = ModelDataLarvalDietOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "Nelder_Mead",
                 optCtrl = list(maxfun = 7e5)))

# Save the model
#saveRDS(oc_m1, file = "LarvalTrophicLevelOccurrenceModel.rds")

# Put model formulas in a list
RandomEffectsModelCalls <- list(
  oc_m1 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgTrophicLevel_larval + (1 | SS) + (1| SpeciesFinal), 
               data = ModelDataLarvalDietOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "Nelder_Mead",
                 optCtrl = list(maxfun = 7e5)
  ))
  ),
  oc_m2 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgTrophicLevel_larval + (1 | SS), 
               data = ModelDataLarvalDietOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "Nelder_Mead",
                 optCtrl = list(maxfun = 7e5)
  ))
  ))

# Run models in parallel
RandomEffectsResults <- mclapply(RandomEffectsModelCalls, function(expr) eval(expr),
                                 mc.cores = 2)

oc_m1 <- RandomEffectsResults[[1]]
oc_m2 <- RandomEffectsResults[[2]]

# Compare models
anova(oc_m1, oc_m2)  # Likelihood ratio test
AIC(oc_m1, oc_m2)
```

## Backwards stepwise selection of fixed effects 
```{r}
# Put model formulas in a list

# Additive model
FixedEffectsModelCalls <- list(
  oc_m3 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal + ScaledAvgTrophicLevel_larval + (1 | SS) + (1| SpeciesFinal), 
               data = ModelDataLarvalDietOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "Nelder_Mead",
                 optCtrl = list(maxfun = 7e5)
  ))
  ),
  # Just LU as a predictor
  oc_m4 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal + (1 | SS) + (1| SpeciesFinal), 
               data = ModelDataLarvalDietOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "Nelder_Mead",
                 optCtrl = list(maxfun = 7e5)
  ))
  ),
  # Just average trophic level as a predictor
  oc_m5 = quote(
    bglmer(PresentOrAbsent ~ ScaledAvgTrophicLevel_larval + (1 | SS) + (1| SpeciesFinal), 
               data = ModelDataLarvalDietOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "Nelder_Mead",
                 optCtrl = list(maxfun = 7e5)
  ))
  ),
  # Make a null, intercept only model
  oc_m6 = quote(
    bglmer(PresentOrAbsent ~ (1 | SS) + (1| SpeciesFinal), 
               data = ModelDataLarvalDietOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "Nelder_Mead",
                 optCtrl = list(maxfun = 7e5)
  ))))


# Run models in parallel
FixedEffectsResults <- mclapply(FixedEffectsModelCalls, function(expr) eval(expr),
                    mc.cores = 4)  # adjust cores


oc_m3 <- FixedEffectsResults[[1]]
oc_m4 <- FixedEffectsResults[[2]]
oc_m5 <- FixedEffectsResults[[3]]
oc_m6 <- FixedEffectsResults[[4]]
```

# Check model convergence and compare
```{r}
summary(oc_m3) # additive model - converged
summary(oc_m4) # just LU as predictor - converged
summary(oc_m5) # just egg size as predictor - converged
summary(oc_m6) # null intercept only - converged

## Compare the models
AIC(oc_m1, oc_m3, oc_m4, oc_m5, oc_m6)
anova(oc_m1, oc_m3, oc_m4, oc_m5, oc_m6)
anova(oc_m4, oc_m1)

# oc_m1 the best
```

# Generate predictions from the model
 
```{r}
# Interaction more significant that just additive effect of trophic level
anova(oc_m1)

# Create new data frame for predictions
NewDataOccurrenceTrophicLarvae <- expand.grid(
  ScaledAvgTrophicLevel_larval = seq(from = min(ModelDataLarvalDietOccurrence$ScaledAvgTrophicLevel_larval, na.rm = TRUE),
                to   = max(ModelDataLarvalDietOccurrence$ScaledAvgTrophicLevel_larval, na.rm = TRUE),
                length.out = 100),
  LandUseFinal = levels(ModelDataLarvalDietOccurrence$LandUseFinal))

# Copy the factor levels from the model data
NewDataOccurrenceTrophicLarvae$LandUseFinal <- factor(NewDataOccurrenceTrophicLarvae$LandUseFinal, 
                              levels = levels(ModelDataLarvalDietOccurrence$LandUseFinal))

# Predicted probabilities on link scale (logit) with SE
OccurrencePredictions <- predict(oc_m1, newdata = NewDataOccurrenceTrophicLarvae, type = "link", se.fit = TRUE, re.form = NA)
       
 # Add predictions into data frame
NewDataOccurrenceTrophicLarvae <- NewDataOccurrenceTrophicLarvae %>%
  mutate(
    OccurrenceFit   = OccurrencePredictions$fit,
    OccurrenceSE    = OccurrencePredictions$se.fit,
    OccurrenceLwr = OccurrenceFit - 1.96 * OccurrenceSE,
    OccurrenceUpr = OccurrenceFit + 1.96 * OccurrenceSE,
    # Back-transform to probability scale
    OccurrenceProb = plogis(OccurrenceFit), # predicted probability
    LowerProb = plogis(OccurrenceLwr), # 95% CI lower bound
    UpperProb = plogis(OccurrenceUpr) # 95% CI upper bound
  )

# Extract coefficient in primary
fixef(oc_m1)["ScaledAvgTrophicLevel_larval"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p(parameters::p_value(oc_m1, method = "eti"))

# Extract mass x land use F ratio
anova(oc_m1)
Anova(oc_m1)

# Extract the mean and sd to add the unscaled values back into the prediction data frame for plotting
mean_val <- attr(ModelDataLarvalDietOccurrence$ScaledAvgTrophicLevel_larval, "scaled:center")
sd_val   <- attr(ModelDataLarvalDietOccurrence$ScaledAvgTrophicLevel_larval, "scaled:scale")

# Add a new column with the back transformed values:
NewDataOccurrenceTrophicLarvae$UnscaledAvgTrophicLevel <- NewDataOccurrenceTrophicLarvae$ScaledAvgTrophicLevel_larval * sd_val + mean_val

PlotOccurrenceTrophicLarvae <- ggplot(NewDataOccurrenceTrophicLarvae, aes(x = UnscaledAvgTrophicLevel, y = OccurrenceProb, color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = LowerProb, ymax = UpperProb), 
              alpha = 0.2, color = NA) +
  labs(x = expression("Larval trophic level"),
       y = "Probability of occurrence (± 95 % CI)",
       color = "Land use",
       fill = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = c(2,3), labels = c("Primary consumer", "Secondary consumer"))

# Remove legend
PlotOccurrenceTrophicLarvaeNoLegend <- PlotOccurrenceTrophicLarvae + theme(legend.position = "none") +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC"))

# Add text box 
FinalPlotOccurrenceTrophicLarvaeNoLegend <- ggdraw(PlotOccurrenceTrophicLarvaeNoLegend) +
  draw_label("Coefficient in primary: 
   - 0.04 [95% CI: -0.07, -0.02], p < 0.001
              Larval TL x LU interaction: 
  F-ratio = 23.72, p < 0.0001", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

# Final singular plot
FinalPlotOccurrenceTrophicLarvae <- ggdraw(PlotOccurrenceTrophicLarvae) +
  draw_label("Coefficient in primary: 
   - 0.04 [95% CI: -0.07, -0.02], p < 0.001
              Larval TL x LU interaction: 
  F-ratio = 23.72, p < 0.0001", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

      #ggsave(filename = "./figures/c4_updated/LarvalTrophicLevelOccurrence.png",  plot = FinalPlotOccurrenceTrophicLarvae, width = 6, height = 6, dpi = 300)

rm(oc_m1, oc_m2, oc_m3, oc_m4, oc_m5, oc_m6, RandomEffectsModelCalls, RandomEffectsResults, plot, OccurrencePredictions, PlotOccurrenceTrophicLarvae, FixedEffectsResults, FixedEffectsModelCalls, FinalPlotOccurrenceTrophicLarvae)
```

# Combine the occurrence and abundance predictions

```{r}
NewDataCombined <- NewDataOccurrenceTrophicLarvae %>%
  bind_cols(NewDataAbundanceTrophicLarvae %>% 
              dplyr::select(ExpAbundFit = ExpAbundFit,
                     AbundSE  = AbundSE)) %>%
  mutate(
    # Backtransform to natural scales
    #OccurrenceProb   = plogis(OccurrenceFit), # probability of occurrence
    #ExpAbundFit  = exp(AbundFit), # predicted abundance when present
    
    # Probability of occurrence * predicted abundance when present = predicted abundance ???
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * ExpAbundFit,
    
    # Delta method for SE of product on natural scale
    PredictedAbundanceSE = sqrt(
      (ExpAbundFit^2) * (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2) +  # variance of plogis(fit_link)
      (OccurrenceProb^2) * (ExpAbundFit^2) * (AbundSE^2)                      # variance of exp(fit_abund)
    ),
    
    # 95% confidence intervals
    PredictedAbundanceLwr = PredictedAbundanceAccountingForOccurrenceProbability - 1.96 * PredictedAbundanceSE,
    PredictedAbundanceUpr = PredictedAbundanceAccountingForOccurrenceProbability + 1.96 * PredictedAbundanceSE
  )
  

# Plot
FinalPlotAbundanceOccurrenceLarvaeTLCombined <- ggplot(NewDataCombined,
       aes(x = UnscaledAvgTrophicLevel, y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = PredictedAbundanceLwr, ymax = PredictedAbundanceUpr),
              alpha = 0.2, color = NA) +
  labs(x = expression("Larval trophic level"),
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC")) 

#ggsave(filename = "./figures/c4_updated/LarvalTrophicLevelAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceLarvaeTLCombined, width = 6, height = 6, dpi = 300)
```

# Combine all three plots 

```{r}
# Combine with shared legend and auto subplot labels
AllThreeTogether <- (FinalPlotAbundanceTrophicLarvaeNoLegend | FinalPlotOccurrenceTrophicLarvaeNoLegend | FinalPlotAbundanceOccurrenceLarvaeTLCombined) +
  plot_layout(guides = "collect") &    # combine legends
  plot_annotation(tag_levels = 'a') &  # adds a, b, c tags
  theme(legend.position = "bottom",
        plot.tag = element_text(size = 18, face = "bold"),
                 plot.tag.position = c(0, 1))    # shared legend position

#ggsave(filename = "./figures/c4_updated/LarvalTrophicLevelAllThreeCombined.png",  plot = AllThreeTogether, width = 14, height = 7, dpi = 300)

rm(ab_m1, AbundancePredictions, AllThreeTogether, FinalPlotOccurrenceTrophicLarvaeNoLegend, FinalPlotOccurrenceTrophicLarvae, FinalPlotAbundanceTrophicLarvaeNoLegend, FinalPlotAbundanceTrophicLarvae, FinalPlotAbundanceOccurrenceLarvaeTLCombined, NewDataAbundanceTrophicLarvae, NewDataCombined, NewDataOccurrenceTrophicLarvae, oc_m1, OccurrencePredictions, PlotAbundanceTrophicLarvae, PlotAbundanceTrophicLarvaeNoLegend, PlotCombinedTrophicLarvae, PlotOccurrenceTrophicLarvae, PlotOccurrenceTrophicLarvaeNoLegend)
```


# Modelling log(abundance of present species) ~ LUFinal * average ADULT trophic level

```{r}
# Maximal random effects structure
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevelAdult + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = ModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop block
ab_m2 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevelAdult + (1 | Source_ID/SS) + (1| Family/SpeciesFinal), 
               data = ModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop SS
ab_m3 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevelAdult + (1 | Source_ID) + (1| Family/SpeciesFinal), 
               data = ModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop Family
ab_m4 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevelAdult + (1 | Source_ID) + (1 | SpeciesFinal), 
               data = ModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Compare models
anova(ab_m1, ab_m2, ab_m3, ab_m4)  # Likelihood ratio test
AIC(ab_m1, ab_m2, ab_m3, ab_m4)


# ab_m1 best!
```

# Backwards stepwise selection of fixed effects

```{r}
# Additive model with the interaction removed but both predictors still present
ab_m5 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + AvgTrophicLevelAdult + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


# Just land use as predictor
ab_m6 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Just ADULT trophic level as the predictor
ab_m7 <- blmer(LogAbundanceWhenPresent ~ AvgTrophicLevelAdult + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Make a null, intercept only model
ab_m8 <- blmer(LogAbundanceWhenPresent ~ (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


## Compare the models
AIC(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m6, ab_m1)
anova(ab_m5, ab_m1)

# ab_m1 marginally better than additive (p = 0.02) and model with just LU as a predictor (p = 0.02)
```

# Check the model
```{r}
model_plot(ab_m1)
summary(ab_m1)
```

# Generate predictions from the model

```{r}
# Create a grid of values
NewDataAbundanceTrophicAdult <- expand.grid(
  AvgTrophicLevelAdult = seq(from = min(ModelDataAdultDietAbundance$AvgTrophicLevelAdult),
                   to = max(ModelDataAdultDietAbundance$AvgTrophicLevelAdult),
                   length.out = 100),
  LandUseFinal = levels(ModelDataAdultDietAbundance$LandUseFinal))
  

# Get predictions without random effects, with SEs
AbundancePredictions <- predict(ab_m1, newdata = NewDataAbundanceTrophicAdult, re.form = NA, se.fit = TRUE)

# Add predictions and 95% CIs
NewDataAbundanceTrophicAdult$AbundFit <- AbundancePredictions$fit
NewDataAbundanceTrophicAdult$AbundSE  <- AbundancePredictions$se.fit
NewDataAbundanceTrophicAdult$ExpAbundFit <- exp(NewDataAbundanceTrophicAdult$AbundFit)  # back-transform
NewDataAbundanceTrophicAdult$AbundLwr <- exp(NewDataAbundanceTrophicAdult$AbundFit - 1.96 * NewDataAbundanceTrophicAdult$AbundSE) # exponentiate log abundance values 
NewDataAbundanceTrophicAdult$AbundUpr <- exp(NewDataAbundanceTrophicAdult$AbundFit + 1.96 * NewDataAbundanceTrophicAdult$AbundSE)

summary(ab_m1)

# Extract coefficient in primary
fixef(ab_m1)["AvgTrophicLevelAdult"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p(parameters::p_value(ab_m1, method = "eti"))

# Extract mass x land use F ratio
anova(ab_m1)
Anova(ab_m1)

# Plot
PlotAbundanceTrophicAdult <- ggplot(NewDataAbundanceTrophicAdult,
       aes(x = AvgTrophicLevelAdult, y = ExpAbundFit,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = AbundLwr, ymax = AbundUpr),
              alpha = 0.2, color = NA) +
  labs(x = expression("Adult trophic level"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = c(2,3), labels = c("Primary consumer", "Secondary consumer")) 

# Remove legend and change x axis tick labels for combining with other plots
PlotAbundanceTrophicAdultNoLegend <- PlotAbundanceTrophicAdult + theme(legend.position = "none") +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC")) # force x axis


# Add text box
FinalPlotAbundanceTrophicAdultNoLegend <- ggdraw(PlotAbundanceTrophicAdultNoLegend) +
  draw_label("Coefficient in primary: 
  0.08 [95% CI: -0.07, 0.24], p = 0.291
              Adult TL x LU interaction: 
  F-ratio = 3.53, p = 0.014", x = 0.97, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)


FinalPlotAbundanceTrophicAdult <- ggdraw(PlotAbundanceTrophicAdult) +
  draw_label("Coefficient in primary: 
  0.08 [95% CI: -0.07, 0.24], p = 0.291
              Adult TL x LU interaction: 
  F-ratio = 3.53, p = 0.014", x = 0.97, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

       #ggsave(filename = "./figures/c4_updated/AdultTrophicLevelAbundanceWhenPresent.png",  plot = FinalPlotAbundanceTrophicAdult, width = 6, height = 6, dpi = 300)
```

# Modelling LogAbundanceWhenPresent ~ LUFinal * ADULT trophic level * order

```{r}
# Filter for big five
BigFiveModelDataAdultDietAbundance <- ModelDataAdultDietAbundance %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))

# Maximal model
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * AvgTrophicLevelAdult * Order + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = BigFiveModelDataAdultDietAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Check the model
model_plot(ab_m1)
summary(ab_m1)
anova(ab_m1)
Anova(ab_m1)

# Calculate the adult trophic level range for predictions
AdultTrophicLevelRange <- range(BigFiveModelDataAdultDietAbundance$AvgTrophicLevelAdult, na.rm = TRUE)

# Generate predictions
# How does abundance scale with adult trophic level for each LU * Order combo
AdultTrophicLevelPredictions <- ggpredict(
  ab_m1,
  terms = c(
    sprintf("AvgTrophicLevelAdult [%.2f:%.2f]", AdultTrophicLevelRange[1], AdultTrophicLevelRange[2]), # Generate predictions using that range
    "LandUseFinal",
    "Order"
  ))

# Plot
PlotAdultTrophicLevelAbundance <- ggplot(AdultTrophicLevelPredictions, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.15, color = NA) +
  facet_wrap(~ facet, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))  +
  labs(x = expression("Adult trophic level"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC"))


#ggsave(filename = "./figures/AbundanceWhenPresentBigFiveAdultTrophicLevel.png",   plot = PlotAdultTrophicLevelAbundance, width = 10, height = 6, dpi = 300)

rm(ab_m1, PlotLarvalTrophicLevelAbundance, LarvalTrophicLevelPredictions)
```

# Modelling PresenceOrAbsent ~ LUFinal * average ADULT trophic level 

```{r}
# basic data diagnostics
str(ModelDataAdultDietOccurrence)
table(ModelDataAdultDietOccurrence$PresentOrAbsent)        # balance of 0/1
table(ModelDataAdultDietOccurrence$LandUseFinal, useNA="ifany")
n_distinct(ModelDataAdultDietOccurrence$Source_ID)
n_distinct(ModelDataAdultDietOccurrence$SS)
n_distinct(ModelDataAdultDietOccurrence$SpeciesFinal)
```

# Scale average ADULT trophic level to help with convergence
```{r}
ModelDataAdultDietOccurrence$ScaledAvgTrophicLevelAdult <- scale(ModelDataAdultDietOccurrence$AvgTrophicLevelAdult, center = TRUE, scale = TRUE)
hist(ModelDataAdultDietOccurrence$ScaledAvgTrophicLevelAdult) ; hist(ModelDataAdultDietOccurrence$AvgTrophicLevelAdult)
```

# Random effects structure
```{r}
# Run just the model
  oc_m1 <- bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgTrophicLevelAdult + (1 | Source_ID) + (1 | SpeciesFinal), 
               data = ModelDataAdultDietOccurrence, family = binomial(link = "logit"),
  control = glmerControl(
    optimizer = "nloptwrap",
    optCtrl = list(
      algorithm = "NLOPT_LN_NELDERMEAD",
      maxeval = 1e4)))
    
# Save the model
#saveRDS(oc_m1, "AdultTrophicLevelOccurrenceModel.rds")

# Maximum random effects structure
# Put model formulas in a list
RandomEffectsModelCalls <- list(
  oc_m1 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgTrophicLevelAdult + (1 | Source_ID) + (1 | SpeciesFinal), 
               data = ModelDataAdultDietOccurrence, family = binomial(link = "logit"),
  control = glmerControl(
    optimizer = "nloptwrap",
    optCtrl = list(
      algorithm = "NLOPT_LN_NELDERMEAD",
      maxeval = 1e4)
  ))
),
oc_m2 = quote(
  bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgTrophicLevelAdult + (1 | Source_ID), 
               data = ModelDataAdultDietOccurrence, family = binomial(link = "logit"),
  control = glmerControl(
    optimizer = "nloptwrap",
    optCtrl = list(
      algorithm = "NLOPT_LN_NELDERMEAD",
      maxeval = 1e4)
))))

# Run models in parallel
RandomEffectsResults <- mclapply(RandomEffectsModelCalls, function(expr) eval(expr),
                    mc.cores = 2)  # adjust cores


oc_m1 <- RandomEffectsResults[[1]]
oc_m2 <- RandomEffectsResults[[2]]

summary(oc_m1)
summary(oc_m2)

# Compare models
anova(oc_m1, oc_m2)  # Likelihood ratio test
AIC(oc_m1, oc_m2)

```

## Backwards stepwise selection of fixed effects 
```{r}
# Put model formulas in a list
# Additive model

FixedfEffectsModelCalls <- list(
  oc_m3 = quote(
     bglmer(PresentOrAbsent ~ LandUseFinal + ScaledAvgTrophicLevelAdult + (1 | Source_ID) + (1|SpeciesFinal), 
               data = ModelDataAdultDietOccurrence, family = binomial(link = "logit"),
  control = glmerControl(
    optimizer = "nloptwrap",
    optCtrl = list(
      algorithm = "NLOPT_LN_NELDERMEAD",
      maxeval = 1e4)))
  ),
  oc_m4 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal + (1 | Source_ID) + (1|SpeciesFinal), 
               data = ModelDataAdultDietOccurrence, family = binomial(link = "logit"),
  control = glmerControl(
    optimizer = "nloptwrap",
    optCtrl = list(
      algorithm = "NLOPT_LN_NELDERMEAD",
      maxeval = 1e4)))
  ),
  oc_m5 = quote(
    bglmer(PresentOrAbsent ~ ScaledAvgTrophicLevelAdult + (1 | Source_ID) + (1|SpeciesFinal), 
               data = ModelDataAdultDietOccurrence, family = binomial(link = "logit"),
  control = glmerControl(
    optimizer = "nloptwrap",
    optCtrl = list(
      algorithm = "NLOPT_LN_NELDERMEAD",
      maxeval = 1e4)))
  ),
  oc_m6 = quote(
    bglmer(PresentOrAbsent ~ (1 | Source_ID) + (1|SpeciesFinal), 
               data = ModelDataAdultDietOccurrence, family = binomial(link = "logit"),
  control = glmerControl(
    optimizer = "nloptwrap",
    optCtrl = list(
      algorithm = "NLOPT_LN_NELDERMEAD",
      maxeval = 1e4)))
  ))

# Run models in parallel
FixedEffectResults <- mclapply(FixedfEffectsModelCalls, function(expr) eval(expr),
                    mc.cores = 4)  # adjust cores


oc_m3 <- FixedEffectResults[[1]]
oc_m4 <- FixedEffectResults[[2]]
oc_m5 <- FixedEffectResults[[3]]
oc_m6 <- FixedEffectResults[[4]]
```

# Check model convergence and compare
```{r}
summary(oc_m3) # additive model - failed to converge
summary(oc_m4) # just LU as predictor - failed to converge
summary(oc_m5) # just egg size as predictor - converged
summary(oc_m6) # null intercept only - failed to converge

## Compare the models
AIC(oc_m1, oc_m5)
anova(oc_m1, oc_m5)

# oc_m1 the best
```

# Generate predictions from the model
 
```{r}
# Interaction more significant that just additive effect of adult trophic level
anova(oc_m1)

# Create new data frame for predictions
NewDataOccurrenceTrophicAdult <- expand.grid(
  ScaledAvgTrophicLevelAdult = seq(from = min(ModelDataAdultDietOccurrence$ScaledAvgTrophicLevelAdult, na.rm = TRUE),
                to   = max(ModelDataAdultDietOccurrence$ScaledAvgTrophicLevelAdult, na.rm = TRUE),
                length.out = 100),
  LandUseFinal = levels(ModelDataAdultDietOccurrence$LandUseFinal))

# Copy the factor levels from the model data
NewDataOccurrenceTrophicAdult$LandUseFinal <- factor(NewDataOccurrenceTrophicAdult$LandUseFinal, 
                              levels = levels(ModelDataAdultDietOccurrence$LandUseFinal))

# Predicted probabilities on link scale (logit) with SE
OccurrencePredictions <- predict(oc_m1, newdata = NewDataOccurrenceTrophicAdult, type = "link", se.fit = TRUE, re.form = NA)

# Add predictions into data frame
NewDataOccurrenceTrophicAdult <- NewDataOccurrenceTrophicAdult %>%
  mutate(
    OccurrenceFit   = OccurrencePredictions$fit,
    OccurrenceSE    = OccurrencePredictions$se.fit,
    OccurrenceLwr = OccurrenceFit - 1.96 * OccurrenceSE,
    OccurrenceUpr = OccurrenceFit + 1.96 * OccurrenceSE,
    # Back-transform to probability scale
    OccurrenceProb   = plogis(OccurrenceFit), # predicted probability - plogis() is the inverse of the logit function.
    # It takes values on the logit (link) scale (from −∞ to +∞) and transforms them into probabilities between 0 and 1.
    LowerProb = plogis(OccurrenceLwr), # 95% CI lower bound
    UpperProb = plogis(OccurrenceUpr) # 95% CI upper bound
  )

# Extract coefficient in primary
fixef(oc_m1)["ScaledAvgTrophicLevelAdult"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p(parameters::p_value(oc_m1, method = "eti"))

# Extract mass x land use F ratio
anova(oc_m1)
Anova(oc_m1)

# Extract the mean and sd to add the unscaled values back into the prediction data frame for plotting
mean_val <- attr(ModelDataAdultDietOccurrence$ScaledAvgTrophicLevelAdult, "scaled:center")
sd_val   <- attr(ModelDataAdultDietOccurrence$ScaledAvgTrophicLevelAdult, "scaled:scale")

# Add a new column with the back transformed values:
NewDataOccurrenceTrophicAdult$UnscaledAvgTrophicLevelAdult <- NewDataOccurrenceTrophicAdult$ScaledAvgTrophicLevelAdult * sd_val + mean_val

# Plot
PlotOccurrenceTrophicAdult <- ggplot(NewDataOccurrenceTrophicAdult, aes(x = UnscaledAvgTrophicLevelAdult, y = OccurrenceProb, color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = LowerProb, ymax = UpperProb, fill = LandUseFinal), 
              alpha = 0.2, color = NA) +
  labs(x = expression("Adult trophic level"),
       y = "Probability of occurrence (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") + 
  theme_minimal() +
  scale_x_continuous(breaks = c(2,3), labels = c("Primary consumer", "Secondary consumer"))

PlotOccurrenceTrophicAdultNoLegend <- PlotOccurrenceTrophicAdult + theme(legend.position = "none") +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC"))


FinalPlotOccurrenceTrophicAdultNoLegend <- ggdraw(PlotOccurrenceTrophicAdultNoLegend) +
  draw_label("Coefficient in primary: 
   -0.07 [95% CI: -0.10, -0.05], p < 0.001
              Adult TL x LU interaction: 
  F-ratio = 71.46, p < 0.0001", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

FinalPlotOccurrenceTrophicAdult <- ggdraw(PlotOccurrenceTrophicAdult) +
  draw_label("Coefficient in primary: 
   -0.07 [95% CI: -0.10, -0.05], p < 0.001
              Adult TL x LU interaction: 
  F-ratio = 71.46, p < 0.0001", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

       #ggsave(filename = "./figures/c4_updated/AdultTrophicLevelOccurrence.png",  plot = FinalPlotOccurrenceTrophicAdult, width = 6, height = 6, dpi = 300)
```

# Create the combined plot 

```{r}
NewDataCombined <- NewDataOccurrenceTrophicAdult %>%
  bind_cols(NewDataAbundanceTrophicAdult %>% 
              dplyr::select(ExpAbundFit = ExpAbundFit,
                     AbundSE  = AbundSE)) %>%
  mutate(
    # Probability of occurrence * predicted abundance when present = predicted abundance ???
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * ExpAbundFit,
    
    # Delta method for SE of product on natural scale
    PredictedAbundanceSE = sqrt(
      (ExpAbundFit^2) * (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2) +  # variance of plogis(fit_link)
      (OccurrenceProb^2) * (ExpAbundFit^2) * (AbundSE^2)                      # variance of exp(fit_abund)
    ),
    
    # 95% confidence intervals
    PredictedAbundanceLwr = PredictedAbundanceAccountingForOccurrenceProbability - 1.96 * PredictedAbundanceSE,
    PredictedAbundanceUpr = PredictedAbundanceAccountingForOccurrenceProbability + 1.96 * PredictedAbundanceSE
  )

# Plot
FinalPlotAbundanceOccurrenceMassCombined <- ggplot(NewDataCombined,
       aes(x = UnscaledAvgTrophicLevelAdult, y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = PredictedAbundanceLwr, ymax = PredictedAbundanceUpr),
              alpha = 0.2, color = NA) +
  labs(x = expression("Adult trophic level"),
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = c(2,3), labels = c("PC", "SC")) 
#ggsave(filename = "./figures/c4_updated/AdultTrophicLevelAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceMassCombined, width = 6, height = 6, dpi = 300)
```

# Combine all three plots 

```{r}
# Combine with shared legend and auto subplot labels
AllThreeTogether <- (FinalPlotAbundanceTrophicAdultNoLegend | FinalPlotOccurrenceTrophicAdultNoLegend | FinalPlotAbundanceOccurrenceMassCombined) +
  plot_layout(guides = "collect") &    # combine legends
  plot_annotation(tag_levels = 'a') &  # adds a, b, c tags
  theme(legend.position = "bottom",
        plot.tag = element_text(size = 18, face = "bold"),
                 plot.tag.position = c(0, 1))    # shared legend position

#ggsave(filename = "./figures/c4_updated/AdultTrophicLevelAllThreeCombined.png",  plot = AllThreeTogether, width = 14, height = 7, dpi = 300)
```

# Calculate average dispersal ability and merge with PREDICTS

```{r}
# Select cols of interest   
dispersal <- traits %>% dplyr::select(c(1:8, 196:200))

# define dispersal columns
dispersal_cols <- c("Dispersal ability | Very low",
                    "Dispersal ability | Low",
                    "Dispersal ability | Medium",
                    "Dispersal ability | High",
                    "Dispersal ability | Very high")

# filter rows with dispersal data
dispersal <- dispersal %>%
  filter(if_any(all_of(dispersal_cols), ~ !is.na(.) & . > 0)) # if any dispersal columns is not NA and the value is greater than 0 - keep!

# Set up the ordinal scale for calculating average dispersal ability on a scale of 0 to 4
ordinal_scale <- c(0, 1, 2, 3, 4)

# calculate average dispersal ability
AverageDispersalFamily <- dispersal %>%
  rowwise() %>%
  mutate(
    avg_dispersal = sum(as.numeric(c_across(all_of(dispersal_cols))) * ordinal_scale, na.rm = TRUE)
  ) %>%
  ungroup() %>% 
  dplyr::select(Family, avg_dispersal)
```

# Read in PREDICTS data and ensure that species level entities in the Taxon_name_entered column such as "Lasioglossum sp." are included for merging average trophic level values - this would be at the genus level, since a taxon such as "Lasioglossum sp" could be assigned a genus but we don't know the specific species name. 

```{r}
# All insect data effort-corrected and merged
taxon_insects <-  qread("/Users/jusi/Desktop/multi-groups/data/predicts-prepared-whole-database.qs")

# Get rid of rows which aren't resolved to order level
taxon_insects <- taxon_insects %>% filter(Order != "") # 1056 rows not resolved to order

# Recreate genus and species column as sometimes BGB is present in PREDICTS but no info for genus or sp.
taxon_insects <- taxon_insects %>%
  mutate(Best_guess_binomial = as.character(Best_guess_binomial)) %>%
  separate(Best_guess_binomial, into = c("Genus", "Species"), sep = " ", remove = FALSE, extra = "merge", fill = "right")

# There are entries in PREDICTS such as "Lasioglossum sp" - rows with such names tend to not have an entry in the genus column (due to issues with the COL API)
# Previously these taxa were getting family level values in the average mass analysis, despite their genus information being available.
# I will extract the genus information from the Taxon_name_entered column for those rows that do not yet have an entry in genus.
# This will improve the precision of the matching, such that these rows will now be matched at the genus level instead of the family level (as they were beforehand)
taxon_insects <- taxon_insects %>%
  mutate(
    Genus = case_when(
      is.na(Genus) | Genus == "" ~ word(Taxon_name_entered, 1), # Check if Genus is empty or missing, if so, extract the first word from Taxon_name_entered and assign it to genus column.
      TRUE ~ Genus # If genus is not empty, leave it as the pre-existing genus entry. 
    ))

# Have a look
taxon_insects %>% dplyr::select(SS, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, Precision_score) %>% View()

# Remove those entries which have a "?" before or after the name of the genus e.g. "?Abalabus" or "Abalabus?"
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove_all(Genus, "\\?"))

# Remove leading and trailing double quotation marks
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_replace_all(Genus, '^"|"$', ""))

# Remove any numbers 
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove_all(Genus, "[0-9]"))

# Remove entries with sp at the end
taxon_insects <- taxon_insects %>%
  mutate(Genus = str_remove(Genus, "(_|\\s)?sp\\.?$"))

# Filter for columns of interest
taxon_insects <- taxon_insects %>% dplyr::select(Source_ID, SS, SSB, SSBS, Order, Family, Genus, Species, Taxon_name_entered, Best_guess_binomial, Diversity_metric, Diversity_metric_unit, Merged_diversity, Predominant_habitat, Use_intensity)
```

# Merge average dispersal ability per family into my overall data frame 

```{r}
AllInsectsDispersal <- taxon_insects %>%
  left_join(
    AverageDispersalFamily,
    by = "Family"
  )

rm(AverageDispersalFamily, dispersal, taxon_insects)
```

# Now that I have an overall data frame for average dispersal ability for all insect taxa in PREDICTS, I will now filter it for species level entities using APs taxon resolution file and then prepare the data frame for analysis.

```{r}
# Filter for diversity metrics of interest - abundance and occurrence data can be used (i.e. no species richness data)
AllInsectsDispersal <- AllInsectsDispersal %>% filter(Diversity_metric %in% c("abundance", "effort-corrected abundance", "density", "group abundance", "occurrence", "occurrence frequency", "relative abundance"))

# Load taxon resolution file
TaxonResolution <- readRDS(here("data", "CuratedTaxonResolution2.rds"))

# Step 1: Filter TaxonResolution for species-level entries - i.e. those entries with a precision score of 4!
TaxonResolutionSp <- TaxonResolution %>%
  filter(PrecisionScore == 4) %>%
  dplyr::select(SS, Taxon_name_entered) # select cols for merging

# Step 2: Keep only matching rows in AllInsectsDispersal
AllInsectsDispersalSpeciesOnly <- AllInsectsDispersal %>%
  semi_join(TaxonResolutionSp, by = c("SS", "Taxon_name_entered")) # keeps only rows in AllInsectsDispersal that have a match in TaxonResolutionSp. based on SS and Taxon_name_entered

# AllInsectsDispersalSpeciesOnly now contains only the rows from AllInsectsDispersal that correspond to species-level entries in TaxonResolutionSp. 

# Create a present or absent column - its present if Merged_diversity > 0 and absent if its 0
AllInsectsDispersalSpeciesOnly <- AllInsectsDispersalSpeciesOnly %>%
  mutate(PresentOrAbsent = ifelse(Merged_diversity > 0, 1, 0))

# Create a log(abundance when present) column
AllInsectsDispersalSpeciesOnly <- AllInsectsDispersalSpeciesOnly %>% 
  mutate(
    LogAbundanceWhenPresent = ifelse(Merged_diversity > 0,
                                    log(Merged_diversity),
                                    NA))

# Have a look at how many rows are present versus absent
table(AllInsectsDispersalSpeciesOnly$PresentOrAbsent)

# Rename predominant habitat since its not really habitat we're looking at
AllInsectsDispersalSpeciesOnly <- rename(AllInsectsDispersalSpeciesOnly,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
AllInsectsDispersalSpeciesOnly <- AllInsectsDispersalSpeciesOnly %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

table(AllInsectsDispersalSpeciesOnly$Predominant_land_use)
table(AllInsectsDispersalSpeciesOnly$Predominant_land_use, AllInsectsDispersalSpeciesOnly$Use_intensity)
```

# Summary statistics

```{r}
# How many species matched to species level data 
length(unique(AllInsectsDispersalSpeciesOnly$Best_guess_binomial[!is.na(AllInsectsDispersalSpeciesOnly$avg_dispersal)]))
#How many species in total
length(unique(AllInsectsDispersalSpeciesOnly$Best_guess_binomial))
# How many families in total
length(unique(AllInsectsTrophicLevelSpeciesOnly$Family))
```


# Recode the land use classes to have Primary, Secondary vegetation, Low-intensity agriculture and high-intensity agriculture

```{r}

AllInsectsDispersalSpeciesOnly <- AllInsectsDispersalSpeciesOnly %>%
  mutate(
    LandUseFinal = case_when(
      # Primary stays as is
      Predominant_land_use == "Primary" ~ "Primary",

      # Any secondary vegetation type -> Secondary Vegetation
      Predominant_land_use %in% c("Young secondary vegetation",
                                  "Intermediate secondary vegetation",
                                  "Mature secondary vegetation") ~ "Secondary vegetation",

      # Plantation forest depends on intensity
      Predominant_land_use == "Plantation forest" & Use_intensity == "Minimal use" ~ "Low-intensity agriculture",
      Predominant_land_use == "Plantation forest" & Use_intensity %in% c("Light use", "Intense use") ~ "High-intensity agriculture",

      # Pasture
      Predominant_land_use == "Pasture" & Use_intensity %in% c("Minimal use", "Light use") ~ "Low-intensity agriculture",
      Predominant_land_use == "Pasture" & Use_intensity == "Intense use" ~ "High-intensity agriculture",

      # Cropland
      Predominant_land_use == "Cropland" & Use_intensity %in% c("Minimal use", "Light use") ~ "Low-intensity agriculture",
      Predominant_land_use == "Cropland" & Use_intensity == "Intense use" ~ "High-intensity agriculture",

      # Drop Urban by assigning NA -these will get dropped when we create the model data
      Predominant_land_use == "Urban" ~ NA_character_
    )
  ) %>%
  # Turn into factor if you want ordered categories
  mutate(LandUseFinal = factor(LandUseFinal,
                               levels = c("Primary", "Secondary vegetation",
                                          "Low-intensity agriculture", "High-intensity agriculture")))

table(AllInsectsDispersalSpeciesOnly$LandUseFinal)
```

# Create random intercept for Species

```{r}
AllInsectsDispersalSpeciesOnly <- AllInsectsDispersalSpeciesOnly %>%
  mutate(
    SpeciesFinal = case_when(
      Species != "" & !is.na(Species) ~ Species,
      (Species == "" | is.na(Species)) & Best_guess_binomial != "" & !is.na(Best_guess_binomial) ~ Best_guess_binomial,
      (Species == "" | is.na(Species)) & (Best_guess_binomial == "" | is.na(Best_guess_binomial)) ~ paste(SS, Taxon_name_entered, sep = " "),
      TRUE ~ NA_character_
    )
  )

sum(is.na(AllInsectsDispersalSpeciesOnly$SpeciesFinal))
AllInsectsDispersalSpeciesOnly %>% filter(SpeciesFinal == "")
sum(is.na(AllInsectsDispersalSpeciesOnly$SpeciesFinal) | AllInsectsDispersalSpeciesOnly$SpeciesFinal == "")


length(unique(AllInsectsDispersalSpeciesOnly$SpeciesFinal))
length(unique(AllInsectsDispersalSpeciesOnly$Taxon_name_entered))
length(unique(AllInsectsDispersalSpeciesOnly$Best_guess_binomial))
length(unique(AllInsectsDispersalSpeciesOnly$Species))
```

# Create model data

```{r}
# drop all rows with NAs for variables of interest
ModelDataDispersalOccurrence <- drop_na(AllInsectsDispersalSpeciesOnly, LandUseFinal, Use_intensity, PresentOrAbsent, avg_dispersal) %>% droplevels()
ModelDataDispersalAbundance <- drop_na(AllInsectsDispersalSpeciesOnly, LandUseFinal, Use_intensity, LogAbundanceWhenPresent, avg_dispersal) %>% droplevels()


length(unique(ModelDataDispersalOccurrence$Best_guess_binomial))

rm(TaxonResolution, TaxonResolutionSp)
```

#### Collinearity

```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(ModelDataDispersalOccurrence[ , c("LandUseFinal", "Use_intensity", "avg_dispersal", "PresentOrAbsent")])
corvif(ModelDataDispersalAbundance[ , c("LandUseFinal", "Use_intensity","avg_dispersal", "LogAbundanceWhenPresent")])
```

# Convert to factors

```{r}
# ModelDataDispersalOccurrence
ModelDataDispersalOccurrence$Order <- as.factor(ModelDataDispersalOccurrence$Order)
ModelDataDispersalOccurrence$Family <- as.factor(ModelDataDispersalOccurrence$Family)
ModelDataDispersalOccurrence$SpeciesFinal <- as.factor(ModelDataDispersalOccurrence$SpeciesFinal)
ModelDataDispersalOccurrence$SS <- as.factor(ModelDataDispersalOccurrence$SS)
ModelDataDispersalOccurrence$SSB <- as.factor(ModelDataDispersalOccurrence$SSB)

# ModelDataDispersalAbundance
ModelDataDispersalAbundance$Order <- as.factor(ModelDataDispersalAbundance$Order)
ModelDataDispersalAbundance$Family <- as.factor(ModelDataDispersalAbundance$Family)
ModelDataDispersalAbundance$SpeciesFinal <- as.factor(ModelDataDispersalAbundance$SpeciesFinal)
ModelDataDispersalAbundance$SS <- as.factor(ModelDataDispersalAbundance$SS)
ModelDataDispersalAbundance$SSB <- as.factor(ModelDataDispersalAbundance$SSB)

#saveRDS(ModelDataDispersalAbundance, "./C4 data/ModelDataDispersalAbundance.rds")
#saveRDS(ModelDataDispersalOccurrence, "./C4 data/ModelDataDispersalOccurrence.rds")
```

# Modelling log(abundance of present species) ~ LUFinal * average dispersal ability

```{r}
# Maximal random effects structure
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * avg_dispersal + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = ModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop block
ab_m2 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * avg_dispersal + (1 | Source_ID/SS) + (1| Family/SpeciesFinal), 
               data = ModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop SS
ab_m3 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * avg_dispersal + (1 | Source_ID) + (1| Family/SpeciesFinal), 
               data = ModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Drop Family
ab_m4 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * avg_dispersal + (1 | Source_ID) + (1 | SpeciesFinal), 
               data = ModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Compare models
anova(ab_m1, ab_m2, ab_m3, ab_m4)  # Likelihood ratio test
AIC(ab_m1, ab_m2, ab_m3, ab_m4)


# ab_m1 best!
```

# Backwards stepwise selection of fixed effects

```{r}
# Additive model with the interaction removed but both predictors still present
ab_m5 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + avg_dispersal + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Just land use as predictor
ab_m6 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Just average dispersal level as the predictor
ab_m7 <- blmer(LogAbundanceWhenPresent ~ avg_dispersal + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Make a null, intercept only model
ab_m8 <- blmer(LogAbundanceWhenPresent ~ (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), data = ModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))


## Compare the models
AIC(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m1, ab_m5, ab_m6, ab_m7, ab_m8)
anova(ab_m6, ab_m1)

# ab_m1 best model
```

# ab_m1 marginally better than ab_m6 (just LU as a predictor) p = 0.04
# Check the model

```{r}
model_plot(ab_m1)
summary(ab_m1)
```

# Generate predictions from the model

```{r}
# Create a grid of values
NewDataAbundanceDispersal <- expand.grid(
  avg_dispersal = seq(from = min(ModelDataDispersalAbundance$avg_dispersal),
                   to = max(ModelDataDispersalAbundance$avg_dispersal),
                   length.out = 100),
  LandUseFinal = levels(ModelDataDispersalAbundance$LandUseFinal))
  

# Get predictions without random effects, with SEs
AbundancePredictions <- predict(ab_m1, newdata = NewDataAbundanceDispersal, re.form = NA, se.fit = TRUE)

# Add predictions and 95% CIs
NewDataAbundanceDispersal$AbundFit <- AbundancePredictions$fit
NewDataAbundanceDispersal$AbundSE  <- AbundancePredictions$se.fit
NewDataAbundanceDispersal$ExpAbundFit <- exp(NewDataAbundanceDispersal$AbundFit)  # back-transform
NewDataAbundanceDispersal$AbundLwr <- exp(NewDataAbundanceDispersal$AbundFit - 1.96 * NewDataAbundanceDispersal$AbundSE) # exponentiate log abundance values 
NewDataAbundanceDispersal$AbundUpr <- exp(NewDataAbundanceDispersal$AbundFit + 1.96 * NewDataAbundanceDispersal$AbundSE)

summary(ab_m1)

# Extract c oefficient in primary
fixef(ab_m1)["avg_dispersal"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p(parameters::p_value(ab_m1, method = "eti"))

# Extract mass x land use F ratio
anova(ab_m1)
Anova(ab_m1)

# Plot
PlotAbundanceDispersal <- ggplot(NewDataAbundanceDispersal,
       aes(x = avg_dispersal, y = ExpAbundFit,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = AbundLwr, ymax = AbundUpr),
              alpha = 0.2, color = NA) +
  #scale_x_log10(labels = scales::label_number()) +
  labs(x = expression("Dispersal ability"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = 0:4, labels = c("very low", "low", "medium", "high", "very high")) 

PlotAbundanceDispersalNoLegend <- PlotAbundanceDispersal + theme(legend.position = "none")

FinalPlotAbundanceDispersalNoLegend <- ggdraw(PlotAbundanceDispersalNoLegend) +
  draw_label("Coefficient in primary: 
  - 0.01 [95% CI: -0.05, 0.03], p = 0.59
              Dispersal ability x LU interaction: 
  F-ratio = 3.29, p = 0.02", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)


FinalPlotAbundanceDispersal <- ggdraw(PlotAbundanceDispersal) +
  draw_label("Coefficient in primary: 
  - 0.01 [95% CI: -0.05, 0.03], p = 0.59
              Dispersal ability x LU interaction: 
  F-ratio = 3.29, p = 0.02", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

       #ggsave(filename = "./figures/c4_updated/DispersalAbilityAbundanceWhenPresent.png",  plot = FinalPlotAbundanceDispersal, width = 6, height = 6, dpi = 300)
```

# Modelling LogAbundanceWhenPresent ~ LUFinal * dispersal ability * order

```{r}
# Filter for big five
BigFiveModelDataDispersalAbundance <- ModelDataDispersalAbundance %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))

# Maximal model
ab_m1 <- blmer(LogAbundanceWhenPresent ~ LandUseFinal * avg_dispersal * Order + (1 | Source_ID/SS/SSB) + (1| Family/SpeciesFinal), 
               data = BigFiveModelDataDispersalAbundance,
                   control = lme4::lmerControl(
                     optimizer = "nloptwrap",
                     optCtrl = list(algorithm = "NLOPT_LN_SBPLX", maxeval = 1e6)))

# Check the model
model_plot(ab_m1)
summary(ab_m1)
anova(ab_m1)
Anova(ab_m1)

# Calculate the dispersal range for predictions
DispersalRange <- range(BigFiveModelDataDispersalAbundance$avg_dispersal, na.rm = TRUE)

# Generate predictions
# How does abundance scale with dispersal ability for each LU * Order combo
DispersalPredictions <- ggpredict(
  ab_m1,
  terms = c(
    sprintf("avg_dispersal [%.2f:%.2f]", DispersalRange[1], DispersalRange[2]), # Generate predictions using that range
    "LandUseFinal",
    "Order"
  ))

# Plot
PlotDispersalAbilityAbundance <- ggplot(DispersalPredictions, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.15, color = NA) +
  facet_wrap(~ facet, scales = "free_y") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))  +
  labs(x = expression("Dispersal ability"),
       y = "Abundance of present species (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  scale_x_continuous(breaks = 0:4, labels = c("very low", "low", "medium", "high", "very high"))


#ggsave(filename = "./figures/AbundanceWhenPresentBigFiveDispersalAbility.png",   plot = PlotDispersalAbilityAbundance, width = 10, height = 6, dpi = 300)

rm(ab_m1, PlotLarvalTrophicLevelAbundance, LarvalTrophicLevelPredictions)
```

# Modelling PresenceOrAbsent ~ LUFinal * average dispersal ability 

# Scale average dispersal ability to help with convergence
```{r}
ModelDataDispersalOccurrence$ScaledAvgDispersal <- scale(ModelDataDispersalOccurrence$avg_dispersal, center = TRUE, scale = TRUE)

hist(ModelDataDispersalOccurrence$ScaledAvgDispersal) ; hist(ModelDataDispersalOccurrence$avg_dispersal)
```

# Random effects structure
```{r}
# Run the converged model on its own
  oc_m4 <- bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgDispersal + (1 | SS) + (1 | Genus) + (1 | SpeciesFinal), 
               data = ModelDataDispersalOccurrence, family = binomial(link = "logit"), 
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e7)))

# Save the model
#saveRDS(oc_m4, "DispersalOccurrenceModel.rds")

# Put model formulas in a list
RandomEffectsModelCalls <- list(
  oc_m1 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgDispersal + (1 | Source_ID)  + (1 | SS) + (1| Order) + (1 | Family) + (1 | Genus) + (1| SpeciesFinal), 
               data = ModelDataDispersalOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e7)
           ))
  ),
  oc_m2 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgDispersal + (1 | SS) + (1| Order) + (1 | Family) + (1 | Genus) + (1| SpeciesFinal), 
               data = ModelDataDispersalOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e7)
           ))
  ),
   oc_m3 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgDispersal + (1 | SS) + (1 | Family) + (1 | Genus) + (1 | SpeciesFinal), 
               data = ModelDataDispersalOccurrence, family = binomial(link = "logit"), 
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e7)
               ))
   ),
  oc_m4 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgDispersal + (1 | SS) + (1 | Genus) + (1 | SpeciesFinal), 
               data = ModelDataDispersalOccurrence, family = binomial(link = "logit"), 
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e7)))
))

# Run models in parallel
RandomEffectsResults <- mclapply(RandomEffectsModelCalls, function(expr) eval(expr),
                    mc.cores = 4)  # adjust cores


oc_m1 <- RandomEffectsResults[[1]]
oc_m2 <- RandomEffectsResults[[2]]
oc_m3 <- RandomEffectsResults[[3]]
oc_m4 <- RandomEffectsResults[[4]]


summary(oc_m1) # 0.004 - no convergence
summary(oc_m2) # 0.003 - no convergence
summary(oc_m3) # 0.0021 - no convergence
summary(oc_m4) # Yay - converged

# oc_m4 only model to converge
```

## Backwards stepwise selection of fixed effects 
```{r}
# Put model formulas in a list

# Additive model
FixedEffectsModelCalls <- list(
  oc_m5 = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal + ScaledAvgDispersal + (1 | SS) + (1 | Genus) + (1 | SpeciesFinal), 
               data = ModelDataDispersalOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e6)
           ))
  ),
# just LU as predictor
  oc_m6 = quote(
  bglmer(PresentOrAbsent ~ LandUseFinal + (1 | SS) + (1 | Genus) + (1 | SpeciesFinal), 
               data = ModelDataDispersalOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e6)
           ))
  ),
# just dispersal as predictor
oc_m7 <- quote(
  bglmer(PresentOrAbsent ~ ScaledAvgDispersal + (1 | SS) + (1 | Genus) + (1 | SpeciesFinal), 
               data = ModelDataDispersalOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e6)
          ))
  ),
# Make a null, intercept only model
oc_m8 <- quote(
  bglmer(PresentOrAbsent ~ + (1 | SS) + (1 | Genus) + (1 | SpeciesFinal), 
               data = ModelDataDispersalOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e6)
         ))
))

# Run models in parallel
FixedResults <- mclapply(FixedEffectsModelCalls, function(expr) eval(expr),
                    mc.cores = 4)  # adjust cores


oc_m5 <- FixedResults[[1]]
oc_m6 <- FixedResults[[2]]
oc_m7 <- FixedResults[[3]]
oc_m8 <- FixedResults[[4]]
```

# Check model convergence and compare
```{r}
summary(oc_m5) # failed to converge
summary(oc_m6) # just LU as predictor - converged
summary(oc_m7) # just dispersal as predictor - converged
summary(oc_m8) # null intercept only - converged

## Compare the models
AIC(oc_m4, oc_m6, oc_m7, oc_m8)
anova(oc_m4, oc_m6, oc_m7, oc_m8)

# oc_m4 the best
```

# Generate predictions from the model
 
```{r}
# Interaction more significant than just additive effect 
anova(oc_m4)

# Create new data frame for predictions
NewDataOccurrenceDispersal <- expand.grid(
  ScaledAvgDispersal = seq(from = min(ModelDataDispersalOccurrence$ScaledAvgDispersal, na.rm = TRUE),
                to   = max(ModelDataDispersalOccurrence$ScaledAvgDispersal, na.rm = TRUE),
                length.out = 100),
  LandUseFinal = levels(ModelDataDispersalOccurrence$LandUseFinal))

# Copy the factor levels from the model data
NewDataOccurrenceDispersal$LandUseFinal <- factor(NewDataOccurrenceDispersal$LandUseFinal, 
                              levels = levels(ModelDataDispersalOccurrence$LandUseFinal))

# Predicted probabilities on link scale (logit) with SE
OccurrencePredictions <- predict(oc_m4, newdata = NewDataOccurrenceDispersal, type = "link", se.fit = TRUE, re.form = NA)

# Add predictions into data frame
NewDataOccurrenceDispersal <- NewDataOccurrenceDispersal %>%
  mutate(
    OccurrenceFit   = OccurrencePredictions$fit,
    OccurrenceSE    = OccurrencePredictions$se.fit,
    OccurrenceLwr = OccurrenceFit - 1.96 * OccurrenceSE,
    OccurrenceUpr = OccurrenceFit + 1.96 * OccurrenceSE,
    # Back-transform to probability scale
    OccurrenceProb   = plogis(OccurrenceFit), # predicted probability - plogis() is the inverse of the logit function.
    # It takes values on the logit (link) scale (from −∞ to +∞) and transforms them into probabilities between 0 and 1.
    LowerProb = plogis(OccurrenceLwr), # 95% CI lower bound
    UpperProb = plogis(OccurrenceUpr) # 95% CI upper bound
  )

# Extract coefficient in primary
fixef(oc_m3)["ScaledAvgDispersal"]

# Significance of body mass effect in primary
bayestestR::convert_pd_to_p(parameters::p_value(oc_m3, method = "eti"))

# Extract mass x land use F ratio
anova(oc_m3)
Anova(oc_m3)

# Extract the mean and sd to add the unscaled values back into the prediction data frame for plotting
mean_val <- attr(ModelDataDispersalOccurrence$ScaledAvgDispersal, "scaled:center")
sd_val   <- attr(ModelDataDispersalOccurrence$ScaledAvgDispersal, "scaled:scale")

# Add a new column with the unscaled values:
NewDataOccurrenceDispersal$UnscaledAvgDispersal <- NewDataOccurrenceDispersal$ScaledAvgDispersal * sd_val + mean_val

# Plot
PlotOccurrenceDispersal <- ggplot(NewDataOccurrenceDispersal,
       aes(x = UnscaledAvgDispersal, y = OccurrenceProb,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = LowerProb, ymax = UpperProb),
              alpha = 0.2, color = NA) +
  labs(x = expression("Dispersal ability"),
       y = "Probability of occurrence (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = 0:4, labels = c("very low", "low", "medium", "high", "very high"))

PlotOccurrenceDispersalNoLegend <- PlotOccurrenceDispersal + theme(legend.position = "none")

FinalPlotOccurrenceDispersalNoLegend <- ggdraw(PlotOccurrenceDispersalNoLegend) +
  draw_label("Coefficient in primary: 
   -0.09 [95% CI: -0.16, -0.03], p = 0.005
              Dispersal ability x LU interaction: 
  F-ratio = 34.98, p < 0.0001", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

FinalPlotOccurrenceDispersal <- ggdraw(PlotOccurrenceDispersal) +
  draw_label("Coefficient in primary: 
   -0.09 [95% CI: -0.16, -0.03], p = 0.005
              Dispersal ability x LU interaction: 
  F-ratio = 34.98, p < 0.0001", x = 0.96, y = 0.96, hjust = 0.9, vjust = 0.9, size = 10)

      #ggsave(filename = "./figures/c4_updated/DispersalAbilityOccurrence.png",  plot = FinalPlotOccurrenceDispersal, width = 6, height = 6, dpi = 300)
```

# Combine the predictions

```{r}
NewDataCombined <- NewDataOccurrenceDispersal %>%
  bind_cols(NewDataAbundanceDispersal %>% 
              dplyr::select(ExpAbundFit = ExpAbundFit,
                     AbundSE  = AbundSE)) %>%
  mutate(
    # Probability of occurrence * predicted abundance when present 
    PredictedAbundanceAccountingForOccurrenceProbability = OccurrenceProb * ExpAbundFit,
    
    # Delta method for SE of product on natural scale
    PredictedAbundanceSE = sqrt(
      (ExpAbundFit^2) * (OccurrenceProb * (1 - OccurrenceProb))^2 * (OccurrenceSE^2) +  # variance of plogis(fit_link)
      (OccurrenceProb^2) * (ExpAbundFit^2) * (AbundSE^2)                      # variance of exp(fit_abund)
    ),
    
    # 95% confidence intervals
    PredictedAbundanceLwr = PredictedAbundanceAccountingForOccurrenceProbability - 1.96 * PredictedAbundanceSE,
    PredictedAbundanceUpr = PredictedAbundanceAccountingForOccurrenceProbability + 1.96 * PredictedAbundanceSE
  )

# Plot
FinalPlotAbundanceOccurrenceDispersalCombined <- ggplot(NewDataCombined,
       aes(x = UnscaledAvgDispersal, y = PredictedAbundanceAccountingForOccurrenceProbability,
           color = LandUseFinal, fill = LandUseFinal)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = PredictedAbundanceLwr, ymax = PredictedAbundanceUpr),
              alpha = 0.2, color = NA) +
  labs(x = expression("Dispersal ability"),
       y = "Predicted abundance = a x b (± 95 % CI)",
       color = "Land use") +
  guides(fill = "none") +
  theme_minimal() +
  scale_x_continuous(breaks = 0:4, labels = c("very low", "low", "medium", "high", "very high")) 

#ggsave(filename = "./figures/c4_updated/DispersalAbilityAbundanceOccurrenceCombined.png",  plot = FinalPlotAbundanceOccurrenceDispersalCombined, width = 6, height = 6, dpi = 300)
```

# Combine all three plots 

```{r}
# Combine with shared legend and auto subplot labels
AllThreeTogether <- (FinalPlotAbundanceDispersalNoLegend | FinalPlotOccurrenceDispersalNoLegend | FinalPlotAbundanceOccurrenceDispersalCombined) +
  plot_layout(guides = "collect") &    # combine legends
  plot_annotation(tag_levels = 'a') &  # adds a, b, c tags
  theme(legend.position = "bottom",
        plot.tag = element_text(size = 18, face = "bold"),
                 plot.tag.position = c(0, 1))    # shared legend position

#ggsave(filename = "./figures/c4_updated/DispersalAbilityAllThreeCombined.png",  plot = AllThreeTogether, width = 14, height = 7, dpi = 300)
```


# Split LUs into natural and agriculture

```{r}
AbundanceCollapsedLandUse <- model_data_abundance %>%
  mutate(
    LandUseCollapsed = case_when(
      Predominant_land_use %in% c("Primary",
                                  "Young secondary vegetation",
                                  "Intermediate secondary vegetation",
                                  "Mature secondary vegetation") ~ "natural",
      Predominant_land_use %in% c("Plantation forest",
                                  "Pasture",
                                  "Cropland") ~ "agriculture",
      TRUE ~ NA_character_   # everything else becomes NA, i.e. Urban sites
    )
  ) %>%
  filter(!is.na(LandUseCollapsed))   # drop Urban sites

```

# Merge AllInsectsSize with PREDICTS site level abundance data (don't need to do this as we are calculating SPECIES level abundances not SITE LEVEL)

```{r}
# All-insect set of site-level log(abundance) data
# Has filtered out accidentally sampled larvae
abundance <-  qread("/Users/jusi/Desktop/multi-groups/data/abundance-data-prepared-all-insects.qs")

# Use dplyr to reshape the abundance data
# For each SSBS: sum the total abundance values then take the mean log_plus_what (they'll all be the same - comes from the minimum non-zero value across the entire study) 
# logAllInsectAbundance = log(sumTotalAbundance + mean_log_plus_what) - an overall abundance value for all of the insects at each site!
abundance_summary <- abundance %>%
  group_by(SSBS) %>%
  summarise(
    sumTotalAbundance = sum(Total_abundance, na.rm = TRUE),
    mean_log_plus_what = mean(Log_plus_what, na.rm = TRUE)
  ) %>%
  mutate(logAllInsectAbundance = log(sumTotalAbundance + mean_log_plus_what))

# Filter PREDICTS for same set of sites as in my AllInsectsSize dataset
filtered_abundance <-  abundance_summary %>%
  filter(SSBS %in% AllInsectsSize$SSBS) %>%
  distinct(SSBS, .keep_all = TRUE)

# Merge abundance values into the AllInsectsSize data frame 
AllInsectsSize <- AllInsectsSize %>%
  left_join(filtered_abundance %>% dplyr::select(SSBS, logAllInsectAbundance), by = "SSBS")

rm(abundance, abundance_summary, filtered_abundance)
```

# Calculate the average mass for each SSBS

```{r}
avg_mass_by_SSBS <- AllInsectsSize %>%
  group_by(SSBS) %>%
  summarise(
    mean_mass_mg = mean(mass_mg, na.rm = TRUE)
  )

avg_mass_by_SSBS <- avg_mass_by_SSBS %>%
  left_join(
    AllInsectsSize %>%
      dplyr::select(Source_ID, SSBS, logAllInsectAbundance, SS, Predominant_habitat, SSB, Use_intensity) %>%
      distinct(),
    by = "SSBS"
  )

hist(avg_mass_by_SSBS$mean_mass_mg, breaks = 75, main = "Distribution of insect mass (mg)", xlab = "Mass (mg)")

avg_mass_by_SSBS <- avg_mass_by_SSBS %>%
  mutate(
    size_category = case_when(
      mean_mass_mg < 2 ~ "small",
      mean_mass_mg >= 2 & mean_mass_mg < 20 ~ "medium",
      mean_mass_mg >= 20 & mean_mass_mg < 200 ~ "large",
      mean_mass_mg >= 200 ~ "very_large"
    )
  )

avg_mass_by_SSBS <- avg_mass_by_SSBS %>%
  mutate(
    size_category = case_when(
      mean_mass_mg > 0 & mean_mass_mg < 10 ~ "small",
      mean_mass_mg >= 10 & mean_mass_mg < 100 ~ "medium",
      mean_mass_mg >= 100 ~ "large"
    )
  )

AllInsectsSize %>% filter(SSBS == "AD1_2009__Vergara 1  6") %>% dplyr::select(SSBS, Order, Best_guess_binomial, mass_mg, logAllInsectAbundance) %>% View()
```

# Tidy the model data 

```{r}
# Rename predominant habitat since its not really habitat we're looking at
avg_mass_by_SSBS <- rename(avg_mass_by_SSBS,
                Predominant_land_use = Predominant_habitat)

# Relevel the land use and use intensity classes
avg_mass_by_SSBS <- avg_mass_by_SSBS %>%
  mutate(
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    # indeterminate secondary veg and cannot decide get NA
   Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(avg_mass_by_SSBS$Predominant_land_use, avg_mass_by_SSBS$Use_intensity)

levels(avg_mass_by_SSBS$Predominant_land_use)
levels(avg_mass_by_SSBS$Use_intensity)
```

# Create LUI (LU + UI)

```{r}
# Append Predominant_land_use and Use_intensity
avg_mass_by_SSBS <- avg_mass_by_SSBS %>%
  mutate(
    LUI = str_c(Predominant_land_use, Use_intensity, sep = "_"),
    LUI = str_replace_all(LUI, "\\s+", "_") 
  )

# MSV intense use has 6 sites, merge with light use
# ISV intense use has 61 sites, merge with light use
avg_mass_by_SSBS <- avg_mass_by_SSBS %>%
  mutate(LUI = recode_factor(LUI,
    "Mature_secondary_vegetation_Intense_use" = "Mature_secondary_vegetation_High_use",
    "Mature_secondary_vegetation_Light_use"   = "Mature_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Intense_use" = "Intermediate_secondary_vegetation_High_use",
    "Intermediate_secondary_vegetation_Light_use" = "Intermediate_secondary_vegetation_High_use"
  ))

# Set the reference levels for LUI
avg_mass_by_SSBS <- avg_mass_by_SSBS %>%
  mutate(
    LUI = factor(
      LUI,
      levels = c("Primary_Minimal_use", "Primary_Light_use", "Primary_Intense_use",
                 "Young_secondary_vegetation_Minimal_use", "Young_secondary_vegetation_Light_use", "Young_secondary_vegetation_Intense_use",
                 "Intermediate_secondary_vegetation_Minimal_use", "Intermediate_secondary_vegetation_High_use", 
                 "Mature_secondary_vegetation_Minimal_use",     "Mature_secondary_vegetation_High_use", 
                 "Plantation_forest_Minimal_use", "Plantation_forest_Light_use", "Plantation_forest_Intense_use",
                 "Pasture_Minimal_use", "Pasture_Light_use", "Pasture_Intense_use",
                 "Cropland_Minimal_use", "Cropland_Light_use","Cropland_Intense_use",
                 "Urban_Minimal_use", "Urban_Light_use", "Urban_Intense_use")
    )
  )
levels(avg_mass_by_SSBS$LUI)
table(avg_mass_by_SSBS$LUI)
```

# Create model data for abundance

```{r}
# drop all rows with NAs for land use, use intensity and log_abundance_px and drop empty levels
model_data_abundance <- drop_na(avg_mass_by_SSBS, Predominant_land_use, Use_intensity, logAllInsectAbundance) %>% droplevels() 

# check distributions
hist(model_data_abundance$logAllInsectAbundance)

# Have a look at sample sizes
table(model_data_abundance$Predominant_land_use, model_data_abundance$Use_intensity)
table(model_data_abundance$LUI)

length(unique(model_data_abundance$SS)) # 444 studies 
length(unique(model_data_abundance$SSBS)) # 11516 sites
```

# Modelling abundance ~ LU * Mass

```{r}
model <- brm(
  formula = logAllInsectAbundance ~ Predominant_land_use * size_category + (1 | SS) + (1 | SSB), 
  data = model_data_abundance, family = gaussian(), chains = 4, cores = 4, iter = 10000)

summary(model)

hist(log(test$Merged_diversity) + 1)

test <- AllInsectsSize %>% filter(Diversity_metric == "occurrence")

test %>% dplyr::select(SSBS, Order, Best_guess_binomial, Diversity_metric, Diversity_metric_unit, Diversity_metric_type, Merged_diversity, log10_mass, logAllInsectAbundance) %>% distinct() %>% View()

hist(test$Effort_corrected_measurement)


```


# Simulate a single trait and its probability

```{r}
OrdersPivot %>% filter(Order == "Hemiptera") %>% filter(Trait == "Dispersal ability ") %>% filter(Family == "Nabidae")

HymMinSize <- OrdersPivot %>% filter(Order == "Hymenoptera") %>% filter(Trait == "Minimum size ") 

HymMinSizeFreq <- HymMinSize %>% group_by(Mode) %>% count() 

HymSizeTest <- c(
  rep("< 0.35mm", 2),
  rep("< 0.71mm", 8),
  rep("< 1.41mm", 11),
  rep("< 2.82mm", 8),
  rep("< 5.62mm", 28),
  rep("< 11.22mm", 21),
  rep("< 22.39mm", 4),
  rep("< 44.67mm", 1)
)

HymSizeTest <- factor(HymSizeTest, levels = c("< 0.35mm", "< 0.71mm", "< 1.41mm", "< 2.82mm", "< 5.62mm", "< 11.22mm", "< 22.39mm", "< 44.67mm"))

s <- sample(HymSizeTest, size = 1000, replace = T)

table(s)

mean(table(s))

trait_values <- c("low", "med", "high", "very high")
trait_probs  <- c(0.42, 0.26, 0.11, 0.21)  # must sum to 1
n_draws <- 1000

c <- readRDS(file.choose())

OrdersPivot %>% filter(Order == "Coleoptera") %>% filter(Trait == "Dispersal ability ") %>% View()

OrdersPivot %>% filter(Order == "Coleoptera") %>% filter(Trait == "Lifelong fecundity ") %>% View()


OrdersPivot %>% filter(Order == "Coleoptera") %>% filter(Trait == "Adult feeding guild ") %>% distinct(Family) %>% View()

traits %>% dplyr::select(Order, Family, "Lifelong fecundity | < 100 | Number of eggs per female",   "Lifelong fecundity | 100-1000 | Number of eggs per female", 
"Lifelong fecundity | 1000-3000 | Number of eggs per female", "Lifelong fecundity | > 3000 | Number of eggs per female") %>% View()

traits %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Lepidoptera", "Hymenoptera", "Odonata", "Orthoptera")) %>% dplyr::select(Order, Family, "Dispersal ability | Very low", "Dispersal ability | Low", "Dispersal ability | Medium", "Dispersal ability | High", "Dispersal ability | Very high") %>% View()

simulated_traits <- sample(
  x = trait_values,
  size = 100,
  replace = TRUE,
  prob = trait_probs
)


dispersal_test <- c(
  rep("low", 42),
  rep("med", 26),
  rep("high", 11),
  rep("very high", 21)
)

dispersal_test <- factor(
  dispersal_test,
  levels = c("low", "med", "high", "very high"),
  ordered = TRUE
)

hist(table(dispersal_simulation))

dispersal_simulation <- sample(x = dispersal_test, size = 1000, replace = TRUE)

table(dispersal_simulation)

mean(dispersal_simulation)

x <- sample(x = c("Herbivore"), size = 1000, replace = TRUE, prob = c(0.33))

mean(table(x))

prop.table(table(AdultDiet$Mode)) * 100

hist(b$simulated_traits)


b <- as.data.frame(table(simulated_traits))


table(sample(x = "low", size = 1000, replace = TRUE, prob = 0.42))


# Inspect results
table(simulated_traits)

mean(table(simulated_traits))

# Plot original distribution
barplot(height = trait_probs, names.arg = trait_values,
        col = "lightblue", main = "Dispersal ability distribution for Nabidae (Hemiptera)",
        ylab = "Probability / Frequency", ylim = c(0, max(trait_probs)*1.2))
```

# Summary of traits

```{r}
# Filter for orders of interest
OrdersWeWant <- traits %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Lepidoptera", "Hymenoptera", "Odonata", "Orthoptera"))

# Identify all traits present for a group
OrdersWeWant %>%
  pivot_longer(cols = 9:ncol(.), # all trait columns
               names_to = "Trait", # new column for trait names
               values_to = "Value") %>%# new column for the values
  filter(!is.na(Value) & Value != 0) %>%  # keep only traits that are present, i.e are not NA or 0
View()

# Pivot longer traits
OrdersPivot <- OrdersWeWant %>%
  pivot_longer(
    cols = 9:ncol(.), # first 8 cols are meta data
    names_to = "Trait",
    values_to = "Value"
  ) %>%
  filter(!is.na(Value) & Value != "0") %>% # if the trait value is not an NA and is not equal to 0
  separate(Trait, into = c("Trait", "Modality", "Unit"), sep = "\\|", fill = "right", extra = "merge") %>% 
  mutate(Mode = ifelse(is.na(Unit), Modality, paste(Modality, Unit))) %>% # what is the trait category
  dplyr::select(-Modality, -Unit) %>% # remove redundant cols
  dplyr::select(-Value, -Taxon_Code, -SubOrder, -InfraOrder, -SuperFamiy, -`Taxon Name`, -`Approx Species Richness`,  Value)   # moves 'Value' to the end
```


```{r}
# For each model data - filter for the big five
BigFiveModelBodyMassOccurrence <- ModelDataSizeOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))
BigFiveModelEggVolumeOccurrence <- ModelDataEggOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))
BigFiveModelAdultDietOccurrence <- ModelDataAdultDietOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))
BigFiveModelDataLarvalDietOccurrence <- ModelDataLarvalDietOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))
BigFiveModelDataDispersalOccurrence <- ModelDataDispersalOccurrence %>% filter(Order %in% c("Coleoptera", "Diptera", "Hemiptera", "Hymenoptera", "Lepidoptera"))

# Scale the predictors
BigFiveModelEggVolumeOccurrence$ScaledLogAvgEggVolume_mm3 <- scale(BigFiveModelEggVolumeOccurrence$LogAvgEggVolume_mm3, center = TRUE, scale = TRUE)

BigFiveModelDataDispersalOccurrence$ScaledAvgDispersal <- scale(BigFiveModelDataDispersalOccurrence$avg_dispersal, center = TRUE, scale = TRUE)

BigFiveModelDataLarvalDietOccurrence$ScaledAvgTrophicLevel_larval <- scale(BigFiveModelDataLarvalDietOccurrence$AvgTrophicLevel_larval, center = TRUE, scale = TRUE)

BigFiveModelAdultDietOccurrence$ScaledAvgTrophicLevelAdult <- scale(BigFiveModelAdultDietOccurrence$AvgTrophicLevelAdult, center = TRUE, scale = TRUE)

BigFiveModelBodyMassOccurrence$ScaledLog10Mass <- scale(BigFiveModelBodyMassOccurrence$log10_mass, center = TRUE, scale = TRUE)


# Put model formulas in a list
model_calls <- list(
  EggVolumeOccurrenceModel = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledLogAvgEggVolume_mm3 * Order +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = BigFiveModelEggVolumeOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "Nelder_Mead",
             optCtrl = list(maxfun = 2e5)
           ))
  ),
LarvalDietOccurrenceModel = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgTrophicLevel_larval * Order + (1 | SS) + (1| SpeciesFinal), 
               data = BigFiveModelDataLarvalDietOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "Nelder_Mead",
                 optCtrl = list(maxfun = 7e5)
                 ))
  ),
  AdultDietOccurrenceModel = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgTrophicLevelAdult * Order + (1 | Source_ID) + (1 | SpeciesFinal), 
               data = BigFiveModelAdultDietOccurrence, family = binomial(link = "logit"),
  control = glmerControl(
    optimizer = "nloptwrap",
    optCtrl = list(
      algorithm = "NLOPT_LN_NELDERMEAD",
      maxeval = 1e4)
  ))
),
  DispersalOccurrenceModel = quote(
    bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgDispersal * Order + (1 | SS) + (1 | Genus) + (1 | SpeciesFinal), 
               data = BigFiveModelDataDispersalOccurrence, family = binomial(link = "logit"), 
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e7)
                 ))
  ),
 BodyMassOccurrenceModel = quote(
   bglmer(PresentOrAbsent ~ LandUseFinal * log10_mass * Order +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = BigFiveModelBodyMassOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "bobyqa",
             optCtrl = list(maxfun = 200000)
             ))
))



# Run models in parallel
results <- mclapply(model_calls, function(expr) eval(expr),
                    mc.cores = 4)  # adjust cores


results <- mclapply(model_calls, function(expr) {
  try(eval(expr), silent = TRUE)
}, mc.cores = 5)



library(future.apply)

# Use separate R sessions instead of forks
plan(multisession, workers = 5)

results <- future_lapply(model_calls, function(expr) {
  library(blme)
  library(lme4)
  try(eval(expr), silent = TRUE)
})

eval(model_calls[[1]])

oc_m1 <- results[[1]]
oc_m2 <- results[[2]]

summary(oc_m1)
summary(oc_m2)

# Compare models
anova(oc_m1, oc_m2)  # Likelihood ratio test
AIC(oc_m1, oc_m2)


 EggVolumeOccurrenceModel <- bglmer(PresentOrAbsent ~ LandUseFinal * ScaledLogAvgEggVolume_mm3 * Order +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = BigFiveModelEggVolumeOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "nlminbwrap",
             optCtrl = list(maxfun = 7e5)))
 
summary(EggVolumeOccurrenceModel)

LarvalDietOccurrenceModel <- bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgTrophicLevel_larval * Order + (1 | SS) + (1| SpeciesFinal), 
               data = BigFiveModelDataLarvalDietOccurrence, family = binomial(link = "logit"),
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 7e5)
                 ))


 BodyMassOccurrenceModel <- bglmer(PresentOrAbsent ~ LandUseFinal * ScaledLog10Mass * Order +
             (1 | Source_ID) + (1 | SpeciesFinal),
           data = BigFiveModelBodyMassOccurrence,
           family = binomial(link = "logit"),
           control = lme4::glmerControl(
             optimizer = "Nelder_Mead",
             optCtrl = list(maxfun = 7e5)),
  cov.prior = gamma(2, 0.5))
 
   DispersalOccurrenceModel <- bglmer(PresentOrAbsent ~ LandUseFinal * ScaledAvgDispersal * Order + (1 | SS) + (1 | Genus) + (1 | SpeciesFinal), 
               data = BigFiveModelDataDispersalOccurrence, family = binomial(link = "logit"), 
               control = lme4::glmerControl(
                 optimizer = "bobyqa",
                 optCtrl = list(maxfun = 1e7)
                 ))

  allFit(show.meth.tab=TRUE)
  
  af1 <- lme4::allFit(BodyMassOccurrenceModel)
  
  saveRDS(af1, file = "AllFitsDispersalOccurrenceModels.rds")
  
  b <- readRDS(file.choose())
  
  round(sapply(AllEggModels, fixef), 3)
  
  summary(a) ; summary(DispersalOccurrenceModel)
  
  af1 <- lme4::allFit(DispersalOccurrenceModel2)
  
  summary(b)
  
  AllInsectsDispersal %>% group_by(Order, Family) %>% summarise(mean(avg_dispersal)) %>% View()
    
    filter(Family %in% c("Papilionidae", "Aphididae")) %>% View()
  
  saveRDS(BigFiveModelDataDispersalOccurrence, file = "BigFiveModelDataDispersalOccurrence.rds")
  
  summary(LarvalDietOccurrenceModel)
  summary(LarvalDietOccurrenceModel2)
```

