---
title: "Untitled"
author: "Justin Isip"
date: "2023-08-22"
output: html_document
---

# This Rmd takes the latest diversity extract (30th August 2023) and checks for any additional studies relevant for my analysis.


### Clear the workspace
```{r}
rm(list=ls())
```

### Load required packages
```{r}
library(tidyverse) # data processing
library(lme4) # for mixed effects models
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(sjPlot) # for visualising results
library(effects) # for extracting model effects
library(merTools) # useful for a few things, but we're using it for extracting estimates for plotting
library(emmeans) # testing multiple comparisons (but also useful for plotting)
library(coefplot) # visualising model coefficients
library(cli)
library(ggeffects)
library(ggpubr)
library(lmerTest)
library(influence.ME)
library(robustlmm)
library(performance)
library(StatisticalModels)
```

### Load in model plot function
```{r}

## Load in a model criticism function - this function is used for checking the residual/model diagnostic plots to test that the model is meeting all of its assumptions
model_plot <-function(mod.for.plot){
  require(lattice)
  
  # set up a 2 x 2 grid for plotting
  par(mfrow = c(2,2))
  par(ask = TRUE)
  
  # qqplot
  qqnorm(resid(mod.for.plot))
  qqline(resid(mod.for.plot), col = 2)
  
  # residuals vs fitted plot
  plot(fitted(mod.for.plot), resid(mod.for.plot),xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs fitted")
  abline(h=0, lty=2)
  lines(smooth.spline(fitted(mod.for.plot), resid(mod.for.plot)), col = "red")
  
  # histogram of residuals
  hist(resid(mod.for.plot))
  
  # random effects distribution
  dotplot(ranef(mod.for.plot,condVar = TRUE))
}
```

### Load add_order_fold function to split the diversity data into exclusive insect orders 
```{r}

add_order_fold <- function(data, folds = "default", verbose = TRUE){
  
  # By default (and nothing else is yet coded up), this puts each row into
  # an insect order that each have a decent, number of rows of data. 
  # This allows site-level values to be obtained within
  # each of the order folds in turn.
  
  # Set up order_fold, a factor that holds the fold identity.
  data$order_fold <- rep(NA, nrow(data))
  
  # Rows are assigned to folds based on the higher taxonomy information.
  if (folds == "default"){
    data$order_fold[data$Higher_taxon == "Diptera"] <- "Diptera"
    data$order_fold[data$Higher_taxon == "Coleoptera"] <- "Coleoptera"
    data$order_fold[data$Higher_taxon == "Lepidoptera"] <- "Lepidoptera"
    data$order_fold[data$Higher_taxon == "Hymenoptera"] <- "Hymenoptera"
    data$order_fold[data$Higher_taxon == "Hemiptera"] <- "Hemiptera"
    data$order_fold[data$Higher_taxon == "Arachnida"] <- "Arachnida"
    data$order_fold[data$Higher_taxon == "Odonata" |
                    data$Higher_taxon == "Trichoptera" |
                    data$Higher_taxon == "Ephemeroptera" |
                    data$Higher_taxon == "Plecoptera" ] <- "Aquatics"
    data$order_fold[data$Phylum == "Arthropoda" & 
                      data$Higher_taxon != "Diptera" & 
                      data$Higher_taxon != "Coleoptera" &
                      data$Higher_taxon != "Lepidoptera" &
                      data$Higher_taxon != "Hymenoptera" &
                      data$Higher_taxon != "Hemiptera" &
                      data$Higher_taxon != "Arachnida" &
                      data$Higher_taxon != "Odonata" &
                      data$Higher_taxon != "Trichoptera" &
                      data$Higher_taxon != "Ephemeroptera" &
                      data$Higher_taxon != "Plecoptera"] <- "OtherArthropods"
  }else{
    stop("Only the default folds have been coded up so far!")
  }
  
  data$order_fold <- as.factor(data$order_fold)
  
  if (verbose == TRUE){
    # By default, report on the breakdown of records per fold.
    cat("Numbers of rows within each order fold are as follows:\n\n")
    print(table(data$order_fold))
  }
  
  return(data)
}

```

### Load get_site_abundances_order function to calculate total abundance for each site x order_fold combination
```{r}
# Calculate total abundance for each site --------------------------------------
# Modified to clarify reporting is on site x fold combinations
get_site_abundances_order <- function(data, 
                                order_folds = FALSE,
                                verbose = TRUE){
  # This calculates site metrics for each site or, if desired, each order fold
  # within each site.
  
  # Make it a data frame to avoid having to deal with a tibble.
  data <- as.data.frame(data)
  
  # Set up a variable for uniquely identifying the site x order_fold combination
  # or, if order_folds == FALSE, the site. This is what grouping is then done
  # by.
  print(table(data$order_fold))
  
  if (order_folds == TRUE){
    data$the_order_folds <- data$order_fold
  }else{
    data$the_order_folds <- ""
  }
  
  # Just for debugging
  print(table(data$the_order_folds))
  
  # Start to calculate site-level metrics.
  sites <- data %>%
    
    # pull out only the merged diversity data
    distinct(merge_ID, .keep_all = TRUE) %>%
    
    # Re-make SSB and SSBS values since we've now dropped a bunch of values; and
    # also add SSBST which is that order x site combo. The trimws is because the
    # SSBST would otherwise end in a space if order_folds == FALSE.
    mutate(SS = paste(Source_ID, Study_number),
           SSB = paste(SS, Block),
           SSBS = paste(SSB, Site_number),
           SSBST = trimws(paste(SSBS, the_order_folds))) %>%
    
    # group by SSBST (each unique value corresponds to a unique site x fold)
    group_by(SSBST) %>%
    
    # Now add up all the abundance measurements within each site and also 
    # estimate coverage.
    mutate(TotalAbundance = ifelse(Diversity_metric_type == "Abundance",
                                   sum(merged_diversity, na.rm = TRUE),
                                   # If the diversity metric type isn't 
                                   # Abundance, then leave the TotalAbundance 
                                   # measurement as NA.
                                   NA),
           
           SpeciesRichness = ifelse(Diversity_metric_type == "Species richness",
                                    merged_diversity,
                                    # for abundance and occurrence measurements,
                                    # count the number of unique species names 
                                    # that are present at the site 
                                    n_distinct(Best_guess_binomial[merged_diversity > 0])),
           
           coverage = get.SC2(merged_diversity)) %>%
    
    # ungroup
    ungroup() %>%
    
    # pull out unique site/fold combinations.
    distinct(SSBST, .keep_all = TRUE) %>%
    
    # now group by Study ID and fold.
    group_by(SS, the_order_folds) %>%
    
        # pull out the maximum abundance for each study/fold combination.
    mutate(MaxAbundance = max(TotalAbundance),
           MinNonZero = min(TotalAbundance[TotalAbundance > 0])) %>%
    
    # ungroup.
    ungroup() %>%
    
    # now rescale total abundance, so that within each study, the maximum is 1.
    mutate(RescaledAbundance = TotalAbundance/MaxAbundance) %>%
    
    mutate(sqrtRescaledAbundance = sqrt(RescaledAbundance)) 
  
    # If a study reports abundance as a count, add 1 to all values; otherwise
  # add 1/2 the minimum non-zero value; then log-transform to get a useful
  # logAbundance value that may be more symmetric than sqrtRescaledAbundance.
  sites$to_add <- ifelse(sites$Diversity_metric_unit %in%
                           c("individuals", "times observed"), 1,
                         sites$MinNonZero/2)
  sites$logAbundance <- log(sites$TotalAbundance + sites$to_add)
  
  
  # Remove NA values and drop levels
  sites <- as.data.frame(sites)
  sites <- droplevels(subset(sites, !is.na(sqrtRescaledAbundance)))
  
  if (verbose == TRUE){
    # Summarise structure of site-level data frame
    cat("\nThere are abundance data for:\n")
    cat(paste("   ", sum(!is.na(sites$sqrtRescaledAbundance)), "sites, from\n"))
    cat(paste("   ", 
              length(unique(sites$SS[!is.na(sites$sqrtRescaledAbundance)])), 
              "studies and\n"))
    cat(paste("   ", 
              length(unique(sites$Source_ID[!is.na(sites$sqrtRescaledAbundance)])),
              "sources.\n"))
  }
  
  # Set provenance attributes
  attr(sites, which = "diversity_extract") <- 
    attr(data, which = "diversity_extract")
  attr(sites, which = "is_merged") <- attr(data, which = "is_merged")
  attr(sites, which = "order_folds") <- order_folds 
  attr(sites, which = "when_Created") <- Sys.time()
  
  return(sites)
}

```

### Load all other functions
```{r}
source("/Applications/PhD/Andy/data_prep_functions.R")
source("/Applications/PhD/Andy/site_comparison_functions.R")
source("https://highstat.com/Books/Book2/HighstatLibV10.R") # collinearity
source("/Applications/PhD/Projects/multi-groups/scripts/combine_and_plot_function.R")
source("/Applications/PhD/Projects/multi-groups/scripts/predict_this_model.R")
```

### Read in the old and new diversity extract
```{r}
diversity <- readRDS("./data/diversity-2022-04-13-02-33-10.rds")
updated_diversity <- readRDS("./data/diversity-2023-08-30-02-33-17.rds")
```

### Correct for sampling effort
```{r}

if(check_sampling_effort_nas(diversity) ==
   FALSE) stop("Some studies have NA for sampling effort in only some sites!")

if(check_sampling_effort_nas(updated_diversity) ==
   FALSE) stop("Some studies have NA for sampling effort in only some sites!")

diversity <- correct_for_sampling_effort(diversity)

# Using the correct_for_sampling_effort() function loses rows for some reason.. 
#updated_diversity <- correct_for_sampling_effort(updated_diversity)

```

# Run it manually through the correct_for_sampling_effort function as I seem to lose data otherwise
```{r}

  div <- attr(updated_diversity, which = "diversity_extract")

  updated_diversity <- as.data.frame(updated_diversity) 

    updated_diversity$Sampling_effort[is.na(data$Sampling_effort)] <- 1
  
    updated_diversity <- updated_diversity %>%

    # Group by Study.
    group_by(SS) %>%
    
    # Check how many sampling efforts there are in each study.
    mutate(n_sample_effort = n_distinct(Sampling_effort),
           # Get the maximum sampling effort for the studies
           max_sample_effort = max(Sampling_effort)
    ) %>%
    
    ungroup() %>%
    
    # If the study has more than one sampling effort, correct the abundance
    
    # So if there's only one sampling effort, then create a 'dummy sampling 
    # effort' of 1 so that we don't change the abundances when we do the 
    # divisions. Otherwise, we give it the maximum sampling effort.
    mutate(DividingEffort = ifelse(n_sample_effort == 1,
                                   1, max_sample_effort)) %>%
    
    # If the diversity metric isn't sensitive to the effort, then we'll change
    # the value to 1 too (so we won't end up changing the measurement), 
    # otherwise leave it as it is.
    mutate(DividingEffort = ifelse(Diversity_metric_is_effort_sensitive == 
                                     FALSE, 1, DividingEffort)) %>%
    
    # Now let's create the effort-corrected measurement by dividing the
    # abundances by the sampling efforts.
    mutate(Corrected_sampling_effort = Sampling_effort/max_sample_effort,
        Effort_corrected_measurement = Measurement * Corrected_sampling_effort) 
```

### Merge sites
```{r}
merged_old <- merge_sites(diversity, verbose = TRUE) 
merged_new <- merge_sites(updated_diversity, verbose = TRUE) 
```

# Filter merged for five orders of interest and then subset for studies that look at more than one of those orders
```{r}
# Start from the merged data not the model data
five_orders <-
  merged_new %>%
  filter(Higher_taxon %in% c("Diptera", "Coleoptera", "Lepidoptera", "Hymenoptera", "Hemiptera")) %>% 
  # filter here for the orders of interest so that when we group by study, we are only looking at studies that observe those five orders
  group_by(SS) %>% # group by SS
  mutate(
    Multiple_groups = 
      case_when(
        (length(unique(Higher_taxon))) > 1 ~ "YES",
        (length(unique(Higher_taxon))) == 1 ~ "NO")) %>%  
  filter(Multiple_groups == "YES") %>% 
  droplevels()

# five_orders only contains studies that look at more than one of those five orders

levels(five_orders$Higher_taxon)
table(five_orders$Higher_taxon)
```

### Add_order_fold to allow taxonomic cross-validation 
```{r}
# Just adds order_fold column
five_orders <- add_order_fold(five_orders, verbose = TRUE)
levels(five_orders$order_fold)
```

### Calculate site level diversity metrics using get_site_abundances_order and order_fold for each site x order combination
```{r}

five_orders <- get_site_abundances_order(five_orders, 
                                  order_folds = TRUE, verbose = TRUE)

table(five_orders$order_fold) # number of rows for each site x order combination

# We have 8065 sites for the five orders in multi group studies

```

### Fix up your explanatory variables 
```{r}

# Rename predominant habitat since its not really habitat we're looking at
five_orders <- rename(five_orders,
                Predominant_land_use = Predominant_habitat)


# Relevel the land use and use intensity classes
five_orders <- five_orders %>%
  
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    
    # indeterminate secondary veg and cannot decide get NA
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"), 
    Use_intensity = na_if(Use_intensity, "Cannot decide"),
    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )

# take a look at the LandUse/Use intensity split
table(five_orders$Predominant_land_use, five_orders$Use_intensity)
```

# Collinearity
```{r}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")
corvif(five_orders[ , c("Predominant_land_use", "Use_intensity")])
```

### Complete cases
```{r}

# This will drop all the rows with NAs for abundance, land use or use intensity so we have complete data for modelling
model_data_ab <- drop_na(five_orders, 
                         RescaledAbundance, Predominant_land_use,
                         Use_intensity)

model_data_sr <- drop_na(five_orders, 
                         SpeciesRichness, Predominant_land_use,
                         Use_intensity)

#sum(is.na(five_orders$RescaledAbundance))
#sum(is.na(five_orders$SpeciesRichness))
#sum(is.na(five_orders$Predominant_land_use))
# There are no NAs for either of our explanatory variables so model_data_ab and model_data_sr are the same dataframe
```

### Prepare the abundance model dataframe
```{r}

# No longer needed as lme4 has improved, can use logAbundance provided in get_site_abundances()
# model_data_ab$logRescaledAbundance <- log(model_data_ab$RescaledAbundance + 1)
# sum(model_data_ab$RescaledAbundance == 1)

# Check the distributions
hist(model_data_ab$RescaledAbundance) ; hist(model_data_ab$logAbundance) ; hist(model_data_ab$sqrtRescaledAbundance)

plot(density(model_data_ab$logAbundance)) 

# Have a look at sample sizes
table(model_data_ab$Predominant_land_use, model_data_ab$Use_intensity)

# Collapse intense and light land use intensities to have a more even spread of the data
model_data_ab <- 
  model_data_ab %>%
  mutate(
    Use_intensity = recode_factor(Use_intensity,
                                  "Intense use" = "High_use",
                                  "Light use" = "High_use"),
# reset reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )
# Have a look at sample sizes
table(model_data_ab$Predominant_land_use, model_data_ab$Use_intensity)
```

# See if there is any additional studies for my analysis in the latest diversity extract
```{r}
# This is the model_data_ab from the old diversity extract - from a separate script, naughty!
compare <- model_data_ab 

# This is the model_data_ab from the latest diversity extract
a <- model_data_ab %>% dplyr::select(Source_ID, SS, Sampling_method) %>% unique()

# Simplify the old extract
b <- compare %>% dplyr::select(Source_ID, SS, Sampling_method) %>% unique()

# Which data is in the latest diversity extract which isn't in the old diversity extract
new_data <- anti_join(b, a, by = c("Source_ID", "SS", "Sampling_method"))
```

