---
title: "Modelling taxon richness"
author: "Andy Purvis"
date: "`r Sys.Date()`"
output: html_document
---

This markdown models how species richness (relative to that in primary 
vegetation) is affected by land use in each biogeographic realm, when the 
effect of the taxonomic level of taxa in PREDICTS is controlled for. To ensure
adequate sample sizes, some land-use/realm combinations are collapsed together.

The result from this markdown is a data frame of relative species richness
in each land-use class in each realm (with model coefficients being repeated
as necessary).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r load-packages}
library(blme)
library(glmmTMB)
library(DHARMa)
library(qqplotr)
library(performance)
library(MuMIn)
library(tictoc)
library(sjPlot)
library(dplyr)
library(afex)
library(ggeffects)
library(ggplot2)
library(GGally)
```

```{r function}
# test for overdispersion: from Bolker GLMM FAQ
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

```

## Site-level data frame
The file `abundance-model-data-prepared.qs` has `Taxon_richness` values along
with all the requisite explanatory variables. However, it needs wrangling before
it is used in modelling, to ensure sample sizes are large enough for reliable
modelling (using PREDICTS' standard threshold of 100 sites); the wrangling also
drops the few sites lacking `Taxon_fold_ss` (which would otherwise be in some
but not all models, confounding AIC comparisons) and all the studies that have
only a single species. Timber sites are also dropped. Here is a tabulation of 
sites per PREDICTS land-use class x Realm combination.

```{r load-data}
ds <- qs::qread("data/abundance-model-data-prepared.qs") %>%
  subset(!is.na(LUH2) & 
           LUH2 != "Timber" &
           !is.na(Taxon_fold_ss)) %>%
  droplevels()

# Drop any studies where the maximum taxon richness is 1 (i.e., single-species 
# studies)

max_tr <- tapply(ds$Taxon_richness, ds$SS, max, na.rm = TRUE)
to_drop <- names(max_tr[max_tr == 1])

cat(paste("Input file prepared on ",
          attr(ds, which = "date_created"),
          "\n"))
cat(paste("Database extract dated ",
          attr(ds, which = "date_database_extract"),
          "\n"))
cat("Single-species studies dropped from the data set:\n")
print(to_drop)
ds <- ds %>%
  filter(! SS %in% to_drop) %>%
  droplevels()

cat(paste("Number of studies remaining",
          length(unique(ds$SS)),
          "\n"))
cat(paste("Number of sites",
          length(unique(ds$SSBS)),
          "\n"))
cat(paste("Number of countries",
          length(unique(ds$Country)),
          "\n"))

summary_table <- ds %>%
  group_by(LUH2, Realm) %>%
  summarise(
    n_rows = n(),
    n_unique_SS = n_distinct(SS),
    .groups = "drop"
  ) %>%
  mutate(cell_value = paste0(n_rows, " / ", n_unique_SS)) %>%  # combine counts
  select(LUH2, Realm, cell_value) %>%
  tidyr::pivot_wider(names_from = Realm, values_from = cell_value)

# Display the table
knitr::kable(summary_table)

# Take a note of the set of realms in the data: used later on
realm_set <- unique(ds$Realm)
```

As agreed with Simon Ferrier, we lump nitrogen-fixing cropland in with annual
cropland. Although the CSIRO data do not differentiate age classes in secondary 
vegetation, they have requested separate estimates for the three. There is no 
class for Timber.

```{r lump-classes}
# Group NFX and ANN together
levels(ds$LUH2)[levels(ds$LUH2) %in% c("Annual", "Nitrogen_fixing")] <- "Annual"

summary_table <- ds %>%
  group_by(LUH2, Realm) %>%
  summarise(
    n_rows = n(),
    n_unique_SS = n_distinct(SS),
    .groups = "drop"
  ) %>%
  mutate(cell_value = paste0(n_rows, " / ", n_unique_SS)) %>%  # combine counts
  select(LUH2, Realm, cell_value) %>%
  tidyr::pivot_wider(names_from = Realm, values_from = cell_value)

# Display the table
knitr::kable(summary_table)
```

Sample sizes are generally too small in Oceania, and there are a few other 
combinations where there are fewer than 100 sites and/or fewer than 10 studies. 
These are fixed by collapsing in what is hopefully a reasonable way, and also 
taking a note of what the new names are made up of (used later on). Oceania is 
always lumped with Indo-Malay (into a combined "realm" labelled "IMO"), and 
sometimes also with Australasia (to give "IMOA"). Nearctic and Palearctic are 
also occasionally lumped (as "Holarctic").

```{r lump-realms}
# Not enough Oceania data to model so lump with Indo-Malay (rather than with
# Australasia, as Oceania and Indo-Malay are mostly tropical)
levels(ds$Realm)[levels(ds$Realm) %in% c("Oceania", "Indo-Malay")] <- "IMO"

# Make a new variable for the LU x Realm combination
ds$LU_Realm <- paste(ds$LUH2, ds$Realm)

# Make sensible aggregations to get all N > 100
ds$LU_Realm[ds$LU_Realm %in% 
              c("Perennial Nearctic",
                "Perennial Palearctic")] <-
  "Perennial Holarctic"
ds$LU_Realm[ds$LU_Realm %in% 
              c("Managed_pasture Nearctic", 
                "Managed_pasture Palearctic")] <-
  "Managed_pasture Holarctic"
ds$LU_Realm[ds$LU_Realm %in% 
              c("Annual Australasia", 
                "Annual IMO")] <-
  "Annual IMOA"
ds$LU_Realm[ds$LU_Realm %in% 
              c("Rangelands Australasia",
                "Rangelands IMO")] <-
  "Rangelands IMOA"
ds$LU_Realm[ds$LU_Realm %in% 
              c("Rangelands Nearctic",
                "Rangelands Palearctic")] <-
  "Rangelands Holarctic"
ds$LU_Realm[ds$LU_Realm %in% 
              c("Managed_pasture Australasia", 
                "Managed_pasture IMO")] <-
  "Managed_pasture IMOA"
ds$LU_Realm[ds$LU_Realm %in% 
              c("Urban Australasia",
                "Urban IMO")] <-
  "Urban IMOA"
ds$LU_Realm[ds$LU_Realm %in% 
              c("Mature_secondary_vegetation Australasia",
                "Mature_secondary_vegetation IMO")] <-
  "Mature_secondary_vegetation IMOA"

# Combine Primary across realm, so all realms share the same baseline
ds$LU_Realm[ds$LUH2 == "Primary"] <- "Primary"

# Make LU_Realm a factor and make Primary the reference level
ds$LU_Realm <- as.factor(ds$LU_Realm)
ds$LU_Realm <- relevel(ds$LU_Realm, "Primary")

# Sample sizes are all now at least 100

summary_table <- ds %>%
  group_by(LU_Realm) %>%
  summarise(
    n_rows = n(),                     # number of rows
    n_unique_SS = n_distinct(SS),     # number of unique SS values
    .groups = "drop"
  )

# Display the table
knitr::kable(summary_table)


```


## Modelling
`Taxon_richness` is hypothesised to be affected by `LU_Realm`, but is also
expected to be affected deterministically by `Mean_taxon_precision`. As we want
the effects of `LU_Realm` when `Mean_taxon_precision` is at its maximal value, 
we first add a new column, `Taxonomic_imprecision`, calculated as (4 - 
`Mean_taxon_precision`) / 4; 4 is the maximum value for `Mean_taxon_precision`,
so this puts it onto a 0-1 scale to ease model convergence.

```{r add-imprecision}
ds$Taxonomic_imprecision <- (4 - ds$Mean_taxon_precision) / 4

```


### Using `glmmTMB`
The `glmmTMB` package allows zero-inflated negative binomial models, likely a 
better choice than a Poisson with an observation-level random intercept. We fit
two negative binomial variants (`zinb1` and `zinb2`), a zero-inflated poisson 
with an observation-level random intercept (`zipo`), and a poisson model with an
observation-level random effect to accommodate overdispersion (`polre`). AIC 
is used to choose among these alternative formulations, and favours `zinb2`.

```{r glmmTMB1}
tictoc::tic("Fitting zero-inflated negative-binomial1 GLMM in glmmTMB")
zinb1 <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LU_Realm + 
                         poly(Taxonomic_imprecision, 2) +
                         (1|SS) + (1|SSB) + (1|Taxon_fold_ss), 
                  data = ds,
                  family = nbinom1(link = "log"), # Neg binom for overdispersion
                  ziformula = ~1,                 # Zero-inflation intercept
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=1e3, eval.max=1e3)))
tictoc::toc()

tictoc::tic("Fitting zero-inflated negative-binomial2 GLMM in glmmTMB")
zinb2 <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LU_Realm + 
                         poly(Taxonomic_imprecision, 2) +
                         (1|SS) + (1|SSB) + (1|Taxon_fold_ss), 
                  data = ds,
                  family = nbinom2(link = "log"), # Neg binom for overdispersion
                  ziformula = ~ 1,                # Zero-inflation intercept
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=1e3, eval.max=1e3)))
tictoc::toc()


tictoc::tic("Fitting zero-inflated poisson GLMM with OLRE in glmmTMB")
zipo <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LU_Realm + 
                         poly(Taxonomic_imprecision, 2) +
                         (1 | SS) + (1 | SSB) + (1 | Taxon_fold_ss) + (1 |SSBS),
                  data = ds,
                  family = poisson(link = "log"), # Poisson
                  ziformula = ~1,                 # Zero-inflation intercept
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=1e3, eval.max=1e3)))
tictoc::toc()

tictoc::tic("Fitting poisson GLMM with OLRE in glmmTMB")
polre <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LU_Realm + 
                         poly(Taxonomic_imprecision, 2) +
                         (1 | SS) + (1 | SSB) + (1 | Taxon_fold_ss) + (1 |SSBS),
                  data = ds,
                  family = poisson(link = "log"), # Poisson
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=1e3, eval.max=1e3)))
tictoc::toc()

AIC(zinb1, zinb2, zipo, polre)
```

Residuals are not very well distributed, suggesting that model fit may be 
improved further by specifying a `dispformula`. Experimentation with several 
formulations favours a dispersion model with `SS` as a random intercept in the 
dispersion model. (Including it as a fixed effect requires too many extra 
parameters for model convergence.)

```{r refine}

tictoc::tic("Refining richness model")
# Make Taxonomic_imprecision linear (poly 2 gives + 1.8 AIC units)
zinb2a <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LU_Realm + 
                         Taxonomic_imprecision +
                         (1|SS) + (1|SSB) + (1|Taxon_fold_ss), 
                  data = ds,
                  family = nbinom2(link = "log"), # Neg binom for overdispersion
                  ziformula = ~ 1,                # Zero-inflation intercept
                  dispformula = ~ (1|SS),       # Definite heterogeneity
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=5e3, eval.max=5e3)))

# Drop zero-inflation
zinb2ab <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LU_Realm + 
                         Taxonomic_imprecision +
                         (1|SS) + (1|Taxon_fold_ss), 
                  data = ds,
                  family = nbinom2(link = "log"), # Neg binom for overdispersion
                  ziformula = ~ 0,                # Zero-inflation intercept
                  dispformula = ~ (1|SS),       # Definite heterogeneity
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=1e3, eval.max=1e3)))


# And also drop (1|SSB)
zinb2b <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LU_Realm + 
                         Taxonomic_imprecision +
                         (1|SS) + (1|Taxon_fold_ss), 
                  data = ds,
                  family = nbinom2(link = "log"), # Neg binom for overdispersion
                  ziformula = ~ 1,                # Zero-inflation intercept
                  dispformula = ~ (1|SS),       # Definite heterogeneity
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=1e3, eval.max=1e3)))

# And instead drop (1|Taxon_fold_ss)
zinb2c <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LU_Realm + 
                         Taxonomic_imprecision +
                         (1|SS) + (1|SSB), 
                  data = ds,
                  family = nbinom2(link = "log"), # Neg binom for overdispersion
                  ziformula = ~ 1,                # Zero-inflation intercept
                  dispformula = ~ (1|SS),       # Definite heterogeneity
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=1e3, eval.max=1e3)))

# And instead drop zero inflation parameter, as the value was -5
zinb2d <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LU_Realm + 
                         Taxonomic_imprecision +
                         (1|SS) + (1|SSB) + (1|Taxon_fold_ss), 
                  data = ds,
                  family = nbinom2(link = "log"), # Neg binom for overdispersion
                  # DROPPING zero inflation as the value was -5
                  dispformula = ~ (1|SS),       # Definite heterogeneity
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=1e3, eval.max=1e3)))
tictoc::toc()

AIC(zinb2a, zinb2ab, zinb2b, zinb2c, zinb2d)


```

```{r choose}
best <- zinb2b
summary(best)
performance::r2(best)
car::Anova(best)
```

```{r diagnostics}
sjPlot::plot_model(best,
                   type = "diag")
glmmTMB::diagnose(best)
sim <- DHARMa::simulateResiduals(best, nSim = 250)
plot(sim)
DHARMa::testZeroInflation(sim)
DHARMa::testDispersion(sim)
DHARMa::plotResiduals(sim, form = ds$LU_Realm)
DHARMa::plotResiduals(sim, form = ds$Taxonomic_imprecision)
DHARMa::plotResiduals(sim, form = ds$Taxon_fold_ss)
DHARMa::plotResiduals(sim, form = ds$Realm)
DHARMa::plotResiduals(sim, form = ds$Biome)
DHARMa::plotResiduals(sim, form = ds$SS)

```

```{r plots}

pred <- ggeffects::ggpredict(best, terms = "LU_Realm", bias_correction = TRUE)
ggplot2::ggplot(pred, aes(x = predicted, y = x)) +
  geom_point(size = 3, color = "blue") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "gray") +
  labs(title = "Predicted Species Richness by LU x Realm",
       x = "Predicted Species Richness", y = "LU x Realm") +
  theme_minimal()

```

### Sensitivity of estimates for LU x Realm combinations
8 models have been fitted, 7 of which have converged. Here is a pairs plot of 
the richness estimates from those 7 (suppressing irrelevant warnings for 
aesthetic reasons). The column and row relating to the best model, `zinb2a`, is
highlighted.

```{r basic, message = FALSE, warnings = FALSE}
pred_zinb1 <- ggeffects::ggpredict(zinb1, terms = "LU_Realm")$predicted
predictions <- data.frame(zinb1 = pred_zinb1)
predictions$zinb2 <- ggeffects::ggpredict(zinb2, terms = "LU_Realm")$predicted
predictions$zipo <- ggeffects::ggpredict(zipo, terms = "LU_Realm")$predicted
predictions$polre <- ggeffects::ggpredict(polre, terms = "LU_Realm")$predicted
predictions$zinb2a <- ggeffects::ggpredict(zinb2a, terms = "LU_Realm")$predicted
predictions$zinb2b <- ggeffects::ggpredict(zinb2b, terms = "LU_Realm")$predicted
predictions$zinb2d <- ggeffects::ggpredict(zinb2d, terms = "LU_Realm")$predicted

```


```{r ggpairs-highlight, message=FALSE, warning=FALSE}

# Suppress GGally progress bar
options(GGally.progress = FALSE)

# Create ggpairs plot
p <- ggpairs(predictions)

# Highlight row and column 6 (zinb2b) with pastel background
for (i in 1:p$nrow) {
  for (j in 1:p$ncol) {
    if (i == 6 || j == 6) {
      p[i, j] <- p[i, j] + theme(panel.background = element_rect(fill = "lavender", color = NA))
    }
  }
}

print(p)
```

## Fit LU-only model
Because estimating BII for so many land use x realm combinations has the 
potential to overstretch the data, we also fit a simpler model with a single
estimate per land-use class.

```{r lu-only}
lu_only <- glmmTMB::glmmTMB(Taxon_richness ~ 
                         LUH2 + 
                         Taxonomic_imprecision +
                         (1|SS) + (1|SSB) + (1|Taxon_fold_ss), 
                  data = ds,
                  family = nbinom2(link = "log"), # Neg binom for overdispersion
                  ziformula = ~ 1,                # Zero-inflation intercept
                  dispformula = ~ (1|SS),       # Definite heterogeneity
                  control = glmmTMB::glmmTMBControl(
                    optCtrl = list(iter.max=5e3, eval.max=5e3)))

summary(lu_only)
performance::r2(lu_only)
car::Anova(lu_only)

lu_pred <- ggeffects::ggpredict(lu_only, 
                                terms = "LUH2", 
                                bias_correction = TRUE)
ggplot2::ggplot(lu_pred, aes(x = predicted, y = x)) +
  geom_point(size = 3, color = "blue") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), 
                 height = 0.2, 
                 color = "gray") +
  labs(title = "Predicted Species Richness by aggregated LUH2 class",
       x = "Predicted Species Richness", y = "Land-Use Class") +
  theme_minimal()
```



## Populating the LUH2 x Realm matrix
Coefficients from the best model are converted into estimates of relative 
species richness (with the baseline, primary vegetation, having a relative 
richness of 1 in all realms) for each LUH2 class in each biogeographic realm. 
Estimates are capped at 1, mirroring their handling in the BII pipeline. The 
resulting data frame also indicates which model coefficient provided each value.

```{r get_matrix}
sec_coefs <- fixef(lu_only)$cond[c("LUH2Intermediate_secondary_vegetation",
                                   "LUH2Mature_secondary_vegetation",
                                   "LUH2Young_secondary_vegetation")]
sec_ests <- exp(sec_coefs)
mean_sec <- mean(sec_ests)

LUH2_set <- c("c3ann", "c3nfx", "c3per",
              "c4ann", "c4per", "pastr",
              "primf", "primn", "range",
              "secdf", "secdn", "urban")

# Set up data frame to hold the values CSIRO wants
rel_rich <- expand.grid(
  realm = realm_set, # From when the data were first read in
  LUH2 = LUH2_set
)
rel_rich$relative_richness <- NA # Relative richness from LU_Realm model
rel_rich$model_term <- NA        # Which coefficient in the model it comes from
rel_rich$lu_rel_richness <- NA   # Relative richness from lu_only model

# Get names of the model factor levels other than the baseline (primary)
coef_names <- paste("LU_Realm", pred$x, sep = "")
coefs <- fixef(best, component = "cond")
lu_coef_names <- paste("LUH2", lu_pred$x, sep = "")
lu_coefs <- fixef(lu_only, component = "cond")

# Estimate richness relative to primary (capping at 1)
ests <- ifelse(pred$predicted > pred$predicted[1],
               1,
               pred$predicted/pred$predicted[1])

# Do the same with the lu_only model
lu_ests <- ifelse(lu_pred$predicted > lu_pred$predicted[1],
               1,
               lu_pred$predicted/lu_pred$predicted[1])

# Primary is the baseline in each realm, so primn and primf get values of 1
rel_rich$relative_richness[rel_rich$LUH2 %in% c("primf", "primn")] <- 1
rel_rich$model_term[rel_rich$LUH2 %in% c("primf", "primn")] <- 
  "Primary (intercept)"
rel_rich$lu_rel_richness[rel_rich$LUH2 %in% c("primf", "primn")] <- 1

# Secondary gets the mean of the three secondary coefficients from the lu_only 
# model (and is NA for the LUxRealm model, as it would be a faff to do more!)
rel_rich$lu_rel_richness[rel_rich$LUH2 %in% c("secdf", "secdn")] <- mean_sec

# Loop through coefficients for the other LU_realm combinations
for (lur in coef_names[-1]) {
  # This loop puts the estimate of relative richness (calculated from the fixed 
  # effect whose name is lur) in the right cell(s) of the rel_rich data frame.
  # It also each time puts the relative richness from the land-use-only model
  # (calculated from the fixed effect whose name is the lu part of lur) in all
  # those rows, as lu_rel_richness; those assignments happen more than once, but
  # that doesn't matter.
  est <- ests[which(coef_names == lur)]
  space_pos <- regexpr(" ", lur)
  lu <- substring(lur, 9, space_pos - 1)
  lu_est <- lu_ests[which(lu_coef_names == paste("LUH2", lu, sep = ""))]
  realm <- substring(lur, space_pos + 1)
  
  # Which LUH2 classes should get each coefficient?
  which_lu <- NA
  if (lu == "Annual") which_lu <- c("c3ann", "c3nfx", "c4ann")
  if (lu == "Managed_pasture") which_lu <- "pastr"
  if (lu == "Perennial") which_lu <- c("c3per", "c4per")
  if (lu == "Rangelands") which_lu <- "range"
  if (lu == "Secondary") which_lu <- c("secdf", "secdn")
  if (lu == "Urban") which_lu <- "urban"
  
  # Which realms should get each coefficient?
  which_realm <- NA
  if (realm == "Afrotropic") which_realm <- "Afrotropic"
  if (realm == "Australasia") which_realm <- "Australasia"
  if (realm == "Holarctic") which_realm <- c("Nearctic", "Palearctic")
  if (realm == "IMO") which_realm <- c("Indo-Malay", "Oceania")
  if (realm == "IMOA") which_realm <- c("Indo-Malay", "Oceania", "Australasia")
  if (realm == "Nearctic") which_realm <- "Nearctic"
  if (realm == "Neotropic") which_realm <- "Neotropic"
  if (realm == "Palearctic") which_realm <- "Palearctic"
  
  # Put the right coefficient in the relevant rows of the data frame, and say 
  # where it came from for traceability.
  matches <- which(rel_rich$realm %in% which_realm &
                   rel_rich$LUH2 %in% which_lu)
  rel_rich$relative_richness[matches] <- est
  rel_rich$model_term[matches] <- substring(lur, 9) # Omits "LU_Realm"
  rel_rich$lu_rel_richness[matches] <- lu_est # Inefficient but OK
}


# Produce attributes for traceability
attr(rel_rich, which = "date_created") <- Sys.Date()
attr(rel_rich, which = "richness_file_date") <- attr(ds, which = "date_created")
attr(rel_rich, which = "date_database_extract") <- 
  attr(ds, which = "date_database_extract")

# Save as an .rds so it has the attributes as part of it
saveRDS(rel_rich, file = "output/relative_richness.rds")

# Show the result in table form
knitr::kable(rel_rich)

```

